### 2013-09-03 notes/notes.tex
added notes.

\documentclass[11p]{paper}

%%%%%%begin preamble
\usepackage[hmargin=1in, vmargin=1in]{geometry} % Margins
\usepackage{url}
\usepackage{hyperref}
\usepackage{times}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{pdfpages}

%\usepackage{fontspec}
%\setmainfont{TimesNewRoman}
%\definetypeface[Serapion][rm][Xserif][Serapion Pro]
%\setupbodyfont[Serapion, 12pt]

%%%
%%%%%% uncomment following 4 lines to adjust title size/shape and
%%%%%% trailing space
%% \usepackage{titling}
%% %\pretitle{\noindent\Large\bfseries}
%% \date{}
%% \setlength{\droptitle}{-1in}
%\posttitle{\\}

\setcounter{tocdepth}{2}
%% headers
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{}
\chead{}
\rhead{Weakly Nonlinear MRI Notes}
\lfoot{}
\cfoot{\thepage}
\rfoot{}

\newcommand{\Reyn}{\ensuremath{\mathrm{Re}}}
\newcommand{\Rmag}{\ensuremath{\mathrm{Rm}}}
\newcommand{\Rmagc}{\ensuremath{\mathrm{Rm}_\mathrm{crit}}}
\newcommand{\Prandtl}{\ensuremath{\mathrm{Pm}}}
\newcommand{\Lund}{\ensuremath{\mathrm{S}}}
\newcommand{\Lundc}{\ensuremath{\mathrm{S}_\mathrm{crit}}}
\newcommand{\yt}{\texttt{yt}}
\newcommand{\enzo}{\texttt{Enzo}}
\newcommand{\nosection}[1]{%
  \refstepcounter{section}%
  \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
  \markright{#1}}
\newcommand{\shellcmd}[1]{\\\indent\indent\texttt{\footnotesize\$ #1}\\}

%%%%%%end preamble

\title{Some Notes on the Weakly Nonlinear MRI Project}
\begin{document}
\maketitle

\section{Mathematical Basics}
\label{sec:basics}

Here are some really basic math stuff that might be useful to you.

\section{Computational Basics}
\label{sec:comp_basics}

We'll be making a lot of use of Python, SymPy, the iPython notebook,
and later, Dedalus 2.0.

\subsection{Software to install}
\label{sec:software}

\begin{itemize}
\item Python3 (latest version)
\item SymPy 
\item ipython 1.0
\item matplotlib 1.3.0
\item mercurial (hg)
\end{itemize}

I know you use a Mac, which I don't, so I'm not an expert. If you
haven't already, one way to do this is to install python3, and then
install pip using the \texttt{get-pip.py} script described at
\url{http://www.pip-installer.org/en/latest/installing.html}. Once you have pip installed, you can just do 
\shellcmd{pip install sympy} 
for example, and everything should just work.

\subsection{Dedalus 2.0}
\label{sec:dedalus2}

Dedalus 2.0 isn't quite ready for us to use yet, and it's very likely
a bunch of stuff will change between now and when we are ready, but if
you're curious, it's here: \url{https://bitbucket.org/jsoishi/dedalus2}.

\subsection{Version Control with Mercurial and Bitbucket}
\label{sec:vc}

In order to facilitate our collaboration, we'll be using the mercurial
version control system. A great intro to is
\url{http://hginit.com}.

Our shared repository will be hosted on \url{https://bitbucket.org},
starting with this very document, which you can find here:
\url{https://bitbucket.org/jsoishi/weakly_nonlinear_MRI/}

\end{document}

### 2013-09-16 notes/notes.tex
loads more stuff.

@@
 \usepackage{hyperref}
-\usepackage{times}
+%\usepackage{times}
 \usepackage{natbib}
@@
 
+\section{Overview of The Problem}
+\label{sec:overview}
+
+We want to construct an analytical understanding of the saturation of
+the MRI, complementary to the linear analysis of \citet{Pessah2010}.
+
+In order to understand saturation of the MRI, we need to consider the
+options available to it.
+
 \section{Mathematical Basics}
@@
 
-Here are some really basic math stuff that might be useful to you.
+Here are some really basic math stuff that might be useful to
+you. First some notation. Derivatives are a fun one. You'll see the
+partial derivative of a function $f$ with respect to $x$, 
+\begin{equation*}
+  \label{eq:partial}
+  \frac{\partial f}{\partial x}
+\end{equation*}
+written as $\partial_x f$, or even $f_x$, especially in the applied
+math literature. I tend to use $\partial_x f$, because it's terse but
+not so terse I forget what it is I meant by it.
+
+In order to solve the equations of magnetohydrodyanmics (MHD), we'll
+need names for the velocity, density, magnetic, and pressure
+fields. Typically, the velocity vector is written as $\mathbf{v}$ or
+$\mathbf{u}$ (I tend to use the latter). The components of
+$\mathbf{u}$ are written either as $(u_x, u_y, u_z)$ (typical in the
+astrophysical literature) or $(u, v, w)$ (more common in geophysical
+fluid dynamics). Density is nearly always $\rho$. The magnetic field
+is usually $\mathbf{B}$ when a vector, with scalar components $(B_x,
+B_y, B_z)$ though you may (very rarely) see $(a, b, c)$. Obviously,
+you'd substitute the subscripts for $r, \theta, \phi$ in spherical and
+$r, \theta, z$ in cylindrical coordinates. Occasionally you will see
+the cylindrical radius written $\altrho$.
+
+In compressible gas dynamics, the sound speed is usually written as
+$c$ or $c_s$. There rarely is confusion with the speed of light, even
+in relativisitic (special or general) MHD. Of course, GR MHD is way
+beyond the scope of this project, but it brings from GR a lot of very
+interesting and rather different notation.
+
+There are a lot of ways to write the equations we need, and they go by
+a lot of names.
+
+Fundamentally, they are
+
+\begin{equation}
+  \label{eq:continuity}
+  \partial \rho
+\end{equation}
 
@@
 
-Our shared repository will be hosted on \url{https://bitbucket.org},
-starting with this very document, which you can find here:
-\url{https://bitbucket.org/jsoishi/weakly_nonlinear_MRI/}
+Our shared repository will be hosted on
+\url{https://bitbucket.org}. You'll need to first setup a bitbucket
+account. Just go to the site, create a username, and email it to
+me. I'll add you to the access list for our, which you can find here:
+\url{https://bitbucket.org/jsoishi/weakly_nonlinear_MRI/}. Until
+you've created a username and emailed me, I'll leave it public so you
+can see how it works without having to wait for me. However, after
+you've gotten your username, we'll make it private, so you don't have
+to worry about anyone seeing our mistakes (which we will make plenty
+of!). Of course, there is tremendous value in open science, so I
+typically open up my repositories after publication. I will encourage
+but not require you to do the same.
 
+\section{Roadmap}
+\label{sec:roadmap}
+
+Here's a very rough, and sure to change outline of the project as I
+see it now.
+
+\begin{enumerate}
+\item Background reading
+\item Linearize the MHD equations in a shearing box
+\item Derive the linear conditions for the MRI with viscosity and resistivity
+\item Derive the Ginzburg-Landau equation from convection
+\item Derive the Ginzburg-Landau equation from the MHD equations in a
+  thin-gap Taylor-Couette flow (rederive \citet{2007PhRvL..98c4501U})
+\item Extend weakly nonlinear analysis to the wide gap case
+\end{enumerate}
+
+Once we get down to the bottom of the list, we can begin to understand
+
+\section{Annotated Bibliography}
+\label{sec:annotated_bib}
+
+Basics on the CGLE: ``The complex Ginzburg-Landau equation: an introduction,'' Garcia-Morales, V \& Krischner, K. \url{http://www.tandfonline.com/doi/pdf/10.1080/00107514.2011.642554}
+
+\bibliographystyle{apj}
+\bibliography{library}
 \end{document}

### 2013-09-27 notes/notes.tex
first draft sent to susan.

@@
 %%%%%%begin preamble
-\usepackage[hmargin=1in, vmargin=1in]{geometry} % Margins
+%\usepackage[hmargin=1in, vmargin=1in]{geometry} % Margins
 \usepackage{url}
@@
 Here are some really basic math stuff that might be useful to
-you. First some notation. Derivatives are a fun one. You'll see the
-partial derivative of a function $f$ with respect to $x$, 
+you. First some notation. You'll see the partial derivative of a
+function $f$ with respect to $x$,
 \begin{equation*}
@@
 $r, \theta, z$ in cylindrical coordinates. Occasionally you will see
-the cylindrical radius written $\altrho$.
+the cylindrical radius written $\varpi$.
 
@@
 $c$ or $c_s$. There rarely is confusion with the speed of light, even
-in relativisitic (special or general) MHD. Of course, GR MHD is way
-beyond the scope of this project, but it brings from GR a lot of very
-interesting and rather different notation.
+in relativistic (special or general) MHD.
+
+\subsection{The Equations of MHD}
+\label{sec:equations}
 
 There are a lot of ways to write the equations we need, and they go by
-a lot of names.
+a lot of names. We will start by writing the equations of
+hydrodynamics, and then add MHD later. Fundamentally, these equations
+express the conservation of mass, momentum, and energy. For MHD, we
+also need to describe the evolution of the magnetic field. Please note
+that we are not saying anything physical about fluids, plasmas, or
+(magneto)hydrodynamics here. We're just laying out the relevant
+equations. There is a lot to say about MHD, but I will only say one
+physical thing here: it is really, really important to remember that
+MHD equations are not like general relativity or Maxwell's equations
+or the Schrodinger Equation: they are themselves an
+approximation. They are \emph{not} always valid: the description of a
+system as a continuum of neutral (hydrodynamics) or charged (MHD)
+particles can break down without the mathematical solution you've
+painstakingly derived telling you a thing about it. This is especially
+important to keep in mind because this whole project is about
+constructing analytic solutions by making loads of approximations.
 
-Fundamentally, they are
+For a fluid, mass conservation is given by 
+\begin{equation}
+  \label{eq:continuity}
+  \partial_t \rho + \nabla \cdot \rho \mathbf{u} = 0.
+\end{equation}
+This is also known as the \textit{continuity} equation, because it not
+only says that mass is conserved, it also states that mass must flow
+continuously from one place to another. In this form, it is the model
+of a class of equations called the conservation laws, because you can
+think of it as saying the following. The time rate of change of the
+mass of fluid within a given volume $\mathcal{V}$ is given by the
+\emph{flux} $\mathcal{F} = \rho \mathbf{u}$ across the surface
+bounding $\mathcal{V}$.
 
 \begin{equation}
-  \label{eq:continuity}
-  \partial \rho
+  \label{eq:navier-stokes}
+  \partial_t \mathbf{u} + \mathbf{u \cdot \nabla u} = -\frac{\nabla p}{\rho} + \nu \nabla^2 \mathbf{u},
 \end{equation}
+which is also called the \textit{Navier-Stokes} equation when $\nu
+\neq 0$ and the \textit{Euler} equation otherwise. This equation is
+essentially $F = ma$ for a fluid: any additional physics we want to
+add will add force terms to the right hand side of this equation. Most
+important for us will be the fictitious rotational forces and the
+Lorentz force of MHD.
 
+Of course, we also need an equation for the conservation of energy (that is, we must satisfy the First Law of Thermodynamics).
+However, we will be working exclusively with \textit{incompressible} fluids, so we can simply enforce the \emph{incompressibility constraint},
+\begin{equation}
+  \label{eq:incompressibility}
+  \mathbf{\nabla \cdot u} = 0.
+\end{equation}
+This is called a ``constraint'' equation because it does not specify
+the time evolution of the system, but instead specifies a certain
+property the solution must obey.
+
+To add the magneto to hydrodynamics, we first introduce the Lorentz
+force per unit volume,
+\begin{equation}
+  \label{eq:lorentz}
+  \mathbf{f_L = \frac{J \times B}{\rho}}.
+\end{equation}
+The current density (current per unit volume) is just
+\begin{equation}
+  \label{eq:current}
+  \mathbf{J = \frac{\nabla \times B}{4\pi}}
+\end{equation}
+in Gaussian units (which we'll use. Sorry.). We then combine the $f_L$
+with the definition of $J$, and using a vector calculus you can easily
+look up, we arrive at 
+\begin{equation}
+  \label{eq:lorentz_final}
+  \mathbf{F_L = -\frac{\nabla B^2}{8 \pi \rho} + \frac{B \cdot \nabla B}{4
+      \pi \rho}}.
+\end{equation}
+This form is very telling, because it splits the Lorentz force into
+two pieces: the \emph{magnetic pressure} ($\nabla B^2/8 \pi$) and the
+\emph{magnetic tension} ($B \cdot \nabla B / 4 \pi$). The physics of
+the MRI is contained in the latter: magnetic fields, when stretched
+act exactly like elastic bands.
+
+Of course, now we have added a new vector field, $\mathbf{B}$, to our
+equation set, so obviously we need three more equations
+
+I'm glossing over a lot here, but for MHD, where the plasma behaves
+like a fluid, the ``Ohm's Law'', or relationship between the Electric
+Field and the current, gives
+\begin{equation}
+  \label{eq:ohms}
+  \mathbf{E} = -\frac{\mathbf{u \times B}}{c} - \frac{\mathbf{J}}{\sigma},
+\end{equation}
+where $\sigma$ is the \emph{conductivity} of the plasma (we typically
+never talk in terms of the conductivity, preferring the resistivity,
+which is implicitly derived below).
+
+Now we can employ one of the Maxwell's equations,
+\begin{equation}
+  \label{eq:maxwell}
+  \partial_t \mathbf{B} = -\frac{c}{4\pi} \mathbf{\nabla \times E},
+\end{equation}
+substituting in equation~(\ref{eq:ohms}), we arrive at the
+\emph{induction} equation, sometimes also known as the \emph{MHD}
+equation,
+\begin{equation}
+  \label{eq:induction}
+  \partial_t \mathbf{B = \nabla \times (u \times B) + \mu \nabla^2 B},
+\end{equation}
+where $\mu$ is the \emph{resistivity} of the plasma.
 \section{Computational Basics}
@@
 \url{https://bitbucket.org}. You'll need to first setup a bitbucket
-account. Just go to the site, create a username, and email it to
+account. Just go to the site, create a user name, and email it to
 me. I'll add you to the access list for our, which you can find here:
 \url{https://bitbucket.org/jsoishi/weakly_nonlinear_MRI/}. Until
-you've created a username and emailed me, I'll leave it public so you
+you've created a user name and emailed me, I'll leave it public so you
 can see how it works without having to wait for me. However, after
-you've gotten your username, we'll make it private, so you don't have
+you've gotten your user name, we'll make it private, so you don't have
 to worry about anyone seeing our mistakes (which we will make plenty
@@
 \end{document}
+%  LocalWords:  Pessah magnetohydrodyanmics MHD Navier SymPy iPython
+%  LocalWords:  Dedalus ipython matplotlib hg py sympy Bitbucket CGLE
+%  LocalWords:  bitbucket Roadmap Linearize Ginzburg Couette
+%  LocalWords:  Krischner

### 2013-12-20 notes/mri_notes.tex
Added beginning of project notes

\documentclass[letterpaper,12pt]{article}

%% Page dimensions for 1 inch margins
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-0in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in} 
\setlength{\headheight}{0in}
\setlength{\headsep}{0in} 
\setlength{\hoffset}{0in}
\setlength{\voffset}{0in}

%% Graphics and math
\usepackage{graphics,graphicx,subfigure}
\usepackage{amsmath}

\newcommand\lsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
        \raise1pt\hbox{$<$}}}
\newcommand\gsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
        \raise1pt\hbox{$>$}}}

%% Bibliographic stuff

%\usepackage[numbers]{natbib}
\usepackage{natbib}
\bibliographystyle{apj}
%\setlength{\bibsep}{0.0pt}

\newcommand\apjl{ApJL}
\newcommand\apj{ApJ}
\newcommand\aj{AJ}
\newcommand\apjs{ApJS}
\newcommand\aap{AAP}
\newcommand\mnras{MNRAS}
\newcommand\araa{ ARA\&A }
\newcommand\nat{ Nature }

%Use this for citing without parentheses.
\newcommand{\citei}[1]{\citeauthor{#1} \citeyear{#1}}

\begin{document}
\pagestyle{plain}
\pagenumbering{arabic}

\title{Weakly Nonlinear MRI: Notes}
\author{Susan Clark}
\maketitle
\vspace{0.2cm}

The basic setup for this investigation will follow \citet{Umurhan:2007dz}, who apply a weakly nonlinear analysis to the MRI in the shearing box approximation, in a thin-gap limit with idealized boundary conditions. We will follow a similar derivation, without the thin-gap restriction and with more realistic boundary conditions.

The shearing box (SB) approximation is one in which one considers the flow in a small box at some fiducial distance ${r_0}$ away from the center of the rotating system. This constitutes only a small portion of an accretion disk, but makes the problem more tractable, both for analytic and numerical analyses.

My understanding is that the thin-gap limit allows one to ignore radially-dependent (``channel") modes. Channel modes are axially symmetric, linear modes, and are exact solutions of the nonlinear MRI in the SB approximation. These are the so-called ``primary MRI" modes discussed in \citet{Pessah:2010ic}. The advantage to analyzing the MRI without channel modes is that the background is steady and has a constant pressure. By widening our gap, we will complicate things because the non-constant background means that we open ourselves to parasitic modes -- modes whose growth rates depend on the amplitudes of the primary modes. In particular, the channel modes are Kelvin-Helmholtz unstable.

Another way to say this is that the wide gap approximation does not allow us to ignore the $\left(\textbf{u} \cdot \nabla \right) \textbf{u}$ term in the Navier-Stokes equation, which couples velocities with each other and thus allows the existence of parasitic modes. This can be seen by considering $\textbf{u}$ as a sum of discrete modes with wavenumber k:

\begin{equation}
\textbf{u} = \sum\limits_{k = 0}^{\infty}a_k e^{ikx} \, dk
\end{equation}

\noindent The Fourier transform of $\left(\textbf{u} \cdot \nabla \right) \textbf{u}$ is:

\begin{equation}
a_k \otimes i \textbf{k} \cdot a_k
\end{equation}

\noindent Clearly, each wavenumber (and its associated amplitude) depends on all other wave numbers.\\

So essentially, our approach is similar to \citet{Umurhan:2007dz}, but our setup is complementary to \citet{Pessah:2010ic}. \\

\noindent The MRI setup in \citet{Umurhan:2007dz} is as follows: 
\begin{itemize}
\item An axisymmetric channel flow with a differential rotation ${\Omega\left(r\right) \propto \Omega_0\left(\frac{r}{r_0}\right)^{-q}}$ 
\item A velocity $\textbf{V}$ = $U\left(x\right)\hat{\bf{y}}$ with a linear shear profile $U\left(x\right) = -qU_0x$
\item Constant, vertical magnetic field $\bf{\hat{B}}$ = $B_0\hat{z}$
\end{itemize}

We will use the incompressible limit ($\nabla \cdot \textbf{u}$ = $0$), which applies because we consider weak magnetic fields (magnetic energy is small compared to thermal energy).

\bibliography{mri.bib}

\end{document}

### 2013-12-20 notes/mri_notes.tex
Added more bullet points

@@
 
-We will use the incompressible limit ($\nabla \cdot \textbf{u}$ = $0$), which applies because we consider weak magnetic fields (magnetic energy is small compared to thermal energy).
+We will use the incompressible limit ($\nabla \cdot \textbf{u}$ = $0$), which applies because we consider weak magnetic fields (magnetic energy is small compared to thermal energy). \\
+
+\noindent We will proceed as follows: \\
+
+\begin{itemize}
+\item Linearize equations for perturbations of the form $e^{st + ik_x x + ik_z z}$. We neglect y-dependence because of axisymmetry.
+\item Write linearized equations in matrix form, use SymPy to obtain dispersion relation. Should be a 4th order polynomial in s.
+\item We are only interested in marginality to the MRI, so we will take s=0. 
+\item This should allow us to isolate the most unstable mode at criticality. 
+\end{itemize}
+
+\noindent Things to include:
+\begin{itemize}
+\item Derivations that I wrote out by hand (scan)
+\item Nondimensionalization 
+\item Multiscale analysis (in analogy to Raleigh B\'enard convection)
+\item Boundary conditions
+\item Amplitude functions
+\item GLE
+\end{itemize}
 

### 2014-01-27 notes/mri_notes.tex
Revised project outline

@@
 
-Another way to say this is that the wide gap approximation does not allow us to ignore the $\left(\textbf{u} \cdot \nabla \right) \textbf{u}$ term in the Navier-Stokes equation, which couples velocities with each other and thus allows the existence of parasitic modes.
+%Another way to say this is that the wide gap approximation does not allow us to
+In general, we cannot ignore the $\left(\textbf{u} \cdot \nabla \right) \textbf{u}$ term in the Navier-Stokes equation, which couples velocities with each other.
+%and thus allows the existence of parasitic modes.
 This can be seen by considering $\textbf{u}$ as a sum of discrete modes with wavenumber k:
@@
 
-\noindent We will proceed as follows: \\
+\noindent \textbf{We will proceed as follows:}
 
@@
 \item Linearize equations for perturbations of the form $e^{st + ik_x x + ik_z z}$. We neglect y-dependence because of axisymmetry.
-\item Write linearized equations in matrix form, use SymPy to obtain dispersion relation. Should be a 4th order polynomial in s.
-\item We are only interested in marginality to the MRI, so we will take s=0. 
-\item This should allow us to isolate the most unstable mode at criticality. 
+\item Write linearized equations in matrix form, use SymPy to obtain dispersion relation. This is a 4th order polynomial in s.
+\end{itemize}
+\textit{The following five bullet points are performed in analogy to Raleigh B\'enard convection:}
+\begin{itemize}
+\item We want to derive an envelope (amplitude) equation, which we will do with a multiscale analysis.
+We choose scalings such that $s$, $k_x$, and $k_z$ in our dispersion relation come in at the same (lowest) order in the small parameter $\epsilon$.
+This determines the scalings for the corresponding slow variables T, X, and Y.
+\item Rewrite equations in terms of quantities that depend on both fast and slow variables, in the form $u(x, y, t; X, Y, T) = A(X, Y, T)e^{i\mathbf{k} \cdot \mathbf{x}}$.
+\item Expand quantities in powers of $\epsilon$.
+\item Group terms by $\mathcal{O}\left(\epsilon\right)$
+\item Solve successively to obtain envelope equation. We solve up to the order where a term appears that can cause saturation.\\
+
+%\item We are only interested in marginality to the MRI, so we will take s=0. 
+%\item This should allow us to isolate the most unstable mode at criticality. 
 \end{itemize}

### 2014-02-06 notes/mri_notes.tex
Derivation of equations

@@
 %\item This should allow us to isolate the most unstable mode at criticality.
+
 \end{itemize}
@@
 
+\vspace{10cm}
+
+\textbf{Derivation of Equations} \\
+
+Begin with the Navier-Stokes equation:
+
+\begin{equation}
+\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, = \, -\frac{1}{\rho}\nabla P \, - \, \nabla\Phi \, + \, \frac{1}{\rho} \left(\mathbf{J}\times\mathbf{B}\right) \, - \, \nu\nabla^2 \mathbf{u} \, - \, 2\mathbf{\Omega} \times \mathbf{u} \, - \, \mathbf{\Omega} \times \left(\mathbf{\Omega} \times \mathbf{r} \right)
+\end{equation}\label{momeqn}
+
+The reference frame is rotating with angular velocity $\Omega_0 \mathbf{\hat{z}}$. \\
+
+The coriolis term becomes: \\
+
+$2\mathbf{\Omega} \times \mathbf{u} \, = \, 2\Omega_0 \mathbf{\hat{z}} \times \mathbf{u} \, = \, 2\Omega_0\left(\mathbf{\hat{z}} \times u_r \mathbf{\hat{r}} \, + \, \mathbf{\hat{z}} \times u_{\phi} \mathbf{\hat{\phi}}\right) \, = \, 2\Omega_0\left(v_r \mathbf{\hat{\phi}} - v_{\phi}\mathbf{\hat{r}}\right) $\\
+
+The centrifugal term becomes: \\
+
+$\mathbf{\Omega} \times \left(\mathbf{\Omega} \times \mathbf{r} \right) \, = \, \Omega_0 \mathbf{\hat{z}} \times \left(\Omega_0 \mathbf{\hat{z}} \times \mathbf{\hat{r}}\right) = \Omega_0^2 \mathbf{\hat{z}} \times \left(r \mathbf{\hat{\phi}}\right) \, = \, -\Omega_0^2 r \mathbf{\hat{r}}$ \\
+
+The gravitational force per unit mass can be written: \\
+
+$\nabla\Phi \, = \, -\frac{1}{\rho_b}\nabla P_b + r\Omega^2 \mathbf{\hat{r}}$ \\
+
+Where $\rho_b$ and $P_b$ are the density and pressure that satisfy a barotropic relation in a steady axisymmetric solution. \\
+
+The current density is $\mathbf{J} = \frac{1}{4\pi} \nabla \times \mathbf{B}$ and thus the Lorentz force becomes: \\
+
+$\frac{1}{\rho} \left(\mathbf{J}\times\mathbf{B}\right) = \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} = - \frac{1}{8 \pi \rho}\nabla B^2 + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}$ \\
+
+Using the identity: \\
+
+$\left( \nabla \times \mathbf{B} \right) \times \mathbf{B} = - \mathbf{B} \times \left( \nabla \times \mathbf{B} \right) = -\nabla \left(\mathbf{B} \cdot \mathbf{B}\right) + \left(\mathbf{B} \cdot \nabla \right)\mathbf{B} + \left(\mathbf{B} \cdot \nabla \right)\mathbf{B} + \mathbf{B} \times \left(\nabla \times \mathbf{B}\right)$ \\
+
+The initial magnetic field is a constant vertical field $\mathbf{B} = B_0 \mathbf{\hat{z}}$, and thus the magnetic pressure term becomes: \\
+
+$- \frac{1}{8 \pi \rho}\nabla B^2 \,= \, - \frac{1}{8 \pi \rho} \left(2 B_0 \partial_z \mathbf{B}\right) \, = \, - \frac{1}{4 \pi \rho} \left(B_0 \partial_z \mathbf{B}\right)$ \\
+
+We now have:
+
+\begin{equation}
+\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(v_r \mathbf{\hat{\phi}} - v_{\phi}\mathbf{\hat{r}}\right) \, = \, - r\left(\Omega^2 - \Omega_0^2\right)\mathbf{\hat{r}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  - \, \frac{1}{4 \pi \rho} \left(B_0 \partial_z \mathbf{B}\right) + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}  \, - \, \nu\nabla^2 \mathbf{u} \,
+\end{equation}
+
+The net radial body force in the rotating frame, $r\left(\Omega^2 - \Omega_0^2\right)$, is Taylor expanded as follows: \\
+
+$-r\left(\Omega^2 - \Omega_0^2\right) \,= \,-2 r_0 \Omega\left(r_0\right)\left(r - r_0\right)\left.\frac{\partial \Omega}{\partial r}\right|_{r_0} \,= \, -2 r_0 \Omega\left(r_0\right)\left(r - r_0\right)\frac{\Omega_0}{r_0} \left.\frac{\partial ln \Omega}{\partial ln r}\right|_{r_0} $ \\
+
+And the definition -$\left.\frac{\partial ln \Omega}{\partial ln r}\right|_{r_0} \equiv q$ is introduced. \\
+
+We now finally make the switch
+
 \bibliography{mri.bib}

### 2014-02-06 notes/mri_notes.tex
More added to derivations

@@
 
-The initial magnetic field is a constant vertical field $\mathbf{B} = B_0 \mathbf{\hat{z}}$, and thus the magnetic pressure term becomes: \\
+%The initial magnetic field is a constant vertical field $\mathbf{B} = B_0 \mathbf{\hat{z}}$, and thus the magnetic pressure term becomes: \\
 
-$- \frac{1}{8 \pi \rho}\nabla B^2 \,= \, - \frac{1}{8 \pi \rho} \left(2 B_0 \partial_z \mathbf{B}\right) \, = \, - \frac{1}{4 \pi \rho} \left(B_0 \partial_z \mathbf{B}\right)$ \\
+%$- \frac{1}{8 \pi \rho}\nabla B^2 \,= \, - \frac{1}{8 \pi \rho} \left(2 B_0 \partial_z \mathbf{B}\right) \, = \, - \frac{1}{4 \pi \rho} \left(B_0 \partial_z \mathbf{B}\right)$ \\
 
@@
 \begin{equation}
-\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(v_r \mathbf{\hat{\phi}} - v_{\phi}\mathbf{\hat{r}}\right) \, = \, - r\left(\Omega^2 - \Omega_0^2\right)\mathbf{\hat{r}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  - \, \frac{1}{4 \pi \rho} \left(B_0 \partial_z \mathbf{B}\right) + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}  \, - \, \nu\nabla^2 \mathbf{u} \,
+\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(v_r \mathbf{\hat{\phi}} - v_{\phi}\mathbf{\hat{r}}\right) \, = \, - r\left(\Omega^2 - \Omega_0^2\right)\mathbf{\hat{r}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  - \, \frac{1}{8 \pi \rho}\nabla B^2 + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}  \, - \, \nu\nabla^2 \mathbf{u} \,
 \end{equation}
@@
 
-And the definition -$\left.\frac{\partial ln \Omega}{\partial ln r}\right|_{r_0} \equiv q$ is introduced. \\
+And the definition -$\left.\frac{\partial ln \Omega}{\partial ln r}\right|_{r_0} \equiv q$ is introduced, so that the term becomes: \\
 
-We now finally make the switch
+$-2 r_0 \Omega\left(r_0\right)\left(r - r_0\right)\frac{\Omega_0}{r_0} \left.\frac{\partial ln \Omega}{\partial ln r}\right|_{r_0} \, = \, 2q\Omega_0^2\left(r - r_0\right)$ \\
+
+We now make the shearing box approximation -- that is, we take a small region of size $\Delta$ centered on a point $\left(r = r_0, \phi = \phi_0, z=0\right)$ in our disk.
+The box has sides of length $\Delta$: \\
+
+\noindent$r_0 - \frac{\Delta}{2} \le r \le r_0 + \frac{\Delta}{2}$ \\
+$\phi_0 - \frac{\Delta}{2 r_0} \le \phi \le \phi_0 + \frac{\Delta}{2 r_0}$ \\
+$ - \frac{\Delta}{2} \le z \le r \frac{\Delta}{2}$ \\
+
+And our momentum equation now reads: 
+\[\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(v_x \mathbf{\hat{y}} - v_{y}\mathbf{\hat{x}}\right) \, = \, 2\Omega_0^2 x q \mathbf{\hat{x}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  - \, \frac{1}{8 \pi \rho}\nabla B^2 + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}  \, - \, \nu\nabla^2 \mathbf{u} \,\]
 

### 2014-02-06 notes/mri_notes.tex
More equations added

@@
 
-Begin with the Navier-Stokes equation:
+An equation for the conservation of energy is easily derived assuming an ideal gas: \\
+
+$P = \left(\gamma - 1\right)\rho e$\\
+
+$\partial_t e + \left(\mathbf{v}\cdot \nabla\right) e + \frac{1}{\rho}P\nabla \cdot \mathbf{u} = 0$ \\
+
+Becomes: \\
+
+$\partial_t P + \left(\mathbf{u} \cdot \nabla\right)P + \left(\gamma -1\right)P\nabla \cdot \mathbf{u} = 0$ \\
+
+But in our set-up, this will become formally equivalent to the incompressible limit ($\nabla \cdot \mathbf{u} \,=\, 0$). Likewise, we can write the equation for conservation of mass: \\
+
+$\partial_t \rho + \nabla \cdot \left(\rho \mathbf{u}\right) = 0$ \\
+
+But later on, the small shearing box approximation and the assumption that the initial density perturbation is zero will render this equation unnecessary. \\
+
+So instead, we begin with the momentum equation:
 
@@
 
-The coriolis term becomes: \\
+The Coriolis term becomes: \\
 
@@
 
+Where $\delta \equiv \frac{\Delta}{r_0} \ll 1 $. \\
+
 And our momentum equation now reads: 

### 2014-02-06 notes/mri_notes.tex
Beginning of nondimensionalization

@@
 So instead, we begin with the momentum equation: 
+%\begin{equation}
+\[\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, = \, -\frac{1}{\rho}\nabla P \, - \, \nabla\Phi \, + \, \frac{1}{\rho} \left(\mathbf{J}\times\mathbf{B}\right) \, - \, \nu\nabla^2 \mathbf{u} \, - \, 2\mathbf{\Omega} \times \mathbf{u} \, - \, \mathbf{\Omega} \times \left(\mathbf{\Omega} \times \mathbf{r} \right)\]
+%\end{equation}\label{momeqn}
 
-\begin{equation}
-\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, = \, -\frac{1}{\rho}\nabla P \, - \, \nabla\Phi \, + \, \frac{1}{\rho} \left(\mathbf{J}\times\mathbf{B}\right) \, - \, \nu\nabla^2 \mathbf{u} \, - \, 2\mathbf{\Omega} \times \mathbf{u} \, - \, \mathbf{\Omega} \times \left(\mathbf{\Omega} \times \mathbf{r} \right)
-\end{equation}\label{momeqn}
+And the induction equation:
+\[\partial_t \mathbf{B} = \nabla \times \left(\mathbf{u} \times \mathbf{B}\right) + \eta\nabla^2\mathbf{B} \]
 
@@
 
-We now have:
+We now have, for the momentum equation:
 
@@
 
+\noindent We nondimensionalize as follows (see Regev \& Umurhan 2008): \\
+
+\noindent x, y, and z are scaled by $\Delta = \delta r_0$ \\
+Velocities are scaled by $\Omega_0 r_0 \delta$ \\
+Magnetic fields are scaled by $B_0$ \\
+Pressure is scaled by $\epsilon^2 \Omega_0^2 r_0^2$ \\
+
+And we introduce the following terms:
+\[R_e \equiv \frac{\Omega_0 r_0^2 \delta^2}{\nu} ; \, \, \, R_m \equiv \frac{\Omega_0 r_0^2 \delta^2}{\eta}\]
+
+And a pseudo-plasma beta parameter: 
+\[\beta \equiv \frac{\epsilon^2\Omega_0^2r_0^2}{v_A^2} \]
+
+Where $v_A$ is the Alfven speed. \\
+
+We will next take perturbations with respect to the background flow.
+
 \bibliography{mri.bib}

### 2014-02-07 notes/mri_notes.tex
Explicit nondimensionalization

@@
 
-We now make the shearing box approximation -- that is, we take a small region of size $\Delta$ centered on a point $\left(r = r_0, \phi = \phi_0, z=0\right)$ in our disk.
+We now make the shearing box approximation \-- that is, we take a small region of size $\Delta$ centered on a point $\left(r = r_0, \phi = \phi_0, z=0\right)$ in our disk.
 The box has sides of length $\Delta$: \\
@@
 And our momentum equation now reads: 
-\[\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(v_x \mathbf{\hat{y}} - v_{y}\mathbf{\hat{x}}\right) \, = \, 2\Omega_0^2 x q \mathbf{\hat{x}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  - \, \frac{1}{8 \pi \rho}\nabla B^2 + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}  \, - \, \nu\nabla^2 \mathbf{u} \,\]
+\[\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(u_x \mathbf{\hat{y}} - u_{y}\mathbf{\hat{x}}\right) \, = \, 2\Omega_0^2 x q \mathbf{\hat{x}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  - \, \frac{1}{8 \pi \rho}\nabla B^2 + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}  \, - \, \nu\nabla^2 \mathbf{u} \,\]
 
@@
 
-Where $v_A$ is the Alfven speed. \\
+Where $v_A$ is the Alfven speed, which is $v_A = \frac{B_0^2}{\rho_0}$. \\
+
+To make it perfectly clear, I will write this out explicitly.
+Dimensional quantities are denoted by a tilde superscript, and I will first rewrite the momentum equation in its dimensional quantities: \\
+
+\[\frac{\partial \mathbf{\widetilde{u}}}{\partial \widetilde{t}} \, + \, \mathbf{\widetilde{u}} \cdot \widetilde{\nabla} \mathbf{\widetilde{u}} \, + 2\Omega_0\left(\widetilde{u_x} \mathbf{\hat{y}} - \widetilde{u_{y}}\mathbf{\hat{x}}\right) \, = \, 2\Omega_0^2 \widetilde{x} q \mathbf{\hat{x}} \, -\frac{1}{\widetilde{\rho}}\widetilde{\nabla} \widetilde{P} \, +\frac{1}{\widetilde{\rho_b}}\widetilde{\nabla} \widetilde{P_b} \,  - \, \frac{1}{8 \pi \widetilde{\rho}}\widetilde{\nabla} \widetilde{B}^2 + \frac{1}{4 \pi \widetilde{\rho}} \left(\mathbf{\widetilde{B}} \cdot \widetilde{\nabla}\right)\mathbf{\widetilde{B}}  \, - \, \nu \widetilde{\nabla}^2 \mathbf{\widetilde{u}} \,\]
+
+The explicit substitutions made are as follows: \\
+
+$\widetilde{u} = \Omega_0 r_0 \delta u$ \\
+
+$\widetilde{x} = \delta r_0 x$ \\
+
+$\widetilde{\nabla} = \frac{\nabla}{\delta r_0}$ \\
+
+$\widetilde{t} = \frac{t}{\Omega_0}$ \\
+
+$\widetilde{B} = B_0 B \, \,; \,  \widetilde{P} = P_0 P \, \, ; \, \widetilde{\rho} = \rho_0 \rho $ \\
+
+The velocity scale is related to the local sound speed scale $\left(c_{s0} \equiv \sqrt{\frac{P_0}{\rho_0}}\right)$ by:\\
+
+$v_0 = \frac{\delta}{\epsilon} c_{s0}$ \\
+
+And thus: $\Omega_0 r_0 \delta u = \frac{\delta}{\epsilon} \sqrt{\frac{P_0}{\rho_0}}$ \\
+
+Making these substitutions, the dimensionless equation becomes: \\
+
+%\begin{equation}
+$\left(\Omega_0^2 r_0 \delta\right) \partial_t \mathbf{u} \, + \, \left(\Omega_0^2 r_0 \delta\right)\mathbf{u} \cdot \nabla \mathbf{u} \,+\, \left(\Omega_0^2 r_0 \delta\right) 2 u_x \mathbf{\hat{y}} \,-\, \left(\Omega_0^2 r_0 \delta\right) 2 u_y \mathbf{\hat{x}} \, = \, \left(\Omega_0^2 r_0 \delta\right) 2 x q \mathbf{\hat{x}} \,-\, \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right)\frac{1}{\rho} \nabla P  \,+\, \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right)\frac{1}{\rho_b} \nabla P_b \, - \, \frac{1}{8\pi}\Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right) \frac{1}{\beta} \frac{1}{\rho} \nabla B^2 \, + \, \frac{1}{4\pi} \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right) \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla \right)\mathbf{B} \, - \, \nu\left(\frac{\Omega_0}{r_0 \delta}\right) \nabla^2 \mathbf{u}$ \\
+%\end{equation}
+
+Dividing out a factor of $\left(\Omega_0^2 r_0 \delta\right)$ and rewriting in terms of $\beta$ and $R_e$ yields the nondimensionalized momentum equation: \\
+
+$\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + \, 2 u_x \mathbf{\hat{y}} - 2 u_y \mathbf{\hat{x}} \, = \, 2 x q \mathbf{\hat{x}} \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \nabla P \, + \,  \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \nabla P_b \, - \, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \nabla B^2 \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla \right)\mathbf{B} \, - \, \frac{1}{R_e} \nabla^2 \mathbf{u}$ \\
 

### 2014-02-07 notes/mri_notes.tex
Nondimensionalized induction eq

@@
 
+We perform the same nondimensionalization for the induction equation: \\
+
+\[\frac{\partial \widetilde{\mathbf{B}}}{\partial \widetilde{t}} \,  = \widetilde{\nabla} \times \left(\widetilde{\mathbf{u}} \times \widetilde{\mathbf{B}}\right) + \eta\widetilde{\nabla}^2 \widetilde{\mathbf{B}} \]
+
+Which expands to become:
+
+\[\frac{\partial \widetilde{\mathbf{B}}}{\partial \widetilde{t}} \, = \, \left(\widetilde{\mathbf{B}} \cdot \widetilde{\nabla} \right) \widetilde{\mathbf{u}} \, - \, \left(\widetilde{\mathbf{u}} \cdot \widetilde{\nabla} \right) \widetilde{\mathbf{B}} \, + \, \eta \widetilde{\nabla}^2\widetilde{\mathbf{B}} \] \\
+
+After using the vector identity:
+
+\[\nabla \times \left(\mathbf{u} \times \mathbf{B}\right) \, = \, \mathbf{u}\left(\nabla \cdot \mathbf{B}\right) \, - \, \mathbf{B}\left(\nabla \cdot \mathbf{u}\right) \, + \, \left(\mathbf{B} \cdot \nabla\right)\mathbf{u} \, - \, \left(\mathbf{u} \cdot \nabla\right)\mathbf{B} \] \\
+
+And applying the incompressibility and solenoidal magnetic field constraints. \\
+
+Substituting for the dimensional variables, the induction equation becomes: \\
+
+\[\left(\Omega_0 B_0\right) \partial_t\mathbf{B} \, = \, \left(\Omega_0 B_0\right)\left(\mathbf{B} \cdot \nabla\right)\mathbf{u} \, - \, \left(\Omega_0 B_0\right)\left(\mathbf{u} \cdot \nabla \right)\mathbf{B} \, + \, \frac{\eta B_0}{\delta^2 r_0^2}\mathbf{B} \]
+
+Which, after dividing by a factor of $\left(\Omega_0 B_0\right)$ and substituting in $R_m$, becomes: \\
+
+\[\partial_t\mathbf{B} \, = \, \left(\mathbf{B} \cdot \nabla \right)\mathbf{u} \, -\, \left(\mathbf{u} \cdot \nabla \right)\mathbf{B} \, + \, \frac{1}{R_m}\mathbf{B} \]
+
 We will next take perturbations with respect to the background flow.

### 2014-02-07 notes/mri_notes.tex
perturbations
more perturbations

@@
 We will next take perturbations with respect to the background flow.
+The steady solution is $\mathbf{u} = -2 q \Omega_0 \mathbf{\hat{y}}$, $P = P_b$, $\rho = \rho_b$, and $\mathbf{B} = B_0 \mathbf{\hat{z}}$, so we find perturbation solutions of the following form (1 subscripts denote perturbations): \\
+
+$\mathbf{u} \, = \, -q \Omega_0 x \mathbf{\hat{y}} \, + \, \mathbf{u_1}$ \\
+
+$\rho \, = \, \rho_b + \rho_1$ \\
+
+$P \, = \, P_b \, + \, P_1$ \\
+
+$\mathbf{B} \, = \, B_0\mathbf{\hat{z}} + \mathbf{B_1}$ \\
+
+We will begin with the momentum equation. To make this more explicit, we write out all 3 equations (pre-perturbation): \\
+
+$\partial_t u_x \, + \, \left(\mathbf{u} \cdot \nabla\right)u_x \, - \, 2 u_y \, = \, 2 x q \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_x P \, + \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_x P_b \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_x \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_x \, - \, \frac{1}{R_e}\nabla^2 u_x  $ \\
+
+$\partial_t u_y \, + \, \left(\mathbf{u} \cdot \nabla\right)u_y \, + \, 2u_x =  \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_y P \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_y \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_y \, - \, \frac{1}{R_e}\nabla^2 u_y  $ \\
+
+$\partial_t u_z \, + \, \left(\mathbf{u} \cdot \nabla\right)u_z \,  =  \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_z P \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_z \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_z \, - \, \frac{1}{R_e}\nabla^2 u_z $ \\
+
+Where we have made the substitution $\partial_y P_b = \partial_y \rho_b = 0$, as the steady barotropic solution must be axisymmetric.
 

### 2014-02-10 notes/mri_notes.tex
fixed sign error in viscous term

--- 2014-02-07 notes/mri_notes.tex
@@
 %\begin{equation}
-\[\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, = \, -\frac{1}{\rho}\nabla P \, - \, \nabla\Phi \, + \, \frac{1}{\rho} \left(\mathbf{J}\times\mathbf{B}\right) \, - \, \nu\nabla^2 \mathbf{u} \, - \, 2\mathbf{\Omega} \times \mathbf{u} \, - \, \mathbf{\Omega} \times \left(\mathbf{\Omega} \times \mathbf{r} \right)\]
+\[\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, = \, -\frac{1}{\rho}\nabla P \, - \, \nabla\Phi \, + \, \frac{1}{\rho} \left(\mathbf{J}\times\mathbf{B}\right) \, + \, \nu\nabla^2 \mathbf{u} \, - \, 2\mathbf{\Omega} \times \mathbf{u} \, - \, \mathbf{\Omega} \times \left(\mathbf{\Omega} \times \mathbf{r} \right)\]
 %\end{equation}\label{momeqn}
@@
 \begin{equation}
-\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(v_r \mathbf{\hat{\phi}} - v_{\phi}\mathbf{\hat{r}}\right) \, = \, - r\left(\Omega^2 - \Omega_0^2\right)\mathbf{\hat{r}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  - \, \frac{1}{8 \pi \rho}\nabla B^2 + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}  \, - \, \nu\nabla^2 \mathbf{u} \,
+\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(v_r \mathbf{\hat{\phi}} - v_{\phi}\mathbf{\hat{r}}\right) \, = \, - r\left(\Omega^2 - \Omega_0^2\right)\mathbf{\hat{r}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  - \, \frac{1}{8 \pi \rho}\nabla B^2 + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}  \, + \, \nu\nabla^2 \mathbf{u} \,
 \end{equation}
@@
 And our momentum equation now reads: 
-\[\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(u_x \mathbf{\hat{y}} - u_{y}\mathbf{\hat{x}}\right) \, = \, 2\Omega_0^2 x q \mathbf{\hat{x}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  - \, \frac{1}{8 \pi \rho}\nabla B^2 + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}  \, - \, \nu\nabla^2 \mathbf{u} \,\]
+\[\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(u_x \mathbf{\hat{y}} - u_{y}\mathbf{\hat{x}}\right) \, = \, 2\Omega_0^2 x q \mathbf{\hat{x}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  - \, \frac{1}{8 \pi \rho}\nabla B^2 + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}  \, + \, \nu\nabla^2 \mathbf{u} \,\]
 
@@
 
-\[\frac{\partial \mathbf{\widetilde{u}}}{\partial \widetilde{t}} \, + \, \mathbf{\widetilde{u}} \cdot \widetilde{\nabla} \mathbf{\widetilde{u}} \, + 2\Omega_0\left(\widetilde{u_x} \mathbf{\hat{y}} - \widetilde{u_{y}}\mathbf{\hat{x}}\right) \, = \, 2\Omega_0^2 \widetilde{x} q \mathbf{\hat{x}} \, -\frac{1}{\widetilde{\rho}}\widetilde{\nabla} \widetilde{P} \, +\frac{1}{\widetilde{\rho_b}}\widetilde{\nabla} \widetilde{P_b} \,  - \, \frac{1}{8 \pi \widetilde{\rho}}\widetilde{\nabla} \widetilde{B}^2 + \frac{1}{4 \pi \widetilde{\rho}} \left(\mathbf{\widetilde{B}} \cdot \widetilde{\nabla}\right)\mathbf{\widetilde{B}}  \, - \, \nu \widetilde{\nabla}^2 \mathbf{\widetilde{u}} \,\]
+\[\frac{\partial \mathbf{\widetilde{u}}}{\partial \widetilde{t}} \, + \, \mathbf{\widetilde{u}} \cdot \widetilde{\nabla} \mathbf{\widetilde{u}} \, + 2\Omega_0\left(\widetilde{u_x} \mathbf{\hat{y}} - \widetilde{u_{y}}\mathbf{\hat{x}}\right) \, = \, 2\Omega_0^2 \widetilde{x} q \mathbf{\hat{x}} \, -\frac{1}{\widetilde{\rho}}\widetilde{\nabla} \widetilde{P} \, +\frac{1}{\widetilde{\rho_b}}\widetilde{\nabla} \widetilde{P_b} \,  - \, \frac{1}{8 \pi \widetilde{\rho}}\widetilde{\nabla} \widetilde{B}^2 + \frac{1}{4 \pi \widetilde{\rho}} \left(\mathbf{\widetilde{B}} \cdot \widetilde{\nabla}\right)\mathbf{\widetilde{B}}  \, + \, \nu \widetilde{\nabla}^2 \mathbf{\widetilde{u}} \,\]
 
@@
 %\begin{equation}
-$\left(\Omega_0^2 r_0 \delta\right) \partial_t \mathbf{u} \, + \, \left(\Omega_0^2 r_0 \delta\right)\mathbf{u} \cdot \nabla \mathbf{u} \,+\, \left(\Omega_0^2 r_0 \delta\right) 2 u_x \mathbf{\hat{y}} \,-\, \left(\Omega_0^2 r_0 \delta\right) 2 u_y \mathbf{\hat{x}} \, = \, \left(\Omega_0^2 r_0 \delta\right) 2 x q \mathbf{\hat{x}} \,-\, \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right)\frac{1}{\rho} \nabla P  \,+\, \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right)\frac{1}{\rho_b} \nabla P_b \, - \, \frac{1}{8\pi}\Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right) \frac{1}{\beta} \frac{1}{\rho} \nabla B^2 \, + \, \frac{1}{4\pi} \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right) \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla \right)\mathbf{B} \, - \, \nu\left(\frac{\Omega_0}{r_0 \delta}\right) \nabla^2 \mathbf{u}$ \\
+$\left(\Omega_0^2 r_0 \delta\right) \partial_t \mathbf{u} \, + \, \left(\Omega_0^2 r_0 \delta\right)\mathbf{u} \cdot \nabla \mathbf{u} \,+\, \left(\Omega_0^2 r_0 \delta\right) 2 u_x \mathbf{\hat{y}} \,-\, \left(\Omega_0^2 r_0 \delta\right) 2 u_y \mathbf{\hat{x}} \, = \, \left(\Omega_0^2 r_0 \delta\right) 2 x q \mathbf{\hat{x}} \,-\, \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right)\frac{1}{\rho} \nabla P  \,+\, \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right)\frac{1}{\rho_b} \nabla P_b \, - \, \frac{1}{8\pi}\Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right) \frac{1}{\beta} \frac{1}{\rho} \nabla B^2 \, + \, \frac{1}{4\pi} \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right) \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla \right)\mathbf{B} \, + \, \nu\left(\frac{\Omega_0}{r_0 \delta}\right) \nabla^2 \mathbf{u}$ \\
 %\end{equation}
@@
 
-$\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + \, 2 u_x \mathbf{\hat{y}} - 2 u_y \mathbf{\hat{x}} \, = \, 2 x q \mathbf{\hat{x}} \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \nabla P \, + \,  \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \nabla P_b \, - \, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \nabla B^2 \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla \right)\mathbf{B} \, - \, \frac{1}{R_e} \nabla^2 \mathbf{u}$ \\
+$\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + \, 2 u_x \mathbf{\hat{y}} - 2 u_y \mathbf{\hat{x}} \, = \, 2 x q \mathbf{\hat{x}} \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \nabla P \, + \,  \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \nabla P_b \, - \, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \nabla B^2 \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla \right)\mathbf{B} \, + \, \frac{1}{R_e} \nabla^2 \mathbf{u}$ \\
 
@@
 We will next take perturbations with respect to the background flow.
+The steady solution is $\mathbf{u} = -2 q \Omega_0 \mathbf{\hat{y}}$, $P = P_b$, $\rho = \rho_b$, and $\mathbf{B} = B_0 \mathbf{\hat{z}}$, so we find perturbation solutions of the following form (1 subscripts denote perturbations): \\
+
+$\mathbf{u} \, = \, -q \Omega_0 x \mathbf{\hat{y}} \, + \, \mathbf{u_1}$ \\
+
+$\rho \, = \, \rho_b + \rho_1$ \\
+
+$P \, = \, P_b \, + \, P_1$ \\
+
+$\mathbf{B} \, = \, B_0\mathbf{\hat{z}} + \mathbf{B_1}$ \\
+
+We will begin with the momentum equation. To make this more explicit, we write out all 3 equations (pre-perturbation): \\
+
+$\partial_t u_x \, + \, \left(\mathbf{u} \cdot \nabla\right)u_x \, - \, 2 u_y \, = \, 2 x q \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_x P \, + \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_x P_b \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_x \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_x \, + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
+
+$\partial_t u_y \, + \, \left(\mathbf{u} \cdot \nabla\right)u_y \, + \, 2u_x =  \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_y P \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_y \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_y \, + \, \frac{1}{R_e}\nabla^2 u_y  $ \\
+
+$\partial_t u_z \, + \, \left(\mathbf{u} \cdot \nabla\right)u_z \,  =  \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_z P \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_z \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_z \, + \, \frac{1}{R_e}\nabla^2 u_z $ \\
+
+Where we have made the substitution $\partial_y P_b = \partial_y \rho_b = 0$, as the steady barotropic solution must be axisymmetric. \\
+
+These become: \\
+
+$\partial_t u_{1x} \, + \, \left(\mathbf{u} \cdot \nabla\right)u_{1x} \, + \, 2 q \Omega_0 x \, - \, 2 u_{1y}\,= \, 2 x q \, - \, \frac{\epsilon^2}{\delta^2} \left( \frac{\partial_x P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_x P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_x \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_x \, + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
 

### 2014-02-10 notes/mri_notes.tex
B perturbation

@@
 
-Where $\delta \equiv \frac{\Delta}{r_0} \ll 1 $. \\
+Where $\delta \equiv \frac{\Delta}{r_0} \ll 1 $.
+The box lives in a disk of vertical height $h_0$, parameterized by $\epsilon \equiv \frac{h_0}{r_0}$.
+The relative scaling of $\delta$ and $\epsilon$ determine whether we are in the small or large shearing box limit.\\
 
@@
 
+The term $\left(\mathbf{u} \cdot \nabla\right)\mathbf{u}$ expands as follows:\\
+
+$\left(-q \Omega_0 x \mathbf{\hat{y}} \, + \, \mathbf{u_1}\right) \cdot \left(-q \Omega_0 x \mathbf{\hat{y}} \, + \, \mathbf{u_1}\right) \, = \, - q \Omega_0 x \mathbf{\hat{y}} \cdot \nabla \mathbf{u_1} + \mathbf{u_1} \cdot \nabla \mathbf{u_1} \, = \, - q \Omega_0 x \partial_y \mathbf{u_1} + \mathbf{u_1} \cdot \nabla \mathbf{u_1}$ \\
+
+Because all of the $\nabla \left(-q \Omega_0 x \mathbf{\hat{y}}\right)$ terms are 0. \\
+
+Similarly, the magnetic tension term expands to become: \\
+
+$\left(\mathbf{B} \cdot \nabla \right)\mathbf{B} \, = \, \left(\left(B_0 \mathbf{\hat{z}} + \mathbf{B_1} \right)\cdot \nabla \right)\left(B_0 \mathbf{\hat{z}} + \mathbf{B_1} \right) \, = \, B_0 \partial_z \mathbf{B_1} \, + \, \mathbf{B_1} \cdot \nabla \mathbf{B_1}$ \\
+
+Because the $\nabla B_0 \mathbf{\hat{z}}$ terms are 0.
+
+The magnetic pressure term expands to become: \\
+
+$\nabla \left(B_0 \mathbf{\hat{z}} + \mathbf{B_1}\right)^2 $ \\
+
 These become: \\
 
-$\partial_t u_{1x} \, + \, \left(\mathbf{u} \cdot \nabla\right)u_{1x} \, + \, 2 q \Omega_0 x \, - \, 2 u_{1y}\,= \, 2 x q \, - \, \frac{\epsilon^2}{\delta^2} \left( \frac{\partial_x P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_x P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_x \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_x \, + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
+$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, + \, 2 q \Omega_0 x \, - \, 2 u_{1y}\,= \, 2 q x \, - \, \frac{\epsilon^2}{\delta^2} \left( \frac{\partial_x P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_x P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_x \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_x \, + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
 

### 2014-02-10 notes/mri_notes.tex
fixed lorentz force term

@@
 
+However, this formulation allows the magnetic pressure term to have a component parallel to the magnetic field.
+Clearly $\mathbf{J} \times \mathbf{B}$ is always perpendicular to the field.
+This will become problematic if we leave these two terms separated \-- terms will cancel in the perturbation expansion that should not.
+So for the time being, we retain the Lorentz force as $\frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B}$.
+\\
+
 %The initial magnetic field is a constant vertical field $\mathbf{B} = B_0 \mathbf{\hat{z}}$, and thus the magnetic pressure term becomes: \\
@@
 \begin{equation}
-\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(v_r \mathbf{\hat{\phi}} - v_{\phi}\mathbf{\hat{r}}\right) \, = \, - r\left(\Omega^2 - \Omega_0^2\right)\mathbf{\hat{r}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  - \, \frac{1}{8 \pi \rho}\nabla B^2 + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}  \, + \, \nu\nabla^2 \mathbf{u} \,
+\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(v_r \mathbf{\hat{\phi}} - v_{\phi}\mathbf{\hat{r}}\right) \, = \, - r\left(\Omega^2 - \Omega_0^2\right)\mathbf{\hat{r}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  + \, \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B}  \, + \, \nu\nabla^2 \mathbf{u} \,
 \end{equation}
@@
 And our momentum equation now reads: 
-\[\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(u_x \mathbf{\hat{y}} - u_{y}\mathbf{\hat{x}}\right) \, = \, 2\Omega_0^2 x q \mathbf{\hat{x}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  - \, \frac{1}{8 \pi \rho}\nabla B^2 + \frac{1}{4 \pi \rho} \left(\mathbf{B} \cdot \nabla\right)\mathbf{B}  \, + \, \nu\nabla^2 \mathbf{u} \,\]
+\[\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + 2\Omega_0\left(u_x \mathbf{\hat{y}} - u_{y}\mathbf{\hat{x}}\right) \, = \, 2\Omega_0^2 x q \mathbf{\hat{x}} \, -\frac{1}{\rho}\nabla P \, +\frac{1}{\rho_b}\nabla P_b \,  + \, \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} + \, \nu\nabla^2 \mathbf{u} \,\]
 
@@
 
-Where $v_A$ is the Alfven speed, which is $v_A = \frac{B_0^2}{\rho_0}$. \\
+Where $v_A$ is the Alfven speed, which is ${v_A}^2 = \frac{B_0^2}{\rho_0}$. \\
 
@@
 
-\[\frac{\partial \mathbf{\widetilde{u}}}{\partial \widetilde{t}} \, + \, \mathbf{\widetilde{u}} \cdot \widetilde{\nabla} \mathbf{\widetilde{u}} \, + 2\Omega_0\left(\widetilde{u_x} \mathbf{\hat{y}} - \widetilde{u_{y}}\mathbf{\hat{x}}\right) \, = \, 2\Omega_0^2 \widetilde{x} q \mathbf{\hat{x}} \, -\frac{1}{\widetilde{\rho}}\widetilde{\nabla} \widetilde{P} \, +\frac{1}{\widetilde{\rho_b}}\widetilde{\nabla} \widetilde{P_b} \,  - \, \frac{1}{8 \pi \widetilde{\rho}}\widetilde{\nabla} \widetilde{B}^2 + \frac{1}{4 \pi \widetilde{\rho}} \left(\mathbf{\widetilde{B}} \cdot \widetilde{\nabla}\right)\mathbf{\widetilde{B}}  \, + \, \nu \widetilde{\nabla}^2 \mathbf{\widetilde{u}} \,\]
+\[\frac{\partial \mathbf{\widetilde{u}}}{\partial \widetilde{t}} \, + \, \mathbf{\widetilde{u}} \cdot \widetilde{\nabla} \mathbf{\widetilde{u}} \, + 2\Omega_0\left(\widetilde{u_x} \mathbf{\hat{y}} - \widetilde{u_{y}}\mathbf{\hat{x}}\right) \, = \, 2\Omega_0^2 \widetilde{x} q \mathbf{\hat{x}} \, -\frac{1}{\widetilde{\rho}}\widetilde{\nabla} \widetilde{P} \, +\frac{1}{\widetilde{\rho_b}}\widetilde{\nabla} \widetilde{P_b} \,  + \, \frac{1}{4 \pi \widetilde{\rho}} \left( \widetilde{\nabla} \times \mathbf{\widetilde{B}} \right) \times \mathbf{\widetilde{B}}  \, + \, \nu \widetilde{\nabla}^2 \mathbf{\widetilde{u}} \,\]
 
@@
 %\begin{equation}
-$\left(\Omega_0^2 r_0 \delta\right) \partial_t \mathbf{u} \, + \, \left(\Omega_0^2 r_0 \delta\right)\mathbf{u} \cdot \nabla \mathbf{u} \,+\, \left(\Omega_0^2 r_0 \delta\right) 2 u_x \mathbf{\hat{y}} \,-\, \left(\Omega_0^2 r_0 \delta\right) 2 u_y \mathbf{\hat{x}} \, = \, \left(\Omega_0^2 r_0 \delta\right) 2 x q \mathbf{\hat{x}} \,-\, \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right)\frac{1}{\rho} \nabla P  \,+\, \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right)\frac{1}{\rho_b} \nabla P_b \, - \, \frac{1}{8\pi}\Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right) \frac{1}{\beta} \frac{1}{\rho} \nabla B^2 \, + \, \frac{1}{4\pi} \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right) \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla \right)\mathbf{B} \, + \, \nu\left(\frac{\Omega_0}{r_0 \delta}\right) \nabla^2 \mathbf{u}$ \\
+$\left(\Omega_0^2 r_0 \delta\right) \partial_t \mathbf{u} \, + \, \left(\Omega_0^2 r_0 \delta\right)\mathbf{u} \cdot \nabla \mathbf{u} \,+\, \left(\Omega_0^2 r_0 \delta\right) 2 u_x \mathbf{\hat{y}} \,-\, \left(\Omega_0^2 r_0 \delta\right) 2 u_y \mathbf{\hat{x}} \, = \, \left(\Omega_0^2 r_0 \delta\right) 2 x q \mathbf{\hat{x}} \,-\, \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right)\frac{1}{\rho} \nabla P  \,+\, \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right)\frac{1}{\rho_b} \nabla P_b \, - \, \frac{\epsilon^2 \Omega_0^2 r_0}{\beta \delta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \nu\left(\frac{\Omega_0}{r_0 \delta}\right) \nabla^2 \mathbf{u}$ \\
 %\end{equation}
@@
 
-$\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + \, 2 u_x \mathbf{\hat{y}} - 2 u_y \mathbf{\hat{x}} \, = \, 2 x q \mathbf{\hat{x}} \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \nabla P \, + \,  \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \nabla P_b \, - \, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \nabla B^2 \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla \right)\mathbf{B} \, + \, \frac{1}{R_e} \nabla^2 \mathbf{u}$ \\
+$\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + \, 2 u_x \mathbf{\hat{y}} - 2 u_y \mathbf{\hat{x}} \, = \, 2 x q \mathbf{\hat{x}} \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \nabla P \, + \,  \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \nabla P_b \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{R_e} \nabla^2 \mathbf{u}$ \\
 
@@
 
-$\partial_t u_x \, + \, \left(\mathbf{u} \cdot \nabla\right)u_x \, - \, 2 u_y \, = \, 2 x q \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_x P \, + \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_x P_b \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_x \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_x \, + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
+$\partial_t u_x \, + \, \left(\mathbf{u} \cdot \nabla\right)u_x \, - \, 2 u_y \, = \, 2 x q \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_x P \, + \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_x P_b \, +\, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
 
-$\partial_t u_y \, + \, \left(\mathbf{u} \cdot \nabla\right)u_y \, + \, 2u_x =  \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_y P \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_y \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_y \, + \, \frac{1}{R_e}\nabla^2 u_y  $ \\
+$\partial_t u_y \, + \, \left(\mathbf{u} \cdot \nabla\right)u_y \, + \, 2u_x =  \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_y P \, +\, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{R_e}\nabla^2 u_y  $ \\
 
-$\partial_t u_z \, + \, \left(\mathbf{u} \cdot \nabla\right)u_z \,  =  \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_z P \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_z \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_z \, + \, \frac{1}{R_e}\nabla^2 u_z $ \\
+$\partial_t u_z \, + \, \left(\mathbf{u} \cdot \nabla\right)u_z \,  =  \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_z P \, +\, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{R_e}\nabla^2 u_z $ \\
 
@@
 
-Similarly, the magnetic tension term expands to become: \\
+%Similarly, the magnetic tension term expands to become: \\
 
-$\left(\mathbf{B} \cdot \nabla \right)\mathbf{B} \, = \, \left(\left(B_0 \mathbf{\hat{z}} + \mathbf{B_1} \right)\cdot \nabla \right)\left(B_0 \mathbf{\hat{z}} + \mathbf{B_1} \right) \, = \, B_0 \partial_z \mathbf{B_1} \, + \, \mathbf{B_1} \cdot \nabla \mathbf{B_1}$ \\
+%$\left(\mathbf{B} \cdot \nabla \right)\mathbf{B} \, = \, \left(\left(B_0 \mathbf{\hat{z}} + \mathbf{B_1} \right)\cdot \nabla \right)\left(B_0 \mathbf{\hat{z}} + \mathbf{B_1} \right) \, = \, B_0 \partial_z \mathbf{B_1} \, + \, \mathbf{B_1} \cdot \nabla \mathbf{B_1}$ \\
 
-Because the $\nabla B_0 \mathbf{\hat{z}}$ terms are 0.
+%Because the $\nabla B_0 \mathbf{\hat{z}}$ terms are 0.
 
-The magnetic pressure term expands to become: \\
+%The magnetic pressure term expands to become: \\
 
-$\nabla \left(B_0 \mathbf{\hat{z}} + \mathbf{B_1}\right)^2 $ \\
+%$\nabla \left(B_0 \mathbf{\hat{z}} + \mathbf{B_1}\right)^2 $ \\
+
+The Lorentz force term $\frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B}$ expands as follows: \\
+
+$ \left( \nabla \times \left(B_0 \mathbf{\hat{z}} + \mathbf{B_1}\right) \right) \times \left(B_0 \mathbf{\hat{z}} + \mathbf{B_1}\right) \, = \, \left( \nabla \times B_0 \mathbf{\hat{z}} + \nabla \times \mathbf{B_1} \right) \times \left(B_0 \mathbf{\hat{z}} + \mathbf{B_1}\right) \, $ \\
+
+$= \, \left( \nabla \times B_0 \mathbf{\hat{z}} + \nabla \times \mathbf{B_1} \right) \times B_0 \mathbf{\hat{z}} + \left( \nabla \times B_0 \mathbf{\hat{z}} + \nabla \times \mathbf{B_1} \right) \times \mathbf{B_1} \, $ \\
+
+$= \, \nabla \times B_0 \mathbf{\hat{z}} \times B_0 \mathbf{\hat{z}} \, +\, \nabla \times \mathbf{B_1} \times B_0 \mathbf{\hat{z}} \, + \, \nabla \times B_0 \mathbf{\hat{z}} \times \mathbf{B_1} \, + \, \nabla \times \mathbf{B_1} \times \mathbf{B_1}$\\
+
+$= \, -\frac{\nabla B_0^2}{2} \, + \, B_0 \mathbf{\hat{z}} \cdot \nabla \left(B_0 \mathbf{\hat{z}}\right) \, - \, \nabla\left(B_0 \mathbf{\hat{z}} \cdot \mathbf{B_1}\right) \, + \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, + \, \mathbf{B_1} \cdot \nabla B_0 \mathbf{\hat{z}} \, + \, \mathbf{B_1} \times \nabla \times B_0 \mathbf{\hat{z}} \, + \, \nabla \times B_0 \mathbf{\hat{z}} \times \mathbf{B_1} \, - \, \frac{\nabla \mathbf{B_1}^2}{2} \, + \, \mathbf{B_1} \cdot \nabla \mathbf{B_1}$ \\
+
+Now the $\nabla B_0 \mathbf{\hat{z}}$ terms are zero, and $\mathbf{B_1} \times \nabla \times B_0 \mathbf{\hat{z}} \, = \, - \nabla \times B_0 \mathbf{\hat{z}} \times \mathbf{B_1}$, so those two terms cancel one another.
+So we have: \\
+
+$ = \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(-\frac{\nabla B_0^2}{2} \, - \, \nabla \left(B_0 \mathbf{\hat{z}}\cdot \mathbf{B_1}\right)\, + \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
+
+And finally, the Lorentz force term becomes: \\
+
+$\frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, = \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(-\frac{\nabla B_0^2}{2} \,  - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
 

### 2014-02-12 notes/mri_notes.tex
fixed lorentz force term
more
better pressure scaling explanation

@@
 
-And thus: $\Omega_0 r_0 \delta u = \frac{\delta}{\epsilon} \sqrt{\frac{P_0}{\rho_0}}$ \\
+And thus: $\Omega_0 r_0 \delta = \frac{\delta}{\epsilon} \sqrt{\frac{P_0}{\rho_0}}$ \\
+
+Why is this a reasonable choice for the pressure scaling?
+The disk is vertically pressure-supported.
+The height of the disk is determined by the sound speed and gravity.
+At the radius of our box, $r_0$, the vertical scale is $h_0 = c_s/\Omega_0$.
+Thus $\epsilon r_0 = c_s/\Omega_0$ and $\epsilon$ is small if the Keplerian velocity is highly supersonic.
+This scaling allows us to substitute out the sound speed, which is meaningless in an incompressible flow anyway.
+\\
 
@@
 
-$\partial_t u_z \, + \, \left(\mathbf{u} \cdot \nabla\right)u_z \,  =  \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_z P \, +\, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{R_e}\nabla^2 u_z $ \\
+$\partial_t u_z \, + \, \left(\mathbf{u} \cdot \nabla\right)u_z \,  =  \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_z P \,  + \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_z P_b \, +\, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{R_e}\nabla^2 u_z $ \\
 
@@
 
+$ = \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(-\frac{\nabla B_0^2}{2} \, - \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, + \, B_0 \cdot \nabla \mathbf{B_1} \,+ \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
+
 And finally, the Lorentz force term becomes: \\
 
-$\frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, = \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(-\frac{\nabla B_0^2}{2} \,  - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
+$\frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, = \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
 
-These become: \\
+Which simplifies to: \\
 
-$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, + \, 2 q \Omega_0 x \, - \, 2 u_{1y}\,= \, 2 q x \, - \, \frac{\epsilon^2}{\delta^2} \left( \frac{\partial_x P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_x P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, -\, \frac{1}{8\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \partial_x \left(\mathbf{B}^2\right) \, + \, \frac{1}{4\pi} \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{\rho} \left(\mathbf{B} \cdot \nabla\right) B_x \, + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
+$\frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, = \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(B_0 \partial_z \mathbf{B_1} - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
+
+The three momentum equations become: \\
+
+$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, + \, 2 q \Omega_0 x \, - \, 2 u_{1y}\,= \, 2 q x \, - \, \frac{\epsilon^2}{\delta^2} \left( \frac{\partial_x P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_x P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1x} \,- \, \frac{B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
+
+$\partial_t u_{1y} \, - \, q \Omega_0 x \partial_y u_{1y} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1y} \,- \, 2 q \Omega_0 x \, + \, 2u_{1y} \, =  $\\
+$\, - \, \frac{\epsilon^2}{\delta^2} \left( \frac{\partial_y P_1}{\rho_1 + \rho_b} \right) \, + \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1y} \,- \, \frac{B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{R_e}\nabla^2 u_y  $ \\
+
+$\partial_t u_{1z} \, - \, q \Omega_0 x \partial_y u_{1z} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1z} \, = $\\
+$\, - \, \frac{\epsilon^2}{\delta^2} \left( \frac{\partial_z P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_z P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1z} \,- \, \frac{B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
+
+In the small shearing box (SSB) limit, $\delta \ll \epsilon \ll 1$. For a rotation law that is close to Keplerian,
 

### 2014-02-12 notes/mri_notes.tex
more scalings
fixed accidentally dropped term

@@
 
-$\partial_t u_{1y} \, - \, q \Omega_0 x \partial_y u_{1y} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1y} \,- \, 2 q \Omega_0 x \, + \, 2u_{1y} \, =  $\\
+$\partial_t u_{1y} \, - \, q \Omega_0 u_{1x} \, - \, q \Omega_0 x \partial_y u_{1y} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1y} \,+ \, 2u_{1x} \, =  $\\
 $\, - \, \frac{\epsilon^2}{\delta^2} \left( \frac{\partial_y P_1}{\rho_1 + \rho_b} \right) \, + \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1y} \,- \, \frac{B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{R_e}\nabla^2 u_y  $ \\
@@
 
-In the small shearing box (SSB) limit, $\delta \ll \epsilon \ll 1$. For a rotation law that is close to Keplerian,
+In the small shearing box (SSB) limit, $\delta \ll \epsilon \ll 1$. For a rotation law that is close to Keplerian, $\frac{1}{\rho_b}\partial_xP_b = \mathcal{O}\left(\delta^3/\epsilon^2\right). $
+
+The following replacements are made: \\
+
+\noindent $\partial_i P_b$ becomes $\frac{\delta^2}{\epsilon^2} \partial_i P_b $ \\
+$\partial_i \rho_b$ becomes $\frac{\delta^2}{\epsilon^2} \partial_i \rho_b $ \\
+
+The pressure perturbation can be expanded: \\
+
+$P_1 = P_0 + \frac{\delta^2}{\epsilon^2}\partial_i p$ where $P_0$ is a constant that gets absorbed into $P_b$. \\
+
+This is equivalent to taking $\delta = \epsilon^2$. Making these substitutions and taking $\rho_1 = 0$, we obtain: \\
+
+$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, - \, 2 u_{1y}\,= \, - \, \frac{1}{\rho_b}\partial_x p  \, + \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1x} \,- \, \frac{B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
+
+$\partial_t u_{1y}\, - \, q \Omega_0 u_{1x}  \, - \, q \Omega_0 x \partial_y u_{1y} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1y} \,+ \, 2u_{1x} \, =  $\\
+$\, - \, \frac{1}{\rho_b}\partial_y p \, + \,  \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1y} \,- \, \frac{B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{R_e}\nabla^2 u_y  $ \\
+
+$\partial_t u_{1z} \, - \, q \Omega_0 x \partial_y u_{1z} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1z} \, = \, - \, \frac{1}{\rho_b}\partial_z p  \, + \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1z} \,- \, \frac{B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
+
+Now we reparameterize in terms of the total (gas and magnetic) pressure: \\
+
+$\Pi \equiv p + \frac{1}{2\beta} B^2$ \\
+
+And take $\rho_b$ = 1. And finally, we have the momentum equation: \\
+
+$\partial_t \mathbf{u_1} \, - \, q\Omega_0 x \partial_y \mathbf{u_1} \, + \, \left(\mathbf{u_1}\cdot \nabla\right) \mathbf{u_1} \, - \, 2 u_{1y} \mathbf{\hat{x}} + $
 

### 2014-02-13 notes/mri_notes.tex
end of derivation
added formal def of incompressible limit

@@
 
-And take $\rho_b$ = 1. And finally, we have the momentum equation: \\
+And take $\rho_b$ = 1. And finally, we have the perturbed, nondimensionalized momentum equation in all its glory: \\
 
-$\partial_t \mathbf{u_1} \, - \, q\Omega_0 x \partial_y \mathbf{u_1} \, + \, \left(\mathbf{u_1}\cdot \nabla\right) \mathbf{u_1} \, - \, 2 u_{1y} \mathbf{\hat{x}} + $
+$\partial_t \mathbf{u_1} \, - \, q\Omega_0 x \partial_y \mathbf{u_1} \, + \, \left(\mathbf{u_1}\cdot \nabla\right) \mathbf{u_1} \, - \, 2 u_{1y} \mathbf{\hat{x}} \, + \, \left(2 - \Omega_0 q\right)u_{1x}\mathbf{\hat{y}} \, = \, \frac{1}{4\pi} \nabla \Pi \, + \, \frac{1}{4\pi}B_0\partial_z\mathbf{B_1} \, + \, \frac{1}{4\pi}\mathbf{B_1} \cdot \nabla \mathbf{B_1}$ \\
+
+We now formally show what happens to the energy equation. If we perform the same steps as above on our energy equation: \\
+
+$\frac{\partial}{\partial\widetilde{t}} \widetilde{P} + \left(\widetilde{\mathbf{u}} \cdot \widetilde{\nabla} \right)\widetilde{P} + \left(\gamma -1\right)\widetilde{P} \widetilde{\nabla} \cdot \widetilde{\mathbf{u}} = 0$ \\
+
+That is, nondimensionalize: \\
+
+$\Omega_0^3 r_0^2 \epsilon^2 \rho_0 \frac{\partial}{\partial t} P \, + \, \Omega_0^3 r_0^2 \epsilon^2 \rho_0 \left(\mathbf{u} \cdot \nabla \right) P \, + \, \Omega_0^3 r_0^2 \epsilon^2 \rho_0 \left(\gamma - 1 \right) P \nabla \cdot \mathbf{u} \, = \, 0 $ \\
+
+$\partial_t P \, + \, \left(\mathbf{u} \cdot \nabla \right) P \, + \, \left(\gamma - 1 \right) P \nabla \cdot \mathbf{u} \, = \, 0 $ \\
+
+Take perturbations: \\
+
+$\partial_t P_1 \, - \, q\Omega_0 x \partial_y P_1 \, + \, u_{1x}\partial_x P_b \, + \, u_{z1}\partial_z P_b \, + \, \left(\gamma - 1 \right) P_b \nabla \cdot \mathbf{u_1} \, + \, \mathbf{u_1} \cdot \nabla P_1 \, \, + \, \left(\gamma - 1\right)P_1 \nabla \cdot \mathbf{u_1} \, = \, 0$ \\
+
+Substitute in $P_1 = \frac{\delta^2}{\epsilon^2} p$ and $\partial_i \rho_b$ = $\frac{\delta^2}{\epsilon^2} \partial_i \rho_b$ and $\partial_i P_b$ = $\frac{\delta^2}{\epsilon^2} \partial_i P_b$: \\
+
+$\frac{\delta^2}{\epsilon^2}\partial_t p \, - \, \frac{\delta^2}{\epsilon^2}q\Omega_0 x \partial_y p \, + \, \frac{\delta^2}{\epsilon^2}u_{1x}\partial_x P_b \, + \, \frac{\delta^2}{\epsilon^2}u_{z1}\partial_z P_b \, + \, \left(\gamma - 1 \right) P_b \nabla \cdot \mathbf{u_1} \, + \, \frac{\delta^2}{\epsilon^2}\mathbf{u_1} \cdot \nabla p \, \, + \, \frac{\delta^2}{\epsilon^2}\left(\gamma - 1\right)p \nabla \cdot \mathbf{u_1} \, = \, 0$ \\
+
+Drop terms of $\mathcal{O}\left(\frac{\delta^2}{\epsilon^2}\right)$, and we recover:\\
+
+$\nabla \cdot \mathbf{u_1} \, = \, 0$ \\
+
+Thus we have formally proven that we are working in the incompressible limit!
 

### 2014-02-17 notes/mri_notes.tex
full equation set derived
changed small parameter notation

--- 2014-02-12 notes/mri_notes.tex
@@
 \newcommand\nat{ Nature }
+
+\newcommand\reye{\mathrel{Re}}
 
@@
 Where $\delta \equiv \frac{\Delta}{r_0} \ll 1 $.
-The box lives in a disk of vertical height $h_0$, parameterized by $\epsilon \equiv \frac{h_0}{r_0}$.
-The relative scaling of $\delta$ and $\epsilon$ determine whether we are in the small or large shearing box limit.\\
+The box lives in a disk of vertical height $h_0$, parameterized by $\varepsilon \equiv \frac{h_0}{r_0}$.
+The relative scaling of $\delta$ and $\varepsilon$ determine whether we are in the small or large shearing box limit.\\
 
@@
 Magnetic fields are scaled by $B_0$ \\
-Pressure is scaled by $\epsilon^2 \Omega_0^2 r_0^2$ \\
+Pressure is scaled by $\varepsilon^2 \Omega_0^2 r_0^2$ \\
 
@@
 And a pseudo-plasma beta parameter: 
-\[\beta \equiv \frac{\epsilon^2\Omega_0^2r_0^2}{v_A^2} \]
+\[\beta \equiv \frac{\varepsilon^2\Omega_0^2r_0^2}{v_A^2} \]
 
@@
 
-$v_0 = \frac{\delta}{\epsilon} c_{s0}$ \\
+$v_0 = \frac{\delta}{\varepsilon} c_{s0}$ \\
 
-And thus: $\Omega_0 r_0 \delta = \frac{\delta}{\epsilon} \sqrt{\frac{P_0}{\rho_0}}$ \\
+And thus: $\Omega_0 r_0 \delta = \frac{\delta}{\varepsilon} \sqrt{\frac{P_0}{\rho_0}}$ \\
 
@@
 At the radius of our box, $r_0$, the vertical scale is $h_0 = c_s/\Omega_0$.
-Thus $\epsilon r_0 = c_s/\Omega_0$ and $\epsilon$ is small if the Keplerian velocity is highly supersonic.
+Thus $\varepsilon r_0 = c_s/\Omega_0$ and $\varepsilon$ is small if the Keplerian velocity is highly supersonic.
 This scaling allows us to substitute out the sound speed, which is meaningless in an incompressible flow anyway.
@@
 %\begin{equation}
-$\left(\Omega_0^2 r_0 \delta\right) \partial_t \mathbf{u} \, + \, \left(\Omega_0^2 r_0 \delta\right)\mathbf{u} \cdot \nabla \mathbf{u} \,+\, \left(\Omega_0^2 r_0 \delta\right) 2 u_x \mathbf{\hat{y}} \,-\, \left(\Omega_0^2 r_0 \delta\right) 2 u_y \mathbf{\hat{x}} \, = \, \left(\Omega_0^2 r_0 \delta\right) 2 x q \mathbf{\hat{x}} \,-\, \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right)\frac{1}{\rho} \nabla P  \,+\, \Omega_0^2 r_0 \left(\frac{\epsilon^2}{\delta}\right)\frac{1}{\rho_b} \nabla P_b \, - \, \frac{\epsilon^2 \Omega_0^2 r_0}{\beta \delta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \nu\left(\frac{\Omega_0}{r_0 \delta}\right) \nabla^2 \mathbf{u}$ \\
+$\left(\Omega_0^2 r_0 \delta\right) \partial_t \mathbf{u} \, + \, \left(\Omega_0^2 r_0 \delta\right)\mathbf{u} \cdot \nabla \mathbf{u} \,+\, \left(\Omega_0^2 r_0 \delta\right) 2 u_x \mathbf{\hat{y}} \,-\, \left(\Omega_0^2 r_0 \delta\right) 2 u_y \mathbf{\hat{x}} \, = \, \left(\Omega_0^2 r_0 \delta\right) 2 x q \mathbf{\hat{x}} \,-\, \Omega_0^2 r_0 \left(\frac{\varepsilon^2}{\delta}\right)\frac{1}{\rho} \nabla P  \,+\, \Omega_0^2 r_0 \left(\frac{\varepsilon^2}{\delta}\right)\frac{1}{\rho_b} \nabla P_b \, - \, \frac{\varepsilon^2 \Omega_0^2 r_0}{\beta \delta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \nu\left(\frac{\Omega_0}{r_0 \delta}\right) \nabla^2 \mathbf{u}$ \\
 %\end{equation}
@@
 
-$\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + \, 2 u_x \mathbf{\hat{y}} - 2 u_y \mathbf{\hat{x}} \, = \, 2 x q \mathbf{\hat{x}} \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \nabla P \, + \,  \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \nabla P_b \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{R_e} \nabla^2 \mathbf{u}$ \\
+$\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + \, 2 u_x \mathbf{\hat{y}} - 2 u_y \mathbf{\hat{x}} \, = \, 2 x q \mathbf{\hat{x}} \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho} \nabla P \, + \,  \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho_b} \nabla P_b \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{\reye} \nabla^2 \mathbf{u}$ \\
 
@@
 
-\[\left(\Omega_0 B_0\right) \partial_t\mathbf{B} \, = \, \left(\Omega_0 B_0\right)\left(\mathbf{B} \cdot \nabla\right)\mathbf{u} \, - \, \left(\Omega_0 B_0\right)\left(\mathbf{u} \cdot \nabla \right)\mathbf{B} \, + \, \frac{\eta B_0}{\delta^2 r_0^2}\mathbf{B} \]
+\[\left(\Omega_0 B_0\right) \partial_t\mathbf{B} \, = \, \left(\Omega_0 B_0\right)\left(\mathbf{B} \cdot \nabla\right)\mathbf{u} \, - \, \left(\Omega_0 B_0\right)\left(\mathbf{u} \cdot \nabla \right)\mathbf{B} \, + \, \frac{\eta B_0}{\delta^2 r_0^2}\nabla^2\mathbf{B} \]
 
@@
 
-\[\partial_t\mathbf{B} \, = \, \left(\mathbf{B} \cdot \nabla \right)\mathbf{u} \, -\, \left(\mathbf{u} \cdot \nabla \right)\mathbf{B} \, + \, \frac{1}{R_m}\mathbf{B} \]
+\[\partial_t\mathbf{B} \, = \, \left(\mathbf{B} \cdot \nabla \right)\mathbf{u} \, -\, \left(\mathbf{u} \cdot \nabla \right)\mathbf{B} \, + \, \frac{1}{R_m}\nabla^2\mathbf{B} \]
 
@@
 
-$\partial_t u_x \, + \, \left(\mathbf{u} \cdot \nabla\right)u_x \, - \, 2 u_y \, = \, 2 x q \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_x P \, + \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_x P_b \, +\, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
+$\partial_t u_x \, + \, \left(\mathbf{u} \cdot \nabla\right)u_x \, - \, 2 u_y \, = \, 2 x q \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho} \partial_x P \, + \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_x P_b \, +\, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
 
-$\partial_t u_y \, + \, \left(\mathbf{u} \cdot \nabla\right)u_y \, + \, 2u_x =  \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_y P \, +\, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{R_e}\nabla^2 u_y  $ \\
+$\partial_t u_y \, + \, \left(\mathbf{u} \cdot \nabla\right)u_y \, + \, 2u_x =  \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho} \partial_y P \, +\, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
 
-$\partial_t u_z \, + \, \left(\mathbf{u} \cdot \nabla\right)u_z \,  =  \, - \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho} \partial_z P \,  + \, \frac{\epsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_z P_b \, +\, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{R_e}\nabla^2 u_z $ \\
+$\partial_t u_z \, + \, \left(\mathbf{u} \cdot \nabla\right)u_z \,  =  \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho} \partial_z P \,  + \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_z P_b \, +\, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{\reye}\nabla^2 u_z $ \\
 
@@
 
-The Lorentz force term $\frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B}$ expands as follows: \\
+The Lorentz force term $\frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B}$ expands as follows: \\
 
@@
 
-$ = \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(-\frac{\nabla B_0^2}{2} \, - \, \nabla \left(B_0 \mathbf{\hat{z}}\cdot \mathbf{B_1}\right)\, + \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
+$ = \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(-\frac{\nabla B_0^2}{2} \, - \, \nabla \left(B_0 \mathbf{\hat{z}}\cdot \mathbf{B_1}\right)\, + \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
 
-$ = \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(-\frac{\nabla B_0^2}{2} \, - \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, + \, B_0 \cdot \nabla \mathbf{B_1} \,+ \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
+$ = \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(-\frac{\nabla B_0^2}{2} \, - \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, + \, B_0 \cdot \nabla \mathbf{B_1} \,+ \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
 
@@
 
-$\frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, = \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
+$\frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, = \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
 
@@
 
-$\frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, = \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(B_0 \partial_z \mathbf{B_1} - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
+$\frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, = \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(B_0 \partial_z \mathbf{B_1} - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
 
@@
 
-$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, + \, 2 q \Omega_0 x \, - \, 2 u_{1y}\,= \, 2 q x \, - \, \frac{\epsilon^2}{\delta^2} \left( \frac{\partial_x P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_x P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1x} \,- \, \frac{B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
+$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, + \, 2 q \Omega_0 x \, - \, 2 u_{1y}\,= \, 2 q x \, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_x P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_x P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1x} \,- \, \frac{B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
 
-$\partial_t u_{1y} \, - \, q \Omega_0 x \partial_y u_{1y} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1y} \,- \, 2 q \Omega_0 x \, + \, 2u_{1y} \, =  $\\
-$\, - \, \frac{\epsilon^2}{\delta^2} \left( \frac{\partial_y P_1}{\rho_1 + \rho_b} \right) \, + \, \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1y} \,- \, \frac{B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{R_e}\nabla^2 u_y  $ \\
+$\partial_t u_{1y} \, - \, q \Omega_0 u_{1x} \, - \, q \Omega_0 x \partial_y u_{1y} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1y} \,+ \, 2u_{1x} \, =  $\\
+$\, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_y P_1}{\rho_1 + \rho_b} \right) \, + \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1y} \,- \, \frac{B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
 
 $\partial_t u_{1z} \, - \, q \Omega_0 x \partial_y u_{1z} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1z} \, = $\\
-$\, - \, \frac{\epsilon^2}{\delta^2} \left( \frac{\partial_z P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_z P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\epsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1z} \,- \, \frac{B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{R_e}\nabla^2 u_x  $ \\
+$\, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_z P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_z P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1z} \,- \, \frac{B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
 
-In the small shearing box (SSB) limit, $\delta \ll \epsilon \ll 1$. For a rotation law that is close to Keplerian,
+In the small shearing box (SSB) limit, $\delta \ll \varepsilon \ll 1$. For a rotation law that is close to Keplerian, $\frac{1}{\rho_b}\partial_xP_b = \mathcal{O}\left(\delta^3/\varepsilon^2\right). $
+
+The following replacements are made: \\
+
+\noindent $\partial_i P_b$ becomes $\frac{\delta^2}{\varepsilon^2} \partial_i P_b $ \\
+$\partial_i \rho_b$ becomes $\frac{\delta^2}{\varepsilon^2} \partial_i \rho_b $ \\
+
+The pressure perturbation can be expanded: \\
+
+$P_1 = P_0 + \frac{\delta^2}{\varepsilon^2}\partial_i p$ where $P_0$ is a constant that gets absorbed into $P_b$. \\
+
+This is equivalent to taking $\delta = \varepsilon^2$. Making these substitutions and taking $\rho_1 = 0$, we obtain: \\
+
+$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, - \, 2 u_{1y}\,= \, - \, \frac{1}{\rho_b}\partial_x p  \, + \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1x} \,- \, \frac{B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
+
+$\partial_t u_{1y}\, - \, q \Omega_0 u_{1x}  \, - \, q \Omega_0 x \partial_y u_{1y} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1y} \,+ \, 2u_{1x} \, =  $\\
+$\, - \, \frac{1}{\rho_b}\partial_y p \, + \,  \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1y} \,- \, \frac{B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
+
+$\partial_t u_{1z} \, - \, q \Omega_0 x \partial_y u_{1z} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1z} \, = \, - \, \frac{1}{\rho_b}\partial_z p  \, + \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1z} \,- \, \frac{B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
+
+Now we reparameterize in terms of the total (gas and magnetic) pressure: \\
+
+$\Pi \equiv p + \frac{1}{2\beta} B^2$ \\
+
+And take $\rho_b$ = 1. And finally, we have the perturbed, nondimensionalized momentum equation in all its glory: \\
+
+$\partial_t \mathbf{u_1} \, - \, q\Omega_0 x \partial_y \mathbf{u_1} \, + \, \left(\mathbf{u_1}\cdot \nabla\right) \mathbf{u_1} \, - \, 2 u_{1y} \mathbf{\hat{x}} \, + \, \left(2 - \Omega_0 q\right)u_{1x}\mathbf{\hat{y}} \, = \, \frac{1}{4\pi} \nabla \Pi \, + \, \frac{1}{4\pi}B_0\partial_z\mathbf{B_1} \, + \, \frac{1}{4\pi}\mathbf{B_1} \cdot \nabla \mathbf{B_1} \,+ \, \frac{1}{\reye}\nabla^2 \mathbf{u_1}$ \\
+
+We now formally show what happens to the energy equation. If we perform the same steps as above on our energy equation: \\
+
+$\frac{\partial}{\partial\widetilde{t}} \widetilde{P} + \left(\widetilde{\mathbf{u}} \cdot \widetilde{\nabla} \right)\widetilde{P} + \left(\gamma -1\right)\widetilde{P} \widetilde{\nabla} \cdot \widetilde{\mathbf{u}} = 0$ \\
+
+That is, nondimensionalize: \\
+
+$\Omega_0^3 r_0^2 \varepsilon^2 \rho_0 \frac{\partial}{\partial t} P \, + \, \Omega_0^3 r_0^2 \varepsilon^2 \rho_0 \left(\mathbf{u} \cdot \nabla \right) P \, + \, \Omega_0^3 r_0^2 \varepsilon^2 \rho_0 \left(\gamma - 1 \right) P \nabla \cdot \mathbf{u} \, = \, 0 $ \\
+
+$\partial_t P \, + \, \left(\mathbf{u} \cdot \nabla \right) P \, + \, \left(\gamma - 1 \right) P \nabla \cdot \mathbf{u} \, = \, 0 $ \\
+
+Take perturbations: \\
+
+$\partial_t P_1 \, - \, q\Omega_0 x \partial_y P_1 \, + \, u_{1x}\partial_x P_b \, + \, u_{z1}\partial_z P_b \, + \, \left(\gamma - 1 \right) P_b \nabla \cdot \mathbf{u_1} \, + \, \mathbf{u_1} \cdot \nabla P_1 \, \, + \, \left(\gamma - 1\right)P_1 \nabla \cdot \mathbf{u_1} \, = \, 0$ \\
+
+Substitute in $P_1 = \frac{\delta^2}{\varepsilon^2} p$ and $\partial_i \rho_b$ = $\frac{\delta^2}{\varepsilon^2} \partial_i \rho_b$ and $\partial_i P_b$ = $\frac{\delta^2}{\varepsilon^2} \partial_i P_b$: \\
+
+$\frac{\delta^2}{\varepsilon^2}\partial_t p \, - \, \frac{\delta^2}{\varepsilon^2}q\Omega_0 x \partial_y p \, + \, \frac{\delta^2}{\varepsilon^2}u_{1x}\partial_x P_b \, + \, \frac{\delta^2}{\varepsilon^2}u_{z1}\partial_z P_b \, + \, \left(\gamma - 1 \right) P_b \nabla \cdot \mathbf{u_1} \, + \, \frac{\delta^2}{\varepsilon^2}\mathbf{u_1} \cdot \nabla p \, \, + \, \frac{\delta^2}{\varepsilon^2}\left(\gamma - 1\right)p \nabla \cdot \mathbf{u_1} \, = \, 0$ \\
+
+Drop terms of $\mathcal{O}\left(\frac{\delta^2}{\varepsilon^2}\right)$, and we recover:\\
+
+$\nabla \cdot \mathbf{u_1} \, = \, 0$ \\
+
+Thus we have formally proven that we are working in the incompressible limit! \\
+
+As for the induction equation, we already nondimensionalized to obtain: \\
+
+$\partial_t\mathbf{B} \, = \, \left(\mathbf{B} \cdot \nabla \right)\mathbf{u} \, -\, \left(\mathbf{u} \cdot \nabla \right)\mathbf{B} \, + \, \frac{1}{R_m}\nabla^2\mathbf{B} $ \\
+
+If we plug in perturbations and all terms involving $\nabla\left(-q\Omega_0 x \mathbf{\hat{y}}\right)$ or $\nabla B_0 \mathbf{\hat{z}}$ go to zero, we arrive at: \\
+
+$\partial_t \mathbf{B_1} \, = \, B_0 \partial_z \mathbf{u_1} \, + \, \mathbf{B_1} \cdot \nabla \mathbf{u_1} \, + \, q \Omega_0 x \partial_y \mathbf{B_1} \, - \, \mathbf{u_1} \cdot \nabla \mathbf{B_1} \, + \, \frac{1}{R_m} \nabla^2 \mathbf{B_1} $ \\
+
+And finally we have our full equation set: \\
+
+$\partial_t \mathbf{u_1} \, - \, q\Omega_0 x \partial_y \mathbf{u_1} \, + \, \left(\mathbf{u_1}\cdot \nabla\right) \mathbf{u_1} \, - \, 2 u_{1y} \mathbf{\hat{x}} \, + \, \left(2 - \Omega_0 q\right)u_{1x}\mathbf{\hat{y}} \, = \, \frac{1}{4\pi} \nabla \Pi \, + \, \frac{1}{4\pi}B_0\partial_z\mathbf{B_1} \, + \, \frac{1}{4\pi}\mathbf{B_1} \cdot \nabla \mathbf{B_1}\,  + \, \frac{1}{\reye}\nabla^2 \mathbf{u_1}$ \\
+
+$\partial_t \mathbf{B_1} \, = \, B_0 \partial_z \mathbf{u_1} \, + \, \mathbf{B_1} \cdot \nabla \mathbf{u_1} \, + \, q \Omega_0 x \partial_y \mathbf{B_1} \, - \, \mathbf{u_1} \cdot \nabla \mathbf{B_1} \, + \, \frac{1}{R_m} \nabla^2 \mathbf{B_1} $ \\
+
+$\nabla \cdot \mathbf{u_1} \, = \, 0$\\
+
+$\nabla \cdot \mathbf{B_1} \, = \, 0$ \\
 

### 2014-02-17 notes/mri_notes.tex
inserted missing nablas

@@
 \newcommand\reye{\mathrel{Re}}
+\newcommand\reym{\mathrel{Rm}}
 
@@
 And we introduce the following terms:
-\[R_e \equiv \frac{\Omega_0 r_0^2 \delta^2}{\nu} ; \, \, \, R_m \equiv \frac{\Omega_0 r_0^2 \delta^2}{\eta}\]
+\[\reye \equiv \frac{\Omega_0 r_0^2 \delta^2}{\nu} ; \, \, \, \reym \equiv \frac{\Omega_0 r_0^2 \delta^2}{\eta}\]
 
@@
 
-Dividing out a factor of $\left(\Omega_0^2 r_0 \delta\right)$ and rewriting in terms of $\beta$ and $R_e$ yields the nondimensionalized momentum equation: \\
+Dividing out a factor of $\left(\Omega_0^2 r_0 \delta\right)$ and rewriting in terms of $\beta$ and $\reye$ yields the nondimensionalized momentum equation: \\
 
@@
 
-Which, after dividing by a factor of $\left(\Omega_0 B_0\right)$ and substituting in $R_m$, becomes: \\
+Which, after dividing by a factor of $\left(\Omega_0 B_0\right)$ and substituting in $\reym$, becomes: \\
 
-\[\partial_t\mathbf{B} \, = \, \left(\mathbf{B} \cdot \nabla \right)\mathbf{u} \, -\, \left(\mathbf{u} \cdot \nabla \right)\mathbf{B} \, + \, \frac{1}{R_m}\nabla^2\mathbf{B} \]
+\[\partial_t\mathbf{B} \, = \, \left(\mathbf{B} \cdot \nabla \right)\mathbf{u} \, -\, \left(\mathbf{u} \cdot \nabla \right)\mathbf{B} \, + \, \frac{1}{\reym}\nabla^2\mathbf{B} \]
 
@@
 
-$ = \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(-\frac{\nabla B_0^2}{2} \, - \, \nabla \left(B_0 \mathbf{\hat{z}}\cdot \mathbf{B_1}\right)\, + \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
+$ = \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(-\frac{\nabla B_0^2}{2} \, - \, \nabla \left(B_0 \mathbf{\hat{z}}\cdot \mathbf{B_1}\right)\, + \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, - \, \frac{\nabla \mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
 
-$ = \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(-\frac{\nabla B_0^2}{2} \, - \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, + \, B_0 \cdot \nabla \mathbf{B_1} \,+ \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
+$ = \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(-\frac{\nabla B_0^2}{2} \, - \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, + \, B_0 \cdot \nabla \mathbf{B_1} \,+ \, B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} \, - \, \frac{\nabla\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
 
@@
 
-$\frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, = \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
+$\frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, = \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(B_0 \mathbf{\hat{z}} \cdot \nabla \mathbf{B_1} - \, \frac{\nabla \mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
 
@@
 
-$\frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, = \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(B_0 \partial_z \mathbf{B_1} - \, \frac{\mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
+$\frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, = \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left(B_0 \partial_z \mathbf{B_1} - \, \frac{\nabla \mathbf{B_1}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla \mathbf{B_1} \, \right)$ \\
 
@@
 
-$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, + \, 2 q \Omega_0 x \, - \, 2 u_{1y}\,= \, 2 q x \, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_x P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_x P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1x} \,- \, \frac{B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
+$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, + \, 2 q \Omega_0 x \, - \, 2 u_{1y}\,= \, 2 q x \, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_x P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_x P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1x} \,- \, \frac{\nabla B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
 
 $\partial_t u_{1y} \, - \, q \Omega_0 u_{1x} \, - \, q \Omega_0 x \partial_y u_{1y} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1y} \,+ \, 2u_{1x} \, =  $\\
-$\, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_y P_1}{\rho_1 + \rho_b} \right) \, + \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1y} \,- \, \frac{B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
+$\, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_y P_1}{\rho_1 + \rho_b} \right) \, + \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1y} \,- \, \frac{\nabla B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
 
 $\partial_t u_{1z} \, - \, q \Omega_0 x \partial_y u_{1z} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1z} \, = $\\
-$\, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_z P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_z P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1z} \,- \, \frac{B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
+$\, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_z P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_z P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1z} \,- \, \frac{\nabla B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
 
@@
 
-$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, - \, 2 u_{1y}\,= \, - \, \frac{1}{\rho_b}\partial_x p  \, + \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1x} \,- \, \frac{B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
+$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, - \, 2 u_{1y}\,= \, - \, \frac{1}{\rho_b}\partial_x p  \, + \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1x} \,- \, \frac{\nabla B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
 
 $\partial_t u_{1y}\, - \, q \Omega_0 u_{1x}  \, - \, q \Omega_0 x \partial_y u_{1y} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1y} \,+ \, 2u_{1x} \, =  $\\
-$\, - \, \frac{1}{\rho_b}\partial_y p \, + \,  \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1y} \,- \, \frac{B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
+$\, - \, \frac{1}{\rho_b}\partial_y p \, + \,  \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1y} \,- \, \frac{\nabla B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
 
-$\partial_t u_{1z} \, - \, q \Omega_0 x \partial_y u_{1z} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1z} \, = \, - \, \frac{1}{\rho_b}\partial_z p  \, + \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1z} \,- \, \frac{B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
+$\partial_t u_{1z} \, - \, q \Omega_0 x \partial_y u_{1z} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1z} \, = \, - \, \frac{1}{\rho_b}\partial_z p  \, + \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1z} \,- \, \frac{\nabla B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
 
@@
 
-$\partial_t\mathbf{B} \, = \, \left(\mathbf{B} \cdot \nabla \right)\mathbf{u} \, -\, \left(\mathbf{u} \cdot \nabla \right)\mathbf{B} \, + \, \frac{1}{R_m}\nabla^2\mathbf{B} $ \\
+$\partial_t\mathbf{B} \, = \, \left(\mathbf{B} \cdot \nabla \right)\mathbf{u} \, -\, \left(\mathbf{u} \cdot \nabla \right)\mathbf{B} \, + \, \frac{1}{\reym}\nabla^2\mathbf{B} $ \\
 
@@
 
-$\partial_t \mathbf{B_1} \, = \, B_0 \partial_z \mathbf{u_1} \, + \, \mathbf{B_1} \cdot \nabla \mathbf{u_1} \, + \, q \Omega_0 x \partial_y \mathbf{B_1} \, - \, \mathbf{u_1} \cdot \nabla \mathbf{B_1} \, + \, \frac{1}{R_m} \nabla^2 \mathbf{B_1} $ \\
+$\partial_t \mathbf{B_1} \, = \, B_0 \partial_z \mathbf{u_1} \, + \, \mathbf{B_1} \cdot \nabla \mathbf{u_1} \, + \, q \Omega_0 x \partial_y \mathbf{B_1} \, - \, \mathbf{u_1} \cdot \nabla \mathbf{B_1} \, + \, \frac{1}{\reym} \nabla^2 \mathbf{B_1} $ \\
 

### 2014-02-17 notes/mri_notes.tex
beginning of streamfunctions

@@
 
+\section*{Streamfunctions}
+
+Because we are working with a flow that is axisymmetric and incompressible, we can reparameterize the velocity in terms of its streamfunction. \\
+
+Expanding the incompressibility requirement, we have: \\
+
+$\nabla \cdot \mathbf{u_1} \, = \, \partial_x u_{1x} \, \, + \partial_z u_{1z} \, = \, 0$ \\
+
+Because the $\partial_y$ term is 0 even if $u_{1y}$ itself is nonzero. So we see that we can enforce this requirement if we rewrite the velocity in terms of a function $\Psi$ such that:\\
+
+$u_{x1} \, = \, \frac{\partial \Psi}{\partial z}$ \, \, and \, \, $u_{z1} \, = \, - \frac{\partial \Psi}{\partial x}$ \\
+
+Because, plugging into the incompressibility condition: \\
+
+\[\frac{\partial \left(\frac{\partial \Psi}{\partial z}\right)}{\partial x} \, + \, \frac{\partial \left(- \frac{\partial \Psi}{\partial x}\right)}{\partial z} \, = \, \frac{\partial^2 \Psi}{\partial x \partial z} \, - \, \frac{\partial^2 \Psi}{\partial x \partial z} \, = \, 0\] \\
+
+This is clearly zero always. Thus, we have the stream function $\Psi$, defined such that $\left(u_{x1}, u_{z1}\right) \, = \, \left(\partial_z \Psi, -\partial_x \Psi\right)$.
+
 \bibliography{mri.bib}

### 2014-02-24 notes/mri_notes.tex
more stream functions

@@
 
-Expanding the incompressibility requirement, we have: \\
+We express the velocity perturbation in terms of the curl of some vector field $\Psi$: \\
+
+ $\mathbf{u_1} \, = - \, \nabla \times \Psi\left(x, z\right) \mathbf{\hat{y}}$ \\
+
+ This expands as follows: \\
+
+ $- \nabla \times \Psi\left(x, z\right) \mathbf{\hat{y}} \, = \, - \Psi\nabla \times \mathbf{\hat{y}} \, + \, \mathbf{\hat{y}} \times \nabla \Psi \, = \, \mathbf{\hat{y}} \times \nabla \Psi \,$ \\
+
+ $= \, \mathbf{\hat{y}} \times \partial_x \Psi \mathbf{\hat{x}} \, + \, \mathbf{\hat{y}} \times \partial_z \Psi \mathbf{\hat{z}} \, = \, -\partial_x \Psi \mathbf{\hat{z}} \, + \, \partial_z \Psi \mathbf{\hat{x}}$ \\
+
+ And thus we have: \\
+
+$u_{x1} = \partial_z \Psi$ \, and \, $u_{z1} = -\partial_x \Psi$ \\
+
+We can check that this works in the following way. Expanding the incompressibility requirement, we have: \\
 
@@
 
-This is clearly zero always. Thus, we have the stream function $\Psi$, defined such that $\left(u_{x1}, u_{z1}\right) \, = \, \left(\partial_z \Psi, -\partial_x \Psi\right)$.
+This is clearly zero always.
+Thus, we have the stream function $\Psi$, defined such that $\left(u_{x1}, u_{z1}\right) \, = \, \left(\partial_z \Psi, -\partial_x \Psi\right)$.
+Analogously, we can use the solenoidal magnetic field constraint $\nabla \cdot \mathbf{B_1} \, = \, 0$ to define a flux function based on the vector potential A: \\
+
+$B_{1x} = \partial_z A$ \, and \, $B_{1z} = - \partial_x A$ \\
+
+Essentially, we plug in the following to our equations: \\
+
+$\mathbf{B_1} \, = \, \left(\begin{matrix}
+
+\partial_zA \\
+B_{1y} \\
+-\partial_xA \\
+
+\end{matrix}\right)$ \, and \, $\mathbf{u_1} \, = \, \left(\begin{matrix}
+
+\partial_z\Psi \\
+u_{1y} \\
+-\partial_x\Psi \\
+\end{matrix}\right)$ \\
 

### 2014-02-25 notes/mri_notes.tex
yet more stream functions
lots more stream functions

@@
 
-If we plug in perturbations and all terms involving $\nabla\left(-q\Omega_0 x \mathbf{\hat{y}}\right)$ or $\nabla B_0 \mathbf{\hat{z}}$ go to zero, we arrive at: \\
+%If we plug in perturbations and all terms involving $\nabla\left(-q\Omega_0 x \mathbf{\hat{y}}\right)$ or $\nabla B_0 \mathbf{\hat{z}}$ go to zero, we arrive at: \\
 
-$\partial_t \mathbf{B_1} \, = \, B_0 \partial_z \mathbf{u_1} \, + \, \mathbf{B_1} \cdot \nabla \mathbf{u_1} \, + \, q \Omega_0 x \partial_y \mathbf{B_1} \, - \, \mathbf{u_1} \cdot \nabla \mathbf{B_1} \, + \, \frac{1}{\reym} \nabla^2 \mathbf{B_1} $ \\
+For the first two terms on the righthand side, plugging in perturbations yields: \\
+
+$\left(\mathbf{B} \cdot \nabla \right)\mathbf{u} \, -\, \left(\mathbf{u} \cdot \nabla \right)\mathbf{B} \, = \, \left(B_0 \mathbf{\hat{z}} + \mathbf{B_1}\right)\cdot\nabla \left(-q\Omega_0 x \mathbf{\hat{y}} + \mathbf{u_1}\right) \, - \, \left(-q\Omega_0 x \mathbf{\hat{y}} + \mathbf{u_1}\right) \cdot \nabla \left(B_0 \mathbf{\hat{z}} + \mathbf{B_1}\right) $ \\
+
+$ = B_0 \partial_z \mathbf{u_1} \, + \, \mathbf{B_1}\cdot\nabla\mathbf{u_1} \, + \, q \Omega_0 x \partial_y \mathbf{B_1} \, - \, \mathbf{u_1}\cdot\nabla\mathbf{B_1} \, + \, \mathbf{B_1}\cdot \nabla\left(-q\Omega_0x \mathbf{\hat{y}}\right) \, + \, B_0\mathbf{\hat{z}} \cdot \nabla\left(-q\Omega_0x \mathbf{\hat{y}}\right) $ \\
+
+Because all terms with $\nabla B_0 \mathbf{\hat{z}}$ go to zero.
+That last term goes to zero, and the term $\mathbf{B_1}\cdot \nabla\left(-q\Omega_0x \mathbf{\hat{y}}\right) \, = \, B_{1x} \partial_x\left(-q \Omega_0 x \right) \mathbf{\hat{y}}$ \\
+
+So the induction equation becomes: \\
+
+$\partial_t \mathbf{B_1} \, = \, B_0 \partial_z \mathbf{u_1} \, + \, \mathbf{B_1} \cdot \nabla \mathbf{u_1} \, + \, q \Omega_0 x \partial_y \mathbf{B_1} \, - \, \mathbf{u_1} \cdot \nabla \mathbf{B_1} \, - \, q\Omega_0 B_{1x} \mathbf{\hat{y}} \,+ \, \frac{1}{\reym} \nabla^2 \mathbf{B_1} $ \\
 
@@
 
-$\partial_t \mathbf{B_1} \, = \, B_0 \partial_z \mathbf{u_1} \, + \, \mathbf{B_1} \cdot \nabla \mathbf{u_1} \, + \, q \Omega_0 x \partial_y \mathbf{B_1} \, - \, \mathbf{u_1} \cdot \nabla \mathbf{B_1} \, + \, \frac{1}{\reym} \nabla^2 \mathbf{B_1} $ \\
+$\partial_t \mathbf{B_1} \, = \, B_0 \partial_z \mathbf{u_1} \, + \, \mathbf{B_1} \cdot \nabla \mathbf{u_1} \, + \, q \Omega_0 x \partial_y \mathbf{B_1} \, - \, \mathbf{u_1} \cdot \nabla \mathbf{B_1} \, - \, q\Omega_0 B_{1x} \mathbf{\hat{y}} \,\, + \, \frac{1}{\reym} \nabla^2 \mathbf{B_1} $ \\
 
@@
 
+For the x component of the momentum equation, we have: \\
+
+$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \, + \, \left(\mathbf{u_1} \cdot \nabla\right) u_{1x} \, - \, 2 u_{1y} \, = \, \frac{1}{4\pi} \partial_x \Pi \, + \, \frac{1}{4\pi} B_0 \partial_z B_{1x} \, + \, \frac{1}{4\pi} \mathbf{B_1} \cdot \nabla B_{1x} \, + \, \frac{1}{\reye}\nabla^2 u_{1x} $ \\
+
+The $\partial_y$ terms go to zero, and this expands to become: \\
+
+$\partial_t\partial_z\Psi \, + \, \partial_z\Psi \partial_x \partial_z \Psi \, - \, \partial_x \Psi \partial_z^2\Psi \, - \, 2 u_{1y} \, = \, \frac{1}{4\pi}\partial_x \Pi \, + \, \frac{1}{4\pi} B_0 \partial_z^2 A \, + \, \frac{1}{4\pi} \partial_z A \partial_x \partial_z A \, - \, \frac{1}{4\pi} \partial_x A \partial_z^2 A \, + \, \frac{1}{\reye} \nabla^2 \partial_z \Psi$ \\
+
+Multiplying through by $\partial_z$ and expanding terms yields: \\
+
+$\partial_t \partial_z^2 \Psi \, + \, \partial_z^2\Psi\partial_x\partial_z\Psi \, + \, \partial_z\Psi \partial_x\partial_z^2 \Psi \, - \, \partial_z\partial_x\Psi\partial_z^2\Psi \, - \, \partial_x \Psi \partial_z^3 \Psi \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_z\partial_x\Pi \, + \, \frac{1}{4\pi} B_0\partial_z^3A \, + \, \frac{1}{4\pi}\partial_z^2 A \partial_x \partial_z A \, + \, \frac{1}{4\pi}\partial_zA\partial_x\partial_z^2 A \, - \, \partial{1}{4\pi}\partial_z\partial_x A \partial_z^2 A \, - \, \frac{1}{4\pi}\partial_xA\partial_z^3A \, + \, \frac{1}{\reye} \nabla^2 \partial_z^2 \Psi $ \\
+
+Several terms cancel, and several can be grouped into Jacobians: \\
+
+$J\left(\Psi, \partial_z^2\Psi\right) \, = \, \partial_z \Psi \partial_x\partial_z^2\Psi \, - \, \partial_x\Psi\partial_z^3\Psi $ \\
+
+$J\left(A, \partial_z^2 A\right) \, = \, \partial_z A \partial_x\partial_z^2A \, - \, \partial_xA\partial_z^3A $
+
+Where $J\left(f, g\right) \equiv \partial_z f \partial_x g - \partial_xf \partial_z g$ \\
+
+The x-component of the momentum equation then becomes: \\
+
+$\partial_t\partial_z^2\Psi \, + \, J\left(\Psi, \partial_z^2\Psi\right) \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_z\partial_z\Pi \, + \, \frac{1}{4\pi}B_0 \partial_z^3A \, + \, J\left(A, \partial_z^2A\right) \, + \, \frac{1}{\reye}\nabla^2 \partial_z^2 \Psi$ \\
+
+The z-component of the momentum equation is: \\
+
+$\partial_t u_{1z} \, - \, q\Omega_0x\partial_y u_{1z} \, + \, \left(\mathbf{u_1} \cdot \nabla \right) u_{1z} \, = \, \frac{1}{4\pi}\partial_z \Pi \, + \, \frac{1}{4\pi} B_0 \partial_z\partial_zB_{1z} \, + \, \frac{1}{4\pi}\mathbf{B_1}\cdot\nabla B_{1z} \, + \, \frac{1}{\reye}\nabla^2 u_{1z} $ \\
+
+Which, after following the steps outlined above in the expansion of the x-component, becomes: \\
+
+$\partial_t \partial_x^2 \Psi \, + \, J\left(\Psi, \partial_x^2 \Psi\right) \, = \, -\frac{1}{4\pi}\partial_x\partial_z\Pi \, + \, \frac{1}{4\pi}B_0\partial_z\partial_x^2 A \, + \, J\left(A, \partial_x^2 A \right) \, + \, \frac{1}{\reye} \nabla^2 \partial_x^2 \Psi$ \\
+
+We can combine the two equations (for the x- and z-components), noting that: \\
+
+$\nabla^2 \, = \, \partial_x^2 \mathbf{\hat{x}} \, \partial_z^2 \mathbf{\hat{z}}$ \\
+
+$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \partial_x^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
+
+Where the term $\frac{1}{4\pi}\partial_x\partial_z\Pi \left(\mathbf{\hat{x}} \, - \, \mathbf{\hat{z}}\right)$ goes to zero. \\
+
+For the induction equation, we begin with the y component: \\
+
+$\partial_t B_{1y} \, = \, B_0 \partial_z u_{1y} \, + \, \mathbf{B_1} \cdot \nabla u_{1y} \, + \, q \Omega_0x\partial_y B_{1y} \, - \, \mathbf{u_1} \cdot \nabla B_{1y} \, + \, \frac{1}{\reym} \nabla^2 B_{1y} \, - \, q\Omega_0 B_{1x} \,$ \\
+
+Where: \\
+
+$\nabla u_{1y} \, = \, \partial_x u_{1y} \mathbf{\hat{x}} \, + \, \partial_z u_{1y}\mathbf{\hat{z}}$ \\
+
+$\nabla B_{1y} \, = \, \partial_x B_{1y}\mathbf{\hat{x}} \, + \, \partial_z B_{1y}\mathbf{\hat{z}}$ \\
+
+The equation expands to become: \\
+
+$\partial_t B_{1y} \, = \, B_0 \partial_z u_{1y} \, + \, \partial_z A \mathbf{\hat{x}} \cdot \partial_x u_{1y}\mathbf{\hat{x}} \, + \, \partial_z A \mathbf{\hat{x}} \cdot \partial_z u_{1y} \mathbf{\hat{z}} \, - \, \partial_x A \mathbf{\hat{z}} \cdot \partial_x u_{1y} \mathbf{\hat{x}} \, - \, \partial_x A \mathbf{\hat{z}} \cdot \partial_z u_{1y} \mathbf{\hat{z}} \, + \, B_{1y}\mathbf{\hat{y}} \cdot \partial_x u_{1y} \mathbf{\hat{x}} \, + \, B_{1y}\mathbf{\hat{y}} \cdot \partial_z u_{1y} \mathbf{\hat{z}} \, + \, q\Omega_0 x \partial_y B_{1y} \mathbf{\hat{y}} \, - \, \partial_z\Psi\mathbf{\hat{x}}\cdot \partial_x B_{1y} \mathbf{\hat{x}} \, - \, \partial_z\Psi\mathbf{\hat{x}}\cdot \partial_z B_{1y} \mathbf{\hat{z}} \, + \, \partial_x\Psi\mathbf{\hat{z}}\cdot \partial_x B_{1y} \mathbf{\hat{x}} \, + \, \partial_x\Psi\mathbf{\hat{z}}\cdot \partial_z B_{1y} \mathbf{\hat{z}} \, - \, u_{1y} \mathbf{\hat{y}} \partial_x B_{1y} \mathbf{\hat{x}} \, - \, u_{1y}\mathbf{\hat{y}} \partial_z B_{1y} \mathbf{\hat{z}} \, + \, \frac{1}{\reym} \nabla^2 B_{1y} \, - \, q\Omega_0 \left(\partial_z A \right)$ \\
+
+Clearly most of these terms cancel in the dot product, and we are left with: \\
+
+$\partial_t B_{1y} \, = \, B_0 \partial_z u_{1y} \, + \, \partial_z A \partial_x u_{1y} \, - \, \partial_x A \partial_z u_{1y} \, + \, q\Omega_0 x \partial_y B_{1y} \, - \, \partial_z\Psi \partial_x B_{1y} \, + \, \partial_x\Psi \partial_z B_{1y} \, + \, \frac{1}{\reym} \nabla^2 B_{1y} \, - \, q \Omega_0 \partial_z A$ \\
+
+The nonlinear terms are Jacobians, and $\partial_y B_{1y} = 0$: \\
+
+$\partial_t B_{1y} \, = \, B_0 \partial_z u_{1y} \, + \, J\left(A, u_{1y}\right) \, - \, J\left(\Psi, B_{1y}\right) \, + \, \frac{1}{\reym} \nabla^2 B_{1y}  \, - \, q \Omega_0 \partial_z A$ \\
+
+The x component of the induction equation: \\
+
+$\partial_t B_{1x} \, = \, B_0 \partial_z u_{1x} \, + \, \mathbf{B_1} \cdot \nabla u_{1x} \, + \, q \Omega_0 x \partial_y B_{1x} \, - \, \mathbf{u_1} \cdot \nabla B_{1x} \, + \, \frac{1}{\reym} \nabla^2 B_{1x}$ \\
+
+Where: \\
+
+$\nabla u_{1x}  \, = \, \partial_x \partial_z \Psi \mathbf{\hat{x}} \, + \, \partial_y \partial_z \Psi \mathbf{\hat{y}} \, + \, \partial_z \partial_z \Psi \mathbf{\hat{z}}$ \\
+
+$\nabla B_{1x} \, = \, \partial_x \partial_z A \mathbf{\hat{x}} \, + \, \partial_y\partial_z A \mathbf{\hat{y}} \, + \, \partial_z \partial_z A \mathbf{\hat{z}}$ \\
+
+Becomes: \\
+
+$\partial_t \partial_z A \, = \, B_0 \partial_z^2 \Psi \, + \, \partial_z A \partial_x \partial_z \Psi \, - \, \partial_x A \partial_z^2 \Psi \, - \, \partial_z \Psi \partial_x \partial_z A \, + \, \partial_x \Psi \partial_z^2 A \, + \, \frac{1}{\reym} \nabla^2 \partial_z A$ \\
+
+After setting all of the $\partial_y$ terms to 0. Using the fact that: \\
+
+$\partial_z\left(\partial_z A \partial_x \Psi\right) \, = \, \partial_z A \partial_z \partial_x \Psi \, + \, \partial_x \Psi \partial_z^2 A$ \\
+
+$\partial_z\left(\partial_x A \partial_z \Psi\right) \, = \, \partial_x A \partial_z^2 \Psi \, + \, \partial_z \Psi \partial_z \partial_x A$ \\
+
+This can be rewritten: \\
+
+$\partial_z \left(\partial_t A \right) \, = \, \partial_z \left(B_0 \partial_z \Psi \right) \, + \partial_z \left(\partial_z A \partial_x \Psi\right) \, - \, \partial_z\left(\partial_x A \partial_z \Psi \right) \, + \, \partial_z \left(\frac{1}{Rm} \nabla^2 A\right)$ \\
+
+And we can divide out a $\partial_z$ term to arrive at: \\
+
+$\partial_t A \, = \, B_0 \partial_z \Psi \, + \partial_z A \partial_x \Psi \, - \, \partial_x A \partial_z \Psi \, + \, \frac{1}{Rm} \nabla^2 A$ \\
+
+Or: \\
+
+$\partial_t A \, = \, B_0 \partial_z \Psi \, + \, J\left(A, \Psi\right) \, + \, \frac{1}{Rm} \nabla^2 A$ \\
+
+Clearly, this is the equation for both the x and z components of the induction equation.
+
 \bibliography{mri.bib}

### 2014-02-25 notes/mri_notes.tex
final equation set

@@
 
-Where $J\left(f, g\right) \equiv \partial_z f \partial_x g - \partial_xf \partial_z g$ \\
+Where $J\left(f, g\right) \equiv \partial_z f \partial_x g - \partial_x f \partial_z g$ \\
 
@@
 Where the term $\frac{1}{4\pi}\partial_x\partial_z\Pi \left(\mathbf{\hat{x}} \, - \, \mathbf{\hat{z}}\right)$ goes to zero. \\
+
+The y-component of the momentum equation is even simpler: \\
+
+$\partial_t u_{1y} \, - q\Omega_0 x \partial_y u_{1y} \, + \, \left(\mathbf{u_1} \cdot \nabla \right)u_{1y} \, + \, \left(2 - \Omega_0 q\right) u_{1x} \, = \, \frac{1}{4\pi} \partial_y \Pi \, + \, \frac{1}{4\pi} B_0 \partial_z B_{1y} \, + \, \frac{1}{4\pi}\mathbf{B_1} \cdot \nabla B_{1y} \, + \, \frac{1}{\reye}\nabla^2 u_{1y}$ \\
+
+Which becomes: \\
+
+$\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, + \, \left(2 - \Omega_0 q\right) \partial_z \Psi \, = \, \frac{1}{4\pi}B_0\partial_z B_{1y} \, + \, \frac{1}{4\pi} J\left(A, B_{1y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{1y}$ \\
 
@@
 
-Clearly, this is the equation for both the x and z components of the induction equation.
+Clearly, this is the equation for both the x and z components of the induction equation. \\
+
+Our final equation set is: \\
+
+$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \partial_x^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
+
+$\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, + \, \left(2 - \Omega_0 q\right) \partial_z \Psi \, = \, \frac{1}{4\pi}B_0\partial_z B_{1y} \, + \, \frac{1}{4\pi} J\left(A, B_{1y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{1y}$ \\
+
+$\partial_t A \, = \, B_0 \partial_z \Psi \, + \, J\left(A, \Psi\right) \, + \, \frac{1}{Rm} \nabla^2 A$ \\
+
+$\partial_t B_{1y} \, = \, B_0 \partial_z u_{1y} \, + \, J\left(A, u_{1y}\right) \, - \, J\left(\Psi, B_{1y}\right) \, + \, \frac{1}{\reym} \nabla^2 B_{1y}  \, - \, q \Omega_0 \partial_z A$ \\
+
+Note that in all of these derivations, the shear terms went to zero because they contained $\partial_y$ terms.
+We could instead have explicitly transformed into a shearing frame moving with the steady-state azimuthal flow, but it does not end up mattering because our perturbations are axisymmetric.
+\\
 

### 2014-03-03 notes/mri_notes.tex
beginning of multiscale analysis

@@
 
-$\partial_t\partial_z^2\Psi \, + \, J\left(\Psi, \partial_z^2\Psi\right) \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_z\partial_z\Pi \, + \, \frac{1}{4\pi}B_0 \partial_z^3A \, + \, J\left(A, \partial_z^2A\right) \, + \, \frac{1}{\reye}\nabla^2 \partial_z^2 \Psi$ \\
+$\partial_t\partial_z^2\Psi \, + \, J\left(\Psi, \partial_z^2\Psi\right) \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_x\partial_z\Pi \, + \, \frac{1}{4\pi}B_0 \partial_z^3A \, + \, J\left(A, \partial_z^2A\right) \, + \, \frac{1}{\reye}\nabla^2 \partial_z^2 \Psi$ \\
 
@@
 
+\section*{Multiscale Analysis}
+
+We now wish to characterize our perturbations in terms of amplitude functions with fast and slow variables.
+To choose our scalings, we first linearize our equations and find the linear dispersion relation.
+We then pick spatial and temporal scalings such that the spatial and temporal eigenvalues all appear in the dispersion relation at the same lowest order.
+\\
+
+Linearizing our four equations, we find a system described by the matrix: \\
+
+%$\left[\begin{smallmatrix}{}s \left(- k_{x}^{2} - k_{z}^{2}\right) - \frac{1}{Re} \left(k_{x}^{4} + 2 k_{x}^{2} k_{z}^{2} + k_{z}^{4}\right) & 2 i k_{z} & \frac{i B_{0} k_{z}}{4 \pi} \left(- k_{x}^{2} - k_{z}^{2}\right) & 0\\\i k_{z} \\left(- \\Omega_{0} q + 2\\right) & s - \\frac{1}{Re} \\left(- k_{x}^{2} - k_{z}^{2}\\right) & 0 & - \\frac{i B_{0} k_{z}}{4 \\pi}\\\\- i B_{0} k_{z} & 0 & s - \\frac{1}{Rm} \\left(- k_{x}^{2} - k_{z}^{2}\\right) & 0\\\\0 & - i B_{0} k_{z} & i \\Omega_{0} k_{z} q & s - \\frac{1}{Rm} \\left(- k_{x}^{2} - k_{z}^{2}\\right)\end{smallmatrix}\right]$
+
+\begin{equation*}
+\left[\begin{matrix}
+\left(- k_{x}^{2} - k_{z}^{2}\right) - \frac{1}{Re} \left(k_{x}^{4} + 2 k_{x}^{2} k_{z}^{2} + k_{z}^{4}\right) & 2 i k_{z} & \frac{i B_{0} k_{z}}{4 \pi} \left(- k_{x}^{2} - k_{z}^{2}\right) & 0\\
+
+i k_{z} \left(- \Omega_{0} q + 2\right) & s - \frac{1}{Re} \left(- k_{x}^{2} - k_{z}^{2}\right) & 0 & - \frac{i B_{0} k_{z}}{4 \pi}\\
+
+-i B_{0} k_{z} & 0 & s - \frac{1}{Rm} \left(- k_{x}^{2} - k_{z}^{2}\right) & 0\\
+
+0 & - i B_{0} k_{z} & i \Omega_{0} k_{z} q & s - \frac{1}{Rm} \left(- k_{x}^{2} - k_{z}^{2}\right)
+
+\end{matrix}\right] \cdot \left[\begin{matrix}
+
+\Psi \\
+u_{1y}\\
+A\\
+B_{1y}
+
+\end{matrix}\right]
+\end{equation*}
+
 \bibliography{mri.bib}

### 2014-03-04 notes/mri_notes.tex
fast and slow vars

@@
 Linearizing our four equations, we find a system described by the matrix: \\
-
-%$\left[\begin{smallmatrix}{}s \left(- k_{x}^{2} - k_{z}^{2}\right) - \frac{1}{Re} \left(k_{x}^{4} + 2 k_{x}^{2} k_{z}^{2} + k_{z}^{4}\right) & 2 i k_{z} & \frac{i B_{0} k_{z}}{4 \pi} \left(- k_{x}^{2} - k_{z}^{2}\right) & 0\\\i k_{z} \\left(- \\Omega_{0} q + 2\\right) & s - \\frac{1}{Re} \\left(- k_{x}^{2} - k_{z}^{2}\\right) & 0 & - \\frac{i B_{0} k_{z}}{4 \\pi}\\\\- i B_{0} k_{z} & 0 & s - \\frac{1}{Rm} \\left(- k_{x}^{2} - k_{z}^{2}\\right) & 0\\\\0 & - i B_{0} k_{z} & i \\Omega_{0} k_{z} q & s - \\frac{1}{Rm} \\left(- k_{x}^{2} - k_{z}^{2}\\right)\end{smallmatrix}\right]$
-
 \begin{equation*}
-\left[\begin{matrix}
+M = \left[\begin{matrix}
 \left(- k_{x}^{2} - k_{z}^{2}\right) - \frac{1}{Re} \left(k_{x}^{4} + 2 k_{x}^{2} k_{z}^{2} + k_{z}^{4}\right) & 2 i k_{z} & \frac{i B_{0} k_{z}}{4 \pi} \left(- k_{x}^{2} - k_{z}^{2}\right) & 0\\
@@
 
-\end{matrix}\right] \cdot \left[\begin{matrix}
+\end{matrix}\right] 
+\end{equation*}
+
+where: 
+$M \cdot \left[\begin{matrix}
 
@@
 
+\end{matrix}\right] = \left[\begin{matrix}
+0 \\
+0\\
+0\\
+0\\
 \end{matrix}\right]
-\end{equation*}
+$ \\
+
+The determinant of M gives the dispersion relation: \\
+
+$\frac{B_{0}^{4} k_{x}^{2} k_{z}^{4}}{16 \pi^{2}} + \frac{B_{0}^{4} k_{z}^{6}}{16 \pi^{2}} - \frac{B_{0}^{2} \Omega_{0} k_{z}^{4} q}{2 \pi} - 2 \Omega_{0} k_{z}^{2} q s^{2} - \frac{4 s}{Rm} \Omega_{0} k_{x}^{2} k_{z}^{2} q - \frac{4 s}{Rm} \Omega_{0} k_{z}^{4} q - \frac{2 \Omega_{0}}{Rm^{2}} k_{x}^{4} k_{z}^{2} q - \frac{4 \Omega_{0}}{Rm^{2}} k_{x}^{2} k_{z}^{4} q - \frac{2 \Omega_{0}}{Rm^{2}} k_{z}^{6} q - k_{x}^{2} s^{4} - k_{z}^{2} s^{4} + 4 k_{z}^{2} s^{2} - \frac{2 s^{3}}{Rm} k_{x}^{4} - \frac{4 s^{3}}{Rm} k_{x}^{2} k_{z}^{2} + \frac{8 s}{Rm} k_{x}^{2} k_{z}^{2} - \frac{2 s^{3}}{Rm} k_{z}^{4} + \frac{8 s}{Rm} k_{z}^{4} - \frac{k_{x}^{6} s^{2}}{Rm^{2}} - \frac{3 s^{2}}{Rm^{2}} k_{x}^{4} k_{z}^{2} + \frac{4 k_{x}^{4}}{Rm^{2}} k_{z}^{2} - \frac{3 s^{2}}{Rm^{2}} k_{x}^{2} k_{z}^{4} + \frac{8 k_{x}^{2}}{Rm^{2}} k_{z}^{4} - \frac{k_{z}^{6} s^{2}}{Rm^{2}} + \frac{4 k_{z}^{6}}{Rm^{2}} - \frac{2 s^{3}}{Re} k_{x}^{4} - \frac{4 s^{3}}{Re} k_{x}^{2} k_{z}^{2} - \frac{2 s^{3}}{Re} k_{z}^{4} - \frac{4 k_{x}^{6} s^{2}}{Re Rm} - \frac{12 k_{x}^{4} k_{z}^{2} s^{2}}{Re Rm} - \frac{12 k_{x}^{2} k_{z}^{4} s^{2}}{Re Rm} - \frac{4 k_{z}^{6} s^{2}}{Re Rm} - \frac{2 k_{x}^{8} s}{Re Rm^{2}} - \frac{8 k_{x}^{6} k_{z}^{2} s}{Re Rm^{2}} - \frac{12 k_{x}^{4} k_{z}^{4} s}{Re Rm^{2}} - \frac{8 k_{x}^{2} k_{z}^{6} s}{Re Rm^{2}} - \frac{2 k_{z}^{8} s}{Re Rm^{2}} - \frac{k_{x}^{6} s^{2}}{Re^{2}} - \frac{3 s^{2}}{Re^{2}} k_{x}^{4} k_{z}^{2} - \frac{3 s^{2}}{Re^{2}} k_{x}^{2} k_{z}^{4} - \frac{k_{z}^{6} s^{2}}{Re^{2}} - \frac{2 k_{x}^{8} s}{Re^{2} Rm} - \frac{8 k_{x}^{6} k_{z}^{2} s}{Re^{2} Rm} - \frac{12 k_{x}^{4} k_{z}^{4} s}{Re^{2} Rm} - \frac{8 k_{x}^{2} k_{z}^{6} s}{Re^{2} Rm} - \frac{2 k_{z}^{8} s}{Re^{2} Rm} - \frac{k_{x}^{10}}{Re^{2} Rm^{2}} - \frac{5 k_{x}^{8} k_{z}^{2}}{Re^{2} Rm^{2}} - \frac{10 k_{x}^{6} k_{z}^{4}}{Re^{2} Rm^{2}} - \frac{10 k_{x}^{4} k_{z}^{6}}{Re^{2} Rm^{2}} - \frac{5 k_{x}^{2} k_{z}^{8}}{Re^{2} Rm^{2}} - \frac{k_{z}^{10}}{Re^{2} Rm^{2}}$ \\
+
+Which, as expected, is fourth order in s.
+We see that the spatial eigenvalues appear as $k_x^2$ and $k_z^2$ at lowest order.
+The temporal eigenvalue appears as $s$ at lowest order.
+We choose the scalings between fast variables $(x, y, z, t)$ and slow variables $(X, Y, Z, T)$ such that all eigenvalues come in at $\mathcal{O}(\epsilon^2)$.
+Hence we define: \\
+
+\noindent $X = \epsilon x$, \, \, $Y = \epsilon y$, \, \, $Z = \epsilon z$, \, \, $T = \epsilon^2 t$ \\
 

### 2014-03-10 notes/mri_notes.tex
corrected mistake
expanded operators

@@
 
-$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \partial_x^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
+$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
 
@@
 
-\noindent $X = \epsilon x$, \, \, $Y = \epsilon y$, \, \, $Z = \epsilon z$, \, \, $T = \epsilon^2 t$ \\
+\noindent $X \equiv \epsilon x$, \, \, $Y \equiv \epsilon y$, \, \, $Z \equiv \epsilon z$, \, \, $T \equiv \epsilon^2 t$ \\
+
+Let $\mathbf{V}$ be the state vector containing all perturbations: $\mathbf{V} \equiv \left[ \Psi, u_{1y}, A, B_{1y}\right]^T$. \\
+
+Then $\mathbf{V}\left(x, z, t; X, Z, T\right) = \alpha \left(X, Z, T\right)e^{st + ik_x x + ik_z z} + c.c.$ Where $\alpha$ is an amplitude function.
+I believe that at marginality to the MRI we will assume the amplitudes are dependent only on Z and T, but I want to check this -- in the mean time I'll retain all possible terms.\\
+
+The operators are expanded as follows: \\
+
+$\partial_t \to \partial_t + \epsilon^2 \partial_T$ \\
+
+$\partial_z \to \partial_z + \epsilon \partial_Z$ \\
+
+$\partial_x \to \partial_x + \epsilon \partial_X$ \\
+
+$\nabla^2 = \partial_x^2 + \partial_z^2 \to \partial_x^2 + \epsilon^2 \partial_X^2 + 2\epsilon \partial^2_x \partial^2_X + \partial_z^2 + \epsilon^2 \partial_Z^2 + 2\epsilon\partial_z\partial_Z$ \\
+
+$\nabla^4 \to \partial_x^4 + \epsilon^2 \partial_x^2 \partial_X^2 + 2\epsilon\partial_x^3\partial_X + \partial_x^2\partial_z^2 + \epsilon^2\partial_x^2\partial_Z^2 + 2\epsilon \partial_x^2 \partial_z \partial_Z \\
+ + \epsilon^2\partial_X^2\partial_x^2 + \epsilon^4\partial_X^2 + 2\epsilon^3\partial_X^3 \partial_x + \epsilon^2\partial_X^2\partial_z^2 + \epsilon^4\partial_X^2\partial_Z^2 + 2\epsilon^3\partial_X^2\partial_z\partial_Z \\
+ + 2\epsilon\partial_x^3\partial_X + 2\epsilon^3\partial_x\partial_X^3 + 4\epsilon^2 \partial_x^2\partial_X^2 + 2\epsilon\partial_x\partial_X\partial_z^2 + 2\epsilon^3\partial_x\partial_X\partial_Z^2 + 4\epsilon^2\partial_x\partial_X\partial_z\partial_Z \\
+ + \partial_z^2\partial_x^2 + \epsilon^2\partial_z^2\partial_X^2 + 2\epsilon\partial_z^2\partial_x\partial_X + \partial_z^4 + \epsilon^2\partial_z^2\partial_Z^2 + 2\epsilon\partial_z^3\partial_Z \\
++ \epsilon^2\partial_Z^2\partial_x^2 + \epsilon^4\partial_X^2\partial_Z^2 + 2\epsilon^3\partial_Z^2\partial_x\partial_X + \epsilon^3\partial_Z^2\partial_z^2 + \epsilon^4\partial_Z^4 + 2\epsilon^3\partial_Z^3\partial_z \\
++ 2\epsilon\partial_z\partial_Z\partial_x^2 + 2\epsilon^3\partial_X^2\partial_z\partial_Z + 4\epsilon^2\partial_z\partial_Z\partial_x\partial_X + 2\epsilon\partial_z^3\partial_Z + 2\epsilon^3\partial_Z^3\partial_z + 4\epsilon^2 \partial_z^2 \partial_Z^2$
 

### 2014-03-10 notes/mri_notes.tex
a little more
new operators

@@
 + \epsilon^2\partial_Z^2\partial_x^2 + \epsilon^4\partial_X^2\partial_Z^2 + 2\epsilon^3\partial_Z^2\partial_x\partial_X + \epsilon^3\partial_Z^2\partial_z^2 + \epsilon^4\partial_Z^4 + 2\epsilon^3\partial_Z^3\partial_z \\
-+ 2\epsilon\partial_z\partial_Z\partial_x^2 + 2\epsilon^3\partial_X^2\partial_z\partial_Z + 4\epsilon^2\partial_z\partial_Z\partial_x\partial_X + 2\epsilon\partial_z^3\partial_Z + 2\epsilon^3\partial_Z^3\partial_z + 4\epsilon^2 \partial_z^2 \partial_Z^2$
++ 2\epsilon\partial_z\partial_Z\partial_x^2 + 2\epsilon^3\partial_X^2\partial_z\partial_Z + 4\epsilon^2\partial_z\partial_Z\partial_x\partial_X + 2\epsilon\partial_z^3\partial_Z + 2\epsilon^3\partial_Z^3\partial_z + 4\epsilon^2 \partial_z^2 \partial_Z^2$ \\
+
+We proceed by replacing all operators in our equation set with the above, and expanding $\mathbf{V}$ in a perturbation series in $\epsilon$.
+We then group terms of the same order in $\epsilon$, to make three equations (because we have up to $\mathcal{O}(\epsilon^3)$).
+\\
+
+In analogy to Rayleigh-B\'enard convection and the hydrodynamic Taylor-Couette flow, we assume there is no slow-scale variation in the radial direction, and so the amplitudes are dependent only on Z and T.
+That is: \\
+
+$\mathbf{V} = \mathbf{V}\left(x, z, Z, T)\right)$ where in detail $\mathbf{V} = \left[\Psi(x), u_{1y}(x), A(x), B_{1y}(x)\right]^T \alpha\left(Z, T\right) e^{st + ik_z z}$.
+Periodicity in $z$ allows us to posit the exponential $z$ dependence.
+\\
+
+Our operators therefore become: \\
+
+$\partial_t \to \partial_t + \epsilon^2 \partial_T$ \\
+
+$\partial_z \to \partial_z + \epsilon \partial_Z$ \\
+
+$\nabla^2 = \partial_x^2 + \partial_z^2 \to \partial_x^2  + \partial_z^2 + \epsilon^2 \partial_Z^2 + 2\epsilon\partial_z\partial_Z$ \\
+
+$\nabla^4 \to \partial_x^4 + \partial_x^2\partial_z^2 + \epsilon^2\partial_x^2\partial_Z^2 + 2\epsilon \partial_x^2 \partial_z \partial_Z \\
+ + \partial_z^2\partial_x^2 + \partial_z^4 + \epsilon^2\partial_z^2\partial_Z^2 + 2\epsilon\partial_z^3\partial_Z \\
++ \epsilon^2\partial_Z^2\partial_x^2 + \epsilon^3\partial_Z^2\partial_z^2 + \epsilon^4\partial_Z^4 + 2\epsilon^3\partial_Z^3\partial_z \\
++ 2\epsilon\partial_z\partial_Z\partial_x^2 + 2\epsilon\partial_z^3\partial_Z + 2\epsilon^3\partial_Z^3\partial_z + 4\epsilon^2 \partial_z^2 \partial_Z^2$ \\
 

### 2014-03-11 notes/mri_notes.tex
matrix expansion

@@
 
+We now rewrite our system of equations, grouping similar terms into matrices. The system of equations we are rewriting is: \\
+
+$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, J\left(A, \nabla^2 A \right) \, =  \, 2 \partial_z u_{1y} \,+ \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
+
+$\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, - \, \frac{1}{4\pi} J\left(A, B_{1y}\right) \, + \, \left(2 - \Omega_0 q\right) \partial_z \Psi \, = \, \frac{1}{4\pi}B_0\partial_z B_{1y} \, + \, \frac{1}{\reye} \nabla^2 u_{1y}$ \\
+
+$\partial_t A \, = \, B_0 \partial_z \Psi \, + \, J\left(A, \Psi\right) \, + \, \frac{1}{Rm} \nabla^2 A$ \\
+
+$\partial_t B_{1y} \, = \, B_0 \partial_z u_{1y} \, + \, J\left(A, u_{1y}\right) \, - \, J\left(\Psi, B_{1y}\right) \, + \, \frac{1}{\reym} \nabla^2 B_{1y}  \, - \, q \Omega_0 \partial_z A$ \\
+
+First we have a vector of nonlinear terms:\\
+
+$\mathbf{N} = \left[ \begin{matrix}  J(\Psi, \nabla^2 \Psi) - J(A, \nabla^2 A)\\
+J(\Psi, u_{1y}) - \frac{1}{4\pi} J(A, B_{1y}) \\ 
+-J(A, \Psi) \\
+J(\Psi, B_{1y} - J(A, u_{1y}) \\
+\end{matrix} \right]$ \\
+
+The time-differential term can be described by $\mathcal{D}\partial_t \mathbf{V}$ where: \\
+
+$\mathcal{D} = \left[\begin{matrix}
+\nabla^2 & 0 & 0 & 0 \\
+0 & 1& 0 & 0 \\
+0 & 0 & 1 & 0\\
+0 & 0 & 0 & 1 \\
+\end{matrix}\right]$ \\
+
+The righthand side of the equations can be split into three matrices: one containing the Reynolds numbers, one containing the magnetic field strength, and one containing the shear terms. That is: \\
+
+$RHS = \left[\begin{matrix}
+\frac{1}{\reye} \nabla^4 & 0 & 0 & 0 \\
+0 & \frac{1}{\reye}\nabla^2 & 0 & 0 \\
+0 & 0 & \frac{1}{\reym}\nabla^2 & 0 \\
+0 & 0 & 0 & \frac{1}{\reym}\nabla^2 \\ 
+\end{matrix}\right] + \left[\begin{matrix}
+0 & 0 & \frac{1}{4\pi}B_0 \partial_z \nabla^2 & 0 \\
+0 & 0 & 0 & \frac{1}{4\pi} B_0 \partial_z \\
+B_0 \partial_z & 0 & 0 & 0 \\
+0 & B_0 \partial_z & 0 & 0 \\ \end{matrix}\right]$ \\
+
+$+ \left[\begin{matrix}
+0 & 2\partial_z & 0 & 0 \\
+\left(\Omega_0 - 2\right)q\partial_z & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & -q\Omega_0\partial_z & 0 \\ \end{matrix}\right]$ \\
+
+We will now regroup matrices by $\mathcal{O}(\partial_z)$. The first matrix is rewritten as three matrices, keeping in mind that $\nabla^4 = \partial_x^4 + \partial_z^4 + 2\partial_x\partial_z^2$: \\
+
+$\left[\begin{matrix}
+\frac{1}{\reye} \nabla^4 & 0 & 0 & 0 \\
+0 & \frac{1}{\reye}\nabla^2 & 0 & 0 \\
+0 & 0 & \frac{1}{\reym}\nabla^2 & 0 \\
+0 & 0 & 0 & \frac{1}{\reym}\nabla^2 \\ 
+\end{matrix}\right] = \left[\begin{matrix}
+\frac{1}{\reye}\partial_x^4 & 0 & 0 & 0 \\
+0 & \frac{1}{\reye}\partial_x^2 & 0 &0 \\
+0 & 0 & \frac{1}{\reym}\partial_x^2 & 0 \\
+0 & 0 & 0 & \frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right] $ \\
+
+$+ \left[\begin{matrix}
+\frac{1}{\reye}\partial_z^4 & 0 & 0 & 0 \\
+0 & \frac{1}{\reye}\partial_z^2 & 0 & 0 \\
+0 & 0 & \frac{1}{\reym} \partial_z^2 & 0 \\
+0 & 0 & 0 & \frac{1}{\reym} \partial_z^2 \\ \end{matrix} \right] + \left[ \begin{matrix}
+\frac{1}{\reye} 2\partial_x^2\partial_z^2 & 0 & 0 & 0 \\
+0 & 1 & 0 & 0 \\
+0 & 0 & 1 & 0 \\
+0 & 0 & 0 & 1\\ \end{matrix}\right]$
+
+The second matrix involves $B_0$, which we rewrite in terms of the tuning parameter $\epsilon^2 \equiv 1 - B_0$.
+
 \bibliography{mri.bib}

### 2014-03-11 notes/mri_notes.tex
more matrices

@@
 0 & 0 & 0 & \frac{1}{\reym} \partial_z^2 \\ \end{matrix} \right] + \left[ \begin{matrix}
-\frac{1}{\reye} 2\partial_x^2\partial_z^2 & 0 & 0 & 0 \\
+2 \frac{1}{\reye} \partial_x^2\partial_z^2 & 0 & 0 & 0 \\
 0 & 1 & 0 & 0 \\
 0 & 0 & 1 & 0 \\
-0 & 0 & 0 & 1\\ \end{matrix}\right]$
+0 & 0 & 0 & 1\\ \end{matrix}\right]$ \\
 
-The second matrix involves $B_0$, which we rewrite in terms of the tuning parameter $\epsilon^2 \equiv 1 - B_0$.
+And rearranging once more, we arrive at: \\
+
+$\left[\begin{matrix}
+\frac{1}{\reye} \nabla^4 & 0 & 0 & 0 \\
+0 & \frac{1}{\reye}\nabla^2 & 0 & 0 \\
+0 & 0 & \frac{1}{\reym}\nabla^2 & 0 \\
+0 & 0 & 0 & \frac{1}{\reym}\nabla^2 \\ 
+\end{matrix}\right] = \left[\begin{matrix}
+\frac{1}{\reye}\partial_x^4 & 0 & 0 & 0 \\
+0 & \frac{1}{\reye}\partial_x^2 & 0 &0 \\
+0 & 0 & \frac{1}{\reym}\partial_x^2 & 0 \\
+0 & 0 & 0 & \frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right] $ \\
+
+$+ \partial_z^2 \left[\begin{matrix}
+2\frac{1}{\reye} \partial_x^2 & 0 & 0 & 0 \\
+0 & \frac{1}{\reye} & 0 & 0 \\
+0 & 0 & \frac{1}{\reym} & 0 \\
+0 & 0 & 0 & \frac{1}{\reym} \\ \end{matrix}\right] + \partial_z^4 \left[\begin{matrix}
+\frac{1}{\reye} & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\ \end{matrix}\right] $ \\
+
+Clearly these three matrices are 0th, 2nd, and 4th order in $\partial_z$, respectively. \\
+
+The second matrix in $RHS$ involves $B_0$, which we rewrite in terms of the tuning parameter $\epsilon^2 \equiv 1 - B_0$. \\
+
+$\left[\begin{matrix}
+0 & 0 & \frac{1}{4\pi}B_0 \partial_z \nabla^2 & 0 \\
+0 & 0 & 0 & \frac{1}{4\pi} B_0 \partial_z \\
+B_0 \partial_z & 0 & 0 & 0 \\
+0 & B_0 \partial_z & 0 & 0 \\ \end{matrix}\right] = (1 - \epsilon^2) \left[\begin{matrix}
+0 & 0 & \frac{1}{4\pi}\partial_z(\partial_x^2 + \partial_z^2) & 0 \\
+0 & 0 & 0 & \frac{1}{4\pi}\partial_z \\
+\partial_z & 0 & 0 & 0 \\
+0 & \partial_z & 0 & 0 \\ \end{matrix}\right] $ \\
+
+$ = (1 - \epsilon^2)\partial_z\left[\begin{matrix}
+0 & 0 & \frac{1}{4\pi}\partial_x^2 & 0 \\
+0 & 0 & 0 & \frac{1}{4\pi} \\
+1 & 0 & 0 & 0 \\
+0 & 1 & 0 & 0 \\ \end{matrix} \right] + (1 - \epsilon^2)\partial_z^3 \left[\begin{matrix}
+0 & 0 & \frac{1}{4\pi} & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\ \end{matrix} \right] $ \\
+
+The $\mathcal{O}(\partial_z)$ term is combined with the 3rd matrix in RHS, and finally we arrive at our system of equations: \\
 

### 2014-03-13 notes/mri_notes.tex
finished matrix expansion
more expanding

@@
 
+$\mathcal{D}\partial_t \mathbf{V} + \mathbf{N} = \mathcal{L} \mathbf{V} - \epsilon^2\partial_z\mathcal{X} \mathbf{V} - \epsilon^2 \partial_z^3 \mathcal{L}_3\mathbf{V}$ \\
+
+Where $\mathcal{L} \equiv \mathcal{L}_0 + \mathcal{L}_1\partial_z + \mathcal{L}_2\partial_z^2 + \mathcal{L}_3\partial_z^3 + \mathcal{L}_4\partial_z^4$ \\
+
+And: \\
+
+$\mathcal{L}_0 = \left[\begin{matrix}
+\frac{1}{\reye}\partial_x^4 & 0 & 0 & 0 \\
+0 & \frac{1}{\reye}\partial_x^2 & 0 &0 \\
+0 & 0 & \frac{1}{\reym}\partial_x^2 & 0 \\
+0 & 0 & 0 & \frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right] $ \\
+
+\vspace{2mm}
+
+$\mathcal{L}_1 = \left[\begin{matrix}
+0 & 2 & \frac{1}{4\pi}\partial_x^2 & 0 \\
+(\Omega_0 - 2)q & 0 & 0 & \frac{1}{4\pi} \\
+1 & 0 & 0 & 0 \\
+0 & 1 & -q\Omega_0 & 0 \\ \end{matrix}\right] \\$
+
+\vspace{2mm}
+
+$\mathcal{L}_2 = \left[\begin{matrix}
+2\frac{1}{\reye} \partial_x^2 & 0 & 0 & 0 \\
+0 & \frac{1}{\reye} & 0 & 0 \\
+0 & 0 & \frac{1}{\reym} & 0 \\
+0 & 0 & 0 & \frac{1}{\reym} \\ \end{matrix}\right]$
+
+\vspace{2mm}
+
+$\mathcal{L}_3 = \left[\begin{matrix}
+0 & 0 & \frac{1}{4\pi} & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\ \end{matrix} \right] $
+
+\vspace{2mm}
+
+$\mathcal{L}_4 = \left[\begin{matrix}
+\frac{1}{\reye} & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\ \end{matrix}\right] $ \\
+
+\vspace{2mm}
+
+$\mathcal{X} = \left[\begin{matrix}
+0 & 0 & \frac{1}{4\pi}\partial_x^2 & 0 \\
+0 & 0 & 0 & \frac{1}{4\pi} \\
+1 & 0 & 0 & 0 \\
+0 & 1 & 0 & 0 \\ \end{matrix} \right]$ \\
+
+\vspace{2mm}
+
+Note that this differs from what Umurhan+ find, although it shouldn't. The sign of the $\chi$ term, the $\partial_z$ in front of it, and the inclusion of the last term are all different. \\
+
+Now we will expand terms.
+Note the benefit of this notation: with the exception of $\mathcal{D}$, none of the matrices contain terms that change in the multiple scales expansion.
+Thus we are primarily concerned with expanding the $\partial_z$ terms: \\
+
+$\partial_z \to \partial_z + \epsilon \partial_Z$ \\
+
+$\partial_z^2 \to \partial_z^2 + 2\epsilon \partial_z \partial_Z + \epsilon^2 \partial_Z^2$ \\
+
+$\partial_z^3 \to \partial_z^3 + 3 \epsilon \partial_z^2 \partial_Z + 3 \epsilon^2 \partial_z \partial_Z^2 + \epsilon^3 \partial_Z^3$ \\
+
 \bibliography{mri.bib}

### 2014-03-14 notes/mri_notes.tex
more expansion
a little more
a lot more
final equation

@@
 
-Note that this differs from what Umurhan+ find, although it shouldn't. The sign of the $\chi$ term, the $\partial_z$ in front of it, and the inclusion of the last term are all different. \\
+Note that this differs from what Umurhan+ find, although it shouldn't. The sign of the $\mathcal{X}$ term, the $\partial_z$ in front of it, and the inclusion of the last term are all different. \\
 
@@
 
+$\partial_z^4 = \partial_z^4 + 4\epsilon \partial_z^3 \partial_Z + 6 \epsilon^2 \partial_z^2 \partial_Z^2 + 4 \epsilon^3 \partial_z \partial_Z^3 + \epsilon^4\partial_Z^4$ \\
+
+Thus $\mathcal{L}$ becomes $\mathcal{L^*}$, defined as: \\
+
+$\mathcal{L^*} = \mathcal{L}_0 \, + \, \mathcal{L}_1\partial_z \,+ \, \epsilon\mathcal{L}_1 \partial_Z \, + \, \mathcal{L}_2\partial_z^2 \, + \, \epsilon^2\mathcal{L}_2\partial_Z^2 \, + \, 2\epsilon\mathcal{L}_2\partial_z\partial_Z \, + \, \mathcal{L}_3\partial_z^3 \, +\, 3\epsilon \mathcal{L}_3 \partial_z^2 \partial_Z \,+\, 3 \epsilon^3\mathcal{L}_3\partial_Z^2\partial_z \,+\, \epsilon^3\mathcal{L}_3\partial_Z^3 \, + \, \mathcal{L}_4\partial_z^4 \, + \, 4 \epsilon \mathcal{L}_4 \partial_z^3 \partial_Z \, + \, 6 \epsilon^2\mathcal{L}_4 \partial_z^2 \partial_Z^2 \, + \, 4 \epsilon^3 \mathcal{L}_4 \partial_Z^3\partial_z \, + \, \epsilon^4\mathcal{L}_4\partial_Z^4$ \\
+
+This is just $\mathcal{L}$ plus 10 terms. Grouping those terms by $\mathcal{O}(\epsilon)$: \\
+
+$\mathcal{L^*} = \mathcal{L} + \epsilon\mathcal{L}_1\partial_Z \, + \, 2\epsilon\mathcal{L}_2\partial_z\partial_Z \, + \, 4 \epsilon\mathcal{L}_4\partial_z^3 \partial_Z \, + \, 3\epsilon\mathcal{L}_3\partial_z^2\partial_Z \, + \, \epsilon^2\mathcal{L}_2\partial_z^2 \, + \, 3\epsilon^2\mathcal{L}_3\partial_Z^2\partial_z \, + \, 6\epsilon^2\mathcal{L}_4 \partial_z^2 \partial_Z^2 \, + \, \epsilon^3\mathcal{L}_3\partial_Z^3 \, + \, 4 \epsilon^3\mathcal{L}_4\partial_z^3 \partial_z \, + \, \epsilon^4\mathcal{L}_4\partial_Z^4 $ \\
+
+Note that we have the same terms as the $\widetilde{\mathcal{L}}_1$ operator in Umurhan+, but we do not have the same terms as their $\widetilde{\mathcal{L}}_2$. \\
+
+The $\partial_z\mathcal{X}$ and $\partial_z^3\mathcal{L}_3$ terms become: \\
+
+$-\epsilon^2\partial_z\mathcal{X} \to \left(-\epsilon^2 - \epsilon^3\partial_Z\right)\mathcal{X}$ \\
+
+$-\epsilon^2\partial_z^3\mathcal{L}_3 \to \left(-\epsilon^2\partial_z^3 - 3\epsilon^3\partial_z^2\partial_Z - 3\epsilon^4\partial_z\partial_Z^2 - \epsilon^5\partial_Z^3\right)\mathcal{L}_3$ \\
+
+And $\mathcal{D}$ becomes: \\
+
+$\mathcal{D} \to \left[\begin{matrix}
+\nabla^2 & 0 & 0 & 0 \\
+0 & 1& 0 & 0 \\
+0 & 0 & 1 & 0\\
+0 & 0 & 0 & 1 \\
+\end{matrix}\right] + \left(\epsilon^2\partial_Z^2 + 2\epsilon\partial_z\partial_Z\right)\left[\begin{matrix}
+1 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\ \end{matrix}\right]$ \\
+
+\vspace{2mm}
+
+We will call this $\mathcal{D} + \left(\epsilon^2\partial_Z^2 + 2\epsilon\partial_z\partial_Z\right)\mathcal{D^*}$. \\
+
+Our final equation is: \\
+
+$\left(\partial_t + \epsilon^2 \partial_T \right)\mathcal{D}\mathbf{V} \, + \, \left(\partial_t + \epsilon^2\partial_T\right)\left(\epsilon^2\partial_Z^2 + 2\epsilon\partial_z\partial_Z\right)\mathcal{D^*}\mathbf{V} \, + \, \mathbf{N} \, = \, \mathcal{L}\mathbf{V} \, + \, \epsilon\widetilde{\mathcal{L}}_1\partial_Z\mathbf{V} \, + \, \epsilon^2\widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V} \, + \, \epsilon^3\widetilde{\mathcal{L}}_3\partial_Z^3\mathbf{V} \, + \, \epsilon^4\widetilde{\mathcal{L}}_4\partial_Z^4\mathbf{V} \, - \, \epsilon^2\partial_z\mathcal{X}\mathbf{V} \, - \epsilon^3\partial_Z\mathcal{X}\mathbf{V} \, - \, \epsilon^2\partial_z^3\mathcal{L}_3\mathbf{V} \, - \, 3\epsilon^3\partial_z^2\partial_Z\mathcal{L}_3\mathbf{V} \, - \, 3\epsilon^4\partial_z\partial_Z^2\mathcal{L}_3\mathbf{V} \, - \, \epsilon^5\partial_Z^3\mathcal{L}_3\mathbf{V}$ \\
+
+Where: \\
+
+$\widetilde{\mathcal{L}}_1 = \mathcal{L}_1 + 2\mathcal{L}_2\partial_z + 3\mathcal{L}_3\partial_z^2 + 4\mathcal{L}_4\partial_z^3$ \\
+
+$\widetilde{\mathcal{L}}_2 = \mathcal{L}_2 + 3\mathcal{L}_3\partial_z + 6\mathcal{L}_4\partial_z^2$ \\
+
+$\widetilde{\mathcal{L}}_3 = \mathcal{L}_3 + 4\mathcal{L}_4\partial_z$ \\
+
+$\widetilde{\mathcal{L}}_4 = \mathcal{L}_4$ \\
+
+To $\mathcal{O}(\epsilon^2)$ this becomes: \\
+
+$\partial_t\mathcal{D}\mathbf{V} \, + \, \epsilon^2 \partial_T \mathcal{D}\mathbf{V} \, + \, \epsilon^2\partial_Z^2\partial_t\mathcal{D^*}\mathbf{V} \, + \, 2\epsilon\partial_z\partial_Z\partial_t\mathcal{D^*}\mathbf{V} \, + \, \mathbf{N} \, = \, \mathcal{L}\mathbf{V} \, + \, \epsilon\widetilde{\mathcal{L}}_1\partial_Z\mathbf{V} \, + \, \epsilon^2\widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V} \, - \, \epsilon^2\partial_z\mathcal{X}\mathbf{V} \, - \, \epsilon^2\partial_z^3\mathcal{L}_3\mathbf{V} \, + \, \mathcal{O}(\epsilon^3)$
+
 \bibliography{mri.bib}

### 2014-03-15 notes/mri_notes.tex
nonlinear vector
more nonlinear expansion

@@
 -J(A, \Psi) \\
-J(\Psi, B_{1y} - J(A, u_{1y}) \\
+J(\Psi, B_{1y}) - J(A, u_{1y}) \\
 \end{matrix} \right]$ \\
@@
 
-$\partial_t\mathcal{D}\mathbf{V} \, + \, \epsilon^2 \partial_T \mathcal{D}\mathbf{V} \, + \, \epsilon^2\partial_Z^2\partial_t\mathcal{D^*}\mathbf{V} \, + \, 2\epsilon\partial_z\partial_Z\partial_t\mathcal{D^*}\mathbf{V} \, + \, \mathbf{N} \, = \, \mathcal{L}\mathbf{V} \, + \, \epsilon\widetilde{\mathcal{L}}_1\partial_Z\mathbf{V} \, + \, \epsilon^2\widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V} \, - \, \epsilon^2\partial_z\mathcal{X}\mathbf{V} \, - \, \epsilon^2\partial_z^3\mathcal{L}_3\mathbf{V} \, + \, \mathcal{O}(\epsilon^3)$
+$\mathcal{D}\partial_t\mathbf{V} \, + \, \epsilon^2  \mathcal{D}\partial_T\mathbf{V} \, + \, \epsilon^2\mathcal{D^*}\partial_Z^2\partial_t\mathbf{V} \, + \, 2\epsilon\mathcal{D^*}\partial_z\partial_Z\partial_t\mathbf{V} \, + \, \mathbf{N} \, = \, \mathcal{L}\mathbf{V} \, + \, \epsilon\widetilde{\mathcal{L}}_1\partial_Z\mathbf{V} \, + \, \epsilon^2\widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V} \, - \, \epsilon^2\partial_z\mathcal{X}\mathbf{V} \, - \, \epsilon^2\partial_z^3\mathcal{L}_3\mathbf{V} \, + \, \mathcal{O}(\epsilon^3)$ \\
+
+What about the vector of nonlinear terms?
+Let $\mathbf{N} = \left[N^{\left(\Psi\right)}, N^{\left(u_{1y}\right)}, N^{\left(A\right)}, N^{\left(B_{1y}\right)} \right]^\mathbf{T}$.
+We begin with the term $N^{\left(\Psi\right)} = J\left(\Psi, \nabla^2\Psi\right) - J\left(A, \nabla^2A\right)$: \\
+
+Making the replacements: \\
+
+$\partial_z \to \partial_z + \epsilon\partial_Z$ \\
+
+$\nabla^2 \to \nabla^2 + \epsilon^2\partial_Z^2 + 2\epsilon\partial_z\partial_Z$ \\
+
+We find that $J\left(\Psi, \nabla^2\Psi\right) = \partial_z\Psi\partial_x\nabla^2\Psi - \partial_x\Psi\partial_z\nabla^2\Psi$ becomes: \\
+
+$\partial_z\Psi\partial_x\nabla^2\Psi \, - \, \partial_x\Psi\partial_z\nabla^2\Psi \, + \, \epsilon^2\partial_z\Psi\partial_x\partial_Z^2\Psi \, - \, \epsilon^2\partial_x\Psi\partial_z\partial_Z^2\Psi \, + \, 2\epsilon\partial_z\Psi\partial_x\partial_z\partial_Z\Psi \, - \, 2\epsilon\partial_x\Psi\partial_z^2\partial_Z\Psi \, + \, \epsilon\partial_Z\Psi\partial_x\nabla^2\Psi \, - \, \epsilon\partial_x\Psi\partial_Z\nabla^2\Psi \, + \, \epsilon^3\partial_Z\Psi\partial_x\partial_Z^2\Psi \, - \, \epsilon^3\partial_x\Psi\partial_Z^3\Psi \, + \, 2\epsilon^2\partial_Z\Psi\partial_x\partial_z\partial_Z\Psi \, - \, 2\epsilon^2\partial_x\Psi\partial_z\partial_Z^2\Psi$ \\
+
+$= \, J\left(\Psi, \nabla^2\Psi\right) \, + \, \epsilon^2J\left(\Psi, \partial_Z^2\Psi\right) \, + \, 2\epsilon J\left(\Psi, \partial_z\partial_Z\Psi\right) \, + \epsilon\widetilde{J}\left(\Psi, \nabla^2\Psi\right) \, + \, \epsilon^3\widetilde{J}\left(\Psi, \partial_Z^2\Psi\right) \, + \, 2\epsilon^2\widetilde{J}\left(\Psi, \partial_z\partial_Z\Psi\right)$ \\
+
+Where $\widetilde{J}\left(f, g\right) \equiv \partial_Z f \partial_x g - \partial_x f \partial_Z g$ \\
+
+When we expand the fluid quantities $\mathbf{V}$ in a perturbation series $\mathbf{V} = \epsilon\mathbf{V_1} + \epsilon^2\mathbf{V_2} + \epsilon^3\mathbf{V_3} + ...$, we will introduce additional $\epsilon$ terms.
+Because we are truncating our expansion at $\mathcal{O}(\epsilon^3)$, we will write out the full perturbation series expansion of the nonlinear vector now.
+\\
+
+Note that $J\left(a + b, c + d\right) = J(a, c) + J(a, d) + J(b, c) + J(b, d)$. \\
+
+Thus the first term of $N^{\left(\Psi\right)}$ expands to become: \\
+
+$J(\epsilon\Psi_1 + \epsilon^2\Psi_2, \epsilon\nabla^2 \Psi_1 + \epsilon^2 \nabla^2\Psi_2)$
 

### 2014-03-15 notes/mri_notes.tex
first nonlinear term fully

@@
 
-$J(\epsilon\Psi_1 + \epsilon^2\Psi_2, \epsilon\nabla^2 \Psi_1 + \epsilon^2 \nabla^2\Psi_2)$
+$J(\epsilon\Psi_1 + \epsilon^2\Psi_2, \epsilon\nabla^2 \Psi_1 + \epsilon^2 \nabla^2\Psi_2)$ \\
+
+$= J(\epsilon\Psi_1, \epsilon\nabla^2\Psi_1) + J(\epsilon\Psi_1, \epsilon^2\nabla^2\Psi_2) + J(\epsilon^2\Psi_2, \epsilon\nabla^2\Psi_1) + J(\epsilon^2\Psi_2, \epsilon^2\nabla^2\Psi_2)$ \\
+
+$= \epsilon^2 J(\Psi_1, \nabla^2\Psi_1) + \epsilon^3J(\Psi_1, \nabla^2\Psi_2) + \epsilon^3J(\Psi_2, \nabla^2\Psi_1) + \epsilon^4J(\Psi_2, \nabla^2\Psi_2)$ \\
+
+And the term $2\epsilon J(\epsilon\Psi_1 + \epsilon^2\Psi_2, \epsilon\partial_z\partial_Z\Psi_1 + \epsilon^2\partial_z\partial_Z\Psi_2)$ becomes: \\
+
+$= 2\epsilon^3 J(\Psi_1, \partial_z\partial_Z\Psi_1) + 2\epsilon^4 J(\Psi_1, \partial_z\partial_Z\Psi_2) + 2\epsilon^4 J(\Psi_2, \partial_z\partial_Z\Psi_1) + 2\epsilon^5 J(\Psi_2, \partial_z\partial_Z \Psi_2)$ \\
+
+And the term $\epsilon \widetilde{J}(\epsilon\Psi_1 + \epsilon\Psi_2, \epsilon\nabla^2\Psi_1 + \epsilon^2\nabla^2\Psi_2)$ becomes: \\
+
+$= \epsilon^3 \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) + \epsilon^4\widetilde{J}(\Psi_1, \nabla^2\Psi_2) + \epsilon^4\widetilde{J}(\Psi_2, \nabla^2\Psi_1) + \epsilon^5\widetilde{J}(\Psi_2, \nabla^2 \Psi_2)$ \\
+
+As for the terms in $N^{(\Psi)}$ that are $\mathcal{O}(\epsilon^2)$ and higher, they will not contain any terms in the expansion that are lower than $\mathcal{O}(\epsilon^4)$. \\
+
+Thus the first term in the pre-expansion nonlinear vector $N^{(\Psi)}$ becomes: \\
+
+$J(\Psi, \nabla^2\Psi) \to \epsilon^2 J(\Psi_1, \nabla^2 \Psi_1) + \epsilon^3 J(\Psi_1, \nabla^2\Psi_2) + \epsilon^3 J(\Psi_2, \nabla^2\Psi_1) + 2\epsilon^3 J(\Psi_1, \partial_z\partial_Z \Psi_1) + \epsilon^3\widetilde{J}(\Psi_1, \nabla^2 \Psi_1) + \mathcal{O}(\epsilon^4)$ \\
+
+The second term behaves similarly, and thus $N^{(\Psi)}$ becomes: \\
+
+$N^{(\Psi)} \to \epsilon^2 J(\Psi_1, \nabla^2 \Psi_1) \,-\, \epsilon^2 J(A_1, \nabla^2 A_1) \,+\, \epsilon^3 J(\Psi_1, \nabla^2\Psi_2) \, - \epsilon^3J(A_1, \nabla^2 A_2) \,+\, \epsilon^3 J(\Psi_2, \nabla^2\Psi_1) \, - \, \epsilon^3 J(A_2, \nabla^2 A_1) \,+\, 2\epsilon^3 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2\epsilon^3 J(A_1, \partial_z\partial_Z A_1) \,+\, \epsilon^3\widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \, \epsilon^3 \widetilde{J}(A_1, \nabla^2 A_1) \, + \, \mathcal{O}(\epsilon^4)$ \\
+
+Note that this is the same as Umurhan+ except for a sign on one term: they have a typo in their Eq. B19. \\
 

### 2014-03-17 notes/mri_notes.tex
two more terms
finished nonlinear terms
perturbation expansion

@@
 
+\section*{Expansion of Nonlinear Terms}
+
 What about the vector of nonlinear terms?
@@
 
+Now we expand $N^{(u_{1y})} = J(\Psi, u_{1y}) - \frac{1}{4\pi} J(A, B_{1y})$: \\
+
+$= J(\Psi, u_{1y}) + \epsilon \widetilde{J}(\Psi, u_{1y}) - \frac{1}{4\pi}J(A, B_{1y}) \, - \, \frac{1}{4\pi}\epsilon \widetilde{J}(A, B_{1y})$ \\
+
+For the purposes of the perturbation series expansions, we will rewrite $u_{1y} \equiv u$ and $B_{1y} \equiv B$, with the understanding that $u$ and $B$ are the y-components of the velocity and magnetic field, respectively.
+\\
+
+$N^{(u_{1y})} \to \epsilon^2 J(\Psi_1, u_1) \, - \, \frac{1}{4\pi} \epsilon^2 J(A_1, B_1) \, + \, \epsilon^3 J(\Psi_1, u_2) \, + \, \epsilon^3 J(\Psi_2, u_1) \, + \, \epsilon^3 \widetilde{J}(\Psi_1, u_1) \, - \, \frac{1}{4\pi} \epsilon^3 J(A_1, B_2) \, - \, \frac{1}{4\pi} \epsilon^3 J(A_2, B_1) \, - \, \frac{1}{4\pi}\epsilon^3 \widetilde{J}(A_1, B_1) \, + \, \mathcal{O}(\epsilon^4)$ \\
+
+The $N^{(A)}$ term is simple: \\
+
+$N^{(A)} = -J(A, \Psi) = -J(A, \Psi) - \epsilon\widetilde{J}(A, \Psi)$ \\
+
+$N^{(A)} \to -\epsilon^2 J(A_1, \Psi_1) \, - \, \epsilon^3 J(A_1, \Psi_2) \, - \, \epsilon^3 J(A_2, \Psi_1) \, - \, \epsilon^4 J(A_2, \Psi_2) \, - \, \epsilon^3 \widetilde{J}(A_1, \Psi_1) \, - \, \epsilon^4 \widetilde{J}(A_1, \Psi_2) \, - \, \epsilon^4 \widetilde{J}(A_2, \Psi_1) \, - \, \epsilon^5 \widetilde{J}(A_2, \Psi_2)$ \\
+
+And finally, we expand $N^{(B_{1y})} = J(\Psi, B_{1y}) - J(A, u_{1y})$: \\
+
+$= J(\Psi, B_{1y}) + \epsilon\widetilde{J}(\Psi, B_{1y}) - J(A, u_{1y}) - \epsilon\widetilde{J}(A, u_{1y})$ \\
+
+$ = \epsilon^2 J(\Psi_1, B_1) \, + \, \epsilon^3 J(\Psi_1, B_2) \, + \, \epsilon^3 J(\Psi_2, B_1) \, + \, \epsilon^4 J(\Psi_2, B_2) \, + \, \epsilon^3\widetilde{J}(\Psi_1, B_1) \, + \, \epsilon^4\widetilde{J}(\Psi_1, B_2) \, + \, \epsilon^4 \widetilde{J}(\Psi_2, B_1) \, + \, \epsilon^5 \widetilde{J}(\Psi_2, B_2) \, - \, \epsilon^2 J(A_1, u_1) \, - \, \epsilon^3 J(A_1, u_2) \, - \, \epsilon^3 J(A_2, u_1) \, - \, \epsilon^4 J(A_2, u_2) \, - \, \epsilon^3 \widetilde{J}(A_1, u_1) \, - \, \epsilon^4 \widetilde{J}(A_1, u_2) \, - \, \epsilon^4 \widetilde{J}(\Psi_2, B_1) \, - \, \epsilon^5 \widetilde{J}(\Psi_2, B_2)$\\
+
+Thus we are left with the following: \\
+
+$\mathbf{N} = \epsilon^2\mathbf{N_1} \, + \, \epsilon^3\mathbf{N_2} \, + \, \mathcal{O}(\epsilon^4)$ \\
+
+Where: \\
+
+\noindent $N^{(\Psi)} = \epsilon^2 N_2^{(\Psi)} \, + \, \epsilon^3 N_3^{(\Psi)} \, + \, \mathcal{O}(\epsilon^4)$ \\
+$N^{(u)} = \epsilon^2 N_2^{(u)} \, + \, \epsilon^3 N_3^{(u)} \, + \, \mathcal{O}(\epsilon^4)$ \\
+$N^{(A)} = \epsilon^2 N_2^{(A)} \, + \, \epsilon^3 N_3^{(A)} \, + \, \mathcal{O}(\epsilon^4)$ \\
+$N^{(B)} = \epsilon^2 N_2^{(B)} \, + \, \epsilon^3 N_3^{(B)} \, + \, \mathcal{O}(\epsilon^4)$ \\
+
+And: \\
+
+$N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, J(A_1, \nabla^2 A_1) $ \\
+
+$N_2^{(u)} = J(\Psi_1, u_1) \, - \, \frac{1}{4\pi} J(A_1, B_1) $ \\
+
+$N_2^{(A)} = - J(A_1, \Psi_1)$ \\
+
+$N_2^{(B)} =  J(\Psi_1, B_1) \, - \, J(A_1, u_1)$\\
+
+And: \\
+
+$N_3^{(\Psi)} = J(\Psi_1, \nabla^2\Psi_2) \, - \, J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) \, - \, J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2 J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \widetilde{J}(A_1, \nabla^2 A_1) $ \\
+
+$N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \frac{1}{4\pi} J(A_1, B_2) \, - \, \frac{1}{4\pi} J(A_2, B_1) \, - \, \frac{1}{4\pi} \widetilde{J}(A_1, B_1)$ \\
+
+$N_3^{(A)} = - J(A_1, \Psi_2) \, - \, J(A_2, \Psi_1) \, - \, \widetilde{J}(A_1, \Psi_1) \,$ \\
+
+$N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1) \, $\\
+
+\section*{Perturbation Series Expansion}
+
+We now expand the fluid quantities $\mathbf{V}$ in a perturbation series $\mathbf{V} = \epsilon\mathbf{V_1} + \epsilon^2\mathbf{V_2} + \epsilon^3\mathbf{V_3} + ...$ \\
+
+We drop the $\partial_t$ terms and our $\mathcal{O}(\epsilon^2)$ equation is: \\
+
+$\epsilon^2 \mathcal{D}\partial_T\mathbf{V} \, + \, \mathbf{N} \, = \, \mathcal{L}\mathbf{V} \, + \, \epsilon\widetilde{\mathcal{L}}_1\partial_Z\mathbf{V} \, + \, \epsilon^2\widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V} \, - \, \epsilon^2\partial_z\mathcal{X}\mathbf{V} \, - \, \epsilon^2\partial_z^3\mathcal{L}_3\mathbf{V} \, + \, \mathcal{O}(\epsilon^3)$ \\
+
+Plugging in our $\mathbf{V}$ expansion, we find: \\
+
+$\epsilon^3 \mathcal{D} \partial_T \mathbf{V_1} \, + \, \epsilon^2\mathbf{N_2} \, + \epsilon^3 \mathbf{N_3} \, = \, \epsilon\mathcal{L}\mathbf{V_1} \, + \, \epsilon^2 \mathcal{L} \mathbf{V_2} \, + \, \epsilon^3 \mathcal{L}\mathbf{V_3} \, + \, \epsilon^2\widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1} \, + \epsilon^3 \widetilde{\mathcal{L}}_1\partial_Z \mathbf{V_2} \, + \, \epsilon^3 \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V_1} \, - \, \epsilon^3\partial_z \mathcal{X} \mathbf{V_1} \, - \, \epsilon^3\partial_z^3\mathcal{L}_3\mathbf{V_1} \, + \, \mathcal{O}(\epsilon^4)$ \\
+
+To $\mathcal{O}(\epsilon)$, we have: \\
+
+$\mathcal{L}\mathbf{V_1} = 0$ \\
+
+To $\mathcal{O}(\epsilon^2)$: \\
+
+$\mathcal{L}\mathbf{V_2} \, = \, \mathbf{N_2} \, - \, \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ \\
+
+To $\mathcal{O}(\epsilon^3)$: \\
+
+$\mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, - \partial_Z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1}$ \\
+
 \bibliography{mri.bib}

### 2014-03-31 notes/mri_notes.tex
second order
complete second order
NEW amplitude equation

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
 
@@
 
-$\partial_t \partial_z^2 \Psi \, + \, \partial_z^2\Psi\partial_x\partial_z\Psi \, + \, \partial_z\Psi \partial_x\partial_z^2 \Psi \, - \, \partial_z\partial_x\Psi\partial_z^2\Psi \, - \, \partial_x \Psi \partial_z^3 \Psi \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_z\partial_x\Pi \, + \, \frac{1}{4\pi} B_0\partial_z^3A \, + \, \frac{1}{4\pi}\partial_z^2 A \partial_x \partial_z A \, + \, \frac{1}{4\pi}\partial_zA\partial_x\partial_z^2 A \, - \, \partial{1}{4\pi}\partial_z\partial_x A \partial_z^2 A \, - \, \frac{1}{4\pi}\partial_xA\partial_z^3A \, + \, \frac{1}{\reye} \nabla^2 \partial_z^2 \Psi $ \\
+$\partial_t \partial_z^2 \Psi \, + \, \partial_z^2\Psi\partial_x\partial_z\Psi \, + \, \partial_z\Psi \partial_x\partial_z^2 \Psi \, - \, \partial_z\partial_x\Psi\partial_z^2\Psi \, - \, \partial_x \Psi \partial_z^3 \Psi \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_z\partial_x\Pi \, + \, \frac{1}{4\pi} B_0\partial_z^3A \, + \, \frac{1}{4\pi}\partial_z^2 A \partial_x \partial_z A \, + \, \frac{1}{4\pi}\partial_zA\partial_x\partial_z^2 A \, - \, \frac{1}{4\pi}\partial_z\partial_x A \partial_z^2 A \, - \, \frac{1}{4\pi}\partial_xA\partial_z^3A \, + \, \frac{1}{\reye} \nabla^2 \partial_z^2 \Psi $ \\
 
@@
 
-$\partial_t\partial_z^2\Psi \, + \, J\left(\Psi, \partial_z^2\Psi\right) \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_x\partial_z\Pi \, + \, \frac{1}{4\pi}B_0 \partial_z^3A \, + \, J\left(A, \partial_z^2A\right) \, + \, \frac{1}{\reye}\nabla^2 \partial_z^2 \Psi$ \\
+$\partial_t\partial_z^2\Psi \, + \, J\left(\Psi, \partial_z^2\Psi\right) \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_x\partial_z\Pi \, + \, \frac{1}{4\pi}B_0 \partial_z^3A \, + \, \frac{1}{4\pi}J\left(A, \partial_z^2A\right) \, + \, \frac{1}{\reye}\nabla^2 \partial_z^2 \Psi$ \\
 
@@
 
-$\partial_t \partial_x^2 \Psi \, + \, J\left(\Psi, \partial_x^2 \Psi\right) \, = \, -\frac{1}{4\pi}\partial_x\partial_z\Pi \, + \, \frac{1}{4\pi}B_0\partial_z\partial_x^2 A \, + \, J\left(A, \partial_x^2 A \right) \, + \, \frac{1}{\reye} \nabla^2 \partial_x^2 \Psi$ \\
+$\partial_t \partial_x^2 \Psi \, + \, J\left(\Psi, \partial_x^2 \Psi\right) \, = \, -\frac{1}{4\pi}\partial_x\partial_z\Pi \, + \, \frac{1}{4\pi}B_0\partial_z\partial_x^2 A \, + \, \frac{1}{4\pi}J\left(A, \partial_x^2 A \right) \, + \, \frac{1}{\reye} \nabla^2 \partial_x^2 \Psi$ \\
 
@@
 
-$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
+$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{4\pi}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
 
@@
 
-$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, J\left(A, \nabla^2 A \right) \, =  \, 2 \partial_z u_{1y} \,+ \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
+$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, \frac{1}{4\pi}J\left(A, \nabla^2 A \right) \, =  \, 2 \partial_z u_{1y} \,+ \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
 
@@
 
-$\mathbf{N} = \left[ \begin{matrix}  J(\Psi, \nabla^2 \Psi) - J(A, \nabla^2 A)\\
+$\mathbf{N} = \left[ \begin{matrix}  J(\Psi, \nabla^2 \Psi) - \frac{1}{4\pi}J(A, \nabla^2 A)\\
 J(\Psi, u_{1y}) - \frac{1}{4\pi} J(A, B_{1y}) \\ 
@@
 
-$N^{(\Psi)} \to \epsilon^2 J(\Psi_1, \nabla^2 \Psi_1) \,-\, \epsilon^2 J(A_1, \nabla^2 A_1) \,+\, \epsilon^3 J(\Psi_1, \nabla^2\Psi_2) \, - \epsilon^3J(A_1, \nabla^2 A_2) \,+\, \epsilon^3 J(\Psi_2, \nabla^2\Psi_1) \, - \, \epsilon^3 J(A_2, \nabla^2 A_1) \,+\, 2\epsilon^3 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2\epsilon^3 J(A_1, \partial_z\partial_Z A_1) \,+\, \epsilon^3\widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \, \epsilon^3 \widetilde{J}(A_1, \nabla^2 A_1) \, + \, \mathcal{O}(\epsilon^4)$ \\
+$N^{(\Psi)} \to \epsilon^2 J(\Psi_1, \nabla^2 \Psi_1) \,-\, \epsilon^2 \frac{1}{4\pi}J(A_1, \nabla^2 A_1) \,+\, \epsilon^3 J(\Psi_1, \nabla^2\Psi_2) \, - \epsilon^3\frac{1}{4\pi}J(A_1, \nabla^2 A_2) \,+\, \epsilon^3 J(\Psi_2, \nabla^2\Psi_1) \, - \, \epsilon^3 \frac{1}{4\pi}J(A_2, \nabla^2 A_1) \,+\, 2\epsilon^3 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2\epsilon^3 \frac{1}{4\pi}J(A_1, \partial_z\partial_Z A_1) \,+\, \epsilon^3\widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \, \epsilon^3 \frac{1}{4\pi}\widetilde{J}(A_1, \nabla^2 A_1) \, + \, \mathcal{O}(\epsilon^4)$ \\
 
@@
 
-$N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, J(A_1, \nabla^2 A_1) $ \\
+$N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \frac{1}{4\pi}J(A_1, \nabla^2 A_1) $ \\
 
@@
 
-$N_3^{(\Psi)} = J(\Psi_1, \nabla^2\Psi_2) \, - \, J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) \, - \, J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2 J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \widetilde{J}(A_1, \nabla^2 A_1) $ \\
+$N_3^{(\Psi)} = J(\Psi_1, \nabla^2\Psi_2) \, - \, \frac{1}{4\pi}J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) \, - \, \frac{1}{4\pi}J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2 \frac{1}{4\pi}J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \frac{1}{4\pi}\widetilde{J}(A_1, \nabla^2 A_1) $ \\
 
@@
 
+The solution to the $\mathcal{O}(\epsilon)$ equation is: \\
+
+$\mathbf{V_1} = \alpha(T, Z) \mathbb{V}_{11} e^{i Q z} + c.c. + \beta(T, Z)\mathbb{U}_{11}$ \\
+
+Where $\alpha$ and $\beta$ are amplitudes, $\mathbb{V}_{11}(x) = (\Psi_{11}, u_{11}, A_{11}, B_{11})^\mathbf{T}$ and $\mathbb{U}_{11} = (0, 0, 0, 1)^\mathbf{T}$. \\
+
+To find the $\mathcal{O}(\epsilon^2)$ solution, must write out $\mathbf{N_2}$. \\
+
+We find: \\
+
+$\mathbf{N_2} = \alpha^2\mathbb{N}_{22} e^{i2Qz} + \left|\alpha\right|^2 \mathbb{N}_{20} + c.c.$ \\
+
+Then: \\
+
+$\mathcal{L}\mathbf{V_2} \, = \, \mathbf{N_2} \, - \, \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ \\
+
+$\mathbf{V_2} \, = \, \alpha^2 e^{i2Qz} \mathcal{L}^{-1}\mathbb{N}_{22} \, + \,  \left|\alpha\right|^2 \mathcal{L}^{-1} \mathbb{N}_{20}\, - \, \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ + c.c.\\
+
+And $\partial_Z \mathbf{V_1} = \partial_Z \alpha \mathbb{V}_{11} e^{i Q z} \, + \, \partial_Z \beta \mathbb{U}_{11} \, + \, c.c.$, so we end up with: \\
+
+$\mathbf{V_2} \, = \, \alpha^2 e^{i2Qz} \mathbb{V}_{22} \, + \, \left|\alpha\right|^2 \mathbb{V}_{20} \, - \, \partial_Z \alpha \mathbb{V}_{21} e^{iQz} \, - \, \partial_Z \beta \mathbb{U}_{20} + c.c.$ \\
+
+Where: \\
+
+$\mathbb{V}_{22} = \mathcal{L}^{-1}\mathbb{N}_{22}$ \\
+
+$\mathbb{V}_{20} = \mathcal{L}^{-1} \mathbb{N}_{20}$ \\
+
+$\mathbb{V}_{21} = \mathcal{L}^{-1} \widetilde{\mathcal{L}}_1 \mathbb{V}_{11}$ \\
+
+$\mathbb{U}_{20} = \mathcal{L}^{-1} \widetilde{\mathcal{L}}_1 \mathbb{U}_{11}$ \\
+
+Note that this is the same as Umurhan+ (B27) except for two signs. \\
+
+Our $\mathcal{O}(\epsilon^3)$ equation is: \\
+
+$\mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, - \partial_Z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1}$ \\
+
+Taking only the terms which are resonant with $e^{iQz}$, we find: \\
+
+$\mathcal{D} \mathbb{V}_{11} e^{iQz}\partial_T \alpha \, + \, \alpha \left| \alpha^2 \right| \mathbf{N}_{31} e^{iQz} \, + \, \alpha \partial_Z \beta \widetilde{\mathbf{N}}_{31} e^{iQz} \, = \, - \partial_Z^2 \alpha \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} e^{iQz} \, + \, \partial_Z^2 \alpha \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{iQz} \, - \, \partial_Z \alpha \mathbb{V}_{11}\mathcal{X}e^{iQz} \, - \, iQ^3 \alpha \mathcal{L}_3 \mathbb{V}_{11} e^{iQz}$ \\
+
+Note that $\mathcal{L}\mathbf{V}_3$ does not enter here.\\
+
+Following Umurhan+'s lead and taking the inner product with $\mathbb{V}^\dagger$, I find the amplitude equation: \\
+
+$a \partial_T \alpha \, + \, c \alpha \left|\alpha^2\right| \, + \, \widetilde{c} \alpha \partial_Z \beta \, = \, - b \partial_Z \alpha \, + \, h \partial_Z^2 \alpha \, - \, g i Q^3 \alpha$ \\
+
+Where: \\
+
+$a \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{D}\mathbb{V}^*_{11}\right>$ \\
+
+$c \equiv \left< \mathbb{V}^\dagger \cdot \mathbf{N}^*_{31}\right>$ \\
+
+$\widetilde{c} \equiv \left< \mathbb{V}^\dagger \cdot \widetilde{\mathbf{N}}^*_{31}\right>$ \\
+
+$b \equiv \left< \mathbb{V}^\dagger \cdot \left(\mathcal{X}\mathbb{V}_{11}\right)^* \right>$ \\
+
+$h \equiv \left< \mathbb{V}^\dagger \cdot \left(\widetilde{\mathcal{L}}_2\mathbb{V}_{11} \, - \, \widetilde{\mathcal{L}}_1\mathbb{V}_{21} \right)^* \right>$ \\
+
+$g \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{L}_{3} \mathbb{V}_{11} \right>$ \\
+
 \bibliography{mri.bib}

### 2014-04-01 notes/mri_notes.tex
more details fille in
explanation of second amplitude equation

@@
 
+Which, substituting in the forms of $\mathbf{V_1}$ and $\mathbf{V_2}$ and taking Umurhan+'s $\mathbf{N3}$, gives us the full amplitude equation:\\
+
+$\mathcal{D}\mathbb{V}_{11} e^{iQz}\partial_T \alpha \, + \, \mathcal{D}\mathbb{U}_{11} \partial_T \beta \, + \, \alpha^3 \mathbf{N_{33}} e^{i3Qz} \, + \, \alpha \partial_Z \alpha \mathbf{N_{32}} e^{i2Qz} \, + \, \alpha\left|\alpha\right|^2 \mathbf{N_{31}} e^{iQz} \, + \, \alpha \partial_Z\beta \widetilde{\mathbf{N}}_{31} e^{iQz} \, + \, \alpha^* \partial_Z \alpha \mathbf{N_{30}} \, = \, \mathcal{L}\mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{22} 2 \alpha \partial_Z \alpha e^{i2Qz} \, + \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{20} 2 \left|\alpha\right| \partial_Z \alpha \, - \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} \partial_Z^2 \alpha e^{iQz} \, - \, \widetilde{\mathcal{L}}_1 \mathbb{U}_{20} \partial_Z^2 \beta \, + \, \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{iQz} \partial_Z^2 \alpha \, + \, \widetilde{\mathcal{L}}_2 \mathbb{U}_{11} \partial_Z^2 \beta \, - \, \mathcal{X}\mathbb{V}_{11} e^{iQz} \partial_Z \alpha \, - \, \mathcal{X} \mathbb{U}_{11} \partial_Z \beta \, + \, \mathcal{L}_3 \mathbb{V}_{11} e^{iQz} iQ^3 \alpha $\\
+
 Taking only the terms which are resonant with $e^{iQz}$, we find: \\
 
-$\mathcal{D} \mathbb{V}_{11} e^{iQz}\partial_T \alpha \, + \, \alpha \left| \alpha^2 \right| \mathbf{N}_{31} e^{iQz} \, + \, \alpha \partial_Z \beta \widetilde{\mathbf{N}}_{31} e^{iQz} \, = \, - \partial_Z^2 \alpha \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} e^{iQz} \, + \, \partial_Z^2 \alpha \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{iQz} \, - \, \partial_Z \alpha \mathbb{V}_{11}\mathcal{X}e^{iQz} \, - \, iQ^3 \alpha \mathcal{L}_3 \mathbb{V}_{11} e^{iQz}$ \\
+$\mathcal{D} \mathbb{V}_{11} e^{iQz}\partial_T \alpha \, + \, \alpha \left| \alpha^2 \right| \mathbf{N}_{31} e^{iQz} \, + \, \alpha \partial_Z \beta \widetilde{\mathbf{N}}_{31} e^{iQz} \, = \, - \partial_Z^2 \alpha \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} e^{iQz} \, + \, \partial_Z^2 \alpha \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{iQz} \, - \, \partial_Z \alpha \mathbb{V}_{11}\mathcal{X}e^{iQz} \, + \, iQ^3 \alpha \mathcal{L}_3 \mathbb{V}_{11} e^{iQz}$ \\
 
@@
 
-$a \partial_T \alpha \, + \, c \alpha \left|\alpha^2\right| \, + \, \widetilde{c} \alpha \partial_Z \beta \, = \, - b \partial_Z \alpha \, + \, h \partial_Z^2 \alpha \, - \, g i Q^3 \alpha$ \\
+$a \partial_T \alpha \, + \, c \alpha \left|\alpha^2\right| \, + \, \widetilde{c} \alpha \partial_Z \beta \, = \, - b \partial_Z \alpha \, + \, h \partial_Z^2 \alpha \, + \, g i Q^3 \alpha$ \\
 
@@
 
+There should be a second amplitude equation, found by taking all of the terms which have no z-dependence. These terms are: \\
+
+$\mathcal{D}\mathbb{U}_{11} \partial_T \beta \, = \, \mathcal{L}\mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{20} 2 \left|\alpha\right| \partial_Z \alpha \, - \, \widetilde{\mathcal{L}}_1 \mathbb{U}_{20} \partial_Z^2 \beta \, + \, \widetilde{\mathcal{L}}_2 \mathbb{U}_{11} \partial_Z^2 \beta \, - \, \mathcal{X} \mathbb{U}_{11} \partial_Z \beta \, $\\
+
+And for the (B)-component of that equation, I find: \\
+
+$\partial_T \beta \, = \, -\widetilde{u}_{20} \partial_Z^2 \beta \, + \, \frac{1}{\reym} \partial_Z^2 \beta \, - \, \alpha^* \partial_Z \alpha N_{30}^{(B)} \, + \, 2 \left|\alpha\right| \partial_Z \left| \alpha \right| \left(u_{20} \, - \, q\Omega_0A_{20}\right) \, + \, \frac{1}{\reym} \partial_x^2 B_{30} \, + \, \frac{1}{\reym} \partial_z^2 B_{30}$ \\
+
+The fourth term on the RHS, which couples to the shear, does not appear in Umurhan+, and we need to investigate whether something in the lower-order solvability criterions allow this term to drop out.
+\\
+
 \bibliography{mri.bib}

### 2014-04-07 notes/mri_notes.tex
Fredholm alternative
Fredholm alternative+

@@
 
+Some thoughts on the solvability conditions: \\
+
+The solvability condition arises from the Fredholm alternative, which I looked up in Guenther \& Lee (1988). It states that if you have: \\
+
+$M Y = F$ and $M^\dagger Z = G$ \\
+
+And the corresponding homogenous equations, and if $M^\dagger Z = 0$ has a solution, then $M Y = F$ has a solution if and only if $\left<F \cdot Z \right> = 0$, where Z is the nontrivial solution to the adjoint homogenous equation.
+\\
+
+In our case, we have: \\
+
+$\mathcal{L}\mathbf{V_2} \, = \mathbf{N_2} - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ \\
+
+Where $\widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ contains a term resonant with $e^{iQz}$. This is a secular term because it is a solution of the homogenous equation. \\
+
 \bibliography{mri.bib}

### 2014-04-07 notes/mri_notes.tex
orthogonality argument

@@
 
-And the corresponding homogenous equations, and if $M^\dagger Z = 0$ has a solution, then $M Y = F$ has a solution if and only if $\left<F \cdot Z \right> = 0$, where Z is the nontrivial solution to the adjoint homogenous equation.
+And \\
+
+$M Y = 0$ and $M^\dagger Z = 0$, \\
+
+If $M^\dagger Z = 0$ has a solution, then $M Y = F$ has a solution if and only if $\left<F \cdot Z \right> = 0$, where Z is the nontrivial solution to the adjoint homogenous equation. \\
+
+In our case, we have: \\
+
+$\mathcal{L}\mathbf{V_2} \, = \mathbf{N_2} - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ \\
+
+$\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$ \\
+
+Where $\widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ contains a term resonant with $e^{iQz}$. This is a secular term because it is a solution of the homogenous equation. \\
+
+Because $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$ has a nontrivial solution, the following solvability criterion must be satisfied in order for $\mathcal{L}\mathbf{V_2} \, = \mathbf{N_2} - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ to have a solution: \\
+
+$\left<\left(\mathbf{N_2} - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}\right) \cdot \mathbf{V}^\dagger\right> = 0$ \\
+
+Which can be rewritten: \\
+
+$\left<\mathbf{N_2} \cdot \mathbf{V}^\dagger \right> \, + \, \left< \mathbf{V}^\dagger \cdot \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}\right> = 0 $ \\
+
+Now, since: \\
+
+$\mathbf{V_1} = \alpha(T, Z) \mathbb{V}_{11} e^{i Q z} + c.c. + \beta(T, Z)\mathbb{U}_{11}$ \\
+
+$\mathbf{V}^\dagger = \mathbb{V}^\dagger e^{iQz} + c.c.$ \\
+
+And: \\
+
+$\mathbf{N_2} = \alpha^2\mathbb{N}_{22} e^{i2Qz} + \left|\alpha\right|^2 \mathbb{N}_{20} + c.c.$ \\
+
+All that matters is that $e^{iQz}$ is orthogonal to $e^{2iQz}$ but NOT orthogonal to $e^{iQz}$, that is: \\
+
+$\left< e^{i2Qz} \cdot e^{iQz} \right> \, = \, \int \limits_{-\pi/Q}^{\pi/Q} \left(e^{i2Qz}e^{-iQz^*}\right) \, dz \, = \, 0$\\
+
+$\left< e^{iQz} \cdot e^{iQz} \right> \, = \, \int \limits_{-\pi/Q}^{\pi/Q} \left(e^{iQz}e^{-iQz^*}\right) \, dz \, \ne \, 0$\\
+
+Thus the solvability condition is automatically satisfied by the nonsecular terms, and so must be enforced on the secular terms only.
+This is functionally equivalent to setting the coefficients of the secular terms to zero, but this provides a rigorous justification for doing so.
 \\
 
-In our case, we have: \\
-
-$\mathcal{L}\mathbf{V_2} \, = \mathbf{N_2} - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ \\
-
-Where $\widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ contains a term resonant with $e^{iQz}$. This is a secular term because it is a solution of the homogenous equation. \\
-
 \bibliography{mri.bib}

### 2014-04-07 notes/mri_notes.tex
symmetry. I hope this makes sense

@@
 \usepackage{amsfonts}
+\usepackage[makeroom]{cancel}
 
@@
 
+Because the inner product integrations are actually performed over both z and x, the even/oddness of the fluid quantity eigenfunctions can force the inner product to be zero.
+Thus for the solvability condition: \\
+
+ $\left< \mathbf{V}^\dagger \cdot \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}\right> = 0 $ \\
+
+ $\left< \left(\mathbb{V}^\dagger e^{iQz} + c.c.\right) \cdot \left(\widetilde{\mathcal{L}}_1 \partial_Z \alpha(T, Z) \mathbb{V}_{11} e^{i Q z} + c.c.
++ \widetilde{\mathcal{L}}_1 \partial_Z \beta(T, Z)\mathbb{U}_{11}\right) \right> = 0$ \\
+
+ $\iint \mathbb{V}^\dagger e^{iQz} \cdot \widetilde{\mathcal{L}}_1 \partial_Z \alpha(T, Z)^* \mathbb{V}_{11}^* e^{-i Q z^*} \, dx dz \, + \, \cancel{\iint \mathbb{V}^\dagger e^{iQz} \cdot \widetilde{\mathcal{L}}_1 \partial_Z \alpha(T, Z) \mathbb{V}_{11} e^{i Q z} \, dx dz} \, + $ \\
+$\cancel{\iint \mathbb{V}^\dagger e^{iQz} \cdot \widetilde{\mathcal{L}}_1 \partial_Z \beta(T, Z)^*\mathbb{U}_{11}^* \, dx dz} \, + \, \cancel{\iint {\mathbb{V}^\dagger}^* e^{-iQz^*} \cdot \widetilde{\mathcal{L}}_1 \partial_Z \alpha(T, Z)^* \mathbb{V}_{11}^* e^{-i Q z^*} \, dx dz} \, + $\\
+\noindent $\iint {\mathbb{V}^\dagger}^* e^{-iQz^*} \cdot \widetilde{\mathcal{L}}_1 \partial_Z \alpha(T, Z) \mathbb{V}_{11} e^{i Q z} \, dxdz \, + \, \cancel{\iint {\mathbb{V}^\dagger}^* e^{-iQz^*} \cdot \widetilde{\mathcal{L}}_1 \partial_Z \beta(T, Z)^*\mathbb{U}_{11}^* \, dx dz} \, = \, 0$ \\
+
+Where terms with $\beta$ and terms with either $e^{iQz}\cdot e^{iQz}$ or  $e^{-iQz^*}\cdot e^{-iQz^*}$ are zero.
+Since $\int \limits_{-\pi/Q}^{\pi/Q} \left(e^{iQz}e^{-iQz^*}\right) \, dz \, \ne \, 0$, we have the condition: \\
+
+$\int \limits_{-1}^1 \mathbb{V}^\dagger \cdot \widetilde{\mathcal{L}}_1 \partial_Z \alpha(T, Z)^* \mathbb{V}_{11}^*\, dx \, + \, \int \limits_{-1}^1 {\mathbb{V}^\dagger}^* \cdot \widetilde{\mathcal{L}}_1 \partial_Z \alpha(T, Z) \mathbb{V}_{11} \, dx \, = \, 0$ \\
+
+I believe one can state this as ``$\mathbb{V}^\dagger$ must be orthogonal to $\widetilde{\mathcal{L}}_1 \mathbb{V}_{11}$".
+Because both $\mathbb{V}^\dagger$ and $\mathbb{V}_{11}$ are even symmetric about x = 0, the requirement is really that $\widetilde{\mathcal{L}}_1 \mathbb{V}_{11}$ is odd symmetric with respect to x, XOR $\partial_Z \alpha(T, Z)$ is odd symmetric with respect to x = 0.
+
 \bibliography{mri.bib}

### 2014-04-07 notes/mri_notes.tex
fixed

@@
 I believe one can state this as ``$\mathbb{V}^\dagger$ must be orthogonal to $\widetilde{\mathcal{L}}_1 \mathbb{V}_{11}$".
-Because both $\mathbb{V}^\dagger$ and $\mathbb{V}_{11}$ are even symmetric about x = 0, the requirement is really that $\widetilde{\mathcal{L}}_1 \mathbb{V}_{11}$ is odd symmetric with respect to x, XOR $\partial_Z \alpha(T, Z)$ is odd symmetric with respect to x = 0.
+Because both $\mathbb{V}^\dagger$ and $\mathbb{V}_{11}$ are even symmetric about x = 0, the requirement is really that $\widetilde{\mathcal{L}}_1 \mathbb{V}_{11}$ is odd symmetric with respect to x, since $\partial_Z \alpha(T, Z)$ is presumably not zero everywhere.
 

### 2014-04-09 notes/mri_notes.tex
a little more

@@
 
-Which, substituting in the forms of $\mathbf{V_1}$ and $\mathbf{V_2}$ and taking Umurhan+'s $\mathbf{N3}$, gives us the full amplitude equation:\\
+Which, substituting in the forms of $\mathbf{V_1}$ and $\mathbf{V_2}$ and taking Umurhan+'s $\mathbf{N_3}$, gives us the full amplitude equation:\\
 
@@
 Because both $\mathbb{V}^\dagger$ and $\mathbb{V}_{11}$ are even symmetric about x = 0, the requirement is really that $\widetilde{\mathcal{L}}_1 \mathbb{V}_{11}$ is odd symmetric with respect to x, since $\partial_Z \alpha(T, Z)$ is presumably not zero everywhere.
+\\
+
+This solvability condition is enforceable on all terms resonant with $e^{iQz}$.
+That is why we take the inner product of $\mathbb{V}^\dagger$ with the $e^{iQz}$ terms in the $\mathcal{O}(\epsilon^3)$ equation to arrive at the first amplitude equation (see above).
+\\
+
+The second solvability condition is found for the terms without $e^{iQz}$ dependence: \\
+
+$\mathcal{D}\mathbb{U}_{11} \partial_T \beta \, = \, \mathcal{L}\mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{20} 2 \left|\alpha\right| \partial_Z \alpha \, - \, \widetilde{\mathcal{L}}_1 \mathbb{U}_{20} \partial_Z^2 \beta \, + \, \widetilde{\mathcal{L}}_2 \mathbb{U}_{11} \partial_Z^2 \beta \, - \, \mathcal{X} \mathbb{U}_{11} \partial_Z \beta \, $\\
 

### 2014-05-29 notes/mri_notes.tex
adjoint homogenous eqn
equations coded in dedalus

@@
 %\setlength{\bibsep}{0.0pt}
+\usepackage{courier}
 
@@
 
+Now we will actually solve the adjoint homogenous first-order equation, that is: \\
+
+$\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$ \\
+
+Where $\mathcal{L}^\dagger = \mathcal{L}_0^T \, - \, \mathcal{L}_1^T \partial_z \, + \, \mathcal{L}_2^T\partial_z^2 \, - \, \mathcal{L}_3^T \partial_z^3 \, + \, \mathcal{L}_4^T \partial_z^4$ \\
+
+And $\mathcal{L}_0^T = \mathcal{L}_0$, $\mathcal{L}_2^T = \mathcal{L}_2$, $\mathcal{L}_4^T = \mathcal{L}_4$ because these are diagonal matrices. We also have:\\
+
+$\mathcal{L}_1^T = \left[\begin{matrix}
+0 & 2-q & 1 & 0 \\
+2 & 0 & 0 & 1 \\
+\frac{1}{4\pi} \partial_x^2 & 0 & 0 & -q\\
+0 & \frac{1}{4\pi} & 0 & 0 \\
+\end{matrix}\right]$ \\
+
+$\mathcal{L}_3^T = \left[\begin{matrix}
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+\frac{1}{4\pi} & 0 & 0 & 0\\
+0 & 0 & 0 & 0 \\
+\end{matrix}\right]$ \\
+
+Writing out the expressions, where $\mathbf{V}^\dagger = \left[\Psi^\dagger, u^\dagger, A^\dagger, B^\dagger\right]^T$, we have: \\
+
+$\frac{1}{\reye}\partial_x^4 \Psi^\dagger \, - \, (2-q)\partial_z u^\dagger \, - \, \partial_z A^\dagger \, + \, \frac{2}{\reye}\partial_x^2 \partial_z^2 \Psi^\dagger \, + \, \frac{1}{\reye}\partial_z^4 \Psi^\dagger \, = \, 0$ \\
+
+$\frac{1}{\reye}\partial_x^2 u^\dagger \, - \, 2 \partial_z \Psi^\dagger \, - \, \partial_z B^\dagger, \, + \, \frac{1}{\reye}\partial_z^2 u^\dagger \, = \, 0$ \\
+
+$\frac{1}{\reym} \partial_x^2 A^\dagger \, - \, \frac{1}{4\pi} \partial_x^2 \partial_z \Psi^\dagger \, + \, q \partial_z B^\dagger  \,+ \, \frac{1}{\reym} \partial_z^2 A^\dagger \, - \, \frac{1}{4\pi} \partial_z^3 \Psi^\dagger \, = \, 0$ \\
+
+$\frac{1}{\reym} \partial_x^2 B^\dagger \, - \, \frac{1}{4\pi} \partial_z u^\dagger \, + \, \frac{1}{\reym} \partial_z^2 B^\dagger \, = \, 0$ \\
+
+We will use the linear eigenvalue solver in Dedalus to determine $\mathbf{V}^\dagger$. \\
+
+At marginality, the z-dependence is given by $e^{iQz}$. Thus the above equations become a system of PDEs in x for the fluid quantities: \\
+
+$\frac{1}{\reye}\partial_x^4 \Psi^\dagger \, - \, (2 - q) i Q u^\dagger \, - \,  i Q A^\dagger \, - \, Q^2 \frac{2}{\reye} \partial_x^2 \Psi^\dagger \, + \, Q^4 \frac{1}{\reye} \Psi^\dagger \, = \, 0$ \\
+
+$\frac{1}{\reye}\partial_x^2 u^\dagger \, - \, 2 i Q \Psi^\dagger \, - \, i Q B \, - \frac{1}{\reye} Q^2 u^\dagger \, = \, 0$ \\
+
+$\frac{1}{\reym} \partial_x^2 A^\dagger \, - \, \frac{1}{4\pi} i A \partial_x^2 \Psi^\dagger \, + \, q i Q B^\dagger \, - \, Q^2 \frac{1}{\reym} A^\dagger \, + \, i Q^3 \frac{1}{4\pi} \Psi^\dagger \, = \, 0$ \\
+
+$\frac{1}{\reym} \partial_x^2 B^\dagger \, - \, \frac{1}{4\pi} i Q u^\dagger \, - \, Q^2 \frac{1}{\reym} B^\dagger \, = \, 0$ \\
+
+The above is coded up in \texttt{adjoint\_homogenous.py}.
+
 \bibliography{mri.bib}

### 2014-06-06 scrap.tex
coarse grid run

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{color}
\usepackage{float}

\begin{document}
The whole dispersion relation:\\
$B_{0}^{4} C^{2} k_{x}^{4} k_{z}^{4} + 2 B_{0}^{4} C^{2} k_{x}^{2} k_{z}^{6} + B_{0}^{4} C^{2} k_{z}^{8} - 2 B_{0}^{2} C \Omega^{2} k_{x}^{2} k_{z}^{4} q - 2 B_{0}^{2} C \Omega^{2} k_{z}^{6} q - \frac{B_{0}^{2} C \Omega k_{x}^{4} k_{z}^{4} q}{Rm} + 2 \frac{B_{0}^{2} C \Omega k_{x}^{4} k_{z}^{4}}{Rm} + \frac{B_{0}^{2} C \Omega k_{x}^{3} k_{z}^{5} q}{Rm} - 2 \frac{B_{0}^{2} C \Omega k_{x}^{3} k_{z}^{5}}{Rm} - \frac{B_{0}^{2} C \Omega k_{x}^{2} k_{z}^{6} q}{Rm} + 2 \frac{B_{0}^{2} C \Omega k_{x}^{2} k_{z}^{6}}{Rm} + \frac{B_{0}^{2} C \Omega k_{x} k_{z}^{7} q}{Rm} - 2 \frac{B_{0}^{2} C \Omega k_{x} k_{z}^{7}}{Rm} + \frac{B_{0}^{2} C \Omega k_{x}^{4} k_{z}^{4} q}{R} - \frac{B_{0}^{2} C \Omega k_{x}^{3} k_{z}^{5} q}{R} + \frac{B_{0}^{2} C \Omega k_{x}^{2} k_{z}^{6} q}{R} - \frac{B_{0}^{2} C \Omega k_{x} k_{z}^{7} q}{R} + 2 B_{0}^{2} C k_{x}^{4} k_{z}^{2} s^{2} + 4 B_{0}^{2} C k_{x}^{2} k_{z}^{4} s^{2} + 2 B_{0}^{2} C k_{z}^{6} s^{2} - \frac{B_{0}^{2} C k_{x}^{5} k_{z}^{3} s}{Rm} + 2 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} s}{Rm} - 2 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} s}{Rm} + 2 \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} s}{Rm} - \frac{B_{0}^{2} C k_{x} k_{z}^{7} s}{Rm} - \frac{B_{0}^{2} C k_{x}^{5} k_{z}^{3} s}{R} + 2 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} s}{R} - 2 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} s}{R} + 2 \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} s}{R} - \frac{B_{0}^{2} C k_{x} k_{z}^{7} s}{R} + \frac{B_{0}^{2} C k_{x}^{6} k_{z}^{4}}{R Rm} - 4 \frac{B_{0}^{2} C k_{x}^{5} k_{z}^{5}}{R Rm} + 6 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{6}}{R Rm} - 4 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{7}}{R Rm} + \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{8}}{R Rm} + 2 \frac{\Omega^{3} k_{x}^{2} k_{z}^{4} q^{2}}{Rm} - 4 \frac{\Omega^{3} k_{x}^{2} k_{z}^{4} q}{Rm} - 2 \frac{\Omega^{3} k_{x} k_{z}^{5} q^{2}}{Rm} + 4 \frac{\Omega^{3} k_{x} k_{z}^{5} q}{Rm} - 2 \Omega^{2} k_{x}^{2} k_{z}^{2} q s^{2} + 4 \Omega^{2} k_{x}^{2} k_{z}^{2} s^{2} - 2 \Omega^{2} k_{z}^{4} q s^{2} + 4 \Omega^{2} k_{z}^{4} s^{2} + 2 \frac{\Omega^{2} k_{x}^{3} k_{z}^{3} q s}{Rm} - 4 \frac{\Omega^{2} k_{x}^{3} k_{z}^{3} s}{Rm} - 4 \frac{\Omega^{2} k_{x}^{2} k_{z}^{4} q s}{Rm} + 8 \frac{\Omega^{2} k_{x}^{2} k_{z}^{4} s}{Rm} + 2 \frac{\Omega^{2} k_{x} k_{z}^{5} q s}{Rm} - 4 \frac{\Omega^{2} k_{x} k_{z}^{5} s}{Rm} - \frac{\Omega^{2} k_{x}^{4} k_{z}^{4} q^{2}}{R Rm} + 2 \frac{\Omega^{2} k_{x}^{4} k_{z}^{4} q}{R Rm} + 2 \frac{\Omega^{2} k_{x}^{3} k_{z}^{5} q^{2}}{R Rm} - 4 \frac{\Omega^{2} k_{x}^{3} k_{z}^{5} q}{R Rm} - \frac{\Omega^{2} k_{x}^{2} k_{z}^{6} q^{2}}{R Rm} + 2 \frac{\Omega^{2} k_{x}^{2} k_{z}^{6} q}{R Rm} - \frac{\Omega k_{x}^{4} k_{z}^{2} q s^{2}}{Rm} + \frac{\Omega k_{x}^{3} k_{z}^{3} q s^{2}}{Rm} - \frac{\Omega k_{x}^{2} k_{z}^{4} q s^{2}}{Rm} + \frac{\Omega k_{x} k_{z}^{5} q s^{2}}{Rm} + \frac{\Omega k_{x}^{4} k_{z}^{2} q s^{2}}{R} - 2 \frac{\Omega k_{x}^{4} k_{z}^{2} s^{2}}{R} - \frac{\Omega k_{x}^{3} k_{z}^{3} q s^{2}}{R} + 2 \frac{\Omega k_{x}^{3} k_{z}^{3} s^{2}}{R} + \frac{\Omega k_{x}^{2} k_{z}^{4} q s^{2}}{R} - 2 \frac{\Omega k_{x}^{2} k_{z}^{4} s^{2}}{R} - \frac{\Omega k_{x} k_{z}^{5} q s^{2}}{R} + 2 \frac{\Omega k_{x} k_{z}^{5} s^{2}}{R} + 2 \frac{\Omega k_{x}^{5} k_{z}^{3} s}{R Rm} - 6 \frac{\Omega k_{x}^{4} k_{z}^{4} s}{R Rm} + 6 \frac{\Omega k_{x}^{3} k_{z}^{5} s}{R Rm} - 2 \frac{\Omega k_{x}^{2} k_{z}^{6} s}{R Rm} + k_{x}^{4} s^{4} + 2 k_{x}^{2} k_{z}^{2} s^{4} + k_{z}^{4} s^{4} - \frac{k_{x}^{5} k_{z} s^{3}}{Rm} + 2 \frac{k_{x}^{4} k_{z}^{2} s^{3}}{Rm} - 2 \frac{k_{x}^{3} k_{z}^{3} s^{3}}{Rm} + 2 \frac{k_{x}^{2} k_{z}^{4} s^{3}}{Rm} - \frac{k_{x} k_{z}^{5} s^{3}}{Rm} - \frac{k_{x}^{5} k_{z} s^{3}}{R} + 2 \frac{k_{x}^{4} k_{z}^{2} s^{3}}{R} - 2 \frac{k_{x}^{3} k_{z}^{3} s^{3}}{R} + 2 \frac{k_{x}^{2} k_{z}^{4} s^{3}}{R} - \frac{k_{x} k_{z}^{5} s^{3}}{R} + \frac{k_{x}^{6} k_{z}^{2} s^{2}}{R Rm} - 4 \frac{k_{x}^{5} k_{z}^{3} s^{2}}{R Rm} + 6 \frac{k_{x}^{4} k_{z}^{4} s^{2}}{R Rm} - 4 \frac{k_{x}^{3} k_{z}^{5} s^{2}}{R Rm} + \frac{k_{x}^{2} k_{z}^{6} s^{2}}{R Rm}$ \\

Dispersion relation minus a4: are there any 0-order s terms?\\
$B_{0}^{4} C^{2} k_{x}^{4} k_{z}^{4} + 2 B_{0}^{4} C^{2} k_{x}^{2} k_{z}^{6} + B_{0}^{4} C^{2} k_{z}^{8} + 2 B_{0}^{2} C k_{x}^{4} k_{z}^{2} s^{2} - 2 B_{0}^{2} C k_{x}^{2} k_{z}^{4} \omega^{2} q + 4 B_{0}^{2} C k_{x}^{2} k_{z}^{4} s^{2} - 2 B_{0}^{2} C k_{z}^{6} \omega^{2} q + 2 B_{0}^{2} C k_{z}^{6} s^{2} - \frac{B_{0}^{2} C k_{x}^{5} k_{z}^{3} s}{Rm} - \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} \omega q}{Rm} + 2 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} \omega}{Rm} + 2 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} s}{Rm} + \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} \omega q}{Rm} - 2 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} \omega}{Rm} - 2 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} s}{Rm} - \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} \omega q}{Rm} + 2 \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} \omega}{Rm} + 2 \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} s}{Rm} + \frac{B_{0}^{2} C k_{x} k_{z}^{7} \omega q}{Rm} - 2 \frac{B_{0}^{2} C k_{x} k_{z}^{7} \omega}{Rm} - \frac{B_{0}^{2} C k_{x} k_{z}^{7} s}{Rm} - \frac{B_{0}^{2} C k_{x}^{5} k_{z}^{3} s}{R} + \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} \omega q}{R} + 2 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} s}{R} - \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} \omega q}{R} - 2 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} s}{R} + \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} \omega q}{R} + 2 \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} s}{R} - \frac{B_{0}^{2} C k_{x} k_{z}^{7} \omega q}{R} - \frac{B_{0}^{2} C k_{x} k_{z}^{7} s}{R} + \frac{B_{0}^{2} C k_{x}^{6} k_{z}^{4}}{R Rm} - 4 \frac{B_{0}^{2} C k_{x}^{5} k_{z}^{5}}{R Rm} + 6 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{6}}{R Rm} - 4 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{7}}{R Rm} + \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{8}}{R Rm} + k_{x}^{4} s^{4} - 2 k_{x}^{2} k_{z}^{2} \omega^{2} q s^{2} + 4 k_{x}^{2} k_{z}^{2} \omega^{2} s^{2} + 2 k_{x}^{2} k_{z}^{2} s^{4} - 2 k_{z}^{4} \omega^{2} q s^{2} + 4 k_{z}^{4} \omega^{2} s^{2} + k_{z}^{4} s^{4} - \frac{k_{x}^{5} k_{z} s^{3}}{Rm} - \frac{k_{x}^{4} k_{z}^{2} \omega q s^{2}}{Rm} + 2 \frac{k_{x}^{4} k_{z}^{2} s^{3}}{Rm} + 2 \frac{k_{x}^{3} k_{z}^{3} \omega^{2} q s}{Rm} - 4 \frac{k_{x}^{3} k_{z}^{3} \omega^{2} s}{Rm} + \frac{k_{x}^{3} k_{z}^{3} \omega q s^{2}}{Rm} - 2 \frac{k_{x}^{3} k_{z}^{3} s^{3}}{Rm} + 2 \frac{k_{x}^{2} k_{z}^{4} \omega^{3} q^{2}}{Rm} - 4 \frac{k_{x}^{2} k_{z}^{4} \omega^{3} q}{Rm} - 4 \frac{k_{x}^{2} k_{z}^{4} \omega^{2} q s}{Rm} + 8 \frac{k_{x}^{2} k_{z}^{4} \omega^{2} s}{Rm} - \frac{k_{x}^{2} k_{z}^{4} \omega q s^{2}}{Rm} + 2 \frac{k_{x}^{2} k_{z}^{4} s^{3}}{Rm} - 2 \frac{k_{x} k_{z}^{5} \omega^{3} q^{2}}{Rm} + 4 \frac{k_{x} k_{z}^{5} \omega^{3} q}{Rm} + 2 \frac{k_{x} k_{z}^{5} \omega^{2} q s}{Rm} - 4 \frac{k_{x} k_{z}^{5} \omega^{2} s}{Rm} + \frac{k_{x} k_{z}^{5} \omega q s^{2}}{Rm} - \frac{k_{x} k_{z}^{5} s^{3}}{Rm} - \frac{k_{x}^{5} k_{z} s^{3}}{R} + \frac{k_{x}^{4} k_{z}^{2} \omega q s^{2}}{R} - 2 \frac{k_{x}^{4} k_{z}^{2} \omega s^{2}}{R} + 2 \frac{k_{x}^{4} k_{z}^{2} s^{3}}{R} - \frac{k_{x}^{3} k_{z}^{3} \omega q s^{2}}{R} + 2 \frac{k_{x}^{3} k_{z}^{3} \omega s^{2}}{R} - 2 \frac{k_{x}^{3} k_{z}^{3} s^{3}}{R} + \frac{k_{x}^{2} k_{z}^{4} \omega q s^{2}}{R} - 2 \frac{k_{x}^{2} k_{z}^{4} \omega s^{2}}{R} + 2 \frac{k_{x}^{2} k_{z}^{4} s^{3}}{R} - \frac{k_{x} k_{z}^{5} \omega q s^{2}}{R} + 2 \frac{k_{x} k_{z}^{5} \omega s^{2}}{R} - \frac{k_{x} k_{z}^{5} s^{3}}{R} + \frac{k_{x}^{6} k_{z}^{2} s^{2}}{R Rm} + 2 \frac{k_{x}^{5} k_{z}^{3} \omega s}{R Rm} - 4 \frac{k_{x}^{5} k_{z}^{3} s^{2}}{R Rm} - \frac{k_{x}^{4} k_{z}^{4} \omega^{2} q^{2}}{R Rm} + 2 \frac{k_{x}^{4} k_{z}^{4} \omega^{2} q}{R Rm} - 6 \frac{k_{x}^{4} k_{z}^{4} \omega s}{R Rm} + 6 \frac{k_{x}^{4} k_{z}^{4} s^{2}}{R Rm} + 2 \frac{k_{x}^{3} k_{z}^{5} \omega^{2} q^{2}}{R Rm} - 4 \frac{k_{x}^{3} k_{z}^{5} \omega^{2} q}{R Rm} + 6 \frac{k_{x}^{3} k_{z}^{5} \omega s}{R Rm} - 4 \frac{k_{x}^{3} k_{z}^{5} s^{2}}{R Rm} - \frac{k_{x}^{2} k_{z}^{6} \omega^{2} q^{2}}{R Rm} + 2 \frac{k_{x}^{2} k_{z}^{6} \omega^{2} q}{R Rm} - 2 \frac{k_{x}^{2} k_{z}^{6} \omega s}{R Rm} + \frac{k_{x}^{2} k_{z}^{6} s^{2}}{R Rm} - \frac{- 2 C^{4} Rm^{4} k_{z}^{4} q + C^{3} Rm^{2} k_{z}^{2} \left(k_{x}^{2} + k_{z}^{2}\right)^{2} \left(- 2 q + 4\right) + C \left(k_{x}^{2} + k_{z}^{2}\right) \left(C^{2} Rm^{2} k_{z}^{2} + \frac{C Rm \left(k_{x}^{2} + k_{z}^{2}\right)^{2}}{R}\right)^{2}}{C^{3} Rm^{4}}$

\end{document}

### 2014-06-06 python/multirun_paramspace.py

@@ -1,0 +1,39 @@
+import os
+import subprocess
+import numpy as np
+import pickle
+
+#Qsearch = np.arange(0.710, 0.760, 0.002)
+#Rmsearch = np.arange(4.88, 4.95, 0.01)
+
+Qsearch = np.arange(0.1, 1.5, 0.1)
+Rmsearch = np.arange(4.6, 5.4, 0.1)
+
+#Qsearch = [0.746, 0.748, 0.750, 0.752, 0.754]
+#Rmsearch = [0.48, 0.49, 0.50]
+
+run_script = "multirun_linear_MRI.py"
+
+processes = {}
+for Rm in Rmsearch:
+    for Q in Qsearch: 
+    
+        print("Starting process")
+        proc_args = ["python3",os.path.join(os.getcwd(),run_script), str(Rm), str(Q)]
+        processes[Rm, Q] = subprocess.Popen(proc_args, stdout=subprocess.PIPE)
+        processes[Rm, Q].wait()
+        print("Ending process")
+
+results = {}
+for k,proc in processes.items():
+    res = proc.communicate()
+    if res is not None:
+        str = res[0].decode("utf-8")
+        results[k] = complex(str)
+
+print(results)
+
+pickle.dump(results, open("results_cowling_coarse.p", "wb"))
+
+#np.save("results_Qrange"+str(Qsearch[0])+"_to_"+str(Qsearch[-1])+"Rrange"+str(Rmsearch[0])+"_to_"+str(Rmsearch[-1])+".npy", results)
+#np.save("results_test.npy", results)

### 2014-06-06 python/runsplottest.py

@@ -1,0 +1,34 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pylab
+import pickle
+
+testdata = pickle.load(open("results_cowling_coarse.p", "rb"))
+
+Rms = np.zeros(len(testdata))
+Qs = np.zeros(len(testdata))
+eval = np.zeros(len(testdata), np.complex128)
+
+for i in range(len(testdata)):
+    jj = testdata.popitem()
+    RmQ = jj[0]
+    Rms[i] = RmQ[0]
+    Qs[i] = RmQ[1]
+    eval[i] = jj[1]
+    
+eval = 1j*eval
+
+fig = plt.figure()
+ax1 = fig.add_subplot(121)
+cb = ax1.scatter(Qs, Rms, c=np.real(eval), marker="s", s=40, vmin=-1E-7, vmax=1E-7, cmap="bone")
+fig.colorbar(cb)
+ax1.set_title("Real")
+ax1.set_xlabel("Q")
+ax1.set_ylabel("Rm")
+
+ax2 = fig.add_subplot(122)
+cb2 = ax2.scatter(Qs, Rms, c=np.imag(eval), marker="s", s=40)
+fig.colorbar(cb2)
+ax2.set_title("Imaginary")
+ax2.set_xlabel("Q")
+ax2.set_ylabel("Rm")

### 2014-06-10 notes/mri_notes.tex
python code added
fixed typo
added coarse grid figure

@@
 
-$\partial_t A \, = \, B_0 \partial_z \Psi \, + \, J\left(A, \Psi\right) \, + \, \frac{1}{Rm} \nabla^2 A$ \\
+$\partial_t A \, = \, B_0 \partial_z \Psi \, + \, J\left(A, \Psi\right) \, + \, \frac{1}{\reym} \nabla^2 A$ \\
 
@@
 
-$\mathbf{V} = \mathbf{V}\left(x, z, Z, T)\right)$ where in detail $\mathbf{V} = \left[\Psi(x), u_{1y}(x), A(x), B_{1y}(x)\right]^T \alpha\left(Z, T\right) e^{st + ik_z z}$.
+$\mathbf{V} = \mathbf{V}\left(x, z, Z, T\right)$ where in detail $\mathbf{V} = \left[\Psi(x), u_{1y}(x), A(x), B_{1y}(x)\right]^T \alpha\left(Z, T\right) e^{st + ik_z z}$.
 Periodicity in $z$ allows us to posit the exponential $z$ dependence.
@@
 + 2\epsilon\partial_z\partial_Z\partial_x^2 + 2\epsilon\partial_z^3\partial_Z + 2\epsilon^3\partial_Z^3\partial_z + 4\epsilon^2 \partial_z^2 \partial_Z^2$ \\
+
+\section*{Matrix Formulation}
 
@@
 
+\section*{Solvability Conditions}
+
 Some thoughts on the solvability conditions: \\
@@
 
-$\frac{1}{\reym} \partial_x^2 A^\dagger \, - \, \frac{1}{4\pi} i A \partial_x^2 \Psi^\dagger \, + \, q i Q B^\dagger \, - \, Q^2 \frac{1}{\reym} A^\dagger \, + \, i Q^3 \frac{1}{4\pi} \Psi^\dagger \, = \, 0$ \\
+$\frac{1}{\reym} \partial_x^2 A^\dagger \, - \, \frac{1}{4\pi} i Q \partial_x^2 \Psi^\dagger \, + \, q i Q B^\dagger \, - \, Q^2 \frac{1}{\reym} A^\dagger \, + \, i Q^3 \frac{1}{4\pi} \Psi^\dagger \, = \, 0$ \\
 
@@
 
-The above is coded up in \texttt{adjoint\_homogenous.py}.
+The above is coded up in \texttt{adjoint\_homogenous.py}. \\
+
+To determine what the critical values $\reym = \reym_{crit}$ and $k_z = Q$, we solve the linearized MRI equations for a number of $\reym$ and $k_z$ values to determine criticality. \\
+
+We reproduce Fig. 1b of Umurhan+: \\
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{s_Re_Im_Co_coarse}
+\caption{Real and Imaginary parts of parameter space for linear MRI system.}
+\label{onoff}
+\end{figure*}
+
+By solving the linear MRI system on a coarse parameter space grid. \\
 

### 2014-06-10 python/runsplottest.py

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 import pylab
 import pickle
 
-testdata = pickle.load(open("results_finegrid6.p", "rb"))
+testdata = pickle.load(open("results_test.p", "rb"))
 
 Rms = np.zeros(len(testdata))
 Qs = np.zeros(len(testdata))

### 2014-06-10 notes/mri_notes.tex
another figure

@@
 \includegraphics[width=\textwidth]{s_Re_Im_Co_coarse}
-\caption{Real and Imaginary parts of parameter space for linear MRI system.}
-\label{onoff}
+\caption{Real and Imaginary parts of parameter space for linear MRI system: coarse grid}
 \end{figure*}
 
-By solving the linear MRI system on a coarse parameter space grid. \\
+By solving the linear MRI system on a coarse parameter space grid. We then zoom in to find the critical values of $\reym$ and $Q$:\\
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{s_Re_Im_Co_finegrid3}
+\caption{Real and Imaginary parts of parameter space for linear MRI system: fine grid}
+\end{figure*}
+
+The code that produces the above is in \texttt{multirun\_paramspace.py}, which runs\\ \texttt{multirun\_linear\_MRI.py}. \\
+
+From this investigation, I have been using $Q = 0.75$, $\reym = 4.877$, rather than the $\reym \sim 4.9$ given in Umurhan+.
+The adjoint homogenous solution, however, still does not resemble the one found in Umurhan+.
+Our solution is shown in Figure \ref{AHsol}:
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{adjoint_soln_incorrect}
+\caption{Adjoint homogenous solution.}
+\label{AHsol}
+\end{figure*}
+
+While this solution is symmetric about the origin, as expected, it does not have the same shape as the solution in Umurhan+.
 

### 2014-06-10 python/adjoint_homogenous.py

@@ -18,7 +18,8 @@ Co = 0.08
 
 #Parameter values found by solving the linear MRI to find most unstable mode
 #Rm = 4.8775
-Rm = 4.89
+#Rm = 4.89
+Rm = 4.877
 iRm = 1./Rm
 Q = 0.75
 
@@ -83,29 +84,29 @@ print(e0)
 x = domain.grid(0)
 LEV.set_state(e0[0])
 
-L = LEV.eigenvalue_pencil.L.todense()
-b = np.zeros_like(LEV.eigenvectors[0])
+#L = LEV.eigenvalue_pencil.L.todense()
+#b = np.zeros_like(LEV.eigenvectors[0])
 
-lsolve = np.linalg.solve(L, b)
-LEV.set_state(lsolve)
+#lsolve = np.linalg.solve(L, b)
+#LEV.set_state(lsolve)
 
 fig = plt.figure()
 ax1 = fig.add_subplot(221)
 ax1.plot(x, LEV.state['psi']['g'].imag)
-ax1.plot(x, LEV.state['psi']['g'].real)
-ax1.set_title(r"$\psi$")
+#ax1.plot(x, LEV.state['psi']['g'].real)
+ax1.set_title(r"Im($\psi^\dagger$)")
 ax1 = fig.add_subplot(222)
-ax1.plot(x, LEV.state['u']['g'].imag)
+#ax1.plot(x, LEV.state['u']['g'].imag)
 ax1.plot(x, LEV.state['u']['g'].real)
-ax1.set_title("u")
+ax1.set_title("Re($u^\dagger$)")
 ax1 = fig.add_subplot(223)
-ax1.plot(x, LEV.state['A']['g'].imag)
+#ax1.plot(x, LEV.state['A']['g'].imag)
 ax1.plot(x, LEV.state['A']['g'].real)
-ax1.set_title("A")
+ax1.set_title("Re($A^\dagger$)")
 ax1 = fig.add_subplot(224)
 ax1.plot(x, LEV.state['B']['g'].imag)
-ax1.plot(x, LEV.state['B']['g'].real)
-ax1.set_title("B")
+#ax1.plot(x, LEV.state['B']['g'].real)
+ax1.set_title("Im($B^\dagger$)")
 fig.savefig("ah1.png")
 
 

### 2014-06-13 python/adjoint_homogenous.py
all positive L operators

@@ -40,16 +40,16 @@ iR = 1./R
 #lv1.add_equation("-1j*iRm*dx(Bx) - -1j*Co*1j*Q*u - -1j*Q*Q*iRm*B + dt(B) = 0")
 
 #Proper equations?
-lv1.add_equation("1j*dt(psi) + -1j*Q*A - 1j*Q*(2-q)*u + iR*Q**4*psi - 2*iR*Q**2*psixx + iR*dx(psixxx) = 0")
-lv1.add_equation("1j*dt(u) + -1j*Q*B - 1j*2*Q*psi - iR*Q**2*u + iR*dx(ux) = 0")
-lv1.add_equation("1j*dt(A) + -iRm*Q**2*A + iRm*dx(Ax) + 1j*Q*q*B + 1j*Co*Q**3*psi - 1j*Co*Q*psixx = 0")
-lv1.add_equation("1j*dt(B) + -iRm*Q**2*B + iRm*dx(Bx) - 1j*Co*Q*u = 0")
+#lv1.add_equation("1j*dt(psi) + -1j*Q*A - 1j*Q*(2-q)*u + iR*Q**4*psi - 2*iR*Q**2*psixx + iR*dx(psixxx) = 0")
+#lv1.add_equation("1j*dt(u) + -1j*Q*B - 1j*2*Q*psi - iR*Q**2*u + iR*dx(ux) = 0")
+#lv1.add_equation("1j*dt(A) + -iRm*Q**2*A + iRm*dx(Ax) + 1j*Q*q*B + 1j*Co*Q**3*psi - 1j*Co*Q*psixx = 0")
+#lv1.add_equation("1j*dt(B) + -iRm*Q**2*B + iRm*dx(Bx) - 1j*Co*Q*u = 0")
 
 #Equations without the sign change on the dz terms:
-#lv1.add_equation("1j*dt(psi) + 1j*Q*A + 1j*Q*(2-q)*u + iR*Q**4*psi - 2*iR*Q**2*psixx + iR*dx(psixxx) = 0")
-#lv1.add_equation("1j*dt(u) + 1j*Q*B + 2*1j*Q*psi - iR*Q**2*u + iR*dx(ux) = 0")
-#lv1.add_equation("1j*dt(A) + -iRm*Q**2*A + iRm*dx(Ax) - 1j*Q*q*B - 1j*Co*Q**3*psi + 1j*Co*Q*psixx = 0")
[…]

### 2014-06-13 notes/mri_notes.tex
more notes
marginal solution
added marginal eigenfunctions
two attempts at AH solution -- figures
all positive L operators
red only figures

@@
 
-While this solution is symmetric about the origin, as expected, it does not have the same shape as the solution in Umurhan+.
+While this solution is symmetric about the origin, as expected, it does not have the same shape as the solution in Umurhan+. Note that the normalization is arbitrary. \\
+
+We also use Dedalus to solve the linear MRI at marginality, that is, Umurhan+ Figure 2. Our solution, which agrees with theirs, is shown in Figure \ref{margsol}:
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{marginal_solution}
+\caption{Marginal mode eigenfunctions.}
+\label{margsol}
+\end{figure*}
+
+\pagebreak
+The correct adjoint homogenous solution is still elusive.
+Here is everything I have tried so far.
+In what follows, red is the solution indicated in the title, and black is supposed to be zero everywhere.
+\\
+
+1. $L = L_0 - L_1 dz + L_2 dz^2 - L_3 dz^3 + L_4 dz^4$ \\
+
+Smallest eigenvalue: [-0.00624309 +1.97781664e-16j] \\
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.5]{ah_attempt1}
+\includegraphics[scale=0.5]{ah_attempt1_redonly}
+\end{figure*}
+
+\pagebreak
+2. $L = L_0 + L_1 dz - L_2 dz^2 - L_3 dz^3 + L_4 dz^4$ \\
+
+Smallest eigenvalue: [-0.00624309 +1.97781664e-16j] \\
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.5]{ah_pmmp}
+\includegraphics[scale=0.5]{ah_pmmp_redonly}
+\end{figure*}
+
+\pagebreak
+3. $L = L_0 + L_1 dz + L_2 dz^2 + L_3 dz^3 + L_4 dz^4$ \\
+
+Smallest eigenvalue: [-0.00624309 +1.97781664e-16j] \\
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.5]{ah_allpos}
+\includegraphics[scale=0.5]{ah_allpos_redonly}
+\end{figure*}
+
+\pagebreak
+I also tried to solve the equations without any iQ's -- that is, without the $\partial_z$ terms: \\
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.5]{ah_nodz}
+\end{figure*}
 

### 2014-06-13 python/adjoint_homogenous.py
a little more

@@ -46,10 +46,10 @@ iR = 1./R
 #lv1.add_equation("1j*dt(B) + -iRm*Q**2*B + iRm*dx(Bx) - 1j*Co*Q*u = 0")
 
 #Equations without the sign change on the dz terms:
-lv1.add_equation("1j*dt(psi) + 1j*Q*A + 1j*Q*(2-q)*u + iR*Q**4*psi - 2*iR*Q**2*psixx + iR*dx(psixxx) = 0")
-lv1.add_equation("1j*dt(u) + 1j*Q*B + 2*1j*Q*psi - iR*Q**2*u + iR*dx(ux) = 0")
-lv1.add_equation("1j*dt(A) + -iRm*Q**2*A + iRm*dx(Ax) - 1j*Q*q*B - 1j*Co*Q**3*psi + 1j*Co*Q*psixx = 0")
-lv1.add_equation("1j*dt(B) + -iRm*Q**2*B + iRm*dx(Bx) + 1j*Co*Q*u = 0")
+#lv1.add_equation("1j*dt(psi) + 1j*Q*A + 1j*Q*(2-q)*u + iR*Q**4*psi - 2*iR*Q**2*psixx + iR*dx(psixxx) = 0")
+#lv1.add_equation("1j*dt(u) + 1j*Q*B + 2*1j*Q*psi - iR*Q**2*u + iR*dx(ux) = 0")
+#lv1.add_equation("1j*dt(A) + -iRm*Q**2*A + iRm*dx(Ax) - 1j*Q*q*B - 1j*Co*Q**3*psi + 1j*Co*Q*psixx = 0")
+#lv1.add_equation("1j*dt(B) + -iRm*Q**2*B + iRm*dx(Bx) + 1j*Co*Q*u = 0")
 
 #without iQ's
 #lv1.add_equation("1j*dt(psi) + A + iR*dx(psixxx) + iR*2*psixx + iR*psi + (2-q)*u = 0")
@@ -64,10 +64,10 @@ lv1.add_equation("1j*dt(B) + -iRm*Q**2*B + iRm*dx(Bx) + 1j*Co*Q*u = 0")
 #lv1.add_equation("1j*dt(B) + -iRm*Q**2*B + iRm*dx(Bx) + 1j*Co*Q*u + iRm*dx(Bx) + iRm*B = 0")
 
 #NON-ADJOINT LV=0
[…]

### 2014-06-17 python/adjoint_homogenous.py
updated code

@@ -2,11 +2,16 @@ import numpy as np
 import matplotlib.pyplot as plt
 from dedalus2.public import *
 from dedalus2.pde.solvers import LinearEigenvalue
+from scipy.linalg import eig, norm
 import pylab
+import copy
+
+def dothis():
+
 lv1 = ParsedProblem(['x'],
                       field_names=['psi','u', 'A', 'B', 'psix', 'psixx', 'psixxx', 'ux', 'Ax', 'Bx'],
                       param_names=['Q', 'iR', 'iRm', 'q', 'Co'])
-                      
+                  
 x_basis = Chebyshev(64)#(256)
 domain = Domain([x_basis])#,grid_dtype=np.float64)
 
@@ -26,75 +31,11 @@ Q = 0.75
 R = Rm/Pm
 iR = 1./R
 
-#Define higher order derivatives of x
-#These are the equations....
-# iR*dx(psixxx) - (2-q)*1j*Q*u - 1j*Q*A - Q*Q*iR*2*dx(psix) + Q**4*iR*psi = 0
-# iR*dx(ux) - 2*1j*Q*psi - 1j*Q*B - iR*Q**2*u = 0
-# iRm*dx(Ax) - ifourpi*1j*Q*dx(psix) + q*1j*Q*B - Q*Q*iRm*A + 1j*Q**3*ifourpi*psi = 0
-# iRm*dx(Bx) - ifourpi*1j*Q*u - Q*Q*iRm*B = 0
-
-#multiply by -1j and add dt's:
-#lv1.add_equation("-1j*iR*dx(psixxx) - -1j*(2-q)*1j*Q*u - -1j*1j*Q*A - -1j*Q*Q*iR*2*psixx + -1j*Q**4*iR*psi + dt(psi) = 0")
-#lv1.add_equation("-1j*iR*dx(psixxx) - -1j*(2-q)*1j*Q*u - -1j*1j*Q*A - -1j*Q*Q*iR*2*psixx + -1j*Q**4*iR*psi + dt(psixx) -Q**2*dt(psi) = 0")
-#lv1.add_equation("-1j*iR*dx(ux) - -1j*2*1j*Q*psi - -1j*1j*Q*B - -1j*iR*Q**2*u + dt(u) = 0")
[…]

### 2014-06-20 python/adjoint_homogenous.py
typo

@@ -39,7 +39,7 @@ def setup():
     #Correct equations
     lv1.add_equation("-1j*Q**2*dt(psi) + 1j*dt(psixx) + 1j*Q*A + 1j*(q - 2)*Q*u + iR*Q**4*psi - 2*iR*Q**2*psixx + iR*dx(psixxx) = 0")
     lv1.add_equation("1j*dt(u) + 1j*Q*B + 2*1j*Q*psi - iR*Q**2*u + iR*dx(ux) = 0")
-    lv1.add_equation("1j*dt(A) - iRm*Q**2*A + iRm*dx(Ax) - 1j*Q*q*B - 1j*Co*Q**3*psi + 1j*Co*Q*dx(psix) = 0")
+    lv1.add_equation("1j*dt(A) - iRm*Q**2*A + iRm*dx(Ax) - 1j*Q*q*B - 1j*Co*Q**3*psi + 1j*Co*Q*psixx = 0")
     lv1.add_equation("1j*dt(B) - iRm*Q**2*B + iRm*dx(Bx) + 1j*Co*Q*u = 0")
 
     lv1.add_equation("dx(psi) - psix = 0")
@@ -99,26 +99,46 @@ def plot_all(x, LEV):
 
     #Normalized to resemble Umurhan+
     norm1 = -0.9/np.min(LEV.state['u']['g'].imag) #norm(LEV.eigenvectors)
-
+    
     fig = plt.figure()
     ax1 = fig.add_subplot(221)
     ax1.plot(x, LEV.state['psi']['g'].imag*norm1, color="red")
     ax1.plot(x, LEV.state['psi']['g'].real*norm1, color="black")
+    
+    #psi_cc = LEV.state['psi']['g'] + LEV.state['psi']['g'].conj()
+    #ax1.plot(x, psi_cc.imag, color="red")
+    #ax1.plot(x, psi_cc.real, color="black")
+    
     ax1.set_title(r"Im($\psi^\dagger$)")
 
     ax2 = fig.add_subplot(222)
     ax2.plot(x, LEV.state['u']['g'].imag*norm1, color="black")
     ax2.plot(x, LEV.state['u']['g'].real*norm1, color="red")
+    
+    #u_cc = LEV.state['u']['g'] + LEV.state['u']['g'].conj()
+    #ax2.plot(x, u_cc.imag, color="black")
+    #ax2.plot(x, u_cc.real, color="red")
+    
     ax2.set_title("Re($u^\dagger$)")
 
     ax3 = fig.add_subplot(223)
[…]

### 2014-06-23 notes/mri_notes.tex
a little more
updated code
if __name__ == main added
typo
second smallest eigenvalue

@@
 Magnetic fields are scaled by $B_0$ \\
-Pressure is scaled by $\varepsilon^2 \Omega_0^2 r_0^2$ \\
+Pressure is scaled by $\varepsilon^2 \Omega_0^2 r_0^2 \rho_0$ \\
 
@@
 
-$\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + \, 2 u_x \mathbf{\hat{y}} - 2 u_y \mathbf{\hat{x}} \, = \, 2 x q \mathbf{\hat{x}} \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho} \nabla P \, + \,  \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho_b} \nabla P_b \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{\reye} \nabla^2 \mathbf{u}$ \\
+$\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, + \, 2 u_x \mathbf{\hat{y}} - 2 u_y \mathbf{\hat{x}} \, = \, 2 x q \mathbf{\hat{x}} \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho} \nabla P \, + \,  \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho_b} \nabla P_b \, - \, \frac{\varepsilon^2}{\delta^2} \frac{2}{\beta \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{\reye} \nabla^2 \mathbf{u}$ \\
 
@@
 
-$\partial_t u_x \, + \, \left(\mathbf{u} \cdot \nabla\right)u_x \, - \, 2 u_y \, = \, 2 x q \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho} \partial_x P \, + \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_x P_b \, +\, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
-
-$\partial_t u_y \, + \, \left(\mathbf{u} \cdot \nabla\right)u_y \, + \, 2u_x =  \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho} \partial_y P \, +\, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
-
-$\partial_t u_z \, + \, \left(\mathbf{u} \cdot \nabla\right)u_z \,  =  \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho} \partial_z P \,  + \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_z P_b \, +\, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{\reye}\nabla^2 u_z $ \\
+$\partial_t u_x \, + \, \left(\mathbf{u} \cdot \nabla\right)u_x \, - \, 2 u_y \, = \, 2 x q \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho} \partial_x P \, + \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_x P_b \, +\, \frac{\varepsilon^2}{\delta^2} \frac{2}{\beta \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
+
+$\partial_t u_y \, + \, \left(\mathbf{u} \cdot \nabla\right)u_y \, + \, 2u_x =  \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho} \partial_y P \, +\, \frac{\varepsilon^2}{\delta^2} \frac{2}{\beta \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
+
+$\partial_t u_z \, + \, \left(\mathbf{u} \cdot \nabla\right)u_z \,  =  \, - \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho} \partial_z P \,  + \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\rho_b} \partial_z P_b \, +\, \frac{\varepsilon^2}{\delta^2} \frac{2}{\beta \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B} \, + \, \frac{1}{\reye}\nabla^2 u_z $ \\
 
@@
 
-$\left(-q \Omega_0 x \mathbf{\hat{y}} \, + \, \mathbf{u_1}\right) \cdot \left(-q \Omega_0 x \mathbf{\hat{y}} \, + \, \mathbf{u_1}\right) \, = \, - q \Omega_0 x \mathbf{\hat{y}} \cdot \nabla \mathbf{u_1} + \mathbf{u_1} \cdot \nabla \mathbf{u_1} \, = \, - q \Omega_0 x \partial_y \mathbf{u_1} + \mathbf{u_1} \cdot \nabla \mathbf{u_1}$ \\
+$\left(-q \Omega_0 x \mathbf{\hat{y}} \, + \, \mathbf{u_1}\right) \cdot \nabla \left(-q \Omega_0 x \mathbf{\hat{y}} \, + \, \mathbf{u_1}\right) \, = \, - q \Omega_0 x \mathbf{\hat{y}} \cdot \nabla \mathbf{u_1} + \mathbf{u_1} \cdot \nabla \mathbf{u_1} \, = \, - q \Omega_0 x \partial_y \mathbf{u_1} + \mathbf{u_1} \cdot \nabla \mathbf{u_1}$ \\
 
@@
 
-The Lorentz force term $\frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4 \pi \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B}$ expands as follows: \\
+The Lorentz force term $\frac{\varepsilon^2}{\delta^2} \frac{2}{\beta \rho} \left( \nabla \times \mathbf{B} \right) \times \mathbf{B}$ expands as follows: \\
 
@@
 
+And here is the solution to the *non-adjoint* $LV = 0$ equation:
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.5]{nonadjoint_redonly}
+\end{figure*}
+
+Finally, the solution to the equations:
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.5]{ah_allpos_fixed}
+\end{figure*}
+
 \bibliography{mri.bib}

### 2014-08-12 notes/mri_notes.tex
fixed that pesky 1/4pi typo

@@
 
-The current density is $\mathbf{J} = \frac{1}{4\pi} \nabla \times \mathbf{B}$ and thus the Lorentz force becomes: \\
+The current density is $\mathbf{J} = \frac{2}{\beta} \nabla \times \mathbf{B}$ and thus the Lorentz force becomes: \\
 
@@
 
-$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, + \, 2 q \Omega_0 x \, - \, 2 u_{1y}\,= \, 2 q x \, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_x P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_x P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1x} \,- \, \frac{\nabla B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
+$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, + \, 2 q \Omega_0 x \, - \, 2 u_{1y}\,= \, 2 q x \, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_x P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_x P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{2}{\beta} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1x} \,- \, \frac{\nabla B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
 
 $\partial_t u_{1y} \, - \, q \Omega_0 u_{1x} \, - \, q \Omega_0 x \partial_y u_{1y} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1y} \,+ \, 2u_{1x} \, =  $\\
-$\, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_y P_1}{\rho_1 + \rho_b} \right) \, + \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1y} \,- \, \frac{\nabla B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
+$\, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_y P_1}{\rho_1 + \rho_b} \right) \, + \, \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{2}{\beta} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1y} \,- \, \frac{\nabla B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
 
 $\partial_t u_{1z} \, - \, q \Omega_0 x \partial_y u_{1z} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1z} \, = $\\
-$\, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_z P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_z P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1z} \,- \, \frac{\nabla B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
+$\, - \, \frac{\varepsilon^2}{\delta^2} \left( \frac{\partial_z P_1}{\rho_1 + \rho_b} - \frac{\rho_1\partial_z P_b}{\rho_b^2 + \rho_1\rho_b}\right) \, + \frac{\varepsilon^2}{\delta^2} \frac{1}{\beta} \frac{2}{\beta} \frac{1}{\rho_b + \rho} \left( B_0 \partial_z B_{1z} \,- \, \frac{\nabla B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
 
@@
 
-$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, - \, 2 u_{1y}\,= \, - \, \frac{1}{\rho_b}\partial_x p  \, + \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1x} \,- \, \frac{\nabla B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
+$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1x} \, - \, 2 u_{1y}\,= \, - \, \frac{1}{\rho_b}\partial_x p  \, + \frac{1}{\beta} \frac{2}{\beta} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1x} \,- \, \frac{\nabla B_{1x}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1x} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
 
 $\partial_t u_{1y}\, - \, q \Omega_0 u_{1x}  \, - \, q \Omega_0 x \partial_y u_{1y} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1y} \,+ \, 2u_{1x} \, =  $\\
-$\, - \, \frac{1}{\rho_b}\partial_y p \, + \,  \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1y} \,- \, \frac{\nabla B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
-
-$\partial_t u_{1z} \, - \, q \Omega_0 x \partial_y u_{1z} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1z} \, = \, - \, \frac{1}{\rho_b}\partial_z p  \, + \frac{1}{\beta} \frac{1}{4\pi} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1z} \,- \, \frac{\nabla B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
+$\, - \, \frac{1}{\rho_b}\partial_y p \, + \,  \frac{1}{\beta} \frac{2}{\beta} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1y} \,- \, \frac{\nabla B_{1y}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1y} \right) \, + \, \frac{1}{\reye}\nabla^2 u_y  $ \\
+
+$\partial_t u_{1z} \, - \, q \Omega_0 x \partial_y u_{1z} \,+ \, \left(\mathbf{u_1} \cdot \nabla\right)u_{1z} \, = \, - \, \frac{1}{\rho_b}\partial_z p  \, + \frac{1}{\beta} \frac{2}{\beta} \frac{1}{\rho_b} \left( B_0 \partial_z B_{1z} \,- \, \frac{\nabla B_{1z}^2}{2} \, + \, \mathbf{B_1}\cdot \nabla B_{1z} \right) \, + \, \frac{1}{\reye}\nabla^2 u_x  $ \\
 
@@
 
-$\partial_t \mathbf{u_1} \, - \, q\Omega_0 x \partial_y \mathbf{u_1} \, + \, \left(\mathbf{u_1}\cdot \nabla\right) \mathbf{u_1} \, - \, 2 u_{1y} \mathbf{\hat{x}} \, + \, \left(2 - \Omega_0 q\right)u_{1x}\mathbf{\hat{y}} \, = \, \frac{1}{4\pi} \nabla \Pi \, + \, \frac{1}{4\pi}B_0\partial_z\mathbf{B_1} \, + \, \frac{1}{4\pi}\mathbf{B_1} \cdot \nabla \mathbf{B_1} \,+ \, \frac{1}{\reye}\nabla^2 \mathbf{u_1}$ \\
+$\partial_t \mathbf{u_1} \, - \, q\Omega_0 x \partial_y \mathbf{u_1} \, + \, \left(\mathbf{u_1}\cdot \nabla\right) \mathbf{u_1} \, - \, 2 u_{1y} \mathbf{\hat{x}} \, + \, \left(2 - \Omega_0 q\right)u_{1x}\mathbf{\hat{y}} \, = \, \frac{2}{\beta} \nabla \Pi \, + \, \frac{2}{\beta}B_0\partial_z\mathbf{B_1} \, + \, \frac{2}{\beta}\mathbf{B_1} \cdot \nabla \mathbf{B_1} \,+ \, \frac{1}{\reye}\nabla^2 \mathbf{u_1}$ \\
 
@@
 
-$\partial_t \mathbf{u_1} \, - \, q\Omega_0 x \partial_y \mathbf{u_1} \, + \, \left(\mathbf{u_1}\cdot \nabla\right) \mathbf{u_1} \, - \, 2 u_{1y} \mathbf{\hat{x}} \, + \, \left(2 - \Omega_0 q\right)u_{1x}\mathbf{\hat{y}} \, = \, \frac{1}{4\pi} \nabla \Pi \, + \, \frac{1}{4\pi}B_0\partial_z\mathbf{B_1} \, + \, \frac{1}{4\pi}\mathbf{B_1} \cdot \nabla \mathbf{B_1}\,  + \, \frac{1}{\reye}\nabla^2 \mathbf{u_1}$ \\
+$\partial_t \mathbf{u_1} \, - \, q\Omega_0 x \partial_y \mathbf{u_1} \, + \, \left(\mathbf{u_1}\cdot \nabla\right) \mathbf{u_1} \, - \, 2 u_{1y} \mathbf{\hat{x}} \, + \, \left(2 - \Omega_0 q\right)u_{1x}\mathbf{\hat{y}} \, = \, \frac{2}{\beta} \nabla \Pi \, + \, \frac{2}{\beta}B_0\partial_z\mathbf{B_1} \, + \, \frac{2}{\beta}\mathbf{B_1} \cdot \nabla \mathbf{B_1}\,  + \, \frac{1}{\reye}\nabla^2 \mathbf{u_1}$ \\
 
@@
 
-$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \, + \, \left(\mathbf{u_1} \cdot \nabla\right) u_{1x} \, - \, 2 u_{1y} \, = \, \frac{1}{4\pi} \partial_x \Pi \, + \, \frac{1}{4\pi} B_0 \partial_z B_{1x} \, + \, \frac{1}{4\pi} \mathbf{B_1} \cdot \nabla B_{1x} \, + \, \frac{1}{\reye}\nabla^2 u_{1x} $ \\
+$\partial_t u_{1x} \, - \, q \Omega_0 x \partial_y u_{1x} \, + \, \left(\mathbf{u_1} \cdot \nabla\right) u_{1x} \, - \, 2 u_{1y} \, = \, \frac{2}{\beta} \partial_x \Pi \, + \, \frac{2}{\beta} B_0 \partial_z B_{1x} \, + \, \frac{2}{\beta} \mathbf{B_1} \cdot \nabla B_{1x} \, + \, \frac{1}{\reye}\nabla^2 u_{1x} $ \\
 
@@
 
-$\partial_t\partial_z\Psi \, + \, \partial_z\Psi \partial_x \partial_z \Psi \, - \, \partial_x \Psi \partial_z^2\Psi \, - \, 2 u_{1y} \, = \, \frac{1}{4\pi}\partial_x \Pi \, + \, \frac{1}{4\pi} B_0 \partial_z^2 A \, + \, \frac{1}{4\pi} \partial_z A \partial_x \partial_z A \, - \, \frac{1}{4\pi} \partial_x A \partial_z^2 A \, + \, \frac{1}{\reye} \nabla^2 \partial_z \Psi$ \\
+$\partial_t\partial_z\Psi \, + \, \partial_z\Psi \partial_x \partial_z \Psi \, - \, \partial_x \Psi \partial_z^2\Psi \, - \, 2 u_{1y} \, = \, \frac{2}{\beta}\partial_x \Pi \, + \, \frac{2}{\beta} B_0 \partial_z^2 A \, + \, \frac{2}{\beta} \partial_z A \partial_x \partial_z A \, - \, \frac{2}{\beta} \partial_x A \partial_z^2 A \, + \, \frac{1}{\reye} \nabla^2 \partial_z \Psi$ \\
 
@@
 
-$\partial_t \partial_z^2 \Psi \, + \, \partial_z^2\Psi\partial_x\partial_z\Psi \, + \, \partial_z\Psi \partial_x\partial_z^2 \Psi \, - \, \partial_z\partial_x\Psi\partial_z^2\Psi \, - \, \partial_x \Psi \partial_z^3 \Psi \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_z\partial_x\Pi \, + \, \frac{1}{4\pi} B_0\partial_z^3A \, + \, \frac{1}{4\pi}\partial_z^2 A \partial_x \partial_z A \, + \, \frac{1}{4\pi}\partial_zA\partial_x\partial_z^2 A \, - \, \frac{1}{4\pi}\partial_z\partial_x A \partial_z^2 A \, - \, \frac{1}{4\pi}\partial_xA\partial_z^3A \, + \, \frac{1}{\reye} \nabla^2 \partial_z^2 \Psi $ \\
+$\partial_t \partial_z^2 \Psi \, + \, \partial_z^2\Psi\partial_x\partial_z\Psi \, + \, \partial_z\Psi \partial_x\partial_z^2 \Psi \, - \, \partial_z\partial_x\Psi\partial_z^2\Psi \, - \, \partial_x \Psi \partial_z^3 \Psi \, - \, 2\partial_z u_{1y} \, = \, \frac{2}{\beta}\partial_z\partial_x\Pi \, + \, \frac{2}{\beta} B_0\partial_z^3A \, + \, \frac{2}{\beta}\partial_z^2 A \partial_x \partial_z A \, + \, \frac{2}{\beta}\partial_zA\partial_x\partial_z^2 A \, - \, \frac{2}{\beta}\partial_z\partial_x A \partial_z^2 A \, - \, \frac{2}{\beta}\partial_xA\partial_z^3A \, + \, \frac{1}{\reye} \nabla^2 \partial_z^2 \Psi $ \\
 
@@
 
-$\partial_t\partial_z^2\Psi \, + \, J\left(\Psi, \partial_z^2\Psi\right) \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_x\partial_z\Pi \, + \, \frac{1}{4\pi}B_0 \partial_z^3A \, + \, \frac{1}{4\pi}J\left(A, \partial_z^2A\right) \, + \, \frac{1}{\reye}\nabla^2 \partial_z^2 \Psi$ \\
+$\partial_t\partial_z^2\Psi \, + \, J\left(\Psi, \partial_z^2\Psi\right) \, - \, 2\partial_z u_{1y} \, = \, \frac{2}{\beta}\partial_x\partial_z\Pi \, + \, \frac{2}{\beta}B_0 \partial_z^3A \, + \, \frac{2}{\beta}J\left(A, \partial_z^2A\right) \, + \, \frac{1}{\reye}\nabla^2 \partial_z^2 \Psi$ \\
 
@@
 
-$\partial_t u_{1z} \, - \, q\Omega_0x\partial_y u_{1z} \, + \, \left(\mathbf{u_1} \cdot \nabla \right) u_{1z} \, = \, \frac{1}{4\pi}\partial_z \Pi \, + \, \frac{1}{4\pi} B_0 \partial_z\partial_zB_{1z} \, + \, \frac{1}{4\pi}\mathbf{B_1}\cdot\nabla B_{1z} \, + \, \frac{1}{\reye}\nabla^2 u_{1z} $ \\
+$\partial_t u_{1z} \, - \, q\Omega_0x\partial_y u_{1z} \, + \, \left(\mathbf{u_1} \cdot \nabla \right) u_{1z} \, = \, \frac{2}{\beta}\partial_z \Pi \, + \, \frac{2}{\beta} B_0 \partial_z\partial_zB_{1z} \, + \, \frac{2}{\beta}\mathbf{B_1}\cdot\nabla B_{1z} \, + \, \frac{1}{\reye}\nabla^2 u_{1z} $ \\
 
@@
 
-$\partial_t \partial_x^2 \Psi \, + \, J\left(\Psi, \partial_x^2 \Psi\right) \, = \, -\frac{1}{4\pi}\partial_x\partial_z\Pi \, + \, \frac{1}{4\pi}B_0\partial_z\partial_x^2 A \, + \, \frac{1}{4\pi}J\left(A, \partial_x^2 A \right) \, + \, \frac{1}{\reye} \nabla^2 \partial_x^2 \Psi$ \\
+$\partial_t \partial_x^2 \Psi \, + \, J\left(\Psi, \partial_x^2 \Psi\right) \, = \, -\frac{2}{\beta}\partial_x\partial_z\Pi \, + \, \frac{2}{\beta}B_0\partial_z\partial_x^2 A \, + \, \frac{2}{\beta}J\left(A, \partial_x^2 A \right) \, + \, \frac{1}{\reye} \nabla^2 \partial_x^2 \Psi$ \\
 
@@
 
-$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{4\pi}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
-
-Where the term $\frac{1}{4\pi}\partial_x\partial_z\Pi \left(\mathbf{\hat{x}} \, - \, \mathbf{\hat{z}}\right)$ goes to zero. \\
+$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
+
+Where the term $\frac{2}{\beta}\partial_x\partial_z\Pi \left(\mathbf{\hat{x}} \, - \, \mathbf{\hat{z}}\right)$ goes to zero. \\
 
@@
 
-$\partial_t u_{1y} \, - q\Omega_0 x \partial_y u_{1y} \, + \, \left(\mathbf{u_1} \cdot \nabla \right)u_{1y} \, + \, \left(2 - \Omega_0 q\right) u_{1x} \, = \, \frac{1}{4\pi} \partial_y \Pi \, + \, \frac{1}{4\pi} B_0 \partial_z B_{1y} \, + \, \frac{1}{4\pi}\mathbf{B_1} \cdot \nabla B_{1y} \, + \, \frac{1}{\reye}\nabla^2 u_{1y}$ \\
+$\partial_t u_{1y} \, - q\Omega_0 x \partial_y u_{1y} \, + \, \left(\mathbf{u_1} \cdot \nabla \right)u_{1y} \, + \, \left(2 - \Omega_0 q\right) u_{1x} \, = \, \frac{2}{\beta} \partial_y \Pi \, + \, \frac{2}{\beta} B_0 \partial_z B_{1y} \, + \, \frac{2}{\beta}\mathbf{B_1} \cdot \nabla B_{1y} \, + \, \frac{1}{\reye}\nabla^2 u_{1y}$ \\
 
@@
 
-$\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, + \, \left(2 - \Omega_0 q\right) \partial_z \Psi \, = \, \frac{1}{4\pi}B_0\partial_z B_{1y} \, + \, \frac{1}{4\pi} J\left(A, B_{1y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{1y}$ \\
+$\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, + \, \left(2 - \Omega_0 q\right) \partial_z \Psi \, = \, \frac{2}{\beta}B_0\partial_z B_{1y} \, + \, \frac{2}{\beta} J\left(A, B_{1y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{1y}$ \\
 
@@
 
-$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{4\pi}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
-
-$\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, + \, \left(2 - \Omega_0 q\right) \partial_z \Psi \, = \, \frac{1}{4\pi}B_0\partial_z B_{1y} \, + \, \frac{1}{4\pi} J\left(A, B_{1y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{1y}$ \\
+$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
+
+$\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, + \, \left(2 - \Omega_0 q\right) \partial_z \Psi \, = \, \frac{2}{\beta}B_0\partial_z B_{1y} \, + \, \frac{2}{\beta} J\left(A, B_{1y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{1y}$ \\
 
@@
 
-$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, \frac{1}{4\pi}J\left(A, \nabla^2 A \right) \, =  \, 2 \partial_z u_{1y} \,+ \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
-
-$\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, - \, \frac{1}{4\pi} J\left(A, B_{1y}\right) \, + \, \left(2 - \Omega_0 q\right) \partial_z \Psi \, = \, \frac{1}{4\pi}B_0\partial_z B_{1y} \, + \, \frac{1}{\reye} \nabla^2 u_{1y}$ \\
+$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, =  \, 2 \partial_z u_{1y} \,+ \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
+
+$\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, - \, \frac{2}{\beta} J\left(A, B_{1y}\right) \, + \, \left(2 - \Omega_0 q\right) \partial_z \Psi \, = \, \frac{2}{\beta}B_0\partial_z B_{1y} \, + \, \frac{1}{\reye} \nabla^2 u_{1y}$ \\
 
@@
 
-$\mathbf{N} = \left[ \begin{matrix}  J(\Psi, \nabla^2 \Psi) - \frac{1}{4\pi}J(A, \nabla^2 A)\\
-J(\Psi, u_{1y}) - \frac{1}{4\pi} J(A, B_{1y}) \\ 
+$\mathbf{N} = \left[ \begin{matrix}  J(\Psi, \nabla^2 \Psi) - \frac{2}{\beta}J(A, \nabla^2 A)\\
+J(\Psi, u_{1y}) - \frac{2}{\beta} J(A, B_{1y}) \\ 
 -J(A, \Psi) \\
@@
 \end{matrix}\right] + \left[\begin{matrix}
-0 & 0 & \frac{1}{4\pi}B_0 \partial_z \nabla^2 & 0 \\
-0 & 0 & 0 & \frac{1}{4\pi} B_0 \partial_z \\
+0 & 0 & \frac{2}{\beta}B_0 \partial_z \nabla^2 & 0 \\
+0 & 0 & 0 & \frac{2}{\beta} B_0 \partial_z \\
 B_0 \partial_z & 0 & 0 & 0 \\
@@
 $\left[\begin{matrix}
-0 & 0 & \frac{1}{4\pi}B_0 \partial_z \nabla^2 & 0 \\
-0 & 0 & 0 & \frac{1}{4\pi} B_0 \partial_z \\
+0 & 0 & \frac{2}{\beta}B_0 \partial_z \nabla^2 & 0 \\
+0 & 0 & 0 & \frac{2}{\beta} B_0 \partial_z \\
 B_0 \partial_z & 0 & 0 & 0 \\
 0 & B_0 \partial_z & 0 & 0 \\ \end{matrix}\right] = (1 - \epsilon^2) \left[\begin{matrix}
-0 & 0 & \frac{1}{4\pi}\partial_z(\partial_x^2 + \partial_z^2) & 0 \\
-0 & 0 & 0 & \frac{1}{4\pi}\partial_z \\
+0 & 0 & \frac{2}{\beta}\partial_z(\partial_x^2 + \partial_z^2) & 0 \\
+0 & 0 & 0 & \frac{2}{\beta}\partial_z \\
 \partial_z & 0 & 0 & 0 \\
@@
 $ = (1 - \epsilon^2)\partial_z\left[\begin{matrix}
-0 & 0 & \frac{1}{4\pi}\partial_x^2 & 0 \\
-0 & 0 & 0 & \frac{1}{4\pi} \\
+0 & 0 & \frac{2}{\beta}\partial_x^2 & 0 \\
+0 & 0 & 0 & \frac{2}{\beta} \\
 1 & 0 & 0 & 0 \\
 0 & 1 & 0 & 0 \\ \end{matrix} \right] + (1 - \epsilon^2)\partial_z^3 \left[\begin{matrix}
-0 & 0 & \frac{1}{4\pi} & 0 \\
+0 & 0 & \frac{2}{\beta} & 0 \\
 0 & 0 & 0 & 0 \\
@@
 $\mathcal{L}_1 = \left[\begin{matrix}
-0 & 2 & \frac{1}{4\pi}\partial_x^2 & 0 \\
-(\Omega_0 - 2)q & 0 & 0 & \frac{1}{4\pi} \\
+0 & 2 & \frac{2}{\beta}\partial_x^2 & 0 \\
+(\Omega_0 - 2)q & 0 & 0 & \frac{2}{\beta} \\
 1 & 0 & 0 & 0 \\
@@
 $\mathcal{L}_3 = \left[\begin{matrix}
-0 & 0 & \frac{1}{4\pi} & 0 \\
+0 & 0 & \frac{2}{\beta} & 0 \\
 0 & 0 & 0 & 0 \\
@@
 $\mathcal{X} = \left[\begin{matrix}
-0 & 0 & \frac{1}{4\pi}\partial_x^2 & 0 \\
-0 & 0 & 0 & \frac{1}{4\pi} \\
+0 & 0 & \frac{2}{\beta}\partial_x^2 & 0 \\
+0 & 0 & 0 & \frac{2}{\beta} \\
 1 & 0 & 0 & 0 \\
@@
 
-$N^{(\Psi)} \to \epsilon^2 J(\Psi_1, \nabla^2 \Psi_1) \,-\, \epsilon^2 \frac{1}{4\pi}J(A_1, \nabla^2 A_1) \,+\, \epsilon^3 J(\Psi_1, \nabla^2\Psi_2) \, - \epsilon^3\frac{1}{4\pi}J(A_1, \nabla^2 A_2) \,+\, \epsilon^3 J(\Psi_2, \nabla^2\Psi_1) \, - \, \epsilon^3 \frac{1}{4\pi}J(A_2, \nabla^2 A_1) \,+\, 2\epsilon^3 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2\epsilon^3 \frac{1}{4\pi}J(A_1, \partial_z\partial_Z A_1) \,+\, \epsilon^3\widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \, \epsilon^3 \frac{1}{4\pi}\widetilde{J}(A_1, \nabla^2 A_1) \, + \, \mathcal{O}(\epsilon^4)$ \\
+$N^{(\Psi)} \to \epsilon^2 J(\Psi_1, \nabla^2 \Psi_1) \,-\, \epsilon^2 \frac{2}{\beta}J(A_1, \nabla^2 A_1) \,+\, \epsilon^3 J(\Psi_1, \nabla^2\Psi_2) \, - \epsilon^3\frac{2}{\beta}J(A_1, \nabla^2 A_2) \,+\, \epsilon^3 J(\Psi_2, \nabla^2\Psi_1) \, - \, \epsilon^3 \frac{2}{\beta}J(A_2, \nabla^2 A_1) \,+\, 2\epsilon^3 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2\epsilon^3 \frac{2}{\beta}J(A_1, \partial_z\partial_Z A_1) \,+\, \epsilon^3\widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \, \epsilon^3 \frac{2}{\beta}\widetilde{J}(A_1, \nabla^2 A_1) \, + \, \mathcal{O}(\epsilon^4)$ \\
 
@@
 
-Now we expand $N^{(u_{1y})} = J(\Psi, u_{1y}) - \frac{1}{4\pi} J(A, B_{1y})$: \\
-
-$= J(\Psi, u_{1y}) + \epsilon \widetilde{J}(\Psi, u_{1y}) - \frac{1}{4\pi}J(A, B_{1y}) \, - \, \frac{1}{4\pi}\epsilon \widetilde{J}(A, B_{1y})$ \\
+Now we expand $N^{(u_{1y})} = J(\Psi, u_{1y}) - \frac{2}{\beta} J(A, B_{1y})$: \\
+
+$= J(\Psi, u_{1y}) + \epsilon \widetilde{J}(\Psi, u_{1y}) - \frac{2}{\beta}J(A, B_{1y}) \, - \, \frac{2}{\beta}\epsilon \widetilde{J}(A, B_{1y})$ \\
 
@@
 
-$N^{(u_{1y})} \to \epsilon^2 J(\Psi_1, u_1) \, - \, \frac{1}{4\pi} \epsilon^2 J(A_1, B_1) \, + \, \epsilon^3 J(\Psi_1, u_2) \, + \, \epsilon^3 J(\Psi_2, u_1) \, + \, \epsilon^3 \widetilde{J}(\Psi_1, u_1) \, - \, \frac{1}{4\pi} \epsilon^3 J(A_1, B_2) \, - \, \frac{1}{4\pi} \epsilon^3 J(A_2, B_1) \, - \, \frac{1}{4\pi}\epsilon^3 \widetilde{J}(A_1, B_1) \, + \, \mathcal{O}(\epsilon^4)$ \\
+$N^{(u_{1y})} \to \epsilon^2 J(\Psi_1, u_1) \, - \, \frac{2}{\beta} \epsilon^2 J(A_1, B_1) \, + \, \epsilon^3 J(\Psi_1, u_2) \, + \, \epsilon^3 J(\Psi_2, u_1) \, + \, \epsilon^3 \widetilde{J}(\Psi_1, u_1) \, - \, \frac{2}{\beta} \epsilon^3 J(A_1, B_2) \, - \, \frac{2}{\beta} \epsilon^3 J(A_2, B_1) \, - \, \frac{2}{\beta}\epsilon^3 \widetilde{J}(A_1, B_1) \, + \, \mathcal{O}(\epsilon^4)$ \\
 
@@
 
-$N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \frac{1}{4\pi}J(A_1, \nabla^2 A_1) $ \\
-
-$N_2^{(u)} = J(\Psi_1, u_1) \, - \, \frac{1}{4\pi} J(A_1, B_1) $ \\
+$N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \frac{2}{\beta}J(A_1, \nabla^2 A_1) $ \\
+
+$N_2^{(u)} = J(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_1) $ \\
 
@@
 
-$N_3^{(\Psi)} = J(\Psi_1, \nabla^2\Psi_2) \, - \, \frac{1}{4\pi}J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) \, - \, \frac{1}{4\pi}J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2 \frac{1}{4\pi}J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \frac{1}{4\pi}\widetilde{J}(A_1, \nabla^2 A_1) $ \\
-
-$N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \frac{1}{4\pi} J(A_1, B_2) \, - \, \frac{1}{4\pi} J(A_2, B_1) \, - \, \frac{1}{4\pi} \widetilde{J}(A_1, B_1)$ \\
+$N_3^{(\Psi)} = J(\Psi_1, \nabla^2\Psi_2) \, - \, \frac{2}{\beta}J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) \, - \, \frac{2}{\beta}J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2 \frac{2}{\beta}J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \frac{2}{\beta}\widetilde{J}(A_1, \nabla^2 A_1) $ \\
+
+$N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_2) \, - \, \frac{2}{\beta} J(A_2, B_1) \, - \, \frac{2}{\beta} \widetilde{J}(A_1, B_1)$ \\
 
@@
 2 & 0 & 0 & 1 \\
-\frac{1}{4\pi} \partial_x^2 & 0 & 0 & -q\\
-0 & \frac{1}{4\pi} & 0 & 0 \\
+\frac{2}{\beta} \partial_x^2 & 0 & 0 & -q\\
+0 & \frac{2}{\beta} & 0 & 0 \\
 \end{matrix}\right]$ \\
@@
 0 & 0 & 0 & 0 \\
-\frac{1}{4\pi} & 0 & 0 & 0\\
+\frac{2}{\beta} & 0 & 0 & 0\\
 0 & 0 & 0 & 0 \\
@@
 
-$\frac{1}{\reym} \partial_x^2 A^\dagger \, - \, \frac{1}{4\pi} \partial_x^2 \partial_z \Psi^\dagger \, + \, q \partial_z B^\dagger  \,+ \, \frac{1}{\reym} \partial_z^2 A^\dagger \, - \, \frac{1}{4\pi} \partial_z^3 \Psi^\dagger \, = \, 0$ \\
-
-$\frac{1}{\reym} \partial_x^2 B^\dagger \, - \, \frac{1}{4\pi} \partial_z u^\dagger \, + \, \frac{1}{\reym} \partial_z^2 B^\dagger \, = \, 0$ \\
+$\frac{1}{\reym} \partial_x^2 A^\dagger \, - \, \frac{2}{\beta} \partial_x^2 \partial_z \Psi^\dagger \, + \, q \partial_z B^\dagger  \,+ \, \frac{1}{\reym} \partial_z^2 A^\dagger \, - \, \frac{2}{\beta} \partial_z^3 \Psi^\dagger \, = \, 0$ \\
+
+$\frac{1}{\reym} \partial_x^2 B^\dagger \, - \, \frac{2}{\beta} \partial_z u^\dagger \, + \, \frac{1}{\reym} \partial_z^2 B^\dagger \, = \, 0$ \\
 
@@
 
-$\frac{1}{\reym} \partial_x^2 A^\dagger \, - \, \frac{1}{4\pi} i Q \partial_x^2 \Psi^\dagger \, + \, q i Q B^\dagger \, - \, Q^2 \frac{1}{\reym} A^\dagger \, + \, i Q^3 \frac{1}{4\pi} \Psi^\dagger \, = \, 0$ \\
-
-$\frac{1}{\reym} \partial_x^2 B^\dagger \, - \, \frac{1}{4\pi} i Q u^\dagger \, - \, Q^2 \frac{1}{\reym} B^\dagger \, = \, 0$ \\
+$\frac{1}{\reym} \partial_x^2 A^\dagger \, - \, \frac{2}{\beta} i Q \partial_x^2 \Psi^\dagger \, + \, q i Q B^\dagger \, - \, Q^2 \frac{1}{\reym} A^\dagger \, + \, i Q^3 \frac{2}{\beta} \Psi^\dagger \, = \, 0$ \\
+
+$\frac{1}{\reym} \partial_x^2 B^\dagger \, - \, \frac{2}{\beta} i Q u^\dagger \, - \, Q^2 \frac{1}{\reym} B^\dagger \, = \, 0$ \\
 

### 2014-08-12 python/allorders.py

@@ -2462,32 +2462,32 @@ def plotN2(n2_obj):
     
     # plot N31
     ax1 = fig.add_subplot(2, 2, 1)
-    ax1.plot(n2_obj.x, n2_obj.N31_psi['g'].imag, 'o', mec="#2c7bb6", mfc='none')
-    ax1.plot(n2_obj.x, n2_obj.N31_psi['g'].imag, color="grey")
+    ax1.plot(n2_obj.x, n2_obj.n20_psi['g'].imag, 'o', mec="#2c7bb6", mfc='none')
+    ax1.plot(n2_obj.x, n2_obj.n20_psi['g'].imag, color="grey")
     #ax1.plot(n3_obj.x, n3_obj.N31_psi['g'].real, color="red")
     #ax1.set_title(r"$Im(N_{31}^{(\psi)})$")
     ax1.set_xlim(-1, 1)
     plt.tick_params(labelsize = 20)
 
     ax2 = fig.add_subplot(2, 2, 2)
-    ax2.plot(n2_obj.x, n2_obj.N31_u['g'].real, 'o', mec="#2c7bb6", mfc='none')
-    ax2.plot(n2_obj.x, n2_obj.N31_u['g'].real, color="grey")
+    ax2.plot(n2_obj.x, n2_obj.n20_u['g'].real, 'o', mec="#2c7bb6", mfc='none')
+    ax2.plot(n2_obj.x, n2_obj.n20_u['g'].real, color="grey")
     #ax2.plot(n3_obj.x, n3_obj.N31_u['g'].imag, color="red")
     #ax2.set_title(r"$Re(N_{31}^{(u)})$")
     ax2.set_xlim(-1, 1)
     plt.tick_params(labelsize = 20)
 
     ax3 = fig.add_subplot(2, 2, 3)
-    ax3.plot(n2_obj.x, n2_obj.N31_A['g'].real, 'o', mec="#2c7bb6", mfc='none')
-    ax3.plot(n2_obj.x, n2_obj.N31_A['g'].real, color="grey")
+    ax3.plot(n2_obj.x, n2_obj.n20_A['g'].real, 'o', mec="#2c7bb6", mfc='none')
+    ax3.plot(n2_obj.x, n2_obj.n20_A['g'].real, color="grey")
     #ax3.plot(n3_obj.x, n3_obj.N31_A['g'].imag, color="red")
     #ax3.set_title(r"$Re(N_{31}^{(A)})$")
[…]

### 2014-08-12 paper/mri paper.tex
beginning of paper draft

\documentclass[12pt]{emulateapj}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\hyd}    {{\rm H}}
\newcommand{\vobs}{v_{\rm obs}}
\newcommand{\hi}{H{\sc i}~}
\newcommand{\hii}{H{\sc ii}~}
\newcommand{\hia}{H{\sc i}}
\newcommand{\hiia}{H{\sc ii}}

\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{color}

\makeatletter
\setlength{\@fptop}{0pt}
\setlength{\@fpbot}{0pt}
\setlength{\@fpsep}{0pt}
\makeatother

\newcommand{\citei}[1]{\citeauthor{#1} \citeyear{#1}}
\newcommand{\unit}[1]{\textrm{ #1}}
\newcommand{\kms}{km ${\rm s^{-1}}$}
\newcommand{\kmsa}{km ${\rm s^{-1}}$}
\newcommand{\citeia}[2]{\citeauthor{#1}~(\citeyear{#1};~#2)}

\newcommand\reye{\mathrel{Re}}
\newcommand\reym{\mathrel{Rm}}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\begin{document}

\title{The Weakly Nonlinear Saturation of the Magnetorotational Instability}
\author{Clark, S.E.\altaffilmark{1}}
\author{Oishi, J. S. \altaffilmark{2,}\altaffilmark{3}}
\author{Mac Low, M.-M.\altaffilmark{1,}\altaffilmark{2}}
\altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
\altaffiltext{2}{Department of Astrophysics, American Museum of Natural History, New York, NY}
\altaffiltext{3}{Department of Physics, SUNY Farmingdale}

\begin{abstract}
The MRI saturates.
\end{abstract}

\section{Introduction}

The magnetorotational instability (MRI) arises in a differentially rotating disk with a vertical magnetic field.

\section{Set-Up}

Our domain is designed to be relevant to the design of the Princeton Plasma Physics Laboratory (PPPL) MRI experiment. The experimental apparatus is tall, in one of many attributes designed to mitigate the meridional flows introduced by the endcaps. We use periodic vertical boundary conditions. On the radial boundaries of our domain --  the inner and outer cylinders -- we apply no-slip, perfectly conducting boundary conditions.

\section{Weakly Nonlinear Analysis}

We solve the equations of non-ideal, axisymmetric, incompressible magnetized Taylor Couette flow. We solve the momentum equation:

\beq
\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, = 
\eeq

$\, -\frac{1}{\rho}\nabla P \, - \, \nabla\Phi \, + \, \frac{1}{\rho} \left(\mathbf{J}\times\mathbf{B}\right) \, + \, \nu\nabla^2 \mathbf{u} \, - \, 2\mathbf{\Omega} \times \mathbf{u} \, - \, \mathbf{\Omega} \times \left(\mathbf{\Omega} \times \mathbf{r} \right)$

and the induction equation:

\beq
\partial_t \mathbf{B} = \nabla \times \left(\mathbf{u} \times \mathbf{B}\right) + \eta\nabla^2\mathbf{B}
\eeq

Subject to the magnetic solenoid and incompressibility constraints:

\beq
\nabla \cdot \mathbf{B} = 0
\eeq

\beq
\nabla \cdot \mathbf{u} = 0
\eeq

We perturb the fluid quantities with three-dimensional, axisymmetric perturbations. We nondimensionalize the equations according to [...]
The fluid symbols $\mathbf{u}$, $\mathbf{B}$, etc. will henceforth be used to refer to the perturbed quantities.

We define the flux function $\mathbf{A}$ and streamfunction $\mathbf{\Psi}$. These are scalar fields defined as the curl of the magnetic field and velocity, respectively, and so automatically satisfy our constraints.

$\mathbf{A}$ is thus related to the magnetic field as

\beq
\mathbf{B} \, = \, \left[\begin{matrix}
\partial_zA \\
B_{y} \\
-\partial_xA \\
\end{matrix}\right],\eeq \\

and $\mathbf{\Psi}$ is defined analogously.

The perturbed, nondimensionalized equations which will be the focus of this equation are as follows:

Our final equation set is: \\

$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\

$\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \, \frac{2}{\beta}B_0\partial_z B_{1y} \, + \, \frac{2}{\beta} J\left(A, B_{1y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{1y}$ \\

$\partial_t A \, = \, B_0 \partial_z \Psi \, + \, J\left(A, \Psi\right) \, + \, \frac{1}{\reym} \nabla^2 A$ \\

$\partial_t B_{1y} \, = \, B_0 \partial_z u_{1y} \, + \, J\left(A, u_{1y}\right) \, - \, J\left(\Psi, B_{1y}\right) \, + \, \frac{1}{\reym} \nabla^2 B_{1y}  \, - \, q \Omega_0 \partial_z A$ \\

Note that working in terms of the flux function raises the order of the first momentum equation.

\section{Conclusions}

\end{document}

### 2014-08-12 paper/mri paper.tex
a bit of multiscale discussion

@@
 
+We conduct a formal multiscale analysis.
+Our perturbations are characterized in terms of fast and slow-moving variables, in order to simultaneously track their evolution on fast and slow scales.
+The relative scalings of the fast and slow scales are determined as follows.
+When our four equations are linearized for perturbations of the form $e^{\sigma t + i k_x x + i k_z z}$, we can derive the linear dispersion relation, which is fourth order in $\sigma$.
+The fast and slow variables are then defined such that each fluid quantity appears at the same lowest order in the linear dispersion relation.
+
 \section{Conclusions}

### 2014-08-12 paper/mri paper.tex
a bit more

@@
 When our four equations are linearized for perturbations of the form $e^{\sigma t + i k_x x + i k_z z}$, we can derive the linear dispersion relation, which is fourth order in $\sigma$.
-The fast and slow variables are then defined such that each fluid quantity appears at the same lowest order in the linear dispersion relation.
+The fast and slow variables are then defined such that each of the temporal and spatial eigenvalues appear at the same lowest order in the linear dispersion relation.
+The scalings are:
+
+\beq
+X \equiv \epsilon x,  \, \, Y \equiv \epsilon y, \, \, Z \equiv \epsilon z, \, \, T \equiv \epsilon^2 t
+\eeq
+
+Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic Taylor Couette flow.
 

### 2014-08-13 paper/mri paper.tex
b field tuning in paper

@@
 Our domain is designed to be relevant to the design of the Princeton Plasma Physics Laboratory (PPPL) MRI experiment.
-The experimental apparatus is tall, in one of many attributes designed to mitigate the meridional flows introduced by the endcaps.
+The experimental apparatus is tall, one of many attributes designed to mitigate the meridional flows introduced by the endcaps.
 We use periodic vertical boundary conditions.
@@
 
-Our final equation set is: \\
+Our final equation set is:
 
-$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
+\begin{multline}
+\label{eqset1}
+\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \\
+\, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi
+\end{multline}
 
-$\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \, \frac{2}{\beta}B_0\partial_z B_{1y} \, + \, \frac{2}{\beta} J\left(A, B_{1y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{1y}$ \\
+\begin{multline}
+\label{eqset2}
+\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \\
+\, \frac{2}{\beta}B_0\partial_z B_{1y} \, + \, \frac{2}{\beta} J\left(A, B_{1y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{1y}
+\end{multline}
 
-$\partial_t A \, = \, B_0 \partial_z \Psi \, + \, J\left(A, \Psi\right) \, + \, \frac{1}{\reym} \nabla^2 A$ \\
+\begin{multline}
+\label{eqset3}
+\partial_t A \, = \, B_0 \partial_z \Psi \, + \, J\left(A, \Psi\right) \, + \, \frac{1}{\reym} \nabla^2 A
+\end{multline}
 
-$\partial_t B_{1y} \, = \, B_0 \partial_z u_{1y} \, + \, J\left(A, u_{1y}\right) \, - \, J\left(\Psi, B_{1y}\right) \, + \, \frac{1}{\reym} \nabla^2 B_{1y}  \, - \, q \Omega_0 \partial_z A$ \\
+\begin{multline}
+\label{eqset4}
+\partial_t B_{1y} \, = \, B_0 \partial_z u_{1y} \, + \, J\left(A, u_{1y}\right) \, - \, J\left(\Psi, B_{1y}\right) \, \\
++ \, \frac{1}{\reym} \nabla^2 B_{1y}  \, - \, q \Omega_0 \partial_z A
+\end{multline}
+
+Where $J$ is the Jacobian operator, 
+\beq
+J\left(f, g\right) \equiv \partial_z f\partial_x g - \partial_x f \partial_z g.
+\eeq
 
@@
 
+$B_0$ appears in Equations \ref{eqset1} - \ref{eqset4} because it is the nondimensionalized background field strength.
+That is, $B = B_0 \equiv 1$.
+In order to study our system in the weakly nonlinear regime, we tune the background magnetic field down away from stability (recall that stronger vertical fields stabilize the MRI).
+We do so by substituting $B = B_0\left(1 - \epsilon^2\right)$.
+The degree of departure from the marginal state is measured by the small parameter $\epsilon$.
+An $\mathcal{O}\left(\epsilon^2\right)$ weakening of the background magnetic field destabilizes a band of wave modes with width of $\mathcal{O}\left(\epsilon\right)$, which interact nonlinearly.
+
 \section{Conclusions}

### 2014-08-13 paper/mri paper.tex
matrices

@@
 
-We define the flux function $\mathbf{A}$ and streamfunction $\mathbf{\Psi}$.
-These are scalar fields defined as the curl of the magnetic field and velocity, respectively, and so automatically satisfy our constraints.
+We define the flux function $A$ and streamfunction $\Psi$. These are scalar fields defined as the curl of the magnetic field and velocity, respectively, and so automatically satisfy our constraints.
 
-$\mathbf{A}$ is thus related to the magnetic field as
+$A$ is thus related to the magnetic field as
 
@@
 
+$B_0$ appears in Equations \ref{eqset1} - \ref{eqset4} because it is the nondimensionalized background field strength.
+That is, $B = B_0 \equiv 1$.
+In order to study our system in the weakly nonlinear regime, we tune the background magnetic field down away from stability (recall that stronger vertical fields stabilize the MRI).
+We do so by substituting $B = B_0\left(1 - \epsilon^2\right)$.
+The degree of departure from the marginal state is measured by the small parameter $\epsilon$.
+An $\mathcal{O}\left(\epsilon^2\right)$ weakening of the background magnetic field destabilizes a band of wave modes with width of $\mathcal{O}\left(\epsilon\right)$, which interact nonlinearly.
+
+The destabilizing substitution is made, and Equations \ref{eqset1} - \ref{eqset4} are rewritten such that the fluid variables are contained in a state vector $\mathbf{V} = \left[\Psi, u_y, A, B\right]^\mathrm{T}$.
+This yields the following system of equations.
+
+\beq
+\mathcal{D}\partial_t \mathbf{V} + \mathbf{N} = \mathcal{L} \mathbf{V} - \epsilon^2\partial_z\mathcal{X} \mathbf{V} - \epsilon^2 \partial_z^3 \mathcal{L}_3\mathbf{V}
+\eeq
+
+Where $\mathcal{L} \equiv \mathcal{L}_0 + \mathcal{L}_1\partial_z + \mathcal{L}_2\partial_z^2 + \mathcal{L}_3\partial_z^3 + \mathcal{L}_4\partial_z^4$ \\
+
+And: \\
+
+$\mathcal{D} = \left[\begin{matrix}
+\nabla^2 & 0 & 0 & 0 \\
+0 & 1& 0 & 0 \\
+0 & 0 & 1 & 0\\
+0 & 0 & 0 & 1 \\
+\end{matrix}\right]$ \\
+
+$\mathcal{L}_0 = \left[\begin{matrix}
+\frac{1}{\reye}\partial_x^4 & 0 & 0 & 0 \\
+0 & \frac{1}{\reye}\partial_x^2 & 0 &0 \\
+0 & 0 & \frac{1}{\reym}\partial_x^2 & 0 \\
+0 & 0 & 0 & \frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right] $ \\
+
+\vspace{2mm}
+
+$\mathcal{L}_1 = \left[\begin{matrix}
+0 & 2 & \frac{2}{\beta}\partial_x^2 & 0 \\
+(\Omega_0 - 2)q & 0 & 0 & \frac{2}{\beta} \\
+1 & 0 & 0 & 0 \\
+0 & 1 & -q\Omega_0 & 0 \\ \end{matrix}\right] \\$
+
+\vspace{2mm}
+
+$\mathcal{L}_2 = \left[\begin{matrix}
+2\frac{1}{\reye} \partial_x^2 & 0 & 0 & 0 \\
+0 & \frac{1}{\reye} & 0 & 0 \\
+0 & 0 & \frac{1}{\reym} & 0 \\
+0 & 0 & 0 & \frac{1}{\reym} \\ \end{matrix}\right]$
+
+\vspace{2mm}
+
+$\mathcal{L}_3 = \left[\begin{matrix}
+0 & 0 & \frac{2}{\beta} & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\ \end{matrix} \right] $
+
+\vspace{2mm}
+
+$\mathcal{L}_4 = \left[\begin{matrix}
+\frac{1}{\reye} & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\ \end{matrix}\right] $ \\
+
+\vspace{2mm}
+
+$\mathcal{X} = \left[\begin{matrix}
+0 & 0 & \frac{2}{\beta}\partial_x^2 & 0 \\
+0 & 0 & 0 & \frac{2}{\beta} \\
+1 & 0 & 0 & 0 \\
+0 & 1 & 0 & 0 \\ \end{matrix} \right].$ \\
+
 We conduct a formal multiscale analysis.
@@
 
-$B_0$ appears in Equations \ref{eqset1} - \ref{eqset4} because it is the nondimensionalized background field strength.
-That is, $B = B_0 \equiv 1$.
-In order to study our system in the weakly nonlinear regime, we tune the background magnetic field down away from stability (recall that stronger vertical fields stabilize the MRI).
-We do so by substituting $B = B_0\left(1 - \epsilon^2\right)$.
-The degree of departure from the marginal state is measured by the small parameter $\epsilon$.
-An $\mathcal{O}\left(\epsilon^2\right)$ weakening of the background magnetic field destabilizes a band of wave modes with width of $\mathcal{O}\left(\epsilon\right)$, which interact nonlinearly.
-
 \section{Conclusions}

### 2014-09-08 paper/mri paper.tex
more on paper

@@
 The relative scalings of the fast and slow scales are determined as follows.
-When our four equations are linearized for perturbations of the form $e^{\sigma t + i k_x x + i k_z z}$, we can derive the linear dispersion relation, which is fourth order in $\sigma$.
+When our four equations are linearized for axisymmetric perturbations of the form $e^{\sigma t + i k_x x + i k_z z}$, we can derive the linear dispersion relation, which is fourth order in $\sigma$.
 The fast and slow variables are then defined such that each of the temporal and spatial eigenvalues appear at the same lowest order in the linear dispersion relation.
@@
 Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic Taylor Couette flow.
+Each operator is expanded to reflect these scalings -- for instance, $\partial_z$ becomes $\partial_z + \epsilon\partial_Z$.
+
+The state vector is expanded in orders of $\epsilon$:
+
+\beq
+\mathbf{V} = \epsilon\mathbf{V_1} + \epsilon^2\mathbf{V_2} + \epsilon^3\mathbf{V_3} + ...
+\eeq
 

### 2014-09-08 python/parser_test.py

@@ -17,7 +17,7 @@ problem = ParsedProblem(axis_names=['Z'],
 
 #problem.add_equation("a*dt(alpha) + b*alphaZ - h*dZ(alphaZ) - g*1j*Q**3*alpha = -c*alpha*Absolute(alpha**2)")
 #problem.add_equation("a*dt(alpha) + b*alphaZ - h*dZ(alphaZ) - g*1j*Q**3*alpha = -c*alpha*MagSq(alpha)")
-problem.add_equation("a*dt(alpha) + b*alphaZ - h*dZ(alphaZ) - g*1j*Q**3*alpha = c*alpha*MagSquared(alpha)")
+problem.add_equation("a*dt(alpha) + b*alphaZ - h*dZ(alphaZ) - g*1j*Q**3*alpha = -c*alpha*MagSquared(alpha)")
 problem.add_equation("alphaZ - dZ(alpha) = 0")
 
 #problem.add_equation("-(a/c)*dt(alpha) - (b/c)*alphaZ + (h/c)*dZ(alphaZ) + (g/c)*1j*Q**3*alpha = alpha*Absolute(alpha**2)")

### 2014-09-08 paper/mri paper.tex
more paper and minor notes clarification

@@
 
+The weakly nonlinear regime is where the MRI system is nonlinearly unstable to only the most unstable mode of the linear solution.
+We find the marginal state, where most unstable linear MRI mode neither grows nor decays, for a given set of dimensionless parameters.
+
 $B_0$ appears in Equations \ref{eqset1} - \ref{eqset4} because it is the nondimensionalized background field strength.
@@
 
+After the multiscale expansion our system becomes
+
+\begin{multline}
+\label{multiscale_system}
+\epsilon^2 \mathcal{D}\partial_T\mathbf{V} \, + \, \mathbf{N} \, = \, \mathcal{L}\mathbf{V} \, + \, \epsilon\widetilde{\mathcal{L}}_1\partial_Z\mathbf{V} \, + \, \epsilon^2\widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V} \, \\
+ - \, \epsilon^2\partial_z\mathcal{X}\mathbf{V} \, - \, \epsilon^2\partial_z^3\mathcal{L}_3\mathbf{V} \, + \, \mathcal{O}(\epsilon^3)
+\end{multline}
+
+Where
+
+\beq
+\widetilde{\mathcal{L}}_1 = \mathcal{L}_1 + 2\mathcal{L}_2\partial_z + 3\mathcal{L}_3\partial_z^2 + 4\mathcal{L}_4\partial_z^3
+\eeq
+
+\beq
+\widetilde{\mathcal{L}}_2 = \mathcal{L}_2 + 3\mathcal{L}_3\partial_z + 6\mathcal{L}_4\partial_z^2
+\eeq
+
+% perturbation expansion
 The state vector is expanded in orders of $\epsilon$:
@@
 \beq
+\label{pert_exp}
 \mathbf{V} = \epsilon\mathbf{V_1} + \epsilon^2\mathbf{V_2} + \epsilon^3\mathbf{V_3} + ...
 \eeq
+
+Substituting Equation \ref{pert_exp} into Equation \ref{multiscale_system}
 

### 2014-09-08 paper/mri paper.tex
more equations

@@
 \author{Clark, S.E.\altaffilmark{1}}
-\author{Oishi, J. S. \altaffilmark{2,}\altaffilmark{3}}
+\author{Oishi, J.S. \altaffilmark{2,}\altaffilmark{3}}
 \author{Mac Low, M.-M.\altaffilmark{1,}\altaffilmark{2}}
@@
 
-And: \\
+And:
 
-$\mathcal{D} = \left[\begin{matrix}
+\beq
+\mathcal{D} = \left[\begin{matrix}
 \nabla^2 & 0 & 0 & 0 \\
@@
 0 & 0 & 0 & 1 \\
-\end{matrix}\right]$ \\
+\end{matrix}\right]
+\eeq
 
-$\mathcal{L}_0 = \left[\begin{matrix}
+\beq
+\mathcal{L}_0 = \left[\begin{matrix}
 \frac{1}{\reye}\partial_x^4 & 0 & 0 & 0 \\
@@
 0 & 0 & \frac{1}{\reym}\partial_x^2 & 0 \\
-0 & 0 & 0 & \frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right] $ \\
+0 & 0 & 0 & \frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right]
+\eeq
 
-\vspace{2mm}
-
-$\mathcal{L}_1 = \left[\begin{matrix}
+\beq
+\mathcal{L}_1 = \left[\begin{matrix}
 0 & 2 & \frac{2}{\beta}\partial_x^2 & 0 \\
@@
 1 & 0 & 0 & 0 \\
-0 & 1 & -q\Omega_0 & 0 \\ \end{matrix}\right] \\$
+0 & 1 & -q\Omega_0 & 0 \\ \end{matrix}\right] 
+\eeq
 
-\vspace{2mm}
-
-$\mathcal{L}_2 = \left[\begin{matrix}
+\beq
+\mathcal{L}_2 = \left[\begin{matrix}
 2\frac{1}{\reye} \partial_x^2 & 0 & 0 & 0 \\
@@
 0 & 0 & \frac{1}{\reym} & 0 \\
-0 & 0 & 0 & \frac{1}{\reym} \\ \end{matrix}\right]$
+0 & 0 & 0 & \frac{1}{\reym} \\ \end{matrix}\right]
+\eeq
 
-\vspace{2mm}
-
-$\mathcal{L}_3 = \left[\begin{matrix}
+\beq
+\mathcal{L}_3 = \left[\begin{matrix}
 0 & 0 & \frac{2}{\beta} & 0 \\
@@
 0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\ \end{matrix} \right] $
+0 & 0 & 0 & 0 \\ \end{matrix} \right]
+\eeq
 
-\vspace{2mm}
-
-$\mathcal{L}_4 = \left[\begin{matrix}
+\beq
+\mathcal{L}_4 = \left[\begin{matrix}
 \frac{1}{\reye} & 0 & 0 & 0 \\
@@
 0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\ \end{matrix}\right] $ \\
+0 & 0 & 0 & 0 \\ \end{matrix}\right] 
+\eeq
 
-\vspace{2mm}
-
-$\mathcal{X} = \left[\begin{matrix}
+\beq
+\mathcal{X} = \left[\begin{matrix}
 0 & 0 & \frac{2}{\beta}\partial_x^2 & 0 \\
@@
 1 & 0 & 0 & 0 \\
-0 & 1 & 0 & 0 \\ \end{matrix} \right].$ \\
+0 & 1 & 0 & 0 \\ \end{matrix} \right]. 
+\eeq
 
@@
 
-After the multiscale expansion our system becomes
+Because we investigate the system at marginality, where the growth rate $\sigma = 0$, the $\partial_t$ terms drop out. Thus after the multiscale expansion our system becomes
 
@@
 
-Substituting Equation \ref{pert_exp} into Equation \ref{multiscale_system}
+Substituting Equation \ref{pert_exp} into Equation \ref{multiscale_system} yields the perturbed, expanded equations. We solve these successively in increasing orders of $\epsilon$.
+
+To $\mathcal{O}(\epsilon)$, we have
+
+\beq
+\mathcal{L}\mathbf{V_1} = 0
+\eeq
+
+To $\mathcal{O}(\epsilon^2)$:
+
+\beq
+\mathcal{L}\mathbf{V_2} \, = \, \mathbf{N_2} \, - \, \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}
+\eeq
+
+To $\mathcal{O}(\epsilon^3)$:
+
+\begin{multline}
+\mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, \\
+- \partial_Z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1}
+\end{multline}
+
+The $\mathcal{O}(\epsilon)$ equation is the equation for the linear MRI at marginality.
 

### 2014-09-10 paper/mri paper.tex
ansatz
more paper

@@
 \begin{abstract}
-The MRI saturates.
+The magnetorotational instability (MRI) is a fundamental process of accretion disk physics, but its saturation mechanism remains poorly understood.
+We present an analytic analysis of the nonideal MRI in the weakly nonlinear regime -- that is, when the MRI system is just unstable to its most unstable mode.
 \end{abstract}
@@
 
+The multiple scale dependencies of our solution are encoded into an ansatz for the linear MRI solution at marginality,
+
+\beq
+\label{V1_ansatz}
+\mathbf{V_1} = \alpha(T, Z) \mathbb{V}_{11}(x) e^{i k_c z} + c.c.,
+\eeq
+
+where $\alpha$ is a slowly-varying amplitude equation.
+The $x$ dependence is contained in $\mathbb{V}_{11}$, and must be solved for subject to the radial boundary conditions.
+The periodic vertical boundary conditions allow us to posit the $z$ dependence, where $k_c$ is the value of the vertical wavenumber at marginality.
+
 Because we investigate the system at marginality, where the growth rate $\sigma = 0$, the $\partial_t$ terms drop out. Thus after the multiscale expansion our system becomes
@@
 \beq
+\label{ordere}
 \mathcal{L}\mathbf{V_1} = 0
@@
 \beq
+\label{ordere2}
 \mathcal{L}\mathbf{V_2} \, = \, \mathbf{N_2} \, - \, \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}
@@
 \begin{multline}
+\label{ordere3}
 \mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, \\
@@
 
-The $\mathcal{O}(\epsilon)$ equation is the equation for the linear MRI at marginality.
+The $\mathcal{O}(\epsilon)$ equation is the equation for the linear MRI at marginality. This is the equation that is solved to find the $\mathbb{V}_{11}$ component of Equation \ref{V1_ansatz}.
+
+The partial differential equations that comprise Equations \ref{ordere} to \ref{ordere3} are solved in succession subject to no-slip, perfectly conducting radial boundary conditions, defined as:
+
+\section{Results}
 

### 2014-09-10 paper/mri paper.tex
more and some tweaking

@@
 
-The destabilizing substitution is made, and Equations \ref{eqset1} - \ref{eqset4} are rewritten such that the fluid variables are contained in a state vector $\mathbf{V} = \left[\Psi, u_y, A, B\right]^\mathrm{T}$.
+The destabilizing substitution is made, and Equations \ref{eqset1} - \ref{eqset4} are rewritten such that the fluid variables are contained in a state vector $\mathbf{V} = \left[\Psi, u_y, A, B_y\right]^\mathrm{T}$.
 This yields the following system of equations.
@@
 
+\beq
+\Psi = \partial_x \Psi = u_y = A = \partial_x B_y = 0
+\eeq
+
+The practical advantage of our ansatz construction (Equation \ref{V1_ansatz}) is clear: the separable x-dependence means that the radial boundary conditions are solved in only one dimension.
+Thus our analytical framework is able to side-step many of the resolution issues faced by multidimensional simulations.
+We are able to resolve even significant structure in the boundary layers of our domain, because we need only resolve it in one dimension.
+
 \section{Results}

### 2014-09-10 paper/mri paper.tex
figure added

@@
 We are able to resolve even significant structure in the boundary layers of our domain, because we need only resolve it in one dimension.
+We solve the radial component of each equation using the open source pseudospectral code Dedalus (Burns et al., in prep).
+We compute the radial components on a grid of Chebyshev polynomials, as is appropriate for bounded one-dimensional domains (e.g.
+CITE).
+
+We solve the system up to $\mathcal{O}(\epsilon^3)$ in order to close the solution at $\mathcal{O}(\epsilon^2)$.
 
 \section{Results}
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.3]{firstorder_velocity.eps}
+\caption{First order velocity perturbations.
+Streamlines represent velocity in the vertical-radial plane, with thicker streamlines corresponding to faster speeds.
+The colorbar represents azimuthal velocity.}
+\end{figure*}
 

### 2014-09-11 paper/mri paper.tex
solvability criterion stuff

@@
 
+To find a bounded solution at each order, we eliminate secular terms -- terms which are resonant with solutions to the homogenous equation, and cause the solution to grow without bound.
+The elimination of secular terms requires enforcement of solvability criteria, which arise as a consequence of the Fredholm alternative (e.g.
+Guenther \& Lee 1988).
+The solvability criterion is found by forcing the inner product of the equation with its adjoint homogenous solution to be zero.
+For instance, Equation \ref{ordere2} contains a term $\widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ that is resonant with $e^{i k_c z}$.
+Because the adjoint homogenous equation
+
+\beq
+\mathcal{L}^\dagger \mathbf{V}^\dagger = 0
+\eeq
+
+has a nontrivial solution, the solvability criterion
+
+\beq
+\left< \mathbf{V}^\dagger \cdot \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}\right> = 0 
+\eeq
+
+must be satisfied in order for Equation \ref{ordere2} to have a solution.
+
 \section{Results}
@@
 \caption{First order velocity perturbations.
-Streamlines represent velocity in the vertical-radial plane, with thicker streamlines corresponding to faster speeds.
+Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds.
 The colorbar represents azimuthal velocity.}

### 2014-09-11 paper/mri paper.tex
more solvability criteria

@@
 
-must be satisfied in order for Equation \ref{ordere2} to have a solution.
+must be satisfied in order for Equation \ref{ordere2} to have a solution. In the above,
+
+\beq
+\mathbf{V}^\dagger = \mathbb{V}^\dagger e^{iQz} + c.c.
+\eeq
+
+The elimination of this inner product must also be enforced the terms in Equation \ref{ordere3} that are resonant with $e^{i k_c z}$, yielding
+
+\beq
+\label{ampeq1}
+a \partial_T \alpha \, + \, c \alpha \left|\alpha^2\right| \, + \, \widetilde{c} \alpha \partial_Z \beta \, = \, - b \partial_Z \alpha \, + \, h \partial_Z^2 \alpha \, + \, g i Q^3 \alpha
+\eeq
+
+Where
+
+\begin{multline}
+a \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{D}\mathbb{V}^*_{11}\right> \\
+c \equiv \left< \mathbb{V}^\dagger \cdot \mathbf{N}^*_{31}\right> \\
+\widetilde{c} \equiv \left< \mathbb{V}^\dagger \cdot \widetilde{\mathbf{N}}^*_{31}\right> \\
+b \equiv \left< \mathbb{V}^\dagger \cdot \left(\mathcal{X}\mathbb{V}_{11}\right)^* \right> \\
+h \equiv \left< \mathbb{V}^\dagger \cdot \left(\widetilde{\mathcal{L}}_2\mathbb{V}_{11} \, - \, \widetilde{\mathcal{L}}_1\mathbb{V}_{21} \right)^* \right> \\
+g \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{L}_{3} \mathbb{V}_{11} \right> \\
+\end{multline}
+
+Equation \ref{ampeq1} is an amplitude equation -- a function of the slow variables $Z$ and $T$.
+We solve this initial value problem using Dedalus to obtain the asymptotic behavior of $\alpha$.
+This is the saturation amplitude.
 

### 2014-09-11 paper/mri paper.tex
changed figure
Pm justification

@@
 
+We conduct a full study of the MRI parameter space, focussed on the behavior with changing magnetic Prandtl number ($\mathrm{Pm} \equiv \frac{\mathrm{Rm}}{\mathrm{Re}} \equiv \frac{\nu}{\eta}$).
+In protoplanetary disks, as well as in liquid metal MRI experiments, the magnetic Prandtl number can be very small ($\mathrm{Pm} \sim 10^{-6}$).
+The scale separations caused by these large ratios of diffusivity to viscosity render the small Prandtl number regime inaccessible to simulations.
+This analytic formulation allows us to explore the behavior down to magnetic Prandtl numbers of [....].
+
 \begin{figure*}[h!]
 \centering
-\includegraphics[scale=0.3]{firstorder_velocity.eps}
-\caption{First order velocity perturbations.
+\includegraphics[scale=0.5]{twoorders_velocity.eps}
+\caption{First order (left) and second order (right) velocity perturbations.
 Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds.

### 2014-09-12 paper/mri_paper.tex
bib file and renamed paper

--- 2014-09-11 paper/mri paper.tex
@@
 
-The magnetorotational instability (MRI) arises in a differentially rotating disk with a vertical magnetic field.
+The magnetorotational instability (MRI) is the leading candidate for rapid angular momentum transport in astrophysical disks.
+The instability arises when a differentially rotating disk is threaded by a vertical magnetic field (\citei{CHANDRASEKHAR:1960wh}).
+The presence of the magnetic field destabilizes the disk gas, driving turbulence and angular momentum transport.
+
+The MRI evolves via a period of linear growth, followed by nonlinear saturation. The saturation mechanism is not well understood.
 
@@
 
+%\bibliographystyle{yahapj}
+\bibliography{mri.bib}
+
 \end{document}

### 2014-09-12 paper/mri_paper.tex
fixed annoying bibtex issue... I think

@@
-\documentclass[12pt]{emulateapj}
+\documentclass{emulateapj}
 \newcommand{\beq}{\begin{equation}}
@@
 
-The magnetorotational instability (MRI) is the leading candidate for rapid angular momentum transport in astrophysical disks.
-The instability arises when a differentially rotating disk is threaded by a vertical magnetic field (\citei{CHANDRASEKHAR:1960wh}).
+For matter to accrete from disks onto their central objects, some mechanism for angular momentum transport beyond mere molecular viscosity must be invoked.
+Since its discovery by \citeauthor{Chandrasekhar:1960wh} (\citeyear{Chandrasekhar:1960wh}) and subsequent rediscovery by \citeauthor{Balbus:1991vs} (\citeyear{Balbus:1991vs}), the MRI remains the leading explanation for rapid angular momentum transport in astrophysical disks.
+The instability arises when a differentially rotating disk is threaded by a vertical magnetic field.
 The presence of the magnetic field destabilizes the disk gas, driving turbulence and angular momentum transport.
@@
 
-%\bibliographystyle{yahapj}
-\bibliography{mri.bib}
+%\bibliographystyle{plain}
+%\bibliography{mri_refs2}
+
+\bibliographystyle{plain}
+
+\begin{thebibliography}{53}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{Chandrasekhar(1960)}]{Chandrasekhar:1960wh}
+Chandrasekhar, S. 1960, Proceedings of the National Academy of Sciences of the United States of America, 46, 253
+
+\bibitem[{Balbus(1991)}]{Balbus:1991vs}
+Balbus, S A and Hawley, J F, The Astrophysical Journal, 376, 214
+
+\end{thebibliography}
 

### 2014-09-12 paper/mri_paper.tex
more intro

@@
 
-For matter to accrete from disks onto their central objects, some mechanism for angular momentum transport beyond mere molecular viscosity must be invoked.
+For matter to accrete from a disk onto a central object, angular momentum must be transported radially outward in the disk.
+The transport mechanism is likely turbulent (\citei{Shakura:1973wg}).
 Since its discovery by \citeauthor{Chandrasekhar:1960wh} (\citeyear{Chandrasekhar:1960wh}) and subsequent rediscovery by \citeauthor{Balbus:1991vs} (\citeyear{Balbus:1991vs}), the MRI remains the leading explanation for rapid angular momentum transport in astrophysical disks.
@@
 
-The MRI evolves via a period of linear growth, followed by nonlinear saturation. The saturation mechanism is not well understood.
+The MRI evolves via a period of linear growth, followed by nonlinear saturation.
+The nonlinear evolution of the MRI is a complex process, and an understanding of the mechanism by which the MRI saturates remains elusive.
+Most previous works on MRI saturation have studied the problem in simulation.
+Indeed, only a few works have tackled MRI saturation analytically.
+
+%more mention of other analytic works
+
+This work builds on the weakly nonlinear analysis of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}).
 
@@
 
-\bibitem[{Balbus(1991)}]{Balbus:1991vs}
+\bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}
 Balbus, S A and Hawley, J F, The Astrophysical Journal, 376, 214
+
+\bibitem[{Shakura \& Sunyaev(1973)}]{Shakura:1973wg}
+Shakura, N I and Sunyaev, R A, Astronomy \& Astrophysics, 24, 337
+
+\bibitem[{Umurhan et al.(2007)}]{Umurhan:2007hs}
+Umurhan, O M, Regev, O, Menou, K, Phys. Rev. E, 76, 036310
 

### 2014-09-12 paper/mri_paper.tex
bib stuff

@@
 
-This work builds on the weakly nonlinear analysis of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}).
+This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime.
 
@@
 
-\bibitem[{Umurhan et al.(2007)}]{Umurhan:2007hs}
+\bibitem[{Umurhan et al.(2007b)}]{Umurhan:2007hs}
 Umurhan, O M, Regev, O, Menou, K, Phys. Rev. E, 76, 036310
+
+\bibitem[{Umurhan et al.(2007a)}]{Umurhan:2007dz}
+Umurhan, O M, Regev, O, Menou, K, Phys. Rev. Letters, 98, 034501
 

### 2014-09-15 paper/mri_paper.tex
beginning of appendix

@@
 \usepackage{color}
+%\usepackage{appendix}
+\usepackage[titletoc,title]{appendix}
 
@@
 
-Our domain is designed to be relevant to the design of the Princeton Plasma Physics Laboratory (PPPL) MRI experiment.
+Our domain is designed to be relevant to the design of the Princeton Plasma Physics Laboratory (PPPL) MRI experiment (\citei{Ji:2001kd}).
 The experimental apparatus is tall, one of many attributes designed to mitigate the meridional flows introduced by the endcaps.
-We use periodic vertical boundary conditions.
+We use periodic vertical boundary conditions, so our set-up represents an idealized domain where endcap contamination does not exist.
 On the radial boundaries of our domain --  the inner and outer cylinders -- we apply no-slip, perfectly conducting boundary conditions.
@@
 \beq
-\mathbf{V}^\dagger = \mathbb{V}^\dagger e^{iQz} + c.c.
+\mathbf{V}^\dagger = \mathbb{V}^\dagger(x) e^{iQz} + c.c.
 \eeq
 
+where, again, $\mathbb{V}^\dagger$ contains the radial contribution.
 The elimination of this inner product must also be enforced the terms in Equation \ref{ordere3} that are resonant with $e^{i k_c z}$, yielding
@@
 
+\newpage
+\begin{appendices}
+\begin{widetext}
+\section*{Appendix A.}
+\begin{multline}
+N_{31}^{(A)} = -i Q \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i Q \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
+ + i Q \left(A_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 Q \left(\partial_x A_{11}^* \cdot \Psi_{22}\right) \\
+ + i Q \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i Q \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i Q \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 Q \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
+\end{multline}
+
+\end{widetext}
+\end{appendices}
+
 %\bibliographystyle{plain}
@@
 
+\bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}
+Balbus, S A and Hawley, J F, The Astrophysical Journal, 376, 214
+
 \bibitem[{Chandrasekhar(1960)}]{Chandrasekhar:1960wh}
@@
 
-\bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}
-Balbus, S A and Hawley, J F, The Astrophysical Journal, 376, 214
+\bibitem[{Ji et al.(2001)}]{Ji:2001kd}
+Ji, H, Goodman, J, Kageyama, A, Mon. Not. R. Astron. Soc., 325, 1
 
@@
 
+\bibitem[{Umurhan et al.(2007a)}]{Umurhan:2007dz}
+Umurhan, O M, Regev, O, Menou, K, Phys. Rev. Letters, 98, 034501
+
 \bibitem[{Umurhan et al.(2007b)}]{Umurhan:2007hs}
 Umurhan, O M, Regev, O, Menou, K, Phys. Rev. E, 76, 036310
-
-\bibitem[{Umurhan et al.(2007a)}]{Umurhan:2007dz}
-Umurhan, O M, Regev, O, Menou, K, Phys. Rev. Letters, 98, 034501
 

### 2014-09-16 paper/mri_paper.tex
more equations + formatting
formatting appendix
N31 (u)
appendix a
N_22(psi)
full nonlinear appendix
added figure
more endcaps discussion

@@
 
-\section{Conclusions}
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.5]{twoorders_Bfield.eps}
+\caption{First order (left) and second order (right) velocity perturbations.
+Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds.
+The colorbar represents azimuthal velocity.}
+\end{figure*}
+
+\section{Discussion}
+
+Our domain is designed to be relevant to the design of the Princeton Plasma Physics Laboratory (PPPL) MRI experiment (\citei{Ji:2001kd}).
+A challenge for all MRI experiments is to reduce all extraneous flows which may stabilize the MRI or obscure its detection.
+In particular, care must be taken to mitigate meridional flows introduced by the endcaps of the experimental apparatus.
+The PPPL MRI experiment employs a novel design using concentric, differentially rotating rings as endcaps (\citei{Schartman:2009df}).
+In this investigation, we use periodic vertical boundary conditions, so our set-up represents an idealized domain where endcap contamination does not exist.
+Future work will address specifically our predictions for the PPPL MRI experiment.
 
@@
 \section*{Appendix A.}
-\begin{multline}
-N_{31}^{(A)} = -i Q \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i Q \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
+
+Second order nonlinear terms: \\
+
+\beq
+\begin{split}
+N_{22}^{(\Psi)} = & \, i Q \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11} - Q^2 \partial_x \Psi_{11}\right) - \partial_x \Psi_{11} \cdot \left(i Q \partial_x^2 \Psi_{11} - i Q^3 \Psi_{11}\right) \\
+& + \frac{2}{\beta}\partial_x A_{11} \cdot \left(i Q \partial_x^2 A_{11} - i Q^3 A_{11}\right) - \frac{2}{\beta} i Q A_{11} \cdot \left(\partial_x^3 A_{11} - Q^2 \partial_x A_{11}\right)
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{22}^{(u)} = & \, i Q \Psi_{11} \cdot \partial_x u_{11} - \partial_x \Psi_{11} \cdot i Q u_{11} - \frac{2}{\beta} i Q A_{11} \cdot \partial_x B_{11} + \frac{2}{\beta} \partial_x A_{11} \cdot i Q B_{11}
+\end{split}
+\eeq
+
+\beq
+N_{22}^{(A)} = - i Q A_{11} \cdot \partial_x \Psi_{11} + \partial_x A_{11} \cdot i Q \Psi_{11}
+\eeq
+
+\beq
+N_{22}^{(B)} = i Q \Psi_{11} \cdot \partial_x B_{11} - \partial_x \Psi_{11} \cdot i Q B_{11} - i Q A_{11} \cdot \partial_x u_{11} + \partial_x A_{11} \cdot i Q u_{11}
+\eeq
+
+\beq
+\begin{split}
+N_{20}^{(\Psi)} = & \, i Q \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - Q^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i Q \partial_x^2 \Psi_{11}^* - i Q^3 \Psi_{11}^*\right) \\
+& \frac{2}{\beta} \partial_x A_{11} \cdot \left(i Q \partial_x^2 A_{11}^* - i Q^3 A_{11}^*\right) - \frac{2}{\beta} i Q A_{11} \cdot \left(\partial_x^3 A_{11}^* - Q^2 \partial_x A_{11}^*\right)
+\end{split}
+\eeq
+
+\beq
+N_{20}^{(u)} = i Q \Psi_{11} \cdot \partial_x u_{11}^* - \partial_x \Psi_{11} \cdot i Q u_{11}^* - \frac{2}{\beta} i Q A_{11} \cdot \partial_x B_{11}^* + \frac{2}{\beta} \partial_x A_{11} \cdot i Q B_{11}^*
+\eeq
+
+\beq
+N_{20}^{(A)} = - i Q A_{11} \cdot \partial_x \Psi_{11}^* - \partial_x A_{11} \cdot i Q \Psi_{11}^*
+\eeq
+
+\beq
+N_{20}^{(B)} = i Q \Psi_{11} \cdot \partial_x B_{11}^* - \partial_x \Psi_{11} \cdot i Q B_{11}^* - i Q A_{11} \cdot \partial_x u_{11}^* + \partial_x A_{11} \cdot i Q u_{11}^*
+\eeq \\
+
+Third order nonlinear terms: \\
+
+\beq
+\begin{split}
+N_{31}^{(\Psi)} = &  \, i Q \left(\Psi_{11} \cdot \partial_x^3 \Psi_{20}\right) + i Q \left(\Psi_{11} \cdot \partial_x^3\Psi_{20}^*\right) - i Q \left(\Psi_{11}^* \cdot \partial_x^3 \Psi_{22}\right) - i 2 Q \left(\partial_x \Psi_{11}^* \cdot \partial_x^2 \Psi_{22}\right) \\
+& + i 8 Q^3 \left(\partial_x\Psi_{11}^* \cdot \Psi_{22}\right) + i 4 Q^3 \left(\Psi_{11}^* \cdot \partial_x \Psi_{22}\right) + \frac{2}{\beta}\left[ - i Q \left(A_{11} \cdot \partial^3 A_{20}\right) - i Q \left(A_{11} \cdot \partial_x^3 A_{20}^*\right) \right]\\
+& + \frac{2}{\beta}\left[i Q \left(A_{11}^* \cdot \partial_x^3 A_{22}\right) + i 2 Q \left(\partial_x A_{11}^* \cdot \partial_x^2 A_{22}\right) - i 8 Q^3 \left(\partial_x A_{11}^* \cdot A_{22}\right) - i 4 Q^3 \left(A_{11}^* \cdot \partial_x A_{22}\right)\right] \\
+& + i 2 Q \left(\Psi_{22} \cdot \partial_x^3 \Psi_{11}^*\right) - i 2 Q^3\left(\Psi_{22} \cdot \partial_x \Psi_{11}^* \right) - i Q \left(\partial_x \Psi_{20} \cdot \partial_x^2 \Psi_{11}\right) + i Q \left(\partial_x \Psi_{22} \cdot \partial_x^2 \Psi_{11}^*\right) \\
+& - i Q \left(\partial_x \Psi_{20}^* \cdot \partial_x^2 \Psi_{11}\right) + i Q^3 \left(\partial_x \Psi_{20} \cdot \Psi_{11}\right) + i Q^3 \left(\partial_x \Psi_{20}^* \cdot \Psi_{11}\right) - i Q^3 \left(\partial_x \Psi_{22} \cdot \Psi_{11}^*\right) \\
+& + \frac{2}{\beta} \left[ - i 2 Q \left(A_{22} \cdot \partial_x^3 A_{11}^*\right) + i 2 Q^3 \left(A_{22} \cdot \partial_x A_{11}^*\right) + i Q \left(\partial_x A_{20} \cdot \partial_x^2 A_{11}\right) - i Q \left(\partial_x A_{22} \cdot \partial_x^2 A_{11}^*\right) \right] \\
+& + \frac{2}{\beta} \left[ i Q \left(\partial_x A_{20}^* \cdot \partial_x^2 A_{11}\right) - i Q^3 \left(\partial_x A_{20} \cdot A_{11}\right) - i Q^3 \left(\partial_x A_{20}^* \cdot A_{11}\right) + i Q^3 \left(\partial_x A_{22} \cdot A_{11}^*\right)\right]
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{31}^{(u)} = \, & i Q \left(\Psi_{11} \cdot \partial_x u_{20}\right) + i Q \left(\Psi_{11} \cdot \partial_x u_{20}^*\right) - i Q \left(\Psi_{11}^* \cdot \partial_x u_{22}\right) - i 2 Q \left(\partial_x \Psi_{11}^* \cdot u_{22}\right) \\
+& - i Q \left(u_{11} \cdot \partial_x \Psi_{20} \right) - i Q \left(u_{11} \cdot \partial_x \Psi_{20}^*\right) + i Q \left(u_{11}^* \cdot \partial_x \Psi_{22} \right) + i 2 Q \left(\partial_x u_{11}^* \cdot \Psi_{22}\right) \\
+& +\frac{2}{\beta}\left[- i Q \left(A_{11} \cdot \partial_x B_{20}\right) - i Q \left(A_{11} \cdot \partial_x B_{20}^*\right) + i Q \left(A_{11}^* \cdot \partial_x B_{22}\right) + i 2 Q \left(\partial_x A_{11}^* \cdot B_{22}\right)\right] \\
+& +\frac{2}{\beta}\left[ i Q \left(B_{11} \cdot \partial_x A_{20}\right) + i Q \left(B_{11} \cdot \partial_x A_{20}^*\right) - i Q \left(B_{11}^* \cdot \partial_x A_{20}\right) - i 2 Q \left(\partial_x B_{11}^* \cdot A_{22}\right)\right]
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{31}^{(A)} =\, & -i Q \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i Q \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
  + i Q \left(A_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 Q \left(\partial_x A_{11}^* \cdot \Psi_{22}\right) \\
- + i Q \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i Q \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i Q \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 Q \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
-\end{multline}
+ & + i Q \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i Q \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i Q \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 Q \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{31}^{(B)} = \, & i Q \left(\Psi_{11} \cdot \partial_x B_{20}\right) + i Q \left(\Psi_{11} \cdot \partial_x B_{20}^*\right) - i Q \left(\Psi_{11}^* \cdot \partial_x B_{22}\right) - i 2 Q\left(\partial_x \Psi_{11}^* \cdot B_{22}\right) \\
+& - i Q \left(B_{11} \cdot \partial_x \Psi_{20}\right) - i Q \left(B_{11} \cdot \partial_x \Psi_{20}^*\right) + i Q\left(B_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 Q \left(\partial_x B_{11}^* \cdot \Psi_{22}\right) \\
+& - i Q \left(A_{11} \cdot \partial_x u_{20}\right) - i Q \left(A_{11} \cdot \partial_x u_{20}^*\right) + i Q \left(A_{11}^* \cdot \partial_x u_{22}\right) + i 2 Q \left(\partial_x A_{11}^* \cdot u_{22}\right) \\
+& i Q \left(u_{11} \cdot \partial_x A_{20} \right) + i Q \left(u_{11} \cdot \partial_x A_{20}^*\right) - i Q \left(u_{11}^* \cdot \partial_x A_{22}\right) - i 2 Q \left(\partial_x u_{11}^* \cdot A_{22}\right)
+\end{split}
+\eeq
 
@@
 
+\bibitem[{Schartman, et al.}]{Schartman:2009df}
+Schartman, E, Ji, H, Burin, M J, Rev Sci Instrum, 80, 024501
+
 \bibitem[{Umurhan et al.(2007a)}]{Umurhan:2007dz}

### 2014-09-16 paper/mri_paper.tex
Q -> k_c

--- 2014-09-15 paper/mri_paper.tex
@@
 \beq
-\mathbf{V}^\dagger = \mathbb{V}^\dagger(x) e^{iQz} + c.c.
+\mathbf{V}^\dagger = \mathbb{V}^\dagger(x) e^{i k_c z} + c.c.
 \eeq
@@
 \label{ampeq1}
-a \partial_T \alpha \, + \, c \alpha \left|\alpha^2\right| \, + \, \widetilde{c} \alpha \partial_Z \beta \, = \, - b \partial_Z \alpha \, + \, h \partial_Z^2 \alpha \, + \, g i Q^3 \alpha
+a \partial_T \alpha \, + \, c \alpha \left|\alpha^2\right| \, + \, \widetilde{c} \alpha \partial_Z \beta \, = \, - b \partial_Z \alpha \, + \, h \partial_Z^2 \alpha \, + \, g i k_c^3 \alpha
 \eeq
@@
 
-\section{Conclusions}
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.5]{twoorders_Bfield.eps}
+\caption{First order (left) and second order (right) velocity perturbations.
+Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds.
+The colorbar represents azimuthal velocity.}
+\end{figure*}
+
+\section{Discussion}
+
+Our domain is designed to be relevant to the design of the Princeton Plasma Physics Laboratory (PPPL) MRI experiment (\citei{Ji:2001kd}).
+A challenge for all MRI experiments is to reduce all extraneous flows which may stabilize the MRI or obscure its detection.
+In particular, care must be taken to mitigate meridional flows introduced by the endcaps of the experimental apparatus.
+The PPPL MRI experiment employs a novel design using concentric, differentially rotating rings as endcaps (\citei{Schartman:2009df}).
+In this investigation, we use periodic vertical boundary conditions, so our set-up represents an idealized domain where endcap contamination does not exist.
+Future work will address specifically our predictions for the PPPL MRI experiment.
+
+One advantage of our
 
@@
 \section*{Appendix A.}
-\begin{multline}
-N_{31}^{(A)} = -i Q \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i Q \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
- + i Q \left(A_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 Q \left(\partial_x A_{11}^* \cdot \Psi_{22}\right) \\
- + i Q \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i Q \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i Q \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 Q \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
-\end{multline}
+
+Second order nonlinear terms: \\
+
+\beq
+\begin{split}
+N_{22}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11} - k_c^2 \partial_x \Psi_{11}\right) - \partial_x \Psi_{11} \cdot \left(i k_c \partial_x^2 \Psi_{11} - i k_c^3 \Psi_{11}\right) \\
+& + \frac{2}{\beta}\partial_x A_{11} \cdot \left(i k_c \partial_x^2 A_{11} - i k_c^3 A_{11}\right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11} - k_c^2 \partial_x A_{11}\right)
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{22}^{(u)} = & \, i k_c \Psi_{11} \cdot \partial_x u_{11} - \partial_x \Psi_{11} \cdot i k_c u_{11} - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11} + \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}
+\end{split}
+\eeq
+
+\beq
+N_{22}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11} + \partial_x A_{11} \cdot i k_c \Psi_{11}
+\eeq
+
+\beq
+N_{22}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11} - \partial_x \Psi_{11} \cdot i k_c B_{11} - i k_c A_{11} \cdot \partial_x u_{11} + \partial_x A_{11} \cdot i k_c u_{11}
+\eeq
+
+\beq
+\begin{split}
+N_{20}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - k_c^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i k_c \partial_x^2 \Psi_{11}^* - i k_c^3 \Psi_{11}^*\right) \\
+& \frac{2}{\beta} \partial_x A_{11} \cdot \left(i k_c \partial_x^2 A_{11}^* - i k_c^3 A_{11}^*\right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
+\end{split}
+\eeq
+
+\beq
+N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* - \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11}^* + \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}^*
+\eeq
+
+\beq
+N_{20}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11}^* - \partial_x A_{11} \cdot i k_c \Psi_{11}^*
+\eeq
+
+\beq
+N_{20}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11}^* - \partial_x \Psi_{11} \cdot i k_c B_{11}^* - i k_c A_{11} \cdot \partial_x u_{11}^* + \partial_x A_{11} \cdot i k_c u_{11}^*
+\eeq \\
+
+Third order nonlinear terms: \\
+
+\beq
+\begin{split}
+N_{31}^{(\Psi)} = &  \, i k_c \left(\Psi_{11} \cdot \partial_x^3 \Psi_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x^3\Psi_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x^3 \Psi_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot \partial_x^2 \Psi_{22}\right) \\
+& + i 8 k_c^3 \left(\partial_x\Psi_{11}^* \cdot \Psi_{22}\right) + i 4 k_c^3 \left(\Psi_{11}^* \cdot \partial_x \Psi_{22}\right) + \frac{2}{\beta}\left[ - i k_c \left(A_{11} \cdot \partial^3 A_{20}\right) - i k_c \left(A_{11} \cdot \partial_x^3 A_{20}^*\right) \right]\\
+& + \frac{2}{\beta}\left[i k_c \left(A_{11}^* \cdot \partial_x^3 A_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \partial_x^2 A_{22}\right) - i 8 k_c^3 \left(\partial_x A_{11}^* \cdot A_{22}\right) - i 4 k_c^3 \left(A_{11}^* \cdot \partial_x A_{22}\right)\right] \\
+& + i 2 k_c \left(\Psi_{22} \cdot \partial_x^3 \Psi_{11}^*\right) - i 2 k_c^3\left(\Psi_{22} \cdot \partial_x \Psi_{11}^* \right) - i k_c \left(\partial_x \Psi_{20} \cdot \partial_x^2 \Psi_{11}\right) + i k_c \left(\partial_x \Psi_{22} \cdot \partial_x^2 \Psi_{11}^*\right) \\
+& - i k_c \left(\partial_x \Psi_{20}^* \cdot \partial_x^2 \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20} \cdot \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20}^* \cdot \Psi_{11}\right) - i k_c^3 \left(\partial_x \Psi_{22} \cdot \Psi_{11}^*\right) \\
+& + \frac{2}{\beta} \left[ - i 2 k_c \left(A_{22} \cdot \partial_x^3 A_{11}^*\right) + i 2 k_c^3 \left(A_{22} \cdot \partial_x A_{11}^*\right) + i k_c \left(\partial_x A_{20} \cdot \partial_x^2 A_{11}\right) - i k_c \left(\partial_x A_{22} \cdot \partial_x^2 A_{11}^*\right) \right] \\
+& + \frac{2}{\beta} \left[ i k_c \left(\partial_x A_{20}^* \cdot \partial_x^2 A_{11}\right) - i k_c^3 \left(\partial_x A_{20} \cdot A_{11}\right) - i k_c^3 \left(\partial_x A_{20}^* \cdot A_{11}\right) + i k_c^3 \left(\partial_x A_{22} \cdot A_{11}^*\right)\right]
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{31}^{(u)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x u_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x u_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x u_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot u_{22}\right) \\
+& - i k_c \left(u_{11} \cdot \partial_x \Psi_{20} \right) - i k_c \left(u_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c \left(u_{11}^* \cdot \partial_x \Psi_{22} \right) + i 2 k_c \left(\partial_x u_{11}^* \cdot \Psi_{22}\right) \\
+& +\frac{2}{\beta}\left[- i k_c \left(A_{11} \cdot \partial_x B_{20}\right) - i k_c \left(A_{11} \cdot \partial_x B_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x B_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot B_{22}\right)\right] \\
+& +\frac{2}{\beta}\left[ i k_c \left(B_{11} \cdot \partial_x A_{20}\right) + i k_c \left(B_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(B_{11}^* \cdot \partial_x A_{20}\right) - i 2 k_c \left(\partial_x B_{11}^* \cdot A_{22}\right)\right]
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{31}^{(A)} =\, & -i k_c \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i k_c \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
+ + i k_c \left(A_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \Psi_{22}\right) \\
+ & + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{31}^{(B)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x B_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x B_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x B_{22}\right) - i 2 k_c\left(\partial_x \Psi_{11}^* \cdot B_{22}\right) \\
+& - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}\right) - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c\left(B_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x B_{11}^* \cdot \Psi_{22}\right) \\
+& - i k_c \left(A_{11} \cdot \partial_x u_{20}\right) - i k_c \left(A_{11} \cdot \partial_x u_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x u_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot u_{22}\right) \\
+& i k_c \left(u_{11} \cdot \partial_x A_{20} \right) + i k_c \left(u_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(u_{11}^* \cdot \partial_x A_{22}\right) - i 2 k_c \left(\partial_x u_{11}^* \cdot A_{22}\right)
+\end{split}
+\eeq
 

### 2014-09-16 paper/mri_paper.tex
added setup diagram

@@
 Indeed, only a few works have tackled MRI saturation analytically.
+We will briefly summarize a few of these analytic contributions.
 
@@
 
-Our domain is designed to be relevant to the design of the Princeton Plasma Physics Laboratory (PPPL) MRI experiment (\citei{Ji:2001kd}).
-The experimental apparatus is tall, one of many attributes designed to mitigate the meridional flows introduced by the endcaps.
-We use periodic vertical boundary conditions, so our set-up represents an idealized domain where endcap contamination does not exist.
+We analyze a magnetized Taylor Couette flow, representing a differentially rotating, conducting fluid between two cylinders which are driven at different speeds.
+We initialize an angular velocity profile which decreases outward from the axis of rotation, and a weak vertical magnetic field.
+These are the only two conditions required to excite MRI.
+
+Our domain is designed to be relevant to the design of the Princeton Plasma Physics Laboratory (PPPL) MRI experiment (see \S\ref{discussion} for further discussion).
+We use periodic vertical boundary conditions.
 On the radial boundaries of our domain --  the inner and outer cylinders -- we apply no-slip, perfectly conducting boundary conditions.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[trim=5cm 0cm 0cm 0cm, scale=.35]{setup_diagram.eps}
+\caption{Schematic diagram of our set-up, an axisymmetric magnetized Taylor Couette flow.
+We investigate a 2D slice of the X-Z (radial-vertical) plane.
+Our domain is the bolded black box, of width L.
+The radial dimension is solved with a basis of Chebyshev polynomials, and the vertical dimension is solved on a Fourier basis (see \S\ref{discussion}.) }
+\end{figure}
 
@@
 
-\begin{multline}
-a \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{D}\mathbb{V}^*_{11}\right> \\
-c \equiv \left< \mathbb{V}^\dagger \cdot \mathbf{N}^*_{31}\right> \\
-\widetilde{c} \equiv \left< \mathbb{V}^\dagger \cdot \widetilde{\mathbf{N}}^*_{31}\right> \\
-b \equiv \left< \mathbb{V}^\dagger \cdot \left(\mathcal{X}\mathbb{V}_{11}\right)^* \right> \\
-h \equiv \left< \mathbb{V}^\dagger \cdot \left(\widetilde{\mathcal{L}}_2\mathbb{V}_{11} \, - \, \widetilde{\mathcal{L}}_1\mathbb{V}_{21} \right)^* \right> \\
-g \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{L}_{3} \mathbb{V}_{11} \right> \\
-\end{multline}
+\beq
+\begin{split}
+a & \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{D}\mathbb{V}^*_{11}\right> \\
+c & \equiv \left< \mathbb{V}^\dagger \cdot \mathbf{N}^*_{31}\right> \\
+\widetilde{c} & \equiv \left< \mathbb{V}^\dagger \cdot \widetilde{\mathbf{N}}^*_{31}\right> \\
+b & \equiv \left< \mathbb{V}^\dagger \cdot \left(\mathcal{X}\mathbb{V}_{11}\right)^* \right> \\
+h & \equiv \left< \mathbb{V}^\dagger \cdot \left(\widetilde{\mathcal{L}}_2\mathbb{V}_{11} \, - \, \widetilde{\mathcal{L}}_1\mathbb{V}_{21} \right)^* \right> \\
+g & \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{L}_{3} \mathbb{V}_{11} \right> \\
+\end{split}
+\eeq
 
@@
 
-\section{Discussion}
+\section{Discussion}\label{discussion}
 
@@
 
-One advantage of our
+We compute the solutions to our derived equations pseudospectrally, using Dedalus.
+One advantage to using Dedalus for these computations is that the code can compute the evolution of the fully nonlinear MRI system on the same grid.
+
+% Jeff's sims!!!!!
 

### 2014-09-17 paper/mri_paper.tex
more intro

@@
 Indeed, only a few works have tackled MRI saturation analytically.
-We will briefly summarize a few of these analytic contributions.
+We briefly summarize a few of these analytic contributions.
 
+%examples of simulations papers
 %more mention of other analytic works
@@
 This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime.
+Umurhan et al.
+find for the second-order perturbation envelope a real Ginzburg-Landau Equation (GLE).
+We will show that the system does not, in fact, yield a GLE, and comment on the implications of this in \S\ref{discussion}.
+
+We develop this investigation with the goal of providing a rigorous analytical framework that will be used to make predictions for laboratory MRI experiments.
+In recent years much progress has been made toward the goal of observing MRI in a laboratory setting, although no experiment has yet yielded a confirmed detection.
+Of most immediate interest to this work is the Princeton Plasma Physics Laboratory (PPPL) MRI experiment, which has a design roughly analogous to the set-up investigated here (see \S\ref{setup} for further details).
+The Madison plasma dynamo experiment is a multi-purpose plasma device which may be a vehicle for MRI studies in the near future (\citei{Cooper:2013to}).
+The helical MRI, an instability related to the standard MRI but easier to excite in the laboratory, is the focus of the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE) in Dresden.
+PROMISE has already detected helical MRI in the laboratory (e.g.
+\citei{Stefani:2006iv}), and we discuss possible applications of our framework to the helical MRI in \S\ref{discussion}.
+
+%relation to hTC and R-B
+%explanation of WNA
 
@@
 
-We analyze a magnetized Taylor Couette flow, representing a differentially rotating, conducting fluid between two cylinders which are driven at different speeds.
-We initialize an angular velocity profile which decreases outward from the axis of rotation, and a weak vertical magnetic field.
-These are the only two conditions required to excite MRI.
-
-Our domain is designed to be relevant to the design of the Princeton Plasma Physics Laboratory (PPPL) MRI experiment (see \S\ref{discussion} for further discussion).
-We use periodic vertical boundary conditions.
-On the radial boundaries of our domain --  the inner and outer cylinders -- we apply no-slip, perfectly conducting boundary conditions.
-
-\begin{figure}[h!]
+\begin{figure}[h!]\label{setup}
 \centering
@@
 
-\section{Weakly Nonlinear Analysis}
+We analyze a magnetized Taylor Couette flow, representing a differentially rotating, conducting fluid between two cylinders which are driven at different speeds.
+We initialize an angular velocity profile which decreases outward from the axis of rotation, and a weak vertical magnetic field.
+These are the only two conditions required to excite MRI.
+The differential rotation profile is
+
+\beq
+\Omega\left(r\right) \propto \Omega_0\left(\frac{r}{r_0}\right)^{-q},
+\eeq
+
+where $q$, the shear parameter, is $3/2$ for Keplerian rotation. The initial background magnetic field is
+
+\beq
+\mathbf{B} = B_0 \hat{z},
+\eeq
+
+a constant, vertical field.
+
+Our domain is designed to be relevant to the design of the PPPL MRI experiment (see \S\ref{discussion} for further discussion).
+We use periodic vertical boundary conditions.
+On the radial boundaries of our domain --  the inner and outer cylinders -- we apply no-slip, perfectly conducting boundary conditions.
+The choice of boundary conditions dictates the functional bases on which our system is spectrally solved.
+This is explained further in \S\ref{wna}.
+A diagram of our set-up is shown in Figure \ref{setup}.
+
+\section{Weakly Nonlinear Analysis}\label{wna}
 
@@
 
+% hMRI
+% reduction in background shear not possible in astrophysical disks
+
 \newpage
@@
 
+\bibitem[{Cooper et al.(2014)}]{Cooper:2013to}
+Cooper, C M, Wallace, J, Brookhart, M, Clark, M, Collins, C, Ding, W X, Flanagan, K, Khalzov, I, Li, Y, Milhone, J, Nornberg, M, Nonn, P, Weisberg, D, Whyte, D G, Zweibel, E, Forest, C B, Physics of Plasmas, 21, 013505
+
 \bibitem[{Ji et al.(2001)}]{Ji:2001kd}
 Ji, H, Goodman, J, Kageyama, A, Mon. Not. R. Astron. Soc., 325, 1
+
+\bibitem[{Schartman et al.(2009)}]{Schartman:2009df}
+Schartman, E, Ji, H, Burin, M J, Rev Sci Instrum, 80, 024501
 
@@
 
-\bibitem[{Schartman, et al.}]{Schartman:2009df}
-Schartman, E, Ji, H, Burin, M J, Rev Sci Instrum, 80, 024501
+\bibitem[{Stefani et al.(2006)}]{Stefani:2006iv}
+Stefani, F, Gundrum, T, Gerbeth, G, RŸdiger, G, Schultz, M, Szklarski, J, Hollerbach, R, Phys Rev Lett, 97, 184502
 

### 2014-09-17 paper/mri_paper.tex
more experimental stuff

@@
 \citei{Stefani:2006iv}), and we discuss possible applications of our framework to the helical MRI in \S\ref{discussion}.
+A spherical Couette experiment claimed to have detected MRI in 2004 (\citeauthor{Sisan:2004ig}), but most likely detected unrelated MHD instabilities instead (CITECITE).
 
@@
 
+\bibitem[{Sisan et al.(2004)}]{Sisan:2004ig}
+Sisan, D R, Mujica, N, Tillotson, W A, Huang Y-M, Dorland, W, Hassam, A B, Antonsen, T M, Lathrop, D P, Phys Rev Lett, 93, 11
+
 \bibitem[{Stefani et al.(2006)}]{Stefani:2006iv}

### 2014-09-17 paper/mri_paper.tex
more

--- 2014-09-17 paper/mri_paper.tex
@@
 We develop this investigation with the goal of providing a rigorous analytical framework that will be used to make predictions for laboratory MRI experiments.
-In recent years much progress has been made toward the goal of observing MRI in a laboratory setting, although no experiment has yet yielded a confirmed detection.
+Observing MRI in a laboratory setting would open a new avenue for studying MRI dynamics and saturation.
+In recent years much experimental progress has been made, although no experiment has yet yielded a confirmed detection.
 Of most immediate interest to this work is the Princeton Plasma Physics Laboratory (PPPL) MRI experiment, which has a design roughly analogous to the set-up investigated here (see \S\ref{setup} for further details).
@@
 \citei{Stefani:2006iv}), and we discuss possible applications of our framework to the helical MRI in \S\ref{discussion}.
+A spherical Couette experiment claimed to have detected MRI in 2004 (\citeauthor{Sisan:2004ig}), but most likely detected unrelated MHD instabilities instead (\citei{Hollerbach:2009ig}; \citei{Gissinger:2011td}).
 
@@
 We compute the solutions to our derived equations pseudospectrally, using Dedalus.
-One advantage to using Dedalus for these computations is that the code can compute the evolution of the fully nonlinear MRI system on the same grid.
+One advantage to using Dedalus for these computations is that the code can compute the evolution of the fully nonlinear MRI system on the same Fourier/Chebyshev grid.
+This allows us to directly compare simulation and theory.
 
 % Jeff's sims!!!!!
+
+The present configuration, that of magnetized Taylor Couette flow with a purely axial magnetic field, represents the initial conditions for the standard MRI.
+The helical MRI is a related, but separate, MHD instability.
 
@@
 \bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}
-Balbus, S A and Hawley, J F, The Astrophysical Journal, 376, 214
+Balbus, S A and Hawley, J F, 1991, The Astrophysical Journal, 376, 214
 
@@
 \bibitem[{Cooper et al.(2014)}]{Cooper:2013to}
-Cooper, C M, Wallace, J, Brookhart, M, Clark, M, Collins, C, Ding, W X, Flanagan, K, Khalzov, I, Li, Y, Milhone, J, Nornberg, M, Nonn, P, Weisberg, D, Whyte, D G, Zweibel, E, Forest, C B, Physics of Plasmas, 21, 013505
+Cooper, C M, Wallace, J, Brookhart, M, Clark, M, Collins, C, Ding, W X, Flanagan, K, Khalzov, I, Li, Y, Milhone, J, Nornberg, M, Nonn, P, Weisberg, D, Whyte, D G, Zweibel, E, Forest, C B, Physics of Plasmas, 2013, 21, 013505
+
+\bibitem[{Gissinger et al.(2011)}]{Gissinger:2011td}
+Gissinger, C, Ji, H, Goodman, J, Phys. Rev. E, 2011, 84, 026308
+
+\bibitem[{Hollerback(2009)}]{Hollerbach:2009ig}
+Hollerback, R, Proc. R. Soc. A, 2009, 465, 2107
 
 \bibitem[{Ji et al.(2001)}]{Ji:2001kd}
-Ji, H, Goodman, J, Kageyama, A, Mon. Not. R. Astron. Soc., 325, 1
+Ji, H, Goodman, J, Kageyama, A, Mon. Not. R. Astron. Soc., 2001, 325, 1
 
 \bibitem[{Schartman et al.(2009)}]{Schartman:2009df}
-Schartman, E, Ji, H, Burin, M J, Rev Sci Instrum, 80, 024501
+Schartman, E, Ji, H, Burin, M J, Rev Sci Instrum, 2009, 80, 024501
 
 \bibitem[{Shakura \& Sunyaev(1973)}]{Shakura:1973wg}
-Shakura, N I and Sunyaev, R A, Astronomy \& Astrophysics, 24, 337
+Shakura, N I and Sunyaev, R A, Astronomy \& Astrophysics, 1973, 24, 337
+
+\bibitem[{Sisan et al.(2004)}]{Sisan:2004ig}
+Sisan, D R, Mujica, N, Tillotson, W A, Huang Y-M, Dorland, W, Hassam, A B, Antonsen, T M, Lathrop, D P, Phys Rev Lett, 2004, 93, 11
 
 \bibitem[{Stefani et al.(2006)}]{Stefani:2006iv}
-Stefani, F, Gundrum, T, Gerbeth, G, RŸdiger, G, Schultz, M, Szklarski, J, Hollerbach, R, Phys Rev Lett, 97, 184502
+Stefani, F, Gundrum, T, Gerbeth, G, RŸdiger, G, Schultz, M, Szklarski, J, Hollerbach, R, Phys Rev Lett, 2006, 97, 184502
 
 \bibitem[{Umurhan et al.(2007a)}]{Umurhan:2007dz}
-Umurhan, O M, Regev, O, Menou, K, Phys. Rev. Letters, 98, 034501
+Umurhan, O M, Regev, O, Menou, K, 2007, Phys. Rev. Letters, 98, 034501
 
 \bibitem[{Umurhan et al.(2007b)}]{Umurhan:2007hs}
-Umurhan, O M, Regev, O, Menou, K, Phys. Rev. E, 76, 036310
+Umurhan, O M, Regev, O, Menou, K, 2007, Phys. Rev. E, 76, 036310
 

### 2014-09-17 paper/mri_paper.tex
lots more

@@
 
+This work analyzes Taylor Couette flow in the thin-gap limit, where the channel width is radially narrow compared to its distance from the center of rotation.
+This simplifies the math by allowing curvature terms to drop out at lowest order.
+The relaxation of the thin-gap approximation is an obvious next step for this work,
+
 The present configuration, that of magnetized Taylor Couette flow with a purely axial magnetic field, represents the initial conditions for the standard MRI.
-The helical MRI is a related, but separate, MHD instability.
+The helical MRI is a related, but separate, MHD instability, and was first discovered by Hollerbach and R\"udiger (CITE).
+It arises in the presence of combined axial and azimuthal fields.
+To seed helical MRI in our current set-up, we need only add an azimuthal component to our initial background magnetic field.
+This will be an avenue of future study, and will yield an analytic theory relevant to the above-mentioned PROMISE experiment.
 
+Our results suggest that the primary driver of saturation is reduction of the background shear.
+We note that while shear modification is relevant in the experimental regime, the gravitational pull of the central body will tend to enforce a Keplerian rotation profile in any astrophysical disk.
+Thus a modification of the rotation profile is of little relevance as a saturation mechanism to the astrophysical regime.
+Nevertheless, understanding laboratory MRI experiments is crucial for allowing experimental results to inform our astrophysical understanding of the MRI.
+%^this sentence sucks
+
+% curvature terms
 % hMRI

### 2014-09-18 paper/mri_paper.tex
more intro

@@
 A spherical Couette experiment claimed to have detected MRI in 2004 (\citeauthor{Sisan:2004ig}), but most likely detected unrelated MHD instabilities instead (\citei{Hollerbach:2009ig}; \citei{Gissinger:2011td}).
+
+The focus of this work is the weakly nonlinear regime of the MRI.
+This is where the initial linear growth phase of the instability begins to give way to nonlinear saturation.
+Weakly nonlinear analysis is the perturbation analysis of a system that is just unstable to its most unstable mode.
+The system in its marginal state, when the most unstable mode neither grows nor decays, is tuned slightly into instability.
+This excites a band of wavemodes around the most unstable mode, which interact nonlinearly.
 
@@
 The weakly nonlinear regime is where the MRI system is nonlinearly unstable to only the most unstable mode of the linear solution.
-We find the marginal state, where most unstable linear MRI mode neither grows nor decays, for a given set of dimensionless parameters.
+We find the marginal state, where the most unstable linear MRI mode neither grows nor decays, for a given set of dimensionless parameters.
 
@@
 This simplifies the math by allowing curvature terms to drop out at lowest order.
-The relaxation of the thin-gap approximation is an obvious next step for this work,
+The relaxation of the thin-gap approximation is an obvious next step for this work.
+The term expansion in the weakly nonlinear analysis will be more formidable in the wide-gap regime, but could prove informative for comparison to experiments.
 

### 2014-09-18 paper/mri_paper.tex
notes typo
updated figures
citations

@@
 This excites a band of wavemodes around the most unstable mode, which interact nonlinearly.
+In order to track the evolution of the system on both rapid and slow scales, we conduct a formal multiscale analysis.
 
@@
 We compute the radial components on a grid of Chebyshev polynomials, as is appropriate for bounded one-dimensional domains (e.g.
-CITE).
+\citei{Boyd:2001aa}).
 
@@
 \centering
-\includegraphics[scale=0.5]{twoorders_velocity.eps}
+\includegraphics[scale=0.5]{allorders_velocity.eps}
 \caption{First order (left) and second order (right) velocity perturbations.
@@
 \centering
-\includegraphics[scale=0.5]{twoorders_Bfield.eps}
+\includegraphics[scale=0.5]{allorders_Bfield.eps}
 \caption{First order (left) and second order (right) velocity perturbations.
@@
 The present configuration, that of magnetized Taylor Couette flow with a purely axial magnetic field, represents the initial conditions for the standard MRI.
-The helical MRI is a related, but separate, MHD instability, and was first discovered by Hollerbach and R\"udiger (CITE).
+The helical MRI is a related, but separate, MHD instability, and was first discovered by Hollerbach and R\"udiger (\citeyear{Hollerbach:2005tr}).
 It arises in the presence of combined axial and azimuthal fields.
@@
 
+\bibitem[{Boyd(2001)}]{Boyd:2001aa}
+Boyd, J P, 2001, Chebyshev and Fourier Spectral Methods, New York, Dover
+
 \bibitem[{Chandrasekhar(1960)}]{Chandrasekhar:1960wh}
@@
 
-\bibitem[{Hollerback(2009)}]{Hollerbach:2009ig}
-Hollerback, R, Proc. R. Soc. A, 2009, 465, 2107
+\bibitem[{Hollerbach(2009)}]{Hollerbach:2009ig}
+Hollerbach, R, Proc. R. Soc. A, 2009, 465, 2107
+
+\bibitem[{Hollerbach \& R\"udiger(2005)}]{Hollerbach:2005tr}
+Hollerbach, R, RŸdiger, G, Phys. Rev. Lett, 2005, 95, 124501
 

### 2014-09-18 paper/mri_paper.tex
little changes

@@
 This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime.
-Umurhan et al.
-find for the second-order perturbation envelope a real Ginzburg-Landau Equation (GLE).
-We will show that the system does not, in fact, yield a GLE, and comment on the implications of this in \S\ref{discussion}.
 
@@
 
-\begin{figure}[h!]\label{setup}
+\begin{figure}[h!]
 \centering
@@
 Our domain is the bolded black box, of width L.
-The radial dimension is solved with a basis of Chebyshev polynomials, and the vertical dimension is solved on a Fourier basis (see \S\ref{discussion}.) }
+The radial dimension is solved with a basis of Chebyshev polynomials, and the vertical dimension is solved on a Fourier basis (see \S\ref{discussion}.) }\label{setup}
 \end{figure}
@@
 
+We analyze this system in the thin-gap regime, where the channel is radially narrow. Formally, this means
+
+\beq
+thin gap equation,
+\eeq
+
+the width of the channel is small compared to its distance from the center of rotation.
+The imposition of the thin-gap approximation eliminates channel modes, the axisymmetric MRI modes that are exact solutions of the incompressible MRI in the shearing box approximation (\citei{Goodman:1994ul}).
+Implications of this approximation are discussed further in \S\ref{discussion}.
+
 Our domain is designed to be relevant to the design of the PPPL MRI experiment (see \S\ref{discussion} for further discussion).
@@
 
-We solve the equations of non-ideal, axisymmetric, incompressible magnetized Taylor Couette flow. We solve the momentum equation:
+We solve the equations of non-ideal, axisymmetric, incompressible magnetized Taylor Couette flow. We solve the momentum equation,
 
-\beq
+\beq\label{momentum}
 \partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, = 
@@
 
-and the induction equation:
+and the induction equation,
 
-\beq
-\partial_t \mathbf{B} = \nabla \times \left(\mathbf{u} \times \mathbf{B}\right) + \eta\nabla^2\mathbf{B}
+\beq\label{induction}
+\partial_t \mathbf{B} = \nabla \times \left(\mathbf{u} \times \mathbf{B}\right) + \eta\nabla^2\mathbf{B}.
 \eeq
 
-Subject to the magnetic solenoid and incompressibility constraints:
+Equations \ref{momentum} and \ref{induction} are subject to the magnetic solenoid and incompressibility constraints,
 
@@
 
+and
+
 \beq
-\nabla \cdot \mathbf{u} = 0
+\nabla \cdot \mathbf{u} = 0.
 \eeq
@@
 
-We define the flux function $A$ and streamfunction $\Psi$. These are scalar fields defined as the curl of the magnetic field and velocity, respectively, and so automatically satisfy our constraints.
+We define the flux function $A$ and streamfunction $\Psi$.
+$A$ is the familiar two-dimensional vector potential.
+$A$ and $\Psi$ are scalar fields defined as the curl of the magnetic field and velocity, respectively, and so automatically satisfy our constraints.
 
@@
 
+\bibitem[Goodman \& Xu(1994)]{Goodman:1994ul}
+Goodman, J, Xu, G, 1994, ApJ, 432, 213
+
 \bibitem[{Hollerbach(2009)}]{Hollerbach:2009ig}

### 2014-09-18 paper/mri_paper.tex
equations -> sentences

@@
 
+We perform a weakly nonlinear analysis following the lead of \citeauthor{Umurhan:2007hs}, and use much of the same notation, with some changes for added simplicity.
+
 We solve the equations of non-ideal, axisymmetric, incompressible magnetized Taylor Couette flow. We solve the momentum equation,
@@
 
-The perturbed, nondimensionalized equations which will be the focus of this equation are as follows:
-
-Our final equation set is:
+Our final equation set is
 
@@
 \label{eqset1}
-\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \\
+\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{y} \, = \\
 \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi
@@
 \label{eqset2}
-\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \\
-\, \frac{2}{\beta}B_0\partial_z B_{1y} \, + \, \frac{2}{\beta} J\left(A, B_{1y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{1y}
+\partial_t u_{1y} \, + \, J\left(\Psi, u_{y}\right) \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \\
+\, \frac{2}{\beta}B_0\partial_z B_{y} \, + \, \frac{2}{\beta} J\left(A, B_{y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{y}
 \end{multline}
@@
 \label{eqset4}
-\partial_t B_{1y} \, = \, B_0 \partial_z u_{1y} \, + \, J\left(A, u_{1y}\right) \, - \, J\left(\Psi, B_{1y}\right) \, \\
-+ \, \frac{1}{\reym} \nabla^2 B_{1y}  \, - \, q \Omega_0 \partial_z A
+\partial_t B_{y} \, = \, B_0 \partial_z u_{y} \, + \, J\left(A, u_{y}\right) \, - \, J\left(\Psi, B_{y}\right) \, \\
++ \, \frac{1}{\reym} \nabla^2 B_{y}  \, - \, q \Omega_0 \partial_z A,
 \end{multline}
 
-Where $J$ is the Jacobian operator, 
+where $J$ is the Jacobian operator, 
 \beq
@@
 $B_0$ appears in Equations \ref{eqset1} - \ref{eqset4} because it is the nondimensionalized background field strength.
-That is, $B = B_0 \equiv 1$.
+Because we nondimensionalize $B$ by the magnitude of the background field strength, $B_0 \equiv 1$.
 In order to study our system in the weakly nonlinear regime, we tune the background magnetic field down away from stability (recall that stronger vertical fields stabilize the MRI).
@@
 The destabilizing substitution is made, and Equations \ref{eqset1} - \ref{eqset4} are rewritten such that the fluid variables are contained in a state vector $\mathbf{V} = \left[\Psi, u_y, A, B_y\right]^\mathrm{T}$.
-This yields the following system of equations.
+This yields the system of equations
 
 \beq
-\mathcal{D}\partial_t \mathbf{V} + \mathbf{N} = \mathcal{L} \mathbf{V} - \epsilon^2\partial_z\mathcal{X} \mathbf{V} - \epsilon^2 \partial_z^3 \mathcal{L}_3\mathbf{V}
+\mathcal{D}\partial_t \mathbf{V} + \mathbf{N} = \mathcal{L} \mathbf{V} - \epsilon^2\partial_z\mathcal{X} \mathbf{V} - \epsilon^2 \partial_z^3 \mathcal{L}_3\mathbf{V},
 \eeq
 
-Where $\mathcal{L} \equiv \mathcal{L}_0 + \mathcal{L}_1\partial_z + \mathcal{L}_2\partial_z^2 + \mathcal{L}_3\partial_z^3 + \mathcal{L}_4\partial_z^4$ \\
+where $\mathcal{L} \equiv \mathcal{L}_0 + \mathcal{L}_1\partial_z + \mathcal{L}_2\partial_z^2 + \mathcal{L}_3\partial_z^3 + \mathcal{L}_4\partial_z^4$, \\
 
-And:
+and
 
@@
 The fast and slow variables are then defined such that each of the temporal and spatial eigenvalues appear at the same lowest order in the linear dispersion relation.
-The scalings are:
+The scalings are
 
 \beq
-X \equiv \epsilon x,  \, \, Y \equiv \epsilon y, \, \, Z \equiv \epsilon z, \, \, T \equiv \epsilon^2 t
+X \equiv \epsilon x,  \, \, Y \equiv \epsilon y, \, \, Z \equiv \epsilon z, \, \, T \equiv \epsilon^2 t.
 \eeq
@@
 
-where $\alpha$ is a slowly-varying amplitude equation.
+where $\alpha$ is a slowly-varying amplitude equation and $c.c.$ denotes the complex conjugate.
 The $x$ dependence is contained in $\mathbb{V}_{11}$, and must be solved for subject to the radial boundary conditions.
@@
 \epsilon^2 \mathcal{D}\partial_T\mathbf{V} \, + \, \mathbf{N} \, = \, \mathcal{L}\mathbf{V} \, + \, \epsilon\widetilde{\mathcal{L}}_1\partial_Z\mathbf{V} \, + \, \epsilon^2\widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V} \, \\
- - \, \epsilon^2\partial_z\mathcal{X}\mathbf{V} \, - \, \epsilon^2\partial_z^3\mathcal{L}_3\mathbf{V} \, + \, \mathcal{O}(\epsilon^3)
+ - \, \epsilon^2\partial_z\mathcal{X}\mathbf{V} \, - \, \epsilon^2\partial_z^3\mathcal{L}_3\mathbf{V} \, + \, \mathcal{O}(\epsilon^3),
 \end{multline}
 
-Where
+where
 
@@
 
+and
+
 \beq
-\widetilde{\mathcal{L}}_2 = \mathcal{L}_2 + 3\mathcal{L}_3\partial_z + 6\mathcal{L}_4\partial_z^2
+\widetilde{\mathcal{L}}_2 = \mathcal{L}_2 + 3\mathcal{L}_3\partial_z + 6\mathcal{L}_4\partial_z^2.
 \eeq
@@
 % perturbation expansion
-The state vector is expanded in orders of $\epsilon$:
+The state vector is expanded a perturbation series in orders of $\epsilon$,
 
@@
 \label{ordere}
-\mathcal{L}\mathbf{V_1} = 0
+\mathcal{L}\mathbf{V_1} = 0.
 \eeq
 
-To $\mathcal{O}(\epsilon^2)$:
+To $\mathcal{O}(\epsilon^2)$,
 
@@
 \label{ordere2}
-\mathcal{L}\mathbf{V_2} \, = \, \mathbf{N_2} \, - \, \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}
+\mathcal{L}\mathbf{V_2} \, = \, \mathbf{N_2} \, - \, \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}.
 \eeq
 
-To $\mathcal{O}(\epsilon^3)$:
+To $\mathcal{O}(\epsilon^3)$,
 
@@
 \mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, \\
-- \partial_Z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1}
+- \partial_Z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1}.
 \end{multline}
@@
 
-The partial differential equations that comprise Equations \ref{ordere} to \ref{ordere3} are solved in succession subject to no-slip, perfectly conducting radial boundary conditions, defined as:
+The partial differential equations that comprise Equations \ref{ordere} to \ref{ordere3} are solved in succession subject to no-slip, perfectly conducting radial boundary conditions, defined as
 
 \beq
-\Psi = \partial_x \Psi = u_y = A = \partial_x B_y = 0
+\Psi = \partial_x \Psi = u_y = A = \partial_x B_y = 0.
 \eeq
@@
 \beq
-\mathbf{V}^\dagger = \mathbb{V}^\dagger(x) e^{i k_c z} + c.c.
+\mathbf{V}^\dagger = \mathbb{V}^\dagger(x) e^{i k_c z} + c.c.,
 \eeq
@@
 where, again, $\mathbb{V}^\dagger$ contains the radial contribution.
-The elimination of this inner product must also be enforced the terms in Equation \ref{ordere3} that are resonant with $e^{i k_c z}$, yielding
+The elimination of this inner product must also be enforced on the terms in Equation \ref{ordere3} that are resonant with $e^{i k_c z}$, yielding
 
@@
 \label{ampeq1}
-a \partial_T \alpha \, + \, c \alpha \left|\alpha^2\right| \, + \, \widetilde{c} \alpha \partial_Z \beta \, = \, - b \partial_Z \alpha \, + \, h \partial_Z^2 \alpha \, + \, g i k_c^3 \alpha
+a \partial_T \alpha \, = - \, c \alpha \left|\alpha^2\right| \, - \, \widetilde{c} \alpha \partial_Z \beta \,  - b i k_c \alpha \, + \, h \partial_Z^2 \alpha \, + \, g i k_c^3 \alpha,
 \eeq
 
-Where
+where
 
@@
 h & \equiv \left< \mathbb{V}^\dagger \cdot \left(\widetilde{\mathcal{L}}_2\mathbb{V}_{11} \, - \, \widetilde{\mathcal{L}}_1\mathbb{V}_{21} \right)^* \right> \\
-g & \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{L}_{3} \mathbb{V}_{11} \right> \\
+g & \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{L}_{3} \mathbb{V}_{11} \right>. \\
 \end{split}

### 2014-09-18 paper/mri_paper.tex
more of jeff's comments, some intro

@@
 This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime.
+The present work adds to this foundation both in problems analyzed, with our analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space.
 
@@
 Thus our analytical framework is able to side-step many of the resolution issues faced by multidimensional simulations.
-We are able to resolve even significant structure in the boundary layers of our domain, because we need only resolve it in one dimension.
+We are able to resolve even small-scale structure in the boundary layers of our domain, because we need only resolve it in one dimension.
 We solve the radial component of each equation using the open source pseudospectral code Dedalus (Burns et al., in prep).
@@
 \citei{Boyd:2001aa}).
+The nonuniform spacing of the Chebyshev grid allows us to resolve the boundary structures very well on 256-point grid.
 
@@
 
-To find a bounded solution at each order, we eliminate secular terms -- terms which are resonant with solutions to the homogenous equation, and cause the solution to grow without bound.
+To find a bounded solution at each order, we eliminate secular terms: terms which are resonant with solutions to the homogenous equation and cause the solution to grow without bound.
 The elimination of secular terms requires enforcement of solvability criteria, which arise as a consequence of the Fredholm alternative (e.g.
@@
 The scale separations caused by these large ratios of diffusivity to viscosity render the small Prandtl number regime inaccessible to simulations.
-This analytic formulation allows us to explore the behavior down to magnetic Prandtl numbers of [....].
+This analytic formulation allows us to explore the behavior down to magnetic Prandtl numbers of [TBD!].
 

### 2014-09-19 paper/mri_paper.tex
knobloch and julien, pessah

@@
 Indeed, only a few works have tackled MRI saturation analytically.
-We briefly summarize a few of these analytic contributions.
+\citeauthor{Knobloch:2005ba} (\citeyear{Knobloch:2005ba}) analyze the MRI in the strongly nonlinear regime, by following the already-developed MRI modes into asymptotic saturation.
+They find that this saturation proceeds via modification of the background shear.
+Another notable analytic work is that of \citeauthor{Pessah:2010ic} (\citeyear{Pessah:2010ic}), who presents an analysis of MRI saturation via secondary instabilities.
+This work suggests that MRI saturation can be caused by parasite instabilities growing off of the primary MRI modes, in particular by Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime.
 
@@
 The present work adds to this foundation both in problems analyzed, with our analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space.
+In particular, we attempt to elucidate the role of boundary layers in the saturation of the MRI, as this is of fundamental importance to laboratory MRI work.
 
@@
 
+\bibitem[{Knobloch \& Julien(2005)}]{Knobloch:2005ba}
+Knobloch, E, Julien K, 2005, Physics of Fluids, 17, 094106
+
+\bibitem[Pessah(2010)]{Pessah:2010ic}
+Pessah, M, 2010, ApJ, 716, 1012
+
 \bibitem[{Schartman et al.(2009)}]{Schartman:2009df}

### 2014-09-19 paper/mri_paper.tex
nondimensionalization, thin gap

@@
 
-\newcommand\reye{\mathrel{Re}}
-\newcommand\reym{\mathrel{Rm}}
+\newcommand\reye{\mathrm{Re}}
+\newcommand\reym{\mathrm{Rm}}
 
@@
 This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime.
-The present work adds to this foundation both in problems analyzed, with our analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space.
+The present work adds to this foundation both in subject, with our additional analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space.
 In particular, we attempt to elucidate the role of boundary layers in the saturation of the MRI, as this is of fundamental importance to laboratory MRI work.
@@
 We investigate a 2D slice of the X-Z (radial-vertical) plane.
-Our domain is the bolded black box, of width L.
+Our domain is represented by the bolded black box, of width L.
 The radial dimension is solved with a basis of Chebyshev polynomials, and the vertical dimension is solved on a Fourier basis (see \S\ref{discussion}.) }\label{setup}
@@
 \beq
-thin gap equation,
+\left(R_2 - R_1\right) \ll \frac{1}{2}\left(R_2 + R_1\right),
 \eeq
 
-the width of the channel is small compared to its distance from the center of rotation.
+where $R_1$ is the radius of the inner cylinder and $R_2$ is the radius of the outer cylinder, such that the width of the channel is small compared to its distance from the center of rotation.
 The imposition of the thin-gap approximation eliminates channel modes, the axisymmetric MRI modes that are exact solutions of the incompressible MRI in the shearing box approximation (\citei{Goodman:1994ul}).
@@
 
-We perturb the fluid quantities with three-dimensional, axisymmetric perturbations. We nondimensionalize the equations according to [...]
-The fluid symbols $\mathbf{u}$, $\mathbf{B}$, etc. will henceforth be used to refer to the perturbed quantities.
+We perturb the fluid quantities with three-dimensional, axisymmetric perturbations.
+We nondimensionalize the equations, with lengths nondimensionalized by $L$, velocities by $\Omega_0 L$, magnetic fields by $B_0$, and pressure by $\Omega_0^2 L^2 \rho_0$, where $L$ is the channel width, $\Omega_0$ is the rotation rate at the center of the channel, and $\rho_0$ is the constant pressure in the base state (see Figure \ref{setup}).
+We define the nondimensional Reynolds number,
+
+\beq
+\reye \equiv \frac{\Omega_0 r_0^2 \delta^2}{\nu},
+\eeq
+
+and magnetic Reynolds number,
+
+\beq
+\reym \equiv \frac{\Omega_0 r_0^2 \delta^2}{\eta}.
+\eeq
+
+The fluid symbols $\mathbf{u}$, $\mathbf{B}$, etc. will henceforth be used to refer to the nondimensional, perturbed quantities.
 

### 2014-09-19 paper/mri_paper.tex
nonlinear stuff
much more nonlinear

@@
 
+The nonlinear terms are contained in a vector $\mathbf{N} = \left[N^{(\Psi)}, N^{(u_y)}, N^{(A)}, N^{(B_y)}\right]^\mathrm{T}$, where
+
+\beq
+\mathbf{N} = \left[ \begin{matrix}  J(\Psi, \nabla^2 \Psi) - \frac{2}{\beta}J(A, \nabla^2 A)\\
+J(\Psi, u_{y}) - \frac{2}{\beta} J(A, B_{y}) \\ 
+-J(A, \Psi) \\
+J(\Psi, B_{y}) - J(A, u_{y}) \\
+\end{matrix} \right].
+\eeq
+
 We conduct a formal multiscale analysis.
@@
 Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic Taylor Couette flow.
+In analogy to these problems, we posit slow variation in both $Z$ and $T$.
 Each operator is expanded to reflect these scalings -- for instance, $\partial_z$ becomes $\partial_z + \epsilon\partial_Z$.
@@
 \mathbf{V} = \epsilon\mathbf{V_1} + \epsilon^2\mathbf{V_2} + \epsilon^3\mathbf{V_3} + ...
+\eeq
+
+The multiscale expansion of the operators and the perturbation series expansion of the fluid quantities combine to make the expansion of the nonlinear vector $\mathbf{N}$ expansive, and we relegate the expansion to Appendix (REF).
+$\mathbf{N}$ becomes
+
+\beq
+\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4),
+\eeq
+
+where
+
+\beq
+\mathbf{N_2} = \alpha^2\mathbb{N}_{22} e^{i2 k_c z} + \left|\alpha\right|^2 \mathbb{N}_{20} + c.c.
 \eeq
@@
 
+\beq
+\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4)
+\eeq
+
+where
+
+\beq
+N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \frac{2}{\beta}J(A_1, \nabla^2 A_1) 
+\eeq
+
+\beq
+N_2^{(u)} = J(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_1) 
+\eeq
+
+\beq
+N_2^{(A)} = - J(A_1, \Psi_1) 
+\eeq
+
+\beq
+N_2^{(B)} =  J(\Psi_1, B_1) \, - \, J(A_1, u_1)
+\eeq
+
+and
+
+\beq
+\begin{split}
+N_3^{(\Psi)} & = J(\Psi_1, \nabla^2\Psi_2) \, - \, \frac{2}{\beta}J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) - \, \frac{2}{\beta}J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, \\
+& - \, 2 \frac{2}{\beta}J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \frac{2}{\beta}\widetilde{J}(A_1, \nabla^2 A_1)\\
+\end{split}
+\eeq
+
+\beq
+N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_2) \, - \, \frac{2}{\beta} J(A_2, B_1) \, - \, \frac{2}{\beta} \widetilde{J}(A_1, B_1)
+\eeq
+
+\beq
+N_3^{(A)} = - J(A_1, \Psi_2) \, - \, J(A_2, \Psi_1) \, - \, \widetilde{J}(A_1, \Psi_1)
+\eeq
+
+\beq
+N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1)
+\eeq
+
 Second order nonlinear terms: \\

### 2014-09-19 paper/mri_paper.tex
appendix stuff

@@
 \usepackage[titletoc,title]{appendix}
+\usepackage{cleveref}
 
@@
 
-The multiscale expansion of the operators and the perturbation series expansion of the fluid quantities combine to make the expansion of the nonlinear vector $\mathbf{N}$ expansive, and we relegate the expansion to Appendix (REF).
+The multiscale expansion of the operators and the perturbation series expansion of the fluid quantities combine to make the expansion of the nonlinear vector $\mathbf{N}$ expansive, and we relegate the expansion to Appendix \ref{app:appendix_nonlinear}.
 $\mathbf{N}$ becomes
@@
 \beq
-\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4),
-\eeq
-
-where
-
-\beq
-\mathbf{N_2} = \alpha^2\mathbb{N}_{22} e^{i2 k_c z} + \left|\alpha\right|^2 \mathbb{N}_{20} + c.c.
+\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4).
 \eeq
@@
 \newpage
-\begin{appendices}
+\begin{subappendices}
 \begin{widetext}
-\section*{Appendix A.}
+\section*{Appendix A. Nonlinear term expansion}\label{app:appendix_nonlinear}
 
@@
 
-Second order nonlinear terms: \\
+$\mathbf{N_2}$ and $\mathbf{N_3}$ expand to become
+
+\beq
+\mathbf{N_2} = \alpha^2\mathbb{N}_{22} e^{i2 k_c z} + \left|\alpha\right|^2 \mathbb{N}_{20} + c.c.
+\eeq
+
+and
+
+\beq
+\mathbf{N_3} = \alpha^3 \mathbb{N}_{33} e^{i 3 k_c z} + \alpha\partial_Z\alpha \mathbb{N}_{32} e^{i 2 k_c z} + \alpha\left|\alpha\right|^2 \mathbb{N}_{31} e^{i k_c z} + \alpha \partial_Z  \beta \mathbb{\widetilde{N}}_{31} e^{i k_c z} + \alpha^*\partial_Z \alpha \mathbb{N}_{30} + c.c.
+\eeq
+
+The second order nonlinear terms are \\
 
@@
 
-Third order nonlinear terms: \\
+and the third order nonlinear terms become \\
 
@@
 \end{widetext}
-\end{appendices}
+\end{subappendices}
 

### 2014-09-19 paper/mri_paper.tex
fixed appendix.
more appendix

@@
 The relative scalings of the fast and slow scales are determined as follows.
-When our four equations are linearized for axisymmetric perturbations of the form $e^{\sigma t + i k_x x + i k_z z}$, we can derive the linear dispersion relation, which is fourth order in $\sigma$.
+When our four equations are linearized for axisymmetric perturbations of the form $e^{\sigma t + i k_x x + i k_z z}$, we can derive the linear dispersion relation, which is fourth order in $\sigma$ (Appendix \ref{app:dispersion}).
 The fast and slow variables are then defined such that each of the temporal and spatial eigenvalues appear at the same lowest order in the linear dispersion relation.
@@
 
-The multiscale expansion of the operators and the perturbation series expansion of the fluid quantities combine to make the expansion of the nonlinear vector $\mathbf{N}$ expansive, and we relegate the expansion to Appendix \ref{app:appendix_nonlinear}.
+The multiscale expansion of the operators and the perturbation series expansion of the fluid quantities combine to make the expansion of the nonlinear vector $\mathbf{N}$ expansive, and we relegate the expansion to Appendix \ref{app:nonlinear}.
 $\mathbf{N}$ becomes
@@
 This is the saturation amplitude.
+We use the saturation amplitude as $\alpha$ and the radial component of computed on the Chebyshev grid to plot $\mathbf{V}$.
+We show both first order ($\epsilon\mathbf{V}_1$), second order ($\epsilon^2\mathbf{V}_2$), and combined first and second order fluid perturbations ($\epsilon\mathbf{V}_1$ + ($\epsilon^2\mathbf{V}_2$).
+Figure \ref{allorders_velocity} shows the velocity perturbations, and Figure \ref{allorders_Bfield} shows the magnetic field perturbations.
 
-\section{Results}
+\section{Parameter Space}
 
@@
 Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds.
-The colorbar represents azimuthal velocity.}
+The colorbar represents azimuthal velocity.}\label{allorders_velocity}
 \end{figure*}
 
-\begin{figure*}[h!]
+\begin{figure*}[h!]\label{allorders_Bfield}
 \centering
@@
 
-\newpage
-\begin{subappendices}
-\begin{widetext}
-\section*{Appendix A. Nonlinear term expansion}\label{app:appendix_nonlinear}
+\clearpage
+\appendix
 
+\section{A. Linear dispersion relation}\label{app:dispersion}
+%Note: I don't expect this to go in the actual paper, but it could be useful in the meeting.
+
+The linear dispersion relation, which determines the variable scalings in the multiscale analysis. \\
+
+\beq
+\begin{split}
+& \frac{B_{0}^{4} k_{x}^{2} k_{z}^{4}}{16 \pi^{2}} + \frac{B_{0}^{4} k_{z}^{6}}{16 \pi^{2}} - \frac{B_{0}^{2} \Omega_{0} k_{z}^{4} q}{2 \pi} - 2 \Omega_{0} k_{z}^{2} q \sigma^{2} - \frac{4 \sigma}{\reym} \Omega_{0} k_{x}^{2} k_{z}^{2} q - \frac{4 \sigma}{\reym} \Omega_{0} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{x}^{4} k_{z}^{2} q - \frac{4 \Omega_{0}}{\reym^{2}} k_{x}^{2} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{z}^{6} q \\
+& - k_{x}^{2} \sigma^{4} - k_{z}^{2} \sigma^{4} + 4 k_{z}^{2} \sigma^{2} - \frac{2 \sigma^{3}}{\reym} k_{x}^{4} - \frac{4 \sigma^{3}}{\reym} k_{x}^{2} k_{z}^{2} + \frac{8 \sigma}{\reym} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reym} k_{z}^{4} + \frac{8 \sigma}{\reym} k_{z}^{4} - \frac{k_{x}^{6} \sigma^{2}}{\reym^{2}} - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{4} k_{z}^{2} + \frac{4 k_{x}^{4}}{\reym^{2}} k_{z}^{2} \\
+& - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{2} k_{z}^{4} + \frac{8 k_{x}^{2}}{\reym^{2}} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reym^{2}} + \frac{4 k_{z}^{6}}{\reym^{2}} - \frac{2 \sigma^{3}}{\reye} k_{x}^{4} - \frac{4 \sigma^{3}}{\reye} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reye} k_{z}^{4} - \frac{4 k_{x}^{6} \sigma^{2}}{\reye \reym} -  \frac{12 k_{x}^{4} k_{z}^{2} \sigma^{2}}{\reye \reym} - \frac{12 k_{x}^{2} k_{z}^{4} \sigma^{2}}{\reye \reym} \\
+& - \frac{4 k_{z}^{6} \sigma^{2}}{\reye \reym} - \frac{2 k_{x}^{8} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye \reym^{2}} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye \reym^{2}} - \frac{2 k_{z}^{8} \sigma}{\reye \reym^{2}} - \frac{k_{x}^{6} \sigma^{2}}{\reye^{2}} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{4} k_{z}^{2} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{2} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reye^{2}} \\
+& - \frac{2 k_{x}^{8} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye^{2} \reym} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye^{2} \reym} - \frac{2 k_{z}^{8} \sigma}{\reye^{2} \reym} - \frac{k_{x}^{10}}{\reye^{2} \reym^{2}} - \frac{5 k_{x}^{8} k_{z}^{2}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{6} k_{z}^{4}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{4} k_{z}^{6}}{\reye^{2} \reym^{2}} \\
+& - \frac{5 k_{x}^{2} k_{z}^{8}}{\reye^{2} \reym^{2}} - \frac{k_{z}^{10}}{\reye^{2} \reym^{2}} = 0
+\end{split}
+\eeq
+
+\section{B. Expansion of Nonlinear Terms}\label{app:nonlinear}
 \beq
@@
 \beq
-N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1)
+N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1).\\
 \eeq
@@
 
-\end{widetext}
-\end{subappendices}
-
 %\bibliographystyle{plain}

### 2014-09-19 paper/mri_paper.tex
cartoon
more

@@
 
-\beq
+\begin{figure}[h!]
+\centering
+\includegraphics[trim=6cm 3cm 3cm 0cm, scale=0.35]{slowfast.eps}
+\caption{Cartoon illustrating fast and slow-moving variables (Equation \ref{scalings}).}\label{allorders_velocity}
+\label{figscalings}
+\end{figure}
+
+\beq\label{scalings}
 X \equiv \epsilon x,  \, \, Y \equiv \epsilon y, \, \, Z \equiv \epsilon z, \, \, T \equiv \epsilon^2 t.
@@
 
+See Figure \ref{figscalings}.
 Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic Taylor Couette flow.
@@
 \eeq
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[trim=0cm 1cm 0cm 5cm, scale=0.45]{allorders_velocity.eps}
+\caption{First order (left) and second order (right) velocity perturbations.
+Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds.
+The colorbar represents azimuthal velocity.}\label{allorders_velocity}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.45]{allorders_Bfield.eps}
+\caption{First order (left) and second order (right) magnetic field perturbations.
+Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to higher magnetic field strengths.
+The colorbar represents azimuthal velocity.}\label{allorders_Bfield}
+\end{figure*}
 
@@
 
+Our analytic formulation is advantageous because we can now analyze the contributions of each fluid quantity at each order.
+This allows us to directly identify the source of saturation.
+Equation \ref{ampeq1} asymptotically approaches a value, so the system is clearly saturating, and our approach now allows us to quantify the individual contribution of each term.
+In Figure \ref{allorders_velocity}, we see significant fluid action in the second-order velocity plot.
+By contrast, the second-order magnetic field perturbations are two orders of magnitude smaller than the first-order perturbations (Figure \ref{allorders_Bfield}).
+It appears that the second-order velocity perturbations are instrumental in the saturation of the MRI.
+The likely mechanism is a reduction in the background shear.
+
 \section{Parameter Space}
@@
 This analytic formulation allows us to explore the behavior down to magnetic Prandtl numbers of [TBD!].
-
-\begin{figure*}[h!]
-\centering
-\includegraphics[scale=0.5]{allorders_velocity.eps}
-\caption{First order (left) and second order (right) velocity perturbations.
-Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds.
-The colorbar represents azimuthal velocity.}\label{allorders_velocity}
-\end{figure*}
-
-\begin{figure*}[h!]\label{allorders_Bfield}
-\centering
-\includegraphics[scale=0.5]{allorders_Bfield.eps}
-\caption{First order (left) and second order (right) velocity perturbations.
-Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds.
-The colorbar represents azimuthal velocity.}
-\end{figure*}
 

### 2014-09-19 paper/mri_paper.tex
more

@@
 \label{V1_ansatz}
-\mathbf{V_1} = \alpha(T, Z) \mathbb{V}_{11}(x) e^{i k_c z} + c.c.,
+\mathbf{V_1} = \alpha(T, Z) \mathbb{V}_{11}(x) e^{i k_c z} + c.c. + \beta(T, Z)\mathbb{U}_{11}(x)\\
 \eeq
@@
 where $\alpha$ is a slowly-varying amplitude equation and $c.c.$ denotes the complex conjugate.
-The $x$ dependence is contained in $\mathbb{V}_{11}$, and must be solved for subject to the radial boundary conditions.
+The $x$ dependence is contained in $\mathbb{V}_{11} = (\Psi_{11}, u_{11}, A_{11}, B_{11})^\mathbf{T}$, and must be solved for subject to the radial boundary conditions.
 The periodic vertical boundary conditions allow us to posit the $z$ dependence, where $k_c$ is the value of the vertical wavenumber at marginality.
+As noted by \cite{Umurhan:2007hs}, there exists a spatially constant neutral mode solution to the $B_y$ equation, with $\mathbb{U}_{11} = (0, 0, 0, 1)^\mathbf{T}$.
 

### 2014-09-19 paper/mri_paper.tex
final ish

@@
 The periodic vertical boundary conditions allow us to posit the $z$ dependence, where $k_c$ is the value of the vertical wavenumber at marginality.
-As noted by \cite{Umurhan:2007hs}, there exists a spatially constant neutral mode solution to the $B_y$ equation, with $\mathbb{U}_{11} = (0, 0, 0, 1)^\mathbf{T}$.
+As noted by \citeauthor{Umurhan:2007hs}, there exists a spatially constant neutral mode solution to the $B_y$ equation, with $\mathbb{U}_{11} = (0, 0, 0, 1)^\mathbf{T}$.
+The slowly-varying amplitude $\beta$ encodes the slow evolution of this mode.
 

### 2014-09-19 paper/mri_paper.tex
second amplitude equation added

@@
 h & \equiv \left< \mathbb{V}^\dagger \cdot \left(\widetilde{\mathcal{L}}_2\mathbb{V}_{11} \, - \, \widetilde{\mathcal{L}}_1\mathbb{V}_{21} \right)^* \right> \\
-g & \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{L}_{3} \mathbb{V}_{11} \right>. \\
+g & \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{L}_{3} \mathbb{V}_{11} \right>, \\
 \end{split}
+\eeq
+
+but $\widetilde{c}$ is equal to zero and drops out of the equation.
+This is the well-known Ginzburg-Landau equation, and is with only slight modification the same amplitude equation derived in \citeauthor{Umurhan:2007hs}.
+
+There exists a second amplitude equation, corresponding to the neutral mode, which comes from the solvability condition that must be imposed on all terms in Equation \ref{ordere3} that have no $z$ dependence.
+These terms are
+
+\beq
+\begin{split}
+\mathcal{D}\mathbb{U}_{11} \partial_T \beta \, & = \, \mathcal{L}\mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{20} 2 \left|\alpha\right| \partial_Z \alpha \, - \, \widetilde{\mathcal{L}}_1 \mathbb{U}_{20} \partial_Z^2 \beta \, \\
+& + \, \widetilde{\mathcal{L}}_2 \mathbb{U}_{11} \partial_Z^2 \beta \, - \, \mathcal{X} \mathbb{U}_{11} \partial_Z \beta,
+\end{split}
+\eeq
+
+and taking the inner product of these terms with the adjoint homogenous solution $\mathbb{U}_{11}$ yields
+
+\beq
+\partial_T \beta = k 2 \left|\alpha\right| \partial_Z \alpha - m \partial_Z^2 \beta + \frac{1}{\reym} \partial_Z^2 \beta - p \alpha^* \partial_Z \alpha,
+\eeq
+
+where
+
+\beq
+\begin{split}
+n & \equiv \left< \mathbb{U}_{11} \cdot \left(\mathcal{L} \mathbb{V}_{30}\right)^*\right> \\
+k & \equiv \left< \mathbb{U}_{11} \cdot \left(\widetilde{\mathcal{L}}_1 \mathbb{V}_{20}\right)^*\right> \\
+m & \equiv \left< \mathbb{U}_{11} \cdot \left(\widetilde{\mathcal{L}}_1 \mathbb{U}_{20}\right)^*\right> \\
+p & \equiv \left< \mathbb{U}_{11} \cdot N_{30}^*\right>, \\
+\end{split}
+\eeq
+
+but both $k$ and $p$ are equal to zero, so this amplitude equation decouples from Equation \ref{ampeq1}, and reveals itself to be a simple diffusion equation,
+
+\beq\label{ampeq2}
+\partial_T \beta = - m \partial_Z^2 \beta + \frac{1}{\reym} \partial_Z^2 \beta.
 \eeq

### 2014-09-19 paper/mri_paper.tex
N20 figure, jeff's edits

@@
 
-We perturb the fluid quantities with three-dimensional, axisymmetric perturbations.
+We axisymmetrically perturb all three vector components of each of the fluid quantities.
 We nondimensionalize the equations, with lengths nondimensionalized by $L$, velocities by $\Omega_0 L$, magnetic fields by $B_0$, and pressure by $\Omega_0^2 L^2 \rho_0$, where $L$ is the channel width, $\Omega_0$ is the rotation rate at the center of the channel, and $\rho_0$ is the constant pressure in the base state (see Figure \ref{setup}).
@@
 
+\begin{figure*}
+\centering
+\includegraphics[trim=0cm 10cm 0cm 5cm, scale=.5]{N20fig.eps}
+\caption{Example of a nonlinear term solved on a 256-point Chebyshev grid. Circles represent values on the Chebyshev grid, grey lines are fits.}\label{N20fig}
+\end{figure*}
+
 \beq\label{scalings}
@@
 The nonuniform spacing of the Chebyshev grid allows us to resolve the boundary structures very well on 256-point grid.
+See Figure \ref{N20fig} for an example of the boundary-layer resolution we achieve in a nonlinear term.
 

### 2014-09-24 paper/mri_paper.tex

@@
 Another notable analytic work is that of \citeauthor{Pessah:2010ic} (\citeyear{Pessah:2010ic}), who presents an analysis of MRI saturation via secondary instabilities.
-This work suggests that MRI saturation can be caused by parasite instabilities growing off of the primary MRI modes, in particular by Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime.
+This work suggests that MRI saturation can be caused by parasitic instabilities growing off of the primary MRI modes, in particular by Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime.
 
@@
 In recent years much experimental progress has been made, although no experiment has yet yielded a confirmed detection.
-Of most immediate interest to this work is the Princeton Plasma Physics Laboratory (PPPL) MRI experiment, which has a design roughly analogous to the set-up investigated here (see \S\ref{setup} for further details).
+Of most immediate interest to this work is the Princeton Plasma Physics Laboratory (PPPL) MRI experiment, which has a design roughly analogous to the set-up investigated here (see \S\ref{sec:setup} for further details).
 The Madison plasma dynamo experiment is a multi-purpose plasma device which may be a vehicle for MRI studies in the near future (\citei{Cooper:2013to}).
@@
 In order to track the evolution of the system on both rapid and slow scales, we conduct a formal multiscale analysis.
+We develop and solve this framework in \S\ref{wna}, conduct a full parameter space study in \S\ref{sec:pspace}, and explore the implications of our findings in \S\ref{discussion}.
 
@@
 \section{Set-Up}
+\label{sec:setup}
 
@@
 
-Our domain is designed to be relevant to the design of the PPPL MRI experiment (see \S\ref{discussion} for further discussion).
+Our domain is designed to be relevant to the apparatus of the PPPL MRI experiment (see \S\ref{discussion} for further discussion).
 We use periodic vertical boundary conditions.
@@
 
-We perform a weakly nonlinear analysis following the lead of \citeauthor{Umurhan:2007hs}, and use much of the same notation, with some changes for added simplicity.
+We perform a weakly nonlinear analysis following the lead of \citeauthor{Umurhan:2007hs}, and use much of the same notation, with some changes for added clarity.
 
@@
 \beq\label{momentum}
-\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, = 
+\begin{split}
+& \partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, = \, -\frac{1}{\rho}\nabla P \, - \, \nabla\Phi \, + \, \frac{1}{\rho} \left(\mathbf{J}\times\mathbf{B}\right) \, \\
+& + \, \nu\nabla^2 \mathbf{u} \, - \, 2\mathbf{\Omega} \times \mathbf{u} \, - \, \mathbf{\Omega} \times \left(\mathbf{\Omega} \times \mathbf{r} \right), \\
+\end{split}
 \eeq
-
-$\, -\frac{1}{\rho}\nabla P \, - \, \nabla\Phi \, + \, \frac{1}{\rho} \left(\mathbf{J}\times\mathbf{B}\right) \, + \, \nu\nabla^2 \mathbf{u} \, - \, 2\mathbf{\Omega} \times \mathbf{u} \, - \, \mathbf{\Omega} \times \left(\mathbf{\Omega} \times \mathbf{r} \right)$
 
@@
 \beq
-\reye \equiv \frac{\Omega_0 r_0^2 \delta^2}{\nu},
+\reye \equiv \frac{\Omega_0 L^2}{\nu},
 \eeq
@@
 \beq
-\reym \equiv \frac{\Omega_0 r_0^2 \delta^2}{\eta}.
+\reym \equiv \frac{\Omega_0 L^2}{\eta}.
 \eeq
@@
 where $\alpha$ is a slowly-varying amplitude equation and $c.c.$ denotes the complex conjugate.
-The $x$ dependence is contained in $\mathbb{V}_{11} = (\Psi_{11}, u_{11}, A_{11}, B_{11})^\mathbf{T}$, and must be solved for subject to the radial boundary conditions.
+The $x$ dependence is contained in $\mathbb{V}_{11} = (\Psi_{11}, u_{11}, A_{11}, B_{11})^\mathbf{T}$, and must be solved subject to the radial boundary conditions.
 The periodic vertical boundary conditions allow us to posit the $z$ dependence, where $k_c$ is the value of the vertical wavenumber at marginality.
@@
 Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to higher magnetic field strengths.
-The colorbar represents azimuthal velocity.}\label{allorders_Bfield}
+The colorbar represents azimuthal magnetic field strength.}\label{allorders_Bfield}
 \end{figure*}
@@
 % perturbation expansion
-The state vector is expanded a perturbation series in orders of $\epsilon$,
+The state vector is expanded in a perturbation series in orders of $\epsilon$,
 
@@
 
-The multiscale expansion of the operators and the perturbation series expansion of the fluid quantities combine to make the expansion of the nonlinear vector $\mathbf{N}$ expansive, and we relegate the expansion to Appendix \ref{app:nonlinear}.
+The multiscale expansion of the operators and the perturbation series expansion of the fluid quantities combine to make the expansion of the nonlinear vector $\mathbf{N}$ formidable, and we relegate the expansion to Appendix \ref{app:nonlinear}.
 $\mathbf{N}$ becomes
@@
 
+In the above, we find that $\mathbf{V_2}$ takes the form
+
+\beq
+\begin{split}
+\mathbf{V_2} \, & = \, \alpha^2 e^{i2Qz} \mathbb{V}_{22}(x) \, + \, \left|\alpha\right|^2 \mathbb{V}_{20}(x) \, \\
+&- \, \partial_Z \alpha \mathbb{V}_{21}(x) e^{iQz} \,- \, \partial_Z \beta \mathbb{U}_{20}(x) + c.c.
+\end{split}
+\eeq
+
 The $\mathcal{O}(\epsilon)$ equation is the equation for the linear MRI at marginality. This is the equation that is solved to find the $\mathbb{V}_{11}$ component of Equation \ref{V1_ansatz}.
@@
 \citei{Boyd:2001aa}).
-The nonuniform spacing of the Chebyshev grid allows us to resolve the boundary structures very well on 256-point grid.
+The nonuniform spacing of the Chebyshev grid allows us to resolve the boundary structures well on a 256-point grid.
 See Figure \ref{N20fig} for an example of the boundary-layer resolution we achieve in a nonlinear term.
@@
 where, again, $\mathbb{V}^\dagger$ contains the radial contribution.
-The elimination of this inner product must also be enforced on the terms in Equation \ref{ordere3} that are resonant with $e^{i k_c z}$, yielding
+The elimination of this inner product must also be enforced on the terms in Equation \ref{ordere3} that are resonant with $e^{i k_c z}$, which are
+
+\beq
+\begin{split}
+& \mathcal{D} \mathbb{V}_{11} e^{iQz}\partial_T \alpha \, + \, \alpha \left| \alpha^2 \right| \mathbf{N}_{31} e^{iQz} \, + \, \alpha \partial_Z \beta \widetilde{\mathbf{N}}_{31} e^{iQz} \, =  \\
+& \, - \partial_Z^2 \alpha \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} e^{iQz} \, + \, \partial_Z^2 \alpha \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{iQz} \, - \, \partial_Z \alpha \mathbb{V}_{11}\mathcal{X}e^{iQz} \, \\
+& + \, iQ^3 \alpha \mathcal{L}_3 \mathbb{V}_{11} e^{iQz}.
+\end{split}
+\eeq
+
+Taking the inner product yields the amplitude equation
 
@@
 
-but $\widetilde{c}$ is equal to zero and drops out of the equation.
-This is the well-known Ginzburg-Landau equation, and is with only slight modification the same amplitude equation derived in \citeauthor{Umurhan:2007hs}.
+where ${}^*$ denotes a complex conjugate, but $\widetilde{c}$ is equal to zero and drops out of the equation.
+This is the well-known Ginzburg-Landau equation, and is with only slight modification the same amplitude equation derived in \citeauthor{Umurhan:2007hs}
 
@@
 
-but both $k$ and $p$ are equal to zero, so this amplitude equation decouples from Equation \ref{ampeq1}, and reveals itself to be a simple diffusion equation,
+but both $k$ and $p$ are equal to zero. Thus this amplitude equation decouples from Equation \ref{ampeq1}, and reveals itself to be a simple diffusion equation,
 
@@
 
-\section{Parameter Space}
+\section{Parameter Space}\label{sec:pspace}
 

### 2014-10-15 paper/mri_paper.tex
debugging in param space search and paper edits

@@
 
-Equations \ref{momentum} and \ref{induction} are subject to the magnetic solenoid and incompressibility constraints,
+In the above, the current density $\mathbf{J} = \frac{2}{\beta}\del\times\mathbf{B}$,  Equations \ref{momentum} and \ref{induction} are subject to the magnetic solenoid and incompressibility constraints,
 
@@
 \reym \equiv \frac{\Omega_0 L^2}{\eta}.
+\eeq
+
+We also define a plasma beta parameter,
+
+\beq
+\beta \equiv \frac{\Omega_0^2r_0^2}{v_A^2},
+\eeq
+
+where the Alfv\'en sped $v_A$ is
+
+\beq
+v_A^2 = \frac{B_0^2}{\rho_0}.
 \eeq

### 2014-10-16 notes/mri_notes.tex
more paper and minor notes clarification
notes typo
nondimensionalization, thin gap
started adding hMRI notes

@@
 And we introduce the following terms:
-\[\reye \equiv \frac{\Omega_0 r_0^2 \delta^2}{\nu} ; \, \, \, \reym \equiv \frac{\Omega_0 r_0^2 \delta^2}{\eta}\]
+\[\reye \equiv \frac{\Omega_0 r_0^2 \delta^2}{\nu} \reym \equiv \frac{\Omega_0 r_0^2 \delta^2}{\eta}\]
 
@@
 
-$-\epsilon^2\partial_z\mathcal{X} \to \left(-\epsilon^2 - \epsilon^3\partial_Z\right)\mathcal{X}$ \\
+$-\epsilon^2\partial_z\mathcal{X} \to \left(-\epsilon^2\partial_z - \epsilon^3\partial_Z\right)\mathcal{X}$ \\
 
@@
 
-We drop the $\partial_t$ terms and our $\mathcal{O}(\epsilon^2)$ equation is: \\
+We drop the $\partial_t$ terms and our equation up to order $\mathcal{O}(\epsilon^2)$ is: \\
 
@@
 
+\pagebreak
+\section*{Helical MRI}
+
+Instead of the standard MRI, we initialize an azimuthal field in addition to the constant vertical field.
+The azimuthal field will linearly decrease radially, since in any reasonable setup it will be created by an axial current through the center of the apparatus.
+
+The base state becomes:
+
+\begin{equation}
+\mathbf{B} = B_0 \hat{z} + \xi B_0 \left(\frac{x_1}{x}\right) \hat{y} + \mathbf{B}_1
+\end{equation}
+
 \bibliography{mri.bib}

### 2014-10-18 notes/mri_notes.tex
more hmri derivation
final momentum terms for hmri
induction eqn hmri

@@
 \newcommand\nat{ Nature }
+
+\newcommand{\beq}{\begin{equation}}
+\newcommand{\eeq}{\end{equation}}
 
@@
 
+The Lorentz term in the momentum equation is the only term that will change. It becomes
+
+\beq
+\left(\nabla \times \left(B_0 \hat{z} + \xi B_0 \left(\frac{x_1}{x}\right) \hat{y} + \mathbf{B}_1
+\right) \right) \times \left(B_0 \hat{z} + \xi B_0 \left(\frac{x_1}{x}\right) \hat{y} + \mathbf{B}_1
+\right)
+\eeq
+
+and we gain a potential 5 terms that were not present before, since there are 5 terms in this expansion that involve the azimuthal term. These are as follows.
+
+\beq
+\left(\nabla \times B_0 \hat{z} \right) \times \xi B_0 \left(\frac{x_1}{x}\right)\hat{y} = 0
+\eeq
+
+because of the constant term.
+
+\beq
+\left(\nabla \times \xi B_0 \left(\frac{x_1}{x}\right) \hat{y} \right) \times B_0 \hat{z} = \left(\partial_x \left(\xi B_0 \left(\frac{x_1}{x}\right)\right)\hat{z}\right) \times B_0 \hat{z} = 0
+\eeq
+
+because $\hat{z} \times \hat{z} = 0$.
+
+\beq
+\left(\nabla \times \xi B_0 \frac{x_1}{x} \hat{y} \right) \times \left( \xi B_0 \frac{x_1}{x}\hat{y}\right) = \xi^2 B_0^2 \frac{x_1^2}{x^3} \hat{x}
+\eeq
+
+And
+
+\beq
+\left(\nabla \times \xi B_0 \frac{x_1}{x} \hat{y}\right) \times \mathbf{B_1} = -\xi B_0 \frac{x_1}{x^2} \left(B_{1x} \hat{y} - B_{1y} \hat{x}\right)
+\eeq
+
+And finally,
+
+\beq
+\left(\nabla \times \mathbf{B}_1 \right) \times \left(\xi B_0 \frac{x_1}{x} \hat{y}\right) = - \partial_z B_{1y} \xi B_0 \frac{x_1}{x} \hat{z} - \partial_x B_{1y} \xi B_0 \frac{x_1}{x} \hat{x}
+\eeq
+
+So our final added terms to the momentum equation are:
+
+\beq
+\xi^2 B_0^2 \frac{x_1^2}{x^3} \hat{x} \, - \,  \xi B_0 \frac{x_1}{x^2} B_{1x} \hat{y} \, + \, \xi B_0 \frac{x_1}{x^2} B_{1y} \hat{x} \, - \, \partial_z B_{1y} \xi B_0 \frac{x_1}{x} \hat{z} \, - \, \partial_x B_{1y} \xi B_0 \frac{x_1}{x} \hat{x}
+\eeq
+
+For the induction equation, we have:
+
+\beq
+\partial_t \mathbf{B} = \left(\mathbf{B} \cdot \nabla \right) \mathbf{u} - \left(\mathbf{u} \cdot \nabla \right) \mathbf{B} + \frac{1}{\reym} \nabla^2 \mathbf{B}
+\eeq
+
+The first term on the righthand side becomes:
+
+\beq
+\partial_t \mathbf{B} = \left( \left( \mathbf{B_0} \hat{z} + \xi B_0 \frac{x_1}{x} \hat{y} + \mathbf{B_1} \right) \cdot \nabla \right) \left(-q \Omega_0 x \hat{y} + \mathbf{u_1}\right) = \xi B_0 \frac{x_1}{x} \partial_y \mathbf{u_1}
+\eeq
+
+The second term becomes:
+
+\beq
+\xi B_0 u_{1x} \frac{x_1}{x^2} \hat{y}
+\eeq
+
 \bibliography{mri.bib}

### 2014-10-20 notes/mri_notes.tex
all hmri terms... where to go from here?

@@
 \beq
+\label{momeq_added}
 \xi^2 B_0^2 \frac{x_1^2}{x^3} \hat{x} \, - \,  \xi B_0 \frac{x_1}{x^2} B_{1x} \hat{y} \, + \, \xi B_0 \frac{x_1}{x^2} B_{1y} \hat{x} \, - \, \partial_z B_{1y} \xi B_0 \frac{x_1}{x} \hat{z} \, - \, \partial_x B_{1y} \xi B_0 \frac{x_1}{x} \hat{x}
@@
 \beq
-\partial_t \mathbf{B} = \left( \left( \mathbf{B_0} \hat{z} + \xi B_0 \frac{x_1}{x} \hat{y} + \mathbf{B_1} \right) \cdot \nabla \right) \left(-q \Omega_0 x \hat{y} + \mathbf{u_1}\right) = \xi B_0 \frac{x_1}{x} \partial_y \mathbf{u_1}
+\left( \left( \mathbf{B_0} \hat{z} + \xi B_0 \frac{x_1}{x} \hat{y} + \mathbf{B_1} \right) \cdot \nabla \right) \left(-q \Omega_0 x \hat{y} + \mathbf{u_1}\right) = \xi B_0 \frac{x_1}{x} \partial_y \mathbf{u_1}
 \eeq
@@
 \beq
-\xi B_0 u_{1x} \frac{x_1}{x^2} \hat{y}
+-\left( \left( -q \Omega_0 x \hat{y} + \mathbf{u_1}\right) \cdot \nabla\right) \left( B_0 \hat{z} + \xi B_0 \frac{x_1}{x} \hat{y} + \mathbf{B_1}\right)   \, = \, \xi B_0 u_{1x} \frac{x_1}{x^2} \hat{y}
 \eeq
 
+And the third term on the righthandside becomes
+
+\beq
+\frac{1}{\reym} \xi B_0 \left( \frac{2 x_1}{x^3}\right) \hat{y}
+\eeq
+
+So that all told, the induction equation gains the terms:
+
+\beq
+\xi B_0 \frac{x_1}{x} \partial_y \mathbf{u_1} \, + \, \xi B_0 u_{1x} \frac{x_1}{x^2} \hat{y} \, + \, \frac{1}{\reym} \xi B_0 \left( \frac{2 x_1}{x^3}\right) \hat{y}
+\eeq
+
+Because the $\partial_y$ terms go to zero, the induction equation only picks up terms in its $y$ component (the first term above goes to zero). Rewritten to use the stream function, these terms are:
+
+\beq
+\label{induc_eq_added}
+\xi B_0 \frac{x_1}{x^2} \partial_z \Psi \hat{y} + \frac{1}{\reym} \xi B_0 \frac{2 x_1}{x^3} \hat{y}
+\eeq
+
+We can now inspect Equation \ref{momeq_added} and see that the $\hat{y}$ component of the momentum equation will pick up a term analogous to the the first term of Equation \ref{induc_eq_added} -- a tantalizing $\partial_z$ term, which has the ability to lend a $\partial_Z$ term in the $\mathcal{O}\left(\epsilon^2\right)$ equation!
+But Equation \ref{momeq_added} can no longer be written in terms of flux/stream functions, because contributions to the $\hat{x}$ and $\hat{z}$ components are uneven.
+Should I rewrite all the equations in terms of primitive variables to proceed?
+
 \bibliography{mri.bib}

### 2014-10-20 python/multirun_paramspace.py

@@ -32,8 +32,12 @@ import pickle
 #Rmsearch = np.arange(200, 600, 10)
 
 #Pm = 0.0001, so Rm ~ 0.5
-Rmsearch = np.arange(0.2, 1.2, 0.1)
-Qsearch = np.arange(0.45, 0.55, 0.01) 
+#Rmsearch = np.arange(0.2, 1.2, 0.1)
+#Rmsearch = np.arange(3.2, 4.2, 0.1)
+#Qsearch = np.arange(0.45, 0.55, 0.01)
+Rmsearch = np.arange(3.5, 5.5, 0.1) 
+Qsearch = np.arange(0.2, 3, 0.2)
+#Qsearch = np.arange(0.15, 0.25, 0.01) 
 
 run_script = "multirun_linear_MRI.py"
 
@@ -61,7 +65,7 @@ for k,proc in processes.items():
 
 #print(results)
 
-pickle.dump(results, open("multirun_Rm_50.p", "wb"))
+pickle.dump(results, open("multirun_Rm_p5_big.p", "wb"))
 
 #np.save("results_Qrange"+str(Qsearch[0])+"_to_"+str(Qsearch[-1])+"Rrange"+str(Rmsearch[0])+"_to_"+str(Rmsearch[-1])+".npy", results)
 #np.save("results_test.npy", results)

### 2014-10-20 python/runsplottest.py

@@ -5,7 +5,7 @@ import pickle
 import matplotlib
 matplotlib.rcParams['backend'] = "Qt4Agg"
 
-testdata = pickle.load(open("multirun_Rm_50.p", "rb"))
+testdata = pickle.load(open("multirun_Rm_p5.p", "rb"))
 
 Rms = np.zeros(len(testdata))
 Qs = np.zeros(len(testdata))
@@ -22,7 +22,7 @@ eval = -1*eval
 
 fig = plt.figure()
 ax1 = fig.add_subplot(121)
-cb = ax1.scatter(Qs, Rms, c=np.real(eval), marker="s", s=40, vmin=-1E-7, vmax=1E-7, cmap="bone")
+cb = ax1.scatter(Qs[eval > -99], Rms[eval > -99], c=np.real(eval[eval > -99]), marker="s", s=40, cmap="bone")#, vmin=-1E-7, vmax=1E-7, cmap="bone")
 fig.colorbar(cb)
 ax1.set_title("Real")
 ax1.set_xlabel("Q")
@@ -36,4 +36,4 @@ ax2.set_xlabel("Q")
 ax2.set_ylabel("Rm")
 
 plt.show()
-pylab.savefig("multirun_Rm_p5_Pm04.png")
+pylab.savefig("multirun_Rm_p5_Pm04_7.png")

### 2014-10-21 notes/mri_notes.tex
resolution of asymmetry issue and derivation of all additional hMRI terms

@@
 
-We can now inspect Equation \ref{momeq_added} and see that the $\hat{y}$ component of the momentum equation will pick up a term analogous to the the first term of Equation \ref{induc_eq_added} -- a tantalizing $\partial_z$ term, which has the ability to lend a $\partial_Z$ term in the $\mathcal{O}\left(\epsilon^2\right)$ equation!
+%We can now inspect Equation \ref{momeq_added} and see that the $\hat{y}$ component of the momentum equation will pick up a term analogous to the the first term of Equation \ref{induc_eq_added} -- a tantalizing $\partial_z$ term, which has the ability to lend a $\partial_Z$ term in the $\mathcal{O}\left(\epsilon^2\right)$ equation!
 But Equation \ref{momeq_added} can no longer be written in terms of flux/stream functions, because contributions to the $\hat{x}$ and $\hat{z}$ components are uneven.
@@
 
+For the momentum equation, we now take $\partial_z$ of the x component and $\partial_x$ of the z component, to obtain, for each component:
+
+\beq
+[x]: \xi B_0 \frac{x_0}{x^2} \partial_z B_{1y} - \partial_x\partial_z B_{1y} \xi B_0 \frac{x_0}{x}
+\eeq
+
+\beq
+[y]: - \xi B_0 \frac{x_0}{x^2} B_{1x}
+\eeq
+
+\beq
+[z]: \xi B_0 \frac{x_0}{x^2} \partial_z B_{1y} - \partial_x\partial_z B_{1y} \xi B_0 \frac{x_0}{x}
+\eeq
+
+Rewritten in terms of flux functions:
+
+\beq
+[x]: \xi B_0 \frac{x_0}{x^2} \partial_z B_{y} - \partial_x\partial_z B_{y} \xi B_0 \frac{x_0}{x}
+\eeq
+
+\beq
+[y]: - \xi B_0 \frac{x_0}{x^2} \partial_z A
+\eeq
+
+\beq
+[z]: \xi B_0 \frac{x_0}{x^2} \partial_z B_{y} - \partial_x\partial_z B_{y} \xi B_0 \frac{x_0}{x}
+\eeq
+
+This means that the total added contributions to the final equation set, in $\left[\Psi, u_y, A, B_y\right]^T$ order, are:
+
+\beq
+[\Psi]: \xi B_0 \frac{x_0}{x^2} \partial_z B_{y} - \xi B_0 \partial_x\partial_z B_{y} \frac{x_0}{x}
+\eeq
+
+\beq
+[u_y]: - \xi B_0 \frac{x_0}{x^2} \partial_z A
+\eeq
+
+\beq
+[A]: 0
+\eeq
+
+\beq
+[B]: \xi B_0 \frac{x_1}{x^2} \partial_z \Psi + \frac{1}{\reym} \xi B_0 \frac{2 x_1}{x^3}
+\eeq
+
+Two exciting things are happening. We have some first-order $\partial_z$ terms, and some linear coupling between the velocity and magnetic field equations.
+
 \bibliography{mri.bib}

### 2014-10-21 python/allorders.py

@@ -1140,7 +1140,7 @@ class N3():
                 self.v11_A = v1.LEV.state['A']
                 self.v11_B = v1.LEV.state['B']
             
-            o2 = OrderE2()
+            o2 = OrderE2(Q = self.Q, Rm = self.Rm, Pm = self.Pm, q = self.q, beta = self.beta)
             o2.solve20()
             o2.solve21(norm = norm)
             o2.solve22()
@@ -1481,6 +1481,8 @@ class AmplitudeAlpha():
         self.q = q
         self.beta = beta
         
+        print("amplitude alpha params: ", self.Q, self.Rm, self.Pm, self.q, self.beta)
+        
         # inverse magnetic reynolds number
         self.iRm = 1./self.Rm
 
@@ -1493,7 +1495,7 @@ class AmplitudeAlpha():
             self.va = AdjointHomogenous()
             self.va.solve(save = False, norm = True)
         
-            v1 = OrderE()
+            v1 = OrderE(Q = self.Q, Rm = self.Rm, Pm = self.Pm, q = self.q, beta = self.beta)
             v1.solve(save=False)
             
             if norm == True:
@@ -1516,7 +1518,7 @@ class AmplitudeAlpha():
                 self.v11_A = v1.LEV.state['A']
                 self.v11_B = v1.LEV.state['B']
             
-            self.o2 = OrderE2(speedy = True, o1 = v1)
+            self.o2 = OrderE2(Q = self.Q, Rm = self.Rm, Pm = self.Pm, q = self.q, beta = self.beta, speedy = True, o1 = v1)
             self.o2.solve20()
             self.o2.solve21(norm = norm)
             self.o2.solve22()
@@ -1549,10 +1551,10 @@ class AmplitudeAlpha():
             self.v22_A = self.o2.LEV22.state['A22']
             self.v22_B = self.o2.LEV22.state['B22']
             
-            self.o3 = OrderE3(speedy = True, o2 = self.o2)
+            self.o3 = OrderE3(Q = self.Q, Rm = self.Rm, Pm = self.Pm, q = self.q, beta = self.beta, speedy = True, o2 = self.o2)
             self.o3.solve()
             
-            self.n3 = N3(speedy = True, o1 = v1, o2 = self.o2)
[…]

### 2014-11-15 notes/mri_notes.tex
x1 -> x0
hMRI matrix form
tuning and multiscale hmri
hmri perturbation series expansion
more equations
hmri additions if the azimuthal term is constant in cartesian coordinates

@@
 \begin{equation}
-\mathbf{B} = B_0 \hat{z} + \xi B_0 \left(\frac{x_1}{x}\right) \hat{y} + \mathbf{B}_1
+\mathbf{B} = B_0 \hat{z} + \xi B_0 \left(\frac{x_0}{x}\right) \hat{y} + \mathbf{B}_1
 \end{equation}
@@
 \beq
-\left(\nabla \times \left(B_0 \hat{z} + \xi B_0 \left(\frac{x_1}{x}\right) \hat{y} + \mathbf{B}_1
-\right) \right) \times \left(B_0 \hat{z} + \xi B_0 \left(\frac{x_1}{x}\right) \hat{y} + \mathbf{B}_1
+\left(\nabla \times \left(B_0 \hat{z} + \xi B_0 \left(\frac{x_0}{x}\right) \hat{y} + \mathbf{B}_1
+\right) \right) \times \left(B_0 \hat{z} + \xi B_0 \left(\frac{x_0}{x}\right) \hat{y} + \mathbf{B}_1
 \right)
@@
 \beq
-\left(\nabla \times B_0 \hat{z} \right) \times \xi B_0 \left(\frac{x_1}{x}\right)\hat{y} = 0
+\left(\nabla \times B_0 \hat{z} \right) \times \xi B_0 \left(\frac{x_0}{x}\right)\hat{y} = 0
 \eeq
@@
 \beq
-\left(\nabla \times \xi B_0 \left(\frac{x_1}{x}\right) \hat{y} \right) \times B_0 \hat{z} = \left(\partial_x \left(\xi B_0 \left(\frac{x_1}{x}\right)\right)\hat{z}\right) \times B_0 \hat{z} = 0
+\left(\nabla \times \xi B_0 \left(\frac{x_0}{x}\right) \hat{y} \right) \times B_0 \hat{z} = \left(\partial_x \left(\xi B_0 \left(\frac{x_0}{x}\right)\right)\hat{z}\right) \times B_0 \hat{z} = 0
 \eeq
@@
 \beq
-\left(\nabla \times \xi B_0 \frac{x_1}{x} \hat{y} \right) \times \left( \xi B_0 \frac{x_1}{x}\hat{y}\right) = \xi^2 B_0^2 \frac{x_1^2}{x^3} \hat{x}
+\left(\nabla \times \xi B_0 \frac{x_0}{x} \hat{y} \right) \times \left( \xi B_0 \frac{x_0}{x}\hat{y}\right) = \xi^2 B_0^2 \frac{x_0^2}{x^3} \hat{x}
 \eeq
@@
 \beq
-\left(\nabla \times \xi B_0 \frac{x_1}{x} \hat{y}\right) \times \mathbf{B_1} = -\xi B_0 \frac{x_1}{x^2} \left(B_{1x} \hat{y} - B_{1y} \hat{x}\right)
+\left(\nabla \times \xi B_0 \frac{x_0}{x} \hat{y}\right) \times \mathbf{B_1} = -\xi B_0 \frac{x_0}{x^2} \left(B_{1x} \hat{y} - B_{1y} \hat{x}\right)
 \eeq
@@
 \beq
-\left(\nabla \times \mathbf{B}_1 \right) \times \left(\xi B_0 \frac{x_1}{x} \hat{y}\right) = - \partial_z B_{1y} \xi B_0 \frac{x_1}{x} \hat{z} - \partial_x B_{1y} \xi B_0 \frac{x_1}{x} \hat{x}
+\left(\nabla \times \mathbf{B}_1 \right) \times \left(\xi B_0 \frac{x_0}{x} \hat{y}\right) = - \partial_z B_{1y} \xi B_0 \frac{x_0}{x} \hat{z} - \partial_x B_{1y} \xi B_0 \frac{x_0}{x} \hat{x}
 \eeq
@@
 \label{momeq_added}
-\xi^2 B_0^2 \frac{x_1^2}{x^3} \hat{x} \, - \,  \xi B_0 \frac{x_1}{x^2} B_{1x} \hat{y} \, + \, \xi B_0 \frac{x_1}{x^2} B_{1y} \hat{x} \, - \, \partial_z B_{1y} \xi B_0 \frac{x_1}{x} \hat{z} \, - \, \partial_x B_{1y} \xi B_0 \frac{x_1}{x} \hat{x}
+\xi^2 B_0^2 \frac{x_0^2}{x^3} \hat{x} \, - \,  \xi B_0 \frac{x_0}{x^2} B_{1x} \hat{y} \, + \, \xi B_0 \frac{x_0}{x^2} B_{1y} \hat{x} \, - \, \partial_z B_{1y} \xi B_0 \frac{x_0}{x} \hat{z} \, - \, \partial_x B_{1y} \xi B_0 \frac{x_0}{x} \hat{x}
 \eeq
@@
 \beq
-\left( \left( \mathbf{B_0} \hat{z} + \xi B_0 \frac{x_1}{x} \hat{y} + \mathbf{B_1} \right) \cdot \nabla \right) \left(-q \Omega_0 x \hat{y} + \mathbf{u_1}\right) = \xi B_0 \frac{x_1}{x} \partial_y \mathbf{u_1}
+\left( \left( \mathbf{B_0} \hat{z} + \xi B_0 \frac{x_0}{x} \hat{y} + \mathbf{B_1} \right) \cdot \nabla \right) \left(-q \Omega_0 x \hat{y} + \mathbf{u_1}\right) = \xi B_0 \frac{x_0}{x} \partial_y \mathbf{u_1}
 \eeq
@@
 \beq
--\left( \left( -q \Omega_0 x \hat{y} + \mathbf{u_1}\right) \cdot \nabla\right) \left( B_0 \hat{z} + \xi B_0 \frac{x_1}{x} \hat{y} + \mathbf{B_1}\right)   \, = \, \xi B_0 u_{1x} \frac{x_1}{x^2} \hat{y}
+-\left( \left( -q \Omega_0 x \hat{y} + \mathbf{u_1}\right) \cdot \nabla\right) \left( B_0 \hat{z} + \xi B_0 \frac{x_0}{x} \hat{y} + \mathbf{B_1}\right)   \, = \, \xi B_0 u_{1x} \frac{x_0}{x^2} \hat{y}
 \eeq
@@
 \beq
-\frac{1}{\reym} \xi B_0 \left( \frac{2 x_1}{x^3}\right) \hat{y}
+\frac{1}{\reym} \xi B_0 \left( \frac{2 x_0}{x^3}\right) \hat{y}
 \eeq
@@
 \beq
-\xi B_0 \frac{x_1}{x} \partial_y \mathbf{u_1} \, + \, \xi B_0 u_{1x} \frac{x_1}{x^2} \hat{y} \, + \, \frac{1}{\reym} \xi B_0 \left( \frac{2 x_1}{x^3}\right) \hat{y}
+\xi B_0 \frac{x_0}{x} \partial_y \mathbf{u_1} \, + \, \xi B_0 u_{1x} \frac{x_0}{x^2} \hat{y} \, + \, \frac{1}{\reym} \xi B_0 \left( \frac{2 x_0}{x^3}\right) \hat{y}
 \eeq
@@
 \label{induc_eq_added}
-\xi B_0 \frac{x_1}{x^2} \partial_z \Psi \hat{y} + \frac{1}{\reym} \xi B_0 \frac{2 x_1}{x^3} \hat{y}
+\xi B_0 \frac{x_0}{x^2} \partial_z \Psi \hat{y} + \frac{1}{\reym} \xi B_0 \frac{2 x_0}{x^3} \hat{y}
 \eeq
@@
 \beq
-[B]: \xi B_0 \frac{x_1}{x^2} \partial_z \Psi + \frac{1}{\reym} \xi B_0 \frac{2 x_1}{x^3}
+[B]: \xi B_0 \frac{x_0}{x^2} \partial_z \Psi + \frac{1}{\reym} \xi B_0 \frac{2 x_0}{x^3}
 \eeq
@@
 
+We now rewrite these terms in matrix form. We will have a matrix for the variable terms and a constant vector.
+
+\beq
+\mathcal{H} = \left[ \begin{matrix}  0 & 0 & 0 & x_0\left(\frac{1}{x^2} - \frac{1}{x}\partial_x\right)\\
+0 & 0 & -\frac{x_0}{x^2} & 0 \\ 
+0 & 0 & 0 & 0 \\
+\frac{x_0}{x^2} & 0 & 0 & 0 \\
+\end{matrix} \right] \\
+\eeq
+
+and
+
+\beq
+\mathbf{C} = \left[\begin{matrix} 0 \\
+0 \\
+0 \\
+\frac{1}{\reym} \frac{2 x_0}{x^3} \\
+\end{matrix} \right] \\
+\eeq
+
+And our additional terms become:
+
+\beq
+\xi B_0 \mathcal{H} \partial_z \mathbf{V} + \xi B_0 \mathbf{C}
+\eeq
+
+Destabilizing substitution $B_0 = 1 - \epsilon^2$:
+
+\beq
+\xi \left(1 - \epsilon^2\right) \mathcal{H} \partial_z \mathbf{V} + \xi \left(1 - \epsilon^2\right) \mathbf{C}
+\eeq
+
+\beq
+= \xi \mathcal{H} \partial_z \mathbf{V}  - \epsilon^2 \xi \mathcal{H} \partial_z \mathbf{V} + \xi \mathbf{C} - \xi \epsilon^2 \mathbf{C}
+\eeq
+
+Multiscale expansion ($\partial_z \rightarrow \partial_z + \epsilon \partial_Z$):
+
+\beq
+= \xi \mathcal{H} \partial_z \mathbf{V} + \epsilon \xi \mathcal{H} \partial_Z \mathbf{V} - \epsilon^2 \xi \mathcal{H} \partial_z \mathbf{V} - \epsilon^3 \xi \mathcal{H} \partial_Z \mathbf{V} + \xi \mathbf{C} - \xi \epsilon^2 \mathbf{C}
+\eeq
+
+And finally, the perturbation series expansion $\mathbf{V} = \epsilon \mathbf{V_1} + \epsilon^2 \mathbf{V_2} + \epsilon^3 \mathbf{V_3}$. I include terms up to $\epsilon^3$:
+
+\beq
+\epsilon \xi \mathcal{H} \partial_z \mathbf{V_1} + \epsilon^2 \xi \mathcal{H} \partial_z \mathbf{V_2} + \epsilon^3 \xi \mathcal{H} \partial_z \mathbf{V_3} + \epsilon^2 \xi \mathcal{H} \partial_Z \mathbf{V_1} + \epsilon^3 \xi \mathcal{H} \partial_Z \mathbf{V_2} - \epsilon^3 \xi \mathcal{H} \partial_z \mathbf{V_1} + \xi \mathbf{C} - \xi \epsilon^2 \mathbf{C}
+\eeq
+
+Separated by order:
+
+$\mathcal{O}(0)$:
+
+\beq
+\xi \mathbf{C}
+\eeq
+
+$\mathcal{O}(\epsilon)$:
+
+\beq
+\epsilon \xi \mathcal{H} \partial_z \mathbf{V_1}
+\eeq
+
+$\mathcal{O}(\epsilon^2)$:
+
+\beq
+\epsilon^2 \xi \mathcal{H} \partial_z \mathbf{V_2} + \epsilon^2 \xi \mathcal{H} \partial_Z \mathbf{V_1} - \xi \epsilon^2 \mathbf{C}
+\eeq
+
+$\mathcal{O}(\epsilon^3)$:
+
+\beq
+\epsilon^3 \xi \mathcal{H} \partial_z \mathbf{V_3} + \epsilon^3 \xi \mathcal{H} \partial_Z \mathbf{V_2} - \epsilon^3 \xi \mathcal{H} \partial_z \mathbf{V_1}
+\eeq
+
+Meaning that the equations for the hMRI become: \\
+
+$\mathcal{O}(\epsilon)$:
+
+\beq
+\mathcal{L}\mathbf{V_1} + \xi \mathcal{H} \partial_z \mathbf{V_1} = 0
+\eeq
+
+$\mathcal{O}(\epsilon^2)$:
+
+\beq
+\mathcal{L} \mathbf{V_2} = \mathbf{N_2} - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1} + \xi \mathcal{H} \partial_z \mathbf{V_2} + \xi \mathcal{H} \partial_Z \mathbf{V_1} - \xi \mathbf{C}
+\eeq
+
+$\mathcal{O}(\epsilon^3)$:
+
+\beq
+\mathcal{D}\partial_T \mathbf{V_1} + \mathbf{N_3} = \mathcal{L} \mathbf{V_3} + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_2} + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V_1} - \partial_z \mathcal{X} \mathbf{V_1} - \partial_z^3 \mathcal{L}_3 \mathbf{V_1} +  \xi \mathcal{H} \partial_z \mathbf{V_3} + \xi \mathcal{H} \partial_Z \mathbf{V_2} -  \xi \mathcal{H} \partial_z \mathbf{V_1}
+\eeq
+
+If instead we Taylor expand the azimuthal field term to obtain a helical MRI background of:
+
+\beq
+\mathbf{B} = B_0 \hat{z} + \xi B_0 \hat{y} + \mathbf{B_1},
+\eeq
+
+we don't gain nearly as many terms with the hMRI. The momentum equation picks up only
+
+\beq
+-\xi B_0 \partial_z B_{1y} \hat{z} - \xi B_0 \partial_x B_{1y} \hat{x},
+\eeq
+
+and the induction equation picks up nothing if the perturbations are axisymmetric (the only term gained is $\xi B_0 \partial_y \mathbf{u_1}$, and $\partial_y \rightarrow 0$).
+
+Taking $\partial_z$ of the [x] momentum equation and $\partial_x$ of the [z] momentum equation, the momentum equation hMRI addition becomes
+
+\beq
+-\xi B_0 \partial_x \partial_z B_{1y} \hat{z} - \xi B_0 \partial_x \partial_z B_{1y} \hat{x}.
+\eeq
+
 \bibliography{mri.bib}

### 2014-11-21 paper/mri_paper.tex
defined terms
typos and define gas pressure
fixed typo in streamfunction description

--- 2014-09-24 paper/mri_paper.tex
@@
 
+In the above, $P$ is the gas pressure, $\nu$ is the kinematic viscosity, $\eta$ is the microscopic diffusivity, $\nabla\Phi$ is the gravitational force per unit mass, and the current density is
+
+\beq
+\mathbf{J} = \frac{2}{\beta}\nabla\times\mathbf{B}.
+\eeq
+
 Equations \ref{momentum} and \ref{induction} are subject to the magnetic solenoid and incompressibility constraints,
@@
 
+We also define a plasma beta parameter,
+
+\beq
+\beta \equiv \frac{\Omega_0^2r_0^2}{v_A^2},
+\eeq
+
+where the Alfv\'en sped $v_A$ is
+
+\beq
+v_A^2 = \frac{B_0^2}{\rho_0}.
+\eeq
+
 The fluid symbols $\mathbf{u}$, $\mathbf{B}$, etc. will henceforth be used to refer to the nondimensional, perturbed quantities.
@@
 $A$ is the familiar two-dimensional vector potential.
-$A$ and $\Psi$ are scalar fields defined as the curl of the magnetic field and velocity, respectively, and so automatically satisfy our constraints.
+$A$ and $\Psi$ are scalar fields.
+The curl of $A$ and the curl of $\Psi$ are defined as the magnetic field and velocity, respectively, and so $A$ and $\Psi$ automatically satisfy our constraints.
 

### 2014-11-21 paper/mri_paper.tex
typos

@@
 \mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, \\
-- \partial_Z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1}.
+- \partial_z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1}.
 \end{multline}
@@
 \begin{split}
-\mathbf{V_2} \, & = \, \alpha^2 e^{i2Qz} \mathbb{V}_{22}(x) \, + \, \left|\alpha\right|^2 \mathbb{V}_{20}(x) \, \\
-&- \, \partial_Z \alpha \mathbb{V}_{21}(x) e^{iQz} \,- \, \partial_Z \beta \mathbb{U}_{20}(x) + c.c.
+\mathbf{V_2} \, & = \, \alpha^2 e^{i2 k_c z} \mathbb{V}_{22}(x) \, + \, \left|\alpha\right|^2 \mathbb{V}_{20}(x) \, \\
+&- \, \partial_Z \alpha \mathbb{V}_{21}(x) e^{i k_c z} \,- \, \partial_Z \beta \mathbb{U}_{20}(x) + c.c.
 \end{split}
@@
 \begin{split}
-& \mathcal{D} \mathbb{V}_{11} e^{iQz}\partial_T \alpha \, + \, \alpha \left| \alpha^2 \right| \mathbf{N}_{31} e^{iQz} \, + \, \alpha \partial_Z \beta \widetilde{\mathbf{N}}_{31} e^{iQz} \, =  \\
-& \, - \partial_Z^2 \alpha \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} e^{iQz} \, + \, \partial_Z^2 \alpha \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{iQz} \, - \, \partial_Z \alpha \mathbb{V}_{11}\mathcal{X}e^{iQz} \, \\
-& + \, iQ^3 \alpha \mathcal{L}_3 \mathbb{V}_{11} e^{iQz}.
+& \mathcal{D} \mathbb{V}_{11} e^{i k_c z}\partial_T \alpha \, + \, \alpha \left| \alpha^2 \right| \mathbf{N}_{31} e^{i k_c z} \, + \, \alpha \partial_Z \beta \widetilde{\mathbf{N}}_{31} e^{i k_c z} \, =  \\
+& \, - \partial_Z^2 \alpha \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} e^{i k_c z} \, + \, \partial_Z^2 \alpha \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{i k_c z} \, - \, \partial_Z \alpha \mathbb{V}_{11}\mathcal{X}e^{i k_c z} \, \\
+& + \, i k_c^3 \alpha \mathcal{L}_3 \mathbb{V}_{11} e^{i k_c z}.
 \end{split}

### 2014-11-28 python/MRI_unpack.py
N3 plots

@@ -21,7 +21,7 @@ beta = 25.0
 coeffs = {}
 
 #Pm = 1E-7: Q = 0.745, Rm = 4.90
-gridnum = 512
+gridnum = 1024#512
 Pm = 1.0E-7
 Q = 0.745
 Rm = 4.90

### 2014-11-28 python/allorders.py

@@ -18,7 +18,7 @@ from matplotlib import rc
 rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
 rc('text', usetex=True)
 
-gridnum = 1024#512
+gridnum = 16#512
 x_basis = Chebyshev(gridnum)
 domain = Domain([x_basis], grid_dtype=np.complex128)
 
@@ -1665,19 +1665,24 @@ class AmplitudeAlpha():
         
         # c = <va . N31*>
         c_psi = self.va.psi*self.N31_psi_star
-        c_psi = c_psi.evaluate()
+        c_psi = c_psi.evaluate() #testing
+        self.c_psi = c_psi
         
         c_u = self.va.u*self.N31_u_star
         c_u = c_u.evaluate()
+        self.c_u = c_u
         
         c_A = self.va.A*self.N31_A_star
         c_A = c_A.evaluate()
+        self.c_A = c_A
         
         c_B = self.va.B*self.N31_B_star # N31 B_star and u_star are nan
         c_B = c_B.evaluate()
+        self.c_B = c_B
         
         call = c_psi + c_u + c_A + c_B
         call = call.evaluate()
+        self.call = call
         
         c = domain.new_field()
         c.name = 'c'

### 2014-12-15 paper/mri_paper.tex
hmri description

@@
 \beq
+\label{backgroundfield1}
 \mathbf{B} = B_0 \hat{z},
@@
 The scale separations caused by these large ratios of diffusivity to viscosity render the small Prandtl number regime inaccessible to simulations.
-This analytic formulation allows us to explore the behavior down to magnetic Prandtl numbers of [TBD!].
+This analytic formulation allows us to explore the behavior down to magnetic Prandtl numbers of $10^{-6}$ or lower.
+
+\section{Helical MRI}
+
+The configuration investigated thus far, that of magnetized Taylor Couette flow with a purely axial magnetic field, constitutes the standard MRI.
+The helical MRI (HMRI) is a related, but separate, MHD instability, first identified by Hollerbach and R\"udiger (\citeyear{Hollerbach:2005tr}).
+It arises in the presence of combined axial and azimuthal fields.
+
+We initialize the helical MRI by adding a constant background toroidal field to Equation \ref{backgroundfield1}, such that our initial magnetic field structure becomes
+
+\beq
+\mathbf{B} = B_0 \hat{z} + \xi B_0 \hat{y},
+\eeq
+
+where $\xi$ is a dimensionless parameter representing the relative strength of the axial field and the vertical field.
+
+This addition leaves Equations \ref{eqset2} - \ref{eqset4} unchanged, but Equation \ref{eqset1} gains a term to become
+
+\begin{multline}
+\label{eqset1_hmri}
+\partial_t \nabla^2 \Psi \, = \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, 2 \partial_z u_{y} \, - \xi B_0 \partial_x \partial_z B_{y}\\
+- \, J\left(\Psi, \nabla^2 \Psi\right) \, + \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi,
+\end{multline}
+
+in other words, the $\Psi$ equation picks up a linear $B_y$ term.
+
+Following the same prescription for weakly nonlinear tuning, multiscale expansion, and perturbation expansion, we find the equations for the HMRI up to order $epsilon^3$.
+
+To $\mathcal{O}(\epsilon)$, we find
+
+\mathcal{L} \mathbf{V_1}
 
@@
 
-The present configuration, that of magnetized Taylor Couette flow with a purely axial magnetic field, represents the initial conditions for the standard MRI.
+%The present configuration, that of magnetized Taylor Couette flow with a purely axial magnetic field, represents the initial conditions for the standard MRI.
 The helical MRI is a related, but separate, MHD instability, and was first discovered by Hollerbach and R\"udiger (\citeyear{Hollerbach:2005tr}).

### 2014-12-15 python/MRI_unpack.py

@@ -188,23 +188,25 @@ def plot5():
     cy2 = coeffs[0]["u_y second order"]
     x = coeffs[0]["x"]    
     jj = np.zeros(len(x))
-        
-    ax1.plot(x, 0.5*cy1[25, :].real, color="purple")
-    ax1.plot(x, 0.5**2*cy2[25, :].real, color="orange")
+    
+    ax1.plot(x, jj, color="grey", ls='--')
+    ax1.plot(x, 0.5*cy1[25, :].real, color="purple", lw=2)
+    ax1.plot(x, 0.5**2*cy2[25, :].real, color="orange", lw=2)
     ax1.plot(x, 0.5*cy1[25, :].real + 0.5**2*cy2[25, :].real, color="black")
-    ax1.plot(x, jj, color="grey")
     
     ax2 = fig.add_subplot(312)
-    ax2.plot(x, 0.5*cy1[128, :].real, color="purple")
-    ax2.plot(x, 0.5**2*cy2[128, :].real, color="orange")
+    ax2.plot(x, jj, color="grey", ls='--')
+    ax2.plot(x, 0.5*cy1[128, :].real, color="purple", lw=2)
+    ax2.plot(x, 0.5**2*cy2[128, :].real, color="orange", lw=2)
     ax2.plot(x, 0.5*cy1[128, :].real + 0.5**2*cy2[128, :].real, color="black")
-    ax2.plot(x, jj, color="grey")
+
     
     ax3 = fig.add_subplot(313)
-    ax3.plot(x, 0.5*cy1[231, :].real, color="purple")
-    ax3.plot(x, 0.5**2*cy2[231, :].real, color="orange")
+    ax3.plot(x, jj, color="grey", ls='--')
+    ax3.plot(x, 0.5*cy1[231, :].real, color="purple", lw=2)
+    ax3.plot(x, 0.5**2*cy2[231, :].real, color="orange", lw=2)
     ax3.plot(x, 0.5*cy1[231, :].real + 0.5**2*cy2[231, :].real, color="black")
-    ax3.plot(x, jj, color="grey")
+    
 
 
 for i in range(num):
@@ -259,9 +261,9 @@ for i in range(num):
         
         plt.tick_params(labelsize = 20)
         rr = np.zeros(len(x))
-        plt.plot(x, rr+z[25], color="purple")
-        plt.plot(x, rr+z[128], color="purple")
-        plt.plot(x, rr+z[231], color="purple")
[…]

### 2014-12-15 paper/mri_paper.tex
hmri equations

@@
 
-Following the same prescription for weakly nonlinear tuning, multiscale expansion, and perturbation expansion, we find the equations for the HMRI up to order $epsilon^3$.
+Following the same prescription for weakly nonlinear tuning, multiscale expansion, and perturbation expansion, we find the equations for the HMRI up to order $\epsilon^3$.
 
@@
 
-\mathcal{L} \mathbf{V_1}
+\beq
+\mathcal{L} \mathbf{V_1} + \xi B_0 \mathcal{H} \partial_z \mathbf{V_1} = 0
+\eeq
+
+To $\mathcal{O}(\epsilon^2)$,
+
+\beq
+\mathcal{L}\mathbf{V_2} = \mathbf{N_2} - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1} + \xi \mathcal{H} \partial_z \mathbf{V_2} - \xi \mathcal{H}\partial_Z \mathbf{V_1}
+\eeq
+
+To $\mathcal{O}(\epsilon^3)$,
+
+\begin{multline}
+\mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, \\
+- \partial_z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1} + \xi \mathcal{H} \partial_z \mathbf{V_3} - \xi \mathcal{H} \partial_Z \mathbf{V_2} + \xi \mathcal{H} \partial_z \mathbf{V_1}.
+\end{multline}
 

### 2014-12-15 paper/mri_paper.tex
amplitude func 1 hmri

@@
 The scale separations caused by these large ratios of diffusivity to viscosity render the small Prandtl number regime inaccessible to simulations.
-This analytic formulation allows us to explore the behavior down to magnetic Prandtl numbers of $10^{-6}$ or lower.
+Because we solve spectrally in only one dimension, rather than in two or three, our analytic formulation allows us to explore the behavior down to magnetic Prandtl numbers of $10^{-6}$ or lower.
 
@@
 \end{multline}
+
+In the above,
+
+\beq
+\mathcal{H} = \left[\begin{matrix}
+0 & 0 & 0 & \partial_x \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\ \end{matrix} \right].
+\eeq
+
+The first amplitude equation (\ref{ampeq1}) therefore becomes
+
+\begin{multline}
+\label{ampeq1_hmri}
+a \partial_T \alpha \, = - \, c \alpha \left|\alpha^2\right| \, - \, \widetilde{c} \alpha \partial_Z \beta \,  - b i k_c \alpha \, + \, h \partial_Z^2 \alpha \, \\
++ \, g i k_c^3 \alpha \, + \, q \xi \partial_Z^2 \alpha \, + \, w i k_c \xi \alpha,
+\end{multline}
+
+where
+
+\beq
+q = \left< \mathbb{V}^\dagger \cdot \left(\mathcal{H} \mathbb{V}_{21}\right)^* \right>,
+\eeq
+
+and
+
+\beq
+w = \left< \mathbb{V}^\dagger \cdot \left(\mathcal{H} \mathbb{V}_{11}\right)^*\right>.
+\eeq
 

### 2014-12-17 paper/mri_paper.tex
hmri amplitude eqn 2
more paper background

@@
 Since its discovery by \citeauthor{Chandrasekhar:1960wh} (\citeyear{Chandrasekhar:1960wh}) and subsequent rediscovery by \citeauthor{Balbus:1991vs} (\citeyear{Balbus:1991vs}), the MRI remains the leading explanation for rapid angular momentum transport in astrophysical disks.
-The instability arises when a differentially rotating disk is threaded by a vertical magnetic field.
+The instability in its simplest geometry arises when a differentially rotating disk is threaded by a vertical magnetic field.
 The presence of the magnetic field destabilizes the disk gas, driving turbulence and angular momentum transport.
@@
 The nonlinear evolution of the MRI is a complex process, and an understanding of the mechanism by which the MRI saturates remains elusive.
+Several possible saturation avenues are available to the MRI.
+The instability could saturate via a reduction in the background shear (e.g.
+\citei{Umurhan:2007hs}), cutting off the supply of free energy on which the instability feeds.
+Similarly, the instability could pour its free energy into the magnetic field, and saturate via field generation (e.g.
+\citei{Ebrahimi:2009ey}).
+Saturation may also occur via ``parasite" instabilities, which feed off and destroy primary MRI modes (\citei{Goodman:1994ul}).
+
 Most previous works on MRI saturation have studied the problem in simulation.
@@
 
+\section{Simulations}
+
+We compute the solutions to our derived equations pseudospectrally, using Dedalus.
+One advantage to using Dedalus for these computations is that the code can compute the evolution of the fully nonlinear MRI system on the same Fourier/Chebyshev grid.
+This allows us to directly compare simulation and theory.
+
+% Jeff's sims!!!!!
+
 \section{Helical MRI}
@@
 
+The second amplitude equation (\ref{ampeq2}) becomes
+
+\beq\label{ampeq2_hmri}
+\partial_T \beta = - m \partial_Z^2 \beta + \frac{1}{\reym} \partial_Z^2 \beta \, - \, s \xi 2 \left|\alpha\right|\partial_Z \alpha \, - \, \xi\partial_Z^2 \beta,
+\eeq
+
+where
+
+\beq
+s = \left<\mathbb{U}_{11} \cdot \left(\mathcal{H}\mathbb{V}_{20}\right)^*\right>
+\eeq
+
+and
+
+\beq
+f = \left< \mathbb{U}_{11} \cdot \left(\mathcal{H}\mathbb{U}_{20}\right)^*\right>.
+\eeq
+
+If $s = 0$, as it is for the standard MRI, the two amplitude equations will again decouple.
+
 \section{Discussion}\label{discussion}
@@
 Future work will address specifically our predictions for the PPPL MRI experiment.
-
-We compute the solutions to our derived equations pseudospectrally, using Dedalus.
-One advantage to using Dedalus for these computations is that the code can compute the evolution of the fully nonlinear MRI system on the same Fourier/Chebyshev grid.
-This allows us to directly compare simulation and theory.
-
-% Jeff's sims!!!!!
 
@@
 
+\bibitem[{Ebrahimi et al.(2009)}]{Ebrahimi:2009ey}
+Ebrahimi, F., Prager, S.C., Schnack, D.D
+
 \bibitem[{Gissinger et al.(2011)}]{Gissinger:2011td}

### 2014-12-17 paper/mri_paper.tex
more intro

@@
 They find that this saturation proceeds via modification of the background shear.
-Another notable analytic work is that of \citeauthor{Pessah:2010ic} (\citeyear{Pessah:2010ic}), who presents an analysis of MRI saturation via secondary instabilities.
-This work suggests that MRI saturation can be caused by parasitic instabilities growing off of the primary MRI modes, in particular by Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime.
+Another notable analytic work is that of \citeauthor{Pessah:2010ic} (\citeyear{Pessah:2010ic}), who presents an analysis of MRI saturation via the above mentioned secondary instabilities.
+This work suggests that MRI saturation can be caused by parasitic Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime.
 
@@
 This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime.
-The present work adds to this foundation both in subject, with our additional analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space.
+The present work adds to this foundation both in subject, with our additional analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space and the comparison of weakly nonlinear analytic work to strongly nonlinear simulation.
 In particular, we attempt to elucidate the role of boundary layers in the saturation of the MRI, as this is of fundamental importance to laboratory MRI work.

### 2014-12-17 paper/mri_paper.tex
rm fig 2
new fig

@@
 
-\begin{figure}[h!]
-\centering
-\includegraphics[trim=6cm 3cm 3cm 0cm, scale=0.35]{slowfast.eps}
-\caption{Cartoon illustrating fast and slow-moving variables (Equation \ref{scalings}).}\label{allorders_velocity}
-\label{figscalings}
-\end{figure}
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[trim=6cm 3cm 3cm 0cm, scale=0.35]{slowfast.eps}
+%\caption{Cartoon illustrating fast and slow-moving variables (Equation \ref{scalings}).}\label{allorders_velocity}
+%\label{figscalings}
+%\end{figure}
 
@@
 
-See Figure \ref{figscalings}.
+%See Figure \ref{figscalings}. 
 Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic Taylor Couette flow.
@@
 By contrast, the second-order magnetic field perturbations are two orders of magnitude smaller than the first-order perturbations (Figure \ref{allorders_Bfield}).
-It appears that the second-order velocity perturbations are instrumental in the saturation of the MRI.
-The likely mechanism is a reduction in the background shear.
+
+The role of the boundary layers changes with Pm.
+As Pm decreases, the width of the boundary layer becomes increasingly narrow (Figure \ref{u_secondorder_by_Pm}).
+The relative importance of the $u_y$ second order perturbation to the first order perturbation increases as well (CHECK, PLOT).
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[trim=6cm 0cm 6cm 0cm, scale=0.4]{u_secondorder_by_Pm_2.eps}
+\caption{(Placeholder) The evolution of the second-order velocity perturbations by Pm.}\label{allorders_velocity}
+\label{u_secondorder_by_Pm}
+\end{figure*}
+
+It appears that the second-order velocity perturbations are instrumental in the saturation of the MRI. The likely mechanism is a reduction in the background shear.
 

### 2014-12-17 paper/mri_paper.tex
placeholder figures
more

@@
 As Pm decreases, the width of the boundary layer becomes increasingly narrow (Figure \ref{u_secondorder_by_Pm}).
+The geometry of the velocity in the $x-z$ plane begins as a series of MRI ``cells" whose structure is radially narrow in the bulk of the flow.
+These central cells then stretch and grow, until at $Pm \sim 10^{-5}$ the bulk cells dominate the flow.
 The relative importance of the $u_y$ second order perturbation to the first order perturbation increases as well (CHECK, PLOT).
 
-\begin{figure*}[h!]
+\begin{figure*}[h]
 \centering
-\includegraphics[trim=6cm 0cm 6cm 0cm, scale=0.4]{u_secondorder_by_Pm_2.eps}
+\includegraphics[trim=2cm 0cm 6cm 0cm, scale=0.4]{u_secondorder_by_Pm_2.eps}
 \caption{(Placeholder) The evolution of the second-order velocity perturbations by Pm.}\label{allorders_velocity}
@@
 
-It appears that the second-order velocity perturbations are instrumental in the saturation of the MRI. The likely mechanism is a reduction in the background shear.
+%It appears that the second-order velocity perturbations are instrumental in the saturation of the MRI. The likely mechanism is a reduction in the background shear.
+
+From the behavior of the second order velocity perturbations (Figure \ref{u_secondorder_by_Pm}) in the bulk of the flow (i.e.
+apart from the boundary layers), we gather that the MRI is saturating via a background shear reduction.
+
+\begin{figure*}
+\centering
+\includegraphics[trim=6cm 4cm 6cm 3cm, scale=0.5]{stress_uy_cuts.pdf}
+\caption{(Placeholder) The total stress (left) and vertical cuts of the azimuthal velocity perturbations (right).
+The $u_y$ perturbations shown are first order (purple), second order (orange), and total (black).
+The blue shaded area represents the region where the second order and total perturbations are opposing the background shear because they are negative, and the red shaded region indicates where the second order and total perturbations are opposing the background shear because they are positive.}\label{allorders_velocity}
+\label{stress_uy_cuts}
+\end{figure*}
 

### 2014-12-17 paper/mri_paper.tex
and more

@@
 
+We plot the total stress, or the sum of the Reynolds and Maxwell stresses,
+
+\beq
+\mathbb{T}(x, z) = u_x u_y - \frac{2}{\beta} B_x B_y.
+\eeq
+
+See Figure \ref{stress_uy_cuts}.
+
 The role of the boundary layers changes with Pm.
@@
 apart from the boundary layers), we gather that the MRI is saturating via a background shear reduction.
+Recall that the linear shear profile means that the background shear is positive for negative values of $x$ and negative for positive values of $x$.
+Thus Figure \ref{u_secondorder_by_Pm} shows the azimuthal velocity perturbations resisting the background shear in the bulk of the flow
 

### 2014-12-17 paper/mri_paper.tex
fixed fig refs

@@
 
-We plot the total stress, or the sum of the Reynolds and Maxwell stresses,
-
-\beq
-\mathbb{T}(x, z) = u_x u_y - \frac{2}{\beta} B_x B_y.
-\eeq
-
-See Figure \ref{stress_uy_cuts}.
-
 The role of the boundary layers changes with Pm.
@@
 
-From the behavior of the second order velocity perturbations (Figure \ref{u_secondorder_by_Pm}) in the bulk of the flow (i.e.
+We plot the total stress, or the sum of the Reynolds and Maxwell stresses,
+
+\beq
+\mathbb{T}(x, z) = u_x u_y - \frac{2}{\beta} B_x B_y.
+\eeq
+
+See Figure \ref{stress_uy_cuts}.
+
+From the behavior of the second order velocity perturbations (Figure \ref{stress_uy_cuts}) in the bulk of the flow (i.e.
 apart from the boundary layers), we gather that the MRI is saturating via a background shear reduction.
 Recall that the linear shear profile means that the background shear is positive for negative values of $x$ and negative for positive values of $x$.
-Thus Figure \ref{u_secondorder_by_Pm} shows the azimuthal velocity perturbations resisting the background shear in the bulk of the flow
+Thus Figure \ref{stress_uy_cuts} shows the azimuthal velocity perturbations resisting the background shear in the bulk of the flow.
 

### 2015-01-11 notes/mri_notes.tex
N3 plots
checked hand calcs for all nonlinear terms

@@
 We now rewrite our system of equations, grouping similar terms into matrices. The system of equations we are rewriting is: \\
-
-$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, =  \, 2 \partial_z u_{1y} \,+ \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
+%1/9 edit: inserted Omega_0 into first eqn's 2 d_z u_y term.
+
+$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, =  \, 2 \Omega_0 \partial_z u_{1y} \,+ \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\
 
@@
 $+ \left[\begin{matrix}
-0 & 2\partial_z & 0 & 0 \\
-\left(\Omega_0 - 2\right)q\partial_z & 0 & 0 & 0 \\
+0 & 2 \Omega_0 \partial_z & 0 & 0 \\
+\left(2 - q\right)\Omega_0\partial_z & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 \\
-0 & 0 & -q\Omega_0\partial_z & 0 \\ \end{matrix}\right]$ \\
-
-We will now regroup matrices by $\mathcal{O}(\partial_z)$. The first matrix is rewritten as three matrices, keeping in mind that $\nabla^4 = \partial_x^4 + \partial_z^4 + 2\partial_x\partial_z^2$: \\
+0 & 0 & -q\Omega_0\partial_z & 0 \\ \end{matrix}\right]$ \\ %added Omega_0 to first column, 2nd row
+
+We will now regroup matrices by $\mathcal{O}(\partial_z)$.
+The first matrix is rewritten as three matrices, keeping in mind that $\nabla^4 = \partial_x^4 + \partial_z^4 + 2\partial_x^2\partial_z^2$: \\
 
@@
 $\mathcal{L}_1 = \left[\begin{matrix}
-0 & 2 & \frac{2}{\beta}\partial_x^2 & 0 \\
-(\Omega_0 - 2)q & 0 & 0 & \frac{2}{\beta} \\
+0 & 2 \Omega_0 & \frac{2}{\beta}\partial_x^2 & 0 \\
+(2 - q)\Omega_0 & 0 & 0 & \frac{2}{\beta} \\
 1 & 0 & 0 & 0 \\
@@
 
-$\mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, - \partial_Z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1}$ \\
+$\mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, - \partial_z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1}$ \\
 
@@
 
-$\mathcal{D}\mathbb{V}_{11} e^{iQz}\partial_T \alpha \, + \, \mathcal{D}\mathbb{U}_{11} \partial_T \beta \, + \, \alpha^3 \mathbf{N_{33}} e^{i3Qz} \, + \, \alpha \partial_Z \alpha \mathbf{N_{32}} e^{i2Qz} \, + \, \alpha\left|\alpha\right|^2 \mathbf{N_{31}} e^{iQz} \, + \, \alpha \partial_Z\beta \widetilde{\mathbf{N}}_{31} e^{iQz} \, + \, \alpha^* \partial_Z \alpha \mathbf{N_{30}} \, = \, \mathcal{L}\mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{22} 2 \alpha \partial_Z \alpha e^{i2Qz} \, + \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{20} 2 \left|\alpha\right| \partial_Z \alpha \, - \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} \partial_Z^2 \alpha e^{iQz} \, - \, \widetilde{\mathcal{L}}_1 \mathbb{U}_{20} \partial_Z^2 \beta \, + \, \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{iQz} \partial_Z^2 \alpha \, + \, \widetilde{\mathcal{L}}_2 \mathbb{U}_{11} \partial_Z^2 \beta \, - \, \mathcal{X}\mathbb{V}_{11} e^{iQz} \partial_Z \alpha \, - \, \mathcal{X} \mathbb{U}_{11} \partial_Z \beta \, + \, \mathcal{L}_3 \mathbb{V}_{11} e^{iQz} iQ^3 \alpha $\\
+$\mathcal{D}\mathbb{V}_{11} e^{iQz}\partial_T \alpha \, + \, \mathcal{D}\mathbb{U}_{11} \partial_T \beta \, + \, \alpha^3 \mathbf{N_{33}} e^{i3Qz} \, + \, \alpha \partial_Z \alpha \mathbf{N_{32}} e^{i2Qz} \, + \, \alpha\left|\alpha\right|^2 \mathbf{N_{31}} e^{iQz} \, + \, \alpha \partial_Z\beta \widetilde{\mathbf{N}}_{31} e^{iQz} \, + \, \alpha^* \partial_Z \alpha \mathbf{N_{30}} \, = \, \mathcal{L}\mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{22} 2 \alpha \partial_Z \alpha e^{i2Qz} \, + \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{20} 2 \left|\alpha\right| \partial_Z \alpha \, - \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} \partial_Z^2 \alpha e^{iQz} \, - \, \widetilde{\mathcal{L}}_1 \mathbb{U}_{20} \partial_Z^2 \beta \, + \, \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{iQz} \partial_Z^2 \alpha \, + \, \widetilde{\mathcal{L}}_2 \mathbb{U}_{11} \partial_Z^2 \beta \, - \, \mathcal{X}\mathbb{V}_{11} e^{iQz} \partial_z \alpha \, - \, \mathcal{X} \mathbb{U}_{11} \partial_z \beta \, + \, \mathcal{L}_3 \mathbb{V}_{11} e^{iQz} iQ^3 \alpha $\\
 
@@
 
-$\mathcal{D} \mathbb{V}_{11} e^{iQz}\partial_T \alpha \, + \, \alpha \left| \alpha^2 \right| \mathbf{N}_{31} e^{iQz} \, + \, \alpha \partial_Z \beta \widetilde{\mathbf{N}}_{31} e^{iQz} \, = \, - \partial_Z^2 \alpha \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} e^{iQz} \, + \, \partial_Z^2 \alpha \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{iQz} \, - \, \partial_Z \alpha \mathbb{V}_{11}\mathcal{X}e^{iQz} \, + \, iQ^3 \alpha \mathcal{L}_3 \mathbb{V}_{11} e^{iQz}$ \\
+$\mathcal{D} \mathbb{V}_{11} e^{iQz}\partial_T \alpha \, + \, \alpha \left| \alpha^2 \right| \mathbf{N}_{31} e^{iQz} \, + \, \alpha \partial_Z \beta \widetilde{\mathbf{N}}_{31} e^{iQz} \, = \, - \partial_Z^2 \alpha \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} e^{iQz} \, + \, \partial_Z^2 \alpha \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{iQz} \, - \, i Q \alpha \mathbb{V}_{11}\mathcal{X}e^{iQz} \, + \, iQ^3 \alpha \mathcal{L}_3 \mathbb{V}_{11} e^{iQz}$ \\
 
@@
 
-$a \partial_T \alpha \, + \, c \alpha \left|\alpha^2\right| \, + \, \widetilde{c} \alpha \partial_Z \beta \, = \, - b \partial_Z \alpha \, + \, h \partial_Z^2 \alpha \, + \, g i Q^3 \alpha$ \\
+$a \partial_T \alpha \, + \, c \alpha \left|\alpha^2\right| \, + \, \widetilde{c} \alpha \partial_Z \beta \, = \, - b i Q \alpha \, + \, h \partial_Z^2 \alpha \, + \, g i Q^3 \alpha$ \\
 

### 2015-01-15 paper/mri_paper.tex
missing f
description of cells with Pm
run correct Pm 1E-4 with 512 grid

@@
 \beq
-\mathbf{J} = \frac{2}{\beta}\nabla\times\mathbf{B}.
-\eeq
+%\mathbf{J} = \frac{2}{\beta}\nabla\times\mathbf{B}.
+\mathbf{J} = \nabla\times\mathbf{B},.
+\eeq
+
+in units where the permittivity of free space is unity.
 
@@
 The geometry of the velocity in the $x-z$ plane begins as a series of MRI ``cells" whose structure is radially narrow in the bulk of the flow.
-These central cells then stretch and grow, until at $Pm \sim 10^{-5}$ the bulk cells dominate the flow.
+As Pm is decreased, the saturation geometry of the central cells stretches radially, until at Pm $\sim 10^{-5}$ the bulk cells dominate the flow.
 The relative importance of the $u_y$ second order perturbation to the first order perturbation increases as well (CHECK, PLOT).
@@
 \beq\label{ampeq2_hmri}
-\partial_T \beta = - m \partial_Z^2 \beta + \frac{1}{\reym} \partial_Z^2 \beta \, - \, s \xi 2 \left|\alpha\right|\partial_Z \alpha \, - \, \xi\partial_Z^2 \beta,
+\partial_T \beta = - m \partial_Z^2 \beta + \frac{1}{\reym} \partial_Z^2 \beta \, - \, s \xi 2 \left|\alpha\right|\partial_Z \alpha \, - \, f\xi\partial_Z^2 \beta,
 \eeq

### 2015-01-15 python/MRI_runs.py

@@ -19,7 +19,7 @@ from matplotlib import rc
 rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
 rc('text', usetex=True)
 
-gridnum = 256
+gridnum = 512
 x_basis = Chebyshev(gridnum)
 domain = Domain([x_basis], grid_dtype=np.complex128)
 
@@ -33,22 +33,22 @@ beta = 25.0
 """
 
 #Pm = 1E-3: Q = 0.75, Rm = 4.8775
-
+"""
 Pm = 1.0E-3
 Q = 0.75
 Rm = 4.8775
 q = 1.5
 beta = 25.0
-
+"""
 
 #Pm = 1E-4: Q = 0.747, Rm = 4.88
-"""
+
 Pm = 1.0E-4
 Q = 0.747
 Rm = 4.88
 q = 1.5
 beta = 25.0
-"""
+
 
 #Pm = 1E-5, Q = 0.747, Rm = 4.88
 """
@@ -87,7 +87,9 @@ q = 1.5
 beta = 25.0
 """
 
-pc = PlotContours(Q = Q, Rm = Rm, Pm = Pm, q = q, beta = beta, run = True, norm = True)
+norm = False
+
+pc = PlotContours(Q = Q, Rm = Rm, Pm = Pm, q = q, beta = beta, run = True, norm = norm)
 pc.plot_streams()
 pc.plot_Bfield()
 
@@ -125,7 +127,7 @@ coeffs["A first order"] = pc.V1_A
 coeffs["Psi first order"] = pc.V1_psi
 
 
-outname = "gridnum_"+str(gridnum)+"_Pm_"+str(Pm)+"_Q_"+str(Q)+"_Rm_"+str(Rm)+"_q_"+str(q)+"_beta_"+str(beta)
+outname = "gridnum_"+str(gridnum)+"_Pm_"+str(Pm)+"_Q_"+str(Q)+"_Rm_"+str(Rm)+"_q_"+str(q)+"_beta_"+str(beta)+"_norm_"+str(norm)
 plot_uy_firstorder(pc, oplot = True, labels = False, outname=outname)
 plot_By_firstorder(pc, oplot = True, labels = False, outname=outname)
 plot_uy_secondorder(pc, oplot = True, labels = False, outname=outname)
[…]

### 2015-01-15 python/MRI_unpack.py

@@ -17,48 +17,68 @@ rc('text', usetex=True)
 
 q = 1.5
 beta = 25.0
+norm = False
 
 coeffs = {}
 
 #Pm = 1E-7: Q = 0.745, Rm = 4.90
-
+"""
 gridnum = 1024#512
 Pm = 1.0E-7
 Q = 0.745
 Rm = 4.90
 
 coeffs[0] = pickle.load(open("pspace/coeffs_gridnum"+str(gridnum)+"_Pm_"+str(Pm)+"_Q_"+str(Q)+"_Rm_"+str(Rm)+"_q_"+str(q)+"_beta_"+str(beta)+".p", "rb"))
-
+"""
+"""
 gridnum = 512
 # Pm = 1E-6: Q = 0.75, Rm = 4.88, beta = 25
 Pm = 1.0E-6
 Q = 0.75
 Rm = 4.88
 coeffs[1] = pickle.load(open("pspace/coeffs_gridnum"+str(gridnum)+"_Pm_"+str(Pm)+"_Q_"+str(Q)+"_Rm_"+str(Rm)+"_q_"+str(q)+"_beta_"+str(beta)+".p", "rb"))
+"""
 
 # all the rest have gridnum 256
-gridnum = 256
+#gridnum = 256
+
+gridnum = 64
+
 #Pm = 1E-5, Q = 0.747, Rm = 4.88
 Pm = 1.0E-5
 Q = 0.747
 Rm = 4.88
 
-coeffs[2] = pickle.load(open("pspace/coeffs_gridnum"+str(gridnum)+"_Pm_"+str(Pm)+"_Q_"+str(Q)+"_Rm_"+str(Rm)+"_q_"+str(q)+"_beta_"+str(beta)+".p", "rb"))
+if norm == False:
+    coeffs[0] = pickle.load(open("pspace/coeffs_gridnum"+str(gridnum)+"_Pm_"+str(Pm)+"_Q_"+str(Q)+"_Rm_"+str(Rm)+"_q_"+str(q)+"_beta_"+str(beta)+"_norm_"+str(norm)+".p", "rb"))
+else:
+    coeffs[0] = pickle.load(open("pspace/coeffs_gridnum"+str(gridnum)+"_Pm_"+str(Pm)+"_Q_"+str(Q)+"_Rm_"+str(Rm)+"_q_"+str(q)+"_beta_"+str(beta)+".p", "rb"))
+
 
[…]

### 2015-01-15 python/allorders.py

@@ -18,7 +18,7 @@ from matplotlib import rc
 rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
 rc('text', usetex=True)
 
-gridnum = 64#256
+gridnum = 512
 x_basis = Chebyshev(gridnum)
 domain = Domain([x_basis], grid_dtype=np.complex128)
 
@@ -1602,7 +1602,7 @@ class AmplitudeAlpha():
         if run == True:
         
             self.va = AdjointHomogenous(Q = self.Q, Rm = self.Rm, Pm = self.Pm, q = self.q, beta = self.beta, norm = self.norm)
-            self.va.solve(save = False, norm = True)
+            self.va.solve(save = False)
         
             self.v1 = OrderE(Q = self.Q, Rm = self.Rm, Pm = self.Pm, q = self.q, beta = self.beta, norm = self.norm)
             self.v1.solve(save=False)
@@ -2117,8 +2117,8 @@ class SolveAmplitudeAlpha():
         #problem.add_equation("a*dt(alpha) + b*alphaZ - h*dZ(alphaZ) - g*1j*Q**3*alpha = -c*alpha*MagSq(alpha)")
         #problem.add_equation("-(a/c)*dt(alpha) - (b/c)*alphaZ + (h/c)*dZ(alphaZ) + (g/c)*1j*Q**3*alpha = alpha*Absolute(alpha**2)")
         
-        #problem.add_equation("ac*dt(alpha) + bc*1j*Q*alpha - hc*dZ(alphaZ) - gc*1j*Q**3*alpha = alpha*Absolute(alpha**2)") #fixed to be gle
-        problem.add_equation("ac*dt(alpha) - bc*1j*Q*alpha - hc*dZ(alphaZ) = alpha*Absolute(alpha**2)") #fixed to be gle
+        problem.add_equation("ac*dt(alpha) - bc*1j*Q*alpha - hc*dZ(alphaZ) + gc*1j*Q**3*alpha = alpha*Absolute(alpha**2)") #fixed to be gle
+        #problem.add_equation("ac*dt(alpha) - bc*1j*Q*alpha - hc*dZ(alphaZ) = alpha*Absolute(alpha**2)") #fixed to be gle
         problem.add_equation("alphaZ - dZ(alpha) = 0")
         
         
@@ -2148,7 +2148,7 @@ class SolveAmplitudeAlpha():
         # stopping criteria
         solver.stop_sim_time = np.inf
[…]

### 2015-01-20 python/allorders.py
edited norm of allorders to not include magic number. Still includes real imag shift

@@ -125,7 +125,9 @@ class AdjointHomogenous():
             a = self.LEV.state['psi']['g'].real[13]/n
             b = self.LEV.state['psi']['g'].imag[13]/n
             scale = 1j*a/(b*(a**2/b+b)) + 1./(a**2/b +b)
-            scale *= -664.4114817
+        
+            # magic number
+            #scale *= -664.4114817
             
             psinorm = LEV.state['psi']*scale
             self.psi = psinorm.evaluate()
@@ -295,7 +297,9 @@ class OrderE():
             a = self.LEV.state['psi']['g'].real[13]/n
             b = self.LEV.state['psi']['g'].imag[13]/n
             scale = 1j*a/(b*(a**2/b+b)) + 1./(a**2/b +b)
-            scale *= -664.4114817 
+            
+            # magic number
+            #scale *= -664.4114817 
             
             self.psi = LEV.state['psi']['g']*scale
             self.u = LEV.state['u']['g']*scale
@@ -467,9 +471,10 @@ class N2():
         N20B = 1j*Q*self.psi_1*self.B_1_star_x + self.psi_1_x*1j*Q*self.B_1_star - 1j*Q*self.A_1*self.u_1_star_x - self.A_1_x*1j*Q*self.u_1_star
         self.N20_B = N20B.evaluate()
         
+        """
         if norm == True:
-        
-            self.n20scale = 0.2/0.087941764519131521
+            # magic number
+            #self.n20scale = 0.2/0.087941764519131521
             
             N20psi = self.N20_psi*self.n20scale
             self.N20_psi = N20psi.evaluate()
@@ -482,6 +487,7 @@ class N2():
             
             N20B = self.N20_B*self.n20scale
             self.N20_B = N20B.evaluate()
+        """
         
     def plot(self):
     
@@ -840,8 +846,8 @@ class OrderE2():
             a = self.LEV21.state['psi21']['g'].real[13]/n
             b = self.LEV21.state['psi21']['g'].imag[13]/n
             scale = 1j*a/(b*(a**2/b+b)) + 1./(a**2/b +b)
-            tt = 1j*132/29.097408658719342
-            scale *= tt 
+            #tt = 1j*132/29.097408658719342
+            #scale *= tt 
             
             self.psi21 = LEV21.state['psi21']['g']*scale
             self.u21 = LEV21.state['u21']['g']*scale

### 2015-11-16 python/allorders_2.py
Pm 2E-3

@@ -31,9 +31,11 @@ class MRI():
 
     """
     Base class for MRI equations.
+    
+    Defaults: For Pm of 0.001 critical Rm is 4.879  critical Q is 0.748
     """
 
-    def __init__(self, Q = 0.752, Rm = 4.91, Pm = 0.001, q = 1.5, beta = 25.0, norm = True):
+    def __init__(self, Q = 0.748, Rm = 4.879, Pm = 0.001, q = 1.5, beta = 25.0, norm = True):
     
         self.Q = Q
         self.Rm = Rm
@@ -527,9 +529,10 @@ class MRI():
         Take inner product < vector2 | vector1 > 
         Defined as integral of (vector2.conj * vector1)
         """
-
-        inner_product = vector1[0]['g']*vector2[0]['g'].conj() + vector1[1]['g']*vector2[1]['g'].conj() + vector1[2]['g']*vector2[2]['g'].conj() + vector1[3]['g']*vector2[3]['g'].conj()
+        print(vector1)
         
+        inner_product = vector1[0]['g']*vector2[0]['g'].conj() + vector1[1]['g']*vector2[1]['g'].conj() + vector1[2]['g']*vector2[2]['g'].conj() + vector1[3]['g']*vector2[3]['g'].conj()
+        print(inner_product)
         
         ip = domain.new_field()
         ip.name = "inner product"
@@ -564,7 +567,7 @@ class AdjointHomogenous(MRI):
     Returns V^dagger
     """
 
-    def __init__(self, Q = 0.752, Rm = 4.91, Pm = 0.001, q = 1.5, beta = 25.0, norm = True):
+    def __init__(self, Q = 0.748, Rm = 4.879, Pm = 0.001, q = 1.5, beta = 25.0, norm = True):
         
         print("initializing Adjoint Homogenous")
         
@@ -678,7 +681,7 @@ class OrderE(MRI):
     Returns V_1
     """
 
-    def __init__(self, Q = 0.752, Rm = 4.91, Pm = 0.001, q = 1.5, beta = 25.0, norm = True):
+    def __init__(self, Q = 0.748, Rm = 4.879, Pm = 0.001, q = 1.5, beta = 25.0, norm = True):
         
[…]

### 2015-11-16 python/parallel_LEV.py

@@ -261,7 +261,8 @@ if __name__ == '__main__':
     #Pm = 3.0E-4
     #Pm = 5.0E-4
     #Pm = 9.0E-5
-    Pm = 5.0E-5
+    #Pm = 5.0E-5
+    Pm = 2.0E-3
     
     q = 1.5
     beta = 25.0

### 2015-11-16 python/parameter_critical_run.py

@@ -187,7 +187,7 @@ if __name__ == "__main__":
     for i, Pm in enumerate(Pms):
         marginal_Rm, marginal_Q = get_critical_parameters_by_Pm(Pm)
         
-        amplitude_obj = AmplitudeAlpha(Pm = Pm, Rm = marginal_Rm, Q = marginal_Q, norm = False)
+        amplitude_obj = AmplitudeAlpha(Pm = Pm, Rm = marginal_Rm, Q = marginal_Q, norm = True)
         coeffs[i] = amplitude_obj.sat_amp_coeffs
         ivpsa[i] = amplitude_obj.saturation_amplitude
         

### 2016-01-11 paper/wide_gap_notes.tex
added notes incomplete on wide gap.

\documentclass{paper}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\hyd}    {{\rm H}}
\newcommand{\vobs}{v_{\rm obs}}
\newcommand{\hi}{H{\sc i}~}
\newcommand{\hii}{H{\sc ii}~}
\newcommand{\hia}{H{\sc i}}
\newcommand{\hiia}{H{\sc ii}}

\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{color}
%\usepackage{appendix}
\usepackage[titletoc,title]{appendix}
\usepackage{cleveref}

\makeatletter
\setlength{\@fptop}{0pt}
\setlength{\@fpbot}{0pt}
\setlength{\@fpsep}{0pt}
\makeatother

\newcommand{\citei}[1]{\citeauthor{#1} \citeyear{#1}}
\newcommand{\unit}[1]{\textrm{ #1}}
\newcommand{\kms}{km ${\rm s^{-1}}$}
\newcommand{\kmsa}{km ${\rm s^{-1}}$}
\newcommand{\citeia}[2]{\citeauthor{#1}~(\citeyear{#1};~#2)}

\newcommand{\uphi}{\ensuremath{u_\phi}}
\newcommand{\rhat}{\ensuremath{\mathbf{\hat{r}}}}
\newcommand{\phihat}{\ensuremath{\mathbf{\hat{\phi}}}}
\newcommand{\zhat}{\ensuremath{\mathbf{\hat{z}}}}

\newcommand\reye{\mathrm{Re}}
\newcommand\reym{\mathrm{Rm}}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\begin{document}

\title{Notes on Wide Gap}

\section{Basic Equations}
\label{sec:equations}
The so-called Stokes stream function, used in axisymmetric situations, is given by 
\begin{equation}
  \label{eq:stokes}
  \mathbf{u} \, = \, \left[\begin{matrix}
\frac{1}{r} \partial_z \psi\ \rhat\\
\uphi \ \phihat\\
-\frac{1}{r} \partial_r \psi\ \zhat\\
\end{matrix}\right];
\end{equation}
here we define $A$ in the same way.

Using the definitions in

\begin{multline}
  \label{eq:psi}
\partial_t \left[ \frac{1}{r} \left(\nabla^2 \psi - \frac{2 \partial_r \psi}{r} \right) \right] + \frac{1}{r^2} J(\psi, \nabla^2 A - \frac{2 \partial_r \psi}{r}) = \frac{\partial_z A}{r^3} \left( \nabla^2 A - \frac{2 \partial_r A}{r}\right) \\
+ \frac{1}{r} J\left(A,\frac{1}{r}\left( \nabla^2 A - \frac{2 \partial_r A}{r}\right)\right) - \frac{2 B_\phi \partial_z B_\phi}{r}\\
+ \nu \left\{ \nabla^2 \left[ \frac{1}{r} \left(\nabla^2 \psi - \frac{2 \partial_r \psi}{r}\right) \right]  -\frac{1}{r^2} \left(\nabla^2 \psi - \frac{2 \partial_r \psi}{r}\right) \right\}
\end{multline}

\begin{equation}
  \label{eq:uy}
\partial_t \uphi + J(\psi, \uphi) + \frac{\uphi \partial_r \psi}{r^2} = J(B_\phi, A) - \frac{B_\phi \partial_z A}{r^2} + \nu \left( \nabla^2 \uphi - \frac{\uphi}{r}\right)
\end{equation}

\begin{equation}
  \label{eq:A}
\partial_t A = \frac{1}{r^2} J(\psi, A) - \eta \left[ \frac{1}{r}\left( \nabla^2 A - \frac{2 \partial_r A}{r} \right)\right]
\end{equation}

\begin{multline}
  \label{eq:By}
\partial_t B_\phi = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)\\
+ \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A + \eta \left(\nabla^2 B_\phi - \frac{1}{r^2} B_\phi \right)
\end{multline}

\section{Recovery of Narrow Gap Equations}
\label{sec:narrow_gap_recovery}

\appendix

\section{Cylindrical derivatives}
\label{sec:cylindrical_deriv}

Everything here follows \url{http://farside.ph.utexas.edu/teaching/336L/Fluidhtml/node177.html#scyl}.

For a scalar field $\psi$,

\begin{equation}
  \label{eq:del_scalar}
  \nabla \psi = \frac{\partial \psi}{\partial r} \rhat + \frac{1}{r} \frac{\partial \psi}{\partial \phi} \phihat + \frac{\partial \psi}{\partial z} \zhat.
\end{equation}

However, for a \emph{vector} field $\mathbf{u}$,
\begin{equation}
  \label{eq:div}
  \mathbf{\nabla \cdot u} = \frac{1}{r} \frac{\partial (r u_r)}{\partial r} + \frac{1}{r} \frac{\partial \uphi}{\partial \phi}  + \frac{\partial u_z}{\partial z}
\end{equation}

and

\begin{equation}
  \label{eq:curl}
  \mathbf{\nabla \times u} = \left( \frac{1}{r} \frac{\partial u_z}{\partial \phi} - \frac{\partial u_\phi}{\partial z}\right)\ \rhat + \left(\frac{\partial u_r}{\partial z} - \frac{\partial u_z}{\partial r}\right)\ \phihat  + \left( \frac{1}{r} \frac{\partial (r \uphi)}{\partial r} - \frac{1}{r}\frac{\partial u_r}{\partial \phi}\right)\ \zhat.
\end{equation}

We also need the $\phi$ component of the convective derivative $\mathbf{u \cdot \nabla u}$,
\begin{equation}
  \label{eq:convective_deriv_phi}
  \left[\mathbf{u \cdot \nabla u} \right]_\phi = \mathbf{u \cdot \nabla} \uphi + \frac{ u_r \uphi}{r},
\end{equation}
and finally, the vector Laplacian,
\begin{equation}
  \label{eq:vec_lap_r}
  (\nabla^2 \mathbf{u})_r = \nabla^2 u_r - \frac{u_r}{r^2} - \frac{2}{r^2} \frac{\partial \uphi}{\partial \phi}
\end{equation}

\begin{equation}
  \label{eq:vec_lap_phi}
  (\nabla^2 \mathbf{u})_\phi = \nabla^2 u_\phi + \frac{2}{r^2} \frac{\partial u_r}{\partial \phi} - \frac{u_\phi}{r^2} 
\end{equation}

\begin{equation}
  \label{eq:vec_lap_z}
  (\nabla^2 \mathbf{u})_z = \nabla^2 u_z,
\end{equation}
where $\nabla$ on the vector components is given by equation~(\ref{eq:del_scalar}).
\end{document}

### 2016-02-27 paper/wide_gap_notes.tex
susan question on definition of streamfunction
susan question about streamfunction answered
B_phi term confirmed correct
got something different for vector potential eqn... will check
susan still gets different terms for A equation.... please check?
fixed typos in u_phi equation. I think this matches Susan's now...
Fixed A equation: Susan was right. I was using a different (and wrong!) definition of B from A.
r component of Psi equation in expanded form
z component of Psi equation in expanded form
missing term
typo
never mind the recent commits they are wrong
detailed derivation of Psi equatio

@@
 \end{equation}
-here we define $A$ in the same way.
+here we define $A$ in the same way. \\
 
@@
 
+For the expanded form of the $\Psi$ equation, Susan gets:
+
 \begin{equation}
   \label{eq:uy}
-\partial_t \uphi + J(\psi, \uphi) + \frac{\uphi \partial_r \psi}{r^2} = J(B_\phi, A) - \frac{B_\phi \partial_z A}{r^2} + \nu \left( \nabla^2 \uphi - \frac{\uphi}{r}\right)
+\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} = \frac{J(A, B_\phi)}{r} + \frac{B_\phi \partial_z A}{r^2} + \nu \left( \nabla^2 \uphi - \frac{\uphi}{r}\right)
 \end{equation}
@@
   \label{eq:A}
-\partial_t A = \frac{1}{r^2} J(\psi, A) - \eta \left[ \frac{1}{r}\left( \nabla^2 A - \frac{2 \partial_r A}{r} \right)\right]
+\partial_t A = \frac{1}{r} J(A, \psi) + \eta \left( \nabla^2 A - \frac{2 \partial_r A}{r} \right)
 \end{equation}
 
+Susan gets:
+\beq
+\partial_t A = \frac{1}{r} J(A, \Psi) + \frac{1}{Rm} \left[ -\frac{1}{r} \partial_r A + \partial_r^2 A + \partial_z^2 A\right]
+\eeq
+%Susan got:
+%beq
+%\partial_t A = \frac{2}{r} J(A, \Psi) + \frac{1}{Rm} \left[ \nabla^2 A - \frac{2}{r} \partial_r A \right]
+%\eeq
+%....will check.
+
 \begin{multline}
-  \label{eq:By}
+  \label{eq:By} % Susan confirms correct 1/14/16
 \partial_t B_\phi = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)\\
@@
 \end{multline}
+
+\section{Detailed Derivation of $\Psi$ Equation}
+The $\Psi$ equation, governing the x- and z-components of the velocity, is particularly tricky to derive so I will write out the steps here.
+
+1. Find x and z components of the momentum equation, i.e.:
+
+\beq
+\partial_t u_z + \left[u \cdot \nabla u \right]_z = \left[ \left( \nabla \times B \right) \times B \right]_z + \frac{1}{\reye}\left[ \nabla^2 u \right] _z
+\eeq
+
+Which becomes
+
+\beq
+\frac{1}{r^2} \partial_t \partial_r \Psi
+\eeq
 
@@
 where $\nabla$ on the vector components is given by equation~(\ref{eq:del_scalar}).
+
+Note that, expanding the definition of the vector Laplacian, where the cylindrical scalar Laplacian is substituted in for $\nabla^2 u_r$ and $\nabla^2 u_z$
+
 \end{document}

### 2016-02-27 paper/wide_gap_notes.tex
z component of Psi equation
rhat momentum equation
LHS of Psi equation

@@
 
-1. Find x and z components of the momentum equation, i.e.:
+1. Find $\rhat$ and $\zhat$ components of the momentum equation, i.e.:
 
@@
 
-Which becomes
+We sub in our stream/flux function notation and expand the operators in cylindrical coordinates. Then take $\partial_r$ of the resulting equation to obtain:
+
+\begin{multline}
+\frac{1}{r^2} \partial_t \partial_r \Psi - \frac{1}{r} \partial_t \partial_r^2 \Psi - \frac{3}{r^4}\partial_z \Psi \partial_r \Psi + \frac{1}{r^3} \partial_r \left( \partial_z \Psi \partial_r \Psi \right) + \frac{2}{r^3} \partial_z \Psi \partial_r^2 \Psi - \frac{1}{r^2} \partial_r \left(\partial_z \Psi \partial_r^2 \Psi \right)\\
+- \frac{2}{r^3} \partial_r \Psi \partial_r\partial_z \Psi + \frac{1}{r^2} \partial_r \left(\partial_r \Psi \partial_r \partial_z \Psi\right) = \\
+\partial_r\left(B_\theta \partial_z B_\theta\right) + \frac{2}{r^3} \partial_z^2 A \partial_z A - \frac{1}{r^2} \partial_r \left(\partial_z^2 A \partial_z A\right) + \frac{3}{r^4} \partial_z A \partial_r A - \frac{1}{r^3} \partial_r \left(\partial_z A \partial_r A \right) - \frac{2}{r^3} \partial_z A \partial_r^2 A \\
++ \frac{1}{r^2} \partial_r \left(\partial_z A \partial_r^2 A \right) 
++ \frac{1}{\reye} \left[\frac{3}{r^4} \partial_r \Psi - \frac{3}{r^3} \partial_r^2 \Psi + \frac{2}{r^2} \partial_r^3 \psi - \frac{1}{r} \partial_r^4 \Psi\right]
+\end{multline}
+
+Repeat this process for the $\rhat$ component of the momentum equation,
 
 \beq
-\frac{1}{r^2} \partial_t \partial_r \Psi
+\partial_t u_r + \left[u \cdot \nabla u \right]_r = \left[ \left( \nabla \times B \right) \times B \right]_r + \frac{1}{\reye}\left[ \nabla^2 u \right] _r
 \eeq
+
+and take $\partial_z$ of the expanded equation to obtain
+
+\begin{multline}
+\frac{1}{r} \partial_t \partial_z^2 \Psi - \frac{1}{r^3} \partial_z \left(\partial_z \Psi \partial_z \Psi\right) + \frac{1}{r^2} \partial_z\left(\partial_z\Psi \partial_z\partial_r \Psi\right) - \frac{1}{r^2} \partial_z \left(\partial_r \Psi \partial_z^2 \Psi\right) - \frac{1}{r} 2 u_\theta \partial_z u_\theta \\
+= - \frac{1}{r^2} \partial_z^3 A \partial_r A - \frac{1}{r^2} \partial_z^2 A \partial_r \partial_z A + \frac{2}{r^3} \partial_r\partial_z A \partial_r A - \frac{1}{r^2} \partial_r^2 \partial_z A \partial_rA - \frac{1}{r^2} \partial_r^2 A \partial_r \partial_z A
+\end{multline}
+
+It is clear from the $\partial_t$ terms that we must combine these equations by subtracting the $\zhat$ equation from the $\rhat$ equation.
+
+When we do, we can simplify the LHS of the equation to:
+
+\beq
+\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\theta \partial_z u_\theta
+\eeq
+
+Note that the relevant quantity appears to be $\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi$, and that the $\frac{1}{r^2}$ in the second term cannot come out of the Jacobian (a point of disagreement with Jeff's equation above).
 

### 2016-02-29 paper/wide_gap_notes.tex
LHS of Psi equation
typos
full Psi equation seemingly correct
full derivation of JxB

@@
 
-\begin{equation}
-  \label{eq:A}
-\partial_t A = \frac{1}{r} J(A, \psi) + \eta \left( \nabla^2 A - \frac{2 \partial_r A}{r} \right)
-\end{equation}
+%\begin{equation}
+ % \label{eq:A}
+%\partial_t A = \frac{1}{r} J(A, \psi) + \eta \left( \nabla^2 A - \frac{2 \partial_r A}{r} \right)
+%\end{equation}
 
-Susan gets:
+%Susan confirms correct (just switched eta -> 1/Rm)
 \beq
-\partial_t A = \frac{1}{r} J(A, \Psi) + \frac{1}{Rm} \left[ -\frac{1}{r} \partial_r A + \partial_r^2 A + \partial_z^2 A\right]
+\partial_t A = \frac{1}{r} J(A, \psi) + \frac{1}{\reym} \left( \nabla^2 A - \frac{2 \partial_r A}{r} \right)
+%\partial_t A = \frac{1}{r} J(A, \Psi) + \frac{1}{Rm} \left[ -\frac{1}{r} \partial_r A + \partial_r^2 A + \partial_z^2 A\right]
 \eeq
@@
 + \frac{1}{r^2} \partial_r \left(\partial_z A \partial_r^2 A \right) 
-+ \frac{1}{\reye} \left[\frac{3}{r^4} \partial_r \Psi - \frac{3}{r^3} \partial_r^2 \Psi + \frac{2}{r^2} \partial_r^3 \psi - \frac{1}{r} \partial_r^4 \Psi\right]
++ \frac{1}{\reye} \left[\frac{3}{r^4} \partial_r \Psi - \frac{3}{r^3} \partial_r^2 \Psi + \frac{2}{r^2} \partial_r^3 \Psi - \frac{1}{r} \partial_r^4 \Psi\right]
 \end{multline}
@@
 \frac{1}{r} \partial_t \partial_z^2 \Psi - \frac{1}{r^3} \partial_z \left(\partial_z \Psi \partial_z \Psi\right) + \frac{1}{r^2} \partial_z\left(\partial_z\Psi \partial_z\partial_r \Psi\right) - \frac{1}{r^2} \partial_z \left(\partial_r \Psi \partial_z^2 \Psi\right) - \frac{1}{r} 2 u_\theta \partial_z u_\theta \\
-= - \frac{1}{r^2} \partial_z^3 A \partial_r A - \frac{1}{r^2} \partial_z^2 A \partial_r \partial_z A + \frac{2}{r^3} \partial_r\partial_z A \partial_r A - \frac{1}{r^2} \partial_r^2 \partial_z A \partial_rA - \frac{1}{r^2} \partial_r^2 A \partial_r \partial_z A
+= - \frac{1}{r^2} \partial_z^3 A \partial_r A - \frac{1}{r^2} \partial_z^2 A \partial_r \partial_z A + \frac{2}{r^3} \partial_r\partial_z A \partial_r A - \frac{1}{r^2} \partial_r^2 \partial_z A \partial_rA - \frac{1}{r^2} \partial_r^2 A \partial_r \partial_z A \\
++ \frac{1}{\reye} \left[ -\frac{1}{r^2} \partial_z^2 \partial_r \Psi + \frac{1}{r} \partial_z^2 \partial_r^2 \Psi + \frac{1}{r} \partial_z^4\Psi \right]
 \end{multline}
@@
 Note that the relevant quantity appears to be $\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi$, and that the $\frac{1}{r^2}$ in the second term cannot come out of the Jacobian (a point of disagreement with Jeff's equation above).
+Also I'm confused why Jeff's has no $u_\theta$ term.
+The RHS of this equation is significantly more complicated.
+
+RHS viscous term:
+
+\beq
+\frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right]
+\eeq
+
+Full $\Psi$ equation according to Susan:
+
+\begin{multline}
+\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\theta \partial_z u_\theta \\
+= J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{r}B_\theta \partial_z B_\theta \\
++ \frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right]
+\end{multline}
+
+Note that this is actually beautifully symmetric. Except the viscous term which still seems clunky.....
+
+The derivation of the non-viscous term on the righthand side of the momentum equation ($J \times B$) is as follows.
+
+\beq
+\partial_z\left(\left[\left(\nabla \times B \right) \times B \right]_r\right) - \partial_r \left(\left[\left(\nabla \times B\right) \times B \right]_z\right)
+\eeq
+
+\beq
+= \partial_z\left(\left[\left(\partial_z B_r - \partial_r B_z \right) B_z - \left(\frac{1}{r}\partial_r\left(r B_\theta\right)\right) B_\theta \right] \right) - \partial_r \left(\left[\left(-\partial_z B_\theta\right)B_\theta - \left(\partial_z B_r - \partial_r B_z \right) B_r\right]\right)
+\eeq
+
+\begin{multline}
+= -\frac{1}{r^2} \partial_z^3 A \partial_r A + \frac{1}{r^3} \partial_r \partial_z A \partial_r A - \frac{1}{r^2} \partial_r^2 \partial_z A \partial_r A - \frac{2}{r^3} \partial_z^2 A \partial_z A \\
++ \frac{1}{r^2} \partial_z^2 \partial_r A \partial_z A + \frac{3}{r^4} \partial_r A \partial_z A - \frac{3}{r^3} \partial_r^2 A \partial_z A + \frac{1}{r^2} \partial_r^3 A \partial_z A - \frac{2}{r} B_\theta \partial_z B_\theta
+\end{multline}
+
+This simplifies to
+
+\beq
+J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{r}B_\theta \partial_z B_\theta
+\eeq
 

### 2016-03-19 python/parallel_LEV.py
check biharmonic operator

@@ -60,7 +60,7 @@ def Pmrun(Pm, q, Co, dQ, dRm, Qsearch, Rmsearch):
     
     print(result_dict)
 
-    pickle.dump(result_dict, open("multirun/gridnum_"+str(gridnum)+"_Pm_"+str(Pm)+"_Q_"+str(Qsearch[0])+"_dQ_"+str(dQ)+"_Rm_"+str(Rmsearch[0])+"_dRm_"+str(dRm)+"_allgoodeigs.p", "wb"))
+    pickle.dump(result_dict, open("multirun/gridnum_"+str(gridnum)+"_Pm_"+str(Pm)+"_Q_"+str(Qsearch[0])+"_dQ_"+str(dQ)+"_Rm_"+str(Rmsearch[0])+"_dRm_"+str(dRm)+"_allgoodeigs_testbiharmonic.p", "wb"))
 
     print("process took %10.5e seconds" % (time.time() - start_time))
     
@@ -205,7 +205,8 @@ def run_mri_solve(Q, Pm, Rm, q, Co, run_id, all_mode=False):
         domain = Domain([x_basis])
 
         # In terms of Co ....multiplied dt terms by -1j
-        lv1.add_equation("-1j*dt(psixx) - -1j*Q**2*dt(psi) - iR*dx(psixxx) + 2*iR*Q**2*psixx - iR*Q**4*psi - 2*1j*Q*u - Co*1j*Q*dx(Ax) + Co*Q**3*1j*A = 0")
+        #lv1.add_equation("-1j*dt(psixx) - -1j*Q**2*dt(psi) - iR*dx(psixxx) + 2*iR*Q**2*psixx - iR*Q**4*psi - 2*1j*Q*u - Co*1j*Q*dx(Ax) + Co*Q**3*1j*A = 0")
+        lv1.add_equation("-1j*dt(psixx) - -1j*Q**2*dt(psi) - iR*dx(psixxx) - iR*Q**4*psi - 2*1j*Q*u - Co*1j*Q*dx(Ax) + Co*Q**3*1j*A = 0") # testing alternative biharmonic operator def
         lv1.add_equation("-1j*dt(u) - iR*dx(ux) + iR*Q**2*u + (2-q)*1j*Q*psi - Co*1j*Q*B = 0") 
         lv1.add_equation("-1j*dt(A) - iRm*dx(Ax) + iRm*Q**2*A - 1j*Q*psi = 0") 
[…]

### 2016-03-23 paper/wide_gap_notes.tex
check biharmonic operator
perturbation of A eqn

--- 2016-02-27 paper/wide_gap_notes.tex
@@
 
-\begin{equation}
+%\begin{equation}
+ % \label{eq:A}
+%\partial_t A = \frac{1}{r} J(A, \psi) + \eta \left( \nabla^2 A - \frac{2 \partial_r A}{r} \right)
+%\end{equation}
+
+%Susan confirms correct (just switched eta -> 1/Rm)
+\beq
  \label{eq:A}
-\partial_t A = \frac{1}{r} J(A, \psi) + \eta \left( \nabla^2 A - \frac{2 \partial_r A}{r} \right)
-\end{equation}
-
-Susan gets:
-\beq
-\partial_t A = \frac{1}{r} J(A, \Psi) + \frac{1}{Rm} \left[ -\frac{1}{r} \partial_r A + \partial_r^2 A + \partial_z^2 A\right]
+\partial_t A = \frac{1}{r} J(A, \psi) + \frac{1}{\reym} \left( \nabla^2 A - \frac{2 \partial_r A}{r} \right)
+%\partial_t A = \frac{1}{r} J(A, \Psi) + \frac{1}{Rm} \left[ -\frac{1}{r} \partial_r A + \partial_r^2 A + \partial_z^2 A\right]
 \eeq
@@
 - \frac{2}{r^3} \partial_r \Psi \partial_r\partial_z \Psi + \frac{1}{r^2} \partial_r \left(\partial_r \Psi \partial_r \partial_z \Psi\right) = \\
-\partial_r\left(B_\theta \partial_z B_\theta\right) + \frac{2}{r^3} \partial_z^2 A \partial_z A - \frac{1}{r^2} \partial_r \left(\partial_z^2 A \partial_z A\right) + \frac{3}{r^4} \partial_z A \partial_r A - \frac{1}{r^3} \partial_r \left(\partial_z A \partial_r A \right) - \frac{2}{r^3} \partial_z A \partial_r^2 A \\
+\partial_r\left(B_\phi \partial_z B_\phi\right) + \frac{2}{r^3} \partial_z^2 A \partial_z A - \frac{1}{r^2} \partial_r \left(\partial_z^2 A \partial_z A\right) + \frac{3}{r^4} \partial_z A \partial_r A - \frac{1}{r^3} \partial_r \left(\partial_z A \partial_r A \right) - \frac{2}{r^3} \partial_z A \partial_r^2 A \\
 + \frac{1}{r^2} \partial_r \left(\partial_z A \partial_r^2 A \right) 
-+ \frac{1}{\reye} \left[\frac{3}{r^4} \partial_r \Psi - \frac{3}{r^3} \partial_r^2 \Psi + \frac{2}{r^2} \partial_r^3 \psi - \frac{1}{r} \partial_r^4 \Psi\right]
++ \frac{1}{\reye} \left[\frac{3}{r^4} \partial_r \Psi - \frac{3}{r^3} \partial_r^2 \Psi + \frac{2}{r^2} \partial_r^3 \Psi - \frac{1}{r} \partial_r^4 \Psi\right]
 \end{multline}
@@
 \begin{multline}
-\frac{1}{r} \partial_t \partial_z^2 \Psi - \frac{1}{r^3} \partial_z \left(\partial_z \Psi \partial_z \Psi\right) + \frac{1}{r^2} \partial_z\left(\partial_z\Psi \partial_z\partial_r \Psi\right) - \frac{1}{r^2} \partial_z \left(\partial_r \Psi \partial_z^2 \Psi\right) - \frac{1}{r} 2 u_\theta \partial_z u_\theta \\
-= - \frac{1}{r^2} \partial_z^3 A \partial_r A - \frac{1}{r^2} \partial_z^2 A \partial_r \partial_z A + \frac{2}{r^3} \partial_r\partial_z A \partial_r A - \frac{1}{r^2} \partial_r^2 \partial_z A \partial_rA - \frac{1}{r^2} \partial_r^2 A \partial_r \partial_z A
+\frac{1}{r} \partial_t \partial_z^2 \Psi - \frac{1}{r^3} \partial_z \left(\partial_z \Psi \partial_z \Psi\right) + \frac{1}{r^2} \partial_z\left(\partial_z\Psi \partial_z\partial_r \Psi\right) - \frac{1}{r^2} \partial_z \left(\partial_r \Psi \partial_z^2 \Psi\right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi \\
+= - \frac{1}{r^2} \partial_z^3 A \partial_r A - \frac{1}{r^2} \partial_z^2 A \partial_r \partial_z A + \frac{2}{r^3} \partial_r\partial_z A \partial_r A - \frac{1}{r^2} \partial_r^2 \partial_z A \partial_rA - \frac{1}{r^2} \partial_r^2 A \partial_r \partial_z A \\
++ \frac{1}{\reye} \left[ -\frac{1}{r^2} \partial_z^2 \partial_r \Psi + \frac{1}{r} \partial_z^2 \partial_r^2 \Psi + \frac{1}{r} \partial_z^4\Psi \right]
 \end{multline}
@@
 \beq
-\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\theta \partial_z u_\theta
+\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi
 \eeq
@@
 Note that the relevant quantity appears to be $\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi$, and that the $\frac{1}{r^2}$ in the second term cannot come out of the Jacobian (a point of disagreement with Jeff's equation above).
+Also I'm confused why Jeff's has no $u_\phi$ term.
+The RHS of this equation is significantly more complicated.
+
+RHS viscous term:
+
+\beq
+\frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right]
+\eeq
+
+Full $\Psi$ equation according to Susan:
+
+\begin{multline}
+\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi \\
+= J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{r}B_\phi \partial_z B_\phi \\
++ \frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right]
+\end{multline}
+
+Note that this is actually beautifully symmetric. Except the viscous term which still seems clunky.....
+
+The derivation of the non-viscous term on the righthand side of the momentum equation ($\mathbf{J} \times \mathbf{B}$) is as follows.
+
+\beq
+\partial_z\left(\left[\left(\nabla \times B \right) \times B \right]_r\right) - \partial_r \left(\left[\left(\nabla \times B\right) \times B \right]_z\right)
+\eeq
+
+\beq
+= \partial_z\left(\left[\left(\partial_z B_r - \partial_r B_z \right) B_z - \left(\frac{1}{r}\partial_r\left(r B_\phi\right)\right) B_\phi \right] \right) - \partial_r \left(\left[\left(-\partial_z B_\phi\right)B_\phi - \left(\partial_z B_r - \partial_r B_z \right) B_r\right]\right)
+\eeq
+
+\begin{multline}
+= -\frac{1}{r^2} \partial_z^3 A \partial_r A + \frac{1}{r^3} \partial_r \partial_z A \partial_r A - \frac{1}{r^2} \partial_r^2 \partial_z A \partial_r A - \frac{2}{r^3} \partial_z^2 A \partial_z A \\
++ \frac{1}{r^2} \partial_z^2 \partial_r A \partial_z A + \frac{3}{r^4} \partial_r A \partial_z A - \frac{3}{r^3} \partial_r^2 A \partial_z A + \frac{1}{r^2} \partial_r^3 A \partial_z A - \frac{2}{r} B_\phi \partial_z B_\phi
+\end{multline}
+
+This simplifies to
+
+\beq
+J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{r}B_\phi \partial_z B_\phi
+\eeq
+
+Full derivation of viscous term:
+
+\beq
+\partial_z \left(\frac{1}{\reye}\left[\nabla^2 u\right]_r\right) - \partial_r \left(\frac{1}{\reye}\left[\nabla^2 u\right]_z\right)
+\eeq
+
+\beq
+= \frac{1}{\reye}\left[\partial_z \left(\nabla^2 u_r - \frac{1}{r^2} u_r\right) - \partial_r \left(\nabla^2 u_z\right)\right]
+\eeq
+
+\begin{multline}
+= \frac{1}{\reye} \left[ - \frac{2}{r^2} \partial_z^2 \partial_r \Psi + \frac{2}{r} \partial_z^2 \partial_r^2 \Psi + \frac{1}{r} \partial_z^4 \Psi - \frac{3}{r^4} \partial_r \Psi + \frac{3}{r^3} \partial_r^2 \Psi - \frac{2}{r^2} \partial_r^3 \Psi + \frac{1}{r} \partial_r^4 \Psi \right]
+\end{multline}
 
@@
 \label{sec:narrow_gap_recovery}
+
+\section{Perturbed Equations}
+
+We perturb the wide gap equations according to
+
+\beq
+\mathbf{B} = B_0 \hat{z} + \mathbf{B_1}
+\eeq
+
+\beq
+\mathbf{u} = u_0 \hat{\phi} + \mathbf{u_1}
+\eeq
+
+The only term gained in the $(\hat{r}, \hat{z})$ component of the induction equation is $(B_0 \hat{z} \cdot \nabla) \mathbf{u_1}$, and Equation \ref{eq:A} becomes
+
+\beq
+ \label{eq:A_perturbed}
+\partial_t A = \frac{1}{r} J(A, \psi) + \frac{1}{\reym} \left( \nabla^2 A - \frac{2 \partial_r A}{r} \right) + B_0 \partial_z \Psi
+\eeq
 

### 2016-03-23 python/plots_allorders.py

@@ -916,10 +916,12 @@ def plot_paramspace(path="/Users/susanclark/weakly_nonlinear_mri/python/multirun
         jj = data.popitem()
         if not quiet:
             print(jj)
-        if np.isnan(jj[1]) == True:
+        #if np.isnan(jj[1]) is True:
+        if np.all(jj == np.nan) is True:
             evals[i] = None
         else:
-            evals[i] = jj[1]
+            #evals[i] = jj[1]
+            evals[i] = jj[1][0]
         #ids[i] = jj[0]
         
         allRms[i] = jj[0][0]

### 2016-03-23 paper/wide_gap_notes.tex
added perturbed B_phi equation
u_y perturbed equation
perturbed psi equation

@@
 
+Note: two things are currently missing. The constant $\beta$ from the nondimensionalization, and the Coriolis term in the momentum equation.
+
 We perturb the wide gap equations according to
@@
 \beq
-\mathbf{u} = u_0 \hat{\phi} + \mathbf{u_1}
+\mathbf{u} = -q \Omega_0 r \hat{\phi} + \mathbf{u_1}
 \eeq
+
+The $\Psi$ equation becomes
+
+\begin{multline}
+\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi \\
+= J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{r}B_\phi \partial_z B_\phi \\
++ \frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] + \frac{1}{r}B_0 \partial_z \left(\nabla^2 A - \frac{2}{r} \partial_r A\right)
+\end{multline}
+
+Equation \ref{eq:uy} becomes
+
+\begin{equation}
+  \label{eq:uy_perturbed}
+\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} - \frac{2}{r}q \Omega_0 \partial_z \Psi= \frac{J(A, B_\phi)}{r} + \frac{B_\phi \partial_z A}{r^2} + \frac{1}{\reye} \left( \nabla^2 \uphi - \frac{\uphi}{r}\right) + B_0 \partial_z B_\phi
+\end{equation}
 
@@
 \eeq
+
+Note that this is perfectly analogous to the thin-gap version of this equation.
+
+The $\hat{\phi}$ component of the induction equation, Equation \ref{eq:By}, becomes
+
+\begin{multline}
+  \label{eq:By_perturbed}
+\partial_t B_\phi = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)\\
++ \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A + \frac{1}{\reym} \left(\nabla^2 B_\phi - \frac{1}{r^2} B_\phi \right) + B_0 \partial_z u_\phi - \frac{1}{r} q \Omega_0 \partial_z A
+\end{multline}
 

### 2016-03-24 paper/wide_gap_notes.tex
nondimensionalization of JxB term.... see if you agree

@@
 \usepackage{cleveref}
+\usepackage{mathtools}
 
@@
 
+\section{Nondimensionalization}
+
+The momentum equation must be nondimensionalized. Here are the definitions of all of the dimensional components:
+
+\begin{equation}
+\begin{multlined}
+\widetilde{u} = \Omega_0 r_0 \delta u \\
+\widetilde{x} = \delta r_0 x \\
+\widetilde{\nabla} = \frac{\nabla}{\delta r_0}\\
+\widetilde{t} = \frac{t}{\Omega_0}\\ 
+\widetilde{B} = B_0 B \\
+\widetilde{P} = P_0 P \\
+\widetilde{\rho} = \rho_0 \rho 
+\end{multlined}
+\end{equation}
+
+If we define the velocity scale $v_0$ as the local sound speed scale $\left(c_{s0} \equiv \sqrt{\frac{P_0}{\rho_0}}\right)$, then $\Omega_0 r_0 \delta = \sqrt{\frac{P_0}{\rho_0}}$.
+
+Nondimensionalizing the momentum equation and dividing by a factor of $\Omega_0^2 r_0 \delta$ yields
+
+\beq
+\partial_t \mathbf{u} + \mathbf{u}\cdot \mathbf{\nabla} \mathbf{u} = \frac{1}{4 \pi \rho} \frac{B_0^2}{\rho_0 \Omega_0^2 r_0^2 \delta^2} (\mathbf{\nabla}\times\mathbf{B})\times\mathbf{B} + \frac{1}{\reye}\nabla^2 \mathbf{u} - 2 \mathbf{\Omega} \times \mathbf{u} - \mathbf{\Omega} \times (\mathbf{\Omega} \times \mathbf{r})
+\eeq
+
+If we define
+
+\beq
+\beta \equiv \frac{P}{P_{mag}} = \frac{\rho_0 \Omega_0^2 r_0^2 8 \pi}{B_0^2}
+\eeq
+
+where $P_{mag} = \frac{B_0^2}{8 \pi}$, then the factor in front of $(\mathbf{\nabla}\times\mathbf{B})\times\mathbf{B} $ should be $\frac{2}{\beta}$. Agreed?
+
 \section{Perturbed Equations}

### 2016-03-25 paper/wide_gap_notes.tex
added beta factor
wait, base velocity state should be different in wide gap

@@
 
-Note: two things are currently missing. The constant $\beta$ from the nondimensionalization, and the Coriolis term in the momentum equation.
+Note: currently missing the Coriolis term in the momentum equation. %The Coriolis and centrifugal terms add a $-2\partial_z u_\phi$ term to the LHS of the $\Psi$ equation and a $+2$
 
@@
 \beq
-\mathbf{B} = B_0 \hat{z} + \mathbf{B_1}
+\mathbf{B} = B_0 \hat{z} + \mathbf{B_1},
 \eeq
+
+which is the same perturbation we used in the thin-gap construction. But wait! We perturbed the thin-gap equations according to
 
@@
 
+but I don't think this is valid in the wide-gap case. To be sure, I'm going to perturb instead by
+
+\beq
+\mathbf{u} = r\Omega(r)\hat{\phi}+ \mathbf{u_1}
+\eeq
+
+where $\Omega(r) = \Omega_0 \left(\frac{r}{r_0}\right)^{-q} = \Omega_0 \left(r\right)^{-q}$ in dimensionless coordinates (but keeping $\Omega_0$ to flag a rotational term...) Agreed?
+
 The $\Psi$ equation becomes
@@
 \frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi \\
-= J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{r}B_\phi \partial_z B_\phi \\
-+ \frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] + \frac{1}{r}B_0 \partial_z \left(\nabla^2 A - \frac{2}{r} \partial_r A\right)
+= \frac{2}{\beta} J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi \\
++ \frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] + \frac{2}{\beta} \frac{1}{r}B_0 \partial_z \left(\nabla^2 A - \frac{2}{r} \partial_r A\right)
 \end{multline}
@@
   \label{eq:uy_perturbed}
-\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} - \frac{2}{r}q \Omega_0 \partial_z \Psi= \frac{J(A, B_\phi)}{r} + \frac{B_\phi \partial_z A}{r^2} + \frac{1}{\reye} \left( \nabla^2 \uphi - \frac{\uphi}{r}\right) + B_0 \partial_z B_\phi
+\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} - \frac{2}{r}q \Omega_0 \partial_z \Psi= \frac{2}{\beta} \frac{J(A, B_\phi)}{r} + \frac{2}{\beta} \frac{B_\phi \partial_z A}{r^2} + \frac{1}{\reye} \left( \nabla^2 \uphi - \frac{\uphi}{r}\right) + \frac{2}{\beta} B_0 \partial_z B_\phi
 \end{equation}

### 2016-03-25 paper/wide_gap_notes.tex
corrected velocity perturbations in Psi eqns
all perturbations

@@
 \begin{multline}
-\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi \\
+\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi - \partial_z\left(u_\phi \Omega(r) \right)\\
 = \frac{2}{\beta} J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi \\
@@
   \label{eq:uy_perturbed}
-\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} - \frac{2}{r}q \Omega_0 \partial_z \Psi= \frac{2}{\beta} \frac{J(A, B_\phi)}{r} + \frac{2}{\beta} \frac{B_\phi \partial_z A}{r^2} + \frac{1}{\reye} \left( \nabla^2 \uphi - \frac{\uphi}{r}\right) + \frac{2}{\beta} B_0 \partial_z B_\phi
+\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} + \partial_z\Psi\left( \frac{2}{r}\Omega(r) + \partial_r \Omega(r) \right) = \frac{2}{\beta} \frac{J(A, B_\phi)}{r} + \frac{2}{\beta} \frac{B_\phi \partial_z A}{r^2} + \frac{1}{\reye} \left( \nabla^2 \uphi - \frac{\uphi}{r}\right) + \frac{2}{\beta} B_0 \partial_z B_\phi
 \end{equation}
 
-The only term gained in the $(\hat{r}, \hat{z})$ component of the induction equation is $(B_0 \hat{z} \cdot \nabla) \mathbf{u_1}$, and Equation \ref{eq:A} becomes
+%$\frac{2}{r} \partial_z \Psi \Omega(r) + \partial_z\Psi \partial_r \Omega(r)
+%=\partial_z\Psi\left( \frac{2}{r}\Omega(r) + \partial_r \Omega(r) \right)$
+
+%\begin{equation}
+ % \label{eq:uy_perturbed}
+%\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} - \frac{2}{r}q \Omega_0 \partial_z \Psi= \frac{2}{\beta} \frac{J(A, B_\phi)}{r} + \frac{2}{\beta} \frac{B_\phi \partial_z A}{r^2} + \frac{1}{\reye} \left( \nabla^2 \uphi - \frac{\uphi}{r}\right) + \frac{2}{\beta} B_0 \partial_z B_\phi
+%\end{equation}
+
+%The only term gained in the $(\hat{r}, \hat{z})$ component of the induction equation is $(B_0 \hat{z} \cdot \nabla) \mathbf{u_1}$, and
+
+Equation \ref{eq:A} picks up $+\Omega(r) B_\phi$ from the term $- ({u_0}\hat{\phi} \cdot \nabla)\mathbf{B_1}$ and $-\Omega(r) B_\phi$ from the term $+(\mathbf{B_1} \cdot \nabla) \hat{\phi}$ and, so those cancel and the equation becomes
 
@@
 \partial_t B_\phi = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)\\
-+ \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A + \frac{1}{\reym} \left(\nabla^2 B_\phi - \frac{1}{r^2} B_\phi \right) + B_0 \partial_z u_\phi - \frac{1}{r} q \Omega_0 \partial_z A
++ \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A + \frac{1}{\reym} \left(\nabla^2 B_\phi - \frac{1}{r^2} B_\phi \right) + B_0 \partial_z u_\phi + \partial_z A \left( \frac{2}{r}\Omega(r) + \partial_r \Omega(r) \right) %- \frac{1}{r} q \Omega_0 \partial_z A
 \end{multline}

### 2016-03-25 paper/wide_gap_notes.tex
rotational terms

@@
 
-Note: currently missing the Coriolis term in the momentum equation. %The Coriolis and centrifugal terms add a $-2\partial_z u_\phi$ term to the LHS of the $\Psi$ equation and a $+2$
-
 We perturb the wide gap equations according to
@@
 
+We also add the Coriolis and centrifugal terms $-2 \mathbf{\Omega} \times \mathbf{u} - \mathbf{\Omega} \times \left(\mathbf{\Omega} \times \mathbf{r}\right)$ which expand as follows:
+
+\beq
+-2\Omega(r)\zhat \times \left( \mathbf{u_1} + r\Omega(r)\phihat\right) = -2\Omega(r)u_{r}\phihat + 2\Omega(r)u_\phi\rhat + 2r\Omega(r)^2\rhat
+\eeq
+
+\beq
+-\Omega(r)\zhat \times \left(\Omega(r)\zhat \times r\rhat\right) = +\Omega(r)^2 r \rhat
+\eeq
+
+The $\rhat$ terms will have $\partial_z$ applied to them, which will destroy the $3\Omega(r)^2 r \rhat$ term and so the righthand side of the $\Psi$ equation will ultimately gain only the term $2 \Omega(r) \partial_z u_\phi \rhat$.
+
 The $\Psi$ equation becomes
@@
 \frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi - \partial_z\left(u_\phi \Omega(r) \right)\\
-= \frac{2}{\beta} J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi \\
+= \frac{2}{\beta} J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi + 2 \Omega(r) \partial_z u_\phi \\
 + \frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] + \frac{2}{\beta} \frac{1}{r}B_0 \partial_z \left(\nabla^2 A - \frac{2}{r} \partial_r A\right)
@@
 
+Note that one of the terms gained from the base state in the $(\mathbf{u}\cdot \nabla)\mathbf{u}$ term can be combined with the term gained from the Coriolis term, but here I write them separately so we can check the equation.
 Equation \ref{eq:uy} becomes

### 2016-03-31 paper/wide_gap_notes.tex
all terms accounted for .. please check equations 28 - 31
wide gap nonlinear vector

@@
 The $\rhat$ terms will have $\partial_z$ applied to them, which will destroy the $3\Omega(r)^2 r \rhat$ term and so the righthand side of the $\Psi$ equation will ultimately gain only the term $2 \Omega(r) \partial_z u_\phi \rhat$.
+The $\phihat$ equation gains the term $-\frac{2}{r} \Omega(r) \partial_z \Psi$ on the righthand side.
 
@@
 \begin{multline}
+\label{eq:Psi_perturbed}
 \frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi - \partial_z\left(u_\phi \Omega(r) \right)\\
@@
 
-\begin{equation}
+\begin{multline}
   \label{eq:uy_perturbed}
-\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} + \partial_z\Psi\left( \frac{2}{r}\Omega(r) + \partial_r \Omega(r) \right) = \frac{2}{\beta} \frac{J(A, B_\phi)}{r} + \frac{2}{\beta} \frac{B_\phi \partial_z A}{r^2} + \frac{1}{\reye} \left( \nabla^2 \uphi - \frac{\uphi}{r}\right) + \frac{2}{\beta} B_0 \partial_z B_\phi
-\end{equation}
+\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} + \partial_z\Psi\left( \frac{2}{r}\Omega(r) + \partial_r \Omega(r) \right) = \\
+\frac{2}{\beta} \frac{J(A, B_\phi)}{r} + \frac{2}{\beta} \frac{B_\phi \partial_z A}{r^2} + \frac{1}{\reye} \left( \nabla^2 \uphi - \frac{\uphi}{r}\right) + \frac{2}{\beta} B_0 \partial_z B_\phi -\frac{2}{r} \Omega(r) \partial_z \Psi
+\end{multline}
 
@@
 
+\section{Matrix Formulation}
+
+This is all pending a rigorous check of Equations \ref{eq:Psi_perturbed} - \ref{eq:By_perturbed}, which Jeff is working on, but I'll start putting this into a matrix construction.
+
+First, the nonlinear vector, on the lefthand side of the equation, is
+
+\beq
+\mathbf{N} = \left[\begin{matrix}
+J\left(\Psi, \frac{1}{r^2} \left(\nabla^2 \Psi - \frac{2}{r}\partial_r \Psi \right) \right) - \frac{2}{\beta} J \left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_r A\right)\right) - \frac{2}{r} u_\phi \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r} B_\phi \partial_z B_\phi \\
+\frac{1}{r} J \left(\Psi, u_\phi\right) - \frac{1}{r} \frac{2}{\beta} J \left(A, B_\phi\right) + \frac{1}{r^2} u_\phi \partial_z \Psi - \frac{2}{\beta} \frac{1}{r^2} B_\phi \partial_z A \\
+-\frac{1}{r} J \left(A, \Psi\right)\\
+-\frac{1}{r} J \left(A, u_\phi\right) - \frac{1}{r} J \left(B_\phi, \Psi\right) - \frac{1}{r^2} B_\phi \partial_z \Psi + \frac{1}{r^2} u_\phi \partial_z A
+\end{matrix}\right] 
+\eeq
+
+Note that these differ from the thin-gap nonlinear terms not only because of the curvature terms in the Jacobians, but also because of the additional advective terms in all but the $\mathbf{A}$ equation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                              APPENDIX                                                             %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \appendix

### 2016-03-31 paper/wide_gap_notes.tex
dt terms
begin RHS terms

@@
 
+The $\partial_t$ terms are grouped together into
+
+\beq
+\partial_t D = \partial_t \left[\begin{matrix}
+\frac{1}{r}\nabla^2 - \frac{2}{r^2} \partial_r & 0 & 0 & 0\\
+0 & 1 & 0 & 0\\
+0 & 0 & 1 & 0\\
+0 & 0 & 0 & 1
+\end{matrix}\right]
+\eeq
+
+For the righthand side of the equation, I'll follow the same process that I did with the thin-gap.
+That is I'll first separate the matrices by physical meaning, and then regroup terms based on $z$-dependence.
+The RHS can be described by three matrices: one representing the rotational terms, one representing the background field terms, and one representing the viscous/resistive terms.
+
+\begin{multline}
+RHS = \left[\begin{matrix}
+3 \Omega(r) \partial_z u_\phi \\
+-\frac{4}{r} \Omega(r)\partial_z \Psi - \partial_r \Omega(r) \partial_z \Psi \\
+0 \\
+\frac{2}{r} \Omega(r) \partial_z A + \partial_r \Omega(r) \partial_z A
+\end{matrix}\right]
++ 
+\left[\begin{matrix}
+\frac{2}{\beta} \frac{1}{r} B_0 \partial_z \left(\nabla^2 A - \frac{2}{r} \partial_r A\right)\\
+\frac{2}{\beta} B_0 \partial_z B_\phi \\
+B_0 \partial_z \Psi\\
+B_0 \partial_z u_\phi
+\end{matrix}\right]
++
+\left[\begin{matrix}
+\frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4} \partial_r \Psi\right]\\
+\frac{1}{\reye} \left(\nabla^2 u_\phi - \frac{1}{r} u_\phi\right)\\
+\frac{1}{\reym} \left(\nabla^2 A - \frac{2}{r} \partial_r A\right)\\
+\frac{1}{\reym} \left(\nabla^2 B_\phi - \frac{1}{r^2} B_\phi\right)
+\end{matrix}\right] \\
+=
+\end{multline}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-04-01 paper/wide_gap_notes.tex
more wide gap rhs matrices
resistive/viscous term separated by order in z
B0 terms
L0 and L1

@@
 =
+\left[\begin{matrix}
+0 & 3\Omega(r) \partial_z & 0 & 0 \\
+-\frac{4}{r} \Omega(r) \partial_z - \partial_r \Omega(r) \partial_z & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & \frac{2}{r} \Omega(r) \partial_z + \partial_r \Omega(r) \partial_z & 0
+\end{matrix}\right]
++\\
+\left[\begin{matrix}
+0 & 0 & \frac{2}{\beta} \frac{1}{r} B_0 \partial_z \left(\nabla^2 - \frac{2}{r} \partial_r\right) & 0\\
+0 & 0 & 0 & \frac{2}{\beta} B_0 \partial_z\\
+B_0 \partial_z & 0 & 0 & 0 \\
+0 & B_0 \partial_z & 0 & 0
+\end{matrix}\right]
++ 
+\left[\begin{matrix}
+\frac{1}{\reye} \left(\nabla^2\left(\frac{1}{r}\nabla^2\right) - \frac{1}{r^3}\partial_r^2 - \frac{1}{r^4}\partial_r\right)\\
+\frac{1}{\reye} \left(\nabla^2  - \frac{1}{r} \right)\\
+\frac{1}{\reym} \left(\nabla^2 - \frac{2}{r} \partial_r \right)\\
+\frac{1}{\reym} \left(\nabla^2 - \frac{1}{r^2} \right)
+\end{matrix}\right]
 \end{multline}
+
+We separate the viscous/resistive matrix out in terms of $\partial_z$ dependence.
+
+\begin{multline}
+\mathrm{viscous}\, \,\mathrm{terms}  = \\
+ \left[\begin{matrix}
+\frac{1}{\reye} \left(-\frac{3}{r^4} \partial_r + \frac{3}{r^3} \partial_r^2 - \frac{2}{r^2} \partial_r^3 + \frac{1}{r} \partial_r^4\right) & 0 & 0 & 0\\
+0 & \frac{1}{\reye} \left(\partial_r^2 + \frac{1}{r} \partial_r - \frac{1}{r} \right) & 0 & 0 \\
+0 & 0 & \frac{1}{\reym} \left(\partial_r^2 - \frac{1}{r} \partial_r\right) & 0 \\
+0 & 0 & 0 & \frac{1}{\reym} \left(\partial_r^2 + \frac{1}{r}\partial_r - \frac{1}{r^2}\right)
+\end{matrix}\right] \\
++
+\partial_z^2 \left[\begin{matrix}
+\frac{1}{\reye}\left(-\frac{2}{r^2}\partial_r + \frac{2}{r}\partial_r^2\right) & 0 & 0 & 0 \\
+0 & \frac{1}{\reye} & 0 & 0 \\
+0 & 0 & \frac{1}{\reym} & 0 \\
+0 & 0 & 0 & \frac{1}{\reym}
+\end{matrix}\right] \\
++
+\partial_z^4\left[\begin{matrix}
+\frac{1}{\reye}\frac{1}{r} & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+\end{matrix}\right]
+\end{multline}
+
+Note that the form of each of the terms is different in the first matrix.
+
+We also separate the terms in the $B0$ matrix by $\partial_z$ dependence:
+
+\beq
+B_0 \, \mathrm{terms} = 
+\partial_z \left[\begin{matrix}
+0 & 0 & \frac{2}{\beta} \frac{1}{r} B_0 \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
+0 & 0 & 0 & \frac{2}{\beta}B_0\\
+B_0 & 0 & 0 & 0\\
+0 & B_0 & 0 & 0
+\end{matrix}\right] 
++
+\partial_z^3 \left[\begin{matrix}
+0 & 0 & \frac{2}{\beta} \frac{1}{r} B_0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+\end{matrix}\right]
+\eeq
+
+We replace $B_0$ with our lower magnetic field strength $(1 - \epsilon^2)$:
+
+\beq
+B_0 \, \mathrm{terms} = 
+(1 - \epsilon^2) \partial_z \left[\begin{matrix}
+0 & 0 & \frac{2}{\beta} \frac{1}{r} \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
+0 & 0 & 0 & \frac{2}{\beta}\\
+1 & 0 & 0 & 0\\
+0 & 1 & 0 & 0
+\end{matrix}\right] 
++
+(1 - \epsilon^2) \partial_z^3 \left[\begin{matrix}
+0 & 0 & \frac{2}{\beta} \frac{1}{r} & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+\end{matrix}\right]
+\eeq
+
+We now define, in analogy to the thin-gap matrices, the following:
+
+\beq
+L_0 = \left[\begin{matrix}
+\frac{1}{\reye} \left(-\frac{3}{r^4} \partial_r + \frac{3}{r^3} \partial_r^2 - \frac{2}{r^2} \partial_r^3 + \frac{1}{r} \partial_r^4\right) & 0 & 0 & 0\\
+0 & \frac{1}{\reye} \left(\partial_r^2 + \frac{1}{r} \partial_r - \frac{1}{r} \right) & 0 & 0 \\
+0 & 0 & \frac{1}{\reym} \left(\partial_r^2 - \frac{1}{r} \partial_r\right) & 0 \\
+0 & 0 & 0 & \frac{1}{\reym} \left(\partial_r^2 + \frac{1}{r}\partial_r - \frac{1}{r^2}\right)
+\end{matrix}\right]
+\eeq
+
+\beq
+L_1 = \left[\begin{matrix}
+0 & 3\Omega(r) & \frac{2}{\beta} \frac{1}{r} \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
+-\frac{4}{r} \Omega(r) - \partial_r \Omega(r) & 0 & 0 & \frac{2}{\beta} \\
+1 & 0 & 0 & 0 \\
+0 & 1 & \frac{2}{r} \Omega(r)  + \partial_r \Omega(r) & 0
+\end{matrix}\right]
+\eeq
 

### 2016-04-01 paper/wide_gap_notes.tex
L2, L3, L4
wide gap eqns all there

@@
 \beq
-L_0 = \left[\begin{matrix}
+\mathcal{L}_0 = \left[\begin{matrix}
 \frac{1}{\reye} \left(-\frac{3}{r^4} \partial_r + \frac{3}{r^3} \partial_r^2 - \frac{2}{r^2} \partial_r^3 + \frac{1}{r} \partial_r^4\right) & 0 & 0 & 0\\
@@
 \beq
-L_1 = \left[\begin{matrix}
+\mathcal{L}_1 = \left[\begin{matrix}
 0 & 3\Omega(r) & \frac{2}{\beta} \frac{1}{r} \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
@@
 \eeq
+
+\beq
+\mathcal{L}_2 = \left[\begin{matrix}
+\frac{1}{\reye}\left(-\frac{2}{r^2}\partial_r + \frac{2}{r}\partial_r^2\right) & 0 & 0 & 0 \\
+0 & \frac{1}{\reye} & 0 & 0 \\
+0 & 0 & \frac{1}{\reym} & 0 \\
+0 & 0 & 0 & \frac{1}{\reym}
+\end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{L}_3 = \left[\begin{matrix}
+0 & 0 & \frac{2}{\beta} \frac{1}{r} & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+\end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{L}_4 = \left[\begin{matrix}
+\frac{1}{\reye}\frac{1}{r} & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+\end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{G} = \left[\begin{matrix}
+0 & 3\Omega(r) & 0 & 0 \\
+-\frac{4}{r} \Omega(r)  - \partial_r \Omega(r)  & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & \frac{2}{r} \Omega(r)  + \partial_r \Omega(r)  & 0
+\end{matrix}\right]
+\eeq
+
+Defining
+
+\beq
+\mathcal{L}  = \mathcal{L}_0 + \mathcal{L}_1\partial_z + \mathcal{L}_2\partial_z^2 + \mathcal{L}_3\partial_z^3 + \mathcal{L}_4\partial_z^4
+\eeq
+
+and
+
+\beq
+\mathcal{\widetilde{G}} = - \mathcal{G} \partial_z - \mathcal{L}_3 \partial_z^3, 
+\eeq
+
+we have
+
+\beq
+\mathcal{D} \partial_t \mathbf{V} + \mathbf{N} = \mathcal{L}\mathbf{V} + \epsilon^2 \mathcal{\widetilde{G}} \mathbf{V}
+\eeq
+
+which we then expand:
+
+\beq
+\begin{split}
+ \partial_t & \rightarrow \epsilon^2 \partial_T \\
+ \partial_z & \rightarrow \partial_z + \epsilon \partial_Z
+\end{split}
+\eeq
+
+and perturb:
+
+\beq
+\mathbf{V} = \epsilon\mathbf{V}_1 + \epsilon^2 \mathbf{V}_2 + \epsilon^3 \mathbf{V}_3
+\eeq
+
+Note that as we have defined these matrices, the matrices are of course different than in the thin-gap case, and the nonlinear terms are different, but the equations for behavior at each order in $\epsilon$ are identical, that is
+
+\beq
+\mathcal{L} \mathbf{V}_1 = 0
+\eeq
+
+\beq
+\mathcal{L} \mathbf{V}_2 = \mathbf{N}_2 - \mathcal{\widetilde{L}}_1 \partial_Z \mathbf{V}_1
+\eeq
+
+\beq
+\mathcal{D}\partial_T\mathbf{V}_1 + \mathbf{N}_3 = \mathcal{L}\mathbf{V}_3 + \mathcal{\widetilde{L}}_1 \partial_Z \mathbf{V}_2 + \mathcal{\widetilde{L}}_2\partial_Z^2 \mathbf{V}_1 + \mathcal{\widetilde{G}}\mathbf{V}_1
+\eeq
+
+are still the correct equations to orders $\epsilon$, $\epsilon^2$, and $\epsilon^3$, respectively.
 

### 2016-04-10 paper/wide_gap_notes.tex
fixed G matrix
controlled gap width nondimensionalization -- see question

@@
 
+\subsection{Controlled gap width nondimensionalization}
+
+The idea is to parameterize the gap width rather than simply nondimensionalizing lengths by a point in the center of the channel $r_0$, as we did below.
+
+There are many ways to do this but one option is to define
+
+\beq
+\widetilde{r} = r_0 + \widetilde{x}
+\eeq
+
+Where tildes indicate dimensional quantities. Then for $\widetilde{x}$ to stay Cartesian, we need
+
+\beq
+\widetilde{r} - r_0 \ll r_2 - r_1 = d
+\eeq
+
+where $r_2 = r_1 = d$ is the gap width. Also define $\widetilde{x} \ll d$ and $d/r_0 << 1$ for the small gap approximation. Thus when we nondimensionalize,
+
+\beq
+r = \frac{\widetilde{r}}{r_0} = 1 + \frac{\widetilde{x}}{r_0}
+\eeq
+
+Note that we could control the gap width in terms of a small parameter $\epsilon x$ instead of a fraction $\frac{x}{r_0}$, but I believe both should work the same way.
+
+In the example Jeff worked out before, we looked at the base velocity profile in the thin gap limit.
+The rotation profile is given by $\Omega(r) = \Omega_0 r^{-q}$ in dimensionless quantities, and the background flow $u = r\Omega(r)$, so in our new notation:
+
+\beq
+u = \left(1 + \frac{\widetilde{x}}{r_0}\right) \Omega_0 \left(1 + \frac{\widetilde{x}}{r_0}\right)^{-q} = \Omega_0 \left(1 + \frac{\widetilde{x}}{r_0}\right)^{1-q}  
+\eeq
+
+Taylor expanding the exp(1-q) quantity, we arrive at
+
+\beq
+u = \Omega_0 \left(1 + \frac{\widetilde{x}}{r_0}\right) - \Omega_0 q \frac{\widetilde{x}}{r_0}
+\eeq
+
+Then if we define this with respect to a rest frame $\Omega_0 \left(1 + \frac{\widetilde{x}}{r_0}\right)$ and define a dimensionless $x = \widetilde{x}/r_0$, we get the Umurhan+ base flow $u = - \Omega_0 q x$.
+
+This makes sense to me but I'm having trouble extending it to the actual equations. This implies, for example, that the term $\frac{1}{\eta} \nabla^2 B$, which in the $\hat{\phi}$ equation is
+
+\beq
+\left[\nabla^2 B\right]_\phi = \nabla^2 B_\phi - \frac{B_\phi}{r^2},
+\eeq
+
+should become
+
+\beq
+\nabla^2 B_\phi - B_\phi \left(1 - \frac{2 \widetilde{x}}{r_0}\right)
+\eeq
+
+after substitution and Taylor expansion of the $r^{-2}$.
+But clearly there's no $- B_\phi \left(1 - \frac{2 \widetilde{x}}{r_0}\right)$ in the thin gap equation.
+What am I missing??
+Is this not the right place to be substituting in the new definition of r??
+
+\subsection{Typical nondimensionalization}
 The momentum equation must be nondimensionalized. Here are the definitions of all of the dimensional components:
@@
 \mathcal{G} = \left[\begin{matrix}
-0 & 3\Omega(r) & 0 & 0 \\
--\frac{4}{r} \Omega(r)  - \partial_r \Omega(r)  & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\
-0 & 0 & \frac{2}{r} \Omega(r)  + \partial_r \Omega(r)  & 0
+0 & 0 & \frac{2}{\beta} \frac{1}{r} \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
+0 & 0 & 0 & \frac{2}{\beta}\\
+1 & 0 & 0 & 0\\
+0 & 1 & 0 & 0
 \end{matrix}\right] 

### 2016-04-10 python/allorders_widegap.py

@@ -83,6 +83,7 @@ class MRI():
         
         problem.expand(domain)
         LEV = LinearEigenvalue(problem, domain)
+        print(LEV.pencils[0].L)
         LEV.solve(LEV.pencils[0])
         
         return LEV
@@ -209,18 +210,19 @@ class OrderE(MRI):
         r = domain.grid(0)
                 
         # equations defined in wide_gap_eqns.ipynb
-        # -1j*((D*dt)*V).subs(dz, 1j*Q) - (L*V).subs(dz, 1j*Q)
-        # Note dt terms multiplied by -1j as necessitated by Dedalus hard-coded temporal eigenvalue definition
+        # 1j*((D*dt)*V).subs(dz, 1j*Q) - (L*V).subs(dz, 1j*Q)
+        # Note dt terms multiplied by 1j as necessitated by Dedalus hard-coded temporal eigenvalue definition
+        
         #lv1.add_equation("1j*(1/r)*Q**2*dt(psi) + -1j*(1/r)*dt(psirr) + 1j*(1/r**2)*dt(psir) - 3*1j*Q*r**(-q)*u - iR*(Q**4/r)*psi + (2/beta)*1j*Q**3*(1/r)*A + 2*Q**2*iR*(1/r)*psirr - iR*2*Q**2*(1/r**2)*psir - (2/beta)*1j*Q*(1/r)*dr(Ar) + (2/beta)*1j*Q*(1/r**2)*Ar - iR*(1/r)*dr(psirrr) + iR*(2/r**2)*psirrr - iR*(3/r**3)*psirr + iR*(3/r**4)*psir = 0")
         #lv1.add_equation("-1j*dt(u) - (1j/r)*Q*q*r**(-q)*psi + 1j*(4/r)*Q*r**(-q)*psi + iR*Q**2*u - (2/beta)*1j*Q*B - iR*dr(ur) - iR*(1/r)*ur + iR*(1/r)*u = 0")
         #lv1.add_equation("-1j*dt(A) + iRm*Q**2*A - 1j*Q*psi - iRm*dr(Ar) + iRm*(1/r)*Ar = 0")
[…]

### 2016-04-11 paper/wide_gap_notes.tex
parasitic mode comment
expanding nonlinear terms: N2 psi
N3 psi
N2 u
N3 u
N2A and N3A
N2B and N3B
updated narrow to wide gap section for chain rule notes.

@@
 
+In order to recover the narrow gap, we take the transformation from dimensional, cylindrical coordinates $(\widetilde{r},\widetilde{z})$ to dimensioneless quantities $(x,z)$.
+In order to do so, we use $\widetilde{r} = r_0 (1 + \delta x)$, where $\delta = (r_{out} - r_{in})/r_0 \to 0$.
+Then, we recall that we must change all derivatives $\partial_{\widetilde{r}} \psi = \partial_x \psi \partial_{\widetilde{r}} x$ and likewise for $\partial_{\widetilde{z}}$.
+In order that the $x$ and $z$ parts of the equations come in at the same order, we must choose $\widetilde{z} = r_0 \delta z$.
+Then, $\partial_{\widetilde{r}} x = \delta^{-1}$ and $\partial_{\widetilde{z}} z = \delta^{-1}$.
+When you finally expand out all terms, you should see that everything that is not $\partial^2_x$ and $\partial^2_z$ will be at higher order in $\delta$, and thus will go to zero.
+For example, just taking the scalar Laplacian of $B_\phi$,
+
+\begin{equation}
+  \nabla^2 B_\phi = \frac{1}{r} \frac{\partial}{\partial r}\left(r \frac{\partial \psi}{\partial r} \right) + \frac{\partial^2 \psi}{\partial z^2}
+\end{equation}
+
+gives
+
+\begin{equation}
+  \frac{1}{(1 + \delta x)} \delta^{-2} \partial_x \left[(1 + \delta x) \partial_x B_\psi \right] + \delta^{-2} \partial_z^2 B_\phi,
+\end{equation}
+and using the binomial expansion,
+
+\begin{equation}
+  (1 - \delta x) \delta^{-2} \left(\partial_x^2 B_\psi + \delta \partial_x B_\psi + \delta x \partial_x^2 B_\psi \right) + \delta^{-2} \partial_z^2 B_\phi.
+\end{equation}
+
+When $\delta \to 0$, this leaves
+
+\begin{equation}
+  \partial_x^2 B_\phi + \partial_z^2 B_\phi,
+\end{equation}
+as expected. Of course, we'd also do $B_\phi \to B_y$ at the same time.
+
 \section{Nondimensionalization}
@@
 
-\beq
+\beq\label{eq:nonlinearvector}
 \mathbf{N} = \left[\begin{matrix}
@@
 
+\section{Expansion of Nonlinear Terms}
+
+Expanding the terms in Equation \ref{eq:nonlinearvector}, we find.
+
+\beq
+\mathbf{N} = \epsilon^2 \mathbf{N}_2 + \epsilon^3 \mathbf{N}_3
+\eeq
+
+\beq
+\begin{split}
+N_2^{\Psi} & = J\left(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1\right) + J\left(\Psi_1, -\frac{2}{r^3}\partial_r\Psi_1\right) \\
+&- \frac{2}{\beta} J \left(A_1, \frac{1}{r^2} \nabla^2 A_1\right) - \frac{2}{\beta} J\left(A_1, -\frac{2}{r^3} \partial_r A_1\right) - \frac{2}{r} u_1 \partial_z u_1 + \frac{2}{\beta} \frac{2}{r} B_1 \partial_z B_1
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_3^{\Psi} & = J\left(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_2\right) + J\left(\Psi_2, \frac{1}{r^2} \nabla^2\Psi_1\right) + 2 J \left(\Psi_1, \frac{1}{r^2}\partial_Z\partial_z \Psi_1\right) \\
+&+ J\left(\Psi_1, -\frac{2}{r^3}\partial_r \Psi_2\right) + J\left(\Psi_2, -\frac{2}{r^3}\partial_r \Psi_1\right) + \widetilde{J}\left(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1\right) + \widetilde{J} \left(\Psi_1, -\frac{2}{r^3}\partial_r \Psi_1\right)\\
+& - \frac{2}{\beta} J\left(A_1, \frac{1}{r^2}\nabla^2 A_2\right) - \frac{2}{\beta} J\left(A_2, \frac{1}{r^2}\nabla^2 A_1\right) - \frac{4}{\beta} J\left(A_1, \frac{1}{r^2}\partial_Z\partial_z A_1\right) - \frac{2}{\beta} J\left(A_1, -\frac{2}{r^3} \partial_r A_2 \right) \\
+& - \frac{2}{\beta} J\left(A_2, -\frac{2}{r^3} \partial_r A_1\right) - \frac{2}{\beta} \widetilde{J} \left(A_1, \frac{1}{r^2} \nabla^2 A_1\right) - \frac{2}{\beta} \widetilde{J} \left(A_1, -\frac{2}{r^3} \partial_r A_1\right) \\
+& - \frac{2}{r} u_1 \partial_z u_2 - \frac{2}{r} u_2 \partial_z u_1 - \frac{2}{r} u_1 \partial_Z u_1 + \frac{2}{\beta}\frac{2}{r} B_1\partial_z B_2 + \frac{2}{\beta}\frac{2}{r} B_2 \partial_z B_1 + \frac{2}{\beta} \frac{2}{r} B_1 \partial_Z B_1
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_2^{u} = \frac{1}{r} J\left(\Psi_1, u_1\right) - \frac{1}{r} \frac{2}{\beta} J\left(A_1, B_1\right) + \frac{1}{r^2} u_1 \partial_z \Psi_1 - \frac{2}{\beta}\frac{1}{r^2} B_1 \partial_z A_1
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_3^u & = \frac{1}{r}J\left(\Psi_1, u_2\right) + \frac{1}{r}J\left(\Psi_2, u_1\right) + \frac{1}{r}\widetilde{J} \left(\Psi_1, u_1\right) \\
+& - \frac{1}{r}\frac{2}{\beta} J\left(A_1, B_2\right) - \frac{1}{r} \frac{2}{\beta} J\left(A_2, B_1\right) - \frac{1}{r}\frac{2}{\beta}\widetilde{J}\left(A_1, B_1\right) \\
+& + \frac{1}{r^2} u_1\partial_z \Psi_2 + \frac{1}{r^2} u_2 \partial_z \Psi_1 + \frac{1}{r^2} u_1 \partial_Z \Psi_1 \\
+& - \frac{2}{\beta} \frac{1}{r^2} B_1 \partial_z A_2 - \frac{2}{\beta} \frac{1}{r^2} B_2 \partial_z A_1 - \frac{2}{\beta} \frac{1}{r^2} B_1 \partial_Z A_1
+\end{split}
+\eeq
+
+\beq
+N_2^A = -\frac{1}{r} J\left(A_1, \Psi_1\right)
+\eeq
+
+\beq
+N_3^A = -\frac{1}{r} J\left(A_1, \Psi_2\right) - \frac{1}{r}J\left(A_2, \Psi_1\right) - \frac{1}{r} \widetilde{J}\left(A_1, \Psi_1\right)
+\eeq
+
+\beq
+N_2^B = -\frac{1}{r} J\left(A_1, u_1\right) - \frac{1}{r} J\left(B_1, \Psi_1\right) - \frac{1}{r^2} B_1 \partial_z \Psi_1 + \frac{1}{r^2} u_1 \partial_z A_1
+\eeq
+
+\beq
+\begin{split}
+N_3^B & = - \frac{1}{r} J\left(A_1, u_2\right) - \frac{1}{r} J\left(A_2, u_1\right) - \frac{1}{r}\widetilde{J}\left(A_1, u_1\right) \\
+& - \frac{1}{r} J\left(B_1, \Psi_2\right) - \frac{1}{r} J\left(B_2, \Psi_1\right) - \frac{1}{r} \widetilde{J} \left(B_1, u_1\right) \\
+& - \frac{1}{r^2} B_1\partial_z \Psi_2 - \frac{1}{r^2} B_2 \partial_z \Psi_1 - \frac{1}{r^2} B_1 \partial_Z \Psi_1 \\
+& + \frac{1}{r^2} u_1 \partial_z A_2 + \frac{1}{r^2} u_2 \partial_z A_1 + \frac{1}{r^2} u_1 \partial_Z A_1
+\end{split}
+\eeq
+
+\section{Notes for paper}
+
+-- Comment on channel mode formation w.r.t. parasitic modes?
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \label{eq:del_scalar}
-  \nabla \psi = \frac{\partial \psi}{\partial r} \rhat + \frac{1}{r} \frac{\partial \psi}{\partial \phi} \phihat + \frac{\partial \psi}{\partial z} \zhat.
+  \nabla \psi = \frac{\partial \psi}{\partial r} \rhat + \frac{1}{r} \frac{\partial \psi}{\partial \phi} \phihat + \frac{\partial \psi}{\partial z} \zhat,
+\end{equation}
+
+and
+
+\begin{equation}
+  \label{eq:del2_scalar}
+  \nabla^2 \psi = \frac{1}{r} \frac{\partial}{\partial r}\left(r \frac{\partial \psi}{\partial r} \right) + \frac{1}{r^2} \frac{\partial^2 \psi}{\partial \phi^2} + \frac{\partial^2 \psi}{\partial z^2},
 \end{equation}

### 2016-04-12 paper/wide_gap_notes.tex
math notes url changed
thin gap limit of curl

@@
 \newcommand{\zhat}{\ensuremath{\mathbf{\hat{z}}}}
+\newcommand{\xhat}{\ensuremath{\mathbf{\hat{x}}}}
+\newcommand{\yhat}{\ensuremath{\mathbf{\hat{y}}}}
 
@@
 as expected. Of course, we'd also do $B_\phi \to B_y$ at the same time.
+
+Another operator we need is the curl $\nabla \times \mathbf{A}$,
+
+\beq
+\nabla \times \mathbf{A} = -\partial_{\widetilde{z}} A_\phi \rhat + \left(\partial_{\widetilde{z}} A_r - \partial_{\widetilde{r}} A_z \right) \phihat + \frac{1}{\widetilde{r}} \partial_{\widetilde{r}} \left( \widetilde{r} A_\phi\right) \zhat
+\eeq
+
+\beq
+= - \frac{1}{\delta} \partial_z A_\phi \rhat + \left(\frac{1}{\delta} \partial_z A_r - \frac{1}{\delta} \partial_x A_z\right) \phihat + \left(1 - \delta x\right) \frac{1}{\delta} \left(\partial_x A_\phi + \delta A_\phi + \delta x \partial_x A_\phi \right) \zhat
+\eeq
+
+Taking $\lim\limits_{\delta \to 0}$ of the above and changing variables $\phi \to y$, etc., we get
+
+\beq
+\nabla \times \mathbf{A} = -\partial_z A_y \xhat + \left(\partial_z A_x - \partial_x A_z\right) \yhat + \partial_x A_y \zhat
+\eeq
 
@@
 
-Everything here follows \url{http://farside.ph.utexas.edu/teaching/336L/Fluidhtml/node177.html#scyl}.
+Everything here follows \url{http://farside.ph.utexas.edu/teaching/336L/Fluidhtml/node257.html}.
 

### 2016-04-13 paper/wide_gap_notes.tex
thin gap reduction of A eqn, which apparently only works for unperturbed quantities (Why?)

@@
 \eeq
+
+The $A$ equation reduces as follows.
+
+\beq
+\partial_t A = \frac{1}{r} J (A, \Psi) + \frac{1}{\reym} \left(\nabla^2 A - \frac{2}{r}\partial_r A \right)
+\eeq
+
+\beq
+\partial_t A = (1 - \delta x)\delta^{-2} J(A, \Psi) + \frac{1}{\reym} \left[(1 - \delta x) \delta^{-2} \left(\partial_x^2 A + \delta\partial_x A + \delta x \partial_x^2 A \right) + \delta^{-2} \partial_z^2 A - \frac{2}{\delta}(1 - \delta x) \partial_x A \right]
+\eeq
+
+This reduces to
+
+\beq
+\partial_t A = J(A, \Psi) + \frac{1}{\reym} \nabla^2 A,
+\eeq
+
+which is the (unperturbed) thin-gap equation, when we drop all terms that are higher order than $\delta^{-2}$.
+Note that this means we have to drop terms of $\mathcal{O}(\delta^{-1})$, which means that this formalism cannot be applied to the perturbed equations (see Equation \ref{eq:A_perturbed}), because the term $B_0\partial_zA$ would drop out, but shouldn't.
+I'm also not sure how the $\partial_t$ term formally picks up a $\delta^{-2}$...?
 

### 2016-04-18 paper/wide_gap_notes.tex
time needs delta too to come in at same order
Bphi thin gap reduction
uphi thin gap reduction
notebooks for wide gap equations

@@
 \begin{multline}
+\label{eq:Psi_unperturbed}
 \frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi \\
@@
 
-The $A$ equation reduces as follows.
+The $u_\phi$ equation (Equation \ref{eq:uy}) reduces as follows:
+\begin{equation}
+\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} = \frac{J(A, B_\phi)}{r} + \frac{B_\phi \partial_z A}{r^2} + \nu \left( \nabla^2 \uphi - \frac{\uphi}{r}\right)
+\end{equation}
+
+\beq
+\begin{split}
+& \delta^{-2} \partial_t u_\phi + (1 - \delta r)\delta^{-2} J(\Psi, u_\phi) + (1 - 2\delta r)\delta^{-1} u_\phi \partial_z \Psi  = \\
+& (1 - \delta r)\delta^{-2} J(A, B_\phi) + (1 - 2\delta r)\delta^{-1} B_\phi \partial_z A + \\
+& \nu \left[(1 - \delta x) \delta^{-2} \left(\partial_x^2 u_\phi + \delta \partial_x u_\phi + \delta x \partial_x^2 u_\phi \right) + \delta^{-2} \partial_z^2 u_\phi \right] - \nu(1 - \delta r) u_\phi
+\end{split}
+\eeq
+
+\beq
+\partial_t u_\phi + J(\Psi, u_\phi) = J(A, B_\phi) + \nu(\partial_x^2 u_\phi + \partial_z^2 u_\phi)
+\eeq
+
+The $A$ equation reduces as follows. We also choose $\widetilde{t} = \delta^2 \Omega_0 t$ so that time comes in at the same order as the other dimensions...
 
@@
 \beq
-\partial_t A = (1 - \delta x)\delta^{-2} J(A, \Psi) + \frac{1}{\reym} \left[(1 - \delta x) \delta^{-2} \left(\partial_x^2 A + \delta\partial_x A + \delta x \partial_x^2 A \right) + \delta^{-2} \partial_z^2 A - \frac{2}{\delta}(1 - \delta x) \partial_x A \right]
+\delta^{-2} \partial_t A = (1 - \delta x)\delta^{-2} J(A, \Psi) + \frac{1}{\reym} \left[(1 - \delta x) \delta^{-2} \left(\partial_x^2 A + \delta\partial_x A + \delta x \partial_x^2 A \right) + \delta^{-2} \partial_z^2 A - \frac{2}{\delta}(1 - \delta x) \partial_x A \right]
 \eeq
@@
 Note that this means we have to drop terms of $\mathcal{O}(\delta^{-1})$, which means that this formalism cannot be applied to the perturbed equations (see Equation \ref{eq:A_perturbed}), because the term $B_0\partial_zA$ would drop out, but shouldn't.
-I'm also not sure how the $\partial_t$ term formally picks up a $\delta^{-2}$...?
+(??)
+
+The $B_\phi$ equation
+\begin{multline}
+\partial_t B_\phi = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)\\
++ \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A + \eta \left(\nabla^2 B_\phi - \frac{1}{r^2} B_\phi \right)
+\end{multline}
+
+\beq
+\begin{split}
+\delta^{-2} \partial_t B_\phi  & = (1 - \delta r) \delta^{-2} J(A, u_\phi) + (1 - \delta r) \delta^{-2} J(B_\phi, \Psi) \\
+& + (1 - 2 \delta r) \delta^{-1} B_\phi \partial_z \Psi - (1 - 2 \delta r) \delta^{-1} u_\phi \partial_z A + \\
+& \eta\left[(1 - \delta x) \delta^{-2} \left(\partial_x^2 B_\phi + \delta \partial_x B_\phi + \delta x \partial_x^2 B_\phi \right) + \delta^{-2} \partial_z^2 B_\phi\right] - \eta (1 - 2\delta r) B_\phi
+\end{split}
+\eeq
+
+\beq
+\partial_t B_\phi = J(A, u_\phi) + J(B_\phi, \Psi) + \eta(\partial_x^2 B_\phi + \partial_z^2 B_\phi)
+\eeq
+
+The $\Psi$ equation (Equation \ref{eq:Psi_unperturbed}) reduces as follows
+
+\beq
+\begin{split}
+& (1 - \delta x) \delta^{-2} \left[(1 - \delta x) \delta^{-2} \left(\partial_x^2 \Psi + \delta \partial_x \Psi + \delta x \partial_x^2 \Psi \right) + \delta^{-2} \partial_z^2 \Psi - 2(1 - \delta x)\delta^{-1}\partial_x \Psi \right] \\
+& + \delta^{-2} J(\Psi, (1-2\delta x)\left((1 - \delta x) \delta^{-2} \left(\partial_x^2 \Psi + \delta \partial_x \Psi + \delta x \partial_x^2 \Psi \right) + \delta^{-2} \partial_z^2 \Psi - 2(1 - \delta x)\delta^{-1}\partial_x \Psi \right)) \\
+& - (1 - \delta x)2\delta^{-1} u_\phi \partial_z u_\phi \\
+& = \delta^{-2} J(A, (1 - 2 \delta)\left((1 - \delta x) \delta^{-2} \left(\partial_x^2 \Psi + \delta \partial_x \Psi + \delta x \partial_x^2 \Psi \right) + \delta^{-2} \partial_z^2 \Psi - 2(1 - \delta x) \delta^{-1} A \right)) \\
+& - 2(1 - \delta x)\delta^{-1} B_\phi \partial_z B_\phi + \frac{1}{\reye} \left[\delta^{-2}(\partial_x^2 + \delta \partial_x + \delta x \partial_x^2 + \partial_z^2)((1 - \delta x)(\partial_x^2\Psi + \delta \partial_x\Psi + \delta x \partial_x^2\Psi + \partial_z^2\Psi))\right] \\
+& + \frac{1}{\reye}\left[(1 - 3\delta x)\delta^{-2} \partial_x^2 \Psi - (1 - 4 \delta x)\delta^{-1}\partial_x \Psi \right]
+\end{split}
+\eeq
+
+The viscous term is
+
+\beq
+\frac{1}{\reye} (1 - 
+\eeq
 

### 2016-04-23 paper/wide_gap_notes.tex

@@
 \title{Notes on Wide Gap}
+
+\section{Base Flow}
+In a viscous, Taylor-Couette device, there is only one base state that
+satisfies no-slip boundary conditions and the Navier-Stokes
+equations. This is called Couette flow, and can be written as
+\begin{equation}
+  \label{eq:couette_flow}
+  \Omega(r) = a + \frac{b}{r^2},
+\end{equation}
+where $a = (\Omega_2 r^2_2 - \Omega_1 r^2_1)/(r^2_2 - r^2_1)$ and
+$b = r^2_1 r^2_2 (\Omega_1 - \Omega_2)/(r^2_2 - r^2_1)$. Here,
+$\Omega_1$ is the rotation rate at the inner cylinder radius $r_1$,
+and likewise for the outer cylinder $r_2$.
+
+Keplerian flow is $\Omega \propto r^{-3/2}$; these appear to be
+incompatible. However, in Couette flow, we can approximate Keplerian
+flow by creating what is called ``quasi-Keplerian'' flow.
+
+Defining $q \equiv -d \ln \Omega/d \ln r$, we can see that for Couette flow,
+\begin{equation}
+  \label{eq:couette_q}
+  q \equiv -\frac{d \ln \Omega}{d \ln r} = \frac{2 b}{a r^2 + b}.
+\end{equation}
+Thus, assuming $r_1$ and $r_2$ are fixed by the experiment, through
+judicious choices of $\Omega_1$ and $\Omega_2$, we can make
+$q(r_0) = 3/2$, where $r_0$ is some reference radius. Goodman \& Ji do
+exactly this with $\bar{\zeta}$.
+
+So, we have to choose $\Omega(r) = a + b/r^2$; everywhere it appears in our equations.
 

### 2016-04-25 paper/wide_gap_notes.tex
added boundary conditions to wide gap notes

@@
 
+\section{Boundary Conditions}
+
+We use boundary conditions that are no-slip and perfectly conducting.
+No-slip means that all components of velocity are zero at the boundary, and perfectly conducting means that the radial component of the magnetic field and the axial component of the current are both zero at the boundary.
+
+No-slip means $u_r = u_\phi = u_z = 0$ and therefore $u_z = - \frac{1}{r}\partial_r \Psi = 0$ and $u_r = \frac{1}{r}\partial_z \Psi = 0$.
+
+Perfectly conducting means $J_z = (\nabla \times \mathbf{B})_z = B_\phi + r\partial_r B_\phi = 0$
+
+and also that $B_r = \frac{1}{r}\partial_z A = 0$.
+
+Thus our final boundary conditions are $\Psi = u = A = \partial_r \Psi = \partial_r(r B_\phi) = 0$, for a total of ten boundary conditions.
+
 \section{Notes for paper}

### 2016-04-25 python/allorders_widegap.py

@@ -14,11 +14,11 @@ from matplotlib import rc
 rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
 rc('text', usetex=True)
 
-nr1 = 64#512
+nr1 = 256#512
 r1 = de.Chebyshev('r', nr1, interval=(5, 15))
 d1 = de.Domain([r1])
 
-nr2 = 128#768
+nr2 = 512#768
 r2 = de.Chebyshev('r', nr2, interval=(5, 15))
 d2 = de.Domain([r2])
 
@@ -78,9 +78,9 @@ class MRI():
         #problem.add_bc('left(psi + r*psir) = 0')
         #problem.add_bc('right(psi + r*psir) = 0')
         problem.add_bc('left(psir) = 0')
-        problem.add_bc('right(psir) = 0') # axial component of current = 0
+        problem.add_bc('right(psir) = 0')
         problem.add_bc('left(B + r*Br) = 0')
-        problem.add_bc('right(B + r*Br) = 0')
+        problem.add_bc('right(B + r*Br) = 0') # axial component of current = 0
         
         return problem
     

### 2016-05-09 paper/wide_gap_notes.tex
bcs fixed
various limits of widegap linear eqns

@@
 
-Thus our final boundary conditions are $\Psi = u = A = \partial_r \Psi = \partial_r(r B_\phi) = 0$, for a total of ten boundary conditions.
+Thus our final boundary conditions are $\Psi = u = A = \partial_r \Psi = \partial_r(r B_\phi) = 0$, for a total of ten boundary condition equations.
 
@@
 -- Comment on channel mode formation w.r.t. parasitic modes?
+
+Discussion:
+
+The MRI is a complicated three-dimensional instability
 

### 2016-05-11 paper/wide_gap_notes.tex
fixed r perturbation to be dr(V) = 1j*kr*V
widegap matrices corrected in notes
coarser Rm search

@@
 \frac{1}{\reye} \left(-\frac{3}{r^4} \partial_r + \frac{3}{r^3} \partial_r^2 - \frac{2}{r^2} \partial_r^3 + \frac{1}{r} \partial_r^4\right) & 0 & 0 & 0\\
-0 & \frac{1}{\reye} \left(\partial_r^2 + \frac{1}{r} \partial_r - \frac{1}{r} \right) & 0 & 0 \\
+0 & \frac{1}{\reye} \left(\partial_r^2 + \frac{1}{r} \partial_r - \frac{1}{r^2} \right) & 0 & 0 \\
 0 & 0 & \frac{1}{\reym} \left(\partial_r^2 - \frac{1}{r} \partial_r\right) & 0 \\
@@
 
+%\beq
+%\mathcal{L}_1 = \left[\begin{matrix}
+%0 & 3\Omega(r) & \frac{2}{\beta} \frac{1}{r} \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
+%-\frac{4}{r} \Omega(r) - \partial_r \Omega(r) & 0 & 0 & \frac{2}{\beta} \\
+%1 & 0 & 0 & 0 \\
+%0 & 1 & \frac{2}{r} \Omega(r)  + \partial_r \Omega(r) & 0
+%\end{matrix}\right]
+%\eeq
+
+% work in inertial frame instead
 \beq
 \mathcal{L}_1 = \left[\begin{matrix}
-0 & 3\Omega(r) & \frac{2}{\beta} \frac{1}{r} \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
--\frac{4}{r} \Omega(r) - \partial_r \Omega(r) & 0 & 0 & \frac{2}{\beta} \\
+0 & \frac{2}{r} u_0 & \frac{2}{\beta} \frac{1}{r} \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
+(-\frac{1}{r^2} u_0 - \frac{1}{r} \partial_r u_0) & 0 & 0 & \frac{2}{\beta} \\
 1 & 0 & 0 & 0 \\
-0 & 1 & \frac{2}{r} \Omega(r)  + \partial_r \Omega(r) & 0
+0 & 1 & (\frac{1}{r}\partial_r u_0 - \frac{1}{r^2} u_0) & 0
 \end{matrix}\right]
@@
 
-The MRI is a complicated three-dimensional instability
+-- We locate threshold in the $\reym$ - $k_z$ plane.
+In the wide-gap case, the growth rate depends much more sensitively on $k_z$ than on $\reym$, while in the thin-gap case the opposite is true.
+We attribute this to the decreased influence of the boundary layers in the wide-gap geometry.
+(maybe?)
+
+The complex nature of the MRI means its properties are best parsed in idealized regimes or in simplified geometries. Many studies take
 

### 2016-05-11 python/widegap/parallel_parameter_exploration_widegap.py

@@ -288,14 +288,20 @@ if __name__ == '__main__':
     #Q = np.pi/10 
 
     dQ = 0.05
-    dRm = 0.05
+    #dQ = 0.1
+    #dRm = 0.05
+    dRm = 0.1
     
     #Qsearch = np.arange(0.6, 0.9, dQ)
     Qsearch = np.arange(0.2, 0.6, dQ)
     #Qsearch = np.arange(-0.2, 0.2, dQ)
     #Rmsearch = np.arange(
-    Rmsearch = np.arange(4.6, 5.1, dRm)
+    #Rmsearch = np.arange(4.6, 5.1, dRm)
+    #Rmsearch = np.arange(4.1, 4.6, dRm)
     #Rmsearch = np.arange(5.1, 6.0, dRm)
+    
+    #coarser Rm search
+    Rmsearch = np.arange(2.0, 4.1, dRm)
     
     #Qsearch = np.arange(0.2, 0.4, dQ)
     #Qsearch = np.arange(0.74, 0.76, dQ)

### 2016-05-12 paper/wide_gap_notes.tex
rederived nonlinear terms from scratch, they are correct in notes
plot of first-order velocity perturbations at Pm = 1E-3 added to widegap notes

@@
 
+% rigorously checked from scratch 5/11/16 -- correct.
 \beq\label{eq:nonlinearvector}
@@
 
+\section{Figures}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/u_near_threshold_test}
+\caption{Wide-gap first-order velocity perturbations near threshold.
+Unnormalized.
+This is with Q = 0.05, Rm = 0.85, but we now know that critical parameters are closer to Rm = 0.845, Q = 0.006 (finer grid running now).
+\label{fig:firstorder_u}}
+\end{figure}
+
 \section{Notes for paper}

### 2016-05-13 paper/wide_gap_notes.tex
preliminary B-field first order plot
notes on lack of neutral mode

@@
 
-\section{Figures}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/u_near_threshold_test}
-\caption{Wide-gap first-order velocity perturbations near threshold.
-Unnormalized.
-This is with Q = 0.05, Rm = 0.85, but we now know that critical parameters are closer to Rm = 0.845, Q = 0.006 (finer grid running now).
-\label{fig:firstorder_u}}
-\end{figure}
-
 \section{Notes for paper}
@@
 The complex nature of the MRI means its properties are best parsed in idealized regimes or in simplified geometries. Many studies take
+
+The purely conducting boundary condition states that the axial component of the current ($\mathbf{J}_z = [\nabla \times \mathbf{B}]_z)$ must be zero at the walls.
+In the thin gap geometry, the purely conducting boundary condition on the azimuthal magnetic field is $\partial_x(B_y) = 0$ for axisymmetric perturbations.
+A spatially constant azimuthal field satisfies both the thin-gap MRI equations and this boundary condition.
+This neutral mode is formally included in the analysis of Umurhan et al 2007 and yields a second amplitude equation in the form of a simple diffusion equation.
+This amplitude equation decouples from the GLE because of the translational symmetry of the thin-gap geometry.
+Because that symmetry is not preserved in the wide-gap case, Umurhan et al postulate that slow variation in the wide-gap geometry will be governed by two coupled amplitude equations.
+However, the wide-gap conducting boundary condition $J_z = \frac{1}{r} \partial_r (r B_\phi) = 0$, together with the purely geometric term in Equation \ref{eq:By_perturbed}, conspire to prevent the wide-gap geometry from sustaining a neutral mode.
 
@@
 
+\section{Figures}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/u_near_threshold_test}
+\caption{Wide-gap first-order velocity perturbations near threshold.
+Unnormalized.
+This is with Q = 0.05, Rm = 0.85, but we now know that critical parameters are closer to Rm = 0.845, Q = 0.006 (finer grid running now).
+\label{fig:firstorder_u}}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/B_near_threshold_test}
+\caption{Wide-gap first-order B-field perturbations near threshold.
+Unnormalized.
+This is with Q = 0.05, Rm = 0.85, but we now know that critical parameters are closer to Rm = 0.845, Q = 0.006 (finer grid running now).
+\label{fig:firstorder_B}}
+\end{figure}
+
 \end{document}

### 2016-05-13 paper/wide_gap_notes.tex
added preliminary N2 figure

@@
 
--- We locate threshold in the $\reym$ - $k_z$ plane.
+%-- We locate threshold in the $\reym$ - $k_z$ plane.
 In the wide-gap case, the growth rate depends much more sensitively on $k_z$ than on $\reym$, while in the thin-gap case the opposite is true.
@@
 
-The complex nature of the MRI means its properties are best parsed in idealized regimes or in simplified geometries. Many studies take
+%The complex nature of the MRI means its properties are best parsed in idealized regimes or in simplified geometries. Many studies take
 
@@
 However, the wide-gap conducting boundary condition $J_z = \frac{1}{r} \partial_r (r B_\phi) = 0$, together with the purely geometric term in Equation \ref{eq:By_perturbed}, conspire to prevent the wide-gap geometry from sustaining a neutral mode.
+
+-- coefficients of GLE come from inner products, therefore must always be real. Could we ever get a complex GLE?
 
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/N2_test}
+\caption{N2 from unnormalized lower-order components. This is with Rm = 0.845, Q = 0.006. Black = real, teal = imaginary. Why is psi not zero at the boundaries?
+\label{fig:firstorder_B}}
+\end{figure}
+
 \end{document}

### 2016-05-13 python/allorders_widegap.py

@@ -422,7 +422,10 @@ class OrderE2(MRI):
         else:
             MRI.__init__(self, Q = o1.Q, Rm = o1.Rm, Pm = o1.Pm, beta = o1.beta, R1 = o1.R1, R2 = o1.R2, Omega1 = o1.Omega1, Omega2 = o1.Omega2, norm = o1.norm)
             n2 = N2(o1 = o1, Q = o1.Q, Rm = o1.Rm, Pm = o1.Pm, beta = o1.beta, R1 = o1.R1, R2 = o1.R2, Omega1 = o1.Omega1, Omega2 = o1.Omega2, norm = o1.norm)
-    
+        
+        self.o1 = o1
+        self.n2 = n2
+        
         V21 = de.LBVP(d1,['psi','u', 'A', 'B', 'psir', 'psirr', 'psirrr', 'ur', 'Ar', 'Br'])
         V20 = de.LBVP(d1,['psi','u', 'A', 'B', 'psir', 'psirr', 'psirrr', 'ur', 'Ar', 'Br'])
         V22 = de.LBVP(d1,['psi','u', 'A', 'B', 'psir', 'psirr', 'psirrr', 'ur', 'Ar', 'Br'])

### 2016-05-13 python/widegap/parallel_parameter_exploration_widegap.py
coarse search at Pm = 1E-4

@@ -255,7 +255,8 @@ if __name__ == '__main__':
     """
     
     # Parameters approximating Umurhan+ 2007
-    Pm = 1.0E-3
+    #Pm = 1.0E-3
+    Pm = 1.0E-2
     beta = 25.0
     R1 = 5
     R2 = 15
@@ -287,15 +288,19 @@ if __name__ == '__main__':
     #Rm = 4.052031250000001
     #Q = np.pi/10 
 
-    dQ = 0.00001
-    #dQ = 0.1
+    #dQ = 0.00001
+    dQ = 0.001
     #dRm = 0.01
-    #dRm = 0.1
-    dRm = 0.001
+    dRm = 0.02
+    #dRm = 0.001
+    
+    # coarse search around Rm = 0.85, Q = 0.006
+    Rmsearch = np.arange(0.8, 1.0, dRm)
+    Qsearch = np.arange(0.002, 0.010, dQ)
     
     # search around Rm = 0.845, Q = 0.006
-    Rmsearch = np.arange(0.840, 0.851, dRm)
-    Qsearch = np.arange(0.005, 0.007, dQ)
+    #Rmsearch = np.arange(0.840, 0.851, dRm)
+    #Qsearch = np.arange(0.005, 0.007, dQ)
     
     # search around Rm = 0.85, Q = 0.01
     #Rmsearch = np.arange(0.84, 0.86, dRm)

### 2016-05-13 paper/wide_gap_notes.tex
coarse search at Pm = 1E-4
fixed typo in B field flux function plot

@@
 \centering
-\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/u_near_threshold_test}
-\caption{Wide-gap first-order velocity perturbations near threshold.
-Unnormalized.
-This is with Q = 0.05, Rm = 0.85, but we now know that critical parameters are closer to Rm = 0.845, Q = 0.006 (finer grid running now).
+\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/u_near_threshold_test_critical}
+\caption{Wide-gap first-order velocity perturbations near threshold. Unnormalized. This is with Rm = 0.845, Q = 0.006.
 \label{fig:firstorder_u}}
@@
 \centering
-\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/B_near_threshold_test}
-\caption{Wide-gap first-order B-field perturbations near threshold.
-Unnormalized.
-This is with Q = 0.05, Rm = 0.85, but we now know that critical parameters are closer to Rm = 0.845, Q = 0.006 (finer grid running now).
+\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/B_near_threshold_test_critical}
+\caption{Wide-gap first-order B-field perturbations near threshold. Unnormalized. This is with Rm = 0.845, Q = 0.006.
 \label{fig:firstorder_B}}
@@
 \centering
+\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/firstorderperturbations_test}
+\caption{First order perturbations, Rm = 0.845, Q = 0.006.
+\label{fig:firstorder_all}}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
 \includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/N2_test}
-\caption{N2 from unnormalized lower-order components. This is with Rm = 0.845, Q = 0.006. Black = real, teal = imaginary. Why is psi not zero at the boundaries?
-\label{fig:firstorder_B}}
+\caption{N2 from unnormalized lower-order components.
+This is with Rm = 0.845, Q = 0.006.
+Black = real, teal = imaginary.
+Why is psi not zero at the boundaries?
+Assuming it's because of the additional B advective term, not present in the thin-gap.
+\label{fig:N2}}
 \end{figure}

### 2016-05-15 python/widegap/parallel_parameter_exploration_widegap.py
finer grid search for Pm=1E-2

@@ -289,14 +289,18 @@ if __name__ == '__main__':
     #Q = np.pi/10 
 
     #dQ = 0.00001
-    dQ = 0.001
+    dQ = 0.0001
     #dRm = 0.01
-    dRm = 0.02
+    dRm = 0.001
     #dRm = 0.001
     
+    # For Pm = 1E-2, search around Rm = 0.86, Q = 0.002
+    Rmsearch = np.arange(0.85, 0.87, dRm)
+    Qsearch = np.arange(0.001, 0.003, dQ)
+    
     # coarse search around Rm = 0.85, Q = 0.006
-    Rmsearch = np.arange(0.8, 1.0, dRm)
-    Qsearch = np.arange(0.002, 0.010, dQ)
+    #Rmsearch = np.arange(0.8, 1.0, dRm)
+    #Qsearch = np.arange(0.002, 0.010, dQ)
     
     # search around Rm = 0.845, Q = 0.006
     #Rmsearch = np.arange(0.840, 0.851, dRm)

### 2016-05-16 python/allorders_widegap.py
N31 term added

@@ -566,7 +566,146 @@ class OrderE2(MRI):
         self.u22 = V22solver.state['u']
         self.A22 = V22solver.state['A']
         self.B22 = V22solver.state['B']
+  
+                self.psi20.name = "psi20"
+        self.u20.name = "u20"
+        self.A20.name = "A20"
+        self.B20.name = "B20"
+            
+        self.psi21.name = "psi21"
+        self.u21.name = "u21"
+        self.A21.name = "A21"
+        self.B21.name = "B21"
+                
+        self.psi22.name = "psi22"
+        self.u22.name = "u22"
+        self.A22.name = "A22"
+        self.B22.name = "B22"  
         
+        # Take relevant derivatives and complex conjugates
+        self.psi20_r = self.get_derivative(self.psi20)
+        self.psi20_rr = self.get_derivative(self.psi20_r)
+        self.psi20_rrr = self.get_derivative(self.psi20_rr)
+        
+        self.psi20_star = self.get_complex_conjugate(self.psi20)
+        
+        self.psi20_star_r = self.get_derivative(self.psi20_star)
+        self.psi20_star_rr = self.get_derivative(self.psi20_star_r)
+        self.psi20_star_rrr = self.get_derivative(self.psi20_star_rr)
+        
+        self.psi21_r = self.get_derivative(self.psi21)
+        self.psi21_rr = self.get_derivative(self.psi21_r)
+        
+        self.psi22_r = self.get_derivative(self.psi22)
+        self.psi22_rr = self.get_derivative(self.psi22_r)
+        self.psi22_rrr = self.get_derivative(self.psi22_rr)
+        
+        # u
+        self.u20_r = self.get_derivative(self.u20)
+        self.u20_star = self.get_complex_conjugate(self.u20)
+        self.u20_star_r = self.get_derivative(self.u20_star)
+        
+        self.u22_r = self.get_derivative(self.u22)
+        self.u22_star = self.get_complex_conjugate(self.u22)
[…]

### 2016-05-16 paper/wide_gap_notes.tex
finer grid search for Pm=1E-2
N31 term added
weakly nonlinear Latter Fromang and Faure eqns

@@
 However, the wide-gap conducting boundary condition $J_z = \frac{1}{r} \partial_r (r B_\phi) = 0$, together with the purely geometric term in Equation \ref{eq:By_perturbed}, conspire to prevent the wide-gap geometry from sustaining a neutral mode.
+We note that this mode would exist in a resistance-free approximation.
 
--- coefficients of GLE come from inner products, therefore must always be real. Could we ever get a complex GLE?
+The nonlinear terms in the wide-gap case differ from their thin-gap analogues mainly in the addition of vertical advective terms.
+These terms derive from the advective derivatives in the momentum and induction equations, but are filtered out in the thin-gap approximation.
+These terms have several interesting consequences.
+First, they allow the nonlinear contributions at both second and third order (i.e.
+N$_2$ and N$_3$) not to individually satisfy the boundary conditions on $\Psi$ and $u$.
+Second, the advection terms break the alternating parity exhibited by the thin-gap nonlinear terms.
 
@@
 \includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/u_near_threshold_test_critical}
-\caption{Wide-gap first-order velocity perturbations near threshold. Unnormalized. This is with Rm = 0.845, Q = 0.006.
+\caption{Wide-gap first-order velocity perturbations near threshold. Unnormalized. This is with Pm = 1E-3, Rm = 0.845, Q = 0.006.
 \label{fig:firstorder_u}}
@@
 \includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/B_near_threshold_test_critical}
-\caption{Wide-gap first-order B-field perturbations near threshold. Unnormalized. This is with Rm = 0.845, Q = 0.006.
+\caption{Wide-gap first-order B-field perturbations near threshold. Unnormalized. This is with Pm = 1E-3, Rm = 0.845, Q = 0.006.
 \label{fig:firstorder_B}}
@@
 \includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/firstorderperturbations_test}
-\caption{First order perturbations, Rm = 0.845, Q = 0.006.
+\caption{First order perturbations, Pm = 1E-3, Rm = 0.845, Q = 0.006.
 \label{fig:firstorder_all}}
@@
 \caption{N2 from unnormalized lower-order components.
-This is with Rm = 0.845, Q = 0.006.
+This is with Pm = 1E-3, Rm = 0.845, Q = 0.006.
 Black = real, teal = imaginary.
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/n3_test}
+\caption{N31 from unnormalized lower-order components. This is with Pm = 1E-3, Rm = 0.845, Q = 0.006. Black = real, teal = imaginary. 
+\label{fig:N31}}
+\end{figure}
+
 \end{document}

### 2016-05-18 paper/wide_gap_notes.tex
test contribution of nonlinear advective terms
still not at critical infl point

@@
 N$_2$ and N$_3$) not to individually satisfy the boundary conditions on $\Psi$ and $u$.
-Second, the advection terms break the alternating parity exhibited by the thin-gap nonlinear terms.
+
+The astrophysical magnetorotational instability operates in accretion disks and in stellar interiors.
+In accretion disks, differential rotation is imposed gravitationally by a central body.
+In a Taylor-Couette flow, differential rotation is created by rotating an inner cylinder faster than an outer cylinder.
+In a non ideal fluid
+
+-- MRI mode can be conceptualized as a quantum mechanical particle in a box
+-- In the ideal thin-gap case, V = constant, and thus invariant under parity. Walls and boundary conditions impose symmetric and antisymmetric states.
+-- In the ideal wide-gap case, V is a function of r
+-- I expect the coefficients of the GLE to be complex in the wide-gap case..
+
+%Second, the advection terms break the alternating parity exhibited by the thin-gap nonlinear terms. % no they don't
 

### 2016-05-18 python/widegap/parallel_parameter_exploration_widegap.py

@@ -289,7 +289,7 @@ if __name__ == '__main__':
     #Q = np.pi/10 
 
     #dQ = 0.00001
-    dQ = 0.0001
+    dQ = 0.001
     #dRm = 0.01
     dRm = 0.001
     #dRm = 0.001
@@ -304,8 +304,11 @@ if __name__ == '__main__':
     #Rmsearch = np.arange(0.82, 0.85, dRm)
     
     #Qsearch = np.arange(0.006, 0.008, dQ)
-    Rmsearch = np.arange(0.84, 0.85, dRm)
-    Qsearch = np.arange(0.008, 0.012, dQ)
+    #Rmsearch = np.arange(0.84, 0.85, dRm)
+    #Qsearch = np.arange(0.008, 0.012, dQ)
+    
+    Rmsearch = np.arange(0.84, 0.842, dRm)
+    Qsearch = np.arange(0.012, 0.022, dQ)
     
     # coarse search around Rm = 0.85, Q = 0.006
     #Rmsearch = np.arange(0.8, 1.0, dRm)

### 2016-05-18 paper/wide_gap_notes.tex
still finding Pm=1E-2 critical params

@@
 However, the wide-gap conducting boundary condition $J_z = \frac{1}{r} \partial_r (r B_\phi) = 0$, together with the purely geometric term in Equation \ref{eq:By_perturbed}, conspire to prevent the wide-gap geometry from sustaining a neutral mode.
-We note that this mode would exist in a resistance-free approximation.
+We note that a neutral mode of the form $B_\phi(r) \propto \frac{1}{r}$ would exist in a resistance-free approximation.
 
@@
 In a Taylor-Couette flow, differential rotation is created by rotating an inner cylinder faster than an outer cylinder.
-In a non ideal fluid
+For a non ideal fluid, the base flow (describe what's currently in section 1)...
 
 -- MRI mode can be conceptualized as a quantum mechanical particle in a box
+
 -- In the ideal thin-gap case, V = constant, and thus invariant under parity. Walls and boundary conditions impose symmetric and antisymmetric states.
+
 -- In the ideal wide-gap case, V is a function of r
+
 -- I expect the coefficients of the GLE to be complex in the wide-gap case..

### 2016-05-19 python/thinsims/animate_sims.py
code to make movies of thin gap simulations

@@ -1,0 +1,65 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib import animation
+import h5py
+
+from matplotlib import rc
+rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+rc('text', usetex=True)
+
+inpath = "/Users/susanclark/weakly_nonlinear_mri/python/thinsims/"
+mrirun = "MRI_run_Rm5.00e+00_eps0.00e+00_Pm1.00e-04_beta2.50e+01_Q7.49e-01_qsh1.50e+00_Omega1.00e+00_nz256/"
+
+slicepath = inpath + mrirun + "/slices/"
+
+fig = plt.figure()
+ax1 = fig.add_subplot(121)
+ax2 = fig.add_subplot(122)
+
+ax1.set_title(r"$\Psi$", size = 20)
+ax2.set_title(r"$u_y$", size = 20)
+
+"""
+for slice_i in xrange(1000, 1123, 1):
+
+    fn = slicepath + "slices_s" + str(slice_i) + ".h5"
+    slice = h5py.File(fn, 'r')
+    
+    psi = slice['tasks']['psi'][0, :, :]
+    u = slice['tasks']['u'][0, :, :]
+    A = slice['tasks']['A'][0, :, :]
+    B = slice['tasks']['b'][0, :, :]
+    
+    ax1.imshow(u)
+    plt.show()
+"""
+
+def init():
+    im1 = ax1.imshow(np.zeros((256, 256), np.float_))
+    im2 = ax2.imshow(np.zeros((256, 256), np.float_))
+    
+    return im1, im2
+    
+def animate(i):
+
+    slice_i = i + 1000
+    
+    try:
+        fn = slicepath + "slices_s" + str(slice_i) + ".h5"
+        slice = h5py.File(fn, 'r')
+    except IOError:
+        print fn
+    
+    psi = slice['tasks']['psi'][0, :, :]
+    u = slice['tasks']['u'][0, :, :]
+    
+    im1 = ax1.imshow(psi, cmap = "RdBu")
+    im2 = ax2.imshow(u, cmap = "RdBu")
+
+    return im1, im2
+    
+anim = animation.FuncAnimation(fig, animate, init_func=init,
+                               frames=122, interval=20, blit=True)
+
+anim.save('test_anim_psi_u.mp4', fps=30, extra_args=['-vcodec', 'libx264'])
+    

### 2016-06-08 paper/widegap_mri_paper.tex
starting to place wide gap notes into paper format and add exposition between eqns

\documentclass[onecolumn]{emulateapj}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\hyd}    {{\rm H}}
\newcommand{\vobs}{v_{\rm obs}}
\newcommand{\hi}{H{\sc i}~}
\newcommand{\hii}{H{\sc ii}~}
\newcommand{\hia}{H{\sc i}}
\newcommand{\hiia}{H{\sc ii}}

\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{color}
%\usepackage{appendix}
\usepackage[titletoc,title]{appendix}
\usepackage{cleveref}

\makeatletter
\setlength{\@fptop}{0pt}
\setlength{\@fpbot}{0pt}
\setlength{\@fpsep}{0pt}
\makeatother

\newcommand{\citei}[1]{\citeauthor{#1} \citeyear{#1}}
\newcommand{\unit}[1]{\textrm{ #1}}
\newcommand{\kms}{km ${\rm s^{-1}}$}
\newcommand{\kmsa}{km ${\rm s^{-1}}$}
\newcommand{\citeia}[2]{\citeauthor{#1}~(\citeyear{#1};~#2)}

\newcommand\reye{\mathrm{Re}}
\newcommand\reym{\mathrm{Rm}}

\newcommand{\uphi}{\ensuremath{u_\phi}}
\newcommand{\rhat}{\ensuremath{\mathbf{\hat{r}}}}
\newcommand{\phihat}{\ensuremath{\mathbf{\hat{\phi}}}}
\newcommand{\zhat}{\ensuremath{\mathbf{\hat{z}}}}
\newcommand{\xhat}{\ensuremath{\mathbf{\hat{x}}}}
\newcommand{\yhat}{\ensuremath{\mathbf{\hat{y}}}}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\begin{document}

\title{The Weakly Nonlinear Magnetorotational Instability in a Wide-Gap Channel Flow}
\author{Clark, S.E.\altaffilmark{1}}
\author{Oishi, J.S. \altaffilmark{2,}\altaffilmark{3}}
%\author{Mac Low, M.-M.\altaffilmark{1,}\altaffilmark{2}}
\altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
\altaffiltext{2}{Department of Astrophysics, American Museum of Natural History, New York, NY}
\altaffiltext{3}{Department of Physics, SUNY Farmingdale}

\begin{abstract}
Wide gap formulation and preliminary results
\end{abstract}

\section{Introduction}

\section{Wide Gap Equations}

Many investigations of the MRI use the ``thin gap" approximation, in which the radial extent of the fluid channel is taken to be much smaller than the radius of curvature. That is, for a channel center $r_0$ bounded by inner and outer radii $r_1$ and $r_2$, respectively, the thin gap approximation applies when $r_0 \gg (r_2 - r_1)$. The thin gap approximation simplifies the MRI equations by excluding curvature terms, because the flow through a thin gap can be taken to be approximately linear in $\phi$, i.e. Cartesian. In this work we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.

The basic equations solved are the momentum and induction equations,

\beq\label{momentum}
%\begin{split}
\partial_t \mathbf{u} + \mathbf{u} \cdot \nabla \mathbf{u} = -\frac{1}{\rho}\nabla P - \nabla\Phi + \frac{1}{\rho} \left(\mathbf{J}\times\mathbf{B}\right) + \nu\nabla^2 \mathbf{u} 
%\end{split}
\eeq

and

\beq\label{induction}
\partial_t \mathbf{B} = \nabla \times \left(\mathbf{u} \times \mathbf{B}\right) + \eta\nabla^2\mathbf{B},
\eeq

where $P$ is the gas pressure, $\nu$ is the kinematic viscosity, $\eta$ is the microscopic diffusivity, $\nabla\Phi$ is the gravitational force per unit mass, and the current density is $\mathbf{J} = \nabla\times\mathbf{B}$. We solve these equations subject to the incompressible fluid and solenoidal magnetic field constraints,

\beq
\label{eq:incompressibility}
\nabla \cdot \mathbf{u} = 0
\eeq

and

\beq
\label{eq:solenoid}
\nabla \cdot \mathbf{B} = 0.
\eeq

We perturb these equations axisymmetrically in a cylindrical $(r, \phi, z)$ geometry, i.e. $\mathbf{u} = \mathbf{u_0} + \mathbf{u_1}$ and $\mathbf{B} = \mathbf{B_0} + \mathbf{B_1}$, where $\mathbf{u_0}$ and $\mathbf{B_0}$ are defined below. We define a Stokes stream function $\Psi$ such that

\beq
  \label{eq:stokes}
  \mathbf{u_1} \, = \, \left[\begin{matrix}
\frac{1}{r} \partial_z \Psi\ \rhat\\
\uphi \ \phihat\\
-\frac{1}{r} \partial_r \Psi\ \zhat\\
\end{matrix}\right],
\eeq

and define the magnetic vector potential $A$ analogously. These definitions automatically satisfy Equations \ref{eq:incompressibility} and \ref{eq:solenoid} for axisymmetric disturbances.

The astrophysical magnetorotational instability operates in accretion disks and in stellar interiors, environments where fluid rotation is strongly regulated by gravity. In accretion disks, differential rotation is imposed gravitationally by a central body, so the rotation profile is forced to be Keplerian. Clearly a gravitationally enforced Keplerian flow is inaccessible to laboratory study, so differential rotation is created by rotating an inner cylinder faster than an outer cylinder (a Taylor-Couette setup). For a nonideal fluid subject to no-slip boundary conditions, the base flow is

\begin{equation}
  \label{eq:couette_flow}
  \Omega(r) = a + \frac{b}{r^2},
\end{equation}
where $a = (\Omega_2 r^2_2 - \Omega_1 r^2_1)/(r^2_2 - r^2_1)$, $b = r^2_1 r^2_2 (\Omega_1 - \Omega_2)/(r^2_2 - r^2_1)$, and $\Omega_1$ and $\Omega_2$ are the rotation rates at the inner and outer cylinder radii, respectively. In the laboratory, $r_1$ and $r_2$ are typically fixed by experimental design. However $\Omega_1$ and $\Omega_2$ may be chosen such that the flow in the center of the channel is approximately Keplerian. Defining a shear parameter $q$, we see that for Couette flow,
\begin{equation}
  \label{eq:couette_q}
  q(r) \equiv -\frac{d \ln \Omega}{d \ln r} = \frac{2 b}{a r^2 + b}.
\end{equation}

Thus through judicious choice of cylinder rotation rates, we can set $q(r_0) = 3/2$, for quasi-Keplerian flow. Note that in the thin gap approximation we have the freedom to choose any rotation profile, as the narrow gap width imposes a linear shear (constant $q$). Our base velocity is $\mathbf{u_0} = r\Omega(r) \phihat$. We initialize an axial magnetic field $\mathbf{B_0} = B_0 \zhat$.

Our perturbed system is

\beq
\begin{split}
\label{eq:Psi_perturbed}
\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z \left(\nabla^2 A - \frac{2}{r} \partial_r A\right) - \frac{2}{r}u_0 \partial_z u_\phi - \frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] \\
= - J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) + \frac{2}{\beta} J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi  + \frac{2}{r} u_\phi \partial_z u_\phi \\
\end{split}
\eeq

\begin{align}
& \partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} \left( \nabla^2 \uphi - \frac{1}{r^2} \uphi \right) = \frac{2}{\beta} \frac{1}{r} J(A, B_\phi) - \frac{1}{r} J(\Psi, \uphi) + \frac{2}{\beta}\frac{1}{r^2} B_\phi \partial_z A - \frac{1}{r^2} \uphi \partial_z \Psi \label{eq:uphi_perturbed} \\
& \partial_t A - B_0 \partial_z \Psi - \frac{1}{\reym} \left( \nabla^2 A - \frac{2}{r} \partial_r A \right)= \frac{1}{r} J(A, \psi)  \label{eq:A_perturbed}\\
  \label{eq:Bphi_perturbed}
& \partial_t B_\phi + \frac{1}{r^2} u_0 \partial_z A - B_0 \partial_z u_\phi - \frac{1}{r} \partial_r u_0 \partial_z A - \frac{1}{\reym} \left(\nabla^2 B_\phi - \frac{1}{r^2} B_\phi \right) = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)
+ \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A 
\end{align}

where J is the Jacobian $J(x, y) \equiv \partial_z x \partial_r y - \partial_r x \partial_z y$. Note that in the above, $\nabla^2 x \equiv \partial_r^2 x + \partial_z^2 x + \frac{1}{r} \partial_r x$. Equations \ref{eq:Psi_perturbed} - \ref{eq:Bphi_perturbed} are nondimensionalized, where lengths have been scaled by $r_0$, velocities by $r_0 \Omega_0$, densities by $\rho_0$, and magnetic fields by $B_0$; where $B_0$ appears in the above it is formally unity. $\Omega_0 = \Omega(r_0)$ is the rotation rate at the center of the channel. We introduce the Reynolds number $\reye = \Omega_0 r_0^2/\nu$, the magnetic Reynolds number $\reym = \Omega_0 r_0^2 / \eta$, and a plasma beta parameter $\beta = \Omega_0^2 r_0^2 \rho_0/B_0^2$. If we define the dimensional cylindrical coordinate $r = r_0(1 + \delta x)$, the thin gap equations are recovered in the limit $\delta \rightarrow 0$.

We solve the system subject to periodic vertical boundary conditions and no-slip, perfectly conducting radial boundary conditions.

\section{Weakly Nonlinear Perturbation Analysis}

Just as in the weakly nonlinear analyses of \citei{Umurhan:2007hs} and \citei{Clark:2016}, we tune the system away from marginality by taking $B_0 \rightarrow B_0\left(1 - \epsilon^2\right)$, where $\epsilon \ll 1$. We parameterize scale separation as $Z = \epsilon z$ and $T = \epsilon^2 t$, where $Z$ and $T$ are slowly varying spatial and temporal scales, respectively. We group the fluid variables into a state vector $\mathbf{V} = \left[\Psi, u, A, B\right]^{\mathrm{T}}$ and expand the variables in a perturbation series $\mathbf{V} = \epsilon \mathbf{V}_1 + \epsilon^2 \mathbf{V}_2 + \epsilon^2 \mathbf{V}_3 + h.o.t.$ The perturbed system can then be expressed at each order by the equations

\begin{align}
\mathcal{O}(\epsilon)&: \mathcal{L} \mathbf{V}_1 + \mathcal{D} \partial_t \mathbf{V}_1 = 0 \label{eq:ordere}\\
\mathcal{O}(\epsilon^2)&: \mathcal{L}_1 \mathbf{V}_2 + \widetilde{\mathcal{L}} \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \label{eq:ordere2}\\
\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3. \label{eq:ordere3}
\end{align}

Here, $\mathcal{L}$, $\mathcal{D}$, etc. are matrices defined in Appendix \ref{app:basic_equations}, and $\mathbf{N}_2$ and $\mathbf{N}_3$ are vectors of nonlinear terms. We emphasize that Equations \ref{eq:ordere} - \ref{eq:ordere3} have the same form as these equations in the thin gap case, although the matrices are significantly different in this wide gap formulation. This is because we do not have slow variation in the radial dimension. The slow variation in $Z$ and $T$ are parameterized as an amplitude function $\alpha(Z, T)$ which modulates the flow in these dimensions. This parameterization coupled with the boundary conditions lead us to an ansatz linear solution $\mathbf{V}_1 = \alpha(Z, T) \mathbb{V}_{11}(r) e^{i k_z z} + c.c.$, where the radial variation is contained in $\mathbb{V}_{11}$. We defer full expressions and derivations thereof to the Appendix, and focus here on features of the wide gap system which distinguish it from the thin gap limit.

First, the result of the weakly nonlinear analysis is a single amplitude equation for $\alpha$. We find

\beq
 \label{eq:gle}
\partial_T \alpha = b \alpha + d \nabla^2 \alpha - c \alpha \left|\alpha^2\right|,
\eeq

a Ginzburg-Landau equation (e.g. \cite{Aranson:2002}). Here we note a departure from the behavior of the thin-gap system. The purely conducting boundary condition states that the axial component of the current ($\mathbf{J}_z = [\nabla \times \mathbf{B}]_z)$ must be zero at the walls. In the thin gap geometry, the purely conducting boundary condition on the azimuthal magnetic field is $\partial_x(B_y) = 0$ for axisymmetric perturbations. A spatially constant azimuthal field satisfies both the thin-gap MRI equations and this boundary condition. This neutral mode is formally included in the analysis of \citei{Umurhan:2007hs} and yields a second amplitude equation in the form of a simple diffusion equation. This amplitude equation decouples from the GLE because of the translational symmetry of the thin-gap geometry. Because that symmetry is not preserved in the wide-gap case, \citeauthor{Umurhan:2007hs} postulate that slow variation in the wide-gap geometry will be governed by two coupled amplitude equations. However, the wide-gap conducting boundary condition $J_z = \frac{1}{r} \partial_r (r B_\phi) = 0$, together with the purely geometric term in Equation \ref{eq:Bphi_perturbed}, conspire to prevent the wide-gap geometry from sustaining a neutral mode. We note that a neutral mode of the form $B_\phi(r) \propto \frac{1}{r}$ would exist in a resistance-free approximation.

The preservation of symmetries in the thin-gap geometry is worth a closer look, as its absence in the wide gap case is the source of many differences in the systems. \citei{Latter:2015} point out that in the ideal limit ($\nu, \eta \rightarrow 0$), the system described by Equations \ref{eq:Psi_perturbed} - \ref{eq:Bphi_perturbed} can be expressed as a Shr{\"o}dinger equation for the radial velocity. We can do the same thing for $\Psi$, and in the thin-gap limit this becomes

\beq
\partial_x^2 \Psi + k_z^2 \mathrm{U}(x) \Psi = 0
\eeq

where $\mathrm{U}(x) = {3}/{v_A^2 k_z^2} + 1$ at marginality. When no-slip radial boundary conditions are applied, the thin-gap MRI system resembles a particle in a box with a radially constant potential well. Thus thin-gap MRI modes must be eigenstates of parity. In the wide gap case, the ``potential" $\mathrm{U}(r)$ is a function of $r$, so symmetric and antisymmetric modes are no longer required.

\section{Channel Modes}
Channel modes are radially independent linear MRI modes which are, under certain conditions, exact nonlinear solutions to the MRI equations.

\clearpage
\appendix

\section{A. Detailed Equations}\label{app:basic_equations}

\bibliographystyle{plain}

\begin{thebibliography}{53}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{Aranson \& Kramer(2002)}]{Aranson:2002}
Aranson, I.S. \& Kramer, L., 2002, Rev. Mod. Phys. 74, 99

\bibitem[{Clark \& Oishi(2016a)}]{Clark:2016}
Clark, S.E. and Oishi, J.S., 2016, in prep

\bibitem[{Latter et al.(2015)}]{Latter:2015}
Latter, H.N., Fromang, S., \& Faure, J., 2015, MNRAS 453, 3257

\bibitem[{Umurhan et al.(2007a)}]{Umurhan:2007dz}
Umurhan, O.M., Regev, O., Menou, K., 2007, Phys. Rev. Letters, 98, 034501

\bibitem[{Umurhan et al.(2007b)}]{Umurhan:2007hs}
Umurhan, O.M., Regev, O., Menou, K., 2007, Phys. Rev. E, 76, 036310

\end{thebibliography}

\end{document}

### 2016-06-08 python/thinsims/animate_sims.py

@@ -2,43 +2,41 @@ import numpy as np
 import matplotlib.pyplot as plt
 from matplotlib import animation
 import h5py
+import pylab
+import streamplot_uneven as su
 
 from matplotlib import rc
 rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
 rc('text', usetex=True)
 
-inpath = "/Users/susanclark/weakly_nonlinear_mri/python/thinsims/"
-mrirun = "MRI_run_Rm5.00e+00_eps0.00e+00_Pm1.00e-04_beta2.50e+01_Q7.49e-01_qsh1.50e+00_Omega1.00e+00_nz256/"
+inpath = "/Volumes/DataDavy/MRI/"
+#mrirun = "MRI_run_Rm5.00e+00_eps0.00e+00_Pm1.00e-04_beta2.50e+01_Q7.49e-01_qsh1.50e+00_Omega1.00e+00_nz256/"
+mrirun = "MRI_run_Rm5.00e+00_eps0.00e+00_Pm1.00e-02_beta2.50e+01_Q7.40e-01_qsh1.50e+00_Omega1.00e+00_nz128"
 
 slicepath = inpath + mrirun + "/slices/"
 
+# Set up figure for movie
 fig = plt.figure()
-ax1 = fig.add_subplot(121)
-ax2 = fig.add_subplot(122)
+ax1 = fig.add_subplot(221)
+ax2 = fig.add_subplot(222)
+ax3 = fig.add_subplot(223)
+ax4 = fig.add_subplot(224)
 
 ax1.set_title(r"$\Psi$", size = 20)
 ax2.set_title(r"$u_y$", size = 20)
+ax3.set_title(r"$A$", size = 20)
+ax4.set_title(r"$B$", size = 20)
 
-"""
-for slice_i in xrange(1000, 1123, 1):
-
-    fn = slicepath + "slices_s" + str(slice_i) + ".h5"
-    slice = h5py.File(fn, 'r')
-    
-    psi = slice['tasks']['psi'][0, :, :]
-    u = slice['tasks']['u'][0, :, :]
-    A = slice['tasks']['A'][0, :, :]
-    B = slice['tasks']['b'][0, :, :]
-    
-    ax1.imshow(u)
-    plt.show()
-"""
+runsize = 128
 
 def init():
-    im1 = ax1.imshow(np.zeros((256, 256), np.float_))
-    im2 = ax2.imshow(np.zeros((256, 256), np.float_))
[…]

### 2016-06-09 paper/widegap_mri_paper.tex
more on the symmetries

@@
 The preservation of symmetries in the thin-gap geometry is worth a closer look, as its absence in the wide gap case is the source of many differences in the systems.
-\citei{Latter:2015} point out that in the ideal limit ($\nu, \eta \rightarrow 0$), the system described by Equations \ref{eq:Psi_perturbed} - \ref{eq:Bphi_perturbed} can be expressed as a Shr{\"o}dinger equation for the radial velocity.
-We can do the same thing for $\Psi$, and in the thin-gap limit this becomes
+\citei{Latter:2015} point out that in the ideal limit ($\nu, \eta \rightarrow 0$), the linearized system described by the lefthand side of Equations \ref{eq:Psi_perturbed} - \ref{eq:Bphi_perturbed} can be expressed as a Shr{\"o}dinger equation for the radial velocity.
+Similarly combining equations to obtain a single expression for $\Psi$, we find that the thin-gap limit linear ideal MRI can be expressed as
 
@@
 When no-slip radial boundary conditions are applied, the thin-gap MRI system resembles a particle in a box with a radially constant potential well.
-Thus thin-gap MRI modes must be eigenstates of parity.
+Thus thin-gap linear MRI modes must be eigenstates of parity.
+These symmetries are preserved in the nonlinear MRI terms because they are nonlinear combinations of lower-order eigenfunctions.
 In the wide gap case, the ``potential" $\mathrm{U}(r)$ is a function of $r$, so symmetric and antisymmetric modes are no longer required.

### 2016-06-09 paper/widegap_mri_paper.tex
nonlinear terms appendix

@@
 \label{eq:Psi_perturbed}
-\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z \left(\nabla^2 A - \frac{2}{r} \partial_r A\right) - \frac{2}{r}u_0 \partial_z u_\phi - \frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] \\
-= - J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) + \frac{2}{\beta} J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi  + \frac{2}{r} u_\phi \partial_z u_\phi \\
+\frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] \\
+= - J(\Psi, \frac{1}{r^2} ( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi) ) + \frac{2}{\beta} J(A, \frac{1}{r^2} ( \nabla^2 A - \frac{2}{r} \partial_rA) ) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi  + \frac{2}{r} u_\phi \partial_z u_\phi \\
 \end{split}
@@
 \begin{align}
-& \partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} \left( \nabla^2 \uphi - \frac{1}{r^2} \uphi \right) = \frac{2}{\beta} \frac{1}{r} J(A, B_\phi) - \frac{1}{r} J(\Psi, \uphi) + \frac{2}{\beta}\frac{1}{r^2} B_\phi \partial_z A - \frac{1}{r^2} \uphi \partial_z \Psi \label{eq:uphi_perturbed} \\
-& \partial_t A - B_0 \partial_z \Psi - \frac{1}{\reym} \left( \nabla^2 A - \frac{2}{r} \partial_r A \right)= \frac{1}{r} J(A, \psi)  \label{eq:A_perturbed}\\
+& \partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} ( \nabla^2 \uphi - \frac{1}{r^2} \uphi ) = \frac{2}{\beta} \frac{1}{r} J(A, B_\phi) - \frac{1}{r} J(\Psi, \uphi) + \frac{2}{\beta}\frac{1}{r^2} B_\phi \partial_z A - \frac{1}{r^2} \uphi \partial_z \Psi \label{eq:uphi_perturbed} \\
+& \partial_t A - B_0 \partial_z \Psi - \frac{1}{\reym} ( \nabla^2 A - \frac{2}{r} \partial_r A )= \frac{1}{r} J(A, \psi)  \label{eq:A_perturbed}\\
   \label{eq:Bphi_perturbed}
-& \partial_t B_\phi + \frac{1}{r^2} u_0 \partial_z A - B_0 \partial_z u_\phi - \frac{1}{r} \partial_r u_0 \partial_z A - \frac{1}{\reym} \left(\nabla^2 B_\phi - \frac{1}{r^2} B_\phi \right) = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)
+& \partial_t B_\phi + \frac{1}{r^2} u_0 \partial_z A - B_0 \partial_z u_\phi - \frac{1}{r} \partial_r u_0 \partial_z A - \frac{1}{\reym} (\nabla^2 B_\phi - \frac{1}{r^2} B_\phi ) = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)
 + \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A 
@@
 We parameterize scale separation as $Z = \epsilon z$ and $T = \epsilon^2 t$, where $Z$ and $T$ are slowly varying spatial and temporal scales, respectively.
-We group the fluid variables into a state vector $\mathbf{V} = \left[\Psi, u, A, B\right]^{\mathrm{T}}$ and expand the variables in a perturbation series $\mathbf{V} = \epsilon \mathbf{V}_1 + \epsilon^2 \mathbf{V}_2 + \epsilon^2 \mathbf{V}_3 + h.o.t.$ The perturbed system can then be expressed at each order by the equations
+We group the fluid variables into a state vector $\mathbf{V} = \left[\Psi, u, A, B\right]^{\mathrm{T}}$, such that the full nonlinear system can be expressed as
+
+\beq\label{eq:unperturbed_matrix_eqns}
+\mathcal{D}\partial_t\mathbf{V} + \mathcal{L}\mathbf{V} + \epsilon^2 \widetilde{\mathcal{G}} \mathbf{V} + \mathbf{N} = 0,
+\eeq
+
+where $\mathcal{D}$, $\mathcal{L}$, and $\widetilde{\mathcal{G}}$ are matrices defined in Appendix \ref{app:basic_equations}, and $\mathbf{N}$ is a vector containing all nonlinear terms.
+We then expand the variables in a perturbation series $\mathbf{V} = \epsilon \mathbf{V}_1 + \epsilon^2 \mathbf{V}_2 + \epsilon^2 \mathbf{V}_3 + h.o.t.$ The perturbed system can then be expressed at each order by the equations
 
@@
 \mathcal{O}(\epsilon)&: \mathcal{L} \mathbf{V}_1 + \mathcal{D} \partial_t \mathbf{V}_1 = 0 \label{eq:ordere}\\
-\mathcal{O}(\epsilon^2)&: \mathcal{L}_1 \mathbf{V}_2 + \widetilde{\mathcal{L}} \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \label{eq:ordere2}\\
+\mathcal{O}(\epsilon^2)&: \mathcal{L} \mathbf{V}_2 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \label{eq:ordere2}\\
 \mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3.
@@
 
-Here, $\mathcal{L}$, $\mathcal{D}$, etc.
-are matrices defined in Appendix \ref{app:basic_equations}, and $\mathbf{N}_2$ and $\mathbf{N}_3$ are vectors of nonlinear terms.
-We emphasize that Equations \ref{eq:ordere} - \ref{eq:ordere3} have the same form as these equations in the thin gap case, although the matrices are significantly different in this wide gap formulation.
+See Appendix \ref{app:basic_equations} for the definition of matrices and nonlinear vectors, and a thorough derivation.
+We emphasize that Equations \ref{eq:ordere} - \ref{eq:ordere3} have the same form as these equations in the thin gap case, although the matrices (which contain all radial derivatives) are significantly different in this wide gap formulation.
 This is because we do not have slow variation in the radial dimension.
@@
 
+The nonlinear terms, detailed in Appendix \ref{app:nonlinear_terms}, represent an interesting departure from the thin-gap theory.
+The thin-gap nonlinear terms at both second and third orders are linear combinations of Jacobians.
+The nonlinear terms in the wide-gap case differ from their thin-gap analogues with in the addition of vertical advective terms.
+These terms derive from the advective derivatives in the momentum and induction equations, but are filtered out in the thin-gap approximation.
+These advective terms allow the nonlinear contributions at both second and third order (i.e.
+N$_2$ and N$_3$) not to individually satisfy the boundary conditions on $\Psi$ and $u$.
+(I'm just guessing but this may lead to higher torque on inner cylinder -- should calculate)
+
 \section{Channel Modes}
@@
 \section{A. Detailed Equations}\label{app:basic_equations}
+
+\section{B. Nonlinear Terms}\label{app:nonlinear_terms}
+
+Here we detail the perturbative expansion of the nonlinear vector $\mathbf{N}$ in Equation \ref{eq:unperturbed_matrix_eqns}.
+
+\beq
+\mathbf{N} = \epsilon^2 \mathbf{N}_2 + \epsilon^3 \mathbf{N}_3
+\eeq
+
+\beq
+N_2^{\Psi}  = J(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1) + J(\Psi_1, -\frac{2}{r^3}\partial_r\Psi_1)
+- \frac{2}{\beta} J (A_1, \frac{1}{r^2} \nabla^2 A_1) - \frac{2}{\beta} J(A_1, -\frac{2}{r^3} \partial_r A_1) - \frac{2}{r} u_1 \partial_z u_1 + \frac{2}{\beta} \frac{2}{r} B_1 \partial_z B_1
+\eeq
+
+\beq
+\begin{split}
+N_2^{u} = \frac{1}{r} J\left(\Psi_1, u_1\right) - \frac{1}{r} \frac{2}{\beta} J\left(A_1, B_1\right) + \frac{1}{r^2} u_1 \partial_z \Psi_1 - \frac{2}{\beta}\frac{1}{r^2} B_1 \partial_z A_1
+\end{split}
+\eeq
+
+\beq
+N_2^A = -\frac{1}{r} J\left(A_1, \Psi_1\right)
+\eeq
+
+\beq
+N_2^B = -\frac{1}{r} J\left(A_1, u_1\right) - \frac{1}{r} J\left(B_1, \Psi_1\right) - \frac{1}{r^2} B_1 \partial_z \Psi_1 + \frac{1}{r^2} u_1 \partial_z A_1
+\eeq
+
+\beq
+\begin{split}
+N_3^{\Psi} & = J(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_2) + J(\Psi_2, \frac{1}{r^2} \nabla^2\Psi_1) + 2 J (\Psi_1, \frac{1}{r^2}\partial_Z\partial_z \Psi_1) + J(\Psi_1, -\frac{2}{r^3}\partial_r \Psi_2) + J(\Psi_2, -\frac{2}{r^3}\partial_r \Psi_1) \\
+& + \widetilde{J}(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1) + \widetilde{J} (\Psi_1, -\frac{2}{r^3}\partial_r \Psi_1) - \frac{2}{\beta} J(A_1, \frac{1}{r^2}\nabla^2 A_2) - \frac{2}{\beta} J(A_2, \frac{1}{r^2}\nabla^2 A_1) - \frac{4}{\beta} J(A_1, \frac{1}{r^2}\partial_Z\partial_z A_1) \\ & - \frac{2}{\beta} J(A_1, -\frac{2}{r^3} \partial_r A_2 )
+ - \frac{2}{\beta} J(A_2, -\frac{2}{r^3} \partial_r A_1) - \frac{2}{\beta} \widetilde{J} (A_1, \frac{1}{r^2} \nabla^2 A_1) - \frac{2}{\beta} \widetilde{J} (A_1, -\frac{2}{r^3} \partial_r A_1) \\
+& - \frac{2}{r} u_1 \partial_z u_2 - \frac{2}{r} u_2 \partial_z u_1 - \frac{2}{r} u_1 \partial_Z u_1 + \frac{2}{\beta}\frac{2}{r} B_1\partial_z B_2 + \frac{2}{\beta}\frac{2}{r} B_2 \partial_z B_1 + \frac{2}{\beta} \frac{2}{r} B_1 \partial_Z B_1
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_3^u & = \frac{1}{r}J\left(\Psi_1, u_2\right) + \frac{1}{r}J\left(\Psi_2, u_1\right) + \frac{1}{r}\widetilde{J} \left(\Psi_1, u_1\right) - \frac{1}{r}\frac{2}{\beta} J\left(A_1, B_2\right) - \frac{1}{r} \frac{2}{\beta} J\left(A_2, B_1\right) - \frac{1}{r}\frac{2}{\beta}\widetilde{J}\left(A_1, B_1\right) \\
+& + \frac{1}{r^2} u_1\partial_z \Psi_2 + \frac{1}{r^2} u_2 \partial_z \Psi_1 + \frac{1}{r^2} u_1 \partial_Z \Psi_1 - \frac{2}{\beta} \frac{1}{r^2} B_1 \partial_z A_2 - \frac{2}{\beta} \frac{1}{r^2} B_2 \partial_z A_1 - \frac{2}{\beta} \frac{1}{r^2} B_1 \partial_Z A_1
+\end{split}
+\eeq
+
+\beq
+N_3^A = -\frac{1}{r} J\left(A_1, \Psi_2\right) - \frac{1}{r}J\left(A_2, \Psi_1\right) - \frac{1}{r} \widetilde{J}\left(A_1, \Psi_1\right)
+\eeq
+
+\beq
+\begin{split}
+N_3^B & = - \frac{1}{r} J\left(A_1, u_2\right) - \frac{1}{r} J\left(A_2, u_1\right) - \frac{1}{r}\widetilde{J}\left(A_1, u_1\right) - \frac{1}{r} J\left(B_1, \Psi_2\right) - \frac{1}{r} J\left(B_2, \Psi_1\right) - \frac{1}{r} \widetilde{J} \left(B_1, u_1\right) \\ & - \frac{1}{r^2} B_1\partial_z \Psi_2 - \frac{1}{r^2} B_2 \partial_z \Psi_1 - \frac{1}{r^2} B_1 \partial_Z \Psi_1 + \frac{1}{r^2} u_1 \partial_z A_2 + \frac{1}{r^2} u_2 \partial_z A_1 + \frac{1}{r^2} u_1 \partial_Z A_1
+\end{split}
+\eeq
 

### 2016-06-15 paper/wide_gap_notes.tex
code to make movies of thin gap simulations
different streamfunction equations added (for hmri/smri case) to widegap notes

@@
 
+\section{Equations in more succinct form}
+
+Defining the stream functions instead as
+
+\begin{equation}
+  \label{eq:stokes}
+  \mathbf{u} \, = \, \left[\begin{matrix}
+\partial_z \psi\ \rhat\\
+\uphi \ \phihat\\
+- \partial_r^* \psi\ \zhat\\
+\end{matrix}\right]
+\end{equation}
+
+where $\partial_r^* = \partial_r + \frac{1}{r}$
+
+And defining the function $D^2 = \nabla^2 - \frac{1}{r^2}$
+
+We obtain a very compact form for the equations. The following include helical MRI terms, and reduce to the SMRI when $\xi \rightarrow 0$.
+
+\beq
+\partial_t D^2 \Psi - \frac{2}{r}u_0\partial_z u_\phi - \frac{2}{\beta} B_0 \partial_z D^2 A + \frac{2}{\beta} \frac{2}{r^2} \xi B_0 \partial_z B_\phi - \frac{1}{\reye} D^4 \Psi = 0
+\eeq
+
+\beq
+\partial_t u_\phi + \frac{1}{r} u_0 \partial_z \Psi + \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} D^2 u_{\phi} = 0
+\eeq
+
+\beq
+\partial_t A - B_0 \partial_z \Psi - \frac{1}{\reym} D^2 A = 0
+\eeq
+
+\beq
+\partial_t B_\phi + \frac{1}{r}u_0 \partial_z A - B_0 \partial_z u_\phi - \partial_r u_0 \partial_z A - \frac{2}{r^2} \xi B_0 \partial_z \Psi - \frac{1}{\reym} D^2 B_\phi = 0
+\eeq
+
 \section{Notes for paper}
@@
 
--- In the ideal thin-gap case, V = constant, and thus invariant under parity. Walls and boundary conditions impose symmetric and antisymmetric states.
+-- In the ideal thin-gap case, V = constant, and thus solutions must be eigenstates of parity. Walls and boundary conditions impose symmetric and antisymmetric states.
 
@@
 -- I expect the coefficients of the GLE to be complex in the wide-gap case..
+
+Channel modes:
+-- exact nonlinear solutions in local approximation,
 

### 2016-07-08 paper/widegap_mri_paper.tex
different streamfunction equations added (for hmri/smri case) to widegap notes
modified widegap param search for hmri

@@
 
-and define the magnetic vector potential $A$ analogously. These definitions automatically satisfy Equations \ref{eq:incompressibility} and \ref{eq:solenoid} for axisymmetric disturbances.
+and define the magnetic vector potential $A$ analogously.
+These definitions automatically satisfy Equations \ref{eq:incompressibility} and \ref{eq:solenoid} for axisymmetric disturbances.
+We note that in the linearized equations, streamfunctions of the form $u_x = \partial_z \Psi$, $u_z = -(\partial_r + \frac{1}{r}) \Psi$, and the corresponding definitions of the magnetic vector potential, are convenient choices, but this does not hold true for the nonlinear terms.
 
@@
  \label{eq:gle}
-\partial_T \alpha = b \alpha + d \nabla^2 \alpha - c \alpha \left|\alpha^2\right|,
+\partial_T \alpha = b \alpha + d \partial_Z^2 \alpha - c \alpha \left|\alpha^2\right|,
 \eeq
@@
 These symmetries are preserved in the nonlinear MRI terms because they are nonlinear combinations of lower-order eigenfunctions.
-In the wide gap case, the ``potential" $\mathrm{U}(r)$ is a function of $r$, so symmetric and antisymmetric modes are no longer required.
+In the wide gap case, the ``potential" $\mathrm{U}(r)$ varies with $r$, so symmetric and antisymmetric modes are no longer required.
 

### 2016-07-08 python/widegap/find_crit.py

@@ -37,7 +37,7 @@ comm = MPI.COMM_WORLD
 #Rm = 4.052
 
 
-def find_crit(R1, R2, Omega1, Omega2, beta, Pm, Rm_min, Rm_max, k_min, k_max, n_Rm, n_k, nr):
+def find_crit(R1, R2, Omega1, Omega2, beta, xi, Pm, Rm_min, Rm_max, k_min, k_max, n_Rm, n_k, nr):
     Rm = Rm_min
     k = k_min
     R = Rm/Pm
@@ -62,7 +62,7 @@ def find_crit(R1, R2, Omega1, Omega2, beta, Pm, Rm_min, Rm_max, k_min, k_max, n_
     widegap.parameters['c2'] = c2
     widegap.parameters['beta'] = beta
     widegap.parameters['B0'] = 1
-    widegap.parameters['xi'] = 0
+    widegap.parameters['xi'] = xi
 
     widegap.substitutions['ru0'] = '(r*r*c1 + c2)' # u0 = r Omega(r) = Ar + B/r
     widegap.substitutions['rrdu0'] = '(c1*r*r-c2)' # du0/dr = A - B/r^2

### 2016-07-08 python/widegap/find_widegap_crit.py

@@ -12,6 +12,7 @@ Options:
     --Omega2=<Omega2>          outer cylinder rotation rate (units of 1/s) [default: 67.0631]
     --Pm=<Pm>                  magnetic Prandtl number [default: 1e-3]
     --beta=<beta>              plasma beta [default: 25.]
+    --xi=<xi>                  helical base field strength for HMRI [default: 0.]
     --Rm_min=<Rm_min>          minimum magnetic Reynolds number [default: 0.5]
     --Rm_max=<Rm_max>          maximum magnetic Reynolds number [default: 2.0]
     --k_min=<k_min>            minimum z wavenumber [default: 0.001]
@@ -33,6 +34,7 @@ Omega1 = float(args['--Omega1'])
 Omega2 = float(args['--Omega2'])
 Pm = float(args['--Pm'])
 beta = float(args['--beta'])
+xi = float(args['--xi'])
 Rm_min = float(args['--Rm_min'])
 Rm_max = float(args['--Rm_max'])
 k_min = float(args['--k_min'])
@@ -41,4 +43,4 @@ n_Rm = int(args['--n_Rm'])
 n_k = int(args['--n_k'])
 n_r = int(args['--n_r'])
 
-find_crit(R1, R2, Omega1, Omega2, beta, Pm, Rm_min, Rm_max, k_min, k_max, n_Rm, n_k, n_r)
+find_crit(R1, R2, Omega1, Omega2, beta, xi, Pm, Rm_min, Rm_max, k_min, k_max, n_Rm, n_k, n_r)

### 2016-07-11 paper/widegap_mri_paper.tex
up to date papers

@@
 These definitions automatically satisfy Equations \ref{eq:incompressibility} and \ref{eq:solenoid} for axisymmetric disturbances.
-We note that in the linearized equations, streamfunctions of the form $u_x = \partial_z \Psi$, $u_z = -(\partial_r + \frac{1}{r}) \Psi$, and the corresponding definitions of the magnetic vector potential, are convenient choices, but this does not hold true for the nonlinear terms.
+We note that in the linearized equations, streamfunctions of the form $u_x = \partial_z \Psi$, $u_z = -(\partial_r + \frac{1}{r}) \Psi$, and the corresponding definitions of the magnetic vector potential, are convenient choices, but this does not hold true for the nonlinear terms because of the incommutability of $\partial_r$ and $\partial_r + \frac{1}{r}$.
 

### 2016-07-11 python/run_IVP.py

@@ -95,11 +95,6 @@ while IVP.ok:
 alpha_array = np.array(alpha_list)
 t_array = np.array(t_list)
 
-#self.alpha = alpha
-#self.alphaZ = alphaZ
-#self.alpha_array = alpha_array
-#self.t_array = t_array
-
 saturation_amplitude = alpha_array[-1, 0]
 print(saturation_amplitude)
 

### 2016-07-11 python/widegap/allorders2_widegap.py

@@ -55,9 +55,11 @@ class MRI():
         self.c2 = (self.R1**2*self.R2**2*(self.Omega1 - self.Omega2))/(self.R2**2 - self.R1**2)
         
         self.zeta_mean = 2*(self.R2**2*self.Omega2 - self.R1**2*self.Omega1)/((self.R2**2 - self.R1**2)*np.sqrt(self.Omega1*self.Omega2))
+        self.R0 = (self.R1 + self.R2)/2.0
+        self.q_R0 = 2*self.c2/(self.R0**2*self.c1 + self.c2)
         
         logger.info("MRI parameters: Q = {}; Rm = {}; Pm = {}; beta = {}; norm = {}, Re = {}".format(self.Q, self.Rm, self.Pm, self.beta, norm, self.R))
-        logger.info("Mean zeta is {}, meaning effective shear parameter q(r0) = 2 - zeta = {}".format(self.zeta_mean, 2 - self.zeta_mean))
+        logger.info("Effective shear parameter q(R0) = {}".format(self.q_R0))
         
         if self.xi != 0:
             logger.info("A nonzero xi means this is the HMRI")

### 2016-07-11 python/widegap/find_crit.py

@@ -125,7 +125,7 @@ def find_crit(R1, R2, Omega1, Omega2, beta, xi, Pm, Rm_min, Rm_max, k_min, k_max
 
     # generating the grid is the longest part
     start = time.time()
-    cf.grid_generator(Rm_min, Rm_max, k_min, k_max, n_Rm, n_k) 
+    cf.grid_generator(Rm_min, Rm_max, k_min, k_max, n_Rm, n_k, complex=True) 
     end = time.time()
     if comm.rank == 0:
         print("grid generation time: {:10.5f} sec".format(end-start))

### 2016-07-13 python/run_IVP.py
added widegap eigenfunction and nonlinear vector figures

@@ -33,7 +33,6 @@ pmcoeffs_fn = '/Users/susanclark/weakly_nonlinear_mri/data/pm_sat_coeffs.h5'
 #coeff = Coefficients(19, pmcoeffs_fn)
 coeff = Coefficients(5, pmcoeffs_fn)
 
-print("ac = {}, bc = {}, hc = {}".format(coeff.a/coeff.c, coeff.b/coeff.c, coeff.h/coeff.c))
 print("saturation amplitude = {}".format(np.sqrt(coeff.b/coeff.c)))
 
 # Actually solve the IVP
@@ -48,6 +47,8 @@ problem = de.IVP(Zdomain, variables=['alpha', 'alphaZ'], ncc_cutoff=1e-10)
 problem.parameters['ac'] = -coeff.a/coeff.c
 problem.parameters['bc'] = coeff.b/coeff.c
 problem.parameters['hc'] = -coeff.h/coeff.c
+
+print("ac = {}, bc = {}, hc = {}".format(-coeff.a/coeff.c, coeff.b/coeff.c, -coeff.h/coeff.c))
 
 problem.add_equation("-ac*dt(alpha) + bc*alpha + hc*dZ(alphaZ) = alpha*abs(alpha**2)") 
 problem.add_equation("alphaZ - dZ(alpha) = 0")

### 2016-07-13 paper/widegap_mri_paper.tex
added widegap eigenfunction and nonlinear vector figures
added widegap (h)mri matrices to paper

@@
 
-\title{The Weakly Nonlinear Magnetorotational Instability in a Wide-Gap Channel Flow}
+\title{The Weakly Nonlinear Magnetorotational Instability in a Taylor Couette Flow}
 \author{Clark, S.E.\altaffilmark{1}}
@@
 
-\section{Wide Gap Equations}
+The magnetorotational instability (MRI) is thought to be a crucial driver of angular momentum transport and turbulence in astrophysical disks. (more MRI intro)
 
-Many investigations of the MRI use the ``thin gap" approximation, in which the radial extent of the fluid channel is taken to be much smaller than the radius of curvature.
-That is, for a channel center $r_0$ bounded by inner and outer radii $r_1$ and $r_2$, respectively, the thin gap approximation applies when $r_0 \gg (r_2 - r_1)$.
-The thin gap approximation simplifies the MRI equations by excluding curvature terms, because the flow through a thin gap can be taken to be approximately linear in $\phi$, i.e.
+%The MRI is fundamental to our understanding of star formation, but its
+
+Many investigations of the MRI use the ``narrow gap" approximation, in which the radial extent of the fluid channel is taken to be much smaller than the radius of curvature.
+That is, for a channel center $r_0$ bounded by inner and outer radii $r_1$ and $r_2$, respectively, the narrow gap approximation applies when $r_0 \gg (r_2 - r_1)$.
+The narrow gap approximation simplifies the MRI equations by excluding curvature terms, because the flow through a narrow gap can be taken to be approximately linear in $\phi$, i.e.
 Cartesian.
 In this work we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.
+
+\section{Wide Gap Equations}
 
@@
 Thus through judicious choice of cylinder rotation rates, we can set $q(r_0) = 3/2$, for quasi-Keplerian flow.
-Note that in the thin gap approximation we have the freedom to choose any rotation profile, as the narrow gap width imposes a linear shear (constant $q$).
+Note that in the narrow gap approximation we have the freedom to choose any rotation profile, as the narrow gap width imposes a linear shear (constant $q$).
 Our base velocity is $\mathbf{u_0} = r\Omega(r) \phihat$.
@@
 We introduce the Reynolds number $\reye = \Omega_0 r_0^2/\nu$, the magnetic Reynolds number $\reym = \Omega_0 r_0^2 / \eta$, and a plasma beta parameter $\beta = \Omega_0^2 r_0^2 \rho_0/B_0^2$.
-If we define the dimensional cylindrical coordinate $r = r_0(1 + \delta x)$, the thin gap equations are recovered in the limit $\delta \rightarrow 0$.
+If we define the dimensional cylindrical coordinate $r = r_0(1 + \delta x)$, the narrow gap equations are recovered in the limit $\delta \rightarrow 0$.
 
@@
 
-\section{Weakly Nonlinear Perturbation Analysis}
+\section{Weakly Nonlinear Perturbation Analysis}\label{sec:wnl_analysis}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../python/widegap/figures/orders_1_2_widegap_fiducial_eigenfunctions.pdf}
+\caption{Eigenfunctions}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../python/widegap/figures/nonlinear_2_3_widegap_fiducial_eigenfunctions.pdf}
+\caption{Nonlinear terms}
+\end{figure}
+
 See Appendix \ref{app:basic_equations} for the definition of matrices and nonlinear vectors, and a thorough derivation.
-We emphasize that Equations \ref{eq:ordere} - \ref{eq:ordere3} have the same form as these equations in the thin gap case, although the matrices (which contain all radial derivatives) are significantly different in this wide gap formulation.
+We emphasize that Equations \ref{eq:ordere} - \ref{eq:ordere3} have the same form as these equations in the narrow gap case, although the matrices (which contain all radial derivatives) are significantly different in this wide gap formulation.
 This is because we do not have slow variation in the radial dimension.
@@
 \section{A. Detailed Equations}\label{app:basic_equations}
+
+Here we detail the perturbation analysis described in Section \ref{sec:wnl_analysis}. The linear system is described by Equation \ref{eq:unperturbed_matrix_eqns}, where
+
+\beq
+\mathcal{L} = \mathcal{L}_0 + \mathcal{L}_1 \partial_z + \mathcal{L}_2 \partial_z^2 + \mathcal{L}_3 \partial_z^3 + \mathcal{L}_4 \partial_z^4,
+\eeq
+
+\beq
+\widetilde{\mathcal{G}} = - \mathcal{G} \partial_z - \mathcal{L}_3 \partial_z^3,
+\eeq
+
+\beq
+\widetilde{\mathcal{H}} = \mathcal{H} \partial_z,
+\eeq
+
+and the constituent matrices are defined as
+
+\beq
+\mathcal{L}_0 = \left[\begin{matrix}
+-\frac{1}{\reye} (-\frac{3}{r^4} \partial_r + \frac{3}{r^3}\partial_r^2 - \frac{2}{r^2}\partial_r^3 + \frac{1}{r}\partial_r^4) & 0 & 0 & 0 \\
+0 & -\frac{1}{\reye} (\partial_r^2 + \frac{1}{r}\partial_r - \frac{1}{r^2}) & 0 & 0 \\
+0 & 0 & -\frac{1}{\reym} (\partial_r^2 - \frac{1}{r} \partial_r) & 0 \\
+0 & 0 & 0 & -\frac{1}{\reym} (\partial_r^2 + \frac{1}{r}\partial_r - \frac{1}{r^2}) \\
+\end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{L}_1 = \left[\begin{matrix}
+0 & -\frac{2}{r} u_0 & \frac{2}{\beta} (\frac{1}{r^2} \partial_r - \frac{1}{r}\partial_r^2) & 0 \\
+\frac{1}{r^2} u_0 + \frac{1}{r}\partial_r u_0 & 0 & 0 & -\frac{2}{\beta} \\
+-1 & 0 & 0 & 0 \\
+0 & -1 & \frac{1}{r^2} u_0 - \frac{1}{r} \partial_r u_0 & 0 \\
+\end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{L}_2 = \left[\begin{matrix}
+-\frac{1}{\reye} (-\frac{2}{r^2}\partial_r + \frac{2}{r}\partial_r^2) & 0 & 0 & 0 \\
+0 & -\frac{1}{\reye} & 0 & 0 \\
+0 & 0 & -\frac{1}{\reym} & 0 \\
+0 & 0 & 0 & -\frac{1}{\reym} \\
+\end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{L}_3 = \left[\begin{matrix}
+0 & 0 & -\frac{2}{\beta}\frac{1}{r} & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+\end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{L}_4 = \left[\begin{matrix}
+-\frac{1}{\reye}\frac{1}{r} & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+\end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{G} = \left[\begin{matrix}
+0 & 0 & \frac{2}{\beta}(\frac{1}{r^2}\partial_r - \frac{1}{r}\partial_r^2) & 0 \\
+0 & 0 & 0 & -\frac{2}{\beta} \\
+-1 & 0 & 0 & 0 \\
+0 & -1 & 0 & 0 \\
+\end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{H} = \left[\begin{matrix}
+0 & 0 & 0 & \frac{2}{\beta} \frac{2}{r^2} \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+-\frac{2}{r^3} & 0 & 0 & 0 \\
+\end{matrix}\right]
+\eeq
 

### 2016-07-13 python/widegap/allorders2_widegap.py

@@ -480,11 +480,11 @@ class N2(MRI):
         self.N20_u = (N20_u_jacobians + N20_u_advectives).evaluate()
         self.N20_u.name = "N20_u"
                 
-        N22_A = (1/rfield)*((1j*Q*o1.psi)*o1.A_r - (1j*Q*o1.A)*o1.psi_r)
+        N22_A = (1/rfield)*((1j*Q*o1.psi)*o1.A_r - (1j*Q*o1.A)*o1.psi_r) # correct - checked 7/13/16
         self.N22_A = N22_A.evaluate()
         self.N22_A.name = "N22_A"
         
-        N20_A = (1/rfield)*((1j*Q*o1.psi)*o1.A_star_r - (-1j*Q*o1.A_star)*o1.psi_r)
+        N20_A = (1/rfield)*((1j*Q*o1.psi)*o1.A_star_r - (-1j*Q*o1.A_star)*o1.psi_r) # correct - checked 7/13/16
         self.N20_A = N20_A.evaluate()
         self.N20_A.name = "N20_A"
         
@@ -982,19 +982,19 @@ class AmplitudeAlpha(MRI):
         b_B_rhs = 1j*self.Q*o1.u + invr**3*2*1j*self.Q*self.xi*o1.psi
         b_B_rhs = b_B_rhs.evaluate()
         
-        # (-L1twiddle V21 + L2twiddle V11 - xi H V21)
-        h_psi_rhs = (-self.iR*invr*4*1j*self.Q**3*o2.psi21 - (2/self.beta)*invr*3*self.Q**2*o2.A21 + self.iR*invr*6*self.Q**2*o1.psi - (2/self.beta)*invr*3*1j*self.Q*o1.A
-                    + self.iR*invr*4*1j*self.Q*o2.psi21_rr - self.iR*invr**2*4*1j*self.Q*o2.psi21_r + invr*2*u0*o2.u21 + (2/self.beta)*invr*o2.A21_rr
-                    - (2/self.beta)*invr**2*o2.A21_r - (2/self.beta)*invr**2*2*self.xi*o2.B21 - self.iR*invr*2*o1.psi_rr + self.iR*invr**2*2*o1.psi_r)
+        # (L1twiddle V21 + L2twiddle V11 + xi H V21)
[…]

### 2016-07-13 paper/widegap_mri_paper.tex
all helical terms added
added fiducial parameters table to widegap paper. unfinished. do we like this?

@@
 %The MRI is fundamental to our understanding of star formation, but its 
+- people make a number of simplifications, incl thin gap, shearing box, etc
 
@@
 Our base velocity is $\mathbf{u_0} = r\Omega(r) \phihat$.
-We initialize an axial magnetic field $\mathbf{B_0} = B_0 \zhat$.
+We initialize a magnetic field $\mathbf{B_0} = B_0 \zhat + B_0 \xi \frac{r_0}{r} \phihat$, so that the base magnetic field is axial when $\xi = 0$ and otherwise helical.
+
+In this work we will focus our findings on two fiducial parameter sets, one for the standard MRI (hereafter SMRI) where $\xi = 0$ and one for the helical MRI (hereafter HMRI).
+Our fiducial parameters are described in Table \ref{table:parameters}.
+
+\begin{table}[ht]\label{table:parameters}
+\normalsize
+\caption{Fiducial parameters for MRI runs} 
+\centering
+%\resizebox{\textwidth}{!}
+{\begin{tabular}{rlrrrrrrr}
+  \hline
+ & $\xi$ & $Pm$ & $\Omega_2/\Omega_1$ & $\beta$ & X \\ 
+  \hline\hline
+Standard MRI & 0 & 25 & 0 & 0 & 0 \\ 
+Helical MRI & 4 & 0 & 0\\ 
+   \hline
+\end{tabular}}
+\end{table}
 
@@
 \label{eq:Psi_perturbed}
-\frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] \\
+\frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r^2}B_0 \xi \partial_z B_\phi - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] \\
 = - J(\Psi, \frac{1}{r^2} ( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi) ) + \frac{2}{\beta} J(A, \frac{1}{r^2} ( \nabla^2 A - \frac{2}{r} \partial_rA) ) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi  + \frac{2}{r} u_\phi \partial_z u_\phi \\
@@
   \label{eq:Bphi_perturbed}
-& \partial_t B_\phi + \frac{1}{r^2} u_0 \partial_z A - B_0 \partial_z u_\phi - \frac{1}{r} \partial_r u_0 \partial_z A - \frac{1}{\reym} (\nabla^2 B_\phi - \frac{1}{r^2} B_\phi ) = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)
+& \partial_t B_\phi + \frac{1}{r^2} u_0 \partial_z A - B_0 \partial_z u_\phi - \frac{1}{r} \partial_r u_0 \partial_z A - \frac{2}{r^3} B_0 \xi \partial_z \Psi - \frac{1}{\reym} (\nabla^2 B_\phi - \frac{1}{r^2} B_\phi ) = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)
 + \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A 
@@
 \beq\label{eq:unperturbed_matrix_eqns}
-\mathcal{D}\partial_t\mathbf{V} + \mathcal{L}\mathbf{V} + \epsilon^2 \widetilde{\mathcal{G}} \mathbf{V} + \mathbf{N} = 0,
+\mathcal{D}\partial_t\mathbf{V} + \mathcal{L}\mathbf{V} + \epsilon^2 \widetilde{\mathcal{G}} \mathbf{V} + \xi \widetilde{\mathcal{H}} \mathbf{V} +  \mathbf{N} = 0,
 \eeq
@@
 \begin{align}
-\mathcal{O}(\epsilon)&: \mathcal{L} \mathbf{V}_1 + \mathcal{D} \partial_t \mathbf{V}_1 = 0 \label{eq:ordere}\\
-\mathcal{O}(\epsilon^2)&: \mathcal{L} \mathbf{V}_2 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \label{eq:ordere2}\\
-\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3.
+\mathcal{O}(\epsilon)&: \mathcal{L} \mathbf{V}_1 + \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \mathcal{D} \partial_t \mathbf{V}_1 = 0 \label{eq:ordere}\\
+\mathcal{O}(\epsilon^2)&: \mathcal{L} \mathbf{V}_2 + \xi \widetilde{\mathcal{H}} \mathbf{V}_2 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_1 + \xi \mathcal{H} \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \label{eq:ordere2}\\
+\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \xi \widetilde{\mathcal{H}} \mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2 + \xi \mathcal{H}\partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 - \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3 = 0.
 \label{eq:ordere3}
@@
 \includegraphics[width=\textwidth]{../python/widegap/figures/orders_1_2_widegap_fiducial_eigenfunctions.pdf}
-\caption{Eigenfunctions}
+\caption{Eigenfunctions of the first order equations, first order adjoint homogenous equations, and second order equations.
+We use our fiducial parameters for the standard MRI ($\xi = 0$).
+Eigenfunctions are normalized such that either the real or imaginary part is zero.}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{../python/widegap/figures/nonlinear_2_3_widegap_fiducial_eigenfunctions.pdf}
-\caption{Nonlinear terms}
+\caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.}
 \end{figure}

### 2016-07-14 paper/widegap_mri_paper.tex
some hmri intro to widegap paper added

@@
 
+Despite the challenges inherent in studying the MRI, much progress has been made nu
+
 Many investigations of the MRI use the ``narrow gap" approximation, in which the radial extent of the fluid channel is taken to be much smaller than the radius of curvature.
@@
 In this work we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.
+
+Because we include curvature terms, our theory is also relevant to the helical magnetorotational instability (HMRI).
+Discovered by \citeauthor{Hollerbach:2005tr} (\citeyear{Hollerbach:2005tr}), the HMRI is an overstability in which the background magnetic field is helical.
+The HMRI currently occupies a special place in the MRI puzzle.
+The HMRI has been proposed as a method of awakening angular momentum transport in the ``dead zones" of protoplanetary disks where the magnetic Prandtl number ($Pm = \nu/\eta$) becomes very small.
+However the rotation profiles needed to excite HMRI may be prohibitively steeper than Keplerian depending on the boundary conditions, and so its role in astrophysical disks is currently debated (\citei{Liu:2006}, \citei{Rudiger:2007}, \citei{Kirillov:2013}).
+However, the HMRI is significantly easier to excite in a laboratory setting, and has already been detected in the laboratory by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
+
+We explore the weakly nonlinear HMRI
 
@@
 In this work we will focus our findings on two fiducial parameter sets, one for the standard MRI (hereafter SMRI) where $\xi = 0$ and one for the helical MRI (hereafter HMRI).
+We choose the SMRI parameters to be comparable to the case explored in \citei{Umurhan:2007hs}, with the geometric parameters of \citei{Goodman:2002ix}.
+The HMRI parameters were chosen to be comparable to \citei{Hollerbach:2005tr}.
 Our fiducial parameters are described in Table \ref{table:parameters}.
 
-\begin{table}[ht]\label{table:parameters}
+\begin{table}[ht]
 \normalsize
-\caption{Fiducial parameters for MRI runs} 
+\caption{Fiducial parameters for MRI runs} \label{table:parameters}
 \centering
@@
   \hline
- & $\xi$ & $Pm$ & $\Omega_2/\Omega_1$ & $\beta$ & X \\ 
+ & $\xi$ & $Pm$ & $\beta$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ \\ 
   \hline\hline
-Standard MRI & 0 & 25 & 0 & 0 & 0 \\ 
-Helical MRI & 4 & 0 & 0\\ 
+Standard MRI & 0 & 1E-3 & 25 & 0.18 & 0.33 \\ 
+Helical MRI & 4 & 1E-6 & 1.7E-2 & 0.27 & 0.5\\ 
    \hline
@@
 
-where J is the Jacobian $J(x, y) \equiv \partial_z x \partial_r y - \partial_r x \partial_z y$.
-Note that in the above, $\nabla^2 x \equiv \partial_r^2 x + \partial_z^2 x + \frac{1}{r} \partial_r x$.
+where J is the Jacobian $J(f, g) \equiv \partial_z f \partial_r g - \partial_r f \partial_z g$.
+Note that in the above, $\nabla^2 f \equiv \partial_r^2 f + \partial_z^2 f + \frac{1}{r} \partial_r f$.
 Equations \ref{eq:Psi_perturbed} - \ref{eq:Bphi_perturbed} are nondimensionalized, where lengths have been scaled by $r_0$, velocities by $r_0 \Omega_0$, densities by $\rho_0$, and magnetic fields by $B_0$; where $B_0$ appears in the above it is formally unity.
@@
 
+\bibitem[{Goodman \& Ji(2002)}]{Goodman:2002ix}
+Goodman, J. \& Ji, H., 2002, JFM, 462, 365
+
+\bibitem[{Hollerbach \& R\"udiger(2005)}]{Hollerbach:2005tr}
+Hollerbach, R. \& R\"udiger, G., 2005, Phys. Rev. Lett. 95, 124501
+
+\bibitem[{Kirillov \& Stefani(2013)}]{Kirillov:2013}
+Kirillov, O, \& Stefani, F, 2013, Phys. Rev. Lett. 111, 061103
+
 \bibitem[{Latter et al.(2015)}]{Latter:2015}
 Latter, H.N., Fromang, S., \& Faure, J., 2015, MNRAS 453, 3257
+
+\bibitem[{Liu et al.(2006)}]{Liu:2006}
+Liu, W, Goodman, J, Herron, I, and Ji, H, 2006, Phys. Rev. E 74, 056302
+
+\bibitem[{R\"udiger \& Hollerbach(2007)}]{Rudiger:2007}
+R\"udiger, G, \& Hollerbach, R, Phys. Rev. E 76, 068301
+
+\bibitem[{Stefani et al.(2006)}]{Stefani:2006iv}
+Stefani, F, Gundrum, T, Gerbeth, G, RŸdiger, G, Schultz, M, Szklarski, J, Hollerbach, R, Phys Rev Lett, 2006, 97, 184502
+
+\bibitem[{Stefani et al.(2009)}]{Stefani:2009hp}
+Stefani, F, Gerbeth, G, Gundrum, T, Hollerbach, R, Priede, J, R\"udiger, G, and Szklarski, J, Phys Rev E, 80, 066303
 

### 2016-07-14 paper/widegap_mri_paper.tex
sign of two N31_B advective terms fixed, comments added

@@
 
-\title{The Weakly Nonlinear Magnetorotational Instability in a Taylor Couette Flow}
+\title{The Weakly Nonlinear Standard and Helical Magnetorotational Instabilities in a Taylor Couette Flow}
 \author{Clark, S.E.\altaffilmark{1}}
@@
 Cartesian.
+Previous investigations into the weakly nonlinear MRI have used this narrow gap approximation (\citei{Umurhan:2007dz}, \citei{Umurhan:2007hs}, \citei{Clark:2016}).
 In this work we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.
@@
 However the rotation profiles needed to excite HMRI may be prohibitively steeper than Keplerian depending on the boundary conditions, and so its role in astrophysical disks is currently debated (\citei{Liu:2006}, \citei{Rudiger:2007}, \citei{Kirillov:2013}).
-However, the HMRI is significantly easier to excite in a laboratory setting, and has already been detected in the laboratory by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
+However, the HMRI is significantly easier to excite in a laboratory setting than the standard MRI, and has already been detected in the laboratory by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
 
@@
 
+\beq
+\mathcal{D} = \left[\begin{matrix}
+\frac{1}{r}\partial_r^2 + \frac{1}{r}\partial_z^2 - \frac{1}{r^2}\partial_r & 0 & 0 & 0 \\
+0 & 1 & 0 & 0 \\
+0 & 0 & 1 & 0 \\
+0 & 0 & 0 & 1 \\
+\end{matrix}\right]
+\eeq
+
 \section{B. Nonlinear Terms}\label{app:nonlinear_terms}

### 2016-07-14 python/widegap/allorders2_widegap.py

@@ -870,29 +870,39 @@ class N3(MRI):
                        
         self.N31_psi_noadvective = (Jacobian1_part1 + Jacobian1_part2 + Jacobian2_part1 + Jacobian2_part2 + Jacobian3_part1 + Jacobian3_part2 
                        + Jacobian4_part1 + Jacobian4_part2).evaluate()
-                       
+        
+        # (1/r) J(psi1, u2)               
         Jacobian1 = invr*(1j*k*o1.psi*(o2.u20_r + o2.u20_star_r) + (-1j*k)*o1.psi_star*o2.u22_r - o1.psi_star_r*(2*1j*k)*o2.u22)
         
+        # -(1/r) (2/beta) J(A1, B2)
         Jacobian2 = -invr*(2/self.beta)*(1j*k*o1.A*(o2.B20_r + o2.B20_star_r) + (-1j*k)*o1.A_star*o2.B22_r - o1.A_star_r*(2*1j*k)*o2.B22)
         
+        # (1/r) J(psi2, u1)
         Jacobian3 = invr*(-o2.psi20_r*1j*k*o1.u - o2.psi20_star_r*1j*k*o1.u + 2*1j*k*o2.psi22*o1.u_star_r - o2.psi22_r*(-1j*k)*o1.u_star)
         
+        # -(1/r) (2/beta) J(A2, B1)
         Jacobian4 = -invr*(2/self.beta)*(-o2.A20_r*1j*k*o1.B - o2.A20_star_r*1j*k*o1.B + 2*1j*k*o2.A22*o1.B_star_r - o2.A22_r*(-1j*k)*o1.B_star)
         
+        # (1/r^2) u1 dz psi2
         advective1 = invr**2*(o1.u_star*2*1j*k*o2.psi22)
         
+        # (1/r^2) u2 dz psi1
         advective2 = invr**2*(o2.u20*1j*k*o1.psi + o2.u20_star*1j*k*o1.psi + o2.u22*(-1j*k)*o1.psi_star)
         
+        # -(2/beta)(1/r^2) B1 dz A2
         advective3 = -(2/self.beta)*invr**2*(o1.B_star*2*1j*k*o2.A22)
         
+        # - (2/beta)(1/r^2) B2 dz A1
         advective4 = -(2/self.beta)*invr**2*(o2.B20*1j*k*o1.A + o2.B20_star*1j*k*o1.A + o2.B22*(-1j*k)*o1.A_star)
         
[…]

### 2016-07-14 paper/widegap_mri_paper.tex
more intro. needs more citations.

@@
 Despite the challenges inherent in studying the MRI, much progress has been made nu
+
+Because of the MRI's complexity, analytical treatments of the MRI use some combination of convenient approximations: ideal MHD, linearized equations, expedient boundary conditions, and simplified geometries.
+In particular, much of the analytical work on MRI saturation has focused on local approximations, often representing a section of a disk by solving the MHD equations in an isolated region subject to shear periodic boundary conditions (the ``shearing box").
+While the MRI is a local instability,
 
@@
   \hline
- & $\xi$ & $Pm$ & $\beta$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ \\ 
+ & $\xi$ & $Pm$ & $\beta$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ & radial magnetic b.c. \\ 
   \hline\hline
-Standard MRI & 0 & 1E-3 & 25 & 0.18 & 0.33 \\ 
-Helical MRI & 4 & 1E-6 & 1.7E-2 & 0.27 & 0.5\\ 
+Standard MRI & 0 & 1E-3 & 25 & 0.18 & 0.33 & conducting \\ 
+Helical MRI & 4 & 1E-6 & 1.7E-2 & 0.27 & 0.5 & insulating\\ 
    \hline
@@
 \bibitem[{Stefani et al.(2006)}]{Stefani:2006iv}
-Stefani, F, Gundrum, T, Gerbeth, G, RŸdiger, G, Schultz, M, Szklarski, J, Hollerbach, R, Phys Rev Lett, 2006, 97, 184502
+Stefani, F, Gundrum, T, Gerbeth, G, R\"udiger, G, Schultz, M, Szklarski, J, Hollerbach, R, Phys Rev Lett, 2006, 97, 184502
 

### 2016-07-14 python/widegap/allorders2_widegap.py

@@ -701,7 +701,7 @@ class OrderE2(MRI):
         bv22.add_equation("-r**2*ru0*dz*2*u - twooverbeta*r**3*dz*dr(Ar) + twooverbeta*r**2*dz*Ar - twooverbeta*r**3*dz**3*A - iR*r**3*dr(psirrr) + 2*iR*r**2*psirrr - 2*iR*r**3*dz**2*psirr - 3*iR*r*psirr + 2*iR*r**2*dz**2*psir + 3*iR*psir - iR*r**3*dz**4*psi + twooverbeta*r**2*2*xi*(2*1j*Q)*B = rhs_psi22")
         bv22.add_equation("rrdu0*dz*psi + ru0*dz*psi - r**3*twooverbeta*dz*B - r**3*iR*dr(ur) - r**2*iR*ur - r**3*iR*dz**2*u + r*iR*u = rhs_u22")
         bv22.add_equation("-dz*psi*r - r*iRm*dr(Ar) + iRm*Ar - r*iRm*dz**2*A = rhs_A22")
-        bv22.add_equation("-dz*rrdu0*A - dz*r**3*u + ru0*dz*A - iRm*r**3*dr(Br) - r**2*iRm*Br - iRm*r**3*dz**2*B + r*iRm*B - 2*xi*(2*1j*Q)*psi = rhs_B22")
+        bv22.add_equation("-dz*rrdu0*A - dz*r**3*u + ru0*dz*A - iRm*r**3*dr(Br) - r**2*iRm*Br - iRm*r**3*dz**2*B + r*iRm*B - 2*xi*(2*1j*Q)*psi = rhs_B22") #checked 7/14/16 - these are all right
         
         bv22.add_equation("dr(psi) - psir = 0")
         bv22.add_equation("dr(psir) - psirr = 0")
@@ -877,6 +877,20 @@ class N3(MRI):
         # (2/beta) (2/r) B2 dz B1
         advective4 = 2*invr*(2/self.beta)*(o2.B20*1j*k*o1.B + o2.B20_star*1j*k*o1.B + o2.B22*(-1j*k)*o1.B_star)
         
+        # diagnostics
+        self.N31_psi_Jacobian1_part1 = Jacobian1_part1.evaluate()
+        self.N31_psi_Jacobian1_part2 = Jacobian1_part2.evaluate()
+        self.N31_psi_Jacobian2_part1 = Jacobian2_part1.evaluate()
[…]

### 2016-07-14 paper/widegap_mri_paper.tex
widegap amplitude run code, save data

@@
 
-\title{The Weakly Nonlinear Standard and Helical Magnetorotational Instabilities in a Taylor Couette Flow}
+\title{Weakly Nonlinear Analysis of the Standard and Helical Magnetorotational Instabilities in a Taylor Couette Flow}
 \author{Clark, S.E.\altaffilmark{1}}
@@
 In particular, much of the analytical work on MRI saturation has focused on local approximations, often representing a section of a disk by solving the MHD equations in an isolated region subject to shear periodic boundary conditions (the ``shearing box").
-While the MRI is a local instability,
+While the MRI is a local instability, certain properties of the local problem are not generic to the global problem.
+Channel modes, a class of MRI mode that dominates MRI evolution and saturation in the local approximation, are artificially over-prominent in the shearing box (e.g.
+\citei{Latter:2015}).
 
@@
 
-a Ginzburg-Landau equation (e.g.
-\cite{Aranson:2002}).
+a Ginzburg-Landau equation.
 Here we note a departure from the behavior of the thin-gap system.
@@
 (I'm just guessing but this may lead to higher torque on inner cylinder -- should calculate)
+
+For both the standard and helical MRI we derive a GLE amplitude.
+For the standard MRI we find a real GLE, in which the coefficients of the GLE are purely real (the amplitude is still complex in general).
+For the helical MRI we find a complex GLE, a source of even richer phase dynamics than its real counterpart (e.g.
+\cite{Aranson:2002}).
+This complex GLE is a natural form for the HMRI, as it is a traveling wave instability.
 

### 2016-07-14 python/widegap/allorders2_widegap.py

@@ -804,11 +804,10 @@ class N3(MRI):
         if o1 == None:
             o1 = OrderE(domain, Q = Q, Rm = Rm, Pm = Pm, beta = beta, Omega1 = Omega1, Omega2 = Omega2, xi = xi, norm = norm)
             MRI.__init__(self, domain, Q = Q, Rm = Rm, Pm = Pm, beta = beta, Omega1 = Omega1, Omega2 = Omega2, xi = xi, norm = norm)
-            n2 = N2(domain, Q = Q, Rm = Rm, Pm = Pm, beta = beta, Omega1 = Omega1, Omega2 = Omega2, xi = xi, norm = norm)
+            
         else:
             MRI.__init__(self, domain, Q = o1.Q, Rm = o1.Rm, Pm = o1.Pm, beta = o1.beta, Omega1 = o1.Omega1, Omega2 = o1.Omega2, xi = o1.xi, norm = o1.norm)
-            n2 = N2(domain, Q = o1.Q, Rm = o1.Rm, Pm = o1.Pm, beta = o1.beta, Omega1 = o1.Omega1, Omega2 = o1.Omega2, xi = o1.xi, norm = o1.norm)
-
+            
         if ah == None:
             ah = AdjointHomogenous(domain, o1 = o1, Q = self.Q, Rm = self.Rm, Pm = self.Pm, beta = self.beta, Omega1 = self.Omega1, Omega2 = self.Omega2, xi = self.xi, norm = self.norm)
 

### 2016-07-14 python/widegap/run_widegap_amplitude.py

@@ -1,0 +1,184 @@
+import dedalus.public as de
+import numpy as np
+from mpi4py import MPI
+from allorders2_widegap import AmplitudeAlpha
+import h5py
+
+# UMR07-like parameters
+
+Q = 0.01269
+Rm = 0.67355
+Pm = 1.0e-3
+beta = 25.0
+Omega1 = 313.55
+Omega2 = 56.43
+xi = 0
+R1 = 5
+R2 = 15
+
+
+# Hollerbach & Rudiger
+"""
+Pm = 1.0E-6
+Re = 1521
+Rm = Pm*Re
+xi = 4
+Q = 2.33
+Ha = 16.3
+beta = (2*Re*Rm)/(Ha**2)
+R1 = 1.
+R2 = 2.
+mu_omega = 0.27
+Omega1 = 313.55
+Omega2 = Omega1*mu_omega
+"""
+
+gridnum = 512#256#128
+
+r_basis = de.Chebyshev('r', gridnum, interval=(R1, R2))
+domain = de.Domain([r_basis], np.complex128, comm=MPI.COMM_SELF)
+print("running at gridnum", gridnum)
+
+aa = AmplitudeAlpha(domain, Q = Q, Rm = Rm, Pm = Pm, xi = xi, Omega1 = Omega1, Omega2 = Omega2, beta = beta)
+aa.print_coeffs()
+
+
+fn_root = "/Users/susanclark/weakly_nonlinear_mri/data/"
+fn = fn_root + "widegap_amplitude_parameters_Q_{:03.2f}_Rm_{:04.4f}_Pm_{:.2e}_Omega1_{:05.2f}_Omega2_{:05.2f}_beta_{:.2f}_xi_{:.2f}_gridnum_{}.h5".format(Q, Rm, Pm, Omega1, Omega2, beta, xi, gridnum)
+with h5py.File(fn,'w') as f:
+    r = f.create_dataset("r", data=aa.r)
+    o1psi = f.create_dataset("psi11", data=aa.o1.psi['g'])
+    o1psi_r = f.create_dataset("psi11_r", data=aa.o1.psi_r['g'])
+    o1psi_rr = f.create_dataset("psi11_rr", data=aa.o1.psi_rr['g'])
+    o1psi_rrr = f.create_dataset("psi11_rrr", data=aa.o1.psi_rrr['g'])
+    o1psi_star = f.create_dataset("psi11_star", data=aa.o1.psi_star['g'])
[…]

### 2016-07-15 paper/mri_paper.tex
edited norm of allorders to not include magic number. Still includes real imag shift
fixed N2 psi terms in paper
fixed N2 u terms in paper
Pm 2E-3
removed hmri section from thin gap paper

@@
 \author{Oishi, J.S. \altaffilmark{2,}\altaffilmark{3}}
-\author{Mac Low, M.-M.\altaffilmark{1,}\altaffilmark{2}}
+%\author{Mac Low, M.-M.\altaffilmark{1,}\altaffilmark{2}}
 \altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
@@
 
-% Jeff's sims!!!!!
-
-\section{Helical MRI}
-
-The configuration investigated thus far, that of magnetized Taylor Couette flow with a purely axial magnetic field, constitutes the standard MRI.
-The helical MRI (HMRI) is a related, but separate, MHD instability, first identified by Hollerbach and R\"udiger (\citeyear{Hollerbach:2005tr}).
-It arises in the presence of combined axial and azimuthal fields.
-
-We initialize the helical MRI by adding a constant background toroidal field to Equation \ref{backgroundfield1}, such that our initial magnetic field structure becomes
-
-\beq
-\mathbf{B} = B_0 \hat{z} + \xi B_0 \hat{y},
-\eeq
-
-where $\xi$ is a dimensionless parameter representing the relative strength of the axial field and the vertical field.
-
-This addition leaves Equations \ref{eqset2} - \ref{eqset4} unchanged, but Equation \ref{eqset1} gains a term to become
-
-\begin{multline}
-\label{eqset1_hmri}
-\partial_t \nabla^2 \Psi \, = \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, 2 \partial_z u_{y} \, - \xi B_0 \partial_x \partial_z B_{y}\\
-- \, J\left(\Psi, \nabla^2 \Psi\right) \, + \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi,
-\end{multline}
-
-in other words, the $\Psi$ equation picks up a linear $B_y$ term.
-
-Following the same prescription for weakly nonlinear tuning, multiscale expansion, and perturbation expansion, we find the equations for the HMRI up to order $\epsilon^3$.
-
-To $\mathcal{O}(\epsilon)$, we find
-
-\beq
-\mathcal{L} \mathbf{V_1} + \xi B_0 \mathcal{H} \partial_z \mathbf{V_1} = 0
-\eeq
-
-To $\mathcal{O}(\epsilon^2)$,
-
-\beq
-\mathcal{L}\mathbf{V_2} = \mathbf{N_2} - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1} + \xi \mathcal{H} \partial_z \mathbf{V_2} - \xi \mathcal{H}\partial_Z \mathbf{V_1}
-\eeq
-
-To $\mathcal{O}(\epsilon^3)$,
-
-\begin{multline}
-\mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, \\
-- \partial_z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1} + \xi \mathcal{H} \partial_z \mathbf{V_3} - \xi \mathcal{H} \partial_Z \mathbf{V_2} + \xi \mathcal{H} \partial_z \mathbf{V_1}.
-\end{multline}
-
-In the above,
-
-\beq
-\mathcal{H} = \left[\begin{matrix}
-0 & 0 & 0 & \partial_x \\
-0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\ \end{matrix} \right].
-\eeq
-
-The first amplitude equation (\ref{ampeq1}) therefore becomes
-
-\begin{multline}
-\label{ampeq1_hmri}
-a \partial_T \alpha \, = - \, c \alpha \left|\alpha^2\right| \, - \, \widetilde{c} \alpha \partial_Z \beta \,  - b i k_c \alpha \, + \, h \partial_Z^2 \alpha \, \\
-+ \, g i k_c^3 \alpha \, + \, q \xi \partial_Z^2 \alpha \, + \, w i k_c \xi \alpha,
-\end{multline}
-
-where
-
-\beq
-q = \left< \mathbb{V}^\dagger \cdot \left(\mathcal{H} \mathbb{V}_{21}\right)^* \right>,
-\eeq
-
-and
-
-\beq
-w = \left< \mathbb{V}^\dagger \cdot \left(\mathcal{H} \mathbb{V}_{11}\right)^*\right>.
-\eeq
-
-The second amplitude equation (\ref{ampeq2}) becomes
-
-\beq\label{ampeq2_hmri}
-\partial_T \beta = - m \partial_Z^2 \beta + \frac{1}{\reym} \partial_Z^2 \beta \, - \, s \xi 2 \left|\alpha\right|\partial_Z \alpha \, - \, f\xi\partial_Z^2 \beta,
-\eeq
-
-where
-
-\beq
-s = \left<\mathbb{U}_{11} \cdot \left(\mathcal{H}\mathbb{V}_{20}\right)^*\right>
-\eeq
-
-and
-
-\beq
-f = \left< \mathbb{U}_{11} \cdot \left(\mathcal{H}\mathbb{U}_{20}\right)^*\right>.
-\eeq
-
-If $s = 0$, as it is for the standard MRI, the two amplitude equations will again decouple.
-
 \section{Discussion}\label{discussion}
@@
 \begin{split}
-N_{20}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - k_c^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i k_c \partial_x^2 \Psi_{11}^* - i k_c^3 \Psi_{11}^*\right) \\
-& \frac{2}{\beta} \partial_x A_{11} \cdot \left(i k_c \partial_x^2 A_{11}^* - i k_c^3 A_{11}^*\right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
+N_{20}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - k_c^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i k_c^3 \Psi_{11}^* - i k_c \partial_x^2 \Psi_{11}^*\right) \\
+& +\frac{2}{\beta} \partial_x A_{11} \cdot \left(i k_c^3 A_{11}^* - i k_c \partial_x^2 A_{11}^* \right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
 \end{split}
@@
 \beq
-N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* - \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11}^* + \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}^*
+N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* + \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11}^* - \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}^*
 \eeq
@@
 \beq
-N_{20}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11}^* - \partial_x \Psi_{11} \cdot i k_c B_{11}^* - i k_c A_{11} \cdot \partial_x u_{11}^* + \partial_x A_{11} \cdot i k_c u_{11}^*
+N_{20}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11}^* + \partial_x \Psi_{11} \cdot i k_c B_{11}^* - i k_c A_{11} \cdot \partial_x u_{11}^* - \partial_x A_{11} \cdot i k_c u_{11}^*
 \eeq \\
@@
 
+%Correct -- checked 11/14/15
 \beq

### 2016-07-15 paper/mri_paper.tex
added final mri sim slice to thingap paper with rms structure along z direction. could stretch these out more?

@@
 The $u_y$ perturbations shown are first order (purple), second order (orange), and total (black).
-The blue shaded area represents the region where the second order and total perturbations are opposing the background shear because they are negative, and the red shaded region indicates where the second order and total perturbations are opposing the background shear because they are positive.}\label{allorders_velocity}
+The blue shaded area represents the region where the second order and total perturbations are opposing the background shear because they are negative, and the red shaded region indicates where the second order and total perturbations are opposing the background shear because they are positive.}
 \label{stress_uy_cuts}
@@
 
-We conduct a full study of the MRI parameter space, focussed on the behavior with changing magnetic Prandtl number ($\mathrm{Pm} \equiv \frac{\mathrm{Rm}}{\mathrm{Re}} \equiv \frac{\nu}{\eta}$).
+We conduct a full study of the MRI parameter space, focused on the behavior with changing magnetic Prandtl number ($\mathrm{Pm} \equiv \frac{\mathrm{Rm}}{\mathrm{Re}} \equiv \frac{\nu}{\eta}$).
 In protoplanetary disks, as well as in liquid metal MRI experiments, the magnetic Prandtl number can be very small ($\mathrm{Pm} \sim 10^{-6}$).
@@
 This allows us to directly compare simulation and theory.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../figures/final_slice_MRI_run_Rm5_00e+01_eps0_00e+00_Pm1_00e-03_beta2_50e+01_Q7_50e-01_qsh1_50e+00_Omega1_00e+00_nx256_nz256_Lz10_CFL.pdf}
+\caption{The final timestamp of each simulation.}\label{fig:final_slices}
+\end{figure*}
 

### 2016-07-15 paper/mri_paper.tex
added streamfunc plots to thingap paper
caption for parameter space plot

@@
 \centering
-\includegraphics[trim=0cm 1cm 0cm 5cm, scale=0.45]{allorders_velocity.eps}
-\caption{First order (left) and second order (right) velocity perturbations.
+\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_velocity.png}%{allorders_velocity.eps}
+\caption{First order (left), second order (center), and total (right) velocity perturbations.
 Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds.
-The colorbar represents azimuthal velocity.}\label{allorders_velocity}
+Colorbar represents azimuthal velocity.}\label{allorders_velocity}
 \end{figure*}
@@
 \centering
-\includegraphics[scale=0.45]{allorders_Bfield.eps}
-\caption{First order (left) and second order (right) magnetic field perturbations.
+\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_Bfield.png}
+\caption{First order (left), second order (center), and total (right) magnetic field perturbations.
 Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to higher magnetic field strengths.
-The colorbar represents azimuthal magnetic field strength.}\label{allorders_Bfield}
+Colorbar represents azimuthal magnetic field strength.}\label{allorders_Bfield}
 \end{figure*}

### 2016-07-15 paper/widegap_mri_paper.tex
parameter space figure added to widegap paper
changed widegap paper to twocolumn format bc we need that for apj....right?
caption for parameter space plot

@@
-\documentclass[onecolumn]{emulateapj}
+%\documentclass[onecolumn]{emulateapj}
+\documentclass{emulateapj}
 \newcommand{\beq}{\begin{equation}}
@@
 
-\begin{table}[ht]
+\begin{widetext}
+\begin{table*}[ht]
 \normalsize
@@
 \end{tabular}}
-\end{table}
+\end{table*}
 
@@
 \end{align}
+\end{widetext}
 
@@
 We solve the system subject to periodic vertical boundary conditions and no-slip, perfectly conducting radial boundary conditions.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../figures/widegap_paramspace_crit_params_Pm_1E-3}
+\caption{Growth rates in the ($\reym, k_z$) plane.
+Color map shows growth rate found by solving the linear eigenvalue problem for each ($\reym, k_z$) in the grid.
+The eigenvalue problem was solved for the widegap parameters listed in Table \ref{table:parameters}.
+Overlaid contours show growth rates at [-2E-3, -1E-3, -1E-5, 1E-5, 1E-3, 2E-3], where dashed contours represent negative values.
+The gray dotted line shows the interpolated marginal stability curve.
+The critical parameters $\reym_c = 0.84$ and $k_c = 0.0127$ correspond to the smallest parameter values that yield a zero growth rate.}
+\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure*}
 \centering
@@
 Eigenfunctions are normalized such that either the real or imaginary part is zero.}
-\end{figure}
+\end{figure*}
 
-\begin{figure}
+\begin{figure*}
 \centering
@@
 \caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.}
-\end{figure}
+\end{figure*}
 

### 2016-07-15 python/run_amplitude.py

@@ -11,13 +11,19 @@ import h5py
 # Rm = 4.876
 # Pm = 5e-4
 
-Q = 0.7470
-Rm = 4.879
-Pm = 1e-3
+#Q = 0.7470
+#Rm = 4.879
+#Pm = 1E-2#1e-3
+
+tgc = h5py.File('/Users/susanclark/weakly_nonlinear_mri/data/pm_sat_coeffs.h5', 'r')
+Pm = tgc['Pm'].value[0]
+Rm = tgc['Rm_c'].value[0]
+Q = tgc['Q_c'].value[0]
+
 q = 1.5
 beta = 25.0
 
-gridnum = 50
+gridnum = 128
 x_basis = de.Chebyshev('x',gridnum, interval=(0, 1))
 domain = de.Domain([x_basis], np.complex128, comm=MPI.COMM_SELF)
 print("running at gridnum", gridnum)

### 2016-07-16 paper/widegap_mri_paper.tex
high res run needs more time

@@
 While the MRI is a local instability, certain properties of the local problem are not generic to the global problem.
-Channel modes, a class of MRI mode that dominates MRI evolution and saturation in the local approximation, are artificially over-prominent in the shearing box (e.g.
+In the shearing box, linear evolution is dominated by a class of MRI mode known as channel modes.
+%Channel modes, a class of MRI mode that dominates MRI evolution and saturation in the local approximation, are artificially over-prominent in the shearing box (e.g.
 \citei{Latter:2015}).
@@
 However the rotation profiles needed to excite HMRI may be prohibitively steeper than Keplerian depending on the boundary conditions, and so its role in astrophysical disks is currently debated (\citei{Liu:2006}, \citei{Rudiger:2007}, \citei{Kirillov:2013}).
-However, the HMRI is significantly easier to excite in a laboratory setting than the standard MRI, and has already been detected in the laboratory by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
+The HMRI is significantly easier to excite in a laboratory setting than the standard MRI, and has already been detected in the laboratory by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
+The standard MRI has yet to be detected in the laboratory, although several plasma experiments are making progress.
 

### 2016-07-23 paper/widegap_mri_paper.tex
fixed widegap neutral mode description

@@
 Because that symmetry is not preserved in the wide-gap case, \citeauthor{Umurhan:2007hs} postulate that slow variation in the wide-gap geometry will be governed by two coupled amplitude equations.
-However, the wide-gap conducting boundary condition $J_z = \frac{1}{r} \partial_r (r B_\phi) = 0$, together with the purely geometric term in Equation \ref{eq:Bphi_perturbed}, conspire to prevent the wide-gap geometry from sustaining a neutral mode.
+%However, the wide-gap conducting boundary condition $J_z = \frac{1}{r} \partial_r (r B_\phi) = 0$, together with the purely geometric term in Equation \ref{eq:Bphi_perturbed}, conspire to prevent the wide-gap geometry from sustaining a neutral mode.
+However, the purely geometric term in Equation \ref{eq:Bphi_perturbed} prevents the wide-gap geometry from sustaining a neutral mode.
 We note that a neutral mode of the form $B_\phi(r) \propto \frac{1}{r}$ would exist in a resistance-free approximation.

### 2016-07-26 python/widegap/run_widegap_amplitude.py
run many more Pms

@@ -16,181 +16,182 @@ xi = 0
 R1 = 5
 R2 = 15
 
-indx = 0
-crit_params = h5py.File('../../data/widegap_crit_Rm_Q.h5', 'r')
-Q = crit_params['Q_c'].value[indx]
-Rm = crit_params['Rm_c'].value[indx]
-Pm = crit_params['Pm'].value[indx]
+for indx in [1, 2, 3, 4, 5, 6, 7, 8, 9]:
 
-# Hollerbach & Rudiger
-"""
-Pm = 1.0E-6
-Re = 1521
-Rm = Pm*Re
-xi = 4
-Q = 2.33
-Ha = 16.3
-beta = (2*Re*Rm)/(Ha**2)
-R1 = 1.
-R2 = 2.
-mu_omega = 0.27
-Omega1 = 313.55
-Omega2 = Omega1*mu_omega
-"""
+    crit_params = h5py.File('../../data/widegap_crit_Rm_Q.h5', 'r')
+    Q = crit_params['Q_c'].value[indx]
+    Rm = crit_params['Rm_c'].value[indx]
+    Pm = crit_params['Pm'].value[indx]
 
-gridnum = 128
+    # Hollerbach & Rudiger
+    """
+    Pm = 1.0E-6
+    Re = 1521
+    Rm = Pm*Re
+    xi = 4
+    Q = 2.33
+    Ha = 16.3
+    beta = (2*Re*Rm)/(Ha**2)
+    R1 = 1.
+    R2 = 2.
+    mu_omega = 0.27
+    Omega1 = 313.55
+    Omega2 = Omega1*mu_omega
+    """
 
-r_basis = de.Chebyshev('r', gridnum, interval=(R1, R2))
-domain = de.Domain([r_basis], np.complex128, comm=MPI.COMM_SELF)
-print("running at gridnum", gridnum)
+    gridnum = 128
 
-aa = AmplitudeAlpha(domain, Q = Q, Rm = Rm, Pm = Pm, xi = xi, Omega1 = Omega1, Omega2 = Omega2, beta = beta)
-aa.print_coeffs()
+    r_basis = de.Chebyshev('r', gridnum, interval=(R1, R2))
+    domain = de.Domain([r_basis], np.complex128, comm=MPI.COMM_SELF)
+    print("running at gridnum", gridnum)
+
[…]

### 2016-07-26 paper/widegap_mri_paper.tex
run many more Pms
run remaining Pms

@@
 
-where $\mathcal{D}$, $\mathcal{L}$, and $\widetilde{\mathcal{G}}$ are matrices defined in Appendix \ref{app:basic_equations}, and $\mathbf{N}$ is a vector containing all nonlinear terms.
+where $\mathcal{D}$, $\mathcal{L}$, and $\widetilde{\mathcal{G}}$ are matrices defined in Appendix \ref{app:basic_equations}, and $\mathbf{N}$ is a vector containing all nonlinear terms defined in Appendix \ref{app:nonlinear_terms}.
 We then expand the variables in a perturbation series $\mathbf{V} = \epsilon \mathbf{V}_1 + \epsilon^2 \mathbf{V}_2 + \epsilon^2 \mathbf{V}_3 + h.o.t.$ The perturbed system can then be expressed at each order by the equations
@@
 \mathcal{O}(\epsilon^2)&: \mathcal{L} \mathbf{V}_2 + \xi \widetilde{\mathcal{H}} \mathbf{V}_2 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_1 + \xi \mathcal{H} \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \label{eq:ordere2}\\
-\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \xi \widetilde{\mathcal{H}} \mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2 + \xi \mathcal{H}\partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 - \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3 = 0.
+\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \xi \widetilde{\mathcal{H}} \mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2 \\ \nonumber
+&+ \xi \mathcal{H}\partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 - \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3 = 0.
 \label{eq:ordere3}
@@
 We use our fiducial parameters for the standard MRI ($\xi = 0$).
-Eigenfunctions are normalized such that either the real or imaginary part is zero.}
+First-order eigenfunctions are normalized such that they are either purely real or purely imaginary, and such that $\int \Psi_{11} \mathrm{d}r = 1$.
+Adjoint homogenous eigenfunctions are normalized such that $\left< V_{11}^\dagger \cdot \mathcal{D} V_{11} \right> = 1$.}
 \end{figure*}
@@
 
-First, the result of the weakly nonlinear analysis is a single amplitude equation for $\alpha$. We find
+First, the result of the weakly nonlinear analysis is a single amplitude equation for $\alpha$.
+This amplitude equation is found by enforcing a solvability criterion on Equation \ref{eq:ordere3}.
+We find
 
@@
 a Ginzburg-Landau equation.
+
 Here we note a departure from the behavior of the thin-gap system.

### 2016-07-26 python/widegap/run_widegap_amplitude.py

@@ -16,7 +16,7 @@ xi = 0
 R1 = 5
 R2 = 15
 
-for indx in [1, 2, 3, 4, 5, 6, 7, 8, 9]:
+for indx in [11, 12, 13, 14, 15]:
 
     crit_params = h5py.File('../../data/widegap_crit_Rm_Q.h5', 'r')
     Q = crit_params['Q_c'].value[indx]

### 2016-07-26 paper/widegap_mri_paper.tex
one more run

@@
 \citei{Latter:2015}).
+These are linear modes which also happen to be exact solutions to the \textit{nonlinear} local MRI equations.
+Runaway growth is avoided in this paradigm by the instability of the channel modes themselves, which are destroyed by parasitic (secondary) instabilities (\citei{Goodman:1994ul}, \citei{Pessah:2010ic}).
 
@@
 
+\bibitem[Goodman \& Xu(1994)]{Goodman:1994ul}
+Goodman, J, Xu, G, 1994, ApJ, 432, 213
+
 \bibitem[{Hollerbach \& R\"udiger(2005)}]{Hollerbach:2005tr}
@@
 Liu, W, Goodman, J, Herron, I, and Ji, H, 2006, Phys. Rev. E 74, 056302
+
+\bibitem[Pessah(2010)]{Pessah:2010ic}
+Pessah, M, 2010, ApJ, 716, 1012
 

### 2016-07-26 python/widegap/allorders2_widegap.py

@@ -169,13 +169,13 @@ class MRI():
         #logger.warn("Normalizing according to norm(psi)")
         #norm = np.linalg.norm(psi['g'])
         
-        #logger.warn("Normalizing according to integral(psi)")
-        #intpsi = psi.integrate('r')
-        #norm = intpsi['g'][0]
+        logger.warn("Normalizing according to integral(psi)")
+        intpsi = psi.integrate('r')
+        norm = intpsi['g'][0]
         
-        logger.warn("Normalizing according to integral(u)")
-        intu = u.integrate('r')
-        norm = intu['g'][0]
+        #logger.warn("Normalizing according to integral(u)")
+        #intu = u.integrate('r')
+        #norm = intu['g'][0]
         
         psi['g'] = psi['g']/norm
         u['g'] = u['g']/norm

### 2016-07-26 python/widegap/run_widegap_amplitude.py

@@ -51,7 +51,7 @@ for indx in [10]:
 
     #fn_root = "/Users/susanclark/weakly_nonlinear_mri/data/"
     fn_root = "../../data/"
-    fn = fn_root + "widegap_amplitude_parameters_Q_{:03.2f}_Rm_{:04.4f}_Pm_{:.2e}_Omega1_{:05.2f}_Omega2_{:05.2f}_beta_{:.2f}_xi_{:.2f}_gridnum_{}_intnormu.h5".format(Q, Rm, Pm, Omega1, Omega2, beta, xi, gridnum)
+    fn = fn_root + "widegap_amplitude_parameters_Q_{:03.2f}_Rm_{:04.4f}_Pm_{:.2e}_Omega1_{:05.2f}_Omega2_{:05.2f}_beta_{:.2f}_xi_{:.2f}_gridnum_{}_intnorm.h5".format(Q, Rm, Pm, Omega1, Omega2, beta, xi, gridnum)
     with h5py.File(fn,'w') as f:
         r = f.create_dataset("r", data=aa.r)
         o1psi = f.create_dataset("psi11", data=aa.o1.psi['g'])

### 2016-07-28 paper/widegap_mri_paper.tex
more widegap intro

@@
 Runaway growth is avoided in this paradigm by the instability of the channel modes themselves, which are destroyed by parasitic (secondary) instabilities (\citei{Goodman:1994ul}, \citei{Pessah:2010ic}).
+The growth of parasitic modes provides a saturation avenue for channel mode-dominated flows, yet this is unlikely to be the dominant saturation mechanism in laboratory experiments or astrophysical disks, as channel modes are artificially over-prominent in the shearing box (e.g.
+\citei{Latter:2015}).
+
+A number of saturation mechanisms have been proposed for the MRI which do not rely on channel modes dominating the flow.
+
+Our investigation is astrophysically motivated, but we also intend our theory to be relevant to laboratory experiments.
+Several experimental efforts are attempting to observe the MRI in the laboratory, which would allow the study of a crucial astrophysical phenomenon in a controlled setting.
+Unfortunately, detection of the MRI has so far proven elusive.
+\citei{Sisan:2004ig} claimed to detect the MRI, but most likely detected unrelated MHD instabilities instead (\citei{Hollerbach:2009ig}, \citei{Gissinger:2011td}).
+Most relevant to our work is the Princeton Plasma Physics Laboratory (PPPL) MRI Experiment, a liquid gallium Taylor-Couette flow exposed to an axial magnetic field (\citei{Ji:2001kd}).
 
@@
 Thus through judicious choice of cylinder rotation rates, we can set $q(r_0) = 3/2$, for quasi-Keplerian flow.
-Note that in the narrow gap approximation we have the freedom to choose any rotation profile, as the narrow gap width imposes a linear shear (constant $q$).
+Note that the narrow gap approximation imposes a linear shear (constant $q$).
 Our base velocity is $\mathbf{u_0} = r\Omega(r) \phihat$.
@@
 \mathcal{O}(\epsilon)&: \mathcal{L} \mathbf{V}_1 + \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \mathcal{D} \partial_t \mathbf{V}_1 = 0 \label{eq:ordere}\\
-\mathcal{O}(\epsilon^2)&: \mathcal{L} \mathbf{V}_2 + \xi \widetilde{\mathcal{H}} \mathbf{V}_2 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_1 + \xi \mathcal{H} \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \label{eq:ordere2}\\
-\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \xi \widetilde{\mathcal{H}} \mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2 \\ \nonumber
+\mathcal{O}(\epsilon^2)&: \mathcal{L} \mathbf{V}_2 + \xi \widetilde{\mathcal{H}} \mathbf{V}_2 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_1 + \xi \mathcal{H} \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \\\label{eq:ordere2}
+\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \xi \widetilde{\mathcal{H}} \mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2\nonumber \\ 
 &+ \xi \mathcal{H}\partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 - \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3 = 0.
-\label{eq:ordere3}
+\\\label{eq:ordere3}
 \end{align}
@@
 
+\bibitem[{Gissinger et al.(2011)}]{Gissinger:2011td}
+Gissinger, C, Ji, H, Goodman, J, Phys. Rev. E, 2011, 84, 026308
+
 \bibitem[{Goodman \& Ji(2002)}]{Goodman:2002ix}
@@
 
+\bibitem[{Hollerbach(2009)}]{Hollerbach:2009ig}
+Hollerbach, R, Proc. R. Soc. A, 2009, 465, 2107
+
 \bibitem[{Hollerbach \& R\"udiger(2005)}]{Hollerbach:2005tr}
 Hollerbach, R. \& R\"udiger, G., 2005, Phys. Rev. Lett. 95, 124501
+
+\bibitem[{Ji et al.(2001)}]{Ji:2001kd}
+Ji, H, Goodman, J, Kageyama, A, Mon. Not. R. Astron. Soc., 2001, 325, 1
 
@@
 
+\bibitem[{Sisan et al.(2004)}]{Sisan:2004ig}
+Sisan, D R, Mujica, N, Tillotson, W A, Huang Y-M, Dorland, W, Hassam, A B, Antonsen, T M, Lathrop, D P, Phys Rev Lett, 2004, 93, 11
+
 \bibitem[{Stefani et al.(2006)}]{Stefani:2006iv}

### 2016-07-29 paper/widegap_mri_paper.tex
more intro on experiments

@@
 Unfortunately, detection of the MRI has so far proven elusive.
-\citei{Sisan:2004ig} claimed to detect the MRI, but most likely detected unrelated MHD instabilities instead (\citei{Hollerbach:2009ig}, \citei{Gissinger:2011td}).
+\citei{Sisan:2004ig} claimed to detect the MRI in a spherical Couette flow, but most likely detected unrelated MHD instabilities instead (\citei{Hollerbach:2009ig}, \citei{Gissinger:2011td}).
 Most relevant to our work is the Princeton Plasma Physics Laboratory (PPPL) MRI Experiment, a liquid gallium Taylor-Couette flow exposed to an axial magnetic field (\citei{Ji:2001kd}).
+Some theoretical work designed to complement the Princeton MRI experiment involves direct simulation of the experimental conditions, much of it focused on the specific challenges in identifying MRI signatures despite apparatus-driven spurious flows (e.g.
+\citei{Gissinger:2012gc}).
+In particular, the vertical endcaps on a laboratory MRI apparatus drives meridional flows which both inhibit the excitement of MRI and obscure its detection.
+The Princeton MRI experiment employs split, independently rotating endcaps to mitigate these flows (\citei{Schartman:2009df}).
+Our work here assumes an infinite vertical domain, an idealization that is theoretically expedient but experimentally impractical.
+However, our setup is designed to be an accurate treatment of the radial dimension of the flow in a Taylor Couette apparatus like the one used in the Princeton MRI experiment.
 
@@
 The HMRI is significantly easier to excite in a laboratory setting than the standard MRI, and has already been detected in the laboratory by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
-The standard MRI has yet to be detected in the laboratory, although several plasma experiments are making progress.
+%The standard MRI has yet to be detected in the laboratory, although several plasma experiments are making progress.
 
@@
 \bibitem[{Gissinger et al.(2011)}]{Gissinger:2011td}
-Gissinger, C, Ji, H, Goodman, J, Phys. Rev. E, 2011, 84, 026308
+Gissinger, C, Ji, H, Goodman, J, 2011, Phys. Rev. E 84, 026308
+
+\bibitem[{Gissinger et al.(2012)}]{Gissinger:2012gc}
+Gissinger, C, Goodman, J, \& Ji, H, 2012, Physics of Fluids 24, 074109
 
@@
 
+\bibitem[{Schartman et al.(2009)}]{Schartman:2009df}
+Schartman, E, Ji, H, Burin, M J, Rev Sci Instrum, 2009, 80, 024501
+
 \bibitem[{Sisan et al.(2004)}]{Sisan:2004ig}

### 2016-07-29 paper/widegap_mri_paper.tex
more intro

@@
 
-Because of the MRI's complexity, analytical treatments of the MRI use some combination of convenient approximations: ideal MHD, linearized equations, expedient boundary conditions, and simplified geometries.
+Because of the MRI's complexity, analytical treatments of the MRI use some combination of convenient approximations, such as ideal MHD, linearized equations, expedient boundary conditions, and simplified geometries.
 In particular, much of the analytical work on MRI saturation has focused on local approximations, often representing a section of a disk by solving the MHD equations in an isolated region subject to shear periodic boundary conditions (the ``shearing box").
@@
 
+The theory we develop here employs radial boundary conditions, and so MRI modes are global and channel modes are not present.
 A number of saturation mechanisms have been proposed for the MRI which do not rely on channel modes dominating the flow.
+The MRI feeds off of the free energy from differential rotation, and so a modification of the background shear may cause saturation (e.g.
+\citei{Knobloch:2005ba}, \citei{Umurhan:2007hs}).
+The MRI may transfer its free energy into the magnetic field, and saturate when the field is too strong to be susceptible to the MRI (e.g.
+\citei{Ebrahimi:2009ey}).
+The MRI may saturate differently depending on the particular parameter regime under investigation, and so our challenge in not only in identifying possible saturation mechanisms, but in understanding how and when each applies in different astrophysical environments.
 
@@
 Kirillov, O, \& Stefani, F, 2013, Phys. Rev. Lett. 111, 061103
+
+\bibitem[{Knobloch \& Julien(2005)}]{Knobloch:2005ba}
+Knobloch, E, Julien K, 2005, Physics of Fluids, 17, 094106
 

### 2016-07-31 python/widegap/animate_vel.py
added 3D velocity visualization

@@ -151,6 +151,10 @@ if widegap is True:
     Vboth_uz1 = eps*((1/rgrid)*satamp*obj['psi11_r'].value*eiqz) + eps**2*(-satamp**2*(1/rgrid)*ei2qz*obj['psi22_r'].value - (np.abs(satamp))**2*(1/rgrid)*obj['psi20_r'].value*ei0qz)
     Vboth_ur1 = eps*((1/rgrid)*satamp*obj['psi11'].value*eiqz_z) + eps**2*(satamp**2*ei2qz_z*obj['psi22'].value)
 
+    #Vboth_Bz1 = eps*((1/r)*satamp*obj['A11_r'].value*eiqz) + eps**2*(-satamp**2*(1/r)*ei2qz*obj['A22_r'].value - (np.abs(satamp))**2*(1/r)*obj['A20_r'].value*ei0qz)
+    #Vboth_Br1 = eps*((1/r)*satamp*obj['A11'].value*eiqz_z) + eps**2*(satamp**2*ei2qz_z*obj['A22'].value)
+
+
 else:
     file_root = "/Users/susanclark/weakly_nonlinear_mri/data/"
     fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_128.h5"
@@ -198,7 +202,11 @@ else:
 
 phi_interpolator = interpolate.interp2d(rgrid, zgrid, Vboth_u)
 z_interpolator = interpolate.interp2d(rgrid, zgrid, Vboth_uz1) 
-r_interpolator = interpolate.interp2d(rgrid, zgrid, Vboth_ur1)        
+r_interpolator = interpolate.interp2d(rgrid, zgrid, Vboth_ur1)   
+
+#B_phi_interpolator = interpolate.interp2d(rgrid, zgrid, Vboth_B)
+#B_z_interpolator = interpolate.interp2d(rgrid, zgrid, Vboth_Bz1) 
+#B_r_interpolator = interpolate.interp2d(rgrid, zgrid, Vboth_Br1)        
 
 def vel_deriv((r, phi, z), t0):
     #Compute the 3D velocities.
@@ -216,6 +224,13 @@ np.random.seed(1)
 x0[:, 0] = np.sort(5 + 10 * np.random.random(N_trajectories))
 x0[:, 1] = 2*np.pi * np.random.random(N_trajectories)
[…]

### 2016-08-01 paper/widegap_mri_paper.tex
added 3D velocity visualization
more text, rearranged eqns

@@
 \beq
-\begin{split}
+%\begin{split}
 \label{eq:Psi_perturbed}
-\frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r^2}B_0 \xi \partial_z B_\phi - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] \\
-= - J(\Psi, \frac{1}{r^2} ( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi) ) + \frac{2}{\beta} J(A, \frac{1}{r^2} ( \nabla^2 A - \frac{2}{r} \partial_rA) ) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi  + \frac{2}{r} u_\phi \partial_z u_\phi \\
-\end{split}
+\frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r^2}B_0 \xi \partial_z B_\phi - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] = N^{(\Psi)}
+%\end{split}
 \eeq
@@
 \begin{align}
-& \partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} ( \nabla^2 \uphi - \frac{1}{r^2} \uphi ) = \frac{2}{\beta} \frac{1}{r} J(A, B_\phi) - \frac{1}{r} J(\Psi, \uphi) + \frac{2}{\beta}\frac{1}{r^2} B_\phi \partial_z A - \frac{1}{r^2} \uphi \partial_z \Psi \label{eq:uphi_perturbed} \\
-& \partial_t A - B_0 \partial_z \Psi - \frac{1}{\reym} ( \nabla^2 A - \frac{2}{r} \partial_r A )= \frac{1}{r} J(A, \psi)  \label{eq:A_perturbed}\\
+& \partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} ( \nabla^2 \uphi - \frac{1}{r^2} \uphi ) = N^{u} \label{eq:uphi_perturbed} \\
+& \partial_t A - B_0 \partial_z \Psi - \frac{1}{\reym} ( \nabla^2 A - \frac{2}{r} \partial_r A )= N^{(A)}  \label{eq:A_perturbed}\\
   \label{eq:Bphi_perturbed}
-& \partial_t B_\phi + \frac{1}{r^2} u_0 \partial_z A - B_0 \partial_z u_\phi - \frac{1}{r} \partial_r u_0 \partial_z A - \frac{2}{r^3} B_0 \xi \partial_z \Psi - \frac{1}{\reym} (\nabla^2 B_\phi - \frac{1}{r^2} B_\phi ) = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)
-+ \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A 
+& \partial_t B_\phi + \frac{1}{r^2} u_0 \partial_z A - B_0 \partial_z u_\phi - \frac{1}{r} \partial_r u_0 \partial_z A - \frac{2}{r^3} B_0 \xi \partial_z \Psi - \frac{1}{\reym} (\nabla^2 B_\phi - \frac{1}{r^2} B_\phi ) = N^{(B)}
 \end{align}
+
+The righthand side of the equations contain the nonlinear terms
+
+\beq
+N^{(\Psi)} = - J(\Psi, \frac{1}{r^2} ( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi) ) + \frac{2}{\beta} J(A, \frac{1}{r^2} ( \nabla^2 A - \frac{2}{r} \partial_rA) ) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi  + \frac{2}{r} u_\phi \partial_z u_\phi
+\eeq
+\beq
+N^{(u)} = \frac{2}{\beta} \frac{1}{r} J(A, B_\phi) - \frac{1}{r} J(\Psi, \uphi) + \frac{2}{\beta}\frac{1}{r^2} B_\phi \partial_z A - \frac{1}{r^2} \uphi \partial_z \Psi 
+\eeq
+\beq
+N^{(A)} = \frac{1}{r} J(A, \psi)
+\eeq
+\beq
+N^{(B)} = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi) + \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A 
+\eeq
+
 \end{widetext}
@@
 First-order eigenfunctions are normalized such that they are either purely real or purely imaginary, and such that $\int \Psi_{11} \mathrm{d}r = 1$.
-Adjoint homogenous eigenfunctions are normalized such that $\left< V_{11}^\dagger \cdot \mathcal{D} V_{11} \right> = 1$.}
+Adjoint homogenous eigenfunctions are normalized such that $\left< V_{11}^\dagger \cdot \mathcal{D} V_{11} \right> = 1$.}\label{fig:linear_eigenfunctions}
 \end{figure*}
@@
 \includegraphics[width=\textwidth]{../python/widegap/figures/nonlinear_2_3_widegap_fiducial_eigenfunctions.pdf}
-\caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.}
+\caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.}\label{fig:N2_N3}
 \end{figure*}
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../figures/widegap_3D_vel_screenshot.png}
+\caption{Visualization of the velocity perturbation field in the fiducial SMRI case.}
+\end{figure}
 
@@
 This parameterization coupled with the boundary conditions lead us to an ansatz linear solution $\mathbf{V}_1 = \alpha(Z, T) \mathbb{V}_{11}(r) e^{i k_z z} + c.c.$, where the radial variation is contained in $\mathbb{V}_{11}$.
-We defer full expressions and derivations thereof to the Appendix, and focus here on features of the wide gap system which distinguish it from the thin gap limit.
+We defer full expressions and derivations thereof to the Appendix, and focus here on features of the wide gap system which distinguish it from the narrow gap limit.
 
-First, the result of the weakly nonlinear analysis is a single amplitude equation for $\alpha$.
-This amplitude equation is found by enforcing a solvability criterion on Equation \ref{eq:ordere3}.
-We find
+The result of the weakly nonlinear analysis is a single amplitude equation for $\alpha$. This amplitude equation is found by enforcing a solvability criterion on Equation \ref{eq:ordere3}. We find
 
@@
 
-a Ginzburg-Landau equation.
+a Ginzburg-Landau equation (GLE).
+The GLE governs the weakly nonlinear amplitude behavior in a wide range of physical systems, including the narrow gap MRI (), Rayleigh-B\'enard convection, and hydrodynamic Taylor Couette flow.
+We emphasize that this is a model equation, valid only near marginality (see ).
+The dynamics of the GLE are determined by its coefficients, which are in turn determined by the linear eigenfunctions and nonlinear vectors plotted in Figures \ref{fig:linear_eigenfunctions} and \ref{fig:N2_N3}.
+Equation \ref{eq:gle} contains three coefficients: $b$, which determines the linear growth rate of the system, $d$, a diffusion coefficient, and $c$, the coefficient of the nonlinear term.
+When all of the coefficients of Equation \ref{eq:gle} are real, this is known as the real GLE, although the amplitude $\alpha$ remains complex.
+The real GLE is subject to several well-studied instabilities, including the Ekhaus and Zig-Zag instabilities.
+When the coefficients are complex, we have the complex GLE, a source of even richer phase dynamics than its real counterpart (e.g.
+\citei{Aranson:2002}).
 
-Here we note a departure from the behavior of the thin-gap system.
+For the standard MRI we derive a real GLE, and for the helical MRI we derive a complex GLE. 
+ This complex GLE is a natural form for the HMRI, as it is a traveling wave instability.
+
+Here we note a departure from the behavior of the narrow gap system.
 The purely conducting boundary condition states that the axial component of the current ($\mathbf{J}_z = [\nabla \times \mathbf{B}]_z)$ must be zero at the walls.
@@
 (I'm just guessing but this may lead to higher torque on inner cylinder -- should calculate)
-
-For both the standard and helical MRI we derive a GLE amplitude.
-For the standard MRI we find a real GLE, in which the coefficients of the GLE are purely real (the amplitude is still complex in general).
-For the helical MRI we find a complex GLE, a source of even richer phase dynamics than its real counterpart (e.g.
-\cite{Aranson:2002}).
-This complex GLE is a natural form for the HMRI, as it is a traveling wave instability.
 

### 2016-08-01 paper/widegap_mri_paper.tex
more widegap paper words

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../figures/widegap_paramspace_crit_params_Pm_1E-3}
+\caption{Growth rates in the ($\reym, k_z$) plane.
+Color map shows growth rate found by solving the linear eigenvalue problem for each ($\reym, k_z$) in the grid.
+The eigenvalue problem was solved for the widegap parameters listed in Table \ref{table:parameters}.
+Overlaid contours show growth rates at [-2E-3, -1E-3, -1E-5, 1E-5, 1E-3, 2E-3], where dashed contours represent negative values.
+The gray dotted line shows the interpolated marginal stability curve.
+The critical parameters $\reym_c = 0.84$ and $k_c = 0.0127$ correspond to the smallest parameter values that yield a zero growth rate.}
+\end{figure}
+
 Thus through judicious choice of cylinder rotation rates, we can set $q(r_0) = 3/2$, for quasi-Keplerian flow.
@@
 \begin{align}
-& \partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} ( \nabla^2 \uphi - \frac{1}{r^2} \uphi ) = N^{u} \label{eq:uphi_perturbed} \\
+& \partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} ( \nabla^2 \uphi - \frac{1}{r^2} \uphi ) = N^{(u)} \label{eq:uphi_perturbed} \\
 & \partial_t A - B_0 \partial_z \Psi - \frac{1}{\reym} ( \nabla^2 A - \frac{2}{r} \partial_r A )= N^{(A)}  \label{eq:A_perturbed}\\
@@
 N^{(B)} = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi) + \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A 
+\label{eq:nonlinear_B}
 \eeq
@@
 Note that in the above, $\nabla^2 f \equiv \partial_r^2 f + \partial_z^2 f + \frac{1}{r} \partial_r f$.
-Equations \ref{eq:Psi_perturbed} - \ref{eq:Bphi_perturbed} are nondimensionalized, where lengths have been scaled by $r_0$, velocities by $r_0 \Omega_0$, densities by $\rho_0$, and magnetic fields by $B_0$; where $B_0$ appears in the above it is formally unity.
+Equations \ref{eq:Psi_perturbed} - \ref{eq:nonlinear_B} are nondimensionalized, where lengths have been scaled by $r_0$, velocities by $r_0 \Omega_0$, densities by $\rho_0$, and magnetic fields by $B_0$; where $B_0$ appears in the above it is formally unity.
 $\Omega_0 = \Omega(r_0)$ is the rotation rate at the center of the channel.
@@
 
-We solve the system subject to periodic vertical boundary conditions and no-slip, perfectly conducting radial boundary conditions.
+We solve the system subject to periodic vertical boundary conditions and no-slip, perfectly conducting radial boundary conditions, namely
 
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{../figures/widegap_paramspace_crit_params_Pm_1E-3}
-\caption{Growth rates in the ($\reym, k_z$) plane.
-Color map shows growth rate found by solving the linear eigenvalue problem for each ($\reym, k_z$) in the grid.
-The eigenvalue problem was solved for the widegap parameters listed in Table \ref{table:parameters}.
-Overlaid contours show growth rates at [-2E-3, -1E-3, -1E-5, 1E-5, 1E-3, 2E-3], where dashed contours represent negative values.
-The gray dotted line shows the interpolated marginal stability curve.
-The critical parameters $\reym_c = 0.84$ and $k_c = 0.0127$ correspond to the smallest parameter values that yield a zero growth rate.}
-\end{figure}
+\beq
+\Psi = \partial_r \Psi = u = A = \partial_r (r B) = 0
+\eeq
+
+at $r = r_1, r_2$.
+
+We note that Equations \ref{eq:Psi_perturbed} - \ref{eq:Bphi_perturbed} are written in a nonstandard form, with the nonlinear terms on the righthand side.
+This choice has a practical motivation.
+As detailed in \S\ref{sec:wnl_analysis}, we expand these equations in a perturbation series and solve them order by order using a pseudospectral code.
+The code solves partial differential equations of the form $\mathrm{M} \partial_t \mathbf{V} + \mathrm{L} \mathbf{V} = \mathbf{F}$, where $\mathrm{M}$ and $\mathrm{L}$ are matrices and $\mathbf{F}$ is a vector containing any inhomogenous terms.
+The nonlinear terms in our perturbation analysis become inhomogenous term inputs to the solver.
 
@@
 This parameterization coupled with the boundary conditions lead us to an ansatz linear solution $\mathbf{V}_1 = \alpha(Z, T) \mathbb{V}_{11}(r) e^{i k_z z} + c.c.$, where the radial variation is contained in $\mathbb{V}_{11}$.
-We defer full expressions and derivations thereof to the Appendix, and focus here on features of the wide gap system which distinguish it from the narrow gap limit.
+
+We solve the equations at each order using Dedalus, an open source pseudospectral code (Burns et al.
+in prep).
+Because our domain is bounded only radially and not vertically, we solve the radial dimension on a basis of Chebyshev polynomials, and the vertical dimension on a Fourier series basis.
+We solve Equation \ref{eq:ordere} as a linear eigenvalue problem, and Equations \ref{eq:ordere2} and \ref{eq:ordere3} as linear boundary value problems.
+%should i mention spurious eigenvals?
+
+%We defer full expressions and derivations thereof to the Appendix, and focus here on features of the wide gap system which distinguish it from the narrow gap limit.
 

### 2016-08-01 paper/widegap_mri_paper.tex
coefficients plot added

@@
 The gray dotted line shows the interpolated marginal stability curve.
-The critical parameters $\reym_c = 0.84$ and $k_c = 0.0127$ correspond to the smallest parameter values that yield a zero growth rate.}
+The critical parameters $\reym_c = 0.84$ and $k_c = 0.0127$ correspond to the smallest parameter values that yield a zero growth rate.}\label{fig:growth_rates}
 \end{figure}
@@
 \eeq
+\end{widetext}
 
-\end{widetext}
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../python/widegap/figures/orders_1_2_widegap_fiducial_eigenfunctions.pdf}
+\caption{Eigenfunctions of the first order equations, first order adjoint homogenous equations, and second order equations.
+We use our fiducial parameters for the standard MRI ($\xi = 0$).
+First-order eigenfunctions are normalized such that they are either purely real or purely imaginary, and such that $\int \Psi_{11} \mathrm{d}r = 1$.
+Adjoint homogenous eigenfunctions are normalized such that $\left< V_{11}^\dagger \cdot \mathcal{D} V_{11} \right> = 1$.}\label{fig:linear_eigenfunctions}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../python/widegap/figures/nonlinear_2_3_widegap_fiducial_eigenfunctions.pdf}
+\caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.}\label{fig:N2_N3}
+\end{figure*}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../figures/widegap_coeffs_by_Pm.pdf}
+\caption{Coefficients.}\label{fig:coefficients}
+\end{figure}
+
 Just as in the weakly nonlinear analyses of \citei{Umurhan:2007hs} and \citei{Clark:2016}, we tune the system away from marginality by taking $B_0 \rightarrow B_0\left(1 - \epsilon^2\right)$, where $\epsilon \ll 1$.
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{../python/widegap/figures/orders_1_2_widegap_fiducial_eigenfunctions.pdf}
-\caption{Eigenfunctions of the first order equations, first order adjoint homogenous equations, and second order equations.
-We use our fiducial parameters for the standard MRI ($\xi = 0$).
-First-order eigenfunctions are normalized such that they are either purely real or purely imaginary, and such that $\int \Psi_{11} \mathrm{d}r = 1$.
-Adjoint homogenous eigenfunctions are normalized such that $\left< V_{11}^\dagger \cdot \mathcal{D} V_{11} \right> = 1$.}\label{fig:linear_eigenfunctions}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{../python/widegap/figures/nonlinear_2_3_widegap_fiducial_eigenfunctions.pdf}
-\caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.}\label{fig:N2_N3}
-\end{figure*}
-
 \begin{figure}
@@
 \includegraphics[width=\columnwidth]{../figures/widegap_3D_vel_screenshot.png}
-\caption{Visualization of the velocity perturbation field in the fiducial SMRI case.}
+\caption{Visualization of the velocity perturbation field in the fiducial SMRI case.
+The three dimensional path of twenty randomly placed tracer particles is plotted, where the particle velocity is defined as the total (first + second order) velocity perturbation.
+This is meant to be a visualization only of the velocity field, so the particle motion feels no contribution from the magnetic field.}
 \end{figure}
@@
 %should i mention spurious eigenvals?
+To analyze the MRI system at marginality, we fix the parameters listed in Table \ref{table:parameters} and determine the critical $\reym$ and $k_z$ by repeatedly solving Equation \ref{eq:ordere} to determine the smallest parameter values for which the fastest growing mode is zero.
+That is, we solve the linear eigenvalue problem for eigenvalues $\sigma = \gamma + i \omega$.
+Figure \ref{fig:growth_rates} shows linear MRI growth rates $\gamma$ in the $(\reym, k_z)$ plane.
+For the fiducial standard MRI parameters in Table \ref{table:parameters} we find critical parameters $\reym_c = 0.84$, $k_c = 0.0127$.
 
@@
 a Ginzburg-Landau equation (GLE).
-The GLE governs the weakly nonlinear amplitude behavior in a wide range of physical systems, including the narrow gap MRI (), Rayleigh-B\'enard convection, and hydrodynamic Taylor Couette flow.
-We emphasize that this is a model equation, valid only near marginality (see ).
+The GLE governs the weakly nonlinear amplitude behavior in a wide range of physical systems, including the narrow gap MRI (\citei{Umurhan:2007hs}), Rayleigh-B\'enard convection, and hydrodynamic Taylor Couette flow.
+We emphasize that this is a model equation, valid only near marginality (see \citei{Cross:1993el}).
 The dynamics of the GLE are determined by its coefficients, which are in turn determined by the linear eigenfunctions and nonlinear vectors plotted in Figures \ref{fig:linear_eigenfunctions} and \ref{fig:N2_N3}.
@@
 
+\bibitem[{Cross \& Hohenberg(1993)}]{Cross:1993el}
+Cross, M. C., \& Hohenberg, P. C., 1993, Rev. Mod. Phys. 65, 851
+
 \bibitem[{Ebrahimi et al.(2009)}]{Ebrahimi:2009ey}

### 2016-08-01 paper/widegap_mri_paper.tex
more words

@@
 \includegraphics[width=\columnwidth]{../figures/widegap_coeffs_by_Pm.pdf}
-\caption{Coefficients.}\label{fig:coefficients}
+\caption{Critical parameters $k_c$ and $\reym$, and coefficients of the Ginzburg-Landau equation (Equation \ref{eq:gle}) as a function of $Pm$.
+Note the very weak dependence of the coefficients on $Pm$.
+The saturation amplitude of the standard MRI system is very insensitive to $Pm$ in the wide gap case.}\label{fig:coefficients}
 \end{figure}
@@
 The GLE governs the weakly nonlinear amplitude behavior in a wide range of physical systems, including the narrow gap MRI (\citei{Umurhan:2007hs}), Rayleigh-B\'enard convection, and hydrodynamic Taylor Couette flow.
-We emphasize that this is a model equation, valid only near marginality (see \citei{Cross:1993el}).
+We emphasize that this is a model equation, valid only near marginality (e.g.
+\citei{Cross:1993el}).
 The dynamics of the GLE are determined by its coefficients, which are in turn determined by the linear eigenfunctions and nonlinear vectors plotted in Figures \ref{fig:linear_eigenfunctions} and \ref{fig:N2_N3}.
@@
 
-For the standard MRI we derive a real GLE, and for the helical MRI we derive a complex GLE. 
- This complex GLE is a natural form for the HMRI, as it is a traveling wave instability.
+%For the standard MRI we derive a real GLE, and for the helical MRI we derive a complex GLE. 
+ %This complex GLE is a natural form for the HMRI, as it is a traveling wave instability.
 
+For the standard MRI we derive a real GLE.
 Here we note a departure from the behavior of the narrow gap system.
@@
 N$_2$ and N$_3$) not to individually satisfy the boundary conditions on $\Psi$ and $u$.
-(I'm just guessing but this may lead to higher torque on inner cylinder -- should calculate)
+%(I'm just guessing but this may lead to higher torque on inner cylinder -- should calculate)
+
+We examine the behavior of the wide gap MRI system as a function of $Pm$.
+Figure \ref{fig:coefficients} shows the critical parameters $k_c$ and $\reym$ as a function of $Pm$, as well as the coefficients $b$, $c$, and $d$.
+The GLE coefficients are remarkably insensitive to $Pm$.
+From Equation \ref{eq:gle} it is readily apparent that the saturation amplitude is $\alpha_{s} = \pm \sqrt{b/c}$, so we conclude that the saturation amplitude of the MRI is only very weakly dependent on $Pm$.
+This is in stark contrast to the narrow gap behavior of the system.
 

### 2016-08-01 paper/widegap_mri_paper.tex
comparison of boundary layers in wide and narrow gap N31_B terms

@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../figures/wide_narrow_gap_N31_B.pdf}
+\caption{caption}\label{fig:wide_narrow_N31B}
+\end{figure*}
+
 Just as in the weakly nonlinear analyses of \citei{Umurhan:2007hs} and \citei{Clark:2016}, we tune the system away from marginality by taking $B_0 \rightarrow B_0\left(1 - \epsilon^2\right)$, where $\epsilon \ll 1$.
@@
 The GLE coefficients are remarkably insensitive to $Pm$.
-From Equation \ref{eq:gle} it is readily apparent that the saturation amplitude is $\alpha_{s} = \pm \sqrt{b/c}$, so we conclude that the saturation amplitude of the MRI is only very weakly dependent on $Pm$.
+From Equation \ref{eq:gle} it is readily apparent that the asymptotic saturation amplitude is $\alpha_{s} = \pm \sqrt{b/c}$, so we conclude that the saturation amplitude of the MRI is only very weakly dependent on $Pm$.
+Note that because $\reym$ is essentially constant as a function of $Pm$, the saturation amplitude is equivalently insensitive to $\reye^{-1}$.
 This is in stark contrast to the narrow gap behavior of the system.
+For these same boundary conditions, \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}) find that the narrow gap saturation amplitude scales as $Pm^{-4/3}$.
+They find that this amplitude dependence is driven by the $Pm^{1/3}$ dependence of the linear boundary layer.
+Boundary layer analysis similarly reveals a $\nu^{1/3}$ dependence for the radial extent of the boundary layer (\citei{Goodman:2002ix}).
+Why does the $Pm$ scaling of the boundary layer width not translate to a stronger $Pm$ scaling for the saturation amplitude in the wide gap case?
+The boundary layer in the wide gap case is strongly localized at the walls, i.e.
+$r_1$ and $r_2$.
 

### 2016-08-01 paper/widegap_mri_paper.tex
Pm dependence words

@@
 $r_1$ and $r_2$.
+Figure \ref{fig:wide_narrow_N31B} shows the structure of the third-order nonlinear term $N_{31}^{(B)}$ as a function of $Pm$ for both the narrow and wide gap standard MRI.
+$N_{31}$ is the vector that determines the GLE coefficient $c$ (see Appendix \ref{app:basic_equations}).
+Clearly, the boundary layers scale with $Pm$ in both the wide and narrow gap MRI.
+However, in the narrow gap case this scaling extends prominently into the center of the channel, whereas for the wide gap case the bulk of the flow is relatively unaffected by the boundary layer scaling.
 

### 2016-08-01 python/widegap/run_widegap_amplitude.py
run Pm = 1E-5 widegap run

@@ -16,182 +16,187 @@ xi = 0
 R1 = 5
 R2 = 15
 
-for indx in [10]:
+#for indx in [10]:
 
-    crit_params = h5py.File('../../data/widegap_crit_Rm_Q.h5', 'r')
-    Q = crit_params['Q_c'].value[indx]
-    Rm = crit_params['Rm_c'].value[indx]
-    Pm = crit_params['Pm'].value[indx]
+crit_params = h5py.File('../../data/widegap_crit_Rm_Q.h5', 'r')
+#Q = crit_params['Q_c'].value[indx]
+#Rm = crit_params['Rm_c'].value[indx]
+#Pm = crit_params['Pm'].value[indx]
 
-    # Hollerbach & Rudiger
-    """
-    Pm = 1.0E-6
-    Re = 1521
-    Rm = Pm*Re
-    xi = 4
-    Q = 2.33
-    Ha = 16.3
-    beta = (2*Re*Rm)/(Ha**2)
-    R1 = 1.
-    R2 = 2.
-    mu_omega = 0.27
-    Omega1 = 313.55
-    Omega2 = Omega1*mu_omega
-    """
+# Pm=1E-5 critical parameters from Jeff
+Pm = 1.0E-5
+Q = 1.26801e-02
+Rm = 8.40273e-01
 
-    gridnum = 128
+# Hollerbach & Rudiger
+"""
+Pm = 1.0E-6
+Re = 1521
+Rm = Pm*Re
+xi = 4
+Q = 2.33
+Ha = 16.3
+beta = (2*Re*Rm)/(Ha**2)
+R1 = 1.
+R2 = 2.
+mu_omega = 0.27
+Omega1 = 313.55
+Omega2 = Omega1*mu_omega
+"""
 
-    r_basis = de.Chebyshev('r', gridnum, interval=(R1, R2))
-    domain = de.Domain([r_basis], np.complex128, comm=MPI.COMM_SELF)
-    print("running at gridnum", gridnum)
+gridnum = 128
 
-    aa = AmplitudeAlpha(domain, Q = Q, Rm = Rm, Pm = Pm, xi = xi, Omega1 = Omega1, Omega2 = Omega2, beta = beta)
-    aa.print_coeffs()
+r_basis = de.Chebyshev('r', gridnum, interval=(R1, R2))
+domain = de.Domain([r_basis], np.complex128, comm=MPI.COMM_SELF)
+print("running at gridnum", gridnum)
+
[…]

### 2016-08-01 paper/widegap_mri_paper.tex
run Pm = 1E-5 widegap run
wide narrow comparison caption

@@
 \includegraphics[width=\textwidth]{../figures/wide_narrow_gap_N31_B.pdf}
-\caption{caption}\label{fig:wide_narrow_N31B}
+\caption{Nonlinear term $N_{31}^{(B)}$ for the wide gap MRI (left) and narrow gap MRI (right).
+Each line represents a run for a different $Pm$, from $Pm = 1E-4$ (darkest) to $Pm \sim 1E-3$ (lightest).
+Inlaid plot in wide gap case shows a zoomed-in view of the boundary layer at the outer boundary ($r_2$).
+In the narrow gap case the boundary layer strongly affects the bulk of the flow, while in the wide gap case the flow in the center of the channel is relatively unaffected by width of the boundary layers.}\label{fig:wide_narrow_N31B}
 \end{figure*}

### 2016-08-01 paper/widegap_mri_paper.tex
widegap ivp run

@@
 \includegraphics[width=\textwidth]{../figures/wide_narrow_gap_N31_B.pdf}
-\caption{Nonlinear term $N_{31}^{(B)}$ for the wide gap MRI (left) and narrow gap MRI (right).
+\caption{Nonlinear term $N_{31}^{(B)}$ for the wide gap (left) and narrow gap (right) standard MRI.
 Each line represents a run for a different $Pm$, from $Pm = 1E-4$ (darkest) to $Pm \sim 1E-3$ (lightest).
@@
 Figure \ref{fig:wide_narrow_N31B} shows the structure of the third-order nonlinear term $N_{31}^{(B)}$ as a function of $Pm$ for both the narrow and wide gap standard MRI.
-$N_{31}$ is the vector that determines the GLE coefficient $c$ (see Appendix \ref{app:basic_equations}).
+$N_{31}$ is the vector that determines the GLE coefficient $c$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs}, \citei{Clark:2016} for the narrow gap equations).
 Clearly, the boundary layers scale with $Pm$ in both the wide and narrow gap MRI.

### 2016-08-01 python/widegap/allorders2_widegap.py

@@ -739,17 +739,26 @@ class OrderE2(MRI):
         u0field = self.domain.new_field()
         u0field['g'] = self.c1*rfield['g'] + self.c2*(1/rfield['g'])
         
+        ru0field = self.domain.new_field()
+        ru0field['g'] = self.c1*rfield['g']**2 + self.c2
+        
         du0field = self.domain.new_field()
         du0field['g'] = self.c1 - self.c2*(1/rfield['g']**2)
         
+        rrdu0field = self.domain.new_field()
+        rrdu0field['g'] = rfield['g']**2*self.c1 - self.c2
+        
         # multiplied by r^4
         rhs_psi21 = (-self.iR*4*1j*self.Q**3*rfield**3*o1.psi - (2/self.beta)*3*rfield**3*self.Q**2*o1.A + self.iR*4*1j*self.Q*rfield**3*o1.psi_rr
-                    -self.iR*4*1j*self.Q*rfield**2*o1.psi_r + 2*rfield**3*u0field*o1.u + (2/self.beta)*rfield**3*o1.A_rr - (2/self.beta)*rfield**2*o1.A_r
+                    -self.iR*4*1j*self.Q*rfield**2*o1.psi_r + 2*rfield**2*ru0field*o1.u + (2/self.beta)*rfield**3*o1.A_rr - (2/self.beta)*rfield**2*o1.A_r
                     -(2/self.beta)*2*xi*rfield**2*o1.B)
         self.rhs_psi21 = rhs_psi21.evaluate()
         
         # multiplied by r^3
-        rhs_u21 = (2*self.iR*1j*self.Q*rfield**3*o1.u - du0field*rfield**2*o1.psi - rfield*u0field*o1.psi + (2/self.beta)*rfield**3*o1.B)
+        #rhs_u21 = (2*self.iR*1j*self.Q*rfield**3*o1.u - du0field*rfield**2*o1.psi - rfield*u0field*o1.psi + (2/self.beta)*rfield**3*o1.B)
+        #self.rhs_u21 = rhs_u21.evaluate()
+        
+        rhs_u21 = (2*self.iR*1j*self.Q*rfield**3*o1.u - rrdu0field*o1.psi - ru0field*o1.psi + (2/self.beta)*rfield**3*o1.B)
         self.rhs_u21 = rhs_u21.evaluate()
         
         # multiplied by r
@@ -757,7 +766,7 @@ class OrderE2(MRI):
[…]

### 2016-08-02 paper/widegap_mri_paper.tex
found sign error in A21 LHS eqn. does not fix coefficient issue

@@
 where $\mathcal{D}$, $\mathcal{L}$, and $\widetilde{\mathcal{G}}$ are matrices defined in Appendix \ref{app:basic_equations}, and $\mathbf{N}$ is a vector containing all nonlinear terms defined in Appendix \ref{app:nonlinear_terms}.
-We then expand the variables in a perturbation series $\mathbf{V} = \epsilon \mathbf{V}_1 + \epsilon^2 \mathbf{V}_2 + \epsilon^2 \mathbf{V}_3 + h.o.t.$ The perturbed system can then be expressed at each order by the equations
+We then expand the variables in a perturbation series $\mathbf{V} = \epsilon \mathbf{V}_1 + \epsilon^2 \mathbf{V}_2 + \epsilon^3 \mathbf{V}_3 + h.o.t.$ The perturbed system can then be expressed at each order by the equations
 

### 2016-08-02 python/run_IVP.py

@@ -33,6 +33,12 @@ pmcoeffs_fn = '/Users/susanclark/weakly_nonlinear_mri/data/pm_sat_coeffs.h5'
 #coeff = Coefficients(19, pmcoeffs_fn)
 coeff = Coefficients(5, pmcoeffs_fn)
 
+print("hack: setting coefficients manually")
+coeff.a = (1.0000000000000002-7.87628310056604e-29j) 
+coeff.c = (113.5995286989022+7.437107283970664e-11j)
+coeff.b = (0.1522176116459535-2.756166265459084e-13j) 
+coeff.h = (0.09295902574657969+1.466113201026659e-12j)
+
 print("saturation amplitude = {}".format(np.sqrt(coeff.b/coeff.c)))
 
 # Actually solve the IVP
@@ -44,17 +50,24 @@ Z_basis = de.Fourier('Z', gridnum, interval=(-(num_critical_wavelengths/2)*lambd
 Zdomain = de.Domain([Z_basis], grid_dtype=np.complex128)
 problem = de.IVP(Zdomain, variables=['alpha', 'alphaZ'], ncc_cutoff=1e-10)
 
-problem.parameters['ac'] = -coeff.a/coeff.c
+#problem.parameters['ac'] = -coeff.a/coeff.c
+#problem.parameters['bc'] = coeff.b/coeff.c
+#problem.parameters['hc'] = -coeff.h/coeff.c
+
+#print("ac = {}, bc = {}, hc = {}".format(-coeff.a/coeff.c, coeff.b/coeff.c, -coeff.h/coeff.c))
+
+#problem.add_equation("-ac*dt(alpha) + bc*alpha + hc*dZ(alphaZ) = alpha*abs(alpha**2)") 
+problem.parameters['ac'] = coeff.a/coeff.c
 problem.parameters['bc'] = coeff.b/coeff.c
-problem.parameters['hc'] = -coeff.h/coeff.c
+problem.parameters['hc'] = coeff.h/coeff.c
 
-print("ac = {}, bc = {}, hc = {}".format(-coeff.a/coeff.c, coeff.b/coeff.c, -coeff.h/coeff.c))
+print("ac = {}, bc = {}, hc = {}".format(coeff.a/coeff.c, coeff.b/coeff.c, coeff.h/coeff.c))
 
[…]

### 2016-08-02 python/widegap/allorders2_widegap.py

@@ -169,13 +169,13 @@ class MRI():
         #logger.warn("Normalizing according to norm(psi)")
         #norm = np.linalg.norm(psi['g'])
         
-        logger.warn("Normalizing according to integral(psi)")
-        intpsi = psi.integrate('r')
-        norm = intpsi['g'][0]
+        #logger.warn("Normalizing according to integral(psi)")
+        #intpsi = psi.integrate('r')
+        #norm = intpsi['g'][0]
         
-        #logger.warn("Normalizing according to integral(u)")
-        #intu = u.integrate('r')
-        #norm = intu['g'][0]
+        logger.warn("Normalizing according to integral(u)")
+        intu = u.integrate('r')
+        norm = intu['g'][0]
         
         psi['g'] = psi['g']/norm
         u['g'] = u['g']/norm
@@ -546,8 +546,8 @@ class N2(MRI):
         self.N20_psi_r4 = N20_psi_r4.evaluate()
         self.N20_psi_r4.name = "N20_psi_r4"
         
-        print("N20 psi r4 + cc:", self.N20_psi_r4['g'] + self.N20_psi_r4['g'].conj())
-        print("rederived:", self.N20_psi_r4_rederived['g'])
+        #print("N20 psi r4 + cc:", self.N20_psi_r4['g'] + self.N20_psi_r4['g'].conj())
+        #print("rederived:", self.N20_psi_r4_rederived['g'])
         
         N22_u_jacobians = ((1/rfield)*((1j*Q*o1.psi)*o1.u_r - (1j*Q*o1.u)*o1.psi_r) - ((1/rfield)*(2/beta)*((1j*Q*o1.A)*o1.B_r - (1j*Q*o1.B)*o1.A_r)))
         N22_u_advectives = ((1/rfield**2)*o1.u*1j*Q*o1.psi - (2/beta)*(1/rfield**2)*o1.B*1j*Q*o1.A)
@@ -728,8 +728,8 @@ class OrderE2(MRI):
         self.A20 = self.BVP20.state['A']
         self.B20 = self.BVP20.state['B']
         
-        print('B20', self.B20['g'])
-        print('N20_B rhs', self.rhs_B20['g'])
+        #print('B20', self.B20['g'])
+        #print('N20_B rhs', self.rhs_B20['g'])
         
[…]

### 2016-08-02 python/widegap/run_widegap_IVP.py

@@ -70,6 +70,10 @@ coeff = Coefficients(coeff_fn)
 #coeff.c = 0.008129082765040497+6.182564137121617e-06j#0.008129112543701307+6.182564137121872e-06j#0.020810184143773249-1.3877885448830667e-05j
 #coeff.h = 0.021707778758499826+3.9384254266509086e-11j
 
+print("hack: changing sign of b, c coeffs")
+coeff.b = -coeff.b
+coeff.c = -coeff.c
+
 print("ac = {}, bc = {}, hc = {}".format(-coeff.a/coeff.c, -coeff.b/coeff.c, -coeff.h/coeff.c)) 
 print("saturation amplitude = {}".format(np.sqrt(-coeff.b/coeff.c)))
 

### 2016-08-03 paper/widegap_mri_paper.tex
a little more intro for widegap

@@
 
-The magnetorotational instability (MRI) is thought to be a crucial driver of angular momentum transport and turbulence in astrophysical disks. (more MRI intro)
+The magnetorotational instability (MRI) drives angular momentum transport and turbulence in astrophysical disks.
+Its discovery by \citeauthor{Balbus:1991vs} (\citeyear{Balbus:1991vs}; actually a rediscovery of \citei{Chandrasekhar:1960wh}) was a breakthrough in the longstanding question of how efficient accretion can exist in the universe: that is, how matter collapsing onto a central body is able to coalesce despite the conservation of specific angular momentum.
+%The ubiquity of disks ...
+
+Magnetic fields were once assumed to have a stabilizing effect
 
@@
 
+\bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}
+Balbus, S A and Hawley, J F, 1991, The Astrophysical Journal, 376, 214
+
+\bibitem[{Chandrasekhar(1960)}]{Chandrasekhar:1960wh}
+Chandrasekhar, S. 1960, Proceedings of the National Academy of Sciences of the United States of America, 46, 253
+
 \bibitem[{Clark \& Oishi(2016a)}]{Clark:2016}

### 2016-08-04 paper/widegap_mri_paper.tex
more widegap intro

@@
 %The ubiquity of disks ...
+The ubiquity of astrophysical disks suggests that systems both experience and overcome this centrifugal barrier, and accretion proceeds even in hydrodynamically stable disks.
+The MRI is excited by weak magnetic fields in differentially rotating fluids, and since its discovery it has been widely invoked to explain accretion in protoplanetary disks () and disks around black holes (), as well as
 
-Magnetic fields were once assumed to have a stabilizing effect
+%Magnetic fields were once assumed to have a stabilizing effect
 
 %The MRI is fundamental to our understanding of star formation, but its 
-- people make a number of simplifications, incl thin gap, shearing box, etc
+%- people make a number of simplifications, incl thin gap, shearing box, etc
 

### 2016-08-04 paper/widegap_mri_paper.tex
widegap stress figure added to paper

@@
 
-\title{Weakly Nonlinear Analysis of the Standard and Helical Magnetorotational Instabilities in a Taylor Couette Flow}
+%\title{Weakly Nonlinear Analysis of the Standard and Helical Magnetorotational Instabilities in a Taylor Couette Flow}
+\title{The Weakly Nonlinear Magnetorotational Instability} % maybe this is punchier?
 \author{Clark, S.E.\altaffilmark{1}}
@@
 The ubiquity of astrophysical disks suggests that systems both experience and overcome this centrifugal barrier, and accretion proceeds even in hydrodynamically stable disks.
-The MRI is excited by weak magnetic fields in differentially rotating fluids, and since its discovery it has been widely invoked to explain accretion in protoplanetary disks () and disks around black holes (), as well as
+The MRI is excited by weak magnetic fields in differentially rotating fluids, and since its discovery it has been widely invoked to explain accretion in protoplanetary disks (see \citei{Armitage:2010} and references therein) and disks around black holes (), as well as
 
@@
 In particular, much of the analytical work on MRI saturation has focused on local approximations, often representing a section of a disk by solving the MHD equations in an isolated region subject to shear periodic boundary conditions (the ``shearing box").
-While the MRI is a local instability, certain properties of the local problem are not generic to the global problem.
+The shearing box is an inexpensive computational framework, so is used by many numerical MRI studies as well.
+However, while the MRI is a local instability, certain properties of the local problem are not generic to the global problem.
 In the shearing box, linear evolution is dominated by a class of MRI mode known as channel modes.
@@
 
-The theory we develop here employs radial boundary conditions, and so MRI modes are global and channel modes are not present.
+The theory we develop here employs realistic radial boundary conditions, and so MRI modes are global and channel modes are not present.
 A number of saturation mechanisms have been proposed for the MRI which do not rely on channel modes dominating the flow.
@@
 \citei{Ebrahimi:2009ey}).
-The MRI may saturate differently depending on the particular parameter regime under investigation, and so our challenge in not only in identifying possible saturation mechanisms, but in understanding how and when each applies in different astrophysical environments.
+The MRI may saturate differently depending on the particular parameter regime under investigation, and so our challenge is not only in identifying possible saturation mechanisms, but in understanding how and when each applies in different astrophysical environments.
 
@@
 \end{align}
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../figures/widegap_rey_max_tot_stresses.png}
+\caption{Reynolds, Maxwell, and total stress for the fiducial standard MRI case. }
+\end{figure*}
 
@@
 
+\bibitem[{Armitage(2010)}]{Armitage:2010}
+Armitage, P. 2010, ARA\&A, 49, 195
+
 \bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}

### 2016-08-05 paper/widegap_mri_paper.tex

@@
 \begin{abstract}
-Wide gap formulation and preliminary results
+We conduct a formal weakly nonlinear analysis of the magnetorotational instability (MRI) in a Taylor Couette flow.
+This is a multiscale perturbative treatment of the nonideal, axisymmetric MRI near threshold, subject to realistic radial boundary conditions.
+We analyze both the standard MRI, initialized by a constant vertical background magnetic field, and the helical MRI, in which the base magnetic field contains an azimuthal component.
+We find that the evolution of the weakly nonlinear perturbation amplitude of the standard MRI is described by a real Ginzburg-Landau equation (GLE), while the amplitude of the helical MRI takes the form of a complex GLE.
 \end{abstract}
@@
 The ubiquity of astrophysical disks suggests that systems both experience and overcome this centrifugal barrier, and accretion proceeds even in hydrodynamically stable disks.
-The MRI is excited by weak magnetic fields in differentially rotating fluids, and since its discovery it has been widely invoked to explain accretion in protoplanetary disks (see \citei{Armitage:2010} and references therein) and disks around black holes (), as well as
+The MRI is excited by weak magnetic fields in differentially rotating fluids, and since its discovery it has been widely invoked to explain accretion in protoplanetary disks (see \citei{Armitage:2010} and references therein), binary systems (), and disks around black holes (), as well as jet and wind launching (\citei{Lesur:2013}), dynamos, etc (cite these)
 
@@
 
-Despite the challenges inherent in studying the MRI, much progress has been made nu
+%Despite the challenges inherent in studying the MRI, much progress has been made nu
+
+Although the MRI is broadly important to many astrophysical systems, its general properties, and in particular its nonlinear saturation mechanism, remain poorly understood.
+The diversity of astrophysical systems which may permit the MRI admits an enormous parameter space to be explored.
+In protoplanetary disks, for example, the behavior and evolution of the MRI may change drastically depending on the properties of the magnetic field, the disk composition, disk geometry, etc.
+Numerical simulation of realistic disk physics is currently an area of intense focus, and is enabling the study of nonideal MHD effects, disk stratification, nonequilibrium chemistry, and other complex physics that does not lend itself easily to analytic study.
+Still, computational costs inevitably constrain numerical approaches.
+MRI saturation is a complicated nonlinear problem which may depend on the assumptions and approximations adopted by simulations in nonobvious ways.
+Analytic methods can play a powerful role
 
@@
 \includegraphics[width=\textwidth]{../figures/widegap_rey_max_tot_stresses.png}
-\caption{Reynolds, Maxwell, and total stress for the fiducial standard MRI case. }
+\caption{Reynolds ($\mathbb{T}_{R} = u_r u_\phi$), Maxwell ($\mathbb{T}_{M} = -\frac{2}{\beta} B_r B_\phi$), and total stress ($\mathbb{T} = \mathbb{T}_{R} + \mathbb{T}_{M}$) for the fiducial standard MRI case.
+}
 \end{figure*}
@@
 
+\bibitem[{Lesur et al.(2013)}]{Lesur:2013}
+Lesur, G., Ferreira, J., \& Ogilvie, G. I. 2013, A\&A 550, A61
+
 \bibitem[{Liu et al.(2006)}]{Liu:2006}

### 2016-08-06 paper/widegap_mri_paper.tex
more intro

@@
 %The ubiquity of disks ...
-The ubiquity of astrophysical disks suggests that systems both experience and overcome this centrifugal barrier, and accretion proceeds even in hydrodynamically stable disks.
+The ubiquity of astrophysical disks suggests that systems both experience and overcome this centrifugal barrier, and accretion proceeds even in hydrodynamically stable disks, where molecular viscosity is vastly insufficient to drive angular momentum transport (\citei{Shakura:1973wg}).
 The MRI is excited by weak magnetic fields in differentially rotating fluids, and since its discovery it has been widely invoked to explain accretion in protoplanetary disks (see \citei{Armitage:2010} and references therein), binary systems (), and disks around black holes (), as well as jet and wind launching (\citei{Lesur:2013}), dynamos, etc (cite these)
@@
 
-Although the MRI is broadly important to many astrophysical systems, its general properties, and in particular its nonlinear saturation mechanism, remain poorly understood.
+Although the MRI is broadly important to many astrophysical systems, many of its general properties, and in particular its nonlinear saturation mechanism, remain poorly understood.
 The diversity of astrophysical systems which may permit the MRI admits an enormous parameter space to be explored.
@@
 MRI saturation is a complicated nonlinear problem which may depend on the assumptions and approximations adopted by simulations in nonobvious ways.
-Analytic methods can play a powerful role
+Analytic methods can play a powerful role in elucidating the mechanisms responsible for MRI saturation.
+For instance, analytical approaches have revealed the mechanism that likely governs saturation in the ``shearing box" approximation.
+The shearing box is an oft-invoked local approximation in which a section of a disk is represented by solving the MHD equations in an isolated region subject to shear periodic boundary conditions.
+The shearing box is an inexpensive computational framework, and has been widely used to study many aspects of the MRI ().
+However, while the MRI is a local instability, certain properties of the local problem are not generic to the global problem.
+In the shearing box, linear evolution is dominated by a class of MRI mode known as channel modes.
+These are linear modes which also happen to be exact solutions to the \textit{nonlinear} local MRI equations.
+Runaway growth is avoided in this paradigm by the instability of the channel modes themselves, which are destroyed by parasitic (secondary) instabilities (\citei{Goodman:1994ul}, \citei{Pessah:2010ic}).
+The growth of parasitic modes provides a saturation avenue for channel mode-dominated flows, yet this is unlikely to be the dominant saturation mechanism in laboratory experiments or astrophysical disks, as channel modes are artificially over-prominent in the shearing box (e.g.
+\citei{Latter:2015}).
+Thus while the shearing box may accurately approximate many features of the global MRI, the saturation mechanism is not among them.
 
-Because of the MRI's complexity, analytical treatments of the MRI use some combination of convenient approximations, such as ideal MHD, linearized equations, expedient boundary conditions, and simplified geometries.
+%Because of the MRI's complexity, analytical treatments of the MRI use some combination of convenient approximations, such as ideal MHD, linearized equations, expedient boundary conditions, and simplified geometries.
 In particular, much of the analytical work on MRI saturation has focused on local approximations, often representing a section of a disk by solving the MHD equations in an isolated region subject to shear periodic boundary conditions (the ``shearing box").
@@
 \citei{Latter:2015}).
-These are linear modes which also happen to be exact solutions to the \textit{nonlinear} local MRI equations.
+%These are linear modes which also happen to be exact solutions to the \textit{nonlinear} local MRI equations.
 Runaway growth is avoided in this paradigm by the instability of the channel modes themselves, which are destroyed by parasitic (secondary) instabilities (\citei{Goodman:1994ul}, \citei{Pessah:2010ic}).
@@
 
-The theory we develop here employs realistic radial boundary conditions, and so MRI modes are global and channel modes are not present.
+The theory we develop here employs realistic radial boundary conditions, and so MRI modes are global and channel modes (at least in their purest, radially independent are not present.
 A number of saturation mechanisms have been proposed for the MRI which do not rely on channel modes dominating the flow.
@@
 
-We explore the weakly nonlinear HMRI
+In this work we explore the behavior of the viscous, dissipative MRI in a cylindrical geometry close to threshold, making explicit comparisons to the standard MRI behavior in the thin-gap regime.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../figures/widegap_paramspace_crit_params_Pm_1E-3}
+\caption{Growth rates in the ($\reym, k_z$) plane.
+Color map shows growth rate found by solving the linear eigenvalue problem for each ($\reym, k_z$) in the grid.
+The eigenvalue problem was solved for the widegap parameters listed in Table \ref{table:parameters}.
+Overlaid contours show growth rates at [-2E-3, -1E-3, -1E-5, 1E-5, 1E-3, 2E-3], where dashed contours represent negative values.
+The gray dotted line shows the interpolated marginal stability curve.
+The critical parameters $\reym_c = 0.84$ and $k_c = 0.0127$ correspond to the smallest parameter values that yield a zero growth rate.}\label{fig:growth_rates}
+\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure*}
 \centering
-\includegraphics[width=\columnwidth]{../figures/widegap_paramspace_crit_params_Pm_1E-3}
-\caption{Growth rates in the ($\reym, k_z$) plane.
-Color map shows growth rate found by solving the linear eigenvalue problem for each ($\reym, k_z$) in the grid.
-The eigenvalue problem was solved for the widegap parameters listed in Table \ref{table:parameters}.
-Overlaid contours show growth rates at [-2E-3, -1E-3, -1E-5, 1E-5, 1E-3, 2E-3], where dashed contours represent negative values.
-The gray dotted line shows the interpolated marginal stability curve.
-The critical parameters $\reym_c = 0.84$ and $k_c = 0.0127$ correspond to the smallest parameter values that yield a zero growth rate.}\label{fig:growth_rates}
-\end{figure}
+\includegraphics[width=\textwidth]{../python/widegap/figures/orders_1_2_widegap_fiducial_eigenfunctions.pdf}
+\caption{Eigenfunctions of the first order equations, first order adjoint homogenous equations, and second order equations.
+We use our fiducial parameters for the standard MRI ($\xi = 0$).
+First-order eigenfunctions are normalized such that they are either purely real or purely imaginary, and such that $\int \Psi_{11} \mathrm{d}r = 1$.
+Adjoint homogenous eigenfunctions are normalized such that $\left< V_{11}^\dagger \cdot \mathcal{D} V_{11} \right> = 1$.}\label{fig:linear_eigenfunctions}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../python/widegap/figures/nonlinear_2_3_widegap_fiducial_eigenfunctions.pdf}
+\caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.}\label{fig:N2_N3}
+\end{figure*}
 
@@
 \end{widetext}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{../python/widegap/figures/orders_1_2_widegap_fiducial_eigenfunctions.pdf}
-\caption{Eigenfunctions of the first order equations, first order adjoint homogenous equations, and second order equations.
-We use our fiducial parameters for the standard MRI ($\xi = 0$).
-First-order eigenfunctions are normalized such that they are either purely real or purely imaginary, and such that $\int \Psi_{11} \mathrm{d}r = 1$.
-Adjoint homogenous eigenfunctions are normalized such that $\left< V_{11}^\dagger \cdot \mathcal{D} V_{11} \right> = 1$.}\label{fig:linear_eigenfunctions}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{../python/widegap/figures/nonlinear_2_3_widegap_fiducial_eigenfunctions.pdf}
-\caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.}\label{fig:N2_N3}
-\end{figure*}
 
@@
 
+\bibitem[{Shakura \& Sunyaev(1973)}]{Shakura:1973wg}
+Shakura, N I and Sunyaev, R A, Astronomy \& Astrophysics, 1973, 24, 337
+
 \bibitem[{Sisan et al.(2004)}]{Sisan:2004ig}

### 2016-08-08 paper/thingap_mri_paper.tex
rewritten thin gap paper

--- 2016-07-15 paper/mri_paper.tex
@@
 
-\title{The Weakly Nonlinear Saturation of the Magnetorotational Instability}
+\title{The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor Couette flow}
 \author{Clark, S.E.\altaffilmark{1}}
@@
 
-For matter to accrete from a disk onto a central object, angular momentum must be transported radially outward in the disk.
-The transport mechanism is likely turbulent (\citei{Shakura:1973wg}).
-Since its discovery by \citeauthor{Chandrasekhar:1960wh} (\citeyear{Chandrasekhar:1960wh}) and subsequent rediscovery by \citeauthor{Balbus:1991vs} (\citeyear{Balbus:1991vs}), the MRI remains the leading explanation for rapid angular momentum transport in astrophysical disks.
-The instability in its simplest geometry arises when a differentially rotating disk is threaded by a vertical magnetic field.
-The presence of the magnetic field destabilizes the disk gas, driving turbulence and angular momentum transport.
+(general MRI intro)
 
-The MRI evolves via a period of linear growth, followed by nonlinear saturation.
-The nonlinear evolution of the MRI is a complex process, and an understanding of the mechanism by which the MRI saturates remains elusive.
-Several possible saturation avenues are available to the MRI.
-The instability could saturate via a reduction in the background shear (e.g.
-\citei{Umurhan:2007hs}), cutting off the supply of free energy on which the instability feeds.
-Similarly, the instability could pour its free energy into the magnetic field, and saturate via field generation (e.g.
-\citei{Ebrahimi:2009ey}).
-Saturation may also occur via ``parasite" instabilities, which feed off and destroy primary MRI modes (\citei{Goodman:1994ul}).
+Weakly nonlinear analysis is a perturbative method used to examine the asymptotic behavior of a system near threshold -- that is, when the system is just barely unstable to its most unstable mode.
+The analytical technique follows the multiscale evolution of fluid variables in a perturbation expansion, allowing the controlled interaction of modes between orders in a perturbation series (see, e.g., Bender \& Orszag for a full mathematical formalism).
+Weakly nonlinear analysis can be a powerful technique for analytically examining systems which in their full generality exhibit such complicated nonlinear behavior that their study is relegated primarily to the simulation domain.
+The MRI is one such phenomenon: while there is a rich literature analytically examining the linear MRI, analytical treatments of the \textit{nonlinear} system are relatively few.
+The weakly nonlinear treatment of the MRI was pioneered by \citeauthor{Umurhan:2007dz} (\citeyear{Umurhan:2007dz}, \citeyear{Umurhan:2007hs}; hereafter URM07).
+The authors undertook the first weakly nonlinear analysis of the MRI in a thin-gap Taylor Couette flow, and found that the marginal MRI system approaches saturation in a manner analogous to that of Rayleigh-B\'enard convection.
+Weakly nonlinear analysis was instrumental in our understanding of Rayleigh-B\'enard convection saturation (\citei{Newell:1969}), and the similarities between convection and the thin-gap MRI are the result of important shared symmetries between the systems.
+The success of URM07 in modeling the MRI system near threshold merits further consideration, but we are unaware of any other attempts to expand upon their theoretical framework.
+In this work we rigorously rederive the theory of URM07, and expand upon their findings.
+We conduct a set of weakly nonlinear simulations and compare these to the analytical predictions.
+Our focus here is on fully characterizing the narrow gap MRI system, both for independent theoretical interest and to have a robust comparison point for extensions of this theory into more complicated geometries.
+In a companion paper we derive for the first time the weakly nonlinear theory of the standard and helical MRI in a wide gap Taylor Couette flow (\citei{Clark:2016}).
+The thin- and wide-gap treatments complement one another theoretically, and both are important regimes for comparison with simulation.
 
-Most previous works on MRI saturation have studied the problem in simulation.
-Indeed, only a few works have tackled MRI saturation analytically.
-\citeauthor{Knobloch:2005ba} (\citeyear{Knobloch:2005ba}) analyze the MRI in the strongly nonlinear regime, by following the already-developed MRI modes into asymptotic saturation.
-They find that this saturation proceeds via modification of the background shear.
-Another notable analytic work is that of \citeauthor{Pessah:2010ic} (\citeyear{Pessah:2010ic}), who presents an analysis of MRI saturation via the above mentioned secondary instabilities.
-This work suggests that MRI saturation can be caused by parasitic Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime.
+This work examines Taylor Couette flow in the thin-gap regime, and idealization in which the radial extent of the channel is very small compared to its distance from the center of rotation, i.e.
+$(r_2 - r_1) \ll \frac{1}{2} (r_1 + r_2)$.
+The thin-gap approximation eliminates curvature terms, so the domain is Cartesian.
+These curvature terms have an explicit dependence on $r$, so they make the problem more challenging both analytically and numerically.
 
-%examples of simulations papers
-%more mention of other analytic works
-
-This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime.
-The present work adds to this foundation both in subject, with our additional analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space and the comparison of weakly nonlinear analytic work to strongly nonlinear simulation.
-In particular, we attempt to elucidate the role of boundary layers in the saturation of the MRI, as this is of fundamental importance to laboratory MRI work.
-
-We develop this investigation with the goal of providing a rigorous analytical framework that will be used to make predictions for laboratory MRI experiments.
-Observing MRI in a laboratory setting would open a new avenue for studying MRI dynamics and saturation.
-In recent years much experimental progress has been made, although no experiment has yet yielded a confirmed detection.
-Of most immediate interest to this work is the Princeton Plasma Physics Laboratory (PPPL) MRI experiment, which has a design roughly analogous to the set-up investigated here (see \S\ref{sec:setup} for further details).
-The Madison plasma dynamo experiment is a multi-purpose plasma device which may be a vehicle for MRI studies in the near future (\citei{Cooper:2013to}).
-The helical MRI, an instability related to the standard MRI but easier to excite in the laboratory, is the focus of the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE) in Dresden.
-PROMISE has already detected helical MRI in the laboratory (e.g.
-\citei{Stefani:2006iv}), and we discuss possible applications of our framework to the helical MRI in \S\ref{discussion}.
-A spherical Couette experiment claimed to have detected MRI in 2004 (\citeauthor{Sisan:2004ig}), but most likely detected unrelated MHD instabilities instead (\citei{Hollerbach:2009ig}; \citei{Gissinger:2011td}).
-
-The focus of this work is the weakly nonlinear regime of the MRI.
-This is where the initial linear growth phase of the instability begins to give way to nonlinear saturation.
-Weakly nonlinear analysis is the perturbation analysis of a system that is just unstable to its most unstable mode.
-The system in its marginal state, when the most unstable mode neither grows nor decays, is tuned slightly into instability.
-This excites a band of wavemodes around the most unstable mode, which interact nonlinearly.
-In order to track the evolution of the system on both rapid and slow scales, we conduct a formal multiscale analysis.
-We develop and solve this framework in \S\ref{wna}, conduct a full parameter space study in \S\ref{sec:pspace}, and explore the implications of our findings in \S\ref{discussion}.
-
-%relation to hTC and R-B
-%explanation of WNA
-
-\section{Set-Up}
-\label{sec:setup}
+\section{Equations}
 
@@
 \includegraphics[trim=5cm 0cm 0cm 0cm, scale=.35]{setup_diagram.eps}
-\caption{Schematic diagram of our set-up, an axisymmetric magnetized Taylor Couette flow.
+\caption{Schematic diagram of our set-up, an axisymmetric thin-gap Taylor Couette flow.
 We investigate a 2D slice of the X-Z (radial-vertical) plane.
 Our domain is represented by the bolded black box, of width L.
-The radial dimension is solved with a basis of Chebyshev polynomials, and the vertical dimension is solved on a Fourier basis (see \S\ref{discussion}.) }\label{setup}
+The radial dimension is solved with a basis of Chebyshev polynomials, and the vertical dimension is solved on a Fourier basis.
+}\label{fig:setup}
 \end{figure}
 
-We analyze a magnetized Taylor Couette flow, representing a differentially rotating, conducting fluid between two cylinders which are driven at different speeds.
-We initialize an angular velocity profile which decreases outward from the axis of rotation, and a weak vertical magnetic field.
-These are the only two conditions required to excite MRI.
-The differential rotation profile is
+%We solve the equations of non-ideal, axisymmetric, incompressible magnetized Taylor Couette flow.
 
-\beq
-\Omega\left(r\right) \propto \Omega_0\left(\frac{r}{r_0}\right)^{-q},
-\eeq
+The evolution of a conducting fluid is governed by the momentum and induction equations,
 
-where $q$, the shear parameter, is $3/2$ for Keplerian rotation. The initial background magnetic field is
-
-\beq
-\label{backgroundfield1}
-\mathbf{B} = B_0 \hat{z},
-\eeq
-
-a constant, vertical field.
-
-We analyze this system in the thin-gap regime, where the channel is radially narrow. Formally, this means
-
-\beq
-\left(R_2 - R_1\right) \ll \frac{1}{2}\left(R_2 + R_1\right),
-\eeq
-
-where $R_1$ is the radius of the inner cylinder and $R_2$ is the radius of the outer cylinder, such that the width of the channel is small compared to its distance from the center of rotation.
-The imposition of the thin-gap approximation eliminates channel modes, the axisymmetric MRI modes that are exact solutions of the incompressible MRI in the shearing box approximation (\citei{Goodman:1994ul}).
-Implications of this approximation are discussed further in \S\ref{discussion}.
-
-Our domain is designed to be relevant to the apparatus of the PPPL MRI experiment (see \S\ref{discussion} for further discussion).
-We use periodic vertical boundary conditions.
-On the radial boundaries of our domain --  the inner and outer cylinders -- we apply no-slip, perfectly conducting boundary conditions.
-The choice of boundary conditions dictates the functional bases on which our system is spectrally solved.
-This is explained further in \S\ref{wna}.
-A diagram of our set-up is shown in Figure \ref{setup}.
-
-\section{Weakly Nonlinear Analysis}\label{wna}
-
-We perform a weakly nonlinear analysis following the lead of \citeauthor{Umurhan:2007hs}, and use much of the same notation, with some changes for added clarity.
-
-We solve the equations of non-ideal, axisymmetric, incompressible magnetized Taylor Couette flow. We solve the momentum equation,
-
-\beq\label{momentum}
+\beq\label{eq:momentum}
 \begin{split}
@@
 
-and the induction equation,
-
-\beq\label{induction}
-\partial_t \mathbf{B} = \nabla \times \left(\mathbf{u} \times \mathbf{B}\right) + \eta\nabla^2\mathbf{B}.
+\beq\label{eq:induction}
+\partial_t \mathbf{B} = \nabla \times \left(\mathbf{u} \times \mathbf{B}\right) + \eta\nabla^2\mathbf{B},
 \eeq
 
-In the above, $P$ is the gas pressure, $\nu$ is the kinematic viscosity, $\eta$ is the microscopic diffusivity, $\nabla\Phi$ is the gravitational force per unit mass, and the current density is
-
-\beq
-%\mathbf{J} = \frac{2}{\beta}\nabla\times\mathbf{B}.
-\mathbf{J} = \nabla\times\mathbf{B},.
-\eeq
-
-in units where the permittivity of free space is unity.
-
+where $P$ is the gas pressure, $\nu$ is the kinematic viscosity, $\eta$ is the microscopic diffusivity, $\nabla\Phi$ is the gravitational force per unit mass, and the current density is $\mathbf{J} = \nabla\times\mathbf{B}$.
+%in units where the permittivity of free space is unity. 
 Equations \ref{momentum} and \ref{induction} are subject to the magnetic solenoid and incompressibility constraints,
 
-\beq
+\beq\label{eq:solenoid_constraint}
 \nabla \cdot \mathbf{B} = 0
@@
 
-and
-
-\beq
+\beq\label{eq:incompressibility_constraint}
 \nabla \cdot \mathbf{u} = 0.
@@
 We axisymmetrically perturb all three vector components of each of the fluid quantities.
-We nondimensionalize the equations, with lengths nondimensionalized by $L$, velocities by $\Omega_0 L$, magnetic fields by $B_0$, and pressure by $\Omega_0^2 L^2 \rho_0$, where $L$ is the channel width, $\Omega_0$ is the rotation rate at the center of the channel, and $\rho_0$ is the constant pressure in the base state (see Figure \ref{setup}).
-We define the nondimensional Reynolds number,
+We nondimensionalize the equations, with lengths nondimensionalized by $L$, velocities by $\Omega_0 L$, magnetic fields by $B_0$, and pressure by $\Omega_0^2 L^2 \rho_0$, where $L$ is the channel width, $\Omega_0$ is the rotation rate at the center of the channel, and $\rho_0$ is the constant pressure in the base state (see Figure \ref{fig:setup}).
+We define the Reynolds number, $\reye \equiv {\Omega_0 L^2}/{\nu}$, magnetic Reynolds number, $\reym \equiv {\Omega_0 L^2}/{\eta}$, and a plasma beta parameter, $\beta \equiv {\Omega_0^2r_0^2}/{v_A^2}$, where the Alfv\'en sped $v_A$ is $v_A^2 = {B_0^2}/{\rho_0}$.
+The fluid symbols $\mathbf{u}$, $\mathbf{B}$, etc.
+will henceforth be used to refer to the nondimensional, perturbed quantities.
 
-\beq
-\reye \equiv \frac{\Omega_0 L^2}{\nu},
-\eeq
-
-and magnetic Reynolds number,
-
-\beq
-\reym \equiv \frac{\Omega_0 L^2}{\eta}.
-\eeq
-
-We also define a plasma beta parameter,
-
-\beq
-\beta \equiv \frac{\Omega_0^2r_0^2}{v_A^2},
-\eeq
-
-where the Alfv\'en sped $v_A$ is
-
-\beq
-v_A^2 = \frac{B_0^2}{\rho_0}.
-\eeq
-
-The fluid symbols $\mathbf{u}$, $\mathbf{B}$, etc. will henceforth be used to refer to the nondimensional, perturbed quantities.
-
-We define the flux function $A$ and streamfunction $\Psi$.
-$A$ is the familiar two-dimensional vector potential.
+We define the flux function $A$ and streamfunction $\Psi$, where $A$ is the familiar two-dimensional vector potential.
 $A$ and $\Psi$ are scalar fields.
-The curl of $A$ and the curl of $\Psi$ are defined as the magnetic field and velocity, respectively, and so $A$ and $\Psi$ automatically satisfy our constraints.
+The curl of $A$ and the curl of $\Psi$ are defined as the magnetic field and velocity, respectively, and so $A$ and $\Psi$ automatically satisfy our constraints (Equations \ref{eq:solenoid_constraint} and \ref{eq:incompressibility_constraint}).
 
@@
 \beq
-\mathcal{D}\partial_t \mathbf{V} + \mathbf{N} = \mathcal{L} \mathbf{V} - \epsilon^2\partial_z\mathcal{X} \mathbf{V} - \epsilon^2 \partial_z^3 \mathcal{L}_3\mathbf{V},
+ \mathcal{D}\partial_t \mathbf{V} +  \mathcal{L} \mathbf{V} + \epsilon^2\widetilde{\mathcal{G}} = \mathbf{N},
 \eeq
 
-where $\mathcal{L} \equiv \mathcal{L}_0 + \mathcal{L}_1\partial_z + \mathcal{L}_2\partial_z^2 + \mathcal{L}_3\partial_z^3 + \mathcal{L}_4\partial_z^4$, \\
-
-and
+where we leave the definition of the matrices $\mathcal{D}$, $\mathcal{L}$, and $\widetilde{\mathcal{G}}$ to Appendix \ref{app:matrices}, and the detailed form of the nonlinear vector $\mathbf{N}$ to Appendix \ref{app:nonlinear_terms}.
+We solve this system subject to no-slip, perfectly conducting radial boundary conditions, defined as
 
 \beq
-\mathcal{D} = \left[\begin{matrix}
-\nabla^2 & 0 & 0 & 0 \\
-0 & 1& 0 & 0 \\
-0 & 0 & 1 & 0\\
-0 & 0 & 0 & 1 \\
-\end{matrix}\right]
+\Psi = \partial_x \Psi = u_y = A = \partial_x B_y = 0.
 \eeq
 
-\beq
-\mathcal{L}_0 = \left[\begin{matrix}
-\frac{1}{\reye}\partial_x^4 & 0 & 0 & 0 \\
-0 & \frac{1}{\reye}\partial_x^2 & 0 &0 \\
-0 & 0 & \frac{1}{\reym}\partial_x^2 & 0 \\
-0 & 0 & 0 & \frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right]
-\eeq
+\section{Weakly Nonlinear Analysis}
 
-\beq
-\mathcal{L}_1 = \left[\begin{matrix}
-0 & 2 & \frac{2}{\beta}\partial_x^2 & 0 \\
-(\Omega_0 - 2)q & 0 & 0 & \frac{2}{\beta} \\
-1 & 0 & 0 & 0 \\
-0 & 1 & -q\Omega_0 & 0 \\ \end{matrix}\right] 
-\eeq
-
-\beq
-\mathcal{L}_2 = \left[\begin{matrix}
-2\frac{1}{\reye} \partial_x^2 & 0 & 0 & 0 \\
-0 & \frac{1}{\reye} & 0 & 0 \\
-0 & 0 & \frac{1}{\reym} & 0 \\
-0 & 0 & 0 & \frac{1}{\reym} \\ \end{matrix}\right]
-\eeq
-
-\beq
-\mathcal{L}_3 = \left[\begin{matrix}
-0 & 0 & \frac{2}{\beta} & 0 \\
-0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\ \end{matrix} \right]
-\eeq
-
-\beq
-\mathcal{L}_4 = \left[\begin{matrix}
-\frac{1}{\reye} & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\ \end{matrix}\right] 
-\eeq
-
-\beq
-\mathcal{X} = \left[\begin{matrix}
-0 & 0 & \frac{2}{\beta}\partial_x^2 & 0 \\
-0 & 0 & 0 & \frac{2}{\beta} \\
-1 & 0 & 0 & 0 \\
-0 & 1 & 0 & 0 \\ \end{matrix} \right]. 
-\eeq
-
-The nonlinear terms are contained in a vector $\mathbf{N} = \left[N^{(\Psi)}, N^{(u_y)}, N^{(A)}, N^{(B_y)}\right]^\mathrm{T}$, where
-
-\beq
-\mathbf{N} = \left[ \begin{matrix}  J(\Psi, \nabla^2 \Psi) - \frac{2}{\beta}J(A, \nabla^2 A)\\
-J(\Psi, u_{y}) - \frac{2}{\beta} J(A, B_{y}) \\ 
--J(A, \Psi) \\
-J(\Psi, B_{y}) - J(A, u_{y}) \\
-\end{matrix} \right].
-\eeq
-
-We conduct a formal multiscale analysis.
-Our perturbations are characterized in terms of fast and slow-moving variables, in order to simultaneously track their evolution on fast and slow scales.
+We conduct a formal multiple scales analysis of this system.
+Our perturbations are characterized in terms of fast and slow-moving variables, in order to simultaneously track their evolution on two scales.
 The relative scalings of the fast and slow scales are determined as follows.
@@
 The scalings are
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[trim=6cm 3cm 3cm 0cm, scale=0.35]{slowfast.eps}
-%\caption{Cartoon illustrating fast and slow-moving variables (Equation \ref{scalings}).}\label{allorders_velocity}
-%\label{figscalings}
-%\end{figure}
-
-\begin{figure*}
-\centering
-\includegraphics[trim=0cm 10cm 0cm 5cm, scale=.5]{N20fig.eps}
-\caption{Example of a nonlinear term solved on a 256-point Chebyshev grid. Circles represent values on the Chebyshev grid, grey lines are fits.}\label{N20fig}
-\end{figure*}
 
@@
 Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic Taylor Couette flow.
+Our $x$ dimension, the direction of angular momentum transport, is analogous to the direction of temperature transport in the convection problem.
 In analogy to these problems, we posit slow variation in both $Z$ and $T$.
@@
 The periodic vertical boundary conditions allow us to posit the $z$ dependence, where $k_c$ is the value of the vertical wavenumber at marginality.
-As noted by \citeauthor{Umurhan:2007hs}, there exists a spatially constant neutral mode solution to the $B_y$ equation, with $\mathbb{U}_{11} = (0, 0, 0, 1)^\mathbf{T}$.
-The slowly-varying amplitude $\beta$ encodes the slow evolution of this mode.
+As noted by URM07, there exists a spatially constant neutral mode solution to the $B_y$ equation, with $\mathbb{U}_{11} = (0, 0, 0, 1)^\mathbf{T}$.
+The amplitude $\beta(T, Z)$ encodes the slow evolution of this mode.
 
-Because we investigate the system at marginality, where the growth rate $\sigma = 0$, the $\partial_t$ terms drop out. Thus after the multiscale expansion our system becomes
-
-\begin{multline}
-\label{multiscale_system}
-\epsilon^2 \mathcal{D}\partial_T\mathbf{V} \, + \, \mathbf{N} \, = \, \mathcal{L}\mathbf{V} \, + \, \epsilon\widetilde{\mathcal{L}}_1\partial_Z\mathbf{V} \, + \, \epsilon^2\widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V} \, \\
- - \, \epsilon^2\partial_z\mathcal{X}\mathbf{V} \, - \, \epsilon^2\partial_z^3\mathcal{L}_3\mathbf{V} \, + \, \mathcal{O}(\epsilon^3),
-\end{multline}
-
-where
+The state vector is expanded in a perturbation series in orders of $\epsilon$,
 
 \beq
-\widetilde{\mathcal{L}}_1 = \mathcal{L}_1 + 2\mathcal{L}_2\partial_z + 3\mathcal{L}_3\partial_z^2 + 4\mathcal{L}_4\partial_z^3
+\label{eq:pert_exp}
+\mathbf{V} = \epsilon\mathbf{V_1} + \epsilon^2\mathbf{V_2} + \epsilon^3\mathbf{V_3} + h.o.t.
 \eeq
 
-and
+And our perturbed system is then expressed order by order as
 
-\beq
-\widetilde{\mathcal{L}}_2 = \mathcal{L}_2 + 3\mathcal{L}_3\partial_z + 6\mathcal{L}_4\partial_z^2.
-\eeq
+\begin{align}
+\mathcal{O}(\epsilon)&: \mathcal{L}\mathbf{V_1} + \mathcal{D}\partial_t \mathbf{V_1} = 0. \\ \label{eq:ordere}
+\mathcal{O}(\epsilon^2)&: \mathcal{L}\mathbf{V_2} + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1} = \mathbf{N_2}\\
+\mathcal{O}(\epsilon^3)&: \mathcal{L} \mathbf{V_3} + \mathcal{D}\partial_T \mathbf{V_1}  + \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} + \widetilde{\mathcal{G}}\mathbf{V_1} = \mathbf{N_3} \, \\ \label{eq:ordere3}
+\end{align}
+
+The partial differential equations that comprise Equations \ref{eq:ordere} to \ref{eq:ordere3} are solved in succession.
+The practical advantage of our ansatz construction (Equation \ref{V1_ansatz}) is clear: the separable x-dependence means that the radial boundary conditions are solved in only one dimension.
+Thus our analytical framework is able to side-step many of the resolution issues faced by multidimensional simulations.
+We are able to resolve even small-scale structure in the boundary layers of our domain, because we need only resolve it in one dimension.
+We solve the radial component of each equation using the open source pseudospectral code Dedalus (Burns et al., in prep).
+We compute the radial components on a grid of Chebyshev polynomials, as is appropriate for bounded one-dimensional domains (e.g.
+\citei{Boyd:2001aa}).
+The nonuniform spacing of the Chebyshev grid allows us to resolve the boundary layers well on a 128-point grid.
+
+\section{Simulations}
+
+We compare the weakly nonlinear theory with simulations of the same system.
+We use the pseudospectral code Dedalus to evolve Equations \ref{eq:momentum} and \ref{eq:induction} in time.
+To analyze the weakly nonlinear regime in simulation, we raise $\reym$ slightly to tune the system just out of marginality.
+We initialize a domain of vertical extent $20 \lambda_c$, where $\lambda_c = 2 \pi / k_c$ is the critical vertical wavelength.
+The radial domain is identical to the analytical domain, and the equations are solved on a hybrid Fourier-Chebyshev domain (Figure \ref{fig:setup}).
 
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_velocity.png}%{allorders_velocity.eps}
+\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_velocity_Pm_1E-3.png}%{allorders_velocity.eps}
 \caption{First order (left), second order (center), and total (right) velocity perturbations.
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_Bfield.png}
+\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_Bfield_Pm_1E-3.png}
 \caption{First order (left), second order (center), and total (right) magnetic field perturbations.
@@
 \end{figure*}
-
-% perturbation expansion
-The state vector is expanded in a perturbation series in orders of $\epsilon$,
-
-\beq
-\label{pert_exp}
-\mathbf{V} = \epsilon\mathbf{V_1} + \epsilon^2\mathbf{V_2} + \epsilon^3\mathbf{V_3} + ...
-\eeq
-
-The multiscale expansion of the operators and the perturbation series expansion of the fluid quantities combine to make the expansion of the nonlinear vector $\mathbf{N}$ formidable, and we relegate the expansion to Appendix \ref{app:nonlinear}.
-$\mathbf{N}$ becomes
-
-\beq
-\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4).
-\eeq
-
-Substituting Equation \ref{pert_exp} into Equation \ref{multiscale_system} yields the perturbed, expanded equations. We solve these successively in increasing orders of $\epsilon$.
-
-To $\mathcal{O}(\epsilon)$, we have
-
-\beq
-\label{ordere}
-\mathcal{L}\mathbf{V_1} = 0.
-\eeq
-
-To $\mathcal{O}(\epsilon^2)$,
-
-\beq
-\label{ordere2}
-\mathcal{L}\mathbf{V_2} \, = \, \mathbf{N_2} \, - \, \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}.
-\eeq
-
-To $\mathcal{O}(\epsilon^3)$,
-
-\begin{multline}
-\label{ordere3}
-\mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, \\
-- \partial_z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1}.
-\end{multline}
-
-In the above, we find that $\mathbf{V_2}$ takes the form
-
-\beq
-\begin{split}
-\mathbf{V_2} \, & = \, \alpha^2 e^{i2 k_c z} \mathbb{V}_{22}(x) \, + \, \left|\alpha\right|^2 \mathbb{V}_{20}(x) \, \\
-&- \, \partial_Z \alpha \mathbb{V}_{21}(x) e^{i k_c z} \,- \, \partial_Z \beta \mathbb{U}_{20}(x) + c.c.
-\end{split}
-\eeq
-
-The $\mathcal{O}(\epsilon)$ equation is the equation for the linear MRI at marginality. This is the equation that is solved to find the $\mathbb{V}_{11}$ component of Equation \ref{V1_ansatz}.
-
-The partial differential equations that comprise Equations \ref{ordere} to \ref{ordere3} are solved in succession subject to no-slip, perfectly conducting radial boundary conditions, defined as
-
-\beq
-\Psi = \partial_x \Psi = u_y = A = \partial_x B_y = 0.
-\eeq
-
-The practical advantage of our ansatz construction (Equation \ref{V1_ansatz}) is clear: the separable x-dependence means that the radial boundary conditions are solved in only one dimension.
-Thus our analytical framework is able to side-step many of the resolution issues faced by multidimensional simulations.
-We are able to resolve even small-scale structure in the boundary layers of our domain, because we need only resolve it in one dimension.
-We solve the radial component of each equation using the open source pseudospectral code Dedalus (Burns et al., in prep).
-We compute the radial components on a grid of Chebyshev polynomials, as is appropriate for bounded one-dimensional domains (e.g.
-\citei{Boyd:2001aa}).
-The nonuniform spacing of the Chebyshev grid allows us to resolve the boundary structures well on a 256-point grid.
-See Figure \ref{N20fig} for an example of the boundary-layer resolution we achieve in a nonlinear term.
-
-We solve the system up to $\mathcal{O}(\epsilon^3)$ in order to close the solution at $\mathcal{O}(\epsilon^2)$.
-
-To find a bounded solution at each order, we eliminate secular terms: terms which are resonant with solutions to the homogenous equation and cause the solution to grow without bound.
-The elimination of secular terms requires enforcement of solvability criteria, which arise as a consequence of the Fredholm alternative (e.g.
-Guenther \& Lee 1988).
-The solvability criterion is found by forcing the inner product of the equation with its adjoint homogenous solution to be zero.
-For instance, Equation \ref{ordere2} contains a term $\widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ that is resonant with $e^{i k_c z}$.
-Because the adjoint homogenous equation
-
-\beq
-\mathcal{L}^\dagger \mathbf{V}^\dagger = 0
-\eeq
-
-has a nontrivial solution, the solvability criterion
-
-\beq
-\left< \mathbf{V}^\dagger \cdot \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}\right> = 0 
-\eeq
-
-must be satisfied in order for Equation \ref{ordere2} to have a solution. In the above,
-
-\beq
-\mathbf{V}^\dagger = \mathbb{V}^\dagger(x) e^{i k_c z} + c.c.,
-\eeq
-
-where, again, $\mathbb{V}^\dagger$ contains the radial contribution.
-The elimination of this inner product must also be enforced on the terms in Equation \ref{ordere3} that are resonant with $e^{i k_c z}$, which are
-
-\beq
-\begin{split}
-& \mathcal{D} \mathbb{V}_{11} e^{i k_c z}\partial_T \alpha \, + \, \alpha \left| \alpha^2 \right| \mathbf{N}_{31} e^{i k_c z} \, + \, \alpha \partial_Z \beta \widetilde{\mathbf{N}}_{31} e^{i k_c z} \, =  \\
-& \, - \partial_Z^2 \alpha \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} e^{i k_c z} \, + \, \partial_Z^2 \alpha \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{i k_c z} \, - \, \partial_Z \alpha \mathbb{V}_{11}\mathcal{X}e^{i k_c z} \, \\
-& + \, i k_c^3 \alpha \mathcal{L}_3 \mathbb{V}_{11} e^{i k_c z}.
-\end{split}
-\eeq
-
-Taking the inner product yields the amplitude equation
-
-\beq
-\label{ampeq1}
-a \partial_T \alpha \, = - \, c \alpha \left|\alpha^2\right| \, - \, \widetilde{c} \alpha \partial_Z \beta \,  - b i k_c \alpha \, + \, h \partial_Z^2 \alpha \, + \, g i k_c^3 \alpha,
-\eeq
-
-where
-
-\beq
-\begin{split}
-a & \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{D}\mathbb{V}^*_{11}\right> \\
-c & \equiv \left< \mathbb{V}^\dagger \cdot \mathbf{N}^*_{31}\right> \\
-\widetilde{c} & \equiv \left< \mathbb{V}^\dagger \cdot \widetilde{\mathbf{N}}^*_{31}\right> \\
-b & \equiv \left< \mathbb{V}^\dagger \cdot \left(\mathcal{X}\mathbb{V}_{11}\right)^* \right> \\
-h & \equiv \left< \mathbb{V}^\dagger \cdot \left(\widetilde{\mathcal{L}}_2\mathbb{V}_{11} \, - \, \widetilde{\mathcal{L}}_1\mathbb{V}_{21} \right)^* \right> \\
-g & \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{L}_{3} \mathbb{V}_{11} \right>, \\
-\end{split}
-\eeq
-
-where ${}^*$ denotes a complex conjugate, but $\widetilde{c}$ is equal to zero and drops out of the equation.
-This is the well-known Ginzburg-Landau equation, and is with only slight modification the same amplitude equation derived in \citeauthor{Umurhan:2007hs}
-
-There exists a second amplitude equation, corresponding to the neutral mode, which comes from the solvability condition that must be imposed on all terms in Equation \ref{ordere3} that have no $z$ dependence.
-These terms are
-
-\beq
-\begin{split}
-\mathcal{D}\mathbb{U}_{11} \partial_T \beta \, & = \, \mathcal{L}\mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1 \mathbb{V}_{20} 2 \left|\alpha\right| \partial_Z \alpha \, - \, \widetilde{\mathcal{L}}_1 \mathbb{U}_{20} \partial_Z^2 \beta \, \\
-& + \, \widetilde{\mathcal{L}}_2 \mathbb{U}_{11} \partial_Z^2 \beta \, - \, \mathcal{X} \mathbb{U}_{11} \partial_Z \beta,
-\end{split}
-\eeq
-
-and taking the inner product of these terms with the adjoint homogenous solution $\mathbb{U}_{11}$ yields
-
-\beq
-\partial_T \beta = k 2 \left|\alpha\right| \partial_Z \alpha - m \partial_Z^2 \beta + \frac{1}{\reym} \partial_Z^2 \beta - p \alpha^* \partial_Z \alpha,
-\eeq
-
-where
-
-\beq
-\begin{split}
-n & \equiv \left< \mathbb{U}_{11} \cdot \left(\mathcal{L} \mathbb{V}_{30}\right)^*\right> \\
-k & \equiv \left< \mathbb{U}_{11} \cdot \left(\widetilde{\mathcal{L}}_1 \mathbb{V}_{20}\right)^*\right> \\
-m & \equiv \left< \mathbb{U}_{11} \cdot \left(\widetilde{\mathcal{L}}_1 \mathbb{U}_{20}\right)^*\right> \\
-p & \equiv \left< \mathbb{U}_{11} \cdot N_{30}^*\right>, \\
-\end{split}
-\eeq
-
-but both $k$ and $p$ are equal to zero. Thus this amplitude equation decouples from Equation \ref{ampeq1}, and reveals itself to be a simple diffusion equation,
-
-\beq\label{ampeq2}
-\partial_T \beta = - m \partial_Z^2 \beta + \frac{1}{\reym} \partial_Z^2 \beta.
-\eeq
-
-Equation \ref{ampeq1} is an amplitude equation -- a function of the slow variables $Z$ and $T$.
-We solve this initial value problem using Dedalus to obtain the asymptotic behavior of $\alpha$.
-This is the saturation amplitude.
-We use the saturation amplitude as $\alpha$ and the radial component of computed on the Chebyshev grid to plot $\mathbf{V}$.
-We show both first order ($\epsilon\mathbf{V}_1$), second order ($\epsilon^2\mathbf{V}_2$), and combined first and second order fluid perturbations ($\epsilon\mathbf{V}_1$ + ($\epsilon^2\mathbf{V}_2$).
-Figure \ref{allorders_velocity} shows the velocity perturbations, and Figure \ref{allorders_Bfield} shows the magnetic field perturbations.
-
-Our analytic formulation is advantageous because we can now analyze the contributions of each fluid quantity at each order.
-This allows us to directly identify the source of saturation.
-Equation \ref{ampeq1} asymptotically approaches a value, so the system is clearly saturating, and our approach now allows us to quantify the individual contribution of each term.
-In Figure \ref{allorders_velocity}, we see significant fluid action in the second-order velocity plot.
-By contrast, the second-order magnetic field perturbations are two orders of magnitude smaller than the first-order perturbations (Figure \ref{allorders_Bfield}).
-
-The role of the boundary layers changes with Pm.
-As Pm decreases, the width of the boundary layer becomes increasingly narrow (Figure \ref{u_secondorder_by_Pm}).
-The geometry of the velocity in the $x-z$ plane begins as a series of MRI ``cells" whose structure is radially narrow in the bulk of the flow.
-As Pm is decreased, the saturation geometry of the central cells stretches radially, until at Pm $\sim 10^{-5}$ the bulk cells dominate the flow.
-The relative importance of the $u_y$ second order perturbation to the first order perturbation increases as well (CHECK, PLOT).
-
-\begin{figure*}[h]
-\centering
-\includegraphics[trim=2cm 0cm 6cm 0cm, scale=0.4]{u_secondorder_by_Pm_2.eps}
-\caption{(Placeholder) The evolution of the second-order velocity perturbations by Pm.}\label{allorders_velocity}
-\label{u_secondorder_by_Pm}
-\end{figure*}
-
-%It appears that the second-order velocity perturbations are instrumental in the saturation of the MRI. The likely mechanism is a reduction in the background shear.
-
-We plot the total stress, or the sum of the Reynolds and Maxwell stresses,
-
-\beq
-\mathbb{T}(x, z) = u_x u_y - \frac{2}{\beta} B_x B_y.
-\eeq
-
-See Figure \ref{stress_uy_cuts}.
-
-From the behavior of the second order velocity perturbations (Figure \ref{stress_uy_cuts}) in the bulk of the flow (i.e.
-apart from the boundary layers), we gather that the MRI is saturating via a background shear reduction.
-Recall that the linear shear profile means that the background shear is positive for negative values of $x$ and negative for positive values of $x$.
-Thus Figure \ref{stress_uy_cuts} shows the azimuthal velocity perturbations resisting the background shear in the bulk of the flow.
-
-\begin{figure*}
-\centering
-\includegraphics[trim=6cm 4cm 6cm 3cm, scale=0.5]{stress_uy_cuts.pdf}
-\caption{(Placeholder) The total stress (left) and vertical cuts of the azimuthal velocity perturbations (right).
-The $u_y$ perturbations shown are first order (purple), second order (orange), and total (black).
-The blue shaded area represents the region where the second order and total perturbations are opposing the background shear because they are negative, and the red shaded region indicates where the second order and total perturbations are opposing the background shear because they are positive.}
-\label{stress_uy_cuts}
-\end{figure*}
-
-\section{Parameter Space}\label{sec:pspace}
-
-We conduct a full study of the MRI parameter space, focused on the behavior with changing magnetic Prandtl number ($\mathrm{Pm} \equiv \frac{\mathrm{Rm}}{\mathrm{Re}} \equiv \frac{\nu}{\eta}$).
-In protoplanetary disks, as well as in liquid metal MRI experiments, the magnetic Prandtl number can be very small ($\mathrm{Pm} \sim 10^{-6}$).
-The scale separations caused by these large ratios of diffusivity to viscosity render the small Prandtl number regime inaccessible to simulations.
-Because we solve spectrally in only one dimension, rather than in two or three, our analytic formulation allows us to explore the behavior down to magnetic Prandtl numbers of $10^{-6}$ or lower.
-
-\section{Simulations}
-
-We compute the solutions to our derived equations pseudospectrally, using Dedalus.
-One advantage to using Dedalus for these computations is that the code can compute the evolution of the fully nonlinear MRI system on the same Fourier/Chebyshev grid.
-This allows us to directly compare simulation and theory.
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{../figures/final_slice_MRI_run_Rm5_00e+01_eps0_00e+00_Pm1_00e-03_beta2_50e+01_Q7_50e-01_qsh1_50e+00_Omega1_00e+00_nx256_nz256_Lz10_CFL.pdf}
-\caption{The final timestamp of each simulation.}\label{fig:final_slices}
-\end{figure*}
-
-\section{Discussion}\label{discussion}
-
-Our domain is designed to be relevant to the design of the Princeton Plasma Physics Laboratory (PPPL) MRI experiment (\citei{Ji:2001kd}).
-A challenge for all MRI experiments is to reduce all extraneous flows which may stabilize the MRI or obscure its detection.
-In particular, care must be taken to mitigate meridional flows introduced by the endcaps of the experimental apparatus.
-The PPPL MRI experiment employs a novel design using concentric, differentially rotating rings as endcaps (\citei{Schartman:2009df}).
-In this investigation, we use periodic vertical boundary conditions, so our set-up represents an idealized domain where endcap contamination does not exist.
-Future work will address specifically our predictions for the PPPL MRI experiment.
-
-This work analyzes Taylor Couette flow in the thin-gap limit, where the channel width is radially narrow compared to its distance from the center of rotation.
-This simplifies the math by allowing curvature terms to drop out at lowest order.
-The relaxation of the thin-gap approximation is an obvious next step for this work.
-The term expansion in the weakly nonlinear analysis will be more formidable in the wide-gap regime, but could prove informative for comparison to experiments.
-
-%The present configuration, that of magnetized Taylor Couette flow with a purely axial magnetic field, represents the initial conditions for the standard MRI.
-The helical MRI is a related, but separate, MHD instability, and was first discovered by Hollerbach and R\"udiger (\citeyear{Hollerbach:2005tr}).
-It arises in the presence of combined axial and azimuthal fields.
-To seed helical MRI in our current set-up, we need only add an azimuthal component to our initial background magnetic field.
-This will be an avenue of future study, and will yield an analytic theory relevant to the above-mentioned PROMISE experiment.
-
-Our results suggest that the primary driver of saturation is reduction of the background shear.
-We note that while shear modification is relevant in the experimental regime, the gravitational pull of the central body will tend to enforce a Keplerian rotation profile in any astrophysical disk.
-Thus a modification of the rotation profile is of little relevance as a saturation mechanism to the astrophysical regime.
-Nevertheless, understanding laboratory MRI experiments is crucial for allowing experimental results to inform our astrophysical understanding of the MRI.
-%^this sentence sucks
-
-% curvature terms
-% hMRI
-% reduction in background shear not possible in astrophysical disks
-
-\clearpage
-\appendix
-
-\section{A. Linear dispersion relation}\label{app:dispersion}
-%Note: I don't expect this to go in the actual paper, but it could be useful in the meeting.
-
-The linear dispersion relation, which determines the variable scalings in the multiscale analysis. \\
-
-\beq
-\begin{split}
-& \frac{B_{0}^{4} k_{x}^{2} k_{z}^{4}}{16 \pi^{2}} + \frac{B_{0}^{4} k_{z}^{6}}{16 \pi^{2}} - \frac{B_{0}^{2} \Omega_{0} k_{z}^{4} q}{2 \pi} - 2 \Omega_{0} k_{z}^{2} q \sigma^{2} - \frac{4 \sigma}{\reym} \Omega_{0} k_{x}^{2} k_{z}^{2} q - \frac{4 \sigma}{\reym} \Omega_{0} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{x}^{4} k_{z}^{2} q - \frac{4 \Omega_{0}}{\reym^{2}} k_{x}^{2} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{z}^{6} q \\
-& - k_{x}^{2} \sigma^{4} - k_{z}^{2} \sigma^{4} + 4 k_{z}^{2} \sigma^{2} - \frac{2 \sigma^{3}}{\reym} k_{x}^{4} - \frac{4 \sigma^{3}}{\reym} k_{x}^{2} k_{z}^{2} + \frac{8 \sigma}{\reym} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reym} k_{z}^{4} + \frac{8 \sigma}{\reym} k_{z}^{4} - \frac{k_{x}^{6} \sigma^{2}}{\reym^{2}} - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{4} k_{z}^{2} + \frac{4 k_{x}^{4}}{\reym^{2}} k_{z}^{2} \\
-& - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{2} k_{z}^{4} + \frac{8 k_{x}^{2}}{\reym^{2}} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reym^{2}} + \frac{4 k_{z}^{6}}{\reym^{2}} - \frac{2 \sigma^{3}}{\reye} k_{x}^{4} - \frac{4 \sigma^{3}}{\reye} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reye} k_{z}^{4} - \frac{4 k_{x}^{6} \sigma^{2}}{\reye \reym} -  \frac{12 k_{x}^{4} k_{z}^{2} \sigma^{2}}{\reye \reym} - \frac{12 k_{x}^{2} k_{z}^{4} \sigma^{2}}{\reye \reym} \\
-& - \frac{4 k_{z}^{6} \sigma^{2}}{\reye \reym} - \frac{2 k_{x}^{8} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye \reym^{2}} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye \reym^{2}} - \frac{2 k_{z}^{8} \sigma}{\reye \reym^{2}} - \frac{k_{x}^{6} \sigma^{2}}{\reye^{2}} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{4} k_{z}^{2} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{2} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reye^{2}} \\
-& - \frac{2 k_{x}^{8} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye^{2} \reym} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye^{2} \reym} - \frac{2 k_{z}^{8} \sigma}{\reye^{2} \reym} - \frac{k_{x}^{10}}{\reye^{2} \reym^{2}} - \frac{5 k_{x}^{8} k_{z}^{2}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{6} k_{z}^{4}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{4} k_{z}^{6}}{\reye^{2} \reym^{2}} \\
-& - \frac{5 k_{x}^{2} k_{z}^{8}}{\reye^{2} \reym^{2}} - \frac{k_{z}^{10}}{\reye^{2} \reym^{2}} = 0
-\end{split}
-\eeq
-
-\section{B. Expansion of Nonlinear Terms}\label{app:nonlinear}
-\beq
-\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4)
-\eeq
-
-where
-
-\beq
-N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \frac{2}{\beta}J(A_1, \nabla^2 A_1) 
-\eeq
-
-\beq
-N_2^{(u)} = J(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_1) 
-\eeq
-
-\beq
-N_2^{(A)} = - J(A_1, \Psi_1) 
-\eeq
-
-\beq
-N_2^{(B)} =  J(\Psi_1, B_1) \, - \, J(A_1, u_1)
-\eeq
-
-and
-
-\beq
-\begin{split}
-N_3^{(\Psi)} & = J(\Psi_1, \nabla^2\Psi_2) \, - \, \frac{2}{\beta}J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) - \, \frac{2}{\beta}J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, \\
-& - \, 2 \frac{2}{\beta}J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \frac{2}{\beta}\widetilde{J}(A_1, \nabla^2 A_1)\\
-\end{split}
-\eeq
-
-\beq
-N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_2) \, - \, \frac{2}{\beta} J(A_2, B_1) \, - \, \frac{2}{\beta} \widetilde{J}(A_1, B_1)
-\eeq
-
-\beq
-N_3^{(A)} = - J(A_1, \Psi_2) \, - \, J(A_2, \Psi_1) \, - \, \widetilde{J}(A_1, \Psi_1)
-\eeq
-
-\beq
-N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1).\\
-\eeq
-
-$\mathbf{N_2}$ and $\mathbf{N_3}$ expand to become
-
-\beq
-\mathbf{N_2} = \alpha^2\mathbb{N}_{22} e^{i2 k_c z} + \left|\alpha\right|^2 \mathbb{N}_{20} + c.c.
-\eeq
-
-and
-
-\beq
-\mathbf{N_3} = \alpha^3 \mathbb{N}_{33} e^{i 3 k_c z} + \alpha\partial_Z\alpha \mathbb{N}_{32} e^{i 2 k_c z} + \alpha\left|\alpha\right|^2 \mathbb{N}_{31} e^{i k_c z} + \alpha \partial_Z  \beta \mathbb{\widetilde{N}}_{31} e^{i k_c z} + \alpha^*\partial_Z \alpha \mathbb{N}_{30} + c.c.
-\eeq
-
-The second order nonlinear terms are \\
-
-\beq
-\begin{split}
-N_{22}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11} - k_c^2 \partial_x \Psi_{11}\right) - \partial_x \Psi_{11} \cdot \left(i k_c \partial_x^2 \Psi_{11} - i k_c^3 \Psi_{11}\right) \\
-& + \frac{2}{\beta}\partial_x A_{11} \cdot \left(i k_c \partial_x^2 A_{11} - i k_c^3 A_{11}\right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11} - k_c^2 \partial_x A_{11}\right)
-\end{split}
-\eeq
-
-\beq
-\begin{split}
-N_{22}^{(u)} = & \, i k_c \Psi_{11} \cdot \partial_x u_{11} - \partial_x \Psi_{11} \cdot i k_c u_{11} - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11} + \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}
-\end{split}
-\eeq
-
-\beq
-N_{22}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11} + \partial_x A_{11} \cdot i k_c \Psi_{11}
-\eeq
-
-\beq
-N_{22}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11} - \partial_x \Psi_{11} \cdot i k_c B_{11} - i k_c A_{11} \cdot \partial_x u_{11} + \partial_x A_{11} \cdot i k_c u_{11}
-\eeq
-
-\beq
-\begin{split}
-N_{20}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - k_c^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i k_c^3 \Psi_{11}^* - i k_c \partial_x^2 \Psi_{11}^*\right) \\
-& +\frac{2}{\beta} \partial_x A_{11} \cdot \left(i k_c^3 A_{11}^* - i k_c \partial_x^2 A_{11}^* \right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
-\end{split}
-\eeq
-
-\beq
-N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* + \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11}^* - \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}^*
-\eeq
-
-\beq
-N_{20}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11}^* - \partial_x A_{11} \cdot i k_c \Psi_{11}^*
-\eeq
-
-\beq
-N_{20}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11}^* + \partial_x \Psi_{11} \cdot i k_c B_{11}^* - i k_c A_{11} \cdot \partial_x u_{11}^* - \partial_x A_{11} \cdot i k_c u_{11}^*
-\eeq \\
-
-and the third order nonlinear terms become \\
-
-\beq
-\begin{split}
-N_{31}^{(\Psi)} = &  \, i k_c \left(\Psi_{11} \cdot \partial_x^3 \Psi_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x^3\Psi_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x^3 \Psi_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot \partial_x^2 \Psi_{22}\right) \\
-& + i 8 k_c^3 \left(\partial_x\Psi_{11}^* \cdot \Psi_{22}\right) + i 4 k_c^3 \left(\Psi_{11}^* \cdot \partial_x \Psi_{22}\right) + \frac{2}{\beta}\left[ - i k_c \left(A_{11} \cdot \partial^3 A_{20}\right) - i k_c \left(A_{11} \cdot \partial_x^3 A_{20}^*\right) \right]\\
-& + \frac{2}{\beta}\left[i k_c \left(A_{11}^* \cdot \partial_x^3 A_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \partial_x^2 A_{22}\right) - i 8 k_c^3 \left(\partial_x A_{11}^* \cdot A_{22}\right) - i 4 k_c^3 \left(A_{11}^* \cdot \partial_x A_{22}\right)\right] \\
-& + i 2 k_c \left(\Psi_{22} \cdot \partial_x^3 \Psi_{11}^*\right) - i 2 k_c^3\left(\Psi_{22} \cdot \partial_x \Psi_{11}^* \right) - i k_c \left(\partial_x \Psi_{20} \cdot \partial_x^2 \Psi_{11}\right) + i k_c \left(\partial_x \Psi_{22} \cdot \partial_x^2 \Psi_{11}^*\right) \\
-& - i k_c \left(\partial_x \Psi_{20}^* \cdot \partial_x^2 \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20} \cdot \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20}^* \cdot \Psi_{11}\right) - i k_c^3 \left(\partial_x \Psi_{22} \cdot \Psi_{11}^*\right) \\
-& + \frac{2}{\beta} \left[ - i 2 k_c \left(A_{22} \cdot \partial_x^3 A_{11}^*\right) + i 2 k_c^3 \left(A_{22} \cdot \partial_x A_{11}^*\right) + i k_c \left(\partial_x A_{20} \cdot \partial_x^2 A_{11}\right) - i k_c \left(\partial_x A_{22} \cdot \partial_x^2 A_{11}^*\right) \right] \\
-& + \frac{2}{\beta} \left[ i k_c \left(\partial_x A_{20}^* \cdot \partial_x^2 A_{11}\right) - i k_c^3 \left(\partial_x A_{20} \cdot A_{11}\right) - i k_c^3 \left(\partial_x A_{20}^* \cdot A_{11}\right) + i k_c^3 \left(\partial_x A_{22} \cdot A_{11}^*\right)\right]
-\end{split}
-\eeq
-
-\beq
-\begin{split}
-N_{31}^{(u)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x u_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x u_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x u_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot u_{22}\right) \\
-& - i k_c \left(u_{11} \cdot \partial_x \Psi_{20} \right) - i k_c \left(u_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c \left(u_{11}^* \cdot \partial_x \Psi_{22} \right) + i 2 k_c \left(\partial_x u_{11}^* \cdot \Psi_{22}\right) \\
-& +\frac{2}{\beta}\left[- i k_c \left(A_{11} \cdot \partial_x B_{20}\right) - i k_c \left(A_{11} \cdot \partial_x B_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x B_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot B_{22}\right)\right] \\
-& +\frac{2}{\beta}\left[ i k_c \left(B_{11} \cdot \partial_x A_{20}\right) + i k_c \left(B_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(B_{11}^* \cdot \partial_x A_{20}\right) - i 2 k_c \left(\partial_x B_{11}^* \cdot A_{22}\right)\right]
-\end{split}
-\eeq
-
-%Correct -- checked 11/14/15
-\beq
-\begin{split}
-N_{31}^{(A)} =\, & -i k_c \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i k_c \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
- + i k_c \left(A_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \Psi_{22}\right) \\
- & + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
-\end{split}
-\eeq
-
-\beq
-\begin{split}
-N_{31}^{(B)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x B_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x B_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x B_{22}\right) - i 2 k_c\left(\partial_x \Psi_{11}^* \cdot B_{22}\right) \\
-& - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}\right) - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c\left(B_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x B_{11}^* \cdot \Psi_{22}\right) \\
-& - i k_c \left(A_{11} \cdot \partial_x u_{20}\right) - i k_c \left(A_{11} \cdot \partial_x u_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x u_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot u_{22}\right) \\
-& i k_c \left(u_{11} \cdot \partial_x A_{20} \right) + i k_c \left(u_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(u_{11}^* \cdot \partial_x A_{22}\right) - i 2 k_c \left(\partial_x u_{11}^* \cdot A_{22}\right)
-\end{split}
-\eeq
-
-%\bibliographystyle{plain}
-%\bibliography{mri_refs2}
 
@@
 
-\bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}
-Balbus, S A and Hawley, J F, 1991, The Astrophysical Journal, 376, 214
-
 \bibitem[{Boyd(2001)}]{Boyd:2001aa}
@@
 
-\bibitem[{Chandrasekhar(1960)}]{Chandrasekhar:1960wh}
-Chandrasekhar, S. 1960, Proceedings of the National Academy of Sciences of the United States of America, 46, 253
+\bibitem[{Clark \& Oishi(2016b)}]{Clark:2016}
+Clark, S.E. and Oishi, J.S., 2016, in prep
 
-\bibitem[{Cooper et al.(2014)}]{Cooper:2013to}
-Cooper, C M, Wallace, J, Brookhart, M, Clark, M, Collins, C, Ding, W X, Flanagan, K, Khalzov, I, Li, Y, Milhone, J, Nornberg, M, Nonn, P, Weisberg, D, Whyte, D G, Zweibel, E, Forest, C B, Physics of Plasmas, 2013, 21, 013505
-
-\bibitem[{Ebrahimi et al.(2009)}]{Ebrahimi:2009ey}
-Ebrahimi, F., Prager, S.C., Schnack, D.D
-
-\bibitem[{Gissinger et al.(2011)}]{Gissinger:2011td}
-Gissinger, C, Ji, H, Goodman, J, Phys. Rev. E, 2011, 84, 026308
-
-\bibitem[Goodman \& Xu(1994)]{Goodman:1994ul}
-Goodman, J, Xu, G, 1994, ApJ, 432, 213
-
-\bibitem[{Hollerbach(2009)}]{Hollerbach:2009ig}
-Hollerbach, R, Proc. R. Soc. A, 2009, 465, 2107
-
-\bibitem[{Hollerbach \& R\"udiger(2005)}]{Hollerbach:2005tr}
-Hollerbach, R, RŸdiger, G, Phys. Rev. Lett, 2005, 95, 124501
-
-\bibitem[{Ji et al.(2001)}]{Ji:2001kd}
-Ji, H, Goodman, J, Kageyama, A, Mon. Not. R. Astron. Soc., 2001, 325, 1
-
-\bibitem[{Knobloch \& Julien(2005)}]{Knobloch:2005ba}
-Knobloch, E, Julien K, 2005, Physics of Fluids, 17, 094106
-
-\bibitem[Pessah(2010)]{Pessah:2010ic}
-Pessah, M, 2010, ApJ, 716, 1012
-
-\bibitem[{Schartman et al.(2009)}]{Schartman:2009df}
-Schartman, E, Ji, H, Burin, M J, Rev Sci Instrum, 2009, 80, 024501
-
-\bibitem[{Shakura \& Sunyaev(1973)}]{Shakura:1973wg}
-Shakura, N I and Sunyaev, R A, Astronomy \& Astrophysics, 1973, 24, 337
-
-\bibitem[{Sisan et al.(2004)}]{Sisan:2004ig}
-Sisan, D R, Mujica, N, Tillotson, W A, Huang Y-M, Dorland, W, Hassam, A B, Antonsen, T M, Lathrop, D P, Phys Rev Lett, 2004, 93, 11
-
-\bibitem[{Stefani et al.(2006)}]{Stefani:2006iv}
-Stefani, F, Gundrum, T, Gerbeth, G, RŸdiger, G, Schultz, M, Szklarski, J, Hollerbach, R, Phys Rev Lett, 2006, 97, 184502
+\bibitem[{Newell \& Whitehead(1969)}]{Newell:1969}
+Newell, A.C. \& Whitehead, J.A. 1969, JFM 38, 279N
 

### 2016-08-08 paper/thingap_mri_paper.tex
discussion of other works

@@
 
-(general MRI intro)
+For matter to accrete from a disk onto a central object, angular momentum must be transported radially outward in the disk.
+The transport mechanism is likely turbulent, as molecular viscosity alone cannot account for the needed angular momentum transfer, and likely magnetic, as this turbulence is excited even in hydrodynamically stable disks (\citei{Shakura:1973wg}).
+Since its discovery by \citeauthor{Chandrasekhar:1960wh} (\citeyear{Chandrasekhar:1960wh}) and subsequent rediscovery by \citeauthor{Balbus:1991vs} (\citeyear{Balbus:1991vs}), the magnetorotational instability (MRI) remains the leading explanation for rapid angular momentum transport in astrophysical disks.
+The instability in its simplest geometry arises when a differentially rotating disk is threaded by a vertical magnetic field.
+The presence of the magnetic field linearly destabilizes the disk gas, driving turbulence and angular momentum transport.
+The MRI likely plays a role in a diverse host of astrophysical systems, including protoplanetary disks and black hole accretion disks, as well as stellar interiors.
+Despite its importance, many aspects of the MRI remain poorly understood.
+In particular, the nonlinear saturation mechanism for the MRI is an open question, and a formidable challenge.
+MRI saturation has been tackled almost exclusively with simulation, with a few notable exceptions detailed below.
+In this work we investigate the weakly nonlinear behavior of the MRI.
 
@@
 This work examines Taylor Couette flow in the thin-gap regime, and idealization in which the radial extent of the channel is very small compared to its distance from the center of rotation, i.e.
-$(r_2 - r_1) \ll \frac{1}{2} (r_1 + r_2)$.
-The thin-gap approximation eliminates curvature terms, so the domain is Cartesian.
-These curvature terms have an explicit dependence on $r$, so they make the problem more challenging both analytically and numerically.
+$(r_2 - r_1) \ll \frac{1}{2} (r_1 + r_2)$ where $r1$ and $r2$ are the radii of the inner and outer flow boundaries, respectively.
+The thin-gap approximation eliminates curvature terms, so the domain geometry is Cartesian rather than cylindrical.
+The excluded curvature terms have an explicit dependence on $r$, so they make the problem more challenging both analytically and numerically.
+In particular, in the wide-gap geometry (i.e.
+true Taylor Couette flow) the base angular velocity is a function of $r$, where in the thin-gap approximation the shear flow reduces to a linear profile.
+
+We note several other important analytical studies of MRI saturation.
+\citeauthor{Knobloch:2005ba} (\citeyear{Knobloch:2005ba}) analyze the MRI in the strongly nonlinear regime, by following the already-developed MRI modes into asymptotic saturation.
+They find that this saturation proceeds via modification of the background shear.
+\citeauthor{Knobloch:2005ba} consider a thin-gap regime as well, and so their theory may be considered the strongly nonlinear analogue to the one developed here.
+\citei{Vasil:2015} examines the weakly nonlinear MRI in a thin-gap regime from a heuristic perspective, finding intriguing similarities between the MRI system and an elastodynamic instability.
+Several authors have investigated the behavior of the MRI when the boundary conditions are shear periodic, and so the MRI has no mechanism by which to modify the background shear flow profile.
+In this paradigm linear MRI growth is dominated by channel modes, a type of MRI mode that, for periodic boundary conditions, are exact solutions of both the linear and nonlinear MRI equations (\citei{Goodman:1994ul}).
+In this regime the MRI saturates via parasitic instabilities, which feed off and destroy the primary MRI modes.
+Analytical investigation of this case reveals that MRI saturation can be caused by parasitic Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime (\citei{Pessah:2010ic}).
+The theory of MRI channel mode parasites is relatively robust (see also \citei{Pessah:2009gm}, \citei{Latter:2010iz}, \citei{Rembiasz:2016}), but they are mostly an artifact of the local approximation, and not germane to global analyses like the one presented here (see \citei{Latter:2015} for a detailed analysis of the relation between local and global MRI modes).
 
@@
 %in units where the permittivity of free space is unity. 
-Equations \ref{momentum} and \ref{induction} are subject to the magnetic solenoid and incompressibility constraints,
+Equations \ref{eq:momentum} and \ref{eq:induction} are subject to the magnetic solenoid and incompressibility constraints,
 
@@
 
+\bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}
+Balbus, S A and Hawley, J F, 1991, The Astrophysical Journal, 376, 214
+
 \bibitem[{Boyd(2001)}]{Boyd:2001aa}
@@
 
+\bibitem[{Chandrasekhar(1960)}]{Chandrasekhar:1960wh}
+Chandrasekhar, S. 1960, Proceedings of the National Academy of Sciences of the United States of America, 46, 253
 \bibitem[{Clark \& Oishi(2016b)}]{Clark:2016}
@@
 
+\bibitem[Goodman \& Xu(1994)]{Goodman:1994ul}
+Goodman, J, Xu, G, 1994, ApJ, 432, 213
+
+\bibitem[{Knobloch \& Julien(2005)}]{Knobloch:2005ba}
+Knobloch, E, Julien K, 2005, Physics of Fluids, 17, 094106
+
+\bibitem[{Latter et al.(2015)}]{Latter:2015}
+Latter, H.N., Fromang, S., \& Faure, J., 2015, MNRAS 453, 3257
+
+\bibitem[{Latter et al.(2010)}]{Latter:2010iz}
+Latter, H.N., Fromang, S., \& Gressel, O., 2010, MNRAS 406, 848
+
 \bibitem[{Newell \& Whitehead(1969)}]{Newell:1969}
 Newell, A.C. \& Whitehead, J.A. 1969, JFM 38, 279N
+
+\bibitem[{Pessah(2010)}]{Pessah:2010ic}
+Pessah, M. E., 2010, ApJ, 716, 1012
+
+\bibitem[{Pessah \& Goodman(2009)}]{Pessah:2009gm}
+Pessah, M. E., \& Goodman, J., 2009, ApJ 698, L72
+
+\bibitem[{Rembiasz et al.(2016)}]{Rembiasz:2016}
+Rembiasz, T., Obergaulinger, M., Cerd\'a-Dur\'an, P., M\"uller, E., \& Aloy, M.A., 2016, MNRAS 456, 3782
+
+\bibitem[{Shakura \& Sunyaev(1973)}]{Shakura:1973wg}
+Shakura, N I and Sunyaev, R A, Astronomy \& Astrophysics, 1973, 24, 337
 
@@
 
+\bibitem[{Vasil(2015)}]{Vasil:2015}
+Vasil, G. M. 2015, RSPSA 471, 20140699
+
 \end{thebibliography}

### 2016-08-09 paper/thingap_mri_paper.tex
run helical mri critical param run w new complex eigenvalue finder

@@
 \citeauthor{Knobloch:2005ba} consider a thin-gap regime as well, and so their theory may be considered the strongly nonlinear analogue to the one developed here.
-\citei{Vasil:2015} examines the weakly nonlinear MRI in a thin-gap regime from a heuristic perspective, finding intriguing similarities between the MRI system and an elastodynamic instability.
+\citei{Vasil:2015} examines the weakly nonlinear MRI in a thin-gap regime from a heuristic perspective, finding deep mathematical similarities between the MRI system and an elastodynamic instability.
 Several authors have investigated the behavior of the MRI when the boundary conditions are shear periodic, and so the MRI has no mechanism by which to modify the background shear flow profile.
-In this paradigm linear MRI growth is dominated by channel modes, a type of MRI mode that, for periodic boundary conditions, are exact solutions of both the linear and nonlinear MRI equations (\citei{Goodman:1994ul}).
+In this approximation linear MRI growth is dominated by channel modes, a type of MRI mode that, for periodic boundary conditions, are exact solutions of both the linear and nonlinear MRI equations (\citei{Goodman:1994ul}).
 In this regime the MRI saturates via parasitic instabilities, which feed off and destroy the primary MRI modes.

### 2016-08-09 paper/widegap_mri_paper.tex
run helical mri critical param run w new complex eigenvalue finder

@@
 The diversity of astrophysical systems which may permit the MRI admits an enormous parameter space to be explored.
-In protoplanetary disks, for example, the behavior and evolution of the MRI may change drastically depending on the properties of the magnetic field, the disk composition, disk geometry, etc.
-Numerical simulation of realistic disk physics is currently an area of intense focus, and is enabling the study of nonideal MHD effects, disk stratification, nonequilibrium chemistry, and other complex physics that does not lend itself easily to analytic study.
+In protoplanetary disks, for example, the behavior and evolution of the MRI may change drastically depending on the properties of the magnetic field, the disk composition, disk geometry, and so forth.
+Numerical simulation of realistic disk physics is currently an area of intense focus, and is enabling the study of nonideal MHD effects, disk stratification, nonequilibrium chemistry, and other complex physics that does not lend itself easily to analytic study (\citei{Fleming:2003fs}, \citei{Bai:2011cm}, \citei{Flock:2013}, \citei{Suzuki:2014vh}, to name only a few).
 Still, computational costs inevitably constrain numerical approaches.
@@
 
-The theory we develop here employs realistic radial boundary conditions, and so MRI modes are global and channel modes (at least in their purest, radially independent are not present.
+The theory we develop here employs realistic radial boundary conditions, and so MRI modes are global and channel modes are not present.
 A number of saturation mechanisms have been proposed for the MRI which do not rely on channel modes dominating the flow.
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{../figures/widegap_rey_max_tot_stresses.png}
-\caption{Reynolds ($\mathbb{T}_{R} = u_r u_\phi$), Maxwell ($\mathbb{T}_{M} = -\frac{2}{\beta} B_r B_\phi$), and total stress ($\mathbb{T} = \mathbb{T}_{R} + \mathbb{T}_{M}$) for the fiducial standard MRI case.
-}
-\end{figure*}
-
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{../figures/widegap_3D_vel_screenshot.png}
-\caption{Visualization of the velocity perturbation field in the fiducial SMRI case.
-The three dimensional path of twenty randomly placed tracer particles is plotted, where the particle velocity is defined as the total (first + second order) velocity perturbation.
-This is meant to be a visualization only of the velocity field, so the particle motion feels no contribution from the magnetic field.}
-\end{figure}
-
 See Appendix \ref{app:basic_equations} for the definition of matrices and nonlinear vectors, and a thorough derivation.
@@
  %This complex GLE is a natural form for the HMRI, as it is a traveling wave instability.
+
+\section{Results and Discussion}
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../figures/widegap_rey_max_tot_stresses.png}
+\caption{Reynolds ($\mathbb{T}_{R} = u_r u_\phi$), Maxwell ($\mathbb{T}_{M} = -\frac{2}{\beta} B_r B_\phi$), and total stress ($\mathbb{T} = \mathbb{T}_{R} + \mathbb{T}_{M}$) for the fiducial standard MRI case.
+}
+\end{figure*}
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../figures/widegap_3D_vel_screenshot.png}
+\caption{Visualization of the velocity perturbation field in the fiducial SMRI case.
+The three dimensional path of twenty randomly placed tracer particles is plotted, where the particle velocity is defined as the total (first + second order) velocity perturbation.
+This is meant to be a visualization only of the velocity field, so the particle motion feels no contribution from the magnetic field.}
+\end{figure}
+
+\subsection{Standard MRI}
 
@@
 
-\section{Channel Modes}
-Channel modes are radially independent linear MRI modes which are, under certain conditions, exact nonlinear solutions to the MRI equations.
+%\section{Channel Modes}
+%Channel modes are radially independent linear MRI modes which are, under certain conditions, exact nonlinear solutions to the MRI equations.
+
+\subsection{Helical MRI}
+
+When $\xi$ in Equation \ref{eq:unperturbed_matrix_eqns} is not equal to zero, the helical MRI arises.
+We examine a single fiducial helical MRI case, for the parameters listed in Table \ref{table:parameters}.
+At the conclusion of the weakly nonlinear analysis, we find that the coefficients of Equation \ref{eq:gle} are $b = $, $d = $, and $c = $, i.e.
+complex.
+The marginal helical MRI is thus described by a complex GLE.
 
@@
 
+\bibitem[{Bai(2011)}]{Bai:2011cm}
+Bai, X. 2011, ApJ 739, 50
+
 \bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}
@@
 Ebrahimi, F., Prager, S.C., Schnack, D.D
+
+\bibitem[{Fleming \& Stone(2003)}]{Fleming:2003fs}
+Fleming, T. \& Stone, J. 2003, ApJ 585, 908
+
+\bibitem[{Flock et al.(2013)}]{Flock:2013}
+Flock, M., Fromang, S., Gonz\'alez, M., Commer\c{c}on, B. A\&A 560, A43
 
@@
 
+\bibitem[{Suzuki \& Inutsuka(2014)}]{Suzuki:2014vh}
+Suzuki, T.K. \& Inutsuka, S. 2014, ApJ 784, 121
+
 \bibitem[{Umurhan et al.(2007a)}]{Umurhan:2007dz}

### 2016-08-09 python/widegap/allorders2_widegap.py
run rhs search in devel

@@ -67,7 +67,7 @@ class MRI():
         self.R0 = (self.R1 + self.R2)/2.0
         self.q_R0 = 2*self.c2/(self.R0**2*self.c1 + self.c2)
         
-        logger.info("MRI parameters: Q = {}; Rm = {}; Pm = {}; beta = {}; norm = {}, Re = {}".format(self.Q, self.Rm, self.Pm, self.beta, norm, self.R))
+        logger.info("MRI parameters: Q = {}; Rm = {}; Pm = {}; beta = {}; norm = {}, Re = {}, xi = {}".format(self.Q, self.Rm, self.Pm, self.beta, norm, self.R, self.xi))
         logger.info("Effective shear parameter q(R0) = {}".format(self.q_R0))
         
         if self.xi != 0:
@@ -93,8 +93,6 @@ class MRI():
             problem.add_bc('right(B + r*Br) = 0') # axial component of current = 0
         else:
             # Insulating boundary conditions
-            problem.parameters['bessel1'] = special.iv(0, self.Q*self.R1)/special.iv(1, self.Q*self.R1)
-            problem.parameters['bessel2'] = special.kn(0, self.Q*self.R2)/special.kn(1, self.Q*self.R2)
             problem.add_bc('left(dr(r*dz*A) - Q*r*bessel1*dz*A) = 0')
             problem.add_bc('right(dr(r*dz*A) + Q*r*bessel2*dz*A) = 0')
             problem.add_bc('left(B) = 0')
@@ -291,6 +289,9 @@ class AdjointHomogenous(MRI):
         lv1.parameters['c2'] = self.c2
         lv1.parameters['B0'] = self.B0
         lv1.parameters['dz'] = 1j*self.Q
+        if conducting is False:
+            lv1.parameters['bessel1'] = special.iv(0, self.Q*self.R1)/special.iv(1, self.Q*self.R1)
+            lv1.parameters['bessel2'] = special.kn(0, self.Q*self.R2)/special.kn(1, self.Q*self.R2)
         
         lv1.substitutions['ru0'] = '(r*r*c1 + c2)' # u0 = r Omega(r) = Ar + B/r
         lv1.substitutions['rrdu0'] = '(c1*r*r-c2)' # du0/dr = A - B/r^2
@@ -378,6 +379,9 @@ class OrderE(MRI):
[…]

### 2016-08-09 paper/widegap_mri_paper.tex
run rhs search in devel
helical crit search nondim by inner radius

@@
   \label{eq:couette_flow}
-  \Omega(r) = a + \frac{b}{r^2},
+  \Omega(r) = c_1 + \frac{c_2}{r^2},
 \end{equation}
-where $a = (\Omega_2 r^2_2 - \Omega_1 r^2_1)/(r^2_2 - r^2_1)$, $b = r^2_1 r^2_2 (\Omega_1 - \Omega_2)/(r^2_2 - r^2_1)$, and $\Omega_1$ and $\Omega_2$ are the rotation rates at the inner and outer cylinder radii, respectively.
+where $c_1 = (\Omega_2 r^2_2 - \Omega_1 r^2_1)/(r^2_2 - r^2_1)$, $c_2 = r^2_1 r^2_2 (\Omega_1 - \Omega_2)/(r^2_2 - r^2_1)$, and $\Omega_1$ and $\Omega_2$ are the rotation rates at the inner and outer cylinder radii, respectively.
 In the laboratory, $r_1$ and $r_2$ are typically fixed by experimental design.
@@
   \label{eq:couette_q}
-  q(r) \equiv -\frac{d \ln \Omega}{d \ln r} = \frac{2 b}{a r^2 + b}.
+  q(r) \equiv -\frac{d \ln \Omega}{d \ln r} = \frac{2 c_2}{c_1 r^2 + c_2}.
 \end{equation}
@@
 The marginal helical MRI is thus described by a complex GLE.
+This difference in character between the amplitude equations that modulate the weakly nonlinear standard and helical MRI is a consequence of the same property that makes the helical MRI an overstability.
 

### 2016-08-09 python/widegap/find_crit.py

@@ -136,7 +136,7 @@ def find_crit(R1, R2, Omega1, Omega2, beta, xi, Pm, Rm_min, Rm_max, k_min, k_max
             if magnetic_bcs == "conducting":
                 gridname = '../../data/hmri_growth_rates_res{0:d}_Rmmin{1:5.02e}_Rmmax{2:5.02e}_kmin{3:5.02e}_kmax{4:5.02e}_nRm{5:5.02e}_nk{6:5.02e}'.format(nr,Rm_min,Rm_max, k_min, k_max, n_Rm, n_k)
             elif magnetic_bcs == "insulating":
-                gridname = '../../data/hmri_growth_rates_res{0:d}_Rmmin{1:5.02e}_Rmmax{2:5.02e}_kmin{3:5.02e}_kmax{4:5.02e}_nRm{5:5.02e}_nk{6:5.02e}_Omega2{7:5.02e}_R1_{}_R2_{}_insulating'.format(nr,Rm_min,Rm_max, k_min, k_max, n_Rm, n_k, Omega2, R1, R2)
+                gridname = '../../data/hmri_growth_rates_res{0:d}_Rmmin{1:5.02e}_Rmmax{2:5.02e}_kmin{3:5.02e}_kmax{4:5.02e}_nRm{5:5.02e}_nk{6:5.02e}_Omega1_{7:5.02e}_Omega2_{7:5.02e}_R1_{}_R2_{}_insulating'.format(nr,Rm_min,Rm_max, k_min, k_max, n_Rm, n_k, Omega1, Omega2, R1, R2)
         
         cf.save_grid(gridname)
 

### 2016-08-09 python/widegap/run_helical_crit_search.sh

@@ -13,5 +13,5 @@ source /u/sclark9/dedalus/bin/activate
 export PYTHONPATH=$PYTHONPATH:$HOME/eigentools
 export MPLBACKEND='Agg'
 date
-mpiexec_mpt -np 400 python3 $dedalus_script.py --R1=1 --R2=2 --Omega1=313.55 --Omega2=87.794 --Pm=1.0E-6 --beta=0.0174 --xi=4.0 --Rm_min=1.0E-4 --Rm_max=1.0E-2 --k_min=1.0E-1 --k_max=10.0 --insulate=1
+mpiexec_mpt -np 400 python3 $dedalus_script.py --R1=1 --R2=2 --Omega1=1.0 --Omega2=0.27 --Pm=1.0E-6 --beta=0.0174 --xi=4.0 --Rm_min=1.0E-4 --Rm_max=1.0E-2 --k_min=1.0E-1 --k_max=10.0 --insulate=1
 date

### 2016-08-11 paper/mri_paper.tex

@@
 
-This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime.
+In this work we investigate the MRI in the weakly nonlinear regime, where the initial linear growth phase of the instability begins to give way to nonlinear saturation.
+Weakly nonlinear analysis is a perturbative method for examining a system at marginality.
+Weakly nonlinear analysis was instrumental in understanding the saturated state of Rayleigh-B\'enard convection (\citei{Newell:1969wr}).
+
+%This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime.
 The present work adds to this foundation both in subject, with our additional analysis of the helical MRI, as well as in scope, with a detailed analysis of MRI saturation properties across parameter space and the comparison of weakly nonlinear analytic work to strongly nonlinear simulation.

### 2016-08-15 paper/widegap_mri_paper.tex
changed nondim description to inner cylinder

@@
 Note that in the above, $\nabla^2 f \equiv \partial_r^2 f + \partial_z^2 f + \frac{1}{r} \partial_r f$.
-Equations \ref{eq:Psi_perturbed} - \ref{eq:nonlinear_B} are nondimensionalized, where lengths have been scaled by $r_0$, velocities by $r_0 \Omega_0$, densities by $\rho_0$, and magnetic fields by $B_0$; where $B_0$ appears in the above it is formally unity.
-$\Omega_0 = \Omega(r_0)$ is the rotation rate at the center of the channel.
-We introduce the Reynolds number $\reye = \Omega_0 r_0^2/\nu$, the magnetic Reynolds number $\reym = \Omega_0 r_0^2 / \eta$, and a plasma beta parameter $\beta = \Omega_0^2 r_0^2 \rho_0/B_0^2$.
-If we define the dimensional cylindrical coordinate $r = r_0(1 + \delta x)$, the narrow gap equations are recovered in the limit $\delta \rightarrow 0$.
+Equations \ref{eq:Psi_perturbed} - \ref{eq:nonlinear_B} are nondimensionalized by inner cylinder quantities: lengths have been scaled by $r_1$, velocities by $r_1 \Omega_1$, and densities by $\rho_0$, where $\rho_0$ is the constant density.
+Magnetic fields are scaled by $B_0$, the constant strength of the initial background field; where $B_0$ appears in the above it is formally unity.
+$\Omega_1 = \Omega(r_1)$ is the rotation rate of the inner cylinder.
+We introduce the Reynolds number $\reye = \Omega_1 r_1^2/\nu$, the magnetic Reynolds number $\reym = \Omega_1 r_1^2 / \eta$, and a plasma beta parameter $\beta = \Omega_1^2 r_1^2 \rho_0/B_0^2$.
+If we define the dimensional cylindrical coordinate $r = r_1(1 + \delta x)$, the narrow gap equations are recovered in the limit $\delta \rightarrow 0$.
 

### 2016-08-15 paper/widegap_mri_paper.tex
widegap added description of adjoint to appendix

@@
   \hline\hline
-Standard MRI & 0 & 1E-3 & 25 & 0.18 & 0.33 & conducting \\ 
+%Standard MRI & 0 & 1E-3 & 25 & 0.18 & 0.33 & conducting \\ 
+Standard MRI & 0 & 1.6E-6 & 41.2 & 0.121 & 0.33 & conducting \\ 
 Helical MRI & 4 & 1E-6 & 1.7E-2 & 0.27 & 0.5 & insulating\\ 
@@
 
+%The GLE is derived from 
+%Equation \ref{eq:ordere3} is resonant with 
+To find a bounded solution at $\mathcal{O}(\epsilon^3)$ we must eliminate secular terms: terms which are resonant with the solution to the homogenous equation $(\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} = 0$ and cause the solution to grow without bound.
+To eliminate these terms we enforce a solvability condition, which arises as a consequence of the Fredholm alternative.
+This solvability condition is the requirement that the inner product of the equation with its adjoint homogenous solution is zero.
+We define the adjoint operator $\mathcal{L}^\dagger$ and solution $\mathbf{V}^\dagger$ as
+
+\beq
+\langle \mathbf{V^\dagger} | (\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} \rangle = \langle (\mathcal{L}^\dagger + \mathcal{H}^\dagger) \mathbf{V}^\dagger | \mathbf{V} \rangle,
+\eeq
+
+where the inner product is defined as
+
+\beq\label{eq:inner_product_def}
+\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \frac{k_c}{2\pi} \int_{-\pi/k_c}^{\pi/k_c} \int_{r_1}^{r_2} \mathbf{V}^{\dagger*} \cdot \mathcal{L} \mathbf{V} \, \mathrm{d} r \mathrm{d} z
+\eeq
+
+We find that
+
+\beq
+\mathcal{L}^\dagger = \mathcal{L}_0^\dagger - \partial_z \mathcal{L}_1^\mathrm{T} + d_z^2 \mathcal{L}_2^\mathrm{T} - \partial_z^3 \mathcal{L}_3^\mathrm{T} + \partial_z^4 L_4^\mathrm{T},
+\eeq
+
+and
+
+\beq
+\mathcal{H}^\dagger = - d_z \mathcal{H}^\mathrm{T}
+\eeq
+
+where
+
+\beq
+\mathcal{L}_0^\dagger = \left[\begin{matrix}
+-\frac{1}{\reye} (-\frac{3}{r^4} \partial_r + \frac{3}{r^3}\partial_r^2 - \frac{2}{r^2}\partial_r^3 + \frac{1}{r}\partial_r^4) & 0 & 0 & 0 \\
+0 & -\frac{1}{\reym} (\partial_r^2 - \frac{1}{r} \partial_r) & 0 & 0 \\
+0 & 0 &-\frac{1}{\reye} (\partial_r^2 + \frac{1}{r}\partial_r - \frac{1}{r^2}) & 0 \\
+0 & 0 & 0 & -\frac{1}{\reym} (\partial_r^2 - \frac{1}{r} \partial_r) \\
+\end{matrix}\right].
+\eeq
+
+The adjoint boundary conditions are selected to satisfy Equation \ref{eq:inner_product_def}, and differ depending on the boundary conditions enforced on the original system.
+For the conducting boundary conditions we apply to the standard MRI, the adjoint equation
+
+\beq
+(\mathcal{L}^\dagger + \xi \widetilde{\mathcal{H}}^\dagger)\mathbf{V}^\dagger = 0
+\eeq
+
+must be solved subject to the boundary conditions
+
+\beq
+\Psi^\dagger = \partial_r \Psi^\dagger = u^\dagger = A^\dagger = \partial_r B^\dagger = 0.
+\eeq
+
 \section{B. Nonlinear Terms}\label{app:nonlinear_terms}

### 2016-08-15 python/widegap/allorders2_widegap.py
fixed typos jeff caught in N20 psi (version multiplied by r^4). rederived coeffificients

@@ -584,7 +584,7 @@ class N2(MRI):
         #if self.xi != 0:
         logger.info("N20 psi r4 term")
         # explicitly include (psi1, psi1star) and (psi1star, psi1)
-        N20_psi_r4_rederived = ((-2*rfield*(1j*Q)**2*o1.psi*(-1j*Q)*o1.psi_star + rfield**2*(1j*Q)*o1.psi*(-1*Q)**2*o1.psi_star_r
+        N20_psi_r4_rederived = ((-2*rfield*(1j*Q)**2*o1.psi*(-1j*Q)*o1.psi_star + rfield**2*(1j*Q)*o1.psi*(-1j*Q)**2*o1.psi_star_r
                                - rfield**2*o1.psi_r*(-1j*Q)**3*o1.psi_star + 3*(1j*Q)*o1.psi*o1.psi_star_r
                                + rfield*(-1j*Q)*o1.psi_r*o1.psi_star_r - 3*rfield*(1j*Q)*o1.psi*o1.psi_star_rr
                                + rfield**2*(1j*Q)*o1.psi*o1.psi_star_rrr - rfield**2*o1.psi_r*(-1j*Q)*o1.psi_star_rr
@@ -594,8 +594,8 @@ class N2(MRI):
                                + rfield*(-1j*Q)*o1.psi_star_r*o1.psi_r - 3*rfield*(-1j*Q)*o1.psi_star*o1.psi_rr
                                + rfield**2*(-1j*Q)*o1.psi_star*o1.psi_rrr - rfield**2*o1.psi_star_r*(1j*Q)*o1.psi_rr
                                - 2*rfield**3*1j*Q*o1.u*o1.u_star)
-                   - (2/beta)*(-2*rfield*(1j*Q)**2*o1.A*(-1j*Q)*o1.A_star + rfield**2*(1j*Q)*o1.A*(-1*Q)**2*o1.A_star_r
-                               - rfield**2*o1.A_star_r*(-1j*Q)**3*o1.A_star + 3*(1j*Q)*o1.A*o1.A_star_r
+                   - (2/beta)*(-2*rfield*(1j*Q)**2*o1.A*(-1j*Q)*o1.A_star + rfield**2*(1j*Q)*o1.A*(-1j*Q)**2*o1.A_star_r
+                               - rfield**2*o1.A_r*(-1j*Q)**3*o1.A_star + 3*(1j*Q)*o1.A*o1.A_star_r
[…]

### 2016-08-16 paper/widegap_mri_paper.tex
fixed typos jeff caught in N20 psi (version multiplied by r^4). rederived coeffificients
added insulating bc description to main text

@@
 
-We solve the system subject to periodic vertical boundary conditions and no-slip, perfectly conducting radial boundary conditions, namely
+We solve the standard MRI system subject to periodic vertical boundary conditions and no-slip, perfectly conducting radial boundary conditions, namely
 
@@
 
-at $r = r_1, r_2$.
+at $r = r_1, r_2$. To the helical MRI system we apply insulating boundary conditions,
+
+\beq
+\partial_r A = k \frac{I_0 (k r)}{I_1 (k r)} A \, \, \mathrm{at} \, \, r = r_1
+\eeq
+
+\beq
+\partial_r A = - k \frac{K_0 (k r)}{K_1 (k r)} A \, \, \mathrm{at} \, \, r = r_2
+\eeq
+
+and $B = 0$ at $r = r_1, r_2$. Here, $I_n$, $K_n$ are the modified Bessel functions of the first and second kind, respectively.
 
@@
 
+Note that the condition on $B^\dagger$ differs from that on $B$. We take the inner product of the adjoint homogenous solution with Equation \ref{eq:ordere3} to find
+
+\beq
+\mathcal{L}\mathbf{V}_3 + \xi \widetilde{\mathcal{H}} \mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2\nonumber 
++ \xi \mathcal{H}\partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 - \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3 = 0.
+\eeq
+
+%%%%% NONLINEAR TERMS %%%%%%%%%%%
 \section{B. Nonlinear Terms}\label{app:nonlinear_terms}

### 2016-08-16 python/widegap/run_widegap_IVP.py
test hmri

@@ -16,47 +16,21 @@ class Coefficients():
         self.c = coeffs.attrs['c']
         self.h = coeffs.attrs['h']
 
-# UMR07-like parameters
-
-Q = 0.01269
-#Rm = 0.67355
-Rm = 0.8403
-Pm = 1.0e-3
-beta = 25.0
-Omega1 = 313.55
-Omega2 = 56.43
-xi = 0
-R1 = 5
-R2 = 15
-gridnum = 128
-
-# Hollerbach & Rudiger
-"""
-Pm = 1.0E-6
-Re = 1521
-Rm = Pm*Re
-xi = 4
-Q = 2.33
-Ha = 16.3
-beta = (2*Re*Rm)/(Ha**2)
-R1 = 1.
-R2 = 2.
-mu_omega = 0.27
-Omega1 = 313.55
-Omega2 = Omega1*mu_omega
-"""
-
+        self.Q = coeffs.attrs['Q']
+         
 # data created in run_widegap_amplitude.py
 fn_root = "/Users/susanclark/weakly_nonlinear_mri/data/"
-fn = "widegap_amplitude_parameters_Q_{:03.2f}_Rm_{:04.4f}_Pm_{:.2e}_Omega1_{:05.2f}_Omega2_{:05.2f}_beta_{:.2f}_xi_{:.2f}_gridnum_{}_Anorm".format(Q, Rm, Pm, Omega1, Omega2, beta, xi, gridnum)
+#fn = "widegap_amplitude_parameters_Q_{:03.2f}_Rm_{:04.4f}_Pm_{:.2e}_Omega1_{:05.2f}_Omega2_{:05.2f}_beta_{:.2f}_xi_{:.2f}_gridnum_{}_Anorm".format(Q, Rm, Pm, Omega1, Omega2, beta, xi, gridnum)
+fn = "widegap_amplitude_parameters_Q_2.33_Rm_0.0015_Pm_1.00e-06_Omega1_01.00_Omega2_00.27_beta_0.02_xi_4.00_gridnum_128_conducting_False_norm_False"
 coeff_fn = fn_root + fn + ".h5"
-
-# For IVP
-gridnum = 128
-lambda_crit = 2*np.pi/Q
 
 # Read coefficients from file
 coeff = Coefficients(coeff_fn)
+
+# For IVP
+Q = coeff.Q
+gridnum = 512
+lambda_crit = 2*np.pi/Q
 
 #print("HACK: all coeffs -> real")
 #coeff.a = coeff.a.real
[…]

### 2016-08-29 paper/widegap_mri_paper.tex
insulating adjoint bc's added to appendix
test hmri
ncc test w jeff's params

@@
 
-and $B = 0$ at $r = r_1, r_2$. Here, $I_n$, $K_n$ are the modified Bessel functions of the first and second kind, respectively.
+and $B = 0$ at $r = r_1, r_2$ (see \citei{Willis:2002bh}). Here, $I_n$, $K_n$ are the modified Bessel functions of the first and second kind, respectively.
 
@@
 This difference in character between the amplitude equations that modulate the weakly nonlinear standard and helical MRI is a consequence of the same property that makes the helical MRI an overstability.
+With the introduction of an azimuthal component, the background magnetic field acquires a handedness that is not present in a purely axial field.
+
+%************* APPENDICES ************************%
 
@@
 \beq
-\langle \mathbf{V^\dagger} | (\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} \rangle = \langle (\mathcal{L}^\dagger + \mathcal{H}^\dagger) \mathbf{V}^\dagger | \mathbf{V} \rangle,
+\langle \mathbf{V^\dagger} | (\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} \rangle = \langle (\mathcal{L}^\dagger + \xi \mathcal{H}^\dagger) \mathbf{V}^\dagger | \mathbf{V} \rangle,
 \eeq
@@
 \beq\label{eq:inner_product_def}
-\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \frac{k_c}{2\pi} \int_{-\pi/k_c}^{\pi/k_c} \int_{r_1}^{r_2} \mathbf{V}^{\dagger*} \cdot \mathcal{L} \mathbf{V} \, \mathrm{d} r \mathrm{d} z
+\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \frac{k_c}{2\pi} \int_{-\pi/k_c}^{\pi/k_c} \int_{r_1}^{r_2} \mathbf{V}^{\dagger*} \cdot \mathcal{L} \mathbf{V} \, r \mathrm{d} r \mathrm{d} z
 \eeq
 
-We find that
+We derive the adjoint operator by successive integration by parts, to find
 
 \beq
-\mathcal{L}^\dagger = \mathcal{L}_0^\dagger - \partial_z \mathcal{L}_1^\mathrm{T} + d_z^2 \mathcal{L}_2^\mathrm{T} - \partial_z^3 \mathcal{L}_3^\mathrm{T} + \partial_z^4 L_4^\mathrm{T},
+\mathcal{L}^\dagger = \mathcal{L}_0^\dagger - \partial_z \mathcal{L}_1^\dagger + d_z^2 \mathcal{L}_2^\dagger - \partial_z^3 \mathcal{L}_3^\dagger + \partial_z^4 L_4^\dagger
 \eeq
@@
 \beq
-\mathcal{H}^\dagger = - d_z \mathcal{H}^\mathrm{T}
+\mathcal{H}^\dagger = - d_z \mathcal{H}^\mathrm{T},
 \eeq
@@
 \mathcal{L}_0^\dagger = \left[\begin{matrix}
--\frac{1}{\reye} (-\frac{3}{r^4} \partial_r + \frac{3}{r^3}\partial_r^2 - \frac{2}{r^2}\partial_r^3 + \frac{1}{r}\partial_r^4) & 0 & 0 & 0 \\
+-\frac{1}{\reye} (-\frac{3}{r^5} + \frac{3}{r^4}\partial_r - \frac{3}{r^3} \partial_r^2 + \frac{2}{r^2}\partial_r^3 + \frac{1}{r}\partial_r^4 ) & 0 & 0 & 0 \\
 0 & -\frac{1}{\reym} (\partial_r^2 - \frac{1}{r} \partial_r) & 0 & 0 \\
@@
 
-\beq
+\beq\label{eq:adjoint}
 (\mathcal{L}^\dagger + \xi \widetilde{\mathcal{H}}^\dagger)\mathbf{V}^\dagger = 0
@@
 
-Note that the condition on $B^\dagger$ differs from that on $B$. We take the inner product of the adjoint homogenous solution with Equation \ref{eq:ordere3} to find
+Note that the condition on $B^\dagger$ differs from that on $B$.
+For the insulating boundary conditions we apply to the helical MRI, the magnetic boundary conditions applied to Equation \ref{eq:adjoint} are
+
+\beq
+B^\dagger = (- k \frac{I_0 (k r)}{I_1 (k r)} + \frac{1}{r} + \partial_r ) A^\dagger = 0 \, \, \mathrm{at} \, \, r = r_1
+\eeq
+
+\beq
+B^\dagger = (k \frac{K_0 (k r)}{K_1 (k r)} + \frac{1}{r} + \partial_r ) A^\dagger = 0 \, \, \mathrm{at} \, \, r = r_2
+\eeq
+
+We take the inner product of the adjoint homogenous solution with Equation \ref{eq:ordere3} to find
 
@@
 
+\bibitem[{Willis \& Barenghi(2002)}]{Willis:2002bh}
+Willis, A.P. \& Barenghi, C.F. 2002, JFM 463, 361
+
 \end{thebibliography}

### 2016-08-29 python/widegap/run_widegap_amplitude.py

@@ -28,18 +28,18 @@ norm = True
 #Pm = 1.60000000e-04
 #Rm=3.34
 #Q=0.905
-#Pm = 1.60000000e-05
-#Rm = 3.31
-#Q = 0.903
+Pm = 1.60000000e-05
+Rm = 3.31
+Q = 0.903
 
 # my newly derived params
 #Pm = 1.6E-4
 #Q = 0.90498707034113368
 #Rm = 3.33105476634
 
-Pm = 1.6E-5
-Rm = 3.31309067365
-Q = 0.90440118783008172
+#Pm = 1.6E-5
+#Rm = 3.31309067365
+#Q = 0.90440118783008172
 
 #Pm = 1.6E-6
 #Rm = 3.31127944845

### 2016-08-30 paper/widegap_mri_paper.tex
widegap appendix: corrected adjoint eqns

@@
 \beq
-\langle \mathbf{V^\dagger} | (\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} \rangle = \langle (\mathcal{L}^\dagger + \xi \mathcal{H}^\dagger) \mathbf{V}^\dagger | \mathbf{V} \rangle,
+\langle \mathbf{V^\dagger} | (\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} \rangle = \langle (\mathcal{L}^\dagger + \xi \widetilde{\mathcal{H}}^\dagger) \mathbf{V}^\dagger | \mathbf{V} \rangle,
 \eeq
@@
 -\frac{1}{\reye} (-\frac{3}{r^5} + \frac{3}{r^4}\partial_r - \frac{3}{r^3} \partial_r^2 + \frac{2}{r^2}\partial_r^3 + \frac{1}{r}\partial_r^4 ) & 0 & 0 & 0 \\
-0 & -\frac{1}{\reym} (\partial_r^2 - \frac{1}{r} \partial_r) & 0 & 0 \\
-0 & 0 &-\frac{1}{\reye} (\partial_r^2 + \frac{1}{r}\partial_r - \frac{1}{r^2}) & 0 \\
-0 & 0 & 0 & -\frac{1}{\reym} (\partial_r^2 - \frac{1}{r} \partial_r) \\
-\end{matrix}\right].
+0 & -\frac{1}{\reye} (\frac{1}{r}\partial_r + \partial_r^2 - \frac{1}{r^2}) & 0 & 0 \\
+0 & 0 &-\frac{1}{\reym} (\frac{3}{r}\partial_r + \partial_r^2) & 0 \\
+0 & 0 & 0 & -\frac{1}{\reym} (\frac{1}{r}\partial_r + \partial_r^2 - \frac{1}{r^2}) \\
+\end{matrix}\right],
 \eeq
 
-The adjoint boundary conditions are selected to satisfy Equation \ref{eq:inner_product_def}, and differ depending on the boundary conditions enforced on the original system.
+\beq
+\mathcal{L}_1^\dagger = \left[\begin{matrix}
+0 & \frac{1}{r^2} u_0 + \frac{1}{r}\partial_r u_0 & -1 & 0 \\
+-\frac{2}{r}u_0 & 0 & 0 & -1 \\
+\frac{2}{\beta}(\frac{1}{r^3} - \frac{1}{r^2}\partial_r - \frac{1}{r}\partial_r^2) & 0 & 0 & \frac{1}{r^2}u_0 - \frac{1}{r}\partial_r u_0 \\
+0 & -\frac{2}{\beta} & 0 & 0 \\
+\end{matrix}\right],
+\eeq
+
+\beq
+\mathcal{L}_2^\dagger = \left[\begin{matrix}
+-\frac{1}{\reye}(-\frac{2}{r^3} + \frac{2}{r^2}\partial_r + \frac{2}{r}\partial_r^2) & 0 & 0 & 0 \\
+0 & -\frac{1}{\reye} & 0 & 0 \\
+0 & 0 & -\frac{1}{\reym} & 0 \\
+0 & 0 & 0 & -\frac{1}{\reym} \\
+\end{matrix}\right],
+\eeq
+
+and $\mathcal{L}_3^\dagger = \mathcal{L}_3^\mathrm{T}$, $\mathcal{L}_4^\dagger = \mathcal{L}_4^\mathrm{T}$.
+The adjoint boundary conditions are selected to satisfy Equation \ref{eq:inner_product_def}, and differ depending on the boundary conditions enforced on the homogenous system.
 For the conducting boundary conditions we apply to the standard MRI, the adjoint equation
@@
 \beq
-\Psi^\dagger = \partial_r \Psi^\dagger = u^\dagger = A^\dagger = \partial_r B^\dagger = 0.
+\Psi^\dagger = \partial_r \Psi^\dagger = u^\dagger = A^\dagger = \partial_r (r B^\dagger) = 0.
 \eeq
 
-Note that the condition on $B^\dagger$ differs from that on $B$.
-For the insulating boundary conditions we apply to the helical MRI, the magnetic boundary conditions applied to Equation \ref{eq:adjoint} are
-
-\beq
-B^\dagger = (- k \frac{I_0 (k r)}{I_1 (k r)} + \frac{1}{r} + \partial_r ) A^\dagger = 0 \, \, \mathrm{at} \, \, r = r_1
-\eeq
-
-\beq
-B^\dagger = (k \frac{K_0 (k r)}{K_1 (k r)} + \frac{1}{r} + \partial_r ) A^\dagger = 0 \, \, \mathrm{at} \, \, r = r_2
-\eeq
-
-We take the inner product of the adjoint homogenous solution with Equation \ref{eq:ordere3} to find
+We similarly define appropriate boundary conditions for the insulating case. We take the inner product of the adjoint homogenous solution with Equation \ref{eq:ordere3} to find
 

### 2016-08-30 paper/widegap_mri_paper.tex
new parameter space figure in paper

@@
 
+%\begin{widetext}
+\begin{table*}[ht]
+\normalsize
+\caption{Fiducial parameters for MRI runs} \label{table:parameters}
+\centering
+\begin{tabular}{rlrrrrrrr}
+  \hline
+ & $\xi$ & $Pm$ & $\beta$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ & radial magnetic b.c. \\ 
+  \hline\hline
+Standard MRI & 0 & 1.6E-6 & 41.2 & 0.121 & 0.33 & conducting \\ 
+Helical MRI & 4 & 1E-6 & 1.7E-2 & 0.27 & 0.5 & insulating\\ 
+   \hline
+\end{tabular}
+\end{table*} 
+%\end{widetext}
+
 Although the MRI is broadly important to many astrophysical systems, many of its general properties, and in particular its nonlinear saturation mechanism, remain poorly understood.
@@
 \centering
-\includegraphics[width=\columnwidth]{../figures/widegap_paramspace_crit_params_Pm_1E-3}
+\includegraphics[width=\columnwidth]{../figures/widegap_paramspace_crit_params_Pm_16E-7}
 \caption{Growth rates in the ($\reym, k_z$) plane.
@@
 The eigenvalue problem was solved for the widegap parameters listed in Table \ref{table:parameters}.
-Overlaid contours show growth rates at [-2E-3, -1E-3, -1E-5, 1E-5, 1E-3, 2E-3], where dashed contours represent negative values.
+Overlaid contours show growth rates at [ -8E-4, -1.3E-4, 1.3E-4, 8E-4, 1.5E-3], where dashed contours represent negative values.
 The gray dotted line shows the interpolated marginal stability curve.
-The critical parameters $\reym_c = 0.84$ and $k_c = 0.0127$ correspond to the smallest parameter values that yield a zero growth rate.}\label{fig:growth_rates}
+The critical parameters $\reym_c = 3.31$ and $k_c = 0.902$ correspond to the smallest parameter values that yield a zero growth rate.}\label{fig:growth_rates}
 \end{figure}
@@
 These definitions automatically satisfy Equations \ref{eq:incompressibility} and \ref{eq:solenoid} for axisymmetric disturbances.
-We note that in the linearized equations, streamfunctions of the form $u_x = \partial_z \Psi$, $u_z = -(\partial_r + \frac{1}{r}) \Psi$, and the corresponding definitions of the magnetic vector potential, are convenient choices, but this does not hold true for the nonlinear terms because of the incommutability of $\partial_r$ and $\partial_r + \frac{1}{r}$.
+We note that in the linearized equations, streamfunctions of the form $u_x = \partial_z \Psi$, $u_z = -(\partial_r + \frac{1}{r}) \Psi$, and the corresponding definitions of the magnetic vector potential, are convenient choices, but we make a different choice for this nonlinear investigation because of the incommutability of $\partial_r$ and $\partial_r + \frac{1}{r}$.
 
@@
 
-Thus through judicious choice of cylinder rotation rates, we can set $q(r_0) = 3/2$, for quasi-Keplerian flow.
-Note that the narrow gap approximation imposes a linear shear (constant $q$).
-Our base velocity is $\mathbf{u_0} = r\Omega(r) \phihat$.
-We initialize a magnetic field $\mathbf{B_0} = B_0 \zhat + B_0 \xi \frac{r_0}{r} \phihat$, so that the base magnetic field is axial when $\xi = 0$ and otherwise helical.
+Thus through judicious choice of cylinder rotation rates, one can set $q(r_0) = 3/2$, for quasi-Keplerian flow.
+Note that the narrow gap approximation imposes a linear shear (constant $q$), and thus the interaction of fluid perturbations with the base velocity profile differs significantly from the case considered here.
+Our base velocity is
+
+\beq
+\label{eq:baseu0}
+\mathbf{u_0} = r\Omega(r) \phihat.
+\eeq
+
+We initialize a magnetic field
+
+\beq
+\label{eq:baseB0}
+\mathbf{B_0} = B_0 \zhat + B_0 \xi \frac{r_0}{r} \phihat,
+\eeq
+
+so that the base magnetic field is axial when $\xi = 0$ and otherwise helical.
 
 In this work we will focus our findings on two fiducial parameter sets, one for the standard MRI (hereafter SMRI) where $\xi = 0$ and one for the helical MRI (hereafter HMRI).
-We choose the SMRI parameters to be comparable to the case explored in \citei{Umurhan:2007hs}, with the geometric parameters of \citei{Goodman:2002ix}.
+We choose the SMRI parameters to be comparable to the case considered in \citei{Goodman:2002ix}.
 The HMRI parameters were chosen to be comparable to \citei{Hollerbach:2005tr}.
@@
 \begin{widetext}
-\begin{table*}[ht]
-\normalsize
-\caption{Fiducial parameters for MRI runs} \label{table:parameters}
-\centering
-%\resizebox{\textwidth}{!}
-{\begin{tabular}{rlrrrrrrr}
-  \hline
- & $\xi$ & $Pm$ & $\beta$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ & radial magnetic b.c. \\ 
-  \hline\hline
-%Standard MRI & 0 & 1E-3 & 25 & 0.18 & 0.33 & conducting \\ 
-Standard MRI & 0 & 1.6E-6 & 41.2 & 0.121 & 0.33 & conducting \\ 
-Helical MRI & 4 & 1E-6 & 1.7E-2 & 0.27 & 0.5 & insulating\\ 
-   \hline
-\end{tabular}}
-\end{table*}
-
 Our perturbed system is
@@
 
-and $B = 0$ at $r = r_1, r_2$ (see \citei{Willis:2002bh}). Here, $I_n$, $K_n$ are the modified Bessel functions of the first and second kind, respectively.
+and $B = 0$ at $r = r_1, r_2$ (see \citei{Willis:2002bh}). Here, $I_n$ and $K_n$ are the modified Bessel functions of the first and second kind, respectively.
 
@@
 
+We find the marginal system as a function of the dimensionless parameters.
+Marginality for our standard MRI system is a hyperplane in $(\reym, Pm, \beta, \Omega_2/\Omega_1, \mathrm{R}_1/\mathrm{R}_2$), but we hold all of these constant except for $\reym$.
+To analyze the MRI system at marginality, we fix the parameters listed in Table \ref{table:parameters} and determine the critical $\reym$ and $k_z$ by repeatedly solving the linear MRI system to determine the smallest parameter values for which the fastest growing mode is zero.
+That is, we solve the linear eigenvalue problem for eigenvalues $\sigma = \gamma + i \omega$.
+Figure \ref{fig:growth_rates} shows linear MRI growth rates $\gamma$ in the $(\reym, k_z)$ plane.
+For the fiducial standard MRI parameters in Table \ref{table:parameters} we find critical parameters $\reym_c = 3.30$ and $k_c = 0.901$.
+
 Just as in the weakly nonlinear analyses of \citei{Umurhan:2007hs} and \citei{Clark:2016}, we tune the system away from marginality by taking $B_0 \rightarrow B_0\left(1 - \epsilon^2\right)$, where $\epsilon \ll 1$.
 We parameterize scale separation as $Z = \epsilon z$ and $T = \epsilon^2 t$, where $Z$ and $T$ are slowly varying spatial and temporal scales, respectively.
-We group the fluid variables into a state vector $\mathbf{V} = \left[\Psi, u, A, B\right]^{\mathrm{T}}$, such that the full nonlinear system can be expressed as
+We group the fluid variables into a state vector $\mathbf{V} = \left[\Psi, u, A, B\right]^{\mathrm{T}}$, such that the full nonlinear system in Equations \ref{eq:Psi_perturbed} - \ref{eq:nonlinear_B} can be expressed as
 
@@
 where $\mathcal{D}$, $\mathcal{L}$, and $\widetilde{\mathcal{G}}$ are matrices defined in Appendix \ref{app:basic_equations}, and $\mathbf{N}$ is a vector containing all nonlinear terms defined in Appendix \ref{app:nonlinear_terms}.
-We then expand the variables in a perturbation series $\mathbf{V} = \epsilon \mathbf{V}_1 + \epsilon^2 \mathbf{V}_2 + \epsilon^3 \mathbf{V}_3 + h.o.t.$ The perturbed system can then be expressed at each order by the equations
+We expand the variables in a perturbation series $\mathbf{V} = \epsilon \mathbf{V}_1 + \epsilon^2 \mathbf{V}_2 + \epsilon^3 \mathbf{V}_3 + h.o.t.$ The perturbed system can then be expressed at each order by the equations
 
@@
 See Appendix \ref{app:basic_equations} for the definition of matrices and nonlinear vectors, and a thorough derivation.
-We emphasize that Equations \ref{eq:ordere} - \ref{eq:ordere3} have the same form as these equations in the narrow gap case, although the matrices (which contain all radial derivatives) are significantly different in this wide gap formulation.
+We emphasize that Equations \ref{eq:ordere} - \ref{eq:ordere3} have the same form as these equations in the narrow gap case, although the matrices, which contain all radial derivatives, are significantly different in this wide gap formulation.
 This is because we do not have slow variation in the radial dimension.
@@
 in prep).
-Because our domain is bounded only radially and not vertically, we solve the radial dimension on a basis of Chebyshev polynomials, and the vertical dimension on a Fourier series basis.
-We solve Equation \ref{eq:ordere} as a linear eigenvalue problem, and Equations \ref{eq:ordere2} and \ref{eq:ordere3} as linear boundary value problems.
-%should i mention spurious eigenvals?
-To analyze the MRI system at marginality, we fix the parameters listed in Table \ref{table:parameters} and determine the critical $\reym$ and $k_z$ by repeatedly solving Equation \ref{eq:ordere} to determine the smallest parameter values for which the fastest growing mode is zero.
+We solve the radial portion of the eigenvectors on a basis of Chebyshev polynomials subject to our radial boundary conditions.
+%Because our domain is bounded only radially and not vertically, we solve the radial dimension on a basis of Chebyshev polynomials, and the vertical dimension on a Fourier series basis.
+We solve Equation \ref{eq:ordere} as a linear eigenvalue problem, and Equation \ref{eq:ordere2} as a linear boundary value problem. %should i mention spurious eigenvals? 
+%To analyze the MRI system at marginality, we fix the parameters listed in Table \ref{table:parameters} and determine the critical $\reym$ and $k_z$ by repeatedly solving Equation \ref{eq:ordere} to determine the smallest parameter values for which the fastest growing mode is zero.
 That is, we solve the linear eigenvalue problem for eigenvalues $\sigma = \gamma + i \omega$.
@@
 
-We similarly define appropriate boundary conditions for the insulating case. We take the inner product of the adjoint homogenous solution with Equation \ref{eq:ordere3} to find
+We similarly define appropriate boundary conditions for the insulating case.
+We take the inner product of the adjoint homogenous solution with the terms in Equation \ref{eq:ordere3} that are resonant with $e^{i k_c z}$.
+This gives us
 
 \beq
-\mathcal{L}\mathbf{V}_3 + \xi \widetilde{\mathcal{H}} \mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2\nonumber 
-+ \xi \mathcal{H}\partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 - \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3 = 0.
+\langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle \partial_T \alpha + \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} - \xi \widetilde{\mathcal{H}} \mathbb{V}_{11} \rangle \alpha +  \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} + \xi \mathcal{H} \mathbb{V}_{21} \rangle \partial_Z^2 \alpha = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle \alpha |\alpha|^2,
+\eeq
+
+or Equation \ref{eq:gle}, the Ginzburg-Landau Equation, where the coefficients are
+
+\beq
+b = \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} - \xi \widetilde{\mathcal{H}} \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
+\eeq
+
+\beq
+h = \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} + \xi \mathcal{H} \mathbb{V}_{21} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
+\eeq
+
+and
+
+\beq
+c = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle. 
 \eeq

### 2016-08-30 paper/widegap_mri_paper.tex

@@
 This is a multiscale perturbative treatment of the nonideal, axisymmetric MRI near threshold, subject to realistic radial boundary conditions.
-We analyze both the standard MRI, initialized by a constant vertical background magnetic field, and the helical MRI, in which the base magnetic field contains an azimuthal component.
+We analyze both the standard MRI, initialized by a constant vertical background magnetic field, and the helical MRI, in which the background magnetic field contains an azimuthal component.
+This is the first weakly nonlinear analysis of the MRI in a Taylor Couette geometry, as well as the first weakly nonlinear analysis of the helical MRI.
 We find that the evolution of the weakly nonlinear perturbation amplitude of the standard MRI is described by a real Ginzburg-Landau equation (GLE), while the amplitude of the helical MRI takes the form of a complex GLE.
+This suggests that the saturated state of the helical MRI may itself be unstable on long spatial and temporal scales.
 \end{abstract}
@@
 %The ubiquity of disks ...
-The ubiquity of astrophysical disks suggests that systems both experience and overcome this centrifugal barrier, and accretion proceeds even in hydrodynamically stable disks, where molecular viscosity is vastly insufficient to drive angular momentum transport (\citei{Shakura:1973wg}).
+The ubiquity of astrophysical accretion disks suggests that a wide variety of systems both experience and overcome this centrifugal barrier.
+Furthermore, accretion proceeds even in hydrodynamically stable disks, where molecular viscosity is vastly insufficient to drive angular momentum transport (\citei{Shakura:1973wg}).
 The MRI is excited by weak magnetic fields in differentially rotating fluids, and since its discovery it has been widely invoked to explain accretion in protoplanetary disks (see \citei{Armitage:2010} and references therein), binary systems (), and disks around black holes (), as well as jet and wind launching (\citei{Lesur:2013}), dynamos, etc (cite these)
@@
 Most relevant to our work is the Princeton Plasma Physics Laboratory (PPPL) MRI Experiment, a liquid gallium Taylor-Couette flow exposed to an axial magnetic field (\citei{Ji:2001kd}).
-Some theoretical work designed to complement the Princeton MRI experiment involves direct simulation of the experimental conditions, much of it focused on the specific challenges in identifying MRI signatures despite apparatus-driven spurious flows (e.g.
+There has been some theoretical work designed to complement the Princeton MRI experiment involving direct simulation of the experimental conditions, much of it focused on the specific challenges in identifying MRI signatures despite apparatus-driven spurious flows (e.g.
 \citei{Gissinger:2012gc}).
-In particular, the vertical endcaps on a laboratory MRI apparatus drives meridional flows which both inhibit the excitement of MRI and obscure its detection.
+In particular, the vertical endcaps on a laboratory MRI apparatus drive meridional flows which both inhibit the excitement of MRI and obscure its detection.
 The Princeton MRI experiment employs split, independently rotating endcaps to mitigate these flows (\citei{Schartman:2009df}).
@@
 
+This realistic radial treatment means that we account for the curvature of the flow in a cylindrical apparatus.
 Many investigations of the MRI use the ``narrow gap" approximation, in which the radial extent of the fluid channel is taken to be much smaller than the radius of curvature.
-That is, for a channel center $r_0$ bounded by inner and outer radii $r_1$ and $r_2$, respectively, the narrow gap approximation applies when $r_0 \gg (r_2 - r_1)$.
+That is, for a center channel radius $r_0$ bounded by inner and outer radii $r_1$ and $r_2$, respectively, the narrow gap approximation applies when $r_0 \gg (r_2 - r_1)$.
 The narrow gap approximation simplifies the MRI equations by excluding curvature terms, because the flow through a narrow gap can be taken to be approximately linear in $\phi$, i.e.
 Cartesian.
-Previous investigations into the weakly nonlinear MRI have used this narrow gap approximation (\citei{Umurhan:2007dz}, \citei{Umurhan:2007hs}, \citei{Clark:2016}).
+Previous investigations into the weakly nonlinear behavior of the MRI have used this narrow gap approximation (\citei{Umurhan:2007dz}, \citei{Umurhan:2007hs}, \citei{Clark:2016}).
 In this work we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.
@@
 In this work we explore the behavior of the viscous, dissipative MRI in a cylindrical geometry close to threshold, making explicit comparisons to the standard MRI behavior in the thin-gap regime.
+We investigate both the standard MRI, in which the background magnetic field is purely axial, as well as the helical MRI, in which the background magnetic field has an azimuthal component as well.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{../figures/widegap_3D_vel_screenshot.png}
-\caption{Visualization of the velocity perturbation field in the fiducial SMRI case.
-The three dimensional path of twenty randomly placed tracer particles is plotted, where the particle velocity is defined as the total (first + second order) velocity perturbation.
-This is meant to be a visualization only of the velocity field, so the particle motion feels no contribution from the magnetic field.}
-\end{figure}
+%\begin{figure}
+%\centering
+%\includegraphics[width=\columnwidth]{../figures/widegap_3D_vel_screenshot.png}
+%\caption{Visualization of the velocity perturbation field in the fiducial SMRI case.
+The three dimensional path of t%wenty randomly placed tracer particles is plotted, where the particle velocity is defined as the total (first + second %order) velocity perturbation.
+This is meant to be a visualization only of the velocity field, so the particle motion f%eels no contribution from the magnetic field.}
+%\end{figure}
 
@@
 %Equation \ref{eq:ordere3} is resonant with 
-To find a bounded solution at $\mathcal{O}(\epsilon^3)$ we must eliminate secular terms: terms which are resonant with the solution to the homogenous equation $(\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} = 0$ and cause the solution to grow without bound.
+To find a bounded solution at $\mathcal{O}(\epsilon^3)$ we must eliminate secular terms: terms which are resonant with the solution to the linear homogenous equation $(\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} = 0$ and cause the solution to grow without bound.
 To eliminate these terms we enforce a solvability condition, which arises as a consequence of the Fredholm alternative.

### 2016-08-30 python/widegap/run_widegap_amplitude.py
Pm=1.6E-6 at 128

@@ -38,13 +38,13 @@ norm = True
 #Rm = 3.33371
 #Q = 0.905152
 
-Pm = 1.6e-5
-Rm = 3.31292
-Q = 0.903172
+#Pm = 1.6e-5
+#Rm = 3.31292
+#Q = 0.903172
 
-#Pm = 1.6E-6
-#Rm = 3.30878
-#Q = 0.902139
+Pm = 1.6E-6
+Rm = 3.30878
+Q = 0.902139
 
 # ********************************************************
 

### 2016-08-30 paper/widegap_mri_paper.tex
Pm=1.6E-6 at 128
hmri run at 128

@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../python/widegap/figures/orders_1_2_widegap_fiducial_eigenfunctions.pdf}
+\caption{Eigenfunctions of the first order equations, first order adjoint homogenous equations, and second order equations.
+We use our fiducial parameters for the standard MRI ($\xi = 0$).
+First-order eigenfunctions are normalized such that they are either purely real or purely imaginary, and such that $\int \Psi_{11} \mathrm{d}r = 1$.
+Adjoint homogenous eigenfunctions are normalized such that $\left< V_{11}^\dagger \cdot \mathcal{D} V_{11} \right> = 1$.}\label{fig:linear_eigenfunctions}
+\end{figure*}
+
 \section{Wide Gap Equations}
@@
 
-and define the magnetic vector potential $A$ analogously.
+and the magnetic vector potential $A$ is
+
+\beq
+  \label{eq:fluxfunc}
+  \mathbf{B_1} \, = \, \left[\begin{matrix}
+\frac{1}{r} \partial_z A\ \rhat\\
+B_{\phi} \ \phihat\\
+-\frac{1}{r} \partial_r A\ \zhat\\
+\end{matrix}\right].
+\eeq
+
+%and define the magnetic vector potential $A$ analogously. 
 These definitions automatically satisfy Equations \ref{eq:incompressibility} and \ref{eq:solenoid} for axisymmetric disturbances.
-We note that in the linearized equations, streamfunctions of the form $u_x = \partial_z \Psi$, $u_z = -(\partial_r + \frac{1}{r}) \Psi$, and the corresponding definitions of the magnetic vector potential, are convenient choices, but we make a different choice for this nonlinear investigation because of the incommutability of $\partial_r$ and $\partial_r + \frac{1}{r}$.
+We note that in the linearized equations, streamfunctions of the form $u_x = \partial_z \Psi$, $u_z = -(\partial_r + \frac{1}{r}) \Psi$, and the corresponding definitions of the magnetic vector potential, are convenient choices, but we define Equations \ref{eq:stokes} and \ref{eq:fluxfunc} for this nonlinear investigation because of the incommutability of $\partial_r$ and $\partial_r + \frac{1}{r}$.
 
@@
 \end{equation}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{../python/widegap/figures/orders_1_2_widegap_fiducial_eigenfunctions.pdf}
-\caption{Eigenfunctions of the first order equations, first order adjoint homogenous equations, and second order equations.
-We use our fiducial parameters for the standard MRI ($\xi = 0$).
-First-order eigenfunctions are normalized such that they are either purely real or purely imaginary, and such that $\int \Psi_{11} \mathrm{d}r = 1$.
-Adjoint homogenous eigenfunctions are normalized such that $\left< V_{11}^\dagger \cdot \mathcal{D} V_{11} \right> = 1$.}\label{fig:linear_eigenfunctions}
-\end{figure*}
 
@@
 
-In this work we will focus our findings on two fiducial parameter sets, one for the standard MRI (hereafter SMRI) where $\xi = 0$ and one for the helical MRI (hereafter HMRI).
+In this work we will focus our findings on two fiducial parameter sets, one for the standard MRI where $\xi = 0$ and one for the helical MRI.
 We choose the SMRI parameters to be comparable to the case considered in \citei{Goodman:2002ix}.
@@
 We introduce the Reynolds number $\reye = \Omega_1 r_1^2/\nu$, the magnetic Reynolds number $\reym = \Omega_1 r_1^2 / \eta$, and a plasma beta parameter $\beta = \Omega_1^2 r_1^2 \rho_0/B_0^2$.
-If we define the dimensional cylindrical coordinate $r = r_1(1 + \delta x)$, the narrow gap equations are recovered in the limit $\delta \rightarrow 0$.
+Note that if we define the dimensional cylindrical coordinate $r = r_1(1 + \delta x)$, we recover the narrow gap approximation of the system in the limit $\delta \rightarrow 0$.
 
@@
 
-\section{Results and Discussion}
+\section{Results}
 
@@
 With the introduction of an azimuthal component, the background magnetic field acquires a handedness that is not present in a purely axial field.
+
+\section{Discussion}
+
+Our work should be placed in the broader context of emergent pattern formation in physical systems.
 

### 2016-08-30 python/widegap/run_hmri_amplitude.py

@@ -31,7 +31,7 @@ aa.print_coeffs()
 
 #fn_root = "/Users/susanclark/weakly_nonlinear_mri/data/"
 fn_root = "../../data/"
-fn = fn_root + "widegap_amplitude_parameters_Q_{:03.2f}_Rm_{:04.4f}_Pm_{:.2e}_Omega1_{:05.2f}_Omega2_{:05.2f}_beta_{:.2f}_xi_{:.2f}_gridnum_{}_conducting_{}_norm_{}.h5".format(Q, Rm, Pm, Omega1, Omega2, beta, xi, gridnum, conducting, norm)
+fn = fn_root + "hmri_amplitude_parameters_Q_{:03.2f}_Rm_{:04.4f}_Pm_{:.2e}_Omega1_{:05.2f}_Omega2_{:05.2f}_beta_{:.2f}_xi_{:.2f}_gridnum_{}_conducting_{}_norm_{}.h5".format(Q, Rm, Pm, Omega1, Omega2, beta, xi, gridnum, conducting, norm)
 with h5py.File(fn,'w') as f:
     r = f.create_dataset("r", data=aa.r)
     o1psi = f.create_dataset("psi11", data=aa.o1.psi['g'])

### 2016-08-31 paper/widegap_mri_paper.tex
appendix fleshed out

@@
 The slow variation in $Z$ and $T$ are parameterized as an amplitude function $\alpha(Z, T)$ which modulates the flow in these dimensions.
-This parameterization coupled with the boundary conditions lead us to an ansatz linear solution $\mathbf{V}_1 = \alpha(Z, T) \mathbb{V}_{11}(r) e^{i k_z z} + c.c.$, where the radial variation is contained in $\mathbb{V}_{11}$.
+This parameterization coupled with the boundary conditions lead us to an ansatz linear solution
+
+\beq
+\label{eq:ansatz}
+\mathbf{V}_1 = \alpha(Z, T) \mathbb{V}_{11}(r) e^{i k_z z} + c.c.,
+\eeq
+
+where the radial variation is contained in $\mathbb{V}_{11}$.
 
@@
 To find a bounded solution at $\mathcal{O}(\epsilon^3)$ we must eliminate secular terms: terms which are resonant with the solution to the linear homogenous equation $(\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} = 0$ and cause the solution to grow without bound.
-To eliminate these terms we enforce a solvability condition, which arises as a consequence of the Fredholm alternative.
-This solvability condition is the requirement that the inner product of the equation with its adjoint homogenous solution is zero.
+Secular terms in our system are those that are resonant with the linear ansatz (Equation \ref{eq:ansatz}), i.e.
+terms with $e^{i k_c z}$ z-dependence.
+To eliminate these terms we enforce a solvability condition, which arises from a corollary to the Fredholm alternative.
+The Fredholm alternative states that if we consider a system of equations $\mathcal{L} \mathbf{V} = \mathbf{b}$ and its adjoint homogenous system $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$, only one of two conditions holds.
+Either there exists one and only one solution to the inhomogenous system, or the homogenous adjoint equation has a nontrivial solution.
+The relevant corollary arises as a consequence of the second condition: if $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$ has a nontrivial solution, then $\mathcal{L} \mathbf{V} = \mathbf{b}$ has a solution if and only if $\langle \mathbf{V}^\dagger | b \rangle = 0$.
+
+% This solvability condition is the requirement that the inner product of the equation with its adjoint homogenous solution is zero. 
 We define the adjoint operator $\mathcal{L}^\dagger$ and solution $\mathbf{V}^\dagger$ as
@@
 
-We similarly define appropriate boundary conditions for the insulating case.
-We take the inner product of the adjoint homogenous solution with the terms in Equation \ref{eq:ordere3} that are resonant with $e^{i k_c z}$.
-This gives us
+For the insulating case, the adjoint boundary conditions are
+
+\beq
+k \frac{I_0 (k r)}{I_1 (k r)} r A^\dagger - 2 A^\dagger - r \partial_r A^\dagger = 0 \, \, \mathrm{at} \, \,   r = r_1
+\eeq
+
+\beq
+- k \frac{K_0 (k r)}{K_1 (k r)} r A^\dagger - 2 A^\dagger - r \partial_r A^\dagger = 0 \, \, \mathrm{at} \, \,   r = r_2
+\eeq
+
+We take the inner product of the adjoint homogenous solution with the terms in Equation \ref{eq:ordere3} that are resonant with $e^{i k_c z}$. This gives us
 

### 2016-08-31 paper/widegap_mri_paper.tex
added some discussion

@@
 Our work should be placed in the broader context of emergent pattern formation in physical systems.
+The real Ginzburg-Landau equation derived here governs the slow-parameter evolution of the standard MRI close to threshold.
+The GLE arises in a number of other physical systems, and in each case it is a consequence not of the particular physics at hand, but of the underlying symmetries in the problem.
+Here we make a brief comparison to two other systems that give rise to a GLE.
+The first and perhaps most famous is Rayleigh-B\'enard convection, in which a fluid between two plates is heated from below.
+If we take the plane of the fluid to be infinite in the horizontal plane, the system is initially translationally symmetric.
+%When the system eventually becomes unstable to convective heat transfer it undergoes a symmetry breaking, allowing the
+At the onset of convection the system undergoes a symmetry breaking, forming rolls, or convection cells, which break the horizontal translational invariance.
+Analogously, the standard MRI system considered here is initially vertically translationally symmetric, because we idealize the Taylor-Couette device as an infinitely long cylinder.
+The MRI breaks this symmetry, forming cells along the vertical length of the domain.
+Just as Rayleigh-B\'enard cells transport heat vertically, the MRI cells transport angular momentum horizontally.
+The symmetry breaking of each of these systems is described near onset by the real GLE.
+
+A real GLE has also been found to describe the formation of zonal flows out of magnetized turbulence in a model system ().
+Zonal flows are axisymmetric structures, large-scale and long-lived, which form spontaneously out of turbulence.
+They have recently been observed in some numerical studies of the MRI, and have generated considerable interest for their possible role in planet formation in protoplanetary disks ().
+The present work is of course an idealized geometry, and we make no attempt to model a realistic protoplanetary disk environment.
+However, it is worth noting that the GLE we derive implies that axisymmetric, large-scale, long-lived structures are a generic feature of the MRI in the weakly nonlinear regime.
+This work provides a mathematical description of the MRI as a pattern-forming process, but much remains to be understood, particularly involving the relevance of these insights to realistic astrophysical disks.
+Our model is most directly relevant to Taylor-Couette flows, and we emphasize that laboratory MRI experiments stand poised to observe the MRI-driven pattern formation predicted here.
 
@@
 \eeq
+
+We solve the $\mathcal{O}(\epsilon)$ (linear) system, followed by the $\mathcal{O}(\epsilon^2)$ system in Equation \ref{eq:ordere2}.
+At second order in $\epsilon$, nonlinear terms arise which are formed by the interaction of first-order MRI modes with themselves and their complex conjugates.
+This mode interaction means that the second-order nonlinear term is
+
+\beq
+\mathbf{N}_2 = |\alpha|^2 \mathbf{N}_{20} + \alpha^2 \mathbf{N}_{22} e^{2 i k_c z},
+\eeq
+
+where terms are grouped by z-dependence.
+See Appendix \ref{app:nonlinear_terms} for the full form of the nonlinear terms.
+Equation \ref{eq:ordere2} must therefore be solved as three separate systems of equations, one for each possible $z$ resonance:
 

### 2016-08-31 paper/widegap_mri_paper.tex
begin long run at Pm=1.6E-5

@@
 a Ginzburg-Landau equation (GLE).
-The GLE governs the weakly nonlinear amplitude behavior in a wide range of physical systems, including the narrow gap MRI (\citei{Umurhan:2007hs}), Rayleigh-B\'enard convection, and hydrodynamic Taylor Couette flow.
+The GLE governs the weakly nonlinear amplitude behavior in a wide range of physical systems, including the narrow gap MRI (\citei{Umurhan:2007hs}), Rayleigh-B\'enard convection (\citei{Newell:1969wr}), and hydrodynamic Taylor Couette flow (e.g.
+\citei{Recktenwald:1993}).
 We emphasize that this is a model equation, valid only near marginality (e.g.
@@
 Equation \ref{eq:gle} contains three coefficients: $b$, which determines the linear growth rate of the system, $d$, a diffusion coefficient, and $c$, the coefficient of the nonlinear term.
-When all of the coefficients of Equation \ref{eq:gle} are real, this is known as the real GLE, although the amplitude $\alpha$ remains complex.
+When all of the coefficients of Equation \ref{eq:gle} are real, this is known as the real GLE, although the amplitude $\alpha$ is in general complex.
 The real GLE is subject to several well-studied instabilities, including the Ekhaus and Zig-Zag instabilities.
-When the coefficients are complex, we have the complex GLE, a source of even richer phase dynamics than its real counterpart (e.g.
-\citei{Aranson:2002}).
+When the coefficients are complex, we have the complex GLE, a source of even richer phase dynamics than its real counterpart (see \citei{Aranson:2002} for a thorough review).
 
@@
 The preservation of symmetries in the thin-gap geometry is worth a closer look, as its absence in the wide gap case is the source of many differences in the systems.
-\citei{Latter:2015} point out that in the ideal limit ($\nu, \eta \rightarrow 0$), the linearized system described by the lefthand side of Equations \ref{eq:Psi_perturbed} - \ref{eq:Bphi_perturbed} can be expressed as a Shr{\"o}dinger equation for the radial velocity.
+\citei{Latter:2015} point out that in the ideal limit ($\nu, \eta \rightarrow 0$), the linearized system described by the lefthand side of Equations \ref{eq:Psi_perturbed} - \ref{eq:Bphi_perturbed} can be expressed as a Schr{\"o}dinger equation for the radial velocity.
 Similarly combining equations to obtain a single expression for $\Psi$, we find that the thin-gap limit linear ideal MRI can be expressed as
@@
 Here we make a brief comparison to two other systems that give rise to a GLE.
-The first and perhaps most famous is Rayleigh-B\'enard convection, in which a fluid between two plates is heated from below.
+The first and perhaps most famous is Rayleigh-B\'enard convection, in which a fluid between two plates is heated from below (\citei{Newell:1969wr}).
 If we take the plane of the fluid to be infinite in the horizontal plane, the system is initially translationally symmetric.
@@
 
-A real GLE has also been found to describe the formation of zonal flows out of magnetized turbulence in a model system ().
+A real GLE has also been found to describe the formation of zonal flows out of magnetized turbulence in a model system (\citei{Parker:2013hy}).
 Zonal flows are axisymmetric structures, large-scale and long-lived, which form spontaneously out of turbulence.
-They have recently been observed in some numerical studies of the MRI, and have generated considerable interest for their possible role in planet formation in protoplanetary disks ().
+They have recently been observed in some numerical studies of the MRI, and have generated considerable interest for their possible role in planet formation in protoplanetary disks (\citei{Johansen:2009uj}, \citei{Kunz:2013}).
 The present work is of course an idealized geometry, and we make no attempt to model a realistic protoplanetary disk environment.
@@
 Our model is most directly relevant to Taylor-Couette flows, and we emphasize that laboratory MRI experiments stand poised to observe the MRI-driven pattern formation predicted here.
+
+\section{Acknowledgements}
+NASA, NSF. The authors thank Mordecai Mac Low, Jeremy Goodman, etc for useful discussion.
 
@@
 
+\bibitem[{Johansen et al.(2009)}]{Johansen:2009uj}
+Johansen, A., Youdin, A., Klahr, H., 2009, ApJ 697, 1269
+
 \bibitem[{Kirillov \& Stefani(2013)}]{Kirillov:2013}
@@
 Knobloch, E, Julien K, 2005, Physics of Fluids, 17, 094106
+
+\bibitem[{Kunz \& Lesur(2013)}]{Kunz:2013}
+Kunz, M.W. \& Lesur, G., 2013, MNRAS 434, 2295
 
@@
 
-\bibitem[Pessah(2010)]{Pessah:2010ic}
+\bibitem[{Newell \& Whitehead(1969)}]{Newell:1969wr}
+Newell, A.C. \& Whitehead, J.A., 1969, JFM 38, 279
+
+\bibitem[{Parker \& Krommes(2013)}]{Parker:2013hy}
+Parker, J.B. \& Krommes, J.A., 2013, Phys of Plasmas 20, 100703
+
+\bibitem[{Pessah(2010)}]{Pessah:2010ic}
 Pessah, M, 2010, ApJ, 716, 1012
+
+\bibitem[{Recktenwald et al.(1993)}]{Recktenwald:1993}
+Recktenwald, A., L\"ucke, M., M\"uller, H.W., 1993, Phys Rev E 48, 4444
 

### 2016-08-31 python/widegap/pleiades_run_widegap_amplitude.sh

@@ -1,5 +1,5 @@
 #PBS -S /bin/bash
-#PBS -N WidegapFinalLong
+#PBS -N WidegapFinalLong5
 #PBS -l select=2:ncpus=10:mpiprocs=10:model=has
 #PBS -l walltime=20:00:00
 #PBS -j oe

### 2016-08-31 python/widegap/run_widegap_amplitude.py

@@ -38,13 +38,13 @@ norm = True
 #Rm = 3.33371
 #Q = 0.905152
 
-#Pm = 1.6e-5
-#Rm = 3.31292
-#Q = 0.903172
+Pm = 1.6e-5
+Rm = 3.31292
+Q = 0.903172
 
-Pm = 1.6E-6
-Rm = 3.30878
-Q = 0.902139
+#Pm = 1.6E-6
+#Rm = 3.30878
+#Q = 0.902139
 
 # ********************************************************
 

### 2016-08-31 paper/widegap_mri_paper.tex
more appendix detail

@@
 The GLE arises in a number of other physical systems, and in each case it is a consequence not of the particular physics at hand, but of the underlying symmetries in the problem.
-Here we make a brief comparison to two other systems that give rise to a GLE.
+Here we make a phenomenological comparison to two other systems that give rise to a GLE.
 The first and perhaps most famous is Rayleigh-B\'enard convection, in which a fluid between two plates is heated from below (\citei{Newell:1969wr}).
@@
 \section{Acknowledgements}
-NASA, NSF. The authors thank Mordecai Mac Low, Jeremy Goodman, etc for useful discussion.
+NASA, NSF. The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, etc for useful discussion.
 
@@
 
+\begin{align}
+\mathcal{L}\mathbf{V}_{20} + \xi \partial_z \mathcal{H} \mathbf{V}_{20} & = \mathbf{N}_{20}\\
+\mathcal{L}\mathbf{V}_{21} + \xi \partial_z \mathcal{H} \mathbf{V}_{21} & = - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_{11} - \xi \partial_Z \mathcal{H} \mathbf{V}_{11} \\
+\mathcal{L}\mathbf{V}_{22} + \xi \partial_z \mathcal{H} \mathbf{V}_{22} & = \mathbf{N}_{22}
+\end{align}
+
 %The GLE is derived from 
@@
 The adjoint boundary conditions are selected to satisfy Equation \ref{eq:inner_product_def}, and differ depending on the boundary conditions enforced on the homogenous system.
+Specifically, the boundary conditions arise from the requirement that the integrands in Equation \ref{eq:inner_product_def} are zero at $r_1$ and $r_2$.
 For the conducting boundary conditions we apply to the standard MRI, the adjoint equation

### 2016-08-31 paper/widegap_mri_paper.tex
store growth rate and frequency in linear eigenvalue and adjoint objects, and in run output

@@
 However, it is worth noting that the GLE we derive implies that axisymmetric, large-scale, long-lived structures are a generic feature of the MRI in the weakly nonlinear regime.
-This work provides a mathematical description of the MRI as a pattern-forming process, but much remains to be understood, particularly involving the relevance of these insights to realistic astrophysical disks.
+This work provides a mathematical description of the MRI as a pattern-forming process, but much remains to be understood, particularly involving the application of this model system to realistic astrophysical disks.
 Our model is most directly relevant to Taylor-Couette flows, and we emphasize that laboratory MRI experiments stand poised to observe the MRI-driven pattern formation predicted here.

### 2016-08-31 python/widegap/allorders2_widegap.py

@@ -157,11 +157,12 @@ class MRI():
         return problem
 
     def fastest_growing(self):
-        gr, largest_eval_indx,freq  = self.EP.growth_rate({})
+        gr, largest_eval_indx, freq  = self.EP.growth_rate({})
         self.largest_eval_indx = largest_eval_indx
         self.EP.solver.set_state(largest_eval_indx)
         
-        logger.info("Fastest mode has growth rate {}".format(gr))
+        logger.info("Fastest mode has growth rate {} and frequency {}".format(gr, freq))
+        return gr, freq
     
     def solve_BVP(self, BVP):
     
@@ -385,7 +386,7 @@ class AdjointHomogenous(MRI):
             self.finalize()
             
     def finalize(self):
-        self.fastest_growing()
+        self.gr, self.freq = self.fastest_growing()
                 
         self.psi = self.EP.solver.state['psi']
         self.u = self.EP.solver.state['u']
@@ -487,7 +488,7 @@ class OrderE(MRI):
             self.finalize()
 
     def finalize(self):
-        self.fastest_growing()        
+        self.gr, self.freq = self.fastest_growing()        
         
         self.psi = self.EP.solver.state['psi']
         self.u = self.EP.solver.state['u']

### 2016-08-31 python/widegap/run_widegap_amplitude.py

@@ -272,6 +272,10 @@ with h5py.File(fn,'w') as f:
     f.attrs["q_R0"] = aa.q_R0
     f.attrs["conducting"] = aa.conducting
     f.attrs["norm"] = aa.norm
+    f.attrs["linear_gr"] = aa.o1.gr
+    f.attrs["linear_freq"] = aa.o1.freq
+    f.attrs["adjoint_gr"] = aa.ah.gr
+    f.attrs["adjoint_freq"] = aa.ah.freq
 
     f.attrs["a"] = aa.a
     f.attrs["b"] = aa.b

### 2016-08-31 paper/widegap_mri_paper.tex
terms from expansion of dz within D

@@
 \mathcal{O}(\epsilon)&: \mathcal{L} \mathbf{V}_1 + \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \mathcal{D} \partial_t \mathbf{V}_1 = 0 \label{eq:ordere}\\
-\mathcal{O}(\epsilon^2)&: \mathcal{L} \mathbf{V}_2 + \xi \widetilde{\mathcal{H}} \mathbf{V}_2 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_1 + \xi \mathcal{H} \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \\\label{eq:ordere2}
-\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \xi \widetilde{\mathcal{H}} \mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2\nonumber \\ 
+\mathcal{O}(\epsilon^2)&: \mathcal{L} \mathbf{V}_2 + \xi \widetilde{\mathcal{H}} \mathbf{V}_2 + \mathcal{D} \partial_t \mathbf{V}_2 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_1 + \xi \mathcal{H} \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \label{eq:ordere2}\\
+\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \xi \widetilde{\mathcal{H}} \mathbf{V}_3 + \mathcal{D} \partial_t \mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2\nonumber \\
 &+ \xi \mathcal{H}\partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 - \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3 = 0.
-\\\label{eq:ordere3}
+\label{eq:ordere3}\\
 \end{align}
@@
 This is because we do not have slow variation in the radial dimension.
+In the standard MRI case, $\sigma = 0$ at marginality and so the $\partial_t$ terms drop out of the equations.
+For the helical MRI case, however, $\sigma$ has a nonzero imaginary component even at threshold, so we must formally include these terms in our perturbation expansion.
 The slow variation in $Z$ and $T$ are parameterized as an amplitude function $\alpha(Z, T)$ which modulates the flow in these dimensions.
@@
 \label{eq:ansatz}
-\mathbf{V}_1 = \alpha(Z, T) \mathbb{V}_{11}(r) e^{i k_z z} + c.c.,
+\mathbf{V}_1 = \alpha(Z, T) \mathbb{V}_{11}(r) e^{i k_z z + \sigma t} + c.c.,
 \eeq
 
-where the radial variation is contained in $\mathbb{V}_{11}$.
+where the radial variation is contained in $\mathbb{V}_{11}$, and $\sigma = \gamma + i\omega$.
 

### 2016-09-01 paper/widegap_mri_paper.tex
run thingap at Pm 1E-6

@@
 \centering
-\includegraphics[width=\textwidth]{../python/widegap/figures/orders_1_2_widegap_fiducial_eigenfunctions.pdf}
+\includegraphics[width=\textwidth]{../figures/widegap_orders1_2_Pm16E-7.pdf}
 \caption{Eigenfunctions of the first order equations, first order adjoint homogenous equations, and second order equations.
@@
 \centering
-\includegraphics[width=\textwidth]{../python/widegap/figures/nonlinear_2_3_widegap_fiducial_eigenfunctions.pdf}
+\includegraphics[width=\textwidth]{../figures/widegap_n2_n3_Pm16E-7.pdf}
 \caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.}\label{fig:N2_N3}
@@
 \centering
-\includegraphics[width=\columnwidth]{../figures/widegap_coeffs_by_Pm.pdf}
-\caption{Critical parameters $k_c$ and $\reym$, and coefficients of the Ginzburg-Landau equation (Equation \ref{eq:gle}) as a function of $Pm$.
-Note the very weak dependence of the coefficients on $Pm$.
-The saturation amplitude of the standard MRI system is very insensitive to $Pm$ in the wide gap case.}\label{fig:coefficients}
+\includegraphics[width=\columnwidth]{../figures/widegap_coeffs_satampscaling.png}
+\caption{Critical parameters $\reym_c$ and $k_c$, and coefficients of the Ginzburg-Landau equation (Equation \ref{eq:gle}) as a function of $Pm$.
+Note the very weak dependence of the linear ($b$) and diffusive ($h$) coefficients on $Pm$.
+The saturation amplitude $\alpha_{sat} = \sqrt{b/c}$ of the standard MRI system has a power law dependence on $Pm$ which we measure to be $\alpha_{sat} \sim Pm^{0.777}$.
+This scaling is driven by the $Pm$ dependence of the nonlinear coefficient $c$.}\label{fig:coefficients}
 \end{figure}

### 2016-09-01 python/pleiades_run_thingap_amplitude.sh

@@ -1,7 +1,7 @@
 #PBS -S /bin/bash
 #PBS -N Widegap_Amplitude
 #PBS -l select=2:ncpus=10:mpiprocs=10:model=has
-#PBS -l walltime=02:00:00
+#PBS -l walltime=00:10:00
 #PBS -j oe
 #PBS -q devel
 

### 2016-09-01 python/widegap/allorders2_widegap.py

@@ -772,7 +772,7 @@ class OrderE2(MRI):
         # multiplied by r^4
         rhs_psi21 = (-self.iR*4*1j*self.Q**3*rfield**3*o1.psi - (2/self.beta)*3*rfield**3*self.Q**2*o1.A + self.iR*4*1j*self.Q*rfield**3*o1.psi_rr
                     - self.iR*4*1j*self.Q*rfield**2*o1.psi_r + 2*rfield**2*ru0field*o1.u + (2/self.beta)*rfield**3*o1.A_rr - (2/self.beta)*rfield**2*o1.A_r
-                    - (2/self.beta)*2*xi*rfield**2*o1.B - rfield**3*2*1j*Q*1j*o1.freq*o1.psi)
+                    - (2/self.beta)*2*xi*rfield**2*o1.B + rfield**3*2*Q*o1.freq*o1.psi)
         self.rhs_psi21 = rhs_psi21.evaluate()
         
         # multiplied by r^3

### 2016-09-01 paper/widegap_mri_paper.tex
thingap run

@@
 We use our fiducial parameters for the standard MRI ($\xi = 0$).
-First-order eigenfunctions are normalized such that they are either purely real or purely imaginary, and such that $\int \Psi_{11} \mathrm{d}r = 1$.
-Adjoint homogenous eigenfunctions are normalized such that $\left< V_{11}^\dagger \cdot \mathcal{D} V_{11} \right> = 1$.}\label{fig:linear_eigenfunctions}
+Eigenfunctions are solved on a 512-element grid of Chebyshev polynomials.
+First-order eigenfunctions are normalized such that $A_{11}(r_0) = 1$.
+Adjoint homogenous eigenfunctions are normalized such that $\langle V_{11}^\dagger \cdot \mathcal{D} V_{11} \rangle = 1$.}\label{fig:linear_eigenfunctions}
 \end{figure*}
@@
 \includegraphics[width=\textwidth]{../figures/widegap_n2_n3_Pm16E-7.pdf}
-\caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.}\label{fig:N2_N3}
+\caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.
+These are nonlinear combinations of lower-order eigenfunctions, representing the weakly nonlinear mode interaction.
+At second order ($N_2$) the most unstable linear MRI mode interacts with itself and its complex conjugate.
+At third order ($N_3$) the first and second order MRI modes interact with each other.}\label{fig:N2_N3}
 \end{figure*}

### 2016-09-01 python/pleiades_run_thingap_amplitude.sh

@@ -1,7 +1,7 @@
 #PBS -S /bin/bash
-#PBS -N Widegap_Amplitude
+#PBS -N Thingap_Amplitude
 #PBS -l select=2:ncpus=10:mpiprocs=10:model=has
-#PBS -l walltime=00:10:00
+#PBS -l walltime=02:00:00
 #PBS -j oe
 #PBS -q devel
 

### 2016-09-01 paper/widegap_mri_paper.tex
thingap run with goodman and ji-like parameters at Pm=1.6E-4

@@
 where $\mathrm{U}(x) = {3}/{v_A^2 k_z^2} + 1$ at marginality.
+This form is not unique to the ideal MHD case, though the ideal approximation simplifies the expression considerably.
 When no-slip radial boundary conditions are applied, the thin-gap MRI system resembles a particle in a box with a radially constant potential well.
@@
 In the wide gap case, the ``potential" $\mathrm{U}(r)$ varies with $r$, so symmetric and antisymmetric modes are no longer required.
+This lack of symmetry is readily apparent in the eigenfunctions and nonlinear vectors in Figures \ref{fig:linear_eigenfunctions} and \ref{fig:N2_N3}, both of which display enhanced boundary layer activity at the inner boundary as compared to the outer boundary.
+The inner and outer boundary layers are symmetric in the thin gap case.
 

### 2016-09-01 python/pleiades_run_thingap_amplitude.sh

@@ -1,7 +1,7 @@
 #PBS -S /bin/bash
 #PBS -N Thingap_Amplitude
 #PBS -l select=2:ncpus=10:mpiprocs=10:model=has
-#PBS -l walltime=02:00:00
+#PBS -l walltime=00:10:00
 #PBS -j oe
 #PBS -q devel
 

### 2016-09-01 python/run_amplitude.py

@@ -21,12 +21,21 @@ import h5py
 #Pm = tgc['Pm'].value[indx].real
 #Rm = tgc['Rm_c'].value[indx].real
 #Q = tgc['Q_c'].value[indx].real
+
+# parameters like Umurhan+
+"""
 Pm = 1.0E-6
 Rm = 4.87322391646 
 Q = 0.74641163216928608
-
 q = 1.5
 beta = 25.0
+"""
+# parameters like Goodman & Ji (but thingapped)
+q = 1.9149705511835795
+beta = 41.2
+Pm = 1.6E-4
+Rm = 2.10717703478
+Q = 0.66135735868343559
 
 gridnum = 128
 x_basis = de.Chebyshev('x',gridnum)

### 2016-09-02 paper/widegap_mri_paper.tex
search widegap critical parameters to flesh out satamp plot. Pm=4E-6

@@
 This is in stark contrast to the narrow gap behavior of the system.
-For these same boundary conditions, \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}) find that the narrow gap saturation amplitude scales as $Pm^{-4/3}$.
+For these same boundary conditions, \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}) find that the narrow gap saturation amplitude scales as $Pm^{2/3}$.
 They find that this amplitude dependence is driven by the $Pm^{1/3}$ dependence of the linear boundary layer.
-Boundary layer analysis similarly reveals a $\nu^{1/3}$ dependence for the radial extent of the boundary layer (\citei{Goodman:2002ix}).
+Boundary layer analysis similarly reveals a $\nu^{1/3}$ dependence for the radial extent of the boundary layer in Taylor Couette flow (\citei{Goodman:2002ix}).
 Why does the $Pm$ scaling of the boundary layer width not translate to a stronger $Pm$ scaling for the saturation amplitude in the wide gap case?

### 2016-09-02 python/widegap/pleiades_run_critical_eigenvalue_search.sh

@@ -1,7 +1,7 @@
 #PBS -S /bin/bash
 #PBS -N CritEvalSearchbyRm
 #PBS -l select=2:ncpus=10:mpiprocs=10:model=has
-#PBS -l walltime=00:10:00
+#PBS -l walltime=02:00:00
 #PBS -j oe
 #PBS -q devel
 

### 2016-09-02 python/widegap/run_widegap_amplitude.py
widegap long run for Pm=4E-5

@@ -48,9 +48,13 @@ norm = True
 
 # ********************************************************
 # extras i derived:
-Pm = 4.0e-06
-Rm=3.3116#3.31157458348
-Q=0.90437#044922674747
+#Pm = 4.0e-06
+#Rm=3.3116#3.31157458348
+#Q=0.90437#044922674747
+
+Pm = 4.0e-05
+Rm = 3.316#12607953
+Q = 0.90449#0.90449003700134456
 
 # ********************************************************
 

### 2016-09-02 paper/widegap_mri_paper.tex
widegap long run for Pm=4E-5
added widegap narrowgap comparison figure

@@
 \centering
-\includegraphics[width=\textwidth]{../figures/wide_narrow_gap_N31_B.pdf}
+\includegraphics[width=\textwidth]{../figures/widegap_narrowgap_comparison_narrowfiducial.png}
 \caption{Nonlinear term $N_{31}^{(B)}$ for the wide gap (left) and narrow gap (right) standard MRI.
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/widegap_rey_max_tot_stresses.png}
+\includegraphics[width=\textwidth]{../figures/fiducial_standard_mri_stressplot_Pm_16E-7.png}
 \caption{Reynolds ($\mathbb{T}_{R} = u_r u_\phi$), Maxwell ($\mathbb{T}_{M} = -\frac{2}{\beta} B_r B_\phi$), and total stress ($\mathbb{T} = \mathbb{T}_{R} + \mathbb{T}_{M}$) for the fiducial standard MRI case.

### 2016-09-02 paper/widegap_mri_paper.tex
added acknowledgements; NASA grants; fixed jso affiliation

@@
 \altaffiltext{2}{Department of Astrophysics, American Museum of Natural History, New York, NY}
-\altaffiltext{3}{Department of Physics, SUNY Farmingdale}
+\altaffiltext{3}{Department of Physics and Astronomy, Bates College}
 
@@
 \section{Acknowledgements}
-NASA, NSF. The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, etc for useful discussion.
+JSO acknowledges support from NASA grant 15-LWS15_2-0063. NSF. The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, etc for useful discussion.
 

### 2016-09-02 paper/widegap_mri_paper.tex
boundary layer edits

@@
 We find the marginal system as a function of the dimensionless parameters.
-Marginality for our standard MRI system is a hyperplane in $(\reym, Pm, \beta, \Omega_2/\Omega_1, \mathrm{R}_1/\mathrm{R}_2$), but we hold all of these constant except for $\reym$.
+The marginal stability curve for our standard MRI system is a hyperplane in $(\reym, Pm, \beta, \Omega_2/\Omega_1, \mathrm{R}_1/\mathrm{R}_2$), but we hold all of these constant except for $\reym$.
 To analyze the MRI system at marginality, we fix the parameters listed in Table \ref{table:parameters} and determine the critical $\reym$ and $k_z$ by repeatedly solving the linear MRI system to determine the smallest parameter values for which the fastest growing mode is zero.
@@
 
-The nonlinear terms, detailed in Appendix \ref{app:nonlinear_terms}, represent an interesting departure from the thin-gap theory.
-The thin-gap nonlinear terms at both second and third orders are linear combinations of Jacobians.
-The nonlinear terms in the wide-gap case differ from their thin-gap analogues with in the addition of vertical advective terms.
+The form of the nonlinear terms, detailed in Appendix \ref{app:nonlinear_terms}, represent a departure from the thin-gap theory.
+The narrow gap nonlinear terms at both second and third orders are linear combinations of Jacobians.
+The nonlinear terms in the wide-gap case differ from their thin-gap analogues with the addition of vertical advective terms.
 These terms derive from the advective derivatives in the momentum and induction equations, but are filtered out in the thin-gap approximation.
-These advective terms allow the nonlinear contributions at both second and third order (i.e.
+The nonlinear terms ultimately determine the dependence of the saturation amplitude on $Pm$, as described below.
+%These advective terms allow the nonlinear contributions at both second and third order (i.e.
 N$_2$ and N$_3$) not to individually satisfy the boundary conditions on $\Psi$ and $u$.
@@
 We examine the behavior of the wide gap MRI system as a function of $Pm$.
-Figure \ref{fig:coefficients} shows the critical parameters $k_c$ and $\reym$ as a function of $Pm$, as well as the coefficients $b$, $c$, and $d$.
+Figure \ref{fig:coefficients} shows the critical parameters $k_c$ and $\reym$ as a function of $Pm$, as well as the GLE linear coefficient $b$ and the diffusion coefficient $d$.
 The GLE coefficients are remarkably insensitive to $Pm$.
@@
 Boundary layer analysis similarly reveals a $\nu^{1/3}$ dependence for the radial extent of the boundary layer in Taylor Couette flow (\citei{Goodman:2002ix}).
-Why does the $Pm$ scaling of the boundary layer width not translate to a stronger $Pm$ scaling for the saturation amplitude in the wide gap case?
+%Why does the $Pm$ scaling of the boundary layer width not translate to a stronger $Pm$ scaling for the saturation amplitude in the wide gap case?
 The boundary layer in the wide gap case is strongly localized at the walls, i.e.
 $r_1$ and $r_2$.
-Figure \ref{fig:wide_narrow_N31B} shows the structure of the third-order nonlinear term $N_{31}^{(B)}$ as a function of $Pm$ for both the narrow and wide gap standard MRI.
-$N_{31}$ is the vector that determines the GLE coefficient $c$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs}, \citei{Clark:2016} for the narrow gap equations).
-Clearly, the boundary layers scale with $Pm$ in both the wide and narrow gap MRI.
-However, in the narrow gap case this scaling extends prominently into the center of the channel, whereas for the wide gap case the bulk of the flow is relatively unaffected by the boundary layer scaling.
+Figure \ref{fig:wide_narrow_N31B} shows the structure of the third-order nonlinear term $N_{31}^{(A)}$ as a function of $Pm$ for both the narrow and wide gap standard MRI.
+$N_{31}$ is the vector that determines the GLE coefficient $c$, and thus the scaling of the saturation amplitude because of the insensitivity of $b$ to $Pm$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs}, \citei{Clark:2016} for the narrow gap equations).
+Clearly, the width of the boundary layers scales with $Pm$ in both the wide and narrow gap MRI.
+This translates to a steeper saturation amplitude $Pm$ dependence in the wide gap case.%However, in the narrow gap case this scaling extends prominently into the center of the channel, whereas for the wide gap case the bulk of the flow is relatively unaffected by the boundary layer scaling.
 

### 2016-09-02 paper/widegap_mri_paper.tex
more hmri, acknowledgments

@@
 When $\xi$ in Equation \ref{eq:unperturbed_matrix_eqns} is not equal to zero, the helical MRI arises.
-We examine a single fiducial helical MRI case, for the parameters listed in Table \ref{table:parameters}.
+We examine a single fiducial helical MRI case, for the parameters used by \citei{Hollerbach:2005tr}, listed in Table \ref{table:parameters}.
+The helical MRI is an overstability, so the ansatz linear eigenvector we consider (Equation \ref{eq:ansatz}) is characterized by a complex temporal eigenvalue $\sigma$.
+For our fiducial parameters, the marginal mode has a frequency $\omega = 0.153$.
+This means that the helical MRI modes are traveling waves, drifting in the $z$ direction.
+
 At the conclusion of the weakly nonlinear analysis, we find that the coefficients of Equation \ref{eq:gle} are $b = $, $d = $, and $c = $, i.e.
@@
 \section{Acknowledgements}
-JSO acknowledges support from NASA grant 15-LWS15_2-0063. NSF. The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, etc for useful discussion.
+S.E.C.
+was supported by a National Science Foundation Graduate Research Fellowship under grant No.
+DGE-11-44155.
+J.S.O.
+acknowledges support from NASA grant 15-LWS15\_2-0063.
+The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, Ellen Zweibel, etc for useful discussion.
 

### 2016-09-02 paper/widegap_mri_paper.tex
hmri added

@@
 \citei{Recktenwald:1993}).
-We emphasize that this is a model equation, valid only near marginality (e.g.
-\citei{Cross:1993el}).
+We emphasize that this is a model equation, valid only near marginality (\citei{Cross:1993el}).
 The dynamics of the GLE are determined by its coefficients, which are in turn determined by the linear eigenfunctions and nonlinear vectors plotted in Figures \ref{fig:linear_eigenfunctions} and \ref{fig:N2_N3}.
@@
 For our fiducial parameters, the marginal mode has a frequency $\omega = 0.153$.
-This means that the helical MRI modes are traveling waves, drifting in the $z$ direction.
+This means that the helical MRI modes are traveling waves, drifting in the $z$ direction with a phase velocity $\omega/k_c$.
 
@@
 With the introduction of an azimuthal component, the background magnetic field acquires a handedness that is not present in a purely axial field.
+The helical MRI eigenvectors are therefore free to be out of phase with one another.
+In our perturbation series, the helical MRI modes interact within and between orders with modes which carry different phases, leading to complex GLE coefficients.
+
+The phase dynamics of the complex GLE are well-studied in a variety of systems, and depend on the values of the GLE coefficients.
+The complex GLE may be unstable to traveling wave instabilities such as the Benjamin-Feir instability, a generalization of the Ekhaus instability which acts on stationary waves, and to which the standard MRI system may be unstable.
+The complex GLE can also admit spatiotemporal chaos, and various classes of coherent structures (\citei{Aranson:2002}).
+%Because the saturation amplitude of the weakly nonlinear helical MRI is governed by a complex GLE, the saturated state of the helical MRI may be unstable to 
+Although a detailed description of the phase dynamics in the helical MRI is beyond the scope of this work, we note that such long-wavelength, long-timescale behavior may be observed in liquid metal helical MRI experiments.
 

### 2016-09-02 paper/widegap_mri_paper.tex
more points on coeff plot

@@
 \centering
-\includegraphics[width=\columnwidth]{../figures/widegap_coeffs_satampscaling.png}
+\includegraphics[width=\columnwidth]{../figures/widegap_coeffs_satampscaling_fivepoints.png}
 \caption{Critical parameters $\reym_c$ and $k_c$, and coefficients of the Ginzburg-Landau equation (Equation \ref{eq:gle}) as a function of $Pm$.
@@
 For our fiducial parameters, the marginal mode has a frequency $\omega = 0.153$.
-This means that the helical MRI modes are traveling waves, drifting in the $z$ direction with a phase velocity $\omega/k_c$.
+This means that the helical MRI modes are traveling waves, moving in the $z$ direction with a phase velocity $\omega/k_c$ = .
 
-At the conclusion of the weakly nonlinear analysis, we find that the coefficients of Equation \ref{eq:gle} are $b = $, $d = $, and $c = $, i.e.
-complex.
+At the conclusion of the weakly nonlinear analysis, we find that the coefficients of Equation \ref{eq:gle} are complex.
 The marginal helical MRI is thus described by a complex GLE.

### 2016-09-02 paper/widegap_mri_paper.tex
summary

@@
 
+In this work we carry out a formal weakly nonlinear multiple scales analysis of the MRI in a Taylor Couette flow.
+We examine both the standard MRI, in which the background magnetic field is purely axial, as well as the helical MRI, in which the background magnetic field also contains an azimuthal component.
+We find that in both cases the amplitude of the weakly nonlinear system is governed by a Ginzburg-Landau equation, but the GLE is real in the case of the standard MRI and complex for the helical MRI.
+We examine the $Pm$ dependence of the saturation amplitude of the standard MRI, making explicit comparison to the $Pm$ scaling behavior of this system in the narrow gap limit.
+
 Our work should be placed in the broader context of emergent pattern formation in physical systems.
@@
 Our model is most directly relevant to Taylor-Couette flows, and we emphasize that laboratory MRI experiments stand poised to observe the MRI-driven pattern formation predicted here.
+
+The theory presented here may be extended in several ways.
 

### 2016-09-02 paper/widegap_mri_paper.tex
edited introduction. changed $Pm$ -> $\Pm$ with requisite definition.

--- 2016-09-02 paper/widegap_mri_paper.tex
@@
 \newcommand{\eeq}{\end{equation}}
+
 \newcommand{\hyd}    {{\rm H}}
@@
 \newcommand\reym{\mathrm{Rm}}
+\newcommand{\Pm}{\mathrm{Pm}}
 
@@
 
-The magnetorotational instability (MRI) drives angular momentum transport and turbulence in astrophysical disks.
-Its discovery by \citeauthor{Balbus:1991vs} (\citeyear{Balbus:1991vs}; actually a rediscovery of \citei{Chandrasekhar:1960wh}) was a breakthrough in the longstanding question of how efficient accretion can exist in the universe: that is, how matter collapsing onto a central body is able to coalesce despite the conservation of specific angular momentum.
-%The ubiquity of disks ...
-The ubiquity of astrophysical accretion disks suggests that a wide variety of systems both experience and overcome this centrifugal barrier.
-Furthermore, accretion proceeds even in hydrodynamically stable disks, where molecular viscosity is vastly insufficient to drive angular momentum transport (\citei{Shakura:1973wg}).
-The MRI is excited by weak magnetic fields in differentially rotating fluids, and since its discovery it has been widely invoked to explain accretion in protoplanetary disks (see \citei{Armitage:2010} and references therein), binary systems (), and disks around black holes (), as well as jet and wind launching (\citei{Lesur:2013}), dynamos, etc (cite these)
-
-%Magnetic fields were once assumed to have a stabilizing effect
-
-%The MRI is fundamental to our understanding of star formation, but its 
-%- people make a number of simplifications, incl thin gap, shearing box, etc
-
-%Despite the challenges inherent in studying the MRI, much progress has been made nu
+The magnetorotational instability (MRI) is widely belived to drive angular momentum transport in astrophysical disks.
+The MRI is a local instability excited by weak magnetic fields in differentially rotating fluids, and since its discovery \citet{Balbus:1991vs} it has been widely invoked to explain accretion in protoplanetary disks (see \citei{Armitage:2010} and references therein), binary systems (), and disks around black holes (), as well as jet and wind launching (\citei{Lesur:2013}), dynamos, etc (cite these).
 
@@
   \hline
- & $\xi$ & $Pm$ & $\beta$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ & radial magnetic b.c. \\ 
+ & $\xi$ & $\Pm$ & $\beta$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ & radial magnetic b.c. \\ 
   \hline\hline
@@
 
-Although the MRI is broadly important to many astrophysical systems, many of its general properties, and in particular its nonlinear saturation mechanism, remain poorly understood.
-The diversity of astrophysical systems which may permit the MRI admits an enormous parameter space to be explored.
-In protoplanetary disks, for example, the behavior and evolution of the MRI may change drastically depending on the properties of the magnetic field, the disk composition, disk geometry, and so forth.
-Numerical simulation of realistic disk physics is currently an area of intense focus, and is enabling the study of nonideal MHD effects, disk stratification, nonequilibrium chemistry, and other complex physics that does not lend itself easily to analytic study (\citei{Fleming:2003fs}, \citei{Bai:2011cm}, \citei{Flock:2013}, \citei{Suzuki:2014vh}, to name only a few).
+The diversity of astrophysical systems which may be MRI unstable yields an enormous parameter space to be explored.
+In protoplanetary disks, for example, the behavior and evolution of the MRI--and even its very existence--may change drastically depending on the properties of the magnetic field, the disk composition, disk geometry, and so forth.
+Multiphysics numerical simulations of such systems is currently an area of intense focus, enabling the study of nonideal MHD effects, disk stratification, nonequilibrium chemistry, and other complex physics that does not lend itself easily to analytic study \citep[e.g.][among many others]{Fleming:2003fs,Bai:2011cm, Flock:2013,Suzuki:2014vh}.
 Still, computational costs inevitably constrain numerical approaches.
 MRI saturation is a complicated nonlinear problem which may depend on the assumptions and approximations adopted by simulations in nonobvious ways.
-Analytic methods can play a powerful role in elucidating the mechanisms responsible for MRI saturation.
-For instance, analytical approaches have revealed the mechanism that likely governs saturation in the ``shearing box" approximation.
-The shearing box is an oft-invoked local approximation in which a section of a disk is represented by solving the MHD equations in an isolated region subject to shear periodic boundary conditions.
-The shearing box is an inexpensive computational framework, and has been widely used to study many aspects of the MRI ().
-However, while the MRI is a local instability, certain properties of the local problem are not generic to the global problem.
-In the shearing box, linear evolution is dominated by a class of MRI mode known as channel modes.
-These are linear modes which also happen to be exact solutions to the \textit{nonlinear} local MRI equations.
-Runaway growth is avoided in this paradigm by the instability of the channel modes themselves, which are destroyed by parasitic (secondary) instabilities (\citei{Goodman:1994ul}, \citei{Pessah:2010ic}).
-The growth of parasitic modes provides a saturation avenue for channel mode-dominated flows, yet this is unlikely to be the dominant saturation mechanism in laboratory experiments or astrophysical disks, as channel modes are artificially over-prominent in the shearing box (e.g.
+For example, the magnetic Prandtl number $\Pm = \nu/\eta \sim 10^{-8}$ in protoplanetary disks \citet[e.g.][]{OishiMacLow:2011} and $\sim 10^{-6}$ in liquid metal experiments \citep[e.g.][]{Goodman:2002ix}.
+Such extreme ratios of viscosity to resistivity far exceed current computational resources.
+However, we can construct asymptotic approximations valid for $\Pm << 1$ using analytic methods.
+Analytic methods can also play a powerful role in elucidating the mechanisms responsible for MRI saturation.
+For instance, analytical approaches have revealed the mechanism that likely governs saturation in the ``shearing box'' approximation.
+The shearing box is an oft-invoked local approximation in which a section of a disk is represented by solving the MHD equations in a rotating, Cartesian box with a linearized background shear, subject to shear periodic boundary conditions in the radial direction.
+The shearing box is a convenient computational framework allowing extreme resolution for local MRI studies and has been extended to include vertical stratification and a wide variety of diffusive effects.
+However, while the MRI is a local instability, there are a number of important problems that require a global treatment.
+Perhaps most importantly, linear evolution in the shearing box is dominated by channel modes, particularly when a net vertical magnetic field threads the box.
+These linear modes are exact solutions to the \textit{nonlinear} local MRI equations.
+The shearing box MRI system avoids runaway growth by a secondary instability of the channel modes themselves (\citei{Goodman:1994ul}, \citei{Pessah:2010ic}).
+The growth of parasitic modes provides a saturation avenue for channel mode-dominated flows, yet this is unlikely to be the dominant saturation mechanism in laboratory experiments or astrophysical disks, as channel modes are artificially over-represented in the shearing box (e.g.
 \citei{Latter:2015}).
-Thus while the shearing box may accurately approximate many features of the global MRI, the saturation mechanism is not among them.
+Thus while the shearing box may accurately approximate many features of the global MRI, the saturation mechanism may not be among them.
 
@@
 
-The theory we develop here employs realistic radial boundary conditions, and so MRI modes are global and channel modes are not present.
+In this paper, we develop an analytic, global theory for the MRI in a Taylor-Couette flow.
+This system precludes channel modes, allowing us to develop an understanding of MRI saturation in their absence.
 A number of saturation mechanisms have been proposed for the MRI which do not rely on channel modes dominating the flow.
@@
 Our investigation is astrophysically motivated, but we also intend our theory to be relevant to laboratory experiments.
-Several experimental efforts are attempting to observe the MRI in the laboratory, which would allow the study of a crucial astrophysical phenomenon in a controlled setting.
+Several experimental efforts are attempting to observe the MRI in the laboratory, which will allow the study of a crucial astrophysical phenomenon in a controlled setting.
 Unfortunately, detection of the MRI has so far proven elusive.
 \citei{Sisan:2004ig} claimed to detect the MRI in a spherical Couette flow, but most likely detected unrelated MHD instabilities instead (\citei{Hollerbach:2009ig}, \citei{Gissinger:2011td}).
-Most relevant to our work is the Princeton Plasma Physics Laboratory (PPPL) MRI Experiment, a liquid gallium Taylor-Couette flow exposed to an axial magnetic field (\citei{Ji:2001kd}).
-There has been some theoretical work designed to complement the Princeton MRI experiment involving direct simulation of the experimental conditions, much of it focused on the specific challenges in identifying MRI signatures despite apparatus-driven spurious flows (e.g.
+Most relevant to our work is the Princeton Plasma Physics Laboratory (PPPL) MRI Experiment, a liquid gallium Taylor-Couette flow with an axial magnetic field (\citei{Ji:2001kd}).
+There has been some theoretical work designed to complement the Princeton MRI experiment involving direct numerical simulation of the experimental conditions, much of it focused on the specific challenges in identifying MRI signatures despite apparatus-driven spurious flows (e.g.
 \citei{Gissinger:2012gc}).
-In particular, the vertical endcaps on a laboratory MRI apparatus drive meridional flows which both inhibit the excitement of MRI and obscure its detection.
+The vertical endcaps on a laboratory MRI apparatus drive meridional flows which both inhibit the excitement of MRI and obscure its detection.
 The Princeton MRI experiment employs split, independently rotating endcaps to mitigate these flows (\citei{Schartman:2009df}).
-Our work here assumes an infinite vertical domain, an idealization that is theoretically expedient but experimentally impractical.
+Our work assumes an infinite vertical domain, an idealization that is theoretically expedient but experimentally impractical.
 However, our setup is designed to be an accurate treatment of the radial dimension of the flow in a Taylor Couette apparatus like the one used in the Princeton MRI experiment.
@@
 
-Because we include curvature terms, our theory is also relevant to the helical magnetorotational instability (HMRI).
-Discovered by \citeauthor{Hollerbach:2005tr} (\citeyear{Hollerbach:2005tr}), the HMRI is an overstability in which the background magnetic field is helical.
+Because we include curvature terms, our treatment also allows us to study the helical magnetorotational instability (HMRI).
+The HMRI is an overstability in which the background magnetic field is helical, $\mathbf{B} = B_0 (\xi \mathbf{\hat{\theta}} + \mathbf{\hat{z}})$ \citet{Hollerbach:2005tr}.
 The HMRI currently occupies a special place in the MRI puzzle.
-The HMRI has been proposed as a method of awakening angular momentum transport in the ``dead zones" of protoplanetary disks where the magnetic Prandtl number ($Pm = \nu/\eta$) becomes very small.
-However the rotation profiles needed to excite HMRI may be prohibitively steeper than Keplerian depending on the boundary conditions, and so its role in astrophysical disks is currently debated (\citei{Liu:2006}, \citei{Rudiger:2007}, \citei{Kirillov:2013}).
-The HMRI is significantly easier to excite in a laboratory setting than the standard MRI, and has already been detected in the laboratory by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
-%The standard MRI has yet to be detected in the laboratory, although several plasma experiments are making progress.
+The HMRI has been proposed as a method of awakening angular momentum transport in the ``dead zones" of protoplanetary disks where the $\Pm$ becomes very small.
+However the rotation profiles needed to excite HMRI may be steeper than Keplerian, depending on the boundary conditions, and so its role in astrophysical disks is currently a matter of debate \citep{Liu:2006,Rudiger:2007,Kirillov:2013}.
+Regardless of its astrophysical role, the HMRI is significantly easier to excite in a laboratory setting than the standard MRI, and has already been detected by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
 
 In this work we explore the behavior of the viscous, dissipative MRI in a cylindrical geometry close to threshold, making explicit comparisons to the standard MRI behavior in the thin-gap regime.
-We investigate both the standard MRI, in which the background magnetic field is purely axial, as well as the helical MRI, in which the background magnetic field has an azimuthal component as well.
+We investigate both the standard MRI, in which the background magnetic field is purely axial, as well as the helical MRI.
+In section~\ref{sec:wide_gap}, we lay out the basic mathematical framework of the problem.
+In section~\ref{sec:wnl_analysis}, we introduce the method of multiple scales we use to construct our theory.
+In section~\ref{sec:results} we describe the basic results, and in section~\ref{sec:discussion} we place them in the context of previous work on other instabilities, discuss their relevance to experiments, and reiterate our final conclusions.
 
@@
 \section{Wide Gap Equations}
-
+\label{sec:wide_gap}
 The basic equations solved are the momentum and induction equations,
@@
 \section{Weakly Nonlinear Perturbation Analysis}\label{sec:wnl_analysis}
-
+\label{sec:Perturbation}
 \begin{figure}
@@
 \includegraphics[width=\columnwidth]{../figures/widegap_coeffs_satampscaling.png}
-\caption{Critical parameters $\reym_c$ and $k_c$, and coefficients of the Ginzburg-Landau equation (Equation \ref{eq:gle}) as a function of $Pm$.
-Note the very weak dependence of the linear ($b$) and diffusive ($h$) coefficients on $Pm$.
-The saturation amplitude $\alpha_{sat} = \sqrt{b/c}$ of the standard MRI system has a power law dependence on $Pm$ which we measure to be $\alpha_{sat} \sim Pm^{0.777}$.
-This scaling is driven by the $Pm$ dependence of the nonlinear coefficient $c$.}\label{fig:coefficients}
+\caption{Critical parameters $\reym_c$ and $k_c$, and coefficients of the Ginzburg-Landau equation (Equation \ref{eq:gle}) as a function of $\Pm$.
+Note the very weak dependence of the linear ($b$) and diffusive ($h$) coefficients on $\Pm$.
+The saturation amplitude $\alpha_{sat} = \sqrt{b/c}$ of the standard MRI system has a power law dependence on $\Pm$ which we measure to be $\alpha_{sat} \sim \Pm^{0.777}$.
+This scaling is driven by the $\Pm$ dependence of the nonlinear coefficient $c$.}\label{fig:coefficients}
 \end{figure}
@@
 \caption{Nonlinear term $N_{31}^{(B)}$ for the wide gap (left) and narrow gap (right) standard MRI.
-Each line represents a run for a different $Pm$, from $Pm = 1E-4$ (darkest) to $Pm \sim 1E-3$ (lightest).
+Each line represents a run for a different $\Pm$, from $\Pm = 1E-4$ (darkest) to $\Pm \sim 1E-3$ (lightest).
 Inlaid plot in wide gap case shows a zoomed-in view of the boundary layer at the outer boundary ($r_2$).
@@
 We find the marginal system as a function of the dimensionless parameters.
-The marginal stability curve for our standard MRI system is a hyperplane in $(\reym, Pm, \beta, \Omega_2/\Omega_1, \mathrm{R}_1/\mathrm{R}_2$), but we hold all of these constant except for $\reym$.
+The marginal stability curve for our standard MRI system is a hyperplane in $(\reym, \Pm, \beta, \Omega_2/\Omega_1, \mathrm{R}_1/\mathrm{R}_2$), but we hold all of these constant except for $\reym$.
 To analyze the MRI system at marginality, we fix the parameters listed in Table \ref{table:parameters} and determine the critical $\reym$ and $k_z$ by repeatedly solving the linear MRI system to determine the smallest parameter values for which the fastest growing mode is zero.
@@
 \citei{Recktenwald:1993}).
-We emphasize that this is a model equation, valid only near marginality (\citei{Cross:1993el}).
+We emphasize that this is a model equation, valid only near marginality (e.g.
+\citei{Cross:1993el}).
 The dynamics of the GLE are determined by its coefficients, which are in turn determined by the linear eigenfunctions and nonlinear vectors plotted in Figures \ref{fig:linear_eigenfunctions} and \ref{fig:N2_N3}.
@@
 \section{Results}
-
+\label{sec:results}
 \begin{figure*}
@@
 \subsection{Standard MRI}
-
+\label{sec:SMRI}
 For the standard MRI we derive a real GLE.
@@
 These terms derive from the advective derivatives in the momentum and induction equations, but are filtered out in the thin-gap approximation.
-The nonlinear terms ultimately determine the dependence of the saturation amplitude on $Pm$, as described below.
+The nonlinear terms ultimately determine the dependence of the saturation amplitude on $\Pm$, as described below.
 %These advective terms allow the nonlinear contributions at both second and third order (i.e.
@@
 
-We examine the behavior of the wide gap MRI system as a function of $Pm$.
-Figure \ref{fig:coefficients} shows the critical parameters $k_c$ and $\reym$ as a function of $Pm$, as well as the GLE linear coefficient $b$ and the diffusion coefficient $d$.
-The GLE coefficients are remarkably insensitive to $Pm$.
-From Equation \ref{eq:gle} it is readily apparent that the asymptotic saturation amplitude is $\alpha_{s} = \pm \sqrt{b/c}$, so we conclude that the saturation amplitude of the MRI is only very weakly dependent on $Pm$.
-Note that because $\reym$ is essentially constant as a function of $Pm$, the saturation amplitude is equivalently insensitive to $\reye^{-1}$.
+We examine the behavior of the wide gap MRI system as a function of $\Pm$.
+Figure \ref{fig:coefficients} shows the critical parameters $k_c$ and $\reym$ as a function of $\Pm$, as well as the GLE linear coefficient $b$ and the diffusion coefficient $d$.
+The GLE coefficients are remarkably insensitive to $\Pm$.
+From Equation \ref{eq:gle} it is readily apparent that the asymptotic saturation amplitude is $\alpha_{s} = \pm \sqrt{b/c}$, so we conclude that the saturation amplitude of the MRI is only very weakly dependent on $\Pm$.
+Note that because $\reym$ is essentially constant as a function of $\Pm$, the saturation amplitude is equivalently insensitive to $\reye^{-1}$.
 This is in stark contrast to the narrow gap behavior of the system.
-For these same boundary conditions, \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}) find that the narrow gap saturation amplitude scales as $Pm^{2/3}$.
-They find that this amplitude dependence is driven by the $Pm^{1/3}$ dependence of the linear boundary layer.
+For these same boundary conditions, \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}) find that the narrow gap saturation amplitude scales as $\Pm^{2/3}$.
+They find that this amplitude dependence is driven by the $\Pm^{1/3}$ dependence of the linear boundary layer.
 Boundary layer analysis similarly reveals a $\nu^{1/3}$ dependence for the radial extent of the boundary layer in Taylor Couette flow (\citei{Goodman:2002ix}).
-%Why does the $Pm$ scaling of the boundary layer width not translate to a stronger $Pm$ scaling for the saturation amplitude in the wide gap case?
+%Why does the $\Pm$ scaling of the boundary layer width not translate to a stronger $\Pm$ scaling for the saturation amplitude in the wide gap case?
 The boundary layer in the wide gap case is strongly localized at the walls, i.e.
 $r_1$ and $r_2$.
-Figure \ref{fig:wide_narrow_N31B} shows the structure of the third-order nonlinear term $N_{31}^{(A)}$ as a function of $Pm$ for both the narrow and wide gap standard MRI.
-$N_{31}$ is the vector that determines the GLE coefficient $c$, and thus the scaling of the saturation amplitude because of the insensitivity of $b$ to $Pm$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs}, \citei{Clark:2016} for the narrow gap equations).
-Clearly, the width of the boundary layers scales with $Pm$ in both the wide and narrow gap MRI.
-This translates to a steeper saturation amplitude $Pm$ dependence in the wide gap case.%However, in the narrow gap case this scaling extends prominently into the center of the channel, whereas for the wide gap case the bulk of the flow is relatively unaffected by the boundary layer scaling.
+Figure \ref{fig:wide_narrow_N31B} shows the structure of the third-order nonlinear term $N_{31}^{(A)}$ as a function of $\Pm$ for both the narrow and wide gap standard MRI.
+$N_{31}$ is the vector that determines the GLE coefficient $c$, and thus the scaling of the saturation amplitude because of the insensitivity of $b$ to $\Pm$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs}, \citei{Clark:2016} for the narrow gap equations).
+Clearly, the width of the boundary layers scales with $\Pm$ in both the wide and narrow gap MRI.
+This translates to a steeper saturation amplitude $\Pm$ dependence in the wide gap case.%However, in the narrow gap case this scaling extends prominently into the center of the channel, whereas for the wide gap case the bulk of the flow is relatively unaffected by the boundary layer scaling.
 
@@
 \subsection{Helical MRI}
-
+\label{sec:HMRI}
 When $\xi$ in Equation \ref{eq:unperturbed_matrix_eqns} is not equal to zero, the helical MRI arises.
-We examine a single fiducial helical MRI case, for the parameters used by \citei{Hollerbach:2005tr}, listed in Table \ref{table:parameters}.
-The helical MRI is an overstability, so the ansatz linear eigenvector we consider (Equation \ref{eq:ansatz}) is characterized by a complex temporal eigenvalue $\sigma$.
-For our fiducial parameters, the marginal mode has a frequency $\omega = 0.153$.
-This means that the helical MRI modes are traveling waves, drifting in the $z$ direction with a phase velocity $\omega/k_c$.
-
+We examine a single fiducial helical MRI case, for the parameters listed in Table \ref{table:parameters}.
 At the conclusion of the weakly nonlinear analysis, we find that the coefficients of Equation \ref{eq:gle} are $b = $, $d = $, and $c = $, i.e.
@@
 With the introduction of an azimuthal component, the background magnetic field acquires a handedness that is not present in a purely axial field.
-The helical MRI eigenvectors are therefore free to be out of phase with one another.
-In our perturbation series, the helical MRI modes interact within and between orders with modes which carry different phases, leading to complex GLE coefficients.
-
-The phase dynamics of the complex GLE are well-studied in a variety of systems, and depend on the values of the GLE coefficients.
-The complex GLE may be unstable to traveling wave instabilities such as the Benjamin-Feir instability, a generalization of the Ekhaus instability which acts on stationary waves, and to which the standard MRI system may be unstable.
-The complex GLE can also admit spatiotemporal chaos, and various classes of coherent structures (\citei{Aranson:2002}).
-%Because the saturation amplitude of the weakly nonlinear helical MRI is governed by a complex GLE, the saturated state of the helical MRI may be unstable to 
-Although a detailed description of the phase dynamics in the helical MRI is beyond the scope of this work, we note that such long-wavelength, long-timescale behavior may be observed in liquid metal helical MRI experiments.
 
 \section{Discussion}
-
+\label{sec:discussion}
 Our work should be placed in the broader context of emergent pattern formation in physical systems.
@@
 \section{Acknowledgements}
-S.E.C.
-was supported by a National Science Foundation Graduate Research Fellowship under grant No.
-DGE-11-44155.
-J.S.O.
-acknowledges support from NASA grant 15-LWS15\_2-0063.
-The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, Ellen Zweibel, etc for useful discussion.
+JSO acknowledges support from NASA grant 15-LWS15\_2-0063. NSF. The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, etc for useful discussion.
 

### 2016-09-02 paper/widegap_mri_paper.tex

@@
 \centering
-\includegraphics[width=\columnwidth]{../figures/widegap_coeffs_satampscaling.png}
+\includegraphics[width=\columnwidth]{../figures/widegap_coeffs_satampscaling_fivepoints.png}
 \caption{Critical parameters $\reym_c$ and $k_c$, and coefficients of the Ginzburg-Landau equation (Equation \ref{eq:gle}) as a function of $\Pm$.
@@
 When $\xi$ in Equation \ref{eq:unperturbed_matrix_eqns} is not equal to zero, the helical MRI arises.
-We examine a single fiducial helical MRI case, for the parameters listed in Table \ref{table:parameters}.
-At the conclusion of the weakly nonlinear analysis, we find that the coefficients of Equation \ref{eq:gle} are $b = $, $d = $, and $c = $, i.e.
-complex.
+We examine a single fiducial helical MRI case, for the parameters used by \citei{Hollerbach:2005tr}, listed in Table \ref{table:parameters}.
+The helical MRI is an overstability, so the ansatz linear eigenvector we consider (Equation \ref{eq:ansatz}) is characterized by a complex temporal eigenvalue $\sigma$.
+For our fiducial parameters, the marginal mode has a frequency $\omega = 0.153$.
+This means that the helical MRI modes are traveling waves, moving in the $z$ direction with a phase velocity $\omega/k_c$ = .
+
+At the conclusion of the weakly nonlinear analysis, we find that the coefficients of Equation \ref{eq:gle} are complex.
 The marginal helical MRI is thus described by a complex GLE.
@@
 
+The theory presented here may be extended in several ways.
+
 \section{Acknowledgements}
-JSO acknowledges support from NASA grant 15-LWS15\_2-0063. NSF. The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, etc for useful discussion.
+S.E.C.
+was supported by a National Science Foundation Graduate Research Fellowship under grant No.
+DGE-11-44155.
+J.S.O.
+acknowledges support from NASA grant 15-LWS15\_2-0063.
+The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, Ellen Zweibel, etc for useful discussion.
 

### 2016-09-02 paper/widegap_mri_paper.tex
taylor couette --> TC

@@
 The shearing box MRI system avoids runaway growth by a secondary instability of the channel modes themselves (\citei{Goodman:1994ul}, \citei{Pessah:2010ic}).
-The growth of parasitic modes provides a saturation avenue for channel mode-dominated flows, yet this is unlikely to be the dominant saturation mechanism in laboratory experiments or astrophysical disks, as channel modes are artificially over-represented in the shearing box (e.g.
-\citei{Latter:2015}).
+The growth of parasitic modes provides a saturation avenue for channel mode-dominated flows, yet this is unlikely to be the dominant saturation mechanism in laboratory experiments or astrophysical disks, as channel modes are artificially over-represented in the shearing box \citep{Latter:2015}.
 Thus while the shearing box may accurately approximate many features of the global MRI, the saturation mechanism may not be among them.
@@
 
-In this paper, we develop an analytic, global theory for the MRI in a Taylor-Couette flow.
+In this paper, we develop a weakly nonlinear, global theory for the MRI in a Taylor-Couette (TC) geometry.
 This system precludes channel modes, allowing us to develop an understanding of MRI saturation in their absence.
 A number of saturation mechanisms have been proposed for the MRI which do not rely on channel modes dominating the flow.
-The MRI feeds off of the free energy from differential rotation, and so a modification of the background shear may cause saturation (e.g.
-\citei{Knobloch:2005ba}, \citei{Umurhan:2007hs}).
-The MRI may transfer its free energy into the magnetic field, and saturate when the field is too strong to be susceptible to the MRI (e.g.
-\citei{Ebrahimi:2009ey}).
+The MRI feeds off of the free energy from differential rotation, and so a modification of the background shear may cause saturation \citep{Knobloch:2005ba, Umurhan:2007hs}.
+The MRI may transfer its free energy into the magnetic field, and saturate when the field is too strong to be susceptible to the MRI \citep{Ebrahimi:2009ey}.
 The MRI may saturate differently depending on the particular parameter regime under investigation, and so our challenge is not only in identifying possible saturation mechanisms, but in understanding how and when each applies in different astrophysical environments.
@@
 Unfortunately, detection of the MRI has so far proven elusive.
-\citei{Sisan:2004ig} claimed to detect the MRI in a spherical Couette flow, but most likely detected unrelated MHD instabilities instead (\citei{Hollerbach:2009ig}, \citei{Gissinger:2011td}).
-Most relevant to our work is the Princeton Plasma Physics Laboratory (PPPL) MRI Experiment, a liquid gallium Taylor-Couette flow with an axial magnetic field (\citei{Ji:2001kd}).
+\citei{Sisan:2004ig} claimed to detect the MRI in a spherical Couette flow, but most likely detected unrelated MHD instabilities instead \citep{Hollerbach:2009ig,Gissinger:2011td}.
+Most relevant to our work is the Princeton Plasma Physics Laboratory (PPPL) MRI Experiment, a liquid gallium TC flow with an axial magnetic field \citep{Ji:2001kd}.
 There has been some theoretical work designed to complement the Princeton MRI experiment involving direct numerical simulation of the experimental conditions, much of it focused on the specific challenges in identifying MRI signatures despite apparatus-driven spurious flows (e.g.
@@
 Our work assumes an infinite vertical domain, an idealization that is theoretically expedient but experimentally impractical.
-However, our setup is designed to be an accurate treatment of the radial dimension of the flow in a Taylor Couette apparatus like the one used in the Princeton MRI experiment.
+Such an approach changes the symmetry properties of the solution significantly, and in the much better studied hydrodynamic case this leads to significant differences even for TC devices with very large aspect ratios \citep{Lopez:2005}.
+However, our setup is designed to be an accurate treatment of the radial dimension of the flow in a TC apparatus like the one used in the Princeton MRI experiment.
 
@@
 Because we include curvature terms, our treatment also allows us to study the helical magnetorotational instability (HMRI).
-The HMRI is an overstability in which the background magnetic field is helical, $\mathbf{B} = B_0 (\xi \mathbf{\hat{\theta}} + \mathbf{\hat{z}})$ \citet{Hollerbach:2005tr}.
+The HMRI is an overstability in which the background magnetic field is helical, $\mathbf{B} = B_0 (\xi r/r_0 \mathbf{\hat{\phi}} + \mathbf{\hat{z}})$ \citet{Hollerbach:2005tr}.
 The HMRI currently occupies a special place in the MRI puzzle.
@@
 In accretion disks, differential rotation is imposed gravitationally by a central body, so the rotation profile is forced to be Keplerian.
-Clearly a gravitationally enforced Keplerian flow is inaccessible to laboratory study, so differential rotation is created by rotating an inner cylinder faster than an outer cylinder (a Taylor-Couette setup).
+Clearly a gravitationally enforced Keplerian flow is inaccessible to laboratory study, so differential rotation is created by rotating an inner cylinder faster than an outer cylinder (a TC setup).
 For a nonideal fluid subject to no-slip boundary conditions, the base flow is
@@
 a Ginzburg-Landau equation (GLE).
-The GLE governs the weakly nonlinear amplitude behavior in a wide range of physical systems, including the narrow gap MRI (\citei{Umurhan:2007hs}), Rayleigh-B\'enard convection (\citei{Newell:1969wr}), and hydrodynamic Taylor Couette flow (e.g.
+The GLE governs the weakly nonlinear amplitude behavior in a wide range of physical systems, including the narrow gap MRI (\citei{Umurhan:2007hs}), Rayleigh-B\'enard convection (\citei{Newell:1969wr}), and hydrodynamic TC flow (e.g.
 \citei{Recktenwald:1993}).
@@
 They find that this amplitude dependence is driven by the $\Pm^{1/3}$ dependence of the linear boundary layer.
-Boundary layer analysis similarly reveals a $\nu^{1/3}$ dependence for the radial extent of the boundary layer in Taylor Couette flow (\citei{Goodman:2002ix}).
+Boundary layer analysis similarly reveals a $\nu^{1/3}$ dependence for the radial extent of the boundary layer in TC flow (\citei{Goodman:2002ix}).
 %Why does the $\Pm$ scaling of the boundary layer width not translate to a stronger $\Pm$ scaling for the saturation amplitude in the wide gap case?
@@
 At the onset of convection the system undergoes a symmetry breaking, forming rolls, or convection cells, which break the horizontal translational invariance.
-Analogously, the standard MRI system considered here is initially vertically translationally symmetric, because we idealize the Taylor-Couette device as an infinitely long cylinder.
+Analogously, the standard MRI system considered here is initially vertically translationally symmetric, because we idealize the TC device as an infinitely long cylinder.
 The MRI breaks this symmetry, forming cells along the vertical length of the domain.
@@
 This work provides a mathematical description of the MRI as a pattern-forming process, but much remains to be understood, particularly involving the application of this model system to realistic astrophysical disks.
-Our model is most directly relevant to Taylor-Couette flows, and we emphasize that laboratory MRI experiments stand poised to observe the MRI-driven pattern formation predicted here.
+Our model is most directly relevant to TC flows, and we emphasize that laboratory MRI experiments stand poised to observe the MRI-driven pattern formation predicted here.
 
@@
 
+\bibitem[Lopez \& Marques(2005)]{Lopez:2005} 
+Lopez, J.~M., \& Marques, F.\ 2005, Physica D, 211, 168
+
 \bibitem[{Newell \& Whitehead(1969)}]{Newell:1969wr}

### 2016-09-02 paper/widegap_mri_paper.tex
more summary

--- 2016-09-02 paper/widegap_mri_paper.tex
@@
 
-where $\mathcal{D}$, $\mathcal{L}$, and $\widetilde{\mathcal{G}}$ are matrices defined in Appendix \ref{app:basic_equations}, and $\mathbf{N}$ is a vector containing all nonlinear terms defined in Appendix \ref{app:nonlinear_terms}.
-We expand the variables in a perturbation series $\mathbf{V} = \epsilon \mathbf{V}_1 + \epsilon^2 \mathbf{V}_2 + \epsilon^3 \mathbf{V}_3 + h.o.t.$ The perturbed system can then be expressed at each order by the equations
+where $\mathcal{D}$, $\mathcal{L}$, and $\widetilde{\mathcal{G}}$ are matrices defined in Appendix \ref{app:basic_equations}, and $\mathbf{N}$ is a vector containing all nonlinear terms.
+We expand the variables in a perturbation series
+
+\beq
+\mathbf{V} = \epsilon \mathbf{V}_1 + \epsilon^2 \mathbf{V}_2 + \epsilon^3 \mathbf{V}_3 + h.o.t.
+\eeq
+
+ The perturbed system can then be expressed at each order by the equations
 
@@
 
-See Appendix \ref{app:basic_equations} for the definition of matrices and nonlinear vectors, and a thorough derivation.
+The nonlinear terms $\mathbf{N}_2$ and $\mathbf{N}_3$ which appear at $\mathcal{O}(\epsilon^2)$ and $\mathcal{O}(\epsilon^3)$, respectively, contain the nonlinear interaction between MRI modes.
+The system is weakly nonlinear because this mode interaction occurs in a controlled way.
+At $\mathcal{O}(\epsilon^2)$, the nonlinear terms represent the interaction of linear ($\mathcal{O}(\epsilon)$) MRI modes with themselves and their complex conjugates.
+At $\mathcal{O}(\epsilon^3)$, the nonlinear terms contain the interaction between first- and second-order MRI modes.
+See Appendix \ref{app:basic_equations} for the definition of matrices and a thorough derivation, and Appendix \ref{app:nonlinear_terms} for the detailed form of the nonlinear vectors.
 We emphasize that Equations \ref{eq:ordere} - \ref{eq:ordere3} have the same form as these equations in the narrow gap case, although the matrices, which contain all radial derivatives, are significantly different in this wide gap formulation.
@@
 
-We solve the equations at each order using Dedalus, an open source pseudospectral code (Burns et al.
-in prep).
+We solve the equations at each order using Dedalus, an open source pseudospectral code.
 We solve the radial portion of the eigenvectors on a basis of Chebyshev polynomials subject to our radial boundary conditions.
 %Because our domain is bounded only radially and not vertically, we solve the radial dimension on a basis of Chebyshev polynomials, and the vertical dimension on a Fourier series basis.
-We solve Equation \ref{eq:ordere} as a linear eigenvalue problem, and Equation \ref{eq:ordere2} as a linear boundary value problem. %should i mention spurious eigenvals? 
+We solve Equation \ref{eq:ordere} as a linear eigenvalue problem, and Equation \ref{eq:ordere2} as a linear boundary value problem. 
+ %should i mention spurious eigenvals? 
 %To analyze the MRI system at marginality, we fix the parameters listed in Table \ref{table:parameters} and determine the critical $\reym$ and $k_z$ by repeatedly solving Equation \ref{eq:ordere} to determine the smallest parameter values for which the fastest growing mode is zero.
@@
 \subsection{Helical MRI}
-\label{sec:HMRI}
 When $\xi$ in Equation \ref{eq:unperturbed_matrix_eqns} is not equal to zero, the helical MRI arises.
@@
 For our fiducial parameters, the marginal mode has a frequency $\omega = 0.153$.
-This means that the helical MRI modes are traveling waves, moving in the $z$ direction with a phase velocity $\omega/k_c$ = .
+This means that the helical MRI modes are traveling waves, moving in the $z$ direction with a phase velocity $\omega/k_c$.
 
@@
 \label{sec:discussion}
+In this work we carry out a formal weakly nonlinear multiscale analysis of the MRI in a Taylor Couette flow.
+We analyze both the standard and helical MRI, which differ only in the geometry of their imposed background magnetic fields.
+We find that the amplitude function, which governs the behavior of the system on long length- and timescales, obeys a real GLE for the case of the standard MRI, and a complex GLE for the helical MRI.
+These two systems are thus subject to different large-scale phase dynamics.
+
 Our work should be placed in the broader context of emergent pattern formation in physical systems.
@@
 Our model is most directly relevant to Taylor-Couette flows, and we emphasize that laboratory MRI experiments stand poised to observe the MRI-driven pattern formation predicted here.
+
+This is the first weakly nonlinear analysis of the MRI in a cylindrical geometry, and the theory developed here should be compared to analytic studies of the MRI in local approximations \citep{Umurhan:2007hs,Vasil:2015}.
+We find that the saturation amplitude of the standard MRI scales somewhat more steeply with $\Pm$ than does this same system system in the narrow gap approximation.
 
@@
 
+\bibitem[{Vasil(2015)}]{Vasil:2015}
+Vasil, G. M. 2015, RSPSA 471, 20140699
+
 \bibitem[{Willis \& Barenghi(2002)}]{Willis:2002bh}

### 2016-09-02 paper/widegap_mri_paper.tex

--- 2016-09-02 paper/widegap_mri_paper.tex
@@
 Our model is most directly relevant to TC flows, and we emphasize that laboratory MRI experiments stand poised to observe the MRI-driven pattern formation predicted here.
+
+This is the first weakly nonlinear analysis of the MRI in a cylindrical geometry, and the theory developed here should be compared to analytic studies of the MRI in local approximations \citep{Umurhan:2007hs,Vasil:2015}.
+We find that the saturation amplitude of the standard MRI scales somewhat more steeply with $\Pm$ than does this same system system in the narrow gap approximation.
 

### 2016-09-02 paper/widegap_mri_paper.tex
possible ways forward in note form

@@
 \label{sec:discussion}
-In this work we carry out a formal weakly nonlinear multiscale analysis of the MRI in a Taylor Couette flow.
+In this work we carry out a formal weakly nonlinear multiscale analysis of the MRI in a Taylor-Couette flow.
 We analyze both the standard and helical MRI, which differ only in the geometry of their imposed background magnetic fields.
@@
 
-This is the first weakly nonlinear analysis of the MRI in a cylindrical geometry, and the theory developed here should be compared to analytic studies of the MRI in local approximations \citep{Umurhan:2007hs,Vasil:2015}.
+This is the first weakly nonlinear analysis of the MRI in a cylindrical geometry, and is thus the global analogue of similar analyses in local approximations \citep{Umurhan:2007hs,Vasil:2015}.
 We find that the saturation amplitude of the standard MRI scales somewhat more steeply with $\Pm$ than does this same system system in the narrow gap approximation.
@@
 The theory presented here may be extended in several ways. 
+- detailed comparison to specific experimental apparatus. will gle dynamics be visible? 
+- the vertical dimension is incompletely modeled - stratification, etc
+- explore saturation property dependence of system on rotation profile 
+- different field geometries (e.g. radially varying)
+- other radial boundary conditions
+- drop axisymmetry
+- comparison to simulation in both weakly and strongly nonlinear regimes
 

### 2016-09-02 paper/widegap_mri_paper.tex
with line breaks

@@
 
-The theory presented here may be extended in several ways. 
-- detailed comparison to specific experimental apparatus. will gle dynamics be visible? 
-- the vertical dimension is incompletely modeled - stratification, etc
-- explore saturation property dependence of system on rotation profile 
-- different field geometries (e.g. radially varying)
-- other radial boundary conditions
-- drop axisymmetry
-- comparison to simulation in both weakly and strongly nonlinear regimes
+The theory presented here may be extended in several ways. \\
+- detailed comparison to specific experimental apparatus. will gle dynamics be visible? \\
+- the vertical dimension is incompletely modeled - stratification, etc\\
+- explore saturation property dependence of system on rotation profile\\ 
+- different field geometries (e.g. radially varying)\\
+- other radial boundary conditions\\
+- drop axisymmetry\\
+- comparison to simulation in both weakly and strongly nonlinear regimes\\
 

### 2016-09-02 paper/widegap_mri_paper.tex
make future work an itemized list.

--- 2016-09-02 paper/widegap_mri_paper.tex
@@
 \makeatletter
-\setlength{\@fptop}{0pt}
+\setlength{\@fptop}{0pt}1
 \setlength{\@fpbot}{0pt}
@@
 Such an approach changes the symmetry properties of the solution significantly, and in the much better studied hydrodynamic case this leads to significant differences even for TC devices with very large aspect ratios \citep{Lopez:2005}.
-However, our setup is designed to be an accurate treatment of the radial dimension of the flow in a TC apparatus like the one used in the Princeton MRI experiment.
-
-This realistic radial treatment means that we account for the curvature of the flow in a cylindrical apparatus.
-Many investigations of the MRI use the ``narrow gap" approximation, in which the radial extent of the fluid channel is taken to be much smaller than the radius of curvature.
+Nevertheless, this study represents a first step in understanding the saturation of global, MRI unstable TC flow without the additional complication of vertical endcap effects.
+
+Our radial treatment includes the curvature of the flow in a cylindrical apparatus.
+Many investigations of the MRI use the ``narrow gap'' approximation (the shearing box is a narrow gap without boundary walls), in which the radial extent of the fluid channel is taken to be much smaller than the radius of curvature.
 That is, for a center channel radius $r_0$ bounded by inner and outer radii $r_1$ and $r_2$, respectively, the narrow gap approximation applies when $r_0 \gg (r_2 - r_1)$.
@@
 
-The theory presented here may be extended in several ways. 
-- detailed comparison to specific experimental apparatus. will gle dynamics be visible? 
-- the vertical dimension is incompletely modeled - stratification, etc
-- explore saturation property dependence of system on rotation profile 
-- different field geometries (e.g. radially varying)
-- other radial boundary conditions
-- drop axisymmetry
-- comparison to simulation in both weakly and strongly nonlinear regimes
+There are several avenues for future work, including
+\begin{itemize}
+\item detailed comparison to specific experimental apparatus. will gle dynamics be visible? 
+\item the vertical dimension is incompletely modeled - stratification, etc
+\item explore saturation property dependence of system on rotation profile 
+\item different field geometries (e.g. radially varying)
+\item other radial boundary conditions
+\item drop axisymmetry
+\item comparison to simulation in both weakly and strongly nonlinear regimes
+\end{itemize}
 

### 2016-09-02 paper/widegap_mri_paper.tex
updated future work bullet points.

@@
 \begin{itemize}
-\item detailed comparison to specific experimental apparatus. will gle dynamics be visible? 
-\item the vertical dimension is incompletely modeled - stratification, etc
-\item explore saturation property dependence of system on rotation profile 
+\item detailed comparison to a specific experimental apparatus. Will gle dynamics be visible? 
+\item inclusion of vertical stratification to the base state
 \item different field geometries (e.g. radially varying)
-\item other radial boundary conditions
-\item drop axisymmetry
+\item comparing the saturation properties of different rotation profiles
+\item expanding radial boundary conditions to mimimic astrophysical disks rather than TC devices
 \item comparison to simulation in both weakly and strongly nonlinear regimes

### 2016-09-02 paper/widegap_mri_paper.tex
fixed typo.
future work bullets expanded

--- 2016-09-02 paper/widegap_mri_paper.tex
@@
 \makeatletter
-\setlength{\@fptop}{0pt}1
+\setlength{\@fptop}{0pt}
 \setlength{\@fpbot}{0pt}
@@
 
-There are several avenues for future work, including
+We detail several avenues for future work, which highlight the application of this theory to both laboratory experiments and astrophysical disks.
+
 \begin{itemize}
-\item detailed comparison to specific experimental apparatus. will gle dynamics be visible? 
-\item the vertical dimension is incompletely modeled - stratification, etc
-\item explore saturation property dependence of system on rotation profile 
-\item different field geometries (e.g. radially varying)
-\item other radial boundary conditions
-\item drop axisymmetry
-\item comparison to simulation in both weakly and strongly nonlinear regimes
+\item This theory may be applied to a specific experimental apparatus to model the predicted saturated state.
+One can then ask whether GLE dynamics should be detectable, especially over apparatus-driven flows.
+\item The saturation properties of different rotation profiles may be compared by direct application of the theory developed here.
+\item Vertical stratification may be added to the base state, which would be a more realistic model of global vertical disk structure. 
+\item Other nonideal MHD effects such as the Hall effect and ambipolar diffusion may be straightforwardly added to this model, and are of particular interest for modeling protoplanetary disks. 
+\item The background magnetic field geometry may be generalized to include radial variation.
+\item The radial boundary conditions considered here may be expanded to mimic astrophysical disks rather than TC devices.
+\item This theory may be compared to simulations in both the weakly and strongly nonlinear regimes. 
 \end{itemize}
 
-\section{Acknowledgements}
+\section{Acknowledgments}
 S.E.C.

### 2016-09-02 paper/widegap_mri_paper.tex
HMRI -> helical MRI

@@
 \begin{document}
-
 %\title{Weakly Nonlinear Analysis of the Standard and Helical Magnetorotational Instabilities in a Taylor Couette Flow}
@@
 
-Because we include curvature terms, our treatment also allows us to study the helical magnetorotational instability (HMRI).
-The HMRI is an overstability in which the background magnetic field is helical, $\mathbf{B} = B_0 (\xi r/r_0 \mathbf{\hat{\phi}} + \mathbf{\hat{z}})$ \citet{Hollerbach:2005tr}.
-The HMRI currently occupies a special place in the MRI puzzle.
-The HMRI has been proposed as a method of awakening angular momentum transport in the ``dead zones" of protoplanetary disks where the $\Pm$ becomes very small.
-However the rotation profiles needed to excite HMRI may be steeper than Keplerian, depending on the boundary conditions, and so its role in astrophysical disks is currently a matter of debate \citep{Liu:2006,Rudiger:2007,Kirillov:2013}.
-Regardless of its astrophysical role, the HMRI is significantly easier to excite in a laboratory setting than the standard MRI, and has already been detected by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
+Because we include curvature terms, our treatment also allows us to study the helical magnetorotational instability.
+The helical MRI is an overstability in which the background magnetic field is helical, $\mathbf{B} = B_0 (\xi r/r_0 \mathbf{\hat{\phi}} + \mathbf{\hat{z}})$ \citet{Hollerbach:2005tr}.
+The helical MRI currently occupies a special place in the MRI puzzle.
+The helical MRI has been proposed as a method of awakening angular momentum transport in the ``dead zones" of protoplanetary disks where the $\Pm$ becomes very small.
+However the rotation profiles needed to excite helical MRI may be steeper than Keplerian, depending on the boundary conditions, and so its role in astrophysical disks is currently a matter of debate \citep{Liu:2006,Rudiger:2007,Kirillov:2013}.
+Regardless of its astrophysical role, the helical MRI is significantly easier to excite in a laboratory setting than the standard MRI, and has already been detected by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
 
@@
 
-\section{Wide Gap Equations}
+\section{Basic Framework}
 \label{sec:wide_gap}
@@
 We choose the SMRI parameters to be comparable to the case considered in \citei{Goodman:2002ix}.
-The HMRI parameters were chosen to be comparable to \citei{Hollerbach:2005tr}.
+The helical MRI parameters were chosen to be comparable to \citei{Hollerbach:2005tr}.
 Our fiducial parameters are described in Table \ref{table:parameters}.

### 2016-09-02 paper/widegap_mri_paper.tex
put equations in gaussian units.

@@
 %\begin{split}
-\partial_t \mathbf{u} + \mathbf{u} \cdot \nabla \mathbf{u} = -\frac{1}{\rho}\nabla P - \nabla\Phi + \frac{1}{\rho} \left(\mathbf{J}\times\mathbf{B}\right) + \nu\nabla^2 \mathbf{u} 
+\partial_t \mathbf{u} + \mathbf{u} \cdot \nabla \mathbf{u} = -\frac{1}{\rho}\nabla P - \nabla\Phi + \frac{1}{c\rho} \left(\mathbf{J}\times\mathbf{B}\right) + \nu\nabla^2 \mathbf{u} 
 %\end{split}
@@
 
-where $P$ is the gas pressure, $\nu$ is the kinematic viscosity, $\eta$ is the microscopic diffusivity, $\nabla\Phi$ is the gravitational force per unit mass, and the current density is $\mathbf{J} = \nabla\times\mathbf{B}$.
+where $P$ is the gas pressure, $\nu$ is the kinematic viscosity, $\eta$ is the microscopic diffusivity, $\nabla\Phi$ is the gravitational force per unit mass, and the current density is $\mathbf{J} = c \nabla\times\mathbf{B}/4\pi$.
 We solve these equations subject to the incompressible fluid and solenoidal magnetic field constraints,

### 2016-09-03 python/pleiades_run_thingap_amplitude.sh
UMR params with Pm=1E-4 thingap at 256

@@ -3,7 +3,7 @@
 #PBS -l select=2:ncpus=10:mpiprocs=10:model=has
 #PBS -l walltime=02:00:00
 #PBS -j oe
-#PBS -q normal
+#PBS -q devel
 
 dedalus_script=run_amplitude
 

### 2016-09-04 paper/widegap_mri_paper.tex
UMR params with Pm=1E-4 thingap at 256
added final versions of figures

@@
 \centering
-\includegraphics[width=\textwidth]{../figures/widegap_narrowgap_comparison_narrowfiducial.png}
+\includegraphics[width=\textwidth]{../figures/widegap_thingap_comparison_zoombox.png}
 \caption{Nonlinear term $N_{31}^{(B)}$ for the wide gap (left) and narrow gap (right) standard MRI.
@@
 \label{sec:results}
-\begin{figure*}
+
+%\begin{figure*}
+%\centering
+%\includegraphics[width=\textwidth]{../figures/fiducial_standard_mri_stressplot_Pm_16E-7_cc.png}
+%\caption{Reynolds ($\mathbb{T}_{R} = u_r u_\phi$), Maxwell ($\mathbb{T}_{M} = -\frac{2}{\beta} B_r B_\phi$), %and total stress ($\mathbb{T} = \mathbb{T}_{R} + \mathbb{T}_{M}$) for the fiducial standard MRI case.
+}
+%\end{figure*}
+
+\begin{figure}
 \centering
-\includegraphics[width=\textwidth]{../figures/fiducial_standard_mri_stressplot_Pm_16E-7.png}
+\includegraphics[width=\columnwidth]{../figures/widegap_totalstress_intstress.png}
 \caption{Reynolds ($\mathbb{T}_{R} = u_r u_\phi$), Maxwell ($\mathbb{T}_{M} = -\frac{2}{\beta} B_r B_\phi$), and total stress ($\mathbb{T} = \mathbb{T}_{R} + \mathbb{T}_{M}$) for the fiducial standard MRI case.
 }
-\end{figure*}
+\end{figure}
 
@@
 %\end{figure}
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../figures/widegap_vel_Bfield_streamfuncs_512_final_wider.png}
+\caption{sdfsdfs }
+\end{figure*}
 

### 2016-09-04 paper/widegap_mri_paper.tex
hmri streamfunc plot

@@
 \author{Oishi, J.S. \altaffilmark{2,}\altaffilmark{3}}
-%\author{Mac Low, M.-M.\altaffilmark{1,}\altaffilmark{2}}
 \altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
@@
 MRI saturation is a complicated nonlinear problem which may depend on the assumptions and approximations adopted by simulations in nonobvious ways.
-For example, the magnetic Prandtl number $\Pm = \nu/\eta \sim 10^{-8}$ in protoplanetary disks \citet[e.g.][]{OishiMacLow:2011} and $\sim 10^{-6}$ in liquid metal experiments \citep[e.g.][]{Goodman:2002ix}.
+For example, the magnetic Prandtl number $\Pm = \nu/\eta \sim 10^{-8}$ in protoplanetary disks \citep[e.g.][]{Oishi:2011ei} and $\sim 10^{-6}$ in liquid metal experiments \citep[e.g.][]{Goodman:2002ix}.
 Such extreme ratios of viscosity to resistivity far exceed current computational resources.
-However, we can construct asymptotic approximations valid for $\Pm << 1$ using analytic methods.
+However, we can construct asymptotic approximations valid for $\Pm \ll 1$ using analytic methods.
 Analytic methods can also play a powerful role in elucidating the mechanisms responsible for MRI saturation.
@@
 The MRI may saturate differently depending on the particular parameter regime under investigation, and so our challenge is not only in identifying possible saturation mechanisms, but in understanding how and when each applies in different astrophysical environments.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../figures/widegap_paramspace_crit_params_Pm_16E-7}
+\caption{Growth rates in the ($\reym, k_z$) plane.
+Color map shows growth rate found by solving the linear eigenvalue problem for each ($\reym, k_z$) in the grid.
+The eigenvalue problem was solved for the widegap parameters listed in Table \ref{table:parameters}.
+Overlaid contours show growth rates at [ -8E-4, -1.3E-4, 1.3E-4, 8E-4, 1.5E-3], where dashed contours represent negative values.
+The gray dotted line shows the interpolated marginal stability curve.
+The critical parameters $\reym_c = 3.31$ and $k_c = 0.902$ correspond to the smallest parameter values that yield a zero growth rate.}\label{fig:growth_rates}
+\end{figure}
 
@@
 
-Our radial treatment includes the curvature of the flow in a cylindrical apparatus.
-Many investigations of the MRI use the ``narrow gap'' approximation (the shearing box is a narrow gap without boundary walls), in which the radial extent of the fluid channel is taken to be much smaller than the radius of curvature.
-That is, for a center channel radius $r_0$ bounded by inner and outer radii $r_1$ and $r_2$, respectively, the narrow gap approximation applies when $r_0 \gg (r_2 - r_1)$.
-The narrow gap approximation simplifies the MRI equations by excluding curvature terms, because the flow through a narrow gap can be taken to be approximately linear in $\phi$, i.e.
-Cartesian.
-Previous investigations into the weakly nonlinear behavior of the MRI have used this narrow gap approximation (\citei{Umurhan:2007dz}, \citei{Umurhan:2007hs}, \citei{Clark:2016}).
-In this work we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.
-
-Because we include curvature terms, our treatment also allows us to study the helical magnetorotational instability.
-The helical MRI is an overstability in which the background magnetic field is helical, $\mathbf{B} = B_0 (\xi r/r_0 \mathbf{\hat{\phi}} + \mathbf{\hat{z}})$ \citet{Hollerbach:2005tr}.
-The helical MRI currently occupies a special place in the MRI puzzle.
-The helical MRI has been proposed as a method of awakening angular momentum transport in the ``dead zones" of protoplanetary disks where the $\Pm$ becomes very small.
-However the rotation profiles needed to excite helical MRI may be steeper than Keplerian, depending on the boundary conditions, and so its role in astrophysical disks is currently a matter of debate \citep{Liu:2006,Rudiger:2007,Kirillov:2013}.
-Regardless of its astrophysical role, the helical MRI is significantly easier to excite in a laboratory setting than the standard MRI, and has already been detected by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
-
-In this work we explore the behavior of the viscous, dissipative MRI in a cylindrical geometry close to threshold, making explicit comparisons to the standard MRI behavior in the thin-gap regime.
-We investigate both the standard MRI, in which the background magnetic field is purely axial, as well as the helical MRI.
-In section~\ref{sec:wide_gap}, we lay out the basic mathematical framework of the problem.
-In section~\ref{sec:wnl_analysis}, we introduce the method of multiple scales we use to construct our theory.
-In section~\ref{sec:results} we describe the basic results, and in section~\ref{sec:discussion} we place them in the context of previous work on other instabilities, discuss their relevance to experiments, and reiterate our final conclusions.
-
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{../figures/widegap_paramspace_crit_params_Pm_16E-7}
-\caption{Growth rates in the ($\reym, k_z$) plane.
-Color map shows growth rate found by solving the linear eigenvalue problem for each ($\reym, k_z$) in the grid.
-The eigenvalue problem was solved for the widegap parameters listed in Table \ref{table:parameters}.
-Overlaid contours show growth rates at [ -8E-4, -1.3E-4, 1.3E-4, 8E-4, 1.5E-3], where dashed contours represent negative values.
-The gray dotted line shows the interpolated marginal stability curve.
-The critical parameters $\reym_c = 3.31$ and $k_c = 0.902$ correspond to the smallest parameter values that yield a zero growth rate.}\label{fig:growth_rates}
-\end{figure}
-
 \begin{figure*}
@@
 Adjoint homogenous eigenfunctions are normalized such that $\langle V_{11}^\dagger \cdot \mathcal{D} V_{11} \rangle = 1$.}\label{fig:linear_eigenfunctions}
+\end{figure*}
+
+Our radial treatment includes the curvature of the flow in a cylindrical apparatus.
+Many investigations of the MRI use the ``narrow gap'' approximation (the shearing box is a narrow gap without boundary walls), in which the radial extent of the fluid channel is taken to be much smaller than the radius of curvature.
+That is, for a center channel radius $r_0$ bounded by inner and outer radii $r_1$ and $r_2$, respectively, the narrow gap approximation applies when $r_0 \gg (r_2 - r_1)$.
+The narrow gap approximation simplifies the MRI equations by excluding curvature terms, because the flow through a narrow gap can be taken to be approximately linear in $\phi$, i.e.
+Cartesian.
+Previous investigations into the weakly nonlinear behavior of the MRI have used this narrow gap approximation (\citei{Umurhan:2007dz}, \citei{Umurhan:2007hs}, \citei{Clark:2016}).
+In this work we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.
+
+Because we include curvature terms, our treatment also allows us to study the helical magnetorotational instability.
+The helical MRI is an overstability in which the background magnetic field is helical, $\mathbf{B} = B_0 (\xi r/r_0 \mathbf{\hat{\phi}} + \mathbf{\hat{z}})$ \citet{Hollerbach:2005tr}.
+The helical MRI currently occupies a special place in the MRI puzzle.
+The helical MRI has been proposed as a method of awakening angular momentum transport in the ``dead zones" of protoplanetary disks where the $\Pm$ becomes very small.
+However the rotation profiles needed to excite helical MRI may be steeper than Keplerian, depending on the boundary conditions, and so its role in astrophysical disks is currently a matter of debate \citep{Liu:2006,Rudiger:2007,Kirillov:2013}.
+Regardless of its astrophysical role, the helical MRI is significantly easier to excite in a laboratory setting than the standard MRI, and has already been detected by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
+
+In this work we explore the behavior of the viscous, dissipative MRI in a cylindrical geometry close to threshold, making explicit comparisons to the standard MRI behavior in the thin-gap regime.
+We investigate both the standard MRI, in which the background magnetic field is purely axial, as well as the helical MRI.
+In section~\ref{sec:wide_gap}, we lay out the basic mathematical framework of the problem.
+In section~\ref{sec:wnl_analysis}, we introduce the method of multiple scales we use to construct our theory.
+In section~\ref{sec:results} we describe the basic results, and in section~\ref{sec:discussion} we place them in the context of previous work on other instabilities, discuss their relevance to experiments, and reiterate our final conclusions.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../figures/widegap_n2_n3_Pm16E-7.pdf}
+\caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.
+These are nonlinear combinations of lower-order eigenfunctions, representing the weakly nonlinear mode interaction.
+At second order ($N_2$) the most unstable linear MRI mode interacts with itself and its complex conjugate.
+At third order ($N_3$) the first and second order MRI modes interact with each other.}\label{fig:N2_N3}
 \end{figure*}
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{../figures/widegap_n2_n3_Pm16E-7.pdf}
-\caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.
-These are nonlinear combinations of lower-order eigenfunctions, representing the weakly nonlinear mode interaction.
-At second order ($N_2$) the most unstable linear MRI mode interacts with itself and its complex conjugate.
-At third order ($N_3$) the first and second order MRI modes interact with each other.}\label{fig:N2_N3}
-\end{figure*}
-
 Thus through judicious choice of cylinder rotation rates, one can set $q(r_0) = 3/2$, for quasi-Keplerian flow.
@@
 \includegraphics[width=\textwidth]{../figures/widegap_vel_Bfield_streamfuncs_512_final_wider.png}
-\caption{sdfsdfs }
+\caption{Perturbation structure for the velocity and magnetic field of the fiducial standard MRI case, including first and second order perturbations.
+Leftmost panel is a radially zoomed-in section of the velocity perturbation structure, to better show the boundary layer-driven structure at the inner cylinder.
+Colors are azimuthal velocity and magnetic field perturbations, and streamfunctions show the perturbation structure in the $r, z$ plane.
+The width of the streamfunctions is proportional to the speed and magnetic field strength in the $r, z$ plane for the velocity and magnetic field, respectively.}
 \end{figure*}
@@
 \subsection{Helical MRI}
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../figures/hmri_velocity_Bfield_streamfuncs.png}
+\caption{sdfsdfs }
+\end{figure*}
+
 When $\xi$ in Equation \ref{eq:unperturbed_matrix_eqns} is not equal to zero, the helical MRI arises.
@@
 
+\bibitem[{Oishi \& Mac Low(2011)}]{Oishi:2011ei}
+Oishi, J.S. \& Mac Low, M.M., 2011, ApJ 740, 18
+
 \bibitem[{Parker \& Krommes(2013)}]{Parker:2013hy}

### 2016-09-05 paper/widegap_mri_paper.tex
final edits

@@
 The magnetorotational instability (MRI) is widely belived to drive angular momentum transport in astrophysical disks.
-The MRI is a local instability excited by weak magnetic fields in differentially rotating fluids, and since its discovery \citet{Balbus:1991vs} it has been widely invoked to explain accretion in protoplanetary disks (see \citei{Armitage:2010} and references therein), binary systems (), and disks around black holes (), as well as jet and wind launching (\citei{Lesur:2013}), dynamos, etc (cite these).
+The MRI is a local instability excited by weak magnetic fields in differentially rotating fluids, and since its discovery \citet{Balbus:1991vs} it has been widely invoked to explain accretion in protoplanetary disks \citep{Armitage:2010} and disks around black holes \citep{Blaes:2014}, as well as jet and wind launching (\citei{Lesur:2013}) anisotropic turbulence \citep{Murphy:2015tn}, and dynamos \citep{Vishniac:2009il} among other phenomena.
 
@@
 \caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.
-These are nonlinear combinations of lower-order eigenfunctions, representing the weakly nonlinear mode interaction.
+These are nonlinear combinations of lower-order eigenfunctions.
 At second order ($N_2$) the most unstable linear MRI mode interacts with itself and its complex conjugate.
@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../figures/widegap_thingap_comparison_zoombox.png}
+\caption{Nonlinear term $N_{31}^{(A)}$ for the wide gap (left) and narrow gap (right) standard MRI, where the wide gap is the TC flow considered in this work.
+Terms shown span three orders of magnitude in $\Pm$.
+The wide gap vectors represent runs using the parameters in Table \ref{table:parameters} and $\Pm = 1.6E-4, 1.6E-5, 1.6E-6$.
+The narrow gap MRI runs use the fiducial parameters in \citet{Umurhan:2007hs}, with $\Pm = 1E-4, 1E-5, 1E-6$.
+Inlaid plots show zoomed-in views of boundary layers at the inner boundary.
+The wide gap case displays dramatic boundary layers only at the inner boundary, but boundary layers in the thin gap approximation are symmetric about the origin because MRI modes in the narrow gap approximation are eigenstates of parity.}\label{fig:thin_wide_comparison}
+\end{figure*}
+
 \section{Weakly Nonlinear Perturbation Analysis}\label{sec:wnl_analysis}
 \label{sec:Perturbation}
+
 \begin{figure}
@@
 
+We find the marginal system as a function of the dimensionless parameters.
+The marginal stability curve for our standard MRI system is a hyperplane in $(\reym, \Pm, \beta, \Omega_2/\Omega_1, \mathrm{R}_1/\mathrm{R}_2$), but we hold all of these constant except for $\reym$.
+To analyze the MRI system at marginality, we fix the parameters listed in Table \ref{table:parameters} and determine the critical $\reym$ and vertical wavenumber $k_z$ by repeatedly solving the linear MRI system to determine the smallest parameter values for which the fastest growing mode has zero growth rate.
+That is, we solve the linear eigenvalue problem for eigenvalues $\sigma = \gamma + i \omega$ and determine the parameters which yield $\gamma = 0$.
+Figure \ref{fig:growth_rates} shows linear MRI growth rates $\gamma$ in the $(\reym, k_z)$ plane.
+For the fiducial standard MRI parameters in Table \ref{table:parameters} we find critical parameters $\reym_c = 3.30$ and $k_c = 0.901$.
+
 \begin{figure*}
 \centering
-\includegraphics[width=\textwidth]{../figures/widegap_thingap_comparison_zoombox.png}
-\caption{Nonlinear term $N_{31}^{(B)}$ for the wide gap (left) and narrow gap (right) standard MRI.
-Each line represents a run for a different $\Pm$, from $\Pm = 1E-4$ (darkest) to $\Pm \sim 1E-3$ (lightest).
-Inlaid plot in wide gap case shows a zoomed-in view of the boundary layer at the outer boundary ($r_2$).
-In the narrow gap case the boundary layer strongly affects the bulk of the flow, while in the wide gap case the flow in the center of the channel is relatively unaffected by width of the boundary layers.}\label{fig:wide_narrow_N31B}
+\includegraphics[width=\textwidth]{../figures/widegap_vel_Bfield_streamfuncs_512_final_wider.png}
+\caption{Perturbation structure for the velocity and magnetic field of the fiducial standard MRI case, including first and second order perturbations.
+Leftmost panel is a radially zoomed-in section of the velocity perturbation structure, to better show the boundary layer-driven structure at the inner cylinder.
+Colors are azimuthal velocity and magnetic field perturbations, and streamfunctions show the perturbation structure in the $r, z$ plane.
+The width of the streamfunctions is proportional to the speed and magnetic field strength in the $r, z$ plane for the velocity and magnetic field, respectively.
+Vertical domain covers one critical wavelength $\lambda_c = 2\pi / k_c$.
+We use the constant saturation amplitude $\alpha_{s} = 3.9 \times 10^{-5}$ derived for this case, and a small parameter $\epsilon = 0.5$.}\label{fig:smri_streamfuncs}
 \end{figure*}
 
-We find the marginal system as a function of the dimensionless parameters.
-The marginal stability curve for our standard MRI system is a hyperplane in $(\reym, \Pm, \beta, \Omega_2/\Omega_1, \mathrm{R}_1/\mathrm{R}_2$), but we hold all of these constant except for $\reym$.
-To analyze the MRI system at marginality, we fix the parameters listed in Table \ref{table:parameters} and determine the critical $\reym$ and $k_z$ by repeatedly solving the linear MRI system to determine the smallest parameter values for which the fastest growing mode is zero.
-That is, we solve the linear eigenvalue problem for eigenvalues $\sigma = \gamma + i \omega$.
-Figure \ref{fig:growth_rates} shows linear MRI growth rates $\gamma$ in the $(\reym, k_z)$ plane.
-For the fiducial standard MRI parameters in Table \ref{table:parameters} we find critical parameters $\reym_c = 3.30$ and $k_c = 0.901$.
-
-Just as in the weakly nonlinear analyses of \citei{Umurhan:2007hs} and \citei{Clark:2016}, we tune the system away from marginality by taking $B_0 \rightarrow B_0\left(1 - \epsilon^2\right)$, where $\epsilon \ll 1$.
+Just as in the weakly nonlinear analyses of \citei{Umurhan:2007hs} and \citei{Clark:2016}, we tune the system away from marginality by taking $B_0 \rightarrow B_0\left(1 - \epsilon^2\right)$, where the small parameter $\epsilon \ll 1$.
 We parameterize scale separation as $Z = \epsilon z$ and $T = \epsilon^2 t$, where $Z$ and $T$ are slowly varying spatial and temporal scales, respectively.
@@
 \mathcal{O}(\epsilon)&: \mathcal{L} \mathbf{V}_1 + \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \mathcal{D} \partial_t \mathbf{V}_1 = 0 \label{eq:ordere}\\
-\mathcal{O}(\epsilon^2)&: \mathcal{L} \mathbf{V}_2 + \xi \widetilde{\mathcal{H}} \mathbf{V}_2 + \mathcal{D} \partial_t \mathbf{V}_2 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_1 + \xi \mathcal{H} \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \label{eq:ordere2}\\
-\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \xi \widetilde{\mathcal{H}} \mathbf{V}_3 + \mathcal{D} \partial_t \mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2\nonumber \\
-&+ \xi \mathcal{H}\partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 - \xi \widetilde{\mathcal{H}} \mathbf{V}_1 + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3 = 0.
-\label{eq:ordere3}\\
+\mathcal{O}(\epsilon^2)&: \mathcal{L} \mathbf{V}_2 + \xi \widetilde{\mathcal{H}} \mathbf{V}_2 + \mathcal{D} \partial_t \mathbf{V}_2 + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_1 \nonumber\\
+& + \xi \mathcal{H} \partial_Z \mathbf{V}_1 + \mathbf{N}_2 = 0 \label{eq:ordere2}\\
+\mathcal{O}(\epsilon^3)&: \mathcal{L}\mathbf{V}_3 + \xi \widetilde{\mathcal{H}} \mathbf{V}_3 + \mathcal{D} \partial_t \mathbf{V}_3 + \mathcal{D} \partial_T \mathbf{V}_1 \nonumber\\
+&+ \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_2 + \xi \mathcal{H}\partial_Z \mathbf{V}_2 + \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V}_1 - \xi \widetilde{\mathcal{H}} \mathbf{V}_1 \nonumber\\
+& + \widetilde{\mathcal{G}} \mathbf{V}_1 + \mathbf{N}_3 = 0.\label{eq:ordere3}
 \end{align}
@@
 
-The result of the weakly nonlinear analysis is a single amplitude equation for $\alpha$. This amplitude equation is found by enforcing a solvability criterion on Equation \ref{eq:ordere3}. We find
+The result of the weakly nonlinear analysis is a single amplitude equation for $\alpha$. This amplitude equation is found by enforcing a solvability condition on Equation \ref{eq:ordere3}. We find
 
@@
 \includegraphics[width=\columnwidth]{../figures/widegap_totalstress_intstress.png}
-\caption{Reynolds ($\mathbb{T}_{R} = u_r u_\phi$), Maxwell ($\mathbb{T}_{M} = -\frac{2}{\beta} B_r B_\phi$), and total stress ($\mathbb{T} = \mathbb{T}_{R} + \mathbb{T}_{M}$) for the fiducial standard MRI case.
-}
+\caption{Total stress $\mathrm{T}(r, z) = u_r u_\phi - \frac{2}{\beta} B_r B_\phi$ , i.e.
+the sum of the Reynolds and Maxwell stresses for the fiducial standard MRI parameters (top panel).
+Bottom panel shows the vertically integrated stress $\int \mathrm{T}(r, z) \, dz$.}\label{fig:total_stress}
 \end{figure}
@@
 %\end{figure}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{../figures/widegap_vel_Bfield_streamfuncs_512_final_wider.png}
-\caption{Perturbation structure for the velocity and magnetic field of the fiducial standard MRI case, including first and second order perturbations.
-Leftmost panel is a radially zoomed-in section of the velocity perturbation structure, to better show the boundary layer-driven structure at the inner cylinder.
-Colors are azimuthal velocity and magnetic field perturbations, and streamfunctions show the perturbation structure in the $r, z$ plane.
-The width of the streamfunctions is proportional to the speed and magnetic field strength in the $r, z$ plane for the velocity and magnetic field, respectively.}
-\end{figure*}
 
@@
 We note that a neutral mode of the form $B_\phi(r) \propto \frac{1}{r}$ would exist in a resistance-free approximation.
+Here, however, this mode does not exist and we derive a single real GLE as the amplitude equation of the standard MRI.
 
@@
 Thus thin-gap linear MRI modes must be eigenstates of parity.
-These symmetries are preserved in the nonlinear MRI terms because they are nonlinear combinations of lower-order eigenfunctions.
+These symmetries are preserved in the nonlinear MRI vectors because they are nonlinear combinations of lower-order eigenfunctions.
 In the wide gap case, the ``potential" $\mathrm{U}(r)$ varies with $r$, so symmetric and antisymmetric modes are no longer required.
 This lack of symmetry is readily apparent in the eigenfunctions and nonlinear vectors in Figures \ref{fig:linear_eigenfunctions} and \ref{fig:N2_N3}, both of which display enhanced boundary layer activity at the inner boundary as compared to the outer boundary.
-The inner and outer boundary layers are symmetric in the thin gap case.
+The inner and outer boundary layers are symmetric in the thin gap case (see Figure \ref{fig:thin_wide_comparison}).
 
@@
 These terms derive from the advective derivatives in the momentum and induction equations, but are filtered out in the thin-gap approximation.
-The nonlinear terms ultimately determine the dependence of the saturation amplitude on $\Pm$, as described below.
+The nonlinear terms ultimately determine the saturation amplitude of the system, as described below.
 %These advective terms allow the nonlinear contributions at both second and third order (i.e.
@@
 
-We examine the behavior of the wide gap MRI system as a function of $\Pm$.
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../figures/hmri_velocity_Bfield_streamfuncs.png}
+\caption{As in Figure \ref{fig:smri_streamfuncs} but for the fiducial helical MRI case, including first and second order perturbations.}
+\end{figure*}
+
+We examine the behavior of the wide gap MRI system as a function of $\Pm$ in the regime $\Pm \ll 1$.
 Figure \ref{fig:coefficients} shows the critical parameters $k_c$ and $\reym$ as a function of $\Pm$, as well as the GLE linear coefficient $b$ and the diffusion coefficient $d$.
-The GLE coefficients are remarkably insensitive to $\Pm$.
-From Equation \ref{eq:gle} it is readily apparent that the asymptotic saturation amplitude is $\alpha_{s} = \pm \sqrt{b/c}$, so we conclude that the saturation amplitude of the MRI is only very weakly dependent on $\Pm$.
-Note that because $\reym$ is essentially constant as a function of $\Pm$, the saturation amplitude is equivalently insensitive to $\reye^{-1}$.
-This is in stark contrast to the narrow gap behavior of the system.
+From Equation \ref{eq:gle} it is readily apparent that the asymptotic saturation amplitude is $\alpha_{s} = \pm \sqrt{b/c}$, and we plot the dependence of $\alpha_{s}$ on $\Pm$ in the bottom panel of \ref{fig:coefficients}.
+%The GLE $b$ and $d$ coefficients are insensitive to $\Pm$, so the $\alpha_{s}$ scaling is driven by the dependence of the nonlinear coefficient $c$ on $\Pm$.
+Note that because $\reym$ is essentially constant as a function of $\Pm$, the saturation amplitude is equivalently sensitive to $\reye^{-1}$.
+We find by fitting the data that the saturation amplitude scales as $\alpha_{s} \sim \Pm^{0.777}$.
 For these same boundary conditions, \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}) find that the narrow gap saturation amplitude scales as $\Pm^{2/3}$.
@@
 $r_1$ and $r_2$.
-Figure \ref{fig:wide_narrow_N31B} shows the structure of the third-order nonlinear term $N_{31}^{(A)}$ as a function of $\Pm$ for both the narrow and wide gap standard MRI.
+Figure \ref{fig:thin_wide_comparison} shows the structure of the third-order nonlinear term $N_{31}^{(A)}$ as a function of $\Pm$ for both the narrow and wide gap standard MRI.
 $N_{31}$ is the vector that determines the GLE coefficient $c$, and thus the scaling of the saturation amplitude because of the insensitivity of $b$ to $\Pm$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs}, \citei{Clark:2016} for the narrow gap equations).
@@
 
+Because it is governed by a real GLE, the saturated standard MRI state may be unstable to the Eckhaus instability, in which the wavelength of the large-scale pattern is adjusted \citep[e.g.][]{Hoyle:2006}.
+Preliminary investigation of the GLE behavior for the standard MRI coefficients derived here indicates that when the simulated vertical domain is large (i.e.
+spans multiple critical wavelengths), the amplitude function is modulated in $Z$, but always be bounded by $\alpha_{s} = \pm \sqrt{b/c}$, as must be the case for the one-dimensional real GLE.
+In Figure \ref{fig:smri_streamfuncs} we plot the saturated state perturbation structure of the fiducial standard MRI, up to and including second order disturbances.
+We use a constant saturation amplitude, but note that a nonconstant $\alpha_{s}$ would introduce more vertical structure.
+We similarly plot the total stress, i.e.
+the sum of the Reynolds and Maxwell stresses in our domain (Figure \ref{fig:total_stress}).
+We find that the saturation mechanism for weakly nonlinear TC flow is a reduction in the background shear profile.
+Thus the bounded cylindrical geometry considered here saturates via the same mechanism as in the narrow gap limit (\citei{Umurhan:2007hs}).
+
 %\section{Channel Modes}
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{../figures/hmri_velocity_Bfield_streamfuncs.png}
-\caption{sdfsdfs }
-\end{figure*}
-
-When $\xi$ in Equation \ref{eq:unperturbed_matrix_eqns} is not equal to zero, the helical MRI arises.
+When $\xi$ in Equation \ref{eq:unperturbed_matrix_eqns} is nonzero, the helical MRI arises.
 We examine a single fiducial helical MRI case, for the parameters used by \citei{Hollerbach:2005tr}, listed in Table \ref{table:parameters}.
@@
 The phase dynamics of the complex GLE are well-studied in a variety of systems, and depend on the values of the GLE coefficients.
-The complex GLE may be unstable to traveling wave instabilities such as the Benjamin-Feir instability, a generalization of the Ekhaus instability which acts on stationary waves, and to which the standard MRI system may be unstable.
+The complex GLE may be unstable to traveling wave instabilities such as the Benjamin-Feir instability, a generalization of the Ekhaus instability.
 The complex GLE can also admit spatiotemporal chaos, and various classes of coherent structures (\citei{Aranson:2002}).
@@
 
-This is the first weakly nonlinear analysis of the MRI in a cylindrical geometry, and is thus the global analogue of similar analyses in local approximations \citep{Umurhan:2007hs,Vasil:2015}.
-We find that the saturation amplitude of the standard MRI scales somewhat more steeply with $\Pm$ than does this same system system in the narrow gap approximation.
-
 We detail several avenues for future work, which highlight the application of this theory to both laboratory experiments and astrophysical disks.
@@
 
+This is the first weakly nonlinear analysis of the MRI in a cylindrical geometry, and is thus the global analogue of similar analyses in local approximations \citep{Umurhan:2007hs,Vasil:2015}.
+Understanding the connection between local and global MRI modes is crucial for interpreting simulation results across domain geometries.
+(The relationship between local and global \textit{linear} MRI modes is investigated in \citet{Latter:2015}.) Phenomena such as saturation and the development of turbulence depend critically on the nature of the underlying MRI modes.
+The formalism presented here describes analytically the weakly nonlinear behavior of global MRI modes.
+This treatment should be expanded to encompass more astrophysically relevant conditions, so that our understanding of complicated MRI phenomena may continue to make contact with analytical theory.
+
+%As this formalism is extended to more astrophysically relevant conditions  
+%The application and extension of the formalism presented here to more astrophysically representative conditions will allow an analytical understanding of weakly nonlinear MRI behavior.
+
+%We find that the saturation amplitude of the standard MRI scales somewhat more steeply with $\Pm$ than does this same system system in the narrow gap approximation.
+
 \section{Acknowledgments}
@@
 acknowledges support from NASA grant 15-LWS15\_2-0063.
-The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, Ellen Zweibel, etc for useful discussion.
+The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
 
@@
 
+\bibitem[{Blaes(2014)}]{Blaes:2014}
+Blaes, O., 2014, SSRv 183, 21
+
 \bibitem[{Chandrasekhar(1960)}]{Chandrasekhar:1960wh}
@@
 
+\bibitem[{Hoyle(2006)}]{Hoyle:2006}
+R. Hoyle, Pattern Formation, 2006, Cambridge U. Press, Cambridge,
+U. K.
+
 \bibitem[{Ji et al.(2001)}]{Ji:2001kd}
@@
 
-\bibitem[Lopez \& Marques(2005)]{Lopez:2005} 
+\bibitem[{Lopez \& Marques(2005)}]{Lopez:2005} 
 Lopez, J.~M., \& Marques, F.\ 2005, Physica D, 211, 168
+
+\bibitem[{Murphy \& Pessah(2015)}]{Murphy:2015tn}
+Murphy, G.C. \& Pessah, M.E., 2015, ApJ 802, 139
 
@@
 
+\bibitem[Vishniac(2009)]{Vishniac:2009il}
+Vishniac, E., 2009, ApJ 696, 1021
+
 \bibitem[{Willis \& Barenghi(2002)}]{Willis:2002bh}

### 2016-09-05 paper/widegap_mri_paper.tex
corrected psi viscous term in paper

@@
 The magnetorotational instability (MRI) is widely belived to drive angular momentum transport in astrophysical disks.
-The MRI is a local instability excited by weak magnetic fields in differentially rotating fluids, and since its discovery \citet{Balbus:1991vs} it has been widely invoked to explain accretion in protoplanetary disks \citep{Armitage:2010} and disks around black holes \citep{Blaes:2014}, as well as jet and wind launching (\citei{Lesur:2013}) anisotropic turbulence \citep{Murphy:2015tn}, and dynamos \citep{Vishniac:2009il} among other phenomena.
+The MRI is a local instability excited by weak magnetic fields in differentially rotating fluids, and since its discovery \citet{Balbus:1991vs} it has been widely invoked to explain accretion in protoplanetary disks \citep{Armitage:2010} and disks around black holes \citep{Blaes:2014}, as well as jet and wind launching (\citei{Lesur:2013}) anisotropic turbulence \citep{Murphy:2015tn}, and dynamo generation \citep{Vishniac:2009il}.
 
@@
 
-\beq
+\begin{align}
 %\begin{split}
-\label{eq:Psi_perturbed}
-\frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r^2}B_0 \xi \partial_z B_\phi - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] = N^{(\Psi)}
+& \frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r^2}B_0 \xi \partial_z B_\phi \nonumber\\
+& - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi - \frac{2}{r^2}\partial_r\Psi) - \frac{1}{r^3} \nabla^2 \Psi + \frac{2}{r^4}\partial_r\Psi\right] = N^{(\Psi)} \label{eq:Psi_perturbed}
 %\end{split}
-\eeq
+\end{align}
 

### 2016-09-05 paper/widegap_mri_paper.tex
changed sign of nonlinear terms in appendix so they represent rhs
updated journal abbreviations

@@
 \beq
-N_2^{\Psi}  = J(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1) + J(\Psi_1, -\frac{2}{r^3}\partial_r\Psi_1)
-- \frac{2}{\beta} J (A_1, \frac{1}{r^2} \nabla^2 A_1) - \frac{2}{\beta} J(A_1, -\frac{2}{r^3} \partial_r A_1) - \frac{2}{r} u_1 \partial_z u_1 + \frac{2}{\beta} \frac{2}{r} B_1 \partial_z B_1
+N_2^{\Psi}  = -J(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1) - J(\Psi_1, -\frac{2}{r^3}\partial_r\Psi_1)
++ \frac{2}{\beta} J (A_1, \frac{1}{r^2} \nabla^2 A_1) + \frac{2}{\beta} J(A_1, -\frac{2}{r^3} \partial_r A_1) + \frac{2}{r} u_1 \partial_z u_1 - \frac{2}{\beta} \frac{2}{r} B_1 \partial_z B_1
 \eeq
@@
 \begin{split}
-N_2^{u} = \frac{1}{r} J\left(\Psi_1, u_1\right) - \frac{1}{r} \frac{2}{\beta} J\left(A_1, B_1\right) + \frac{1}{r^2} u_1 \partial_z \Psi_1 - \frac{2}{\beta}\frac{1}{r^2} B_1 \partial_z A_1
+N_2^{u} = -\frac{1}{r} J\left(\Psi_1, u_1\right) + \frac{1}{r} \frac{2}{\beta} J\left(A_1, B_1\right) - \frac{1}{r^2} u_1 \partial_z \Psi_1 + \frac{2}{\beta}\frac{1}{r^2} B_1 \partial_z A_1
 \end{split}
@@
 \beq
-N_2^A = -\frac{1}{r} J\left(A_1, \Psi_1\right)
-\eeq
-
-\beq
-N_2^B = -\frac{1}{r} J\left(A_1, u_1\right) - \frac{1}{r} J\left(B_1, \Psi_1\right) - \frac{1}{r^2} B_1 \partial_z \Psi_1 + \frac{1}{r^2} u_1 \partial_z A_1
+N_2^A = \frac{1}{r} J\left(A_1, \Psi_1\right)
+\eeq
+
+\beq
+N_2^B = \frac{1}{r} J\left(A_1, u_1\right) - \frac{1}{r} J\left(\Psi_1, B_1\right) + \frac{1}{r^2} B_1 \partial_z \Psi_1 - \frac{1}{r^2} u_1 \partial_z A_1
 \eeq
@@
 \begin{split}
-N_3^{\Psi} & = J(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_2) + J(\Psi_2, \frac{1}{r^2} \nabla^2\Psi_1) + 2 J (\Psi_1, \frac{1}{r^2}\partial_Z\partial_z \Psi_1) + J(\Psi_1, -\frac{2}{r^3}\partial_r \Psi_2) + J(\Psi_2, -\frac{2}{r^3}\partial_r \Psi_1) \\
-& + \widetilde{J}(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1) + \widetilde{J} (\Psi_1, -\frac{2}{r^3}\partial_r \Psi_1) - \frac{2}{\beta} J(A_1, \frac{1}{r^2}\nabla^2 A_2) - \frac{2}{\beta} J(A_2, \frac{1}{r^2}\nabla^2 A_1) - \frac{4}{\beta} J(A_1, \frac{1}{r^2}\partial_Z\partial_z A_1) \\ & - \frac{2}{\beta} J(A_1, -\frac{2}{r^3} \partial_r A_2 )
- - \frac{2}{\beta} J(A_2, -\frac{2}{r^3} \partial_r A_1) - \frac{2}{\beta} \widetilde{J} (A_1, \frac{1}{r^2} \nabla^2 A_1) - \frac{2}{\beta} \widetilde{J} (A_1, -\frac{2}{r^3} \partial_r A_1) \\
-& - \frac{2}{r} u_1 \partial_z u_2 - \frac{2}{r} u_2 \partial_z u_1 - \frac{2}{r} u_1 \partial_Z u_1 + \frac{2}{\beta}\frac{2}{r} B_1\partial_z B_2 + \frac{2}{\beta}\frac{2}{r} B_2 \partial_z B_1 + \frac{2}{\beta} \frac{2}{r} B_1 \partial_Z B_1
+N_3^{\Psi} & = -J(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_2) - J(\Psi_2, \frac{1}{r^2} \nabla^2\Psi_1) - 2 J (\Psi_1, \frac{1}{r^2}\partial_Z\partial_z \Psi_1) - J(\Psi_1, -\frac{2}{r^3}\partial_r \Psi_2) - J(\Psi_2, -\frac{2}{r^3}\partial_r \Psi_1) \\
+& - \widetilde{J}(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1) - \widetilde{J} (\Psi_1, -\frac{2}{r^3}\partial_r \Psi_1) + \frac{2}{\beta} J(A_1, \frac{1}{r^2}\nabla^2 A_2) + \frac{2}{\beta} J(A_2, \frac{1}{r^2}\nabla^2 A_1) + \frac{4}{\beta} J(A_1, \frac{1}{r^2}\partial_Z\partial_z A_1) \\
+& + \frac{2}{\beta} J(A_1, -\frac{2}{r^3} \partial_r A_2 ) + \frac{2}{\beta} J(A_2, -\frac{2}{r^3} \partial_r A_1) + \frac{2}{\beta} \widetilde{J} (A_1, \frac{1}{r^2} \nabla^2 A_1) + \frac{2}{\beta} \widetilde{J} (A_1, -\frac{2}{r^3} \partial_r A_1) \\
+& + \frac{2}{r} u_1 \partial_z u_2 + \frac{2}{r} u_2 \partial_z u_1 + \frac{2}{r} u_1 \partial_Z u_1 - \frac{2}{\beta}\frac{2}{r} B_1\partial_z B_2 - \frac{2}{\beta}\frac{2}{r} B_2 \partial_z B_1 - \frac{2}{\beta} \frac{2}{r} B_1 \partial_Z B_1
 \end{split}
@@
 \begin{split}
-N_3^u & = \frac{1}{r}J\left(\Psi_1, u_2\right) + \frac{1}{r}J\left(\Psi_2, u_1\right) + \frac{1}{r}\widetilde{J} \left(\Psi_1, u_1\right) - \frac{1}{r}\frac{2}{\beta} J\left(A_1, B_2\right) - \frac{1}{r} \frac{2}{\beta} J\left(A_2, B_1\right) - \frac{1}{r}\frac{2}{\beta}\widetilde{J}\left(A_1, B_1\right) \\
-& + \frac{1}{r^2} u_1\partial_z \Psi_2 + \frac{1}{r^2} u_2 \partial_z \Psi_1 + \frac{1}{r^2} u_1 \partial_Z \Psi_1 - \frac{2}{\beta} \frac{1}{r^2} B_1 \partial_z A_2 - \frac{2}{\beta} \frac{1}{r^2} B_2 \partial_z A_1 - \frac{2}{\beta} \frac{1}{r^2} B_1 \partial_Z A_1
+N_3^u & = -\frac{1}{r}J\left(\Psi_1, u_2\right) - \frac{1}{r}J\left(\Psi_2, u_1\right) - \frac{1}{r}\widetilde{J} \left(\Psi_1, u_1\right) + \frac{1}{r}\frac{2}{\beta} J\left(A_1, B_2\right) + \frac{1}{r} \frac{2}{\beta} J\left(A_2, B_1\right) + \frac{1}{r}\frac{2}{\beta}\widetilde{J}\left(A_1, B_1\right) \\
+& - \frac{1}{r^2} u_1\partial_z \Psi_2 - \frac{1}{r^2} u_2 \partial_z \Psi_1 - \frac{1}{r^2} u_1 \partial_Z \Psi_1 + \frac{2}{\beta} \frac{1}{r^2} B_1 \partial_z A_2 + \frac{2}{\beta} \frac{1}{r^2} B_2 \partial_z A_1 + \frac{2}{\beta} \frac{1}{r^2} B_1 \partial_Z A_1
 \end{split}
@@
 \beq
-N_3^A = -\frac{1}{r} J\left(A_1, \Psi_2\right) - \frac{1}{r}J\left(A_2, \Psi_1\right) - \frac{1}{r} \widetilde{J}\left(A_1, \Psi_1\right)
+N_3^A = \frac{1}{r} J\left(A_1, \Psi_2\right) + \frac{1}{r}J\left(A_2, \Psi_1\right) + \frac{1}{r} \widetilde{J}\left(A_1, \Psi_1\right)
 \eeq
@@
 \begin{split}
-N_3^B & = - \frac{1}{r} J\left(A_1, u_2\right) - \frac{1}{r} J\left(A_2, u_1\right) - \frac{1}{r}\widetilde{J}\left(A_1, u_1\right) - \frac{1}{r} J\left(B_1, \Psi_2\right) - \frac{1}{r} J\left(B_2, \Psi_1\right) - \frac{1}{r} \widetilde{J} \left(B_1, u_1\right) \\ & - \frac{1}{r^2} B_1\partial_z \Psi_2 - \frac{1}{r^2} B_2 \partial_z \Psi_1 - \frac{1}{r^2} B_1 \partial_Z \Psi_1 + \frac{1}{r^2} u_1 \partial_z A_2 + \frac{1}{r^2} u_2 \partial_z A_1 + \frac{1}{r^2} u_1 \partial_Z A_1
+N_3^B & = \frac{1}{r} J\left(A_1, u_2\right) + \frac{1}{r} J\left(A_2, u_1\right) + \frac{1}{r}\widetilde{J}\left(A_1, u_1\right) + \frac{1}{r} J\left(B_1, \Psi_2\right) + \frac{1}{r} J\left(B_2, \Psi_1\right) + \frac{1}{r} \widetilde{J} \left(B_1, u_1\right) \\
+& + \frac{1}{r^2} B_1\partial_z \Psi_2 + \frac{1}{r^2} B_2 \partial_z \Psi_1 + \frac{1}{r^2} B_1 \partial_Z \Psi_1 - \frac{1}{r^2} u_1 \partial_z A_2 - \frac{1}{r^2} u_2 \partial_z A_1 - \frac{1}{r^2} u_1 \partial_Z A_1
 \end{split}
@@
 \bibitem[{Aranson \& Kramer(2002)}]{Aranson:2002}
-Aranson, I.S. \& Kramer, L., 2002, Rev. Mod. Phys. 74, 99
+Aranson, I.S. \& Kramer, L., 2002, RvMP 74, 99
 
@@
 \bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}
-Balbus, S A and Hawley, J F, 1991, The Astrophysical Journal, 376, 214
+Balbus, S A and Hawley, J F, 1991, ApJ 376, 214
 
@@
 \bibitem[{Chandrasekhar(1960)}]{Chandrasekhar:1960wh}
-Chandrasekhar, S. 1960, Proceedings of the National Academy of Sciences of the United States of America, 46, 253
+Chandrasekhar, S. 1960, PNAS 46, 253
 
@@
 \bibitem[{Cross \& Hohenberg(1993)}]{Cross:1993el}
-Cross, M. C., \& Hohenberg, P. C., 1993, Rev. Mod. Phys. 65, 851
+Cross, M. C., \& Hohenberg, P. C., 1993, RvMP 65, 851
 
 \bibitem[{Ebrahimi et al.(2009)}]{Ebrahimi:2009ey}
-Ebrahimi, F., Prager, S.C., Schnack, D.D
+Ebrahimi, F., Prager, S.C., Schnack, D.D, 2009, ApJ 698, 233
 
@@
 \bibitem[{Gissinger et al.(2011)}]{Gissinger:2011td}
-Gissinger, C, Ji, H, Goodman, J, 2011, Phys. Rev. E 84, 026308
+Gissinger, C, Ji, H, Goodman, J, 2011, PhRvE 84, 026308
 
 \bibitem[{Gissinger et al.(2012)}]{Gissinger:2012gc}
-Gissinger, C, Goodman, J, \& Ji, H, 2012, Physics of Fluids 24, 074109
+Gissinger, C, Goodman, J, \& Ji, H, 2012, PhFl 24, 074109
 
@@
 \bibitem[{Hollerbach(2009)}]{Hollerbach:2009ig}
-Hollerbach, R, Proc. R. Soc. A, 2009, 465, 2107
+Hollerbach, R., RSPSA 2009, 465, 2107
 
 \bibitem[{Hollerbach \& R\"udiger(2005)}]{Hollerbach:2005tr}
-Hollerbach, R. \& R\"udiger, G., 2005, Phys. Rev. Lett. 95, 124501
+Hollerbach, R. \& R\"udiger, G., 2005, PhRvL 95, 124501
 
@@
 \bibitem[{Ji et al.(2001)}]{Ji:2001kd}
-Ji, H, Goodman, J, Kageyama, A, Mon. Not. R. Astron. Soc., 2001, 325, 1
+Ji, H, Goodman, J, Kageyama, A, MNRAS 2001, 325, 1
 
@@
 \bibitem[{Kirillov \& Stefani(2013)}]{Kirillov:2013}
-Kirillov, O, \& Stefani, F, 2013, Phys. Rev. Lett. 111, 061103
+Kirillov, O, \& Stefani, F, 2013, PhRvL 111, 061103
 
 \bibitem[{Knobloch \& Julien(2005)}]{Knobloch:2005ba}
-Knobloch, E, Julien K, 2005, Physics of Fluids, 17, 094106
+Knobloch, E, Julien K, 2005, PhFl, 17, 094106
 
@@
 \bibitem[{Liu et al.(2006)}]{Liu:2006}
-Liu, W, Goodman, J, Herron, I, and Ji, H, 2006, Phys. Rev. E 74, 056302
+Liu, W, Goodman, J, Herron, I, and Ji, H, 2006, PhRvE 74, 056302
 
 \bibitem[{Lopez \& Marques(2005)}]{Lopez:2005} 
-Lopez, J.~M., \& Marques, F.\ 2005, Physica D, 211, 168
+Lopez, J.~M., \& Marques, F.\ 2005, PhyD, 211, 168
 
@@
 \bibitem[{Parker \& Krommes(2013)}]{Parker:2013hy}
-Parker, J.B. \& Krommes, J.A., 2013, Phys of Plasmas 20, 100703
+Parker, J.B. \& Krommes, J.A., 2013, PhPl 20, 100703
 
@@
 \bibitem[{Recktenwald et al.(1993)}]{Recktenwald:1993}
-Recktenwald, A., L\"ucke, M., M\"uller, H.W., 1993, Phys Rev E 48, 4444
+Recktenwald, A., L\"ucke, M., M\"uller, H.W., 1993, PhRvE 48, 4444
 
 \bibitem[{R\"udiger \& Hollerbach(2007)}]{Rudiger:2007}
-R\"udiger, G, \& Hollerbach, R, Phys. Rev. E 76, 068301
+R\"udiger, G, \& Hollerbach, R, PhRvE 76, 068301
 
 \bibitem[{Schartman et al.(2009)}]{Schartman:2009df}
-Schartman, E, Ji, H, Burin, M J, Rev Sci Instrum, 2009, 80, 024501
+Schartman, E, Ji, H, Burin, M J, RScI, 2009, 80, 024501
 
 \bibitem[{Shakura \& Sunyaev(1973)}]{Shakura:1973wg}
-Shakura, N I and Sunyaev, R A, Astronomy \& Astrophysics, 1973, 24, 337
+Shakura, N I and Sunyaev, R A, A\&A, 1973, 24, 337
 
 \bibitem[{Sisan et al.(2004)}]{Sisan:2004ig}
-Sisan, D R, Mujica, N, Tillotson, W A, Huang Y-M, Dorland, W, Hassam, A B, Antonsen, T M, Lathrop, D P, Phys Rev Lett, 2004, 93, 11
+Sisan, D R, Mujica, N, Tillotson, W A, Huang Y-M, Dorland, W, Hassam, A B, Antonsen, T M, Lathrop, D P, PhRvL, 2004, 93, 11
 
 \bibitem[{Stefani et al.(2006)}]{Stefani:2006iv}
-Stefani, F, Gundrum, T, Gerbeth, G, R\"udiger, G, Schultz, M, Szklarski, J, Hollerbach, R, Phys Rev Lett, 2006, 97, 184502
+Stefani, F, Gundrum, T, Gerbeth, G, R\"udiger, G, Schultz, M, Szklarski, J, Hollerbach, R, PhRvL, 2006, 97, 184502
 
 \bibitem[{Stefani et al.(2009)}]{Stefani:2009hp}
-Stefani, F, Gerbeth, G, Gundrum, T, Hollerbach, R, Priede, J, R\"udiger, G, and Szklarski, J, Phys Rev E, 80, 066303
+Stefani, F, Gerbeth, G, Gundrum, T, Hollerbach, R, Priede, J, R\"udiger, G, and Szklarski, J, PhRvE, 80, 066303
 
@@
 \bibitem[{Umurhan et al.(2007a)}]{Umurhan:2007dz}
-Umurhan, O.M., Regev, O., Menou, K., 2007, Phys. Rev. Letters, 98, 034501
+Umurhan, O.M., Regev, O., Menou, K., 2007, PhRvL, 98, 034501
 
 \bibitem[{Umurhan et al.(2007b)}]{Umurhan:2007hs}
-Umurhan, O.M., Regev, O., Menou, K., 2007, Phys. Rev. E, 76, 036310
+Umurhan, O.M., Regev, O., Menou, K., 2007, PhRvE, 76, 036310
 

### 2016-09-06 paper/widegap_mri_paper.tex
a few final edits.

@@
 %\title{Weakly Nonlinear Analysis of the Standard and Helical Magnetorotational Instabilities in a Taylor Couette Flow}
-\title{The Weakly Nonlinear Magnetorotational Instability} % maybe this is punchier?
+\title{The Weakly Nonlinear Magnetorotational Instability in a Global, Cylindrical Taylor Couette Flow} % maybe this is punchier?
 \author{Clark, S.E.\altaffilmark{1}}
@@
 \begin{abstract}
-We conduct a formal weakly nonlinear analysis of the magnetorotational instability (MRI) in a Taylor Couette flow.
-This is a multiscale perturbative treatment of the nonideal, axisymmetric MRI near threshold, subject to realistic radial boundary conditions.
-We analyze both the standard MRI, initialized by a constant vertical background magnetic field, and the helical MRI, in which the background magnetic field contains an azimuthal component.
-This is the first weakly nonlinear analysis of the MRI in a Taylor Couette geometry, as well as the first weakly nonlinear analysis of the helical MRI.
-We find that the evolution of the weakly nonlinear perturbation amplitude of the standard MRI is described by a real Ginzburg-Landau equation (GLE), while the amplitude of the helical MRI takes the form of a complex GLE.
+We conduct a global, weakly nonlinear analysis of the magnetorotational instability (MRI) in a Taylor Couette flow.
+This is a multiscale perturbative treatment of the nonideal, axisymmetric MRI near threshold, subject to realistic radial boundary conditions and cylindrical geometry.
+We analyze both the standard MRI, initialized by a constant vertical background magnetic field, and the helical MRI, with an azimuthal background field component.
+This is the first weakly nonlinear analysis of the MRI in a global Taylor Couette geometry, as well as the first weakly nonlinear analysis of the helical MRI.
+We find that the evolution of the amplitude of the standard MRI is described by a real Ginzburg-Landau equation (GLE), while the amplitude of the helical MRI takes the form of a complex GLE.
 This suggests that the saturated state of the helical MRI may itself be unstable on long spatial and temporal scales.
@@
 Most relevant to our work is the Princeton Plasma Physics Laboratory (PPPL) MRI Experiment, a liquid gallium TC flow with an axial magnetic field \citep{Ji:2001kd}.
-There has been some theoretical work designed to complement the Princeton MRI experiment involving direct numerical simulation of the experimental conditions, much of it focused on the specific challenges in identifying MRI signatures despite apparatus-driven spurious flows (e.g.
+There has been a significant amount of theoretical work designed to complement the Princeton MRI experiment involving direct numerical simulation of the experimental conditions, much of it focused on the specific challenges in identifying MRI signatures despite spurious, apparatus-driven flows (e.g.
 \citei{Gissinger:2012gc}).
@@
 
-We detail several avenues for future work, which highlight the application of this theory to both laboratory experiments and astrophysical disks.
+We detail several avenues for future work, which highlight the application of this theory to both laboratory experiments and astrophysical disks:
 
 \begin{itemize}
-\item This theory may be applied to a specific experimental apparatus to model the predicted saturated state.
-One can then ask whether GLE dynamics should be detectable, especially over apparatus-driven flows.
+\item Our theory may be applied to a specific experimental apparatus to model the predicted saturated state.
+One can then ask whether GLE dynamics should be detectable, especially over endcap-driven flows.
 \item The saturation properties of different rotation profiles may be compared by direct application of the theory developed here.
-\item Vertical stratification may be added to the base state, which would be a more realistic model of global vertical disk structure. 
-\item Other nonideal MHD effects such as the Hall effect and ambipolar diffusion may be straightforwardly added to this model, and are of particular interest for modeling protoplanetary disks. 
-\item The background magnetic field geometry may be generalized to include radial variation.
+\item Vertical stratification may be added to the base state, constructing a more realistic model of global vertical disk structure. 
+\item Other nonideal MHD effects such as the Hall effect and ambipolar diffusion are straightforward addions to this model, and are of particular interest for understanding protoplanetary disks. 
+\item The background magnetic field geometry may be generalized to include radial variation, another feature relevant to astrophysical disks.
 \item The radial boundary conditions considered here may be expanded to mimic astrophysical disks rather than TC devices.
-\item This theory may be compared to simulations in both the weakly and strongly nonlinear regimes. 
+\item Our theory can be compared to simulations in both the weakly and strongly nonlinear regimes: both in the pattern selection at saturation and the $\Pm^{0.777}$ scaling can be directly tested. 
 \end{itemize}

### 2016-09-11 paper/widegap_mri_paper.tex
updated nsf grant no

@@
 
-A real GLE has also been found to describe the formation of zonal flows out of magnetized turbulence in a model system (\citei{Parker:2013hy}).
+A real GLE has also been found to describe the formation of zonal flows out of magnetized turbulence in a model system \citep{Parker:2013hy, Parker:inpress}.
 Zonal flows are axisymmetric structures, large-scale and long-lived, which form spontaneously out of turbulence.
@@
 \item Vertical stratification may be added to the base state, constructing a more realistic model of global vertical disk structure. 
-\item Other nonideal MHD effects such as the Hall effect and ambipolar diffusion are straightforward addions to this model, and are of particular interest for understanding protoplanetary disks. 
+\item Other nonideal MHD effects such as the Hall effect and ambipolar diffusion are straightforward additions to this model, and are of particular interest for understanding protoplanetary disks. 
 \item The background magnetic field geometry may be generalized to include radial variation, another feature relevant to astrophysical disks.
 \item The radial boundary conditions considered here may be expanded to mimic astrophysical disks rather than TC devices.
-\item Our theory can be compared to simulations in both the weakly and strongly nonlinear regimes: both in the pattern selection at saturation and the $\Pm^{0.777}$ scaling can be directly tested. 
+\item Our theory can be compared to simulations in both the weakly and strongly nonlinear regimes: both the pattern selection at saturation and the $\Pm^{0.777}$ scaling can be directly tested. 
 \end{itemize}
@@
 was supported by a National Science Foundation Graduate Research Fellowship under grant No.
-DGE-11-44155.
+DGE-16-44869.
 J.S.O.
@@
 
+\bibitem[{Parker \& Krommes(in press)}]{Parker:inpress}
+Parker, J.B. \& Krommes, J.A., in press, in Zonal Jets, ed. B. Galperin \& P. Read (Cambridge, U.K.: Cambridge)
+
 \bibitem[{Pessah(2010)}]{Pessah:2010ic}

### 2016-09-26 python/widegap/plot_widegap_dissipation_investigation.py
code to plot initial and final u, B profiles, and dissipation quantities

@@ -1,0 +1,161 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import h5py 
+from matplotlib import pyplot, lines
+from scipy import interpolate, optimize
+import matplotlib.ticker as ticker
+from mpl_toolkits.axes_grid1 import make_axes_locatable, axes_size
+from matplotlib import rc
+import dedalus.public as de
+rc('text', usetex=True)
+
+file_root = "/Users/susanclark/weakly_nonlinear_mri/data/"
+#fn = "widegap_amplitude_parameters_Q_0.01_Rm_0.8403_Pm_1.00e-03_Omega1_313.55_Omega2_56.43_beta_25.00_xi_0.00_gridnum_128_Anorm.h5"
+fn = "widegap_amplitude_parameters_Q_0.90_Rm_3.3088_Pm_1.60e-06_Omega1_01.00_Omega2_00.12_beta_41.20_xi_0.00_gridnum_512_norm_True_Amidnorm_final.h5"
+obj = h5py.File(file_root + fn, "r")
+
+Q = obj.attrs['Q']
+r = obj['r'].value
+rgrid = r
+
+# epsilon (small parameter)
+eps = 0.5
+
+# saturation amplitude -- for now just constant, coefficient-determined
+satamp = np.sqrt(obj.attrs['b']/obj.attrs['c']) #1
+
+# create z grid
+nz = obj.attrs['gridnum']
+Lz = 2*np.pi/Q
+z = np.linspace(0, Lz, nz, endpoint=False)
+zz = z.reshape(nz, 1)
+
+dz = z[1] - z[0]
+
+eiqz = np.exp(1j*Q*zz)
+ei2qz = np.exp(2*1j*Q*zz)
+ei0qz = np.exp(0*1j*Q*zz)
+eiqz_z = 1j*Q*np.exp(1j*Q*zz)
+ei2qz_z = 2*1j*Q*np.exp(2*1j*Q*zz)
+
+eiqzstar = np.exp(-1j*Q*zz)
+eiqzstar_z = -1j*Q*np.exp(-1j*Q*zz)
+ei2qzstar = np.exp(-2*1j*Q*zz)
+ei2qzstar_z = -2*1j*Q*np.exp(-2*1j*Q*zz)
+ei0qzstar = np.exp(-0*1j*Q*zz)
+
+# two-dimensional u and Bstructure
[…]

### 2016-09-30 paper/thingap_mri_paper.tex
added acknowledgements; NASA grants; fixed jso affiliation
code to plot initial and final u, B profiles, and dissipation quantities
added a bunch of text to discussion section. Some minor edits throughout.

@@
 In this work we rigorously rederive the theory of URM07, and expand upon their findings.
-We conduct a set of weakly nonlinear simulations and compare these to the analytical predictions.
 Our focus here is on fully characterizing the narrow gap MRI system, both for independent theoretical interest and to have a robust comparison point for extensions of this theory into more complicated geometries.
-In a companion paper we derive for the first time the weakly nonlinear theory of the standard and helical MRI in a wide gap Taylor Couette flow (\citei{Clark:2016}).
+In a companion paper we derive for the first time the weakly nonlinear theory of the standard and helical MRI in a wide gap Taylor Couette flow \citep[][hereafter Paper II]{Clark:2016}.
 The thin- and wide-gap treatments complement one another theoretically, and both are important regimes for comparison with simulation.
 
-This work examines Taylor Couette flow in the thin-gap regime, and idealization in which the radial extent of the channel is very small compared to its distance from the center of rotation, i.e.
-$(r_2 - r_1) \ll \frac{1}{2} (r_1 + r_2)$ where $r1$ and $r2$ are the radii of the inner and outer flow boundaries, respectively.
+This work examines Taylor Couette flow in the thin-gap regime, an idealization in which the radial extent of the channel is very small compared to its distance from the center of rotation, i.e.
+$(r_2 - r_1) \ll \frac{1}{2} (r_1 + r_2)$ where $r_1$ and $r_2$ are the radii of the inner and outer flow boundaries, respectively.
 The thin-gap approximation eliminates curvature terms, so the domain geometry is Cartesian rather than cylindrical.
@@
 Analytical investigation of this case reveals that MRI saturation can be caused by parasitic Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime (\citei{Pessah:2010ic}).
-The theory of MRI channel mode parasites is relatively robust (see also \citei{Pessah:2009gm}, \citei{Latter:2010iz}, \citei{Rembiasz:2016}), but they are mostly an artifact of the local approximation, and not germane to global analyses like the one presented here (see \citei{Latter:2015} for a detailed analysis of the relation between local and global MRI modes).
+The theory of MRI channel mode parasites is robust (see also \citei{Pessah:2009gm}, \citei{Latter:2010iz}, \citei{Rembiasz:2016}), but their importance may be overestimated by the local approximation \citei{Latter:2015}, and not germane to global analyses like the one presented here.
+\citei{Latter:2015} gives a  detailed analysis of the relation between local and global MRI modes.
 
@@
 \beq
-J\left(f, g\right) \equiv \partial_z f\partial_x g - \partial_x f \partial_z g.
+J\left(f, g\right) \equiv \partial_z f\partial_x g - \partial_x f \partial_z g,
 \eeq  
+and $q = - d \ln \Omega/ \ln R = 3/2$ is the dimensionless shear parameter such that the background velocity profile is $v_0 = q \Omega_0 x$.
 
@@
 
-\section{Simulations}
-
-We compare the weakly nonlinear theory with simulations of the same system.
-We use the pseudospectral code Dedalus to evolve Equations \ref{eq:momentum} and \ref{eq:induction} in time.
-To analyze the weakly nonlinear regime in simulation, we raise $\reym$ slightly to tune the system just out of marginality.
-We initialize a domain of vertical extent $20 \lambda_c$, where $\lambda_c = 2 \pi / k_c$ is the critical vertical wavelength.
-The radial domain is identical to the analytical domain, and the equations are solved on a hybrid Fourier-Chebyshev domain (Figure \ref{fig:setup}).
-
 \begin{figure*}[h!]
@@
 \end{figure*}
+
+\section{Discussion}
+\label{sec:discussion}
+
+The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to $k$ \citep{Hoyle:2006}.
+In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane.
+In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well.
+Figure~\ref{fig:u_saturation} shows the saturated mean radial profiles of $u_0 - u_{\phi} = -q \Omega_0 x - u_{\phi}$ and each term in the steady state force balance (i.e.
+equation~(\ref{eqset2} with $\d_t u_{1y} = 0$).
+In the bulk of the fluid away from the boundary layers, the saturated state shows reduced shear, with little diffusive contribution.
+However, the steep scaling of transport with $\Pm$
+
+Figure~\ref{fig:b_saturation} shows $B_z$ and the terms corresponding to steady state inductive balance (equation~(\ref{eqset2} with $\d_t B_{y} = 0$).
+Here, the instability acts to push magnetic field toward the boundaries in both the bulk and the boundary layers.
+Note also that the total magnetic energy in the $B_z$ component is significantly larger than the initial condition.
+\citet{Ebrahimi:2009} considered the saturation of a single, strongly super-critical MRI mode allowed to interact non-linearly only with itself and the mean.
+They considered two important cases, one in which the mean flow was forced to remain at its initial, quasi-Keplerian state for all time, and one in which the background flow was allowed to evolve.
+This is a crucial difference between the shearing box and our narrow-gap TC flow: perturbations in our simulation can adjust the background flow, whereas in a shearing box, the shear periodicity forbids perturbations from affecting the mean flow.
+In the case with a freely evolving background flow, \citet{Ebrahimi:2009} found that a saturated state quite similar to ours: field pushed to the boundaries, and a reduction in shear in the bulk of the flow.
+Their flows have less pronounced boundary layers, likely because of their much larger $\Pm = 0.1 -1$.
+
+In the high $\reye$ and  $\reym$  limit, \citet{Vasil:2015} derives an amplitude equation considerably different than the one found here.
+By averaging in the $z$ direction, he computes a mean-field equation with striking similarity to the buckling of an elastic beam under load.
+The most salient feature of his equation is its \emph{non-local} character.
+Unlike the present work, which focuses on Keplerian rotation profiles with $q = 3/2$ with a critical $B_0$, \citet{Vasil:2015} focuses on a fixed $B$ and a weakly destabilized shear profile.
+However, these differences are minor; the destabilizing parameter $\epsilon$ enters the analysis in the same quadratic proportion.
+Whether and how \citet{Vasil:2015}'s amplitude equation is equivalent to our own in the limit of dynamically important resistive and viscous effects is beyond the scope of this work.
+
+\section{Acknowledgements}
+JSO acknowledges support from NASA grant 15-LWS15\_2-0063. NSF. The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, etc for useful discussion.
 
@@
 Goodman, J, Xu, G, 1994, ApJ, 432, 213
+
+\bibitem[Hoyle(2006)]{Hoyle:2006} Hoyle, R.\ 2006, Pattern
+  Formation, by Rebecca Hoyle, pp.~432.~Cambridge University Press,
+  March 2006.~ISBN-10: 0521817501.~ISBN-13: 9780521817509, 432
 

### 2016-09-30 paper/thingap_mri_paper.tex
latex corrections

@@
 Figure~\ref{fig:u_saturation} shows the saturated mean radial profiles of $u_0 - u_{\phi} = -q \Omega_0 x - u_{\phi}$ and each term in the steady state force balance (i.e.
-equation~(\ref{eqset2} with $\d_t u_{1y} = 0$).
+equation~(\ref{eqset2} with $\partial_t u_{1y} = 0$).
 In the bulk of the fluid away from the boundary layers, the saturated state shows reduced shear, with little diffusive contribution.
@@
 
-Figure~\ref{fig:b_saturation} shows $B_z$ and the terms corresponding to steady state inductive balance (equation~(\ref{eqset2} with $\d_t B_{y} = 0$).
+Figure~\ref{fig:b_saturation} shows $B_z$ and the terms corresponding to steady state inductive balance (equation~(\ref{eqset2} with $\partial_t B_{y} = 0$).
 Here, the instability acts to push magnetic field toward the boundaries in both the bulk and the boundary layers.

### 2016-09-30 paper/thingap_mri_paper.tex
thingap appendices

@@
 
-\section{Equations}
+\section{Equations}\label{sec:equations}
 
@@
 We are able to resolve even small-scale structure in the boundary layers of our domain, because we need only resolve it in one dimension.
-We solve the radial component of each equation using the open source pseudospectral code Dedalus (Burns et al., in prep).
+We solve the radial component of each equation using the open source pseudospectral code Dedalus.
 We compute the radial components on a grid of Chebyshev polynomials, as is appropriate for bounded one-dimensional domains (e.g.
@@
 JSO acknowledges support from NASA grant 15-LWS15\_2-0063. NSF. The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, etc for useful discussion.
+
+%************* APPENDICES ************************%
+
+\clearpage
+\appendix
+
+\section{A. Detailed Equations}\label{app:matrices}
+
+Here we detail the perturbation analysis described in Section \ref{sec:equations}. The linear system is described by Equation \ref{eq:unperturbed_matrix_eqns}, where
+
+\beq
+\mathcal{L} = \mathcal{L}_0 + \mathcal{L}_1 \partial_z + \mathcal{L}_2 \partial_z^2 + \mathcal{L}_3 \partial_z^3 + \mathcal{L}_4 \partial_z^4,
+\eeq
+
+\beq
+\widetilde{\mathcal{G}} = - \mathcal{G} \partial_z - \mathcal{L}_3 \partial_z^3,
+\eeq
+
+and the constituent matrices are defined as
+
+\beq
+\mathcal{D} = \left[\begin{matrix}
+\nabla^2 & 0 & 0 & 0 \\
+0 & 1& 0 & 0 \\
+0 & 0 & 1 & 0\\
+0 & 0 & 0 & 1 \\
+\end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{L}_0 = \left[\begin{matrix}
+-\frac{1}{\reye}\partial_x^4 & 0 & 0 & 0 \\
+0 & -\frac{1}{\reye}\partial_x^2 & 0 &0 \\
+0 & 0 & -\frac{1}{\reym}\partial_x^2 & 0 \\
+0 & 0 & 0 & -\frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{L}_1 = \left[\begin{matrix}
+0 & -2 & -\frac{2}{\beta}\partial_x^2 & 0 \\
+(2-q)\Omega_0 & 0 & 0 & -\frac{2}{\beta} \\
+-1 & 0 & 0 & 0 \\
+0 & -1 & q\Omega_0 & 0 \\ \end{matrix}\right] 
+\eeq
+
+\beq
+\mathcal{L}_2 = \left[\begin{matrix}
+-2\frac{1}{\reye} \partial_x^2 & 0 & 0 & 0 \\
+0 & -\frac{1}{\reye} & 0 & 0 \\
+0 & 0 & -\frac{1}{\reym} & 0 \\
+0 & 0 & 0 & -\frac{1}{\reym} \\ \end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{L}_3 = \left[\begin{matrix}
+0 & 0 & -\frac{2}{\beta} & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\ \end{matrix} \right]
+\eeq
+
+\beq
+\mathcal{L}_4 = \left[\begin{matrix}
+-\frac{1}{\reye} & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\ \end{matrix}\right] 
+\eeq
+
+\beq
+\mathcal{G} = \left[\begin{matrix}
+0 & 0 & -\frac{2}{\beta}\partial_x^2 & 0 \\
+0 & 0 & 0 & -\frac{2}{\beta} \\
+-1 & 0 & 0 & 0 \\
+0 & -1 & 0 & 0 \\ \end{matrix} \right]
+\eeq
+
+\section{B. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}
+\beq
+\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4)
+\eeq
+
+where
+
+\beq
+N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \frac{2}{\beta}J(A_1, \nabla^2 A_1) 
+\eeq
+
+\beq
+N_2^{(u)} = J(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_1) 
+\eeq
+
+\beq
+N_2^{(A)} = - J(A_1, \Psi_1) 
+\eeq
+
+\beq
+N_2^{(B)} =  J(\Psi_1, B_1) \, - \, J(A_1, u_1)
+\eeq
+
+and
+
+\beq
+\begin{split}
+N_3^{(\Psi)} & = J(\Psi_1, \nabla^2\Psi_2) \, - \, \frac{2}{\beta}J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) - \, \frac{2}{\beta}J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, \\
+& - \, 2 \frac{2}{\beta}J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \frac{2}{\beta}\widetilde{J}(A_1, \nabla^2 A_1)\\
+\end{split}
+\eeq
+
+\beq
+N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_2) \, - \, \frac{2}{\beta} J(A_2, B_1) \, - \, \frac{2}{\beta} \widetilde{J}(A_1, B_1)
+\eeq
+
+\beq
+N_3^{(A)} = - J(A_1, \Psi_2) \, - \, J(A_2, \Psi_1) \, - \, \widetilde{J}(A_1, \Psi_1)
+\eeq
+
+\beq
+N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1).\\
+\eeq
+
+$\mathbf{N_2}$ and $\mathbf{N_3}$ expand to become
+
+\beq
+\mathbf{N_2} = \alpha^2\mathbb{N}_{22} e^{i2 k_c z} + \left|\alpha\right|^2 \mathbb{N}_{20} + c.c.
+\eeq
+
+and
+
+\beq
+\mathbf{N_3} = \alpha^3 \mathbb{N}_{33} e^{i 3 k_c z} + \alpha\partial_Z\alpha \mathbb{N}_{32} e^{i 2 k_c z} + \alpha\left|\alpha\right|^2 \mathbb{N}_{31} e^{i k_c z} + \alpha \partial_Z  \beta \mathbb{\widetilde{N}}_{31} e^{i k_c z} + \alpha^*\partial_Z \alpha \mathbb{N}_{30} + c.c.
+\eeq
+
+The second order nonlinear terms are \\
+
+\beq
+\begin{split}
+N_{22}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11} - k_c^2 \partial_x \Psi_{11}\right) - \partial_x \Psi_{11} \cdot \left(i k_c \partial_x^2 \Psi_{11} - i k_c^3 \Psi_{11}\right) \\
+& + \frac{2}{\beta}\partial_x A_{11} \cdot \left(i k_c \partial_x^2 A_{11} - i k_c^3 A_{11}\right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11} - k_c^2 \partial_x A_{11}\right)
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{22}^{(u)} = & \, i k_c \Psi_{11} \cdot \partial_x u_{11} - \partial_x \Psi_{11} \cdot i k_c u_{11} - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11} + \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}
+\end{split}
+\eeq
+
+\beq
+N_{22}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11} + \partial_x A_{11} \cdot i k_c \Psi_{11}
+\eeq
+
+\beq
+N_{22}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11} - \partial_x \Psi_{11} \cdot i k_c B_{11} - i k_c A_{11} \cdot \partial_x u_{11} + \partial_x A_{11} \cdot i k_c u_{11}
+\eeq
+
+\beq
+\begin{split}
+N_{20}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - k_c^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i k_c^3 \Psi_{11}^* - i k_c \partial_x^2 \Psi_{11}^*\right) \\
+& +\frac{2}{\beta} \partial_x A_{11} \cdot \left(i k_c^3 A_{11}^* - i k_c \partial_x^2 A_{11}^* \right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
+\end{split}
+\eeq
+
+\beq
+N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* + \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11}^* - \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}^*
+\eeq
+
+\beq
+N_{20}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11}^* - \partial_x A_{11} \cdot i k_c \Psi_{11}^*
+\eeq
+
+\beq
+N_{20}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11}^* + \partial_x \Psi_{11} \cdot i k_c B_{11}^* - i k_c A_{11} \cdot \partial_x u_{11}^* - \partial_x A_{11} \cdot i k_c u_{11}^*
+\eeq \\
+
+and the third order nonlinear terms become \\
+
+\beq
+\begin{split}
+N_{31}^{(\Psi)} = &  \, i k_c \left(\Psi_{11} \cdot \partial_x^3 \Psi_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x^3\Psi_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x^3 \Psi_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot \partial_x^2 \Psi_{22}\right) \\
+& + i 8 k_c^3 \left(\partial_x\Psi_{11}^* \cdot \Psi_{22}\right) + i 4 k_c^3 \left(\Psi_{11}^* \cdot \partial_x \Psi_{22}\right) + \frac{2}{\beta}\left[ - i k_c \left(A_{11} \cdot \partial^3 A_{20}\right) - i k_c \left(A_{11} \cdot \partial_x^3 A_{20}^*\right) \right]\\
+& + \frac{2}{\beta}\left[i k_c \left(A_{11}^* \cdot \partial_x^3 A_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \partial_x^2 A_{22}\right) - i 8 k_c^3 \left(\partial_x A_{11}^* \cdot A_{22}\right) - i 4 k_c^3 \left(A_{11}^* \cdot \partial_x A_{22}\right)\right] \\
+& + i 2 k_c \left(\Psi_{22} \cdot \partial_x^3 \Psi_{11}^*\right) - i 2 k_c^3\left(\Psi_{22} \cdot \partial_x \Psi_{11}^* \right) - i k_c \left(\partial_x \Psi_{20} \cdot \partial_x^2 \Psi_{11}\right) + i k_c \left(\partial_x \Psi_{22} \cdot \partial_x^2 \Psi_{11}^*\right) \\
+& - i k_c \left(\partial_x \Psi_{20}^* \cdot \partial_x^2 \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20} \cdot \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20}^* \cdot \Psi_{11}\right) - i k_c^3 \left(\partial_x \Psi_{22} \cdot \Psi_{11}^*\right) \\
+& + \frac{2}{\beta} \left[ - i 2 k_c \left(A_{22} \cdot \partial_x^3 A_{11}^*\right) + i 2 k_c^3 \left(A_{22} \cdot \partial_x A_{11}^*\right) + i k_c \left(\partial_x A_{20} \cdot \partial_x^2 A_{11}\right) - i k_c \left(\partial_x A_{22} \cdot \partial_x^2 A_{11}^*\right) \right] \\
+& + \frac{2}{\beta} \left[ i k_c \left(\partial_x A_{20}^* \cdot \partial_x^2 A_{11}\right) - i k_c^3 \left(\partial_x A_{20} \cdot A_{11}\right) - i k_c^3 \left(\partial_x A_{20}^* \cdot A_{11}\right) + i k_c^3 \left(\partial_x A_{22} \cdot A_{11}^*\right)\right]
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{31}^{(u)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x u_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x u_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x u_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot u_{22}\right) \\
+& - i k_c \left(u_{11} \cdot \partial_x \Psi_{20} \right) - i k_c \left(u_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c \left(u_{11}^* \cdot \partial_x \Psi_{22} \right) + i 2 k_c \left(\partial_x u_{11}^* \cdot \Psi_{22}\right) \\
+& +\frac{2}{\beta}\left[- i k_c \left(A_{11} \cdot \partial_x B_{20}\right) - i k_c \left(A_{11} \cdot \partial_x B_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x B_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot B_{22}\right)\right] \\
+& +\frac{2}{\beta}\left[ i k_c \left(B_{11} \cdot \partial_x A_{20}\right) + i k_c \left(B_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(B_{11}^* \cdot \partial_x A_{20}\right) - i 2 k_c \left(\partial_x B_{11}^* \cdot A_{22}\right)\right]
+\end{split}
+\eeq
+
+%Correct -- checked 11/14/15
+\beq
+\begin{split}
+N_{31}^{(A)} =\, & -i k_c \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i k_c \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
+ + i k_c \left(A_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \Psi_{22}\right) \\
+ & + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{31}^{(B)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x B_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x B_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x B_{22}\right) - i 2 k_c\left(\partial_x \Psi_{11}^* \cdot B_{22}\right) \\
+& - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}\right) - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c\left(B_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x B_{11}^* \cdot \Psi_{22}\right) \\
+& - i k_c \left(A_{11} \cdot \partial_x u_{20}\right) - i k_c \left(A_{11} \cdot \partial_x u_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x u_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot u_{22}\right) \\
+& i k_c \left(u_{11} \cdot \partial_x A_{20} \right) + i k_c \left(u_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(u_{11}^* \cdot \partial_x A_{22}\right) - i 2 k_c \left(\partial_x u_{11}^* \cdot A_{22}\right)
+\end{split}
+\eeq
 

### 2016-09-30 paper/thingap_mri_paper.tex
edits

@@
 \beq
+\label{eq:unperturbed_matrix_equations}
  \mathcal{D}\partial_t \mathbf{V} +  \mathcal{L} \mathbf{V} + \epsilon^2\widetilde{\mathcal{G}} = \mathbf{N},
@@
 This is a crucial difference between the shearing box and our narrow-gap TC flow: perturbations in our simulation can adjust the background flow, whereas in a shearing box, the shear periodicity forbids perturbations from affecting the mean flow.
-In the case with a freely evolving background flow, \citet{Ebrahimi:2009} found that a saturated state quite similar to ours: field pushed to the boundaries, and a reduction in shear in the bulk of the flow.
+In the case with a freely evolving background flow, \citet{Ebrahimi:2009ey} found a saturated state quite similar to ours: field pushed to the boundaries, and a reduction in shear in the bulk of the flow.
 Their flows have less pronounced boundary layers, likely because of their much larger $\Pm = 0.1 -1$.
@@
 \section{Acknowledgements}
-JSO acknowledges support from NASA grant 15-LWS15\_2-0063. NSF. The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, etc for useful discussion.
+S.E.C.
+was supported by a National Science Foundation Graduate Research Fellowship under grant No.
+DGE-16-44869.
+J.S.O.
+acknowledges support from NASA grant 15-LWS15\_2-0063.
+The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
 
@@
 
+\bibitem[{Ebrahimi et al.(2009)}]{Ebrahimi:2009ey}
+Ebrahimi, F., Prager, S.C., Schnack, D.D, 2009, ApJ 698, 233
+
 \bibitem[Goodman \& Xu(1994)]{Goodman:1994ul}

### 2016-09-30 paper/thingap_mri_paper.tex
very short draft of conclusion section. need to add more here...

@@
 
+\section{Conclusion}
+\label{sec:conclusion}
+
+In this paper, we have constructed a weakly nonlinear analysis of the MRI using multiple-scales analysis.
+We find that the saturated state is a complex balance in which reduction of shear and amplification and redistribution of $B_z$ combine to saturate the instability.
+We confirm the previous results of \citet{Umurhan:2007hs}, including their $A^2 \propto \Pm^{4/3} $ scaling.
+
 \section{Acknowledgements}

### 2016-10-03 paper/thingap_mri_paper.tex
IVP figures added to thingap paper

@@
 
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_2.png}%{allorders_velocity.eps}
+\caption{IVP for two critical wavelengths.}\label{fig:IVP_2lambdacrit}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_10.png}%{allorders_velocity.eps}
+\caption{IVP for 10 critical wavelengths.}\label{fig:IVP_10lambdacrit}
+\end{figure*}
+
 \section{Conclusion}

### 2016-10-03 paper/thingap_mri_paper.tex
IVP/GLE text

@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_velocity_Pm_1E-3.png}%{allorders_velocity.eps}
+\caption{First order (left), second order (center), and total (right) velocity perturbations.
+Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds.
+Colorbar represents azimuthal velocity.}\label{allorders_velocity}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_Bfield_Pm_1E-3.png}
+\caption{First order (left), second order (center), and total (right) magnetic field perturbations.
+Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to higher magnetic field strengths.
+Colorbar represents azimuthal magnetic field strength.}\label{allorders_Bfield}
+\end{figure*}
+
 The weakly nonlinear regime is where the MRI system is nonlinearly unstable to only the most unstable mode of the linear solution.
@@
 
-\begin{figure*}[h!]
-\centering
-\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_velocity_Pm_1E-3.png}%{allorders_velocity.eps}
-\caption{First order (left), second order (center), and total (right) velocity perturbations.
-Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds.
-Colorbar represents azimuthal velocity.}\label{allorders_velocity}
-\end{figure*}
+To close the perturbation series we enforce a solvability criterion on Equation \ref{eq:ordere3} (see Appendix \ref{app:matrices}).
+This leads to an amplitude equation for $\alpha(T, Z)$ that governs the slow length- and timescale evolution of the system.
+This amplitude equation is
 
-\begin{figure*}[h!]
-\centering
-\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_Bfield_Pm_1E-3.png}
-\caption{First order (left), second order (center), and total (right) magnetic field perturbations.
-Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to higher magnetic field strengths.
-Colorbar represents azimuthal magnetic field strength.}\label{allorders_Bfield}
-\end{figure*}
+\beq
+\label{eq:gle}
+\partial_T \alpha = b \alpha + d \partial_Z^2 \alpha - c \alpha \left|\alpha^2\right|,
+\eeq
+
+a real Ginzburg-Landau equation.
 
@@
 
-Figure~\ref{fig:b_saturation} shows $B_z$ and the terms corresponding to steady state inductive balance (equation~(\ref{eqset2} with $\partial_t B_{y} = 0$).
+Figure~\ref{fig:b_saturation} shows $B_z$ and the terms corresponding to steady state inductive balance (Equation~\ref{eqset2} with $\partial_t B_{y} = 0$).
 Here, the instability acts to push magnetic field toward the boundaries in both the bulk and the boundary layers.
@@
 
+The real Ginzburg-Landau equation describes the amplitude behavior of our system close to threshold.
+Although the form of the equation is generic, its coefficients depend on the specific physics of our system and govern its detailed evolution (see Appendix \ref{app:matrices}).
+We simulate the evolution of the MRI amplitude equation by solving Equation \ref{eq:gle} on a Fourier basis in $Z$ using Dedalus.
+We initialize uniform random noise of amplitude $-10^{-3}$ to $+10^{-3}$ in $Z$, and timestep the system using a four-stage, fourth-order Runge-Kutta integrator.
+We evolve the system for $100 \Omega_0^{-1}$ in timesteps of $0.02 \Omega_0^{-1}$.
+Results are shown in Figures \ref{fig:IVP_2lambdacrit} and \ref{fig:IVP_10lambdacrit}.
+The system quickly organizes itself into rolls in $Z$ bounded by the analytic saturation amplitude $\alpha_s = \sqrt{b/c}$.
+The specific geometry depends on the number of critical wavelengths $\lambda_{crit} = 2\pi/k_c$ that are initialized in $Z$.
+Figure \ref{fig:IVP_2lambdacrit} shows that a system with a height equal to two critical wavelengths will be modulated by simple rolls of sinusoidal amplitude.
+The saturation amplitude pattern becomes more complicated when more modes are allowed to interact.
+Figure \ref{fig:IVP_10lambdacrit} shows the evolution of a system of height $10 \lambda_{crit}$.
+While still bounded by $\alpha_s$, the saturation amplitude exhibits a nonlinear phase geometry due to the nonlinear interaction of modes in $Z$.
+
 \begin{figure*}[h!]
@@
 \includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_2.png}%{allorders_velocity.eps}
-\caption{IVP for two critical wavelengths.}\label{fig:IVP_2lambdacrit}
-\end{figure*}
+\caption{Evolution of the Ginzburg-Landau amplitude equation (Equation \ref{eq:gle}) on a Fourier $Z$ domain of length $2 \lambda_{crit}$, where $\lambda_{crit} = 2\pi/k_c$ is the critical wavelength of the system.
+Top left panel shows the evolution of the amplitude observable $\alpha \alpha^*$ on the full $Z$ domain as a function of time $T$.
+Bottom left panel shows the amplitude $alpha$ at the final timestep shown, where the black line is the real part $\mathrm{Re}\{\alpha(t = t_{final})\}$ and the gray line is the imaginary part $\mathrm{Im}\{\alpha(t = t_{final})\}$.
+The final amplitude is bounded by the analytic saturation amplitude $\alpha_{saturation} = \pm \sqrt{b/c}$.
+Top right panel shows the evolution of the phase angle $\phi = \mathrm{arctan} (\mathrm{Im}(\alpha)/\mathrm{Re}(\alpha))$ on the same domain.
+Bottom panel shows the phase angle as a function of $Z$ for the final timestep.
+Note that the phase angle is wrapped on a $2\pi$ domain, such that $\pi$ = $-\pi$, as indicated by the circular colorbar.}\label{fig:IVP_2lambdacrit}
+%\end{figure*}
 
-\begin{figure*}[h!]
+%\begin{figure*}[h!]
 \centering
 \includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_10.png}%{allorders_velocity.eps}
-\caption{IVP for 10 critical wavelengths.}\label{fig:IVP_10lambdacrit}
+\caption{Same as Figure \ref{fig:IVP_2lambdacrit} but for a $Z$ domain of length $10 \lambda_{crit}$.}\label{fig:IVP_10lambdacrit}
 \end{figure*}

### 2016-10-03 paper/thingap_mri_paper.tex
added linear dispersion relation. note that our growth rate dependence is quartic

@@
 \begin{align}
-\mathcal{O}(\epsilon)&: \mathcal{L}\mathbf{V_1} + \mathcal{D}\partial_t \mathbf{V_1} = 0. \\ \label{eq:ordere}
-\mathcal{O}(\epsilon^2)&: \mathcal{L}\mathbf{V_2} + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1} = \mathbf{N_2}\\
-\mathcal{O}(\epsilon^3)&: \mathcal{L} \mathbf{V_3} + \mathcal{D}\partial_T \mathbf{V_1}  + \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} + \widetilde{\mathcal{G}}\mathbf{V_1} = \mathbf{N_3} \, \\ \label{eq:ordere3}
+\mathcal{O}(\epsilon)&: \mathcal{L}\mathbf{V_1} + \mathcal{D}\partial_t \mathbf{V_1} = 0. \label{eq:ordere}\\
+\mathcal{O}(\epsilon^2)&: \mathcal{L}\mathbf{V_2} + \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1} = \mathbf{N_2} \label{eq:ordere2}\\
+\mathcal{O}(\epsilon^3)&: \mathcal{L} \mathbf{V_3} + \mathcal{D}\partial_T \mathbf{V_1}  + \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} + \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} + \widetilde{\mathcal{G}}\mathbf{V_1} = \mathbf{N_3} \label{eq:ordere3}
 \end{align}
@@
 
-The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to $k$ \citep{Hoyle:2006}.
+The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic (or in our case quartic) maximum in growth rate with respect to $k$ \citep{Hoyle:2006}.
 In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane.
@@
 
-\section{A. Detailed Equations}\label{app:matrices}
+\section{A. Linear dispersion relation}\label{app:dispersion}
 
-Here we detail the perturbation analysis described in Section \ref{sec:equations}. The linear system is described by Equation \ref{eq:unperturbed_matrix_eqns}, where
+The linear dispersion relation, which determines the variable scalings in the multiple scales analysis.
+This relation is found by perturbing the linear system (Equation \ref{eq:ordere}) with a small perturbation of the form $e^{\sigma t + i k_x x + i k_z z}$.
+
+\beq
+\begin{split}
+& \frac{B_{0}^{4} k_{x}^{2} k_{z}^{4}}{16 \pi^{2}} + \frac{B_{0}^{4} k_{z}^{6}}{16 \pi^{2}} - \frac{B_{0}^{2} \Omega_{0} k_{z}^{4} q}{2 \pi} - 2 \Omega_{0} k_{z}^{2} q \sigma^{2} - \frac{4 \sigma}{\reym} \Omega_{0} k_{x}^{2} k_{z}^{2} q - \frac{4 \sigma}{\reym} \Omega_{0} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{x}^{4} k_{z}^{2} q - \frac{4 \Omega_{0}}{\reym^{2}} k_{x}^{2} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{z}^{6} q \\
+& - k_{x}^{2} \sigma^{4} - k_{z}^{2} \sigma^{4} + 4 k_{z}^{2} \sigma^{2} - \frac{2 \sigma^{3}}{\reym} k_{x}^{4} - \frac{4 \sigma^{3}}{\reym} k_{x}^{2} k_{z}^{2} + \frac{8 \sigma}{\reym} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reym} k_{z}^{4} + \frac{8 \sigma}{\reym} k_{z}^{4} - \frac{k_{x}^{6} \sigma^{2}}{\reym^{2}} - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{4} k_{z}^{2} + \frac{4 k_{x}^{4}}{\reym^{2}} k_{z}^{2} \\
+& - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{2} k_{z}^{4} + \frac{8 k_{x}^{2}}{\reym^{2}} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reym^{2}} + \frac{4 k_{z}^{6}}{\reym^{2}} - \frac{2 \sigma^{3}}{\reye} k_{x}^{4} - \frac{4 \sigma^{3}}{\reye} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reye} k_{z}^{4} - \frac{4 k_{x}^{6} \sigma^{2}}{\reye \reym} -  \frac{12 k_{x}^{4} k_{z}^{2} \sigma^{2}}{\reye \reym} - \frac{12 k_{x}^{2} k_{z}^{4} \sigma^{2}}{\reye \reym} \\
+& - \frac{4 k_{z}^{6} \sigma^{2}}{\reye \reym} - \frac{2 k_{x}^{8} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye \reym^{2}} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye \reym^{2}} - \frac{2 k_{z}^{8} \sigma}{\reye \reym^{2}} - \frac{k_{x}^{6} \sigma^{2}}{\reye^{2}} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{4} k_{z}^{2} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{2} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reye^{2}} \\
+& - \frac{2 k_{x}^{8} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye^{2} \reym} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye^{2} \reym} - \frac{2 k_{z}^{8} \sigma}{\reye^{2} \reym} - \frac{k_{x}^{10}}{\reye^{2} \reym^{2}} - \frac{5 k_{x}^{8} k_{z}^{2}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{6} k_{z}^{4}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{4} k_{z}^{6}}{\reye^{2} \reym^{2}} \\
+& - \frac{5 k_{x}^{2} k_{z}^{8}}{\reye^{2} \reym^{2}} - \frac{k_{z}^{10}}{\reye^{2} \reym^{2}} = 0
+\end{split}
+\eeq
+
+\section{B. Detailed Equations}\label{app:matrices}
+
+Here we detail the perturbation analysis described in Section \ref{sec:equations}. The linear system is described by Equation \ref{eq:unperturbed_matrix_equations}, where
 
@@
 
-\section{B. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}
+\section{C. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}
 \beq

### 2016-10-03 paper/thingap_mri_paper.tex
note order of kz, kx in disp reln

@@
 
-The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic (or in our case quartic) maximum in growth rate with respect to $k$ \citep{Hoyle:2006}.
+The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to $k$ \citep{Hoyle:2006}.
 In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane.
@@
 This relation is found by perturbing the linear system (Equation \ref{eq:ordere}) with a small perturbation of the form $e^{\sigma t + i k_x x + i k_z z}$.
+Note that the spatial eigenvalues appear as $k_z^2$ and $k_x^2$ at lowest order.
 

### 2016-10-03 paper/thingap_mri_paper.tex
lots of edits

@@
 Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds.
-Colorbar represents azimuthal velocity.}\label{allorders_velocity}
+Colorbar represents azimuthal velocity.
+We use a constant amplitude $\alpha = \alpha_{saturation}$ and a small parameter $\epsilon = 0.5$.}\label{fig:allorders_velocity}
 \end{figure*}
@@
 \includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_Bfield_Pm_1E-3.png}
-\caption{First order (left), second order (center), and total (right) magnetic field perturbations.
-Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to higher magnetic field strengths.
-Colorbar represents azimuthal magnetic field strength.}\label{allorders_Bfield}
+\caption{As in Figure \ref{fig:allorders_velocity} but for the magnetic field.
+Streamlines represent magnetic field structure in the vertical-radial plane, where thicker streamlines correspond to higher magnetic field strengths.
+Colorbar represents azimuthal magnetic field strength.}\label{fig:allorders_Bfield}
 \end{figure*}
@@
 Our perturbations are characterized in terms of fast and slow-moving variables, in order to simultaneously track their evolution on two scales.
-The relative scalings of the fast and slow scales are determined as follows.
-When our four equations are linearized for axisymmetric perturbations of the form $e^{\sigma t + i k_x x + i k_z z}$, we can derive the linear dispersion relation, which is fourth order in $\sigma$ (Appendix \ref{app:dispersion}).
-The fast and slow variables are then defined such that each of the temporal and spatial eigenvalues appear at the same lowest order in the linear dispersion relation.
+The relative scalings of the fast and slow variables are chosen such that each of the temporal and spatial eigenvalues appear at the same lowest order in the linear dispersion relation (Appendix \ref{app:dispersion}).
 The scalings are
@@
 Each operator is expanded to reflect these scalings -- for instance, $\partial_z$ becomes $\partial_z + \epsilon\partial_Z$.
+The fast and slow variables are treated as independent in the perturbation series analysis that follows.
 
@@
 a real Ginzburg-Landau equation.
+The saturated solution to Equation \ref{eq:gle} is evidently $\alpha_{saturation} = \pm \sqrt{b/c}$.
+We plot the first order, second order, and total perturbation structure of the fluid variables in Figures \ref{fig:allorders_velocity} and \ref{fig:allorders_Bfield} with a constant $\alpha_{saturation}$.
 
@@
 \includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_10.png}%{allorders_velocity.eps}
-\caption{Same as Figure \ref{fig:IVP_2lambdacrit} but for a $Z$ domain of length $10 \lambda_{crit}$.}\label{fig:IVP_10lambdacrit}
+\caption{As in Figure \ref{fig:IVP_2lambdacrit} but for a $Z$ domain of length $10 \lambda_{crit}$.}\label{fig:IVP_10lambdacrit}
 \end{figure*}
@@
 
-\section{A. Linear dispersion relation}\label{app:dispersion}
-
-The linear dispersion relation, which determines the variable scalings in the multiple scales analysis.
-This relation is found by perturbing the linear system (Equation \ref{eq:ordere}) with a small perturbation of the form $e^{\sigma t + i k_x x + i k_z z}$.
-Note that the spatial eigenvalues appear as $k_z^2$ and $k_x^2$ at lowest order.
-
-\beq
-\begin{split}
-& \frac{B_{0}^{4} k_{x}^{2} k_{z}^{4}}{16 \pi^{2}} + \frac{B_{0}^{4} k_{z}^{6}}{16 \pi^{2}} - \frac{B_{0}^{2} \Omega_{0} k_{z}^{4} q}{2 \pi} - 2 \Omega_{0} k_{z}^{2} q \sigma^{2} - \frac{4 \sigma}{\reym} \Omega_{0} k_{x}^{2} k_{z}^{2} q - \frac{4 \sigma}{\reym} \Omega_{0} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{x}^{4} k_{z}^{2} q - \frac{4 \Omega_{0}}{\reym^{2}} k_{x}^{2} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{z}^{6} q \\
-& - k_{x}^{2} \sigma^{4} - k_{z}^{2} \sigma^{4} + 4 k_{z}^{2} \sigma^{2} - \frac{2 \sigma^{3}}{\reym} k_{x}^{4} - \frac{4 \sigma^{3}}{\reym} k_{x}^{2} k_{z}^{2} + \frac{8 \sigma}{\reym} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reym} k_{z}^{4} + \frac{8 \sigma}{\reym} k_{z}^{4} - \frac{k_{x}^{6} \sigma^{2}}{\reym^{2}} - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{4} k_{z}^{2} + \frac{4 k_{x}^{4}}{\reym^{2}} k_{z}^{2} \\
-& - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{2} k_{z}^{4} + \frac{8 k_{x}^{2}}{\reym^{2}} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reym^{2}} + \frac{4 k_{z}^{6}}{\reym^{2}} - \frac{2 \sigma^{3}}{\reye} k_{x}^{4} - \frac{4 \sigma^{3}}{\reye} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reye} k_{z}^{4} - \frac{4 k_{x}^{6} \sigma^{2}}{\reye \reym} -  \frac{12 k_{x}^{4} k_{z}^{2} \sigma^{2}}{\reye \reym} - \frac{12 k_{x}^{2} k_{z}^{4} \sigma^{2}}{\reye \reym} \\
-& - \frac{4 k_{z}^{6} \sigma^{2}}{\reye \reym} - \frac{2 k_{x}^{8} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye \reym^{2}} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye \reym^{2}} - \frac{2 k_{z}^{8} \sigma}{\reye \reym^{2}} - \frac{k_{x}^{6} \sigma^{2}}{\reye^{2}} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{4} k_{z}^{2} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{2} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reye^{2}} \\
-& - \frac{2 k_{x}^{8} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye^{2} \reym} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye^{2} \reym} - \frac{2 k_{z}^{8} \sigma}{\reye^{2} \reym} - \frac{k_{x}^{10}}{\reye^{2} \reym^{2}} - \frac{5 k_{x}^{8} k_{z}^{2}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{6} k_{z}^{4}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{4} k_{z}^{6}}{\reye^{2} \reym^{2}} \\
-& - \frac{5 k_{x}^{2} k_{z}^{8}}{\reye^{2} \reym^{2}} - \frac{k_{z}^{10}}{\reye^{2} \reym^{2}} = 0
-\end{split}
-\eeq
-
-\section{B. Detailed Equations}\label{app:matrices}
+\section{A. Detailed Equations}\label{app:matrices}
 
@@
 
-\section{C. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}
+\section{B. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}
 \beq
@@
 
+\section{C. Linear dispersion relation}\label{app:dispersion}
+
+The linear dispersion relation, which determines the variable scalings in the multiple scales analysis.
+This relation is found by perturbing the linear system (Equation \ref{eq:ordere}) with a small perturbation of the form $e^{\sigma t + i k_x x + i k_z z}$.
+Note that the spatial eigenvalues appear as $k_z^2$ and $k_x^2$ at lowest order.
+
+\beq
+\begin{split}
+& \frac{B_{0}^{4} k_{x}^{2} k_{z}^{4}}{16 \pi^{2}} + \frac{B_{0}^{4} k_{z}^{6}}{16 \pi^{2}} - \frac{B_{0}^{2} \Omega_{0} k_{z}^{4} q}{2 \pi} - 2 \Omega_{0} k_{z}^{2} q \sigma^{2} - \frac{4 \sigma}{\reym} \Omega_{0} k_{x}^{2} k_{z}^{2} q - \frac{4 \sigma}{\reym} \Omega_{0} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{x}^{4} k_{z}^{2} q - \frac{4 \Omega_{0}}{\reym^{2}} k_{x}^{2} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{z}^{6} q \\
+& - k_{x}^{2} \sigma^{4} - k_{z}^{2} \sigma^{4} + 4 k_{z}^{2} \sigma^{2} - \frac{2 \sigma^{3}}{\reym} k_{x}^{4} - \frac{4 \sigma^{3}}{\reym} k_{x}^{2} k_{z}^{2} + \frac{8 \sigma}{\reym} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reym} k_{z}^{4} + \frac{8 \sigma}{\reym} k_{z}^{4} - \frac{k_{x}^{6} \sigma^{2}}{\reym^{2}} - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{4} k_{z}^{2} + \frac{4 k_{x}^{4}}{\reym^{2}} k_{z}^{2} \\
+& - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{2} k_{z}^{4} + \frac{8 k_{x}^{2}}{\reym^{2}} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reym^{2}} + \frac{4 k_{z}^{6}}{\reym^{2}} - \frac{2 \sigma^{3}}{\reye} k_{x}^{4} - \frac{4 \sigma^{3}}{\reye} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reye} k_{z}^{4} - \frac{4 k_{x}^{6} \sigma^{2}}{\reye \reym} -  \frac{12 k_{x}^{4} k_{z}^{2} \sigma^{2}}{\reye \reym} - \frac{12 k_{x}^{2} k_{z}^{4} \sigma^{2}}{\reye \reym} \\
+& - \frac{4 k_{z}^{6} \sigma^{2}}{\reye \reym} - \frac{2 k_{x}^{8} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye \reym^{2}} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye \reym^{2}} - \frac{2 k_{z}^{8} \sigma}{\reye \reym^{2}} - \frac{k_{x}^{6} \sigma^{2}}{\reye^{2}} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{4} k_{z}^{2} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{2} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reye^{2}} \\
+& - \frac{2 k_{x}^{8} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye^{2} \reym} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye^{2} \reym} - \frac{2 k_{z}^{8} \sigma}{\reye^{2} \reym} - \frac{k_{x}^{10}}{\reye^{2} \reym^{2}} - \frac{5 k_{x}^{8} k_{z}^{2}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{6} k_{z}^{4}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{4} k_{z}^{6}}{\reye^{2} \reym^{2}} \\
+& - \frac{5 k_{x}^{2} k_{z}^{8}}{\reye^{2} \reym^{2}} - \frac{k_{z}^{10}}{\reye^{2} \reym^{2}} = 0
+\end{split}
+\eeq
+
 \bibliographystyle{plain}

### 2016-10-03 paper/widegap_mri_paper.tex
mordecai's edits incorporated into widegap draft

@@
 %\title{Weakly Nonlinear Analysis of the Standard and Helical Magnetorotational Instabilities in a Taylor Couette Flow}
-\title{The Weakly Nonlinear Magnetorotational Instability in a Global, Cylindrical Taylor Couette Flow} % maybe this is punchier?
+\title{The Weakly Nonlinear Magnetorotational Instability in a Global, Cylindrical Taylor-Couette Flow} % maybe this is punchier?
 \author{Clark, S.E.\altaffilmark{1}}
@@
 \begin{abstract}
-We conduct a global, weakly nonlinear analysis of the magnetorotational instability (MRI) in a Taylor Couette flow.
+We conduct a global, weakly nonlinear analysis of the magnetorotational instability (MRI) in a Taylor-Couette flow.
 This is a multiscale perturbative treatment of the nonideal, axisymmetric MRI near threshold, subject to realistic radial boundary conditions and cylindrical geometry.
 We analyze both the standard MRI, initialized by a constant vertical background magnetic field, and the helical MRI, with an azimuthal background field component.
-This is the first weakly nonlinear analysis of the MRI in a global Taylor Couette geometry, as well as the first weakly nonlinear analysis of the helical MRI.
+This is the first weakly nonlinear analysis of the MRI in a global Taylor-Couette geometry, as well as the first weakly nonlinear analysis of the helical MRI.
 We find that the evolution of the amplitude of the standard MRI is described by a real Ginzburg-Landau equation (GLE), while the amplitude of the helical MRI takes the form of a complex GLE.
@@
 The magnetorotational instability (MRI) is widely belived to drive angular momentum transport in astrophysical disks.
-The MRI is a local instability excited by weak magnetic fields in differentially rotating fluids, and since its discovery \citet{Balbus:1991vs} it has been widely invoked to explain accretion in protoplanetary disks \citep{Armitage:2010} and disks around black holes \citep{Blaes:2014}, as well as jet and wind launching (\citei{Lesur:2013}) anisotropic turbulence \citep{Murphy:2015tn}, and dynamo generation \citep{Vishniac:2009il}.
+The MRI is a local instability excited by weak magnetic fields in differentially rotating fluids, and since first applied to an astrophysical context \citep{Balbus:1991vs} it has been widely invoked to explain accretion in protoplanetary disks \citep{Armitage:2010} and disks around black holes \citep{Blaes:2014}, as well as jet and wind launching (\citei{Lesur:2013}), anisotropic turbulence \citep{Murphy:2015tn}, and dynamo generation \citep{Brandenburg:1995, Vishniac:2009il}.
 
@@
 In protoplanetary disks, for example, the behavior and evolution of the MRI--and even its very existence--may change drastically depending on the properties of the magnetic field, the disk composition, disk geometry, and so forth.
-Multiphysics numerical simulations of such systems is currently an area of intense focus, enabling the study of nonideal MHD effects, disk stratification, nonequilibrium chemistry, and other complex physics that does not lend itself easily to analytic study \citep[e.g.][among many others]{Fleming:2003fs,Bai:2011cm, Flock:2013,Suzuki:2014vh}.
+Multiphysics numerical simulations of such systems are currently an area of intense focus, enabling the study of nonideal MHD effects, disk stratification, nonequilibrium chemistry, and other complex physics that does not lend itself easily to analytic study \citep[e.g.][among many others]{Fleming:2003fs,Bai:2011cm, Flock:2013,Suzuki:2014vh}.
 Still, computational costs inevitably constrain numerical approaches.
@@
 However, we can construct asymptotic approximations valid for $\Pm \ll 1$ using analytic methods.
+
 Analytic methods can also play a powerful role in elucidating the mechanisms responsible for MRI saturation.
@@
 The shearing box is a convenient computational framework allowing extreme resolution for local MRI studies and has been extended to include vertical stratification and a wide variety of diffusive effects.
+
 However, while the MRI is a local instability, there are a number of important problems that require a global treatment.
@@
 The helical MRI currently occupies a special place in the MRI puzzle.
-The helical MRI has been proposed as a method of awakening angular momentum transport in the ``dead zones" of protoplanetary disks where the $\Pm$ becomes very small.
+The helical MRI has been proposed as a method of awakening angular momentum transport in the ``dead zones" of protoplanetary disks where the $\reym$ becomes very small.
 However the rotation profiles needed to excite helical MRI may be steeper than Keplerian, depending on the boundary conditions, and so its role in astrophysical disks is currently a matter of debate \citep{Liu:2006,Rudiger:2007,Kirillov:2013}.
@@
 In this work we will focus our findings on two fiducial parameter sets, one for the standard MRI where $\xi = 0$ and one for the helical MRI.
-We choose the SMRI parameters to be comparable to the case considered in \citei{Goodman:2002ix}.
+We choose the standard MRI parameters to be comparable to the case considered in \citei{Goodman:2002ix}.
 The helical MRI parameters were chosen to be comparable to \citei{Hollerbach:2005tr}.
@@
 
-We solve the standard MRI system subject to periodic vertical boundary conditions and no-slip, perfectly conducting radial boundary conditions, namely
+We solve the standard MRI system subject to the same boundary conditions used in \citet{Goodman:2002ix}.
+These are periodic vertical boundary conditions and no-slip, perfectly conducting radial boundary conditions, namely
 
@@
 
-at $r = r_1, r_2$. To the helical MRI system we apply insulating boundary conditions,
+at $r = r_1, r_2$. To the helical MRI system we apply insulating boundary conditions as used in \citet{Hollerbach:2005tr}:
 
@@
 A spatially constant azimuthal field satisfies both the thin-gap MRI equations and this boundary condition.
-This neutral mode is formally included in the analysis of \citei{Umurhan:2007hs} and yields a second amplitude equation in the form of a simple diffusion equation.
+This neutral mode is formally included in the analysis of \citet{Umurhan:2007hs} and yields a second amplitude equation in the form of a simple diffusion equation.
 This amplitude equation decouples from the GLE because of the translational symmetry of the thin-gap geometry.
-Because that symmetry is not preserved in the wide-gap case, \citeauthor{Umurhan:2007hs} postulate that slow variation in the wide-gap geometry will be governed by two coupled amplitude equations.
+Because that symmetry is not preserved in the wide-gap case, \citet{Umurhan:2007hs} postulate that slow variation in the wide-gap geometry will be governed by two coupled amplitude equations.
 %However, the wide-gap conducting boundary condition $J_z = \frac{1}{r} \partial_r (r B_\phi) = 0$, together with the purely geometric term in Equation \ref{eq:Bphi_perturbed}, conspire to prevent the wide-gap geometry from sustaining a neutral mode.
@@
 The preservation of symmetries in the thin-gap geometry is worth a closer look, as its absence in the wide gap case is the source of many differences in the systems.
-\citei{Latter:2015} point out that in the ideal limit ($\nu, \eta \rightarrow 0$), the linearized system described by the lefthand side of Equations \ref{eq:Psi_perturbed} - \ref{eq:Bphi_perturbed} can be expressed as a Schr{\"o}dinger equation for the radial velocity.
-Similarly combining equations to obtain a single expression for $\Psi$, we find that the thin-gap limit linear ideal MRI can be expressed as
+\citet{Latter:2015} point out that in the ideal limit ($\nu, \eta \rightarrow 0$), the linearized system described by the lefthand side of Equations \ref{eq:Psi_perturbed} - \ref{eq:Bphi_perturbed} can be expressed as a Schr{\"o}dinger equation for the radial velocity.
+Similarly combining equations to obtain a single expression for $\Psi$, we find that the thin-gap limit, linear, ideal MRI can be expressed as
 
@@
 When $\xi$ in Equation \ref{eq:unperturbed_matrix_eqns} is nonzero, the helical MRI arises.
-We examine a single fiducial helical MRI case, for the parameters used by \citei{Hollerbach:2005tr}, listed in Table \ref{table:parameters}.
+We examine a single fiducial helical MRI case, for the parameters used by \citet{Hollerbach:2005tr}, listed in Table \ref{table:parameters}.
 The helical MRI is an overstability, so the ansatz linear eigenvector we consider (Equation \ref{eq:ansatz}) is characterized by a complex temporal eigenvalue $\sigma$.
@@
 \bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}
-Balbus, S A and Hawley, J F, 1991, ApJ 376, 214
+Balbus, S. A. and Hawley, J. F., 1991, ApJ 376, 214
 
@@
 
-\bibitem[{Chandrasekhar(1960)}]{Chandrasekhar:1960wh}
-Chandrasekhar, S. 1960, PNAS 46, 253
+\bibitem[{Brandenburg et al.(1995)}]{Brandenburg:1995}
+Brandenburg, A., Norlund, A., Stein, R. F., Torkelsson, U., 1995, ApJ 446, 741
+
+%\bibitem[{Chandrasekhar(1960)}]{Chandrasekhar:1960wh}
+%Chandrasekhar, S. 1960, PNAS 46, 253
 
@@
 \bibitem[{Oishi \& Mac Low(2011)}]{Oishi:2011ei}
-Oishi, J.S. \& Mac Low, M.M., 2011, ApJ 740, 18
+Oishi, J.S. \& Mac Low, M.-M., 2011, ApJ 740, 18
 

### 2016-10-04 paper/thingap_mri_paper.tex
two ways to compute nablasq

@@
 
-\section{Weakly Nonlinear Analysis}
+\section{Weakly Nonlinear Analysis}\label{sec:wnl}
 
@@
 
-Here we detail the perturbation analysis described in Section \ref{sec:equations}. The linear system is described by Equation \ref{eq:unperturbed_matrix_equations}, where
+Here we detail the perturbation analysis described in Section \ref{sec:wnl}. The perturbation series is described by Equations \ref{eq:ordere} - \ref{eq:ordere3}, where
 
@@
 \mathcal{L} = \mathcal{L}_0 + \mathcal{L}_1 \partial_z + \mathcal{L}_2 \partial_z^2 + \mathcal{L}_3 \partial_z^3 + \mathcal{L}_4 \partial_z^4,
+\eeq
+
+\beq
+\widetilde{\mathcal{L}}_1 =  \mathcal{L}_1 + 2\mathcal{L}_2\partial_z + 3\mathcal{L}_3\partial_z^2 + 4\mathcal{L}_4\partial_z^3
+\eeq
+
+\beq
+\widetilde{\mathcal{L}}_2 = \mathcal{L}_2 + 3\mathcal{L}_3\partial_z + 6\mathcal{L}_4\partial_z^2
 \eeq
@@
 \eeq
+
+Once perturbed, the system is solved for successive orders of $\epsilon$ (Equations \ref{eq:ordere} - \ref{eq:ordere3}).
+$\mathcal{O}(\epsilon)$ is the linear system.
+At $\mathcal{O}(\epsilon^2)$, first-order MRI modes nonlinearly interact with themselves and with their complex conjugates, and so the term $\mathbf{N}_2$ in Equation \ref{eq:ordere2} has the form
+
+\beq
+\mathbf{N}_2 = |\alpha|^2 \mathbf{N}_{20} + \alpha^2 \mathbf{N}_{22} e^{2 i k_c z}
+\eeq
+
+(see Appendix \ref{app:nonlinear_terms} for the full form of $\mathbf{N}_{20}$ and $\mathbf{N}_{22}$).
 

### 2016-10-04 python/plot_thingap_dissipation_investigation.py

@@ -9,8 +9,8 @@ from matplotlib import rc
 import dedalus.public as de
 rc('text', usetex=True)
 
-file_root = "/home/joishi/hg-projects/weakly_nonlinear_mri/data/"
-#file_root = "/Users/susanclark/weakly_nonlinear_MRI/data/"
+#file_root = "/home/joishi/hg-projects/weakly_nonlinear_mri/data/"
+file_root = "/Users/susanclark/weakly_nonlinear_MRI/data/"
 fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_128"
 obj = h5py.File(file_root + fn + ".h5", "r")
 
@@ -76,11 +76,11 @@ A22_star_x = obj['A22_x'].value.conj()
 
 # Vboth Psi, A
 V1_Psi = eps*satamp*obj['psi11'].value*eiqz + eps*satamp.conj()*obj['psi11_star'].value*eiqzstar
-V2_Psi = eps**2*satamp**2*obj['psi22'].value*ei2qz + eps**2*satamp.conj()**2*psi22_star*ei2qzstar + np.abs(satamp)**2*(obj['psi20'].value*ei0qz + obj['psi20_star'].value*ei0qzstar)
+V2_Psi = eps**2*satamp**2*obj['psi22'].value*ei2qz + eps**2*satamp.conj()**2*psi22_star*ei2qzstar + eps**2*np.abs(satamp)**2*(obj['psi20'].value*ei0qz + obj['psi20_star'].value*ei0qzstar)
 Vboth_Psi = V1_Psi + V2_Psi
 
 V1_A = eps*satamp*obj['A11'].value*eiqz + eps*satamp.conj()*obj['A11_star'].value*eiqzstar
-V2_A = eps**2*satamp**2*obj['psi22'].value*ei2qz + eps**2*satamp.conj()**2*psi22_star*ei2qzstar + np.abs(satamp)**2*(obj['psi20'].value*ei0qz + obj['psi20_star'].value*ei0qzstar)
+V2_A = eps**2*satamp**2*obj['psi22'].value*ei2qz + eps**2*satamp.conj()**2*psi22_star*ei2qzstar + eps**2*np.abs(satamp)**2*(obj['psi20'].value*ei0qz + obj['psi20_star'].value*ei0qzstar)
 Vboth_A = V1_A + V2_A
 
 # first order perturbations
[…]

### 2016-10-04 python/plot_thingap_saturation_mechanisms.py

@@ -8,8 +8,8 @@ from mpl_toolkits.axes_grid1 import make_axes_locatable, axes_size
 from matplotlib import rc
 rc('text', usetex=True)
 
-fn_root = "/home/joishi/hg-projects/weakly_nonlinear_mri/data/"
-#fn_root = "/Users/susanclark/weakly_nonlinear_MRI/data/"
+#fn_root = "/home/joishi/hg-projects/weakly_nonlinear_mri/data/"
+fn_root = "/Users/susanclark/weakly_nonlinear_MRI/data/"
 fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_128"
 thingap_fn = fn_root + "zavg_quantities_" + fn 
 obj = h5py.File(thingap_fn + ".h5", "r")
@@ -79,7 +79,7 @@ nablasqu_zavg = obj["nablasqu_zavg"].value
 shearu_zavg = obj["shearu_zavg"].value
 dzBphi_zavg = obj["dzBphi_zavg"].value
 
-fig = plt.figure(figsize=(8, 8), facecolor="white")
+fig = plt.figure(figsize=(6, 6), facecolor="white")
 ax1 = fig.add_subplot(211)
 ax2 = fig.add_subplot(212)
 
@@ -87,20 +87,24 @@ ax1.plot(xgrid, uphifinal_zavg - base_flow_zavg, color="black", label=r"$u_\phi^
 #ax1.plot(xgrid, uphifinal_zavg, color="black", label=r"$u_\phi^{sat}$", lw=lw)
 ax1.set_title(r"$u_\phi^{final} - u_\phi^{0}$")
 
-#ax2.plot(xgrid, JPsiu_zavg, label=r"$J(\Psi, u)$")
-#ax2.plot(xgrid, JAB_zavg, label=r"$-\frac{2}{\beta} J(A, B)$")
-#ax2.plot(xgrid, nablasqu_zavg, label=r"$-\frac{1}{\mathrm{Re}} \nabla^2 u$")
+ax2.plot(xgrid, JPsiu_zavg, label=r"$J(\Psi, u)$")
+ax2.plot(xgrid, JAB_zavg, label=r"$-\frac{2}{\beta} J(A, B)$")
[…]

### 2016-10-04 python/plot_thingap_dissipation_investigation.py
saturation components velocity added

@@ -80,7 +80,7 @@ V2_Psi = eps**2*satamp**2*obj['psi22'].value*ei2qz + eps**2*satamp.conj()**2*psi
 Vboth_Psi = V1_Psi + V2_Psi
 
 V1_A = eps*satamp*obj['A11'].value*eiqz + eps*satamp.conj()*obj['A11_star'].value*eiqzstar
-V2_A = eps**2*satamp**2*obj['psi22'].value*ei2qz + eps**2*satamp.conj()**2*psi22_star*ei2qzstar + eps**2*np.abs(satamp)**2*(obj['psi20'].value*ei0qz + obj['psi20_star'].value*ei0qzstar)
+V2_A = eps**2*satamp**2*obj['A22'].value*ei2qz + eps**2*satamp.conj()**2*A22_star*ei2qzstar + eps**2*np.abs(satamp)**2*(obj['A20'].value*ei0qz + obj['A20_star'].value*ei0qzstar)
 Vboth_A = V1_A + V2_A
 
 # first order perturbations
@@ -158,13 +158,13 @@ nabla_B_zavg = np.mean(diff_tens_B['g'], axis=0)
 JPsiu = (Vboth_ur_field*Vboth_uphi_field.differentiate('x') + Vboth_uz_field*Vboth_uphi_field.differentiate('z')).evaluate()
 
 # J (A, B)
-JAB = (-(2/beta)*(Vboth_Br_field*Vboth_Bphi_field.differentiate('x') + Vboth_Bz_field*Vboth_Bphi_field.differentiate('z'))).evaluate()
+JAB = (-(2.0/beta)*(Vboth_Br_field*Vboth_Bphi_field.differentiate('x') + Vboth_Bz_field*Vboth_Bphi_field.differentiate('z'))).evaluate()
 
-nablasqu = (-(1/Re)*(Vboth_uphi_field.differentiate('x').differentiate('x') + Vboth_uphi_field.differentiate('z').differentiate('z'))).evaluate()
+nablasqu = (-(1.0/Re)*(Vboth_uphi_field.differentiate('x').differentiate('x') + Vboth_uphi_field.differentiate('z').differentiate('z'))).evaluate()
 
 shearu = ((2 - q)*Vboth_ur_field).evaluate()
 
-dzBphi = (-(2/beta)*Vboth_Bphi_field.differentiate('z')).evaluate()
[…]

### 2016-10-04 python/plot_thingap_saturation_mechanisms.py

@@ -1,6 +1,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
 import h5py 
+import matplotlib
 from matplotlib import pyplot, lines
 from scipy import interpolate, optimize
 import matplotlib.ticker as ticker
@@ -87,15 +88,38 @@ ax1.plot(xgrid, uphifinal_zavg - base_flow_zavg, color="black", label=r"$u_\phi^
 #ax1.plot(xgrid, uphifinal_zavg, color="black", label=r"$u_\phi^{sat}$", lw=lw)
 ax1.set_title(r"$u_\phi^{final} - u_\phi^{0}$")
 
-ax2.plot(xgrid, JPsiu_zavg, label=r"$J(\Psi, u)$")
-ax2.plot(xgrid, JAB_zavg, label=r"$-\frac{2}{\beta} J(A, B)$")
-ax2.plot(xgrid, nablasqu_zavg, label=r"$-\frac{1}{\mathrm{Re}} \nabla^2 u$")
-ax2.plot(xgrid, shearu_zavg, label=r"$(2-q)\Omega_0 \partial_z \Psi$" )
-ax2.plot(xgrid, dzBphi_zavg, label=r"$-\frac{2}{\beta} B_0 \partial_z B$")
-plt.legend(prop={'size':6})
+ax2.plot(xgrid, JPsiu_zavg, "--", color="slateblue", label=r"$J(\Psi, u)$")
+ax2.plot(xgrid, JAB_zavg, ":", color="green", label=r"$-\frac{2}{\beta} J(A, B)$")
+ax2.plot(xgrid, nablasqu_zavg, color="tomato", label=r"$-\frac{1}{\mathrm{Re}} \nabla^2 u$")
+ax2.plot(xgrid, dzBphi_zavg, "-.", color="royalblue", label=r"$-\frac{2}{\beta} B_0 \partial_z B$")
+ax2.plot(xgrid, shearu_zavg, "_", color="orangered", label=r"$(2-q)\Omega_0 \partial_z \Psi$" )
+ax2.set_xlabel(r"$x$", size=15)
+plt.legend(prop={'size':10}, bbox_to_anchor=(0.95, 0.35))
 
 ax2.set_xlim(-0.2, 0.2)
 ax2.set_ylim(-0.00008, 0.00008)
+ax2.set_xticks([-0.2, 0, 0.2])
+ax2.set_yticks([-0.00008, 0, 0.00008])
[…]

### 2016-10-04 paper/thingap_mri_paper.tex
saturation components velocity added
slice plots added to text
added more conclusions and discussion of force balance. still more to come...

@@
 \usepackage[titletoc,title]{appendix}
-\usepackage{cleveref}
+%\usepackage{cleveref}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../figures/thingap_saturation_mechanism_plots_velocity_slice60.png}
+\caption{.}\label{fig:sat_comp_vel}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../figures/thingap_saturation_mechanism_plots_Bfield_slice60.png}
+\caption{.}\label{fig:sat_comp_Bfield}
+\end{figure}
+
 The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to $k$ \citep{Hoyle:2006}.
-In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane.
+In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane; the quadratic maximum is a consequence of the linear dispersion relation given in Appendix~\ref{app:matrices}.
 In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well.
-Figure~\ref{fig:u_saturation} shows the saturated mean radial profiles of $u_0 - u_{\phi} = -q \Omega_0 x - u_{\phi}$ and each term in the steady state force balance (i.e.
+Figure~\ref{fig:sat_comp_vel} shows the saturated mean radial profiles of $u_0 - u_{\phi} = -q \Omega_0 x - u_{\phi}$ and each term in the steady state force balance (i.e.
 equation~(\ref{eqset2} with $\partial_t u_{1y} = 0$).
 In the bulk of the fluid away from the boundary layers, the saturated state shows reduced shear, with little diffusive contribution.
-However, the steep scaling of transport with $\Pm$
+This demonstrates that even in a case where diffusive effects are important, the bulk of the fluid saturates by balancing shear and magnetic tension.
+As discussed at length in \citet{Vasil:2015}, when diffusive effects are not important, it is impossible to rearrange momentum without also rearranging magnetic fields.
+His model demonstrates saturation without diffusive effects; our results show that outside of the boundary layers, a simultaneous rearrangement of momentum and field occurs.
+In the boundary layers, the non-linear advection balances viscous dissipation.
 
-Figure~\ref{fig:b_saturation} shows $B_z$ and the terms corresponding to steady state inductive balance (Equation~\ref{eqset2} with $\partial_t B_{y} = 0$).
+Figure~\ref{fig:sat_comp_Bfield} shows $B_z$ and the terms corresponding to steady state inductive balance (Equation~\ref{eqset2} with $\partial_t B_{y} = 0$).
 Here, the instability acts to push magnetic field toward the boundaries in both the bulk and the boundary layers.
@@
 
-In this paper, we have constructed a weakly nonlinear analysis of the MRI using multiple-scales analysis.
-We find that the saturated state is a complex balance in which reduction of shear and amplification and redistribution of $B_z$ combine to saturate the instability.
-We confirm the previous results of \citet{Umurhan:2007hs}, including their $A^2 \propto \Pm^{4/3} $ scaling.
+In this paper, we have constructed a weakly nonlinear analysis of the MRI using multiple-scales analysis, leading to a real Ginzburg Landau equation for the nonlinear amplitude, confirming the previous results of \citet{Umurhan:2007hs}.
+We also confirm their $\alpha^2 \propto \Pm^{4/3} $ scaling results.
+We extend their results by constructing a detailed force and inductive balance for the saturated $u_y$ and $B_z$ components.
+In doing so, we find the saturated state is a complex balance in which reduction of shear and amplification and redistribution of $B_z$ combine to saturate the instability.
+We also perform direct numerical simulations of the amplitude equation demonstrating that complex patterns can for on long length scales $Z$ though the maximum magnitude of the amplitude $\alpha$ is well predicted by the steady state solution.
+In paper II, we make use of the techniques developed here to extend the weakly nonlinear analysis of the MRI to a full cylindrical geometry appropriate for a wide gap experiment.
 

### 2016-10-04 paper/thingap_mri_paper.tex
removed comments

@@
 \usepackage{color}
-%\usepackage{appendix}
+
 \usepackage[titletoc,title]{appendix}
-%\usepackage{cleveref}
 
@@
 \author{Oishi, J.S. \altaffilmark{2,}\altaffilmark{3}}
-%\author{Mac Low, M.-M.\altaffilmark{1,}\altaffilmark{2}}
+
 \altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
 \altaffiltext{2}{Department of Astrophysics, American Museum of Natural History, New York, NY}
-\altaffiltext{3}{Department of Physics and Astronomy, Bates College}
+\altaffiltext{3}{Department of Physics and Astronomy, Bates College, Lewiston, ME}
 
@@
 \end{figure}
-
-%We solve the equations of non-ideal, axisymmetric, incompressible magnetized Taylor Couette flow.
 
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_velocity_Pm_1E-3.png}%{allorders_velocity.eps}
+\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_velocity_Pm_1E-3.png}
 \caption{First order (left), second order (center), and total (right) velocity perturbations.
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_2.png}%{allorders_velocity.eps}
+\includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_2.png}
 \caption{Evolution of the Ginzburg-Landau amplitude equation (Equation \ref{eq:gle}) on a Fourier $Z$ domain of length $2 \lambda_{crit}$, where $\lambda_{crit} = 2\pi/k_c$ is the critical wavelength of the system.
@@
 
-%Correct -- checked 11/14/15
 \beq

### 2016-10-04 python/plot_thingap_saturation_mechanisms.py
readability edit for figures:

@@ -89,13 +89,13 @@ shearu_slice = obj["shearu_slice"].value
 dzBphi_slice = obj["dzBphi_slice"].value
 
 
-fig = plt.figure(figsize=(6, 6), facecolor="white")
+fig = plt.figure(figsize=(8, 8), facecolor="white")
 ax1 = fig.add_subplot(211)
 ax2 = fig.add_subplot(212)
 
 ax1.plot(xgrid, uphifinal_zavg - base_flow_zavg, color="black", label=r"$u_\phi^{final} - u_\phi^{0}$", lw=lw)
 #ax1.plot(xgrid, uphifinal_zavg, color="black", label=r"$u_\phi^{sat}$", lw=lw)
-ax1.set_title(r"$u_\phi^{final} - u_\phi^{0}$")
+ax1.set_title(r"$u_\phi^{saturated} - u_\phi^{0}$", size=18)
 
 colors = ["#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3", "#a6d854"]
 lw = 2
@@ -114,7 +114,7 @@ else:
     ax2.plot(xgrid, dzBphi_slice, "-.", color=colors[3], lw=lw, label=r"$-\frac{2}{\beta} B_0 \partial_z B$")
     ax2.plot(xgrid, shearu_slice, dashes = [5,2,10,5], color=colors[4], lw=lw, label=r"$(2-q)\Omega_0 \partial_z \Psi$" )
 ax2.set_xlabel(r"$x$", size=15)
-plt.legend(prop={'size':10}, bbox_to_anchor=(0.9, 0.35))
+plt.legend(prop={'size':15}, bbox_to_anchor=(0.9, 0.4))
 
 #ax2.set_xlim(-0.2, 0.2)
 #ax2.set_ylim(-0.00008, 0.00008)
@@ -150,7 +150,7 @@ JAPsi_dx_slice = obj["JAPsi_dx_slice"].value
 nablasqAterm_dx_slice = obj["nablasqAterm_dx_slice"].value
 Aterm1_dx_slice = obj["Aterm1_dx_slice"].value
 
-fig = plt.figure(figsize=(6, 6), facecolor="white")
+fig = plt.figure(figsize=(8, 8), facecolor="white")
 ax1 = fig.add_subplot(211)
 ax2 = fig.add_subplot(212)
 
[…]

### 2016-10-04 paper/thingap_mri_paper.tex
readability edit for figures:
u_phi -> u_y

--- 2016-10-04 paper/thingap_mri_paper.tex
@@
 \label{eqset2}
-\partial_t u_{1y} \, + \, J\left(\Psi, u_{y}\right) \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \\
+\partial_t u_{y} \, + \, J\left(\Psi, u_{y}\right) \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \\
 \, \frac{2}{\beta}B_0\partial_z B_{y} \, + \, \frac{2}{\beta} J\left(A, B_{y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{y}
@@
 \includegraphics[width=\columnwidth]{../figures/thingap_saturation_mechanism_plots_velocity_slice60.png}
-\caption{.}\label{fig:sat_comp_vel}
+\caption{Top panel: Mean radial profile for $u_y^{saturated} - u_y^0$, the difference between the saturated azimuthal velocity profile and the initial azimuthal velocity profile (TC flow).
+Bottom panel: each term in the azimuthal velocity steady state force balance (Equation \ref{eqset2} with $\partial_t u_{y} = 0$).} \label{fig:sat_comp_vel}
 \end{figure}
@@
 In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well.
-Figure~\ref{fig:sat_comp_vel} shows the saturated mean radial profiles of $u_0 - u_{\phi} = -q \Omega_0 x - u_{\phi}$ and each term in the steady state force balance (i.e.
-equation~(\ref{eqset2} with $\partial_t u_{1y} = 0$).
+Figure~\ref{fig:sat_comp_vel} shows the saturated mean radial profiles of $u_0 - u_{y} = -q \Omega_0 x - u_{y}$ and each term in the steady state force balance (i.e.
+equation~\ref{eqset2} with $\partial_t u_{y} = 0$).
 In the bulk of the fluid away from the boundary layers, the saturated state shows reduced shear, with little diffusive contribution.

### 2016-10-04 paper/thingap_mri_paper.tex
updated intro and abstract.

--- 2016-10-04 paper/thingap_mri_paper.tex
@@
 \begin{abstract}
-The magnetorotational instability (MRI) is a fundamental process of accretion disk physics, but its saturation mechanism remains poorly understood.
-We present an analytic analysis of the nonideal MRI in the weakly nonlinear regime -- that is, when the MRI system is just unstable to its most unstable mode.
+The magnetorotational instability (MRI) is a fundamental process of accretion disk physics, but its saturation mechanism remains poorly understood despite considerable theoretical and computational effort.
+We present a multiple-scales analysis of the nonideal MRI in the weakly nonlinear regime -- that is, when the the most unstable MRI mode has a growth rate asymptotically approaching zero from above.
+Here, we develop our theory in a thin-gap, Cartesian channel.
+Our results confirm Umurhan et al.'s findings that the saturation amplitude follows a Ginzburg-Landau equation.
+We extend these results by performing a detailed force balance for the saturated $u_y$ and $B_z$ fields, demonstrating that even when diffusive effects are important, the bulk flow saturates by a combination of reducing the background shear and rearranging and strengthening the background $B_z$ field.
 \end{abstract}
@@
 The transport mechanism is likely turbulent, as molecular viscosity alone cannot account for the needed angular momentum transfer, and likely magnetic, as this turbulence is excited even in hydrodynamically stable disks (\citei{Shakura:1973wg}).
-Since its discovery by \citeauthor{Chandrasekhar:1960wh} (\citeyear{Chandrasekhar:1960wh}) and subsequent rediscovery by \citeauthor{Balbus:1991vs} (\citeyear{Balbus:1991vs}), the magnetorotational instability (MRI) remains the leading explanation for rapid angular momentum transport in astrophysical disks.
+Discovered by \citet{Chandrasekhar:1960wh} and \citet{Velikhov:1959} in global geometry, the MRI was subsequent rediscoverd and applied to accretion disks by \citet{Balbus:1991vs}.
+Since then, the magnetorotational instability (MRI) remains the leading explanation for rapid angular momentum transport in astrophysical disks.
 The instability in its simplest geometry arises when a differentially rotating disk is threaded by a vertical magnetic field.
@@
 Note also that the total magnetic energy in the $B_z$ component is significantly larger than the initial condition.
-\citet{Ebrahimi:2009} considered the saturation of a single, strongly super-critical MRI mode allowed to interact non-linearly only with itself and the mean.
+\citet{Ebrahimi:2009ey} considered the saturation of a single, strongly super-critical MRI mode allowed to interact non-linearly only with itself and the mean.
 They considered two important cases, one in which the mean flow was forced to remain at its initial, quasi-Keplerian state for all time, and one in which the background flow was allowed to evolve.
@@
 
+\bibitem[{Velikhov(1959)}]{Velikhov:1959}
+Velikhov, E. P., 1959, Sov. Phys. JETP 36, 995.
+
 \end{thebibliography}

### 2016-10-04 paper/thingap_mri_paper.tex
captions, edits, Taylor Couette -> Taylor-Couette

@@
 
-\title{The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor Couette flow}
+\title{The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor-Couette flow}
 \author{Clark, S.E.\altaffilmark{1}}
@@
 The transport mechanism is likely turbulent, as molecular viscosity alone cannot account for the needed angular momentum transfer, and likely magnetic, as this turbulence is excited even in hydrodynamically stable disks (\citei{Shakura:1973wg}).
-Discovered by \citet{Chandrasekhar:1960wh} and \citet{Velikhov:1959} in global geometry, the MRI was subsequent rediscoverd and applied to accretion disks by \citet{Balbus:1991vs}.
+Discovered by \citet{Chandrasekhar:1960wh} and \citet{Velikhov:1959} in global geometry, the MRI was subsequently rediscovered and applied to accretion disks by \citet{Balbus:1991vs}.
 Since then, the magnetorotational instability (MRI) remains the leading explanation for rapid angular momentum transport in astrophysical disks.
@@
 The weakly nonlinear treatment of the MRI was pioneered by \citeauthor{Umurhan:2007dz} (\citeyear{Umurhan:2007dz}, \citeyear{Umurhan:2007hs}; hereafter URM07).
-The authors undertook the first weakly nonlinear analysis of the MRI in a thin-gap Taylor Couette flow, and found that the marginal MRI system approaches saturation in a manner analogous to that of Rayleigh-B\'enard convection.
+The authors undertook the first weakly nonlinear analysis of the MRI in a thin-gap Taylor-Couette flow, and found that the marginal MRI system approaches saturation in a manner analogous to that of Rayleigh-B\'enard convection.
 Weakly nonlinear analysis was instrumental in our understanding of Rayleigh-B\'enard convection saturation (\citei{Newell:1969}), and the similarities between convection and the thin-gap MRI are the result of important shared symmetries between the systems.
@@
 Our focus here is on fully characterizing the narrow gap MRI system, both for independent theoretical interest and to have a robust comparison point for extensions of this theory into more complicated geometries.
-In a companion paper we derive for the first time the weakly nonlinear theory of the standard and helical MRI in a wide gap Taylor Couette flow \citep[][hereafter Paper II]{Clark:2016}.
+In a companion paper we derive for the first time the weakly nonlinear theory of the standard and helical MRI in a wide gap Taylor-Couette flow \citep[][hereafter Paper II]{Clark:2016}.
 The thin- and wide-gap treatments complement one another theoretically, and both are important regimes for comparison with simulation.
 
-This work examines Taylor Couette flow in the thin-gap regime, an idealization in which the radial extent of the channel is very small compared to its distance from the center of rotation, i.e.
+This work examines Taylor-Couette flow in the thin-gap regime, an idealization in which the radial extent of the channel is very small compared to its distance from the center of rotation, i.e.
 $(r_2 - r_1) \ll \frac{1}{2} (r_1 + r_2)$ where $r_1$ and $r_2$ are the radii of the inner and outer flow boundaries, respectively.
@@
 In particular, in the wide-gap geometry (i.e.
-true Taylor Couette flow) the base angular velocity is a function of $r$, where in the thin-gap approximation the shear flow reduces to a linear profile.
+true Taylor-Couette flow) the base angular velocity is a function of $r$, where in the thin-gap approximation the shear flow reduces to a linear profile.
 
@@
 \includegraphics[trim=5cm 0cm 0cm 0cm, scale=.35]{setup_diagram.eps}
-\caption{Schematic diagram of our set-up, an axisymmetric thin-gap Taylor Couette flow.
+\caption{Schematic diagram of our set-up, an axisymmetric thin-gap Taylor-Couette flow.
 We investigate a 2D slice of the X-Z (radial-vertical) plane.
@@
 %See Figure \ref{figscalings}. 
-Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic Taylor Couette flow.
+Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic Taylor-Couette flow.
 Our $x$ dimension, the direction of angular momentum transport, is analogous to the direction of temperature transport in the convection problem.
@@
 \caption{Top panel: Mean radial profile for $u_y^{saturated} - u_y^0$, the difference between the saturated azimuthal velocity profile and the initial azimuthal velocity profile (TC flow).
-Bottom panel: each term in the azimuthal velocity steady state force balance (Equation \ref{eqset2} with $\partial_t u_{y} = 0$).} \label{fig:sat_comp_vel}
+Bottom panel: each term in the steady state force balance (Equation \ref{eqset2} with $\partial_t u_{y} = 0$).
+Saturated quantities are computed with $\alpha_{saturation} = \sqrt{b/c}$ and $\epsilon = 0.5$.
+The saturated state shows reduced shear in the bulk of the flow, outside of the boundary layers.} \label{fig:sat_comp_vel}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{../figures/thingap_saturation_mechanism_plots_Bfield_slice60.png}
-\caption{.}\label{fig:sat_comp_Bfield}
+\caption{Top panel: Mean radial profile for $B_z^{saturated}$, the saturated vertical magnetic field (black line).
+$B_z^0 = 1$ is the constant background magnetic field (gray line).
+Bottom panel: each term in the steady state inductive balance ($\partial_x$ of Equation \ref{eqset4} with $\partial_t B_y = 0$).
+Saturated quantities are computed with $\alpha_{saturation} = \sqrt{b/c}$ and $\epsilon = 0.5$.
+The saturated field is pushed to the radial domain boundaries.}\label{fig:sat_comp_Bfield}
 \end{figure}
@@
 
-Figure~\ref{fig:sat_comp_Bfield} shows $B_z$ and the terms corresponding to steady state inductive balance (Equation~\ref{eqset2} with $\partial_t B_{y} = 0$).
+Figure~\ref{fig:sat_comp_Bfield} shows $B_z$ and the terms corresponding to steady state inductive balance ($\partial_x$ of Equation~\ref{eqset4} with $\partial_t B_{y} = 0$).
 Here, the instability acts to push magnetic field toward the boundaries in both the bulk and the boundary layers.

### 2016-10-04 paper/thingap_mri_paper.tex
appendices finished

@@
 
+Note that, following the notation of \citet{Umurhan:2007hs}, the subscripts refer to $\epsilon$ order, $z$ order, successively, such that $\mathbf{N}_{22}$ is the second-order nonlinear term which corresponds to $e^{2 i k_c z}$ z-dependence.
+
+Equation \ref{eq:ordere2} is solved as three separate systems of equations, one for each possible $z$ resonance:
+
+\begin{align}
+\mathcal{L}\mathbf{V}_{20} & = \mathbf{N}_{20}\\
+\mathcal{L}\mathbf{V}_{21} & = - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_{11} \\
+\mathcal{L}\mathbf{V}_{22} & = \mathbf{N}_{22}
+\end{align}
+
+Finally, at $\mathcal{O}(\epsilon^3)$ we eliminate secular terms to close the system.
+Secular terms are terms which are resonant with the solution to the homogenous linear equation (Equation \ref{eq:ordere}), and which cause the higher-order solutions to grow without bound.
+The solvability criterion we enforce to eliminate these terms is the vanishing of the inner product of the solution to the adjoint linear homogenous equation $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$ with the nonhomogenous terms in Equation \ref{eq:ordere3}, namely
+
+\beq\label{eq:innerprod_gle}
+\langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle \partial_T \alpha + \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} \rangle \alpha +  \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} \rangle \partial_Z^2 \alpha = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle \alpha |\alpha|^2.
+\eeq
+
+This solvability criterion derives from a corollary to the Fredholm Alternative \citep[see][for a formal definition]{Clark:2016}.
+
+Equation \ref{eq:innerprod_gle} can be rewritten as Equation \ref{eq:gle}, the Ginzburg-Landau Equation, where the coefficients are
+
+\beq
+b = \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
+\eeq
+
+\beq
+h = \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
+\eeq
+
+and
+
+\beq
+c = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle. 
+\eeq
+
+We define the adjoint operator $\mathcal{L}^\dagger$ and solution $\mathbf{V}^\dagger$ as
+
+\beq
+\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \langle \mathcal{L}^\dagger \mathbf{V}^\dagger | \mathbf{V} \rangle,
+\eeq
+
+where the inner product is defined as
+
+\beq\label{eq:inner_product_def}
+\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \frac{k_c}{2\pi} \int_{-\pi/k_c}^{\pi/k_c} \int_{x_1}^{x_2} \mathbf{V}^{\dagger*} \cdot \mathcal{L} \mathbf{V} \, \mathrm{d} x \mathrm{d} z.
+\eeq
+
+The solution to the adjoint homogenous equation has the form
+
+\beq
+\mathbf{V^\dagger} = \mathbb{V}^\dagger(x) e^{i k_c z} + c.c.
+\eeq
+
 \section{B. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}

### 2016-10-04 paper/thingap_mri_paper.tex
a little more text in nonlinear appendix

@@
 \section{B. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}
+
+At each order in our perturbation series, lower-order MRI modes nonlinearly interact.
+Thus there is a nonlinear term contribution at $\mathcal{O}(\epsilon^2)$ and $\mathcal{O}(\epsilon^3)$.
+Here we detail the form of these nonlinear terms.
+
+The overall nonlinear contribution to our system, written as a vector $\mathbf{N}$ in Equation \ref{eq:unperturbed_matrix_equations}, is
+
 \beq

### 2016-10-04 paper/thingap_mri_paper.tex
added references; more Vasil discussion; added to abstract; wrote pointless "table of contents"

--- 2016-10-04 paper/thingap_mri_paper.tex
@@
 
-\title{The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor-Couette flow}
-\author{Clark, S.E.\altaffilmark{1}}
-\author{Oishi, J.S. \altaffilmark{2,}\altaffilmark{3}}
+\title{The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor Couette flow}
+\author{Susan E. Clark\altaffilmark{1}}
+\author{Jeffrey S. Oishi\altaffilmark{2,}\altaffilmark{3}}
 
@@
 The magnetorotational instability (MRI) is a fundamental process of accretion disk physics, but its saturation mechanism remains poorly understood despite considerable theoretical and computational effort.
-We present a multiple-scales analysis of the nonideal MRI in the weakly nonlinear regime -- that is, when the the most unstable MRI mode has a growth rate asymptotically approaching zero from above.
+We present a multiple-scales analysis of the non-ideal MRI in the weakly nonlinear regime -- that is, when the the most unstable MRI mode has a growth rate asymptotically approaching zero from above.
 Here, we develop our theory in a thin-gap, Cartesian channel.
@@
 The transport mechanism is likely turbulent, as molecular viscosity alone cannot account for the needed angular momentum transfer, and likely magnetic, as this turbulence is excited even in hydrodynamically stable disks (\citei{Shakura:1973wg}).
-Discovered by \citet{Chandrasekhar:1960wh} and \citet{Velikhov:1959} in global geometry, the MRI was subsequently rediscovered and applied to accretion disks by \citet{Balbus:1991vs}.
+Discovered by \citet{Chandrasekhar:1960wh} and \citet{Velikhov:1959} in global geometry, the MRI was subsequent rediscovered and applied to accretion disks by \citet{Balbus:1991vs}.
 Since then, the magnetorotational instability (MRI) remains the leading explanation for rapid angular momentum transport in astrophysical disks.
 The instability in its simplest geometry arises when a differentially rotating disk is threaded by a vertical magnetic field.
-The presence of the magnetic field linearly destabilizes the disk gas, driving turbulence and angular momentum transport.
-The MRI likely plays a role in a diverse host of astrophysical systems, including protoplanetary disks and black hole accretion disks, as well as stellar interiors.
+The presence of the magnetic field linearly destabilizes the disk gas, driving turbulence and angular momentum transport \citep[e.g.][]{2011ApJ...738...84H,2014MNRAS.438.2513P,2013MNRAS.435.2281P}.
+The MRI likely plays a role in a diverse host of astrophysical systems, including protoplanetary disks \citep[e.g.]{2015ApJ...798...84B} and black hole accretion disks \citep[e.g.][]{2013ApJ...769..156S}, as well as stellar interiors \citep{2015ApJ...799...85W}.
 Despite its importance, many aspects of the MRI remain poorly understood.
@@
 Weakly nonlinear analysis is a perturbative method used to examine the asymptotic behavior of a system near threshold -- that is, when the system is just barely unstable to its most unstable mode.
-The analytical technique follows the multiscale evolution of fluid variables in a perturbation expansion, allowing the controlled interaction of modes between orders in a perturbation series (see, e.g., Bender \& Orszag for a full mathematical formalism).
+The analytical technique follows the multiscale evolution of fluid variables in a perturbation expansion, allowing the controlled interaction of modes between orders in a perturbation series \citep{1978amms.book.....B}.
 Weakly nonlinear analysis can be a powerful technique for analytically examining systems which in their full generality exhibit such complicated nonlinear behavior that their study is relegated primarily to the simulation domain.
 The MRI is one such phenomenon: while there is a rich literature analytically examining the linear MRI, analytical treatments of the \textit{nonlinear} system are relatively few.
-The weakly nonlinear treatment of the MRI was pioneered by \citeauthor{Umurhan:2007dz} (\citeyear{Umurhan:2007dz}, \citeyear{Umurhan:2007hs}; hereafter URM07).
-The authors undertook the first weakly nonlinear analysis of the MRI in a thin-gap Taylor-Couette flow, and found that the marginal MRI system approaches saturation in a manner analogous to that of Rayleigh-B\'enard convection.
+The weakly nonlinear treatment of the MRI was pioneered by \citet{Knobloch:2005ba} and \citet[][hereafter URM07]{Umurhan:2007dz}.
+The latter authors undertook the first weakly nonlinear analysis of the MRI in a thin-gap Taylor Couette flow with strong dissipation (as is appropriate to liquid metal experiments), and found that the marginal MRI system approaches saturation in a manner analogous to that of Rayleigh-B\'enard convection.
 Weakly nonlinear analysis was instrumental in our understanding of Rayleigh-B\'enard convection saturation (\citei{Newell:1969}), and the similarities between convection and the thin-gap MRI are the result of important shared symmetries between the systems.
 The success of URM07 in modeling the MRI system near threshold merits further consideration, but we are unaware of any other attempts to expand upon their theoretical framework.
-In this work we rigorously rederive the theory of URM07, and expand upon their findings.
+In this work we rederive the theory of URM07, and expand upon their findings.
 Our focus here is on fully characterizing the narrow gap MRI system, both for independent theoretical interest and to have a robust comparison point for extensions of this theory into more complicated geometries.
-In a companion paper we derive for the first time the weakly nonlinear theory of the standard and helical MRI in a wide gap Taylor-Couette flow \citep[][hereafter Paper II]{Clark:2016}.
+In a companion paper we derive for the first time the weakly nonlinear theory of the standard and helical MRI in a wide gap Taylor Couette flow \citep[][hereafter Paper II]{Clark:2016}.
 The thin- and wide-gap treatments complement one another theoretically, and both are important regimes for comparison with simulation.
 
-This work examines Taylor-Couette flow in the thin-gap regime, an idealization in which the radial extent of the channel is very small compared to its distance from the center of rotation, i.e.
+This work examines Taylor Couette flow in the thin-gap regime, an idealization in which the radial extent of the channel is very small compared to its distance from the center of rotation, i.e.
 $(r_2 - r_1) \ll \frac{1}{2} (r_1 + r_2)$ where $r_1$ and $r_2$ are the radii of the inner and outer flow boundaries, respectively.
@@
 In particular, in the wide-gap geometry (i.e.
-true Taylor-Couette flow) the base angular velocity is a function of $r$, where in the thin-gap approximation the shear flow reduces to a linear profile.
+true Taylor Couette flow) the base angular velocity is a function of $r$, where in the thin-gap approximation the shear flow reduces to a linear profile.
 
@@
 \citeauthor{Knobloch:2005ba} (\citeyear{Knobloch:2005ba}) analyze the MRI in the strongly nonlinear regime, by following the already-developed MRI modes into asymptotic saturation.
-They find that this saturation proceeds via modification of the background shear.
-\citeauthor{Knobloch:2005ba} consider a thin-gap regime as well, and so their theory may be considered the strongly nonlinear analogue to the one developed here.
-\citei{Vasil:2015} examines the weakly nonlinear MRI in a thin-gap regime from a heuristic perspective, finding deep mathematical similarities between the MRI system and an elastodynamic instability.
+They consider a thin-gap regime as well, and so their theory may be considered the strongly nonlinear analogue to the one developed here.
+\citei{Vasil:2015} examines the weakly nonlinear MRI in a thin-gap regime in a minimal model, finding deep mathematical similarities between the MRI system and the elastodynamic instability of a buckling beam.
+We discuss his results and their relation to ours in section~\ref{sec:discussion}.
 Several authors have investigated the behavior of the MRI when the boundary conditions are shear periodic, and so the MRI has no mechanism by which to modify the background shear flow profile.
@@
 The theory of MRI channel mode parasites is robust (see also \citei{Pessah:2009gm}, \citei{Latter:2010iz}, \citei{Rembiasz:2016}), but their importance may be overestimated by the local approximation \citei{Latter:2015}, and not germane to global analyses like the one presented here.
-\citei{Latter:2015} gives a  detailed analysis of the relation between local and global MRI modes.
+\citei{Latter:2015} gives a detailed analysis of the relation between local and global linear MRI modes.
+
+We begin with an overview of our basic model equations for the thin gap MRI in section~\ref{sec:equations} and then describe our weakly non-linear analysis and give basic results in section~\ref{sec:wnl}.
+We then place our results in the ocntext of previous results from both analytic and computational studies in section~\ref{sec:discussion} and draw conclusions in section~\ref{sec:conclusion}.
 
@@
 \includegraphics[trim=5cm 0cm 0cm 0cm, scale=.35]{setup_diagram.eps}
-\caption{Schematic diagram of our set-up, an axisymmetric thin-gap Taylor-Couette flow.
+\caption{Schematic diagram of our set-up, an axisymmetric thin-gap Taylor Couette flow.
 We investigate a 2D slice of the X-Z (radial-vertical) plane.
@@
 %See Figure \ref{figscalings}. 
-Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic Taylor-Couette flow.
+Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic Taylor Couette flow.
 Our $x$ dimension, the direction of angular momentum transport, is analogous to the direction of temperature transport in the convection problem.
@@
 \caption{Top panel: Mean radial profile for $u_y^{saturated} - u_y^0$, the difference between the saturated azimuthal velocity profile and the initial azimuthal velocity profile (TC flow).
-Bottom panel: each term in the steady state force balance (Equation \ref{eqset2} with $\partial_t u_{y} = 0$).
-Saturated quantities are computed with $\alpha_{saturation} = \sqrt{b/c}$ and $\epsilon = 0.5$.
-The saturated state shows reduced shear in the bulk of the flow, outside of the boundary layers.} \label{fig:sat_comp_vel}
+Bottom panel: each term in the azimuthal velocity steady state force balance (Equation \ref{eqset2} with $\partial_t u_{y} = 0$).} \label{fig:sat_comp_vel}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{../figures/thingap_saturation_mechanism_plots_Bfield_slice60.png}
-\caption{Top panel: Mean radial profile for $B_z^{saturated}$, the saturated vertical magnetic field (black line).
-$B_z^0 = 1$ is the constant background magnetic field (gray line).
-Bottom panel: each term in the steady state inductive balance ($\partial_x$ of Equation \ref{eqset4} with $\partial_t B_y = 0$).
-Saturated quantities are computed with $\alpha_{saturation} = \sqrt{b/c}$ and $\epsilon = 0.5$.
-The saturated field is pushed to the radial domain boundaries.}\label{fig:sat_comp_Bfield}
+\caption{.}\label{fig:sat_comp_Bfield}
 \end{figure}
@@
 
-Figure~\ref{fig:sat_comp_Bfield} shows $B_z$ and the terms corresponding to steady state inductive balance ($\partial_x$ of Equation~\ref{eqset4} with $\partial_t B_{y} = 0$).
+Figure~\ref{fig:sat_comp_Bfield} shows $B_z$ and the terms corresponding to steady state inductive balance (Equation~\ref{eqset2} with $\partial_t B_{y} = 0$).
 Here, the instability acts to push magnetic field toward the boundaries in both the bulk and the boundary layers.
@@
 Whether and how \citet{Vasil:2015}'s amplitude equation is equivalent to our own in the limit of dynamically important resistive and viscous effects is beyond the scope of this work.
+Nevertheless, he identifies the non-linear term responsible for saturation as consisting of flux and field transport and notes these are the only mechanisms that can produce saturation.
+Our results also demonstrate a combination of flux and field transport.
+This suggests that despite our formulation displaying different saturation dynamics (Ginzburg-Landau in our case; a network of coupled Duffing oscillators in his), there may indeed be an underlying unification.
 
@@
 
-\section{Acknowledgements}
+\section{Acknowledgments}
 S.E.C.
@@
 
+\bibitem[Bai(2015)]{2015ApJ...798...84B} Bai, X.-N.\ 2015, \apj, 798, 84
+
 \bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}
 Balbus, S A and Hawley, J F, 1991, The Astrophysical Journal, 376, 214
+
+\bibitem[Bender \& Orszag(1978)]{1978amms.book.....B} Bender, C.~M., \& Orszag, S.~A.\ 1978, Advanced Mathematical Methods for Scientists and Engineers, New York: McGraw-Hill, 1978,
 
@@
 Goodman, J, Xu, G, 1994, ApJ, 432, 213
+
+\bibitem[Hawley et al.(2011)]{2011ApJ...738...84H} Hawley, J.~F., Guan, X., \& Krolik, J.~H.\ 2011, \apj, 738, 84
 
@@
 
+\bibitem[Parkin(2014)]{2014MNRAS.438.2513P} Parkin, E.~R.\ 2014, \mnras, 438, 2513
+
+\bibitem[Parkin \& Bicknell(2013)]{2013MNRAS.435.2281P} Parkin, E.~R., \& Bicknell, G.~V.\ 2013, \mnras, 435, 2281
+
 \bibitem[{Pessah(2010)}]{Pessah:2010ic}
@@
 Rembiasz, T., Obergaulinger, M., Cerd\'a-Dur\'an, P., M\"uller, E., \& Aloy, M.A., 2016, MNRAS 456, 3782
+
+\bibitem[Schnittman et al.(2013)]{2013ApJ...769..156S} Schnittman, J.~D., Krolik, J.~H., \& Noble, S.~C.\ 2013, \apj, 769, 156
 
@@
 
+\bibitem[Wheeler et al.(2015)]{2015ApJ...799...85W} Wheeler, J.~C., Kagan, D., \& Chatzopoulos, E.\ 2015, \apj, 799, 85
+
 \end{thebibliography}
@@
 \end{document}
+
+%  LocalWords:  Magnetorotational Couette Oishi Lewiston Umurhan et
+%  LocalWords:  magnetorotational al Ginzburg accrete Shakura wg wh
+%  LocalWords:  hydrodynamically Velikhov Balbus protoplanetary ba ul
+%  LocalWords:  perturbative analytical multiscale Orszag Knobloch ic
+%  LocalWords:  URM enard Newell Vasil elastodynamic Pessah iz Alfv
+%  LocalWords:  Rembiasz axisymmetric Chebyshev diffusivity zA xA aa
+%  LocalWords:  incompressibility axisymmetrically nondimensionalize
+%  LocalWords:  nondimensionalized nondimensional streamfunction ey
+%  LocalWords:  Jacobian Colorbar azimuthal ansatz marginality Runge
+%  LocalWords:  wavenumber pseudospectral Dedalus solvability Kutta
+%  LocalWords:  axisymmetry advection Ebrahimi Keplerian timestep Im
+%  LocalWords:  timesteps crit arctan colorbar hs DGE LWS Mordecai Xu
+%  LocalWords:  Krommes Geoff Zweibel natexlab Bai Hawley Prager ApJ
+%  LocalWords:  Schnack Guan Krolik Rebecca Julien Fromang Faure JFM
+%  LocalWords:  MNRAS Gressel Parkin Bicknell Obergaulinger Cerd Dur
+%  LocalWords:  uller Aloy Schnittman Sunyaev Regev Menou RSPSA Sov
+%  LocalWords:  JETP Kagan Chatzopoulos

### 2016-10-04 paper/thingap_mri_paper.tex

@@
 
-This work examines Taylor Couette flow in the thin-gap regime, an idealization in which the radial extent of the channel is very small compared to its distance from the center of rotation, i.e.
+This work examines Taylor-Couette flow in the thin-gap regime, an idealization in which the radial extent of the channel is very small compared to its distance from the center of rotation, i.e.
 $(r_2 - r_1) \ll \frac{1}{2} (r_1 + r_2)$ where $r_1$ and $r_2$ are the radii of the inner and outer flow boundaries, respectively.
@@
 
+Note that, following the notation of \citet{Umurhan:2007hs}, the subscripts refer to $\epsilon$ order, $z$ order, successively, such that $\mathbf{N}_{22}$ is the second-order nonlinear term which corresponds to $e^{2 i k_c z}$ z-dependence.
+
+Equation \ref{eq:ordere2} is solved as three separate systems of equations, one for each possible $z$ resonance:
+
+\begin{align}
+\mathcal{L}\mathbf{V}_{20} & = \mathbf{N}_{20}\\
+\mathcal{L}\mathbf{V}_{21} & = - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_{11} \\
+\mathcal{L}\mathbf{V}_{22} & = \mathbf{N}_{22}
+\end{align}
+
+Finally, at $\mathcal{O}(\epsilon^3)$ we eliminate secular terms to close the system.
+Secular terms are terms which are resonant with the solution to the homogenous linear equation (Equation \ref{eq:ordere}), and which cause the higher-order solutions to grow without bound.
+The solvability criterion we enforce to eliminate these terms is the vanishing of the inner product of the solution to the adjoint linear homogenous equation $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$ with the nonhomogenous terms in Equation \ref{eq:ordere3}, namely
+
+\beq\label{eq:innerprod_gle}
+\langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle \partial_T \alpha + \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} \rangle \alpha +  \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} \rangle \partial_Z^2 \alpha = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle \alpha |\alpha|^2.
+\eeq
+
+This solvability criterion derives from a corollary to the Fredholm Alternative \citep[see][for a formal definition]{Clark:2016}.
+
+Equation \ref{eq:innerprod_gle} can be rewritten as Equation \ref{eq:gle}, the Ginzburg-Landau Equation, where the coefficients are
+
+\beq
+b = \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
+\eeq
+
+\beq
+h = \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
+\eeq
+
+and
+
+\beq
+c = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle. 
+\eeq
+
+We define the adjoint operator $\mathcal{L}^\dagger$ and solution $\mathbf{V}^\dagger$ as
+
+\beq
+\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \langle \mathcal{L}^\dagger \mathbf{V}^\dagger | \mathbf{V} \rangle,
+\eeq
+
+where the inner product is defined as
+
+\beq\label{eq:inner_product_def}
+\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \frac{k_c}{2\pi} \int_{-\pi/k_c}^{\pi/k_c} \int_{x_1}^{x_2} \mathbf{V}^{\dagger*} \cdot \mathcal{L} \mathbf{V} \, \mathrm{d} x \mathrm{d} z.
+\eeq
+
+The solution to the adjoint homogenous equation has the form
+
+\beq
+\mathbf{V^\dagger} = \mathbb{V}^\dagger(x) e^{i k_c z} + c.c.
+\eeq
+
 \section{B. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}
+
+At each order in our perturbation series, lower-order MRI modes nonlinearly interact.
+Thus there is a nonlinear term contribution at $\mathcal{O}(\epsilon^2)$ and $\mathcal{O}(\epsilon^3)$.
+Here we detail the form of these nonlinear terms.
+
+The overall nonlinear contribution to our system, written as a vector $\mathbf{N}$ in Equation \ref{eq:unperturbed_matrix_equations}, is
+
 \beq

### 2016-10-04 paper/widegap_mri_paper.tex
two ways to compute nablasq
minor changes.

@@
-%\documentclass[onecolumn]{emulateapj}
 \documentclass{emulateapj}
@@
 \usepackage{color}
-%\usepackage{appendix}
 \usepackage[titletoc,title]{appendix}
-\usepackage{cleveref}
 
@@
 \begin{document}
-%\title{Weakly Nonlinear Analysis of the Standard and Helical Magnetorotational Instabilities in a Taylor Couette Flow}
-\title{The Weakly Nonlinear Magnetorotational Instability in a Global, Cylindrical Taylor-Couette Flow} % maybe this is punchier?
-\author{Clark, S.E.\altaffilmark{1}}
-\author{Oishi, J.S. \altaffilmark{2,}\altaffilmark{3}}
+\title{The Weakly Nonlinear Magnetorotational Instability in a Global, Cylindrical Taylor-Couette Flow} 
+\author{Susan E. Clark\altaffilmark{1}}
+\author{Jeffrey S. Oishi\altaffilmark{2,}\altaffilmark{3}}
 \altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
@@
 
-%\begin{widetext}
 \begin{table*}[ht]
@@
 \end{table*}
-%\end{widetext}
 
@@
 Thus while the shearing box may accurately approximate many features of the global MRI, the saturation mechanism may not be among them.
-
-%Because of the MRI's complexity, analytical treatments of the MRI use some combination of convenient approximations, such as ideal MHD, linearized equations, expedient boundary conditions, and simplified geometries.
-In particular, much of the analytical work on MRI saturation has focused on local approximations, often representing a section of a disk by solving the MHD equations in an isolated region subject to shear periodic boundary conditions (the ``shearing box").
-The shearing box is an inexpensive computational framework, so is used by many numerical MRI studies as well.
-However, while the MRI is a local instability, certain properties of the local problem are not generic to the global problem.
-In the shearing box, linear evolution is dominated by a class of MRI mode known as channel modes.
-%Channel modes, a class of MRI mode that dominates MRI evolution and saturation in the local approximation, are artificially over-prominent in the shearing box (e.g.
-\citei{Latter:2015}).
-%These are linear modes which also happen to be exact solutions to the \textit{nonlinear} local MRI equations.
-Runaway growth is avoided in this paradigm by the instability of the channel modes themselves, which are destroyed by parasitic (secondary) instabilities (\citei{Goodman:1994ul}, \citei{Pessah:2010ic}).
-The growth of parasitic modes provides a saturation avenue for channel mode-dominated flows, yet this is unlikely to be the dominant saturation mechanism in laboratory experiments or astrophysical disks, as channel modes are artificially over-prominent in the shearing box (e.g.
-\citei{Latter:2015}).
 
@@
 \beq\label{momentum}
-%\begin{split}
 \partial_t \mathbf{u} + \mathbf{u} \cdot \nabla \mathbf{u} = -\frac{1}{\rho}\nabla P - \nabla\Phi + \frac{1}{c\rho} \left(\mathbf{J}\times\mathbf{B}\right) + \nu\nabla^2 \mathbf{u} 
-%\end{split}
 \eeq
@@
 
-%and define the magnetic vector potential $A$ analogously. 
 These definitions automatically satisfy Equations \ref{eq:incompressibility} and \ref{eq:solenoid} for axisymmetric disturbances.
@@
 \begin{align}
-%\begin{split}
 & \frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r^2}B_0 \xi \partial_z B_\phi \nonumber\\
 & - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi - \frac{2}{r^2}\partial_r\Psi) - \frac{1}{r^3} \nabla^2 \Psi + \frac{2}{r^4}\partial_r\Psi\right] = N^{(\Psi)} \label{eq:Psi_perturbed}
-%\end{split}
 \end{align}
@@
 We solve the equations at each order using Dedalus, an open source pseudospectral code.
-We solve the radial portion of the eigenvectors on a basis of Chebyshev polynomials subject to our radial boundary conditions.
-%Because our domain is bounded only radially and not vertically, we solve the radial dimension on a basis of Chebyshev polynomials, and the vertical dimension on a Fourier series basis.
+We solve the radial portion of the eigenvectors on a basis of Chebyshev polynomials subject to our radial boundary conditions
 We solve Equation \ref{eq:ordere} as a linear eigenvalue problem, and Equation \ref{eq:ordere2} as a linear boundary value problem. 
- %should i mention spurious eigenvals? 
-%To analyze the MRI system at marginality, we fix the parameters listed in Table \ref{table:parameters} and determine the critical $\reym$ and $k_z$ by repeatedly solving Equation \ref{eq:ordere} to determine the smallest parameter values for which the fastest growing mode is zero.
-That is, we solve the linear eigenvalue problem for eigenvalues $\sigma = \gamma + i \omega$.
-Figure \ref{fig:growth_rates} shows linear MRI growth rates $\gamma$ in the $(\reym, k_z)$ plane.
-For the fiducial standard MRI parameters in Table \ref{table:parameters} we find critical parameters $\reym_c = 0.84$, $k_c = 0.0127$.
-
-%We defer full expressions and derivations thereof to the Appendix, and focus here on features of the wide gap system which distinguish it from the narrow gap limit.
-
 The result of the weakly nonlinear analysis is a single amplitude equation for $\alpha$. This amplitude equation is found by enforcing a solvability condition on Equation \ref{eq:ordere3}. We find
@@
 
-%For the standard MRI we derive a real GLE, and for the helical MRI we derive a complex GLE. 
- %This complex GLE is a natural form for the HMRI, as it is a traveling wave instability.
-
 \section{Results}
 \label{sec:results}
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=\textwidth]{../figures/fiducial_standard_mri_stressplot_Pm_16E-7_cc.png}
-%\caption{Reynolds ($\mathbb{T}_{R} = u_r u_\phi$), Maxwell ($\mathbb{T}_{M} = -\frac{2}{\beta} B_r B_\phi$), %and total stress ($\mathbb{T} = \mathbb{T}_{R} + \mathbb{T}_{M}$) for the fiducial standard MRI case.
-}
-%\end{figure*}
 
@@
 \end{figure}
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=\columnwidth]{../figures/widegap_3D_vel_screenshot.png}
-%\caption{Visualization of the velocity perturbation field in the fiducial SMRI case.
-The three dimensional path of t%wenty randomly placed tracer particles is plotted, where the particle velocity is defined as the total (first + second %order) velocity perturbation.
-This is meant to be a visualization only of the velocity field, so the particle motion f%eels no contribution from the magnetic field.}
-%\end{figure}
 
@@
 Because that symmetry is not preserved in the wide-gap case, \citet{Umurhan:2007hs} postulate that slow variation in the wide-gap geometry will be governed by two coupled amplitude equations.
-%However, the wide-gap conducting boundary condition $J_z = \frac{1}{r} \partial_r (r B_\phi) = 0$, together with the purely geometric term in Equation \ref{eq:Bphi_perturbed}, conspire to prevent the wide-gap geometry from sustaining a neutral mode.
 However, the purely geometric term in Equation \ref{eq:Bphi_perturbed} prevents the wide-gap geometry from sustaining a neutral mode.
@@
 The nonlinear terms ultimately determine the saturation amplitude of the system, as described below.
-%These advective terms allow the nonlinear contributions at both second and third order (i.e.
-N$_2$ and N$_3$) not to individually satisfy the boundary conditions on $\Psi$ and $u$.
-%(I'm just guessing but this may lead to higher torque on inner cylinder -- should calculate)
 
@@
 From Equation \ref{eq:gle} it is readily apparent that the asymptotic saturation amplitude is $\alpha_{s} = \pm \sqrt{b/c}$, and we plot the dependence of $\alpha_{s}$ on $\Pm$ in the bottom panel of \ref{fig:coefficients}.
-%The GLE $b$ and $d$ coefficients are insensitive to $\Pm$, so the $\alpha_{s}$ scaling is driven by the dependence of the nonlinear coefficient $c$ on $\Pm$.
 Note that because $\reym$ is essentially constant as a function of $\Pm$, the saturation amplitude is equivalently sensitive to $\reye^{-1}$.
@@
 Boundary layer analysis similarly reveals a $\nu^{1/3}$ dependence for the radial extent of the boundary layer in TC flow (\citei{Goodman:2002ix}).
-%Why does the $\Pm$ scaling of the boundary layer width not translate to a stronger $\Pm$ scaling for the saturation amplitude in the wide gap case?
-The boundary layer in the wide gap case is strongly localized at the walls, i.e.
-$r_1$ and $r_2$.
 Figure \ref{fig:thin_wide_comparison} shows the structure of the third-order nonlinear term $N_{31}^{(A)}$ as a function of $\Pm$ for both the narrow and wide gap standard MRI.
@@
 Clearly, the width of the boundary layers scales with $\Pm$ in both the wide and narrow gap MRI.
-This translates to a steeper saturation amplitude $\Pm$ dependence in the wide gap case.%However, in the narrow gap case this scaling extends prominently into the center of the channel, whereas for the wide gap case the bulk of the flow is relatively unaffected by the boundary layer scaling.
+This translates to a steeper saturation amplitude $\Pm$ dependence in the wide gap case.
 
@@
 
-%\section{Channel Modes}
-%Channel modes are radially independent linear MRI modes which are, under certain conditions, exact nonlinear solutions to the MRI equations.
-
 \subsection{Helical MRI}
@@
 The complex GLE may be unstable to traveling wave instabilities such as the Benjamin-Feir instability, a generalization of the Ekhaus instability.
-The complex GLE can also admit spatiotemporal chaos, and various classes of coherent structures (\citei{Aranson:2002}).
-%Because the saturation amplitude of the weakly nonlinear helical MRI is governed by a complex GLE, the saturated state of the helical MRI may be unstable to 
-Although a detailed description of the phase dynamics in the helical MRI is beyond the scope of this work, we note that such long-wavelength, long-timescale behavior may be observed in liquid metal helical MRI experiments.
+The complex GLE can also admit spatiotemporal chaos, and various classes of coherent structures (\citei{Aranson:2002}).Although a detailed description of the phase dynamics in the helical MRI is beyond the scope of this work, we note that such long-wavelength, long-timescale behavior may be observed in liquid metal helical MRI experiments.
 
@@
 If we take the plane of the fluid to be infinite in the horizontal plane, the system is initially translationally symmetric.
-%When the system eventually becomes unstable to convective heat transfer it undergoes a symmetry breaking, allowing the
 At the onset of convection the system undergoes a symmetry breaking, forming rolls, or convection cells, which break the horizontal translational invariance.
@@
 
-%As this formalism is extended to more astrophysically relevant conditions  
-%The application and extension of the formalism presented here to more astrophysically representative conditions will allow an analytical understanding of weakly nonlinear MRI behavior.
-
-%We find that the saturation amplitude of the standard MRI scales somewhat more steeply with $\Pm$ than does this same system system in the narrow gap approximation.
-
 \section{Acknowledgments}
@@
 
-Here we detail the perturbation analysis described in Section \ref{sec:wnl_analysis}. The linear system is described by Equation \ref{eq:unperturbed_matrix_eqns}, where
+Here we detail the perturbation analysis described in Section \ref{sec:wnl_analysis}. The perturbation series is described by Equations \ref{eq:ordere} - \ref{eq:ordere3}, where
 
@@
 
-%The GLE is derived from 
-%Equation \ref{eq:ordere3} is resonant with 
 To find a bounded solution at $\mathcal{O}(\epsilon^3)$ we must eliminate secular terms: terms which are resonant with the solution to the linear homogenous equation $(\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} = 0$ and cause the solution to grow without bound.
@@
 
-% This solvability condition is the requirement that the inner product of the equation with its adjoint homogenous solution is zero. 
 We define the adjoint operator $\mathcal{L}^\dagger$ and solution $\mathbf{V}^\dagger$ as
@@
 
-%\bibitem[{Chandrasekhar(1960)}]{Chandrasekhar:1960wh}
-%Chandrasekhar, S. 1960, PNAS 46, 253
-
 \bibitem[{Clark \& Oishi(2016a)}]{Clark:2016}

### 2016-10-04 paper/thingap_mri_paper.tex
conclusion edits

@@
 
-$A$ is thus related to the magnetic field as
+$A$ and $\Psi$ are thus related to the magnetic field and velocity perturbations, respectively, as
 
@@
 -\partial_xA \\
-\end{matrix}\right],\eeq \\
+\end{matrix}\right],\eeq
 
-and $\mathbf{\Psi}$ is defined analogously.
+\beq
+\mathbf{u} \, = \, \left[\begin{matrix}
+\partial_z \Psi \\
+u_{y} \\
+-\partial_x \Psi \\
+\end{matrix}\right].\eeq
+
+%and $\mathbf{\Psi}$ is defined analogously.
 
@@
 
+The weakly nonlinear regime is where the MRI system is nonlinearly unstable to only the most unstable mode of the linear solution.
+We find the marginal state, where the most unstable linear MRI mode neither grows nor decays, for a given set of dimensionless parameters.
+
 \begin{figure*}
@@
 We use a constant amplitude $\alpha = \alpha_{saturation}$ and a small parameter $\epsilon = 0.5$.}\label{fig:allorders_velocity}
-\end{figure*}
-
-\begin{figure*}
+%\end{figure*}
+\vspace*{\floatsep}
+%\begin{figure*}[h!]
 \centering
@@
 \end{figure*}
-
-The weakly nonlinear regime is where the MRI system is nonlinearly unstable to only the most unstable mode of the linear solution.
-We find the marginal state, where the most unstable linear MRI mode neither grows nor decays, for a given set of dimensionless parameters.
 
@@
 
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_2.png}
+\caption{Evolution of the Ginzburg-Landau amplitude equation (Equation \ref{eq:gle}) on a Fourier $Z$ domain of length $2 \lambda_{crit}$, where $\lambda_{crit} = 2\pi/k_c$ is the critical wavelength of the system.
+Top left panel shows the evolution of the amplitude observable $\alpha \alpha^*$ on the full $Z$ domain as a function of time $T$.
+Bottom left panel shows the amplitude $alpha$ at the final timestep shown, where the black line is the real part $\mathrm{Re}\{\alpha(t = t_{final})\}$ and the gray line is the imaginary part $\mathrm{Im}\{\alpha(t = t_{final})\}$.
+The final amplitude is bounded by the analytic saturation amplitude $\alpha_{saturation} = \pm \sqrt{b/c}$.
+Top right panel shows the evolution of the phase angle $\phi = \mathrm{arctan} (\mathrm{Im}(\alpha)/\mathrm{Re}(\alpha))$ on the same domain.
+Bottom panel shows the phase angle as a function of $Z$ for the final timestep.
+Note that the phase angle is wrapped on a $2\pi$ domain, such that $\pi$ = $-\pi$, as indicated by the circular colorbar.}\label{fig:IVP_2lambdacrit}
+%\end{figure*}
+
+%\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_10.png}%{allorders_velocity.eps}
+\caption{As in Figure \ref{fig:IVP_2lambdacrit} but for a $Z$ domain of length $10 \lambda_{crit}$.}\label{fig:IVP_10lambdacrit}
+\end{figure*}
+
 Figure~\ref{fig:sat_comp_Bfield} shows $B_z$ and the terms corresponding to steady state inductive balance ($\partial_x$ of Equation~\ref{eqset4} with $\partial_t B_{y} = 0$).
@@
 
-\begin{figure*}[h!]
-\centering
-\includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_2.png}
-\caption{Evolution of the Ginzburg-Landau amplitude equation (Equation \ref{eq:gle}) on a Fourier $Z$ domain of length $2 \lambda_{crit}$, where $\lambda_{crit} = 2\pi/k_c$ is the critical wavelength of the system.
-Top left panel shows the evolution of the amplitude observable $\alpha \alpha^*$ on the full $Z$ domain as a function of time $T$.
-Bottom left panel shows the amplitude $alpha$ at the final timestep shown, where the black line is the real part $\mathrm{Re}\{\alpha(t = t_{final})\}$ and the gray line is the imaginary part $\mathrm{Im}\{\alpha(t = t_{final})\}$.
-The final amplitude is bounded by the analytic saturation amplitude $\alpha_{saturation} = \pm \sqrt{b/c}$.
-Top right panel shows the evolution of the phase angle $\phi = \mathrm{arctan} (\mathrm{Im}(\alpha)/\mathrm{Re}(\alpha))$ on the same domain.
-Bottom panel shows the phase angle as a function of $Z$ for the final timestep.
-Note that the phase angle is wrapped on a $2\pi$ domain, such that $\pi$ = $-\pi$, as indicated by the circular colorbar.}\label{fig:IVP_2lambdacrit}
-%\end{figure*}
-
-%\begin{figure*}[h!]
-\centering
-\includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_10.png}%{allorders_velocity.eps}
-\caption{As in Figure \ref{fig:IVP_2lambdacrit} but for a $Z$ domain of length $10 \lambda_{crit}$.}\label{fig:IVP_10lambdacrit}
-\end{figure*}
-
 \section{Conclusion}
@@
 In doing so, we find the saturated state is a complex balance in which reduction of shear and amplification and redistribution of $B_z$ combine to saturate the instability.
-We also perform direct numerical simulations of the amplitude equation demonstrating that complex patterns can for on long length scales $Z$ though the maximum magnitude of the amplitude $\alpha$ is well predicted by the steady state solution.
-In paper II, we make use of the techniques developed here to extend the weakly nonlinear analysis of the MRI to a full cylindrical geometry appropriate for a wide gap experiment.
+We also perform direct numerical simulations of the amplitude equation.
+We demonstrate that complex patterns can organize the flow on long length scales $Z$, though the maximum magnitude of the amplitude $\alpha$ is well predicted by the steady state solution.
+In paper II, we make use of the techniques developed here to extend the weakly nonlinear analysis of the MRI to a full cylindrical geometry appropriate for a Taylor-Couette experiment.
 
@@
 
-This solvability criterion derives from a corollary to the Fredholm Alternative \citep[see][for a formal definition]{Clark:2016}.
+This solvability criterion derives from a corollary to the Fredholm Alternative (see Paper II for a formal definition).
 
@@
 \bibitem[{Clark \& Oishi(2016b)}]{Clark:2016}
-Clark, S.E. and Oishi, J.S., 2016, in prep
+Clark, S.E. and Oishi, J.S., 2016, submitted
 

### 2016-10-04 paper/thingap_mri_paper.tex
affiliations changes.

@@
 \title{The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor Couette flow}
-\author{Susan E. Clark\altaffilmark{1}}
+\author{S. E. Clark\altaffilmark{1}}
 \author{Jeffrey S. Oishi\altaffilmark{2,}\altaffilmark{3}}
@@
 \altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
-\altaffiltext{2}{Department of Astrophysics, American Museum of Natural History, New York, NY}
-\altaffiltext{3}{Department of Physics and Astronomy, Bates College, Lewiston, ME}
+\altaffiltext{2}{Department of Physics and Astronomy, Bates College, Lewiston, ME}
+\altaffiltext{3}{Department of Astrophysics, American Museum of Natural History, New York, NY}
 

### 2016-10-04 paper/widegap_mri_paper.tex
affiliations changes.

@@
 \title{The Weakly Nonlinear Magnetorotational Instability in a Global, Cylindrical Taylor-Couette Flow} 
-\author{Susan E. Clark\altaffilmark{1}}
+\author{S. E. Clark\altaffilmark{1}}
 \author{Jeffrey S. Oishi\altaffilmark{2,}\altaffilmark{3}}
 \altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
-\altaffiltext{2}{Department of Astrophysics, American Museum of Natural History, New York, NY}
-\altaffiltext{3}{Department of Physics and Astronomy, Bates College}
+\altaffiltext{2}{Department of Physics and Astronomy, Bates College}
+\altaffiltext{3}{Department of Astrophysics, American Museum of Natural History, New York, NY}
 

### 2016-10-04 paper/widegap_mri_paper.tex
edited saturation mechanism text to reflect our findings in paper I

@@
 the sum of the Reynolds and Maxwell stresses in our domain (Figure \ref{fig:total_stress}).
-We find that the saturation mechanism for weakly nonlinear TC flow is a reduction in the background shear profile.
-Thus the bounded cylindrical geometry considered here saturates via the same mechanism as in the narrow gap limit (\citei{Umurhan:2007hs}).
+As in Paper I, we find that the saturation mechanism for weakly nonlinear TC flow is a combination of reduced shear and redistributed and amplified background $B_z$.
+This strongly suggests that the underlying physics remains the same in the wide gap geometry, despite the addition of curvature terms.
 

### 2016-10-04 paper/thingap_mri_paper.tex

@@
 Here, we develop our theory in a thin-gap, Cartesian channel.
-Our results confirm Umurhan et al.'s findings that the saturation amplitude follows a Ginzburg-Landau equation.
+Our results confirm Umurhan et al.
+2007's findings that the saturation amplitude follows a Ginzburg-Landau equation.
 We extend these results by performing a detailed force balance for the saturated $u_y$ and $B_z$ fields, demonstrating that even when diffusive effects are important, the bulk flow saturates by a combination of reducing the background shear and rearranging and strengthening the background $B_z$ field.
@@
 We plot the first order, second order, and total perturbation structure of the fluid variables in Figures \ref{fig:allorders_velocity} and \ref{fig:allorders_Bfield} with a constant $\alpha_{saturation}$.
+This is the Ginzburg-Landau equation that was previously found by URM07.
+Those authors investigated the behavior of this MRI system as a function of $\Pm$.
+We note that we reproduce the URM07 result that the analytic saturation amplitude scales as $\alpha^2 \propto \Pm^{4/3}$.
 
@@
 The real Ginzburg-Landau equation describes the amplitude behavior of our system close to threshold.
-Although the form of the equation is generic, its coefficients depend on the specific physics of our system and govern its detailed evolution (see Appendix \ref{app:matrices}).
+Although the form of the equation is generic to many systems, its coefficients depend on the specific physics of our system and govern its detailed evolution (see Appendix \ref{app:matrices}).
 We simulate the evolution of the MRI amplitude equation by solving Equation \ref{eq:gle} on a Fourier basis in $Z$ using Dedalus.
@@
 While still bounded by $\alpha_s$, the saturation amplitude exhibits a nonlinear phase geometry due to the nonlinear interaction of modes in $Z$.
+%The saturation amplitude, while not required to be trivially constant, has the same maximum magnitude as the steady-sate system.
 
@@
 
-In this paper, we have constructed a weakly nonlinear analysis of the MRI using multiple-scales analysis, leading to a real Ginzburg Landau equation for the nonlinear amplitude, confirming the previous results of \citet{Umurhan:2007hs}.
+In this paper, we have constructed a weakly nonlinear analysis of the MRI using multiple-scales analysis, leading to a real Ginzburg-Landau equation for the nonlinear amplitude, confirming the previous results of \citet{Umurhan:2007hs}.
 We also confirm their $\alpha^2 \propto \Pm^{4/3} $ scaling results.
@@
 
-Equation \ref{eq:innerprod_gle} can be rewritten as Equation \ref{eq:gle}, the Ginzburg-Landau Equation, where the coefficients are
+Equation \ref{eq:innerprod_gle} can be rewritten as Equation \ref{eq:gle}, the Ginzburg-Landau equation, where the coefficients are
 

### 2016-10-04 paper/widegap_mri_paper.tex
added paper I notation.

@@
 Cartesian.
-Previous investigations into the weakly nonlinear behavior of the MRI have used this narrow gap approximation (\citei{Umurhan:2007dz}, \citei{Umurhan:2007hs}, \citei{Clark:2016}).
-In this work we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.
+Previous investigations into the weakly nonlinear behavior of the MRI have used this narrow gap approximation \citep{Umurhan:2007dz,Umurhan:2007hs,Clark:2016}.Building on our work in \citet[][hereafter Paper I]{Clark:2016}, here we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.
 
@@
 
-Just as in the weakly nonlinear analyses of \citei{Umurhan:2007hs} and \citei{Clark:2016}, we tune the system away from marginality by taking $B_0 \rightarrow B_0\left(1 - \epsilon^2\right)$, where the small parameter $\epsilon \ll 1$.
+Just as in the weakly nonlinear analyses of \citei{Umurhan:2007hs} and Paper I, we tune the system away from marginality by taking $B_0 \rightarrow B_0\left(1 - \epsilon^2\right)$, where the small parameter $\epsilon \ll 1$.
 We parameterize scale separation as $Z = \epsilon z$ and $T = \epsilon^2 t$, where $Z$ and $T$ are slowly varying spatial and temporal scales, respectively.
@@
 Figure \ref{fig:thin_wide_comparison} shows the structure of the third-order nonlinear term $N_{31}^{(A)}$ as a function of $\Pm$ for both the narrow and wide gap standard MRI.
-$N_{31}$ is the vector that determines the GLE coefficient $c$, and thus the scaling of the saturation amplitude because of the insensitivity of $b$ to $\Pm$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs}, \citei{Clark:2016} for the narrow gap equations).
+$N_{31}$ is the vector that determines the GLE coefficient $c$, and thus the scaling of the saturation amplitude because of the insensitivity of $b$ to $\Pm$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs} nad Paper I for the narrow gap equations).
 Clearly, the width of the boundary layers scales with $\Pm$ in both the wide and narrow gap MRI.

### 2016-10-05 paper/thingap_mri_paper.tex
tidied up thingap abstract

@@
 Here, we develop our theory in a thin-gap, Cartesian channel.
-Our results confirm Umurhan et al.
-2007's findings that the saturation amplitude follows a Ginzburg-Landau equation.
-We extend these results by performing a detailed force balance for the saturated $u_y$ and $B_z$ fields, demonstrating that even when diffusive effects are important, the bulk flow saturates by a combination of reducing the background shear and rearranging and strengthening the background $B_z$ field.
+Our results confirm the finding by Umurhan et al.
+(2007) that the perturbation amplitude follows a Ginzburg-Landau equation.
+We extend these results by performing a detailed force balance for the saturated azimuthal velocity and vertical magnetic field, demonstrating that even when diffusive effects are important, the bulk flow saturates via the combined processes of reducing the background shear and rearranging and strengthening the background vertical magnetic field.
+We directly simulate the Ginzburg-Landau amplitude evolution for our system and demonstrate the pattern formation our model predicts on long length and time scales.
 \end{abstract}
@@
 Those authors investigated the behavior of this MRI system as a function of $\Pm$.
-We note that we reproduce the URM07 result that the analytic saturation amplitude scales as $\alpha^2 \propto \Pm^{4/3}$.
+We note that we reproduce the URM07 result that the analytic saturation amplitude scales as $\alpha^2 \propto \Pm^{4/3}$ in a thin-gap geometry.
 
@@
 
-The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to $k$ \citep{Hoyle:2006}.
-In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane; the quadratic maximum is a consequence of the linear dispersion relation given in Appendix~\ref{app:matrices}.
+The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to the wavenumber \citep{Hoyle:2006}.
+In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane, and the quadratic maximum is a consequence of the linear dispersion relation given in Appendix~\ref{app:dispersion}.
 In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well.
+
+Here our focus is on a physical description of the saturation mechanism.
 Figure~\ref{fig:sat_comp_vel} shows the saturated mean radial profiles of $u_0 - u_{y} = -q \Omega_0 x - u_{y}$ and each term in the steady state force balance (i.e.
@@
 This demonstrates that even in a case where diffusive effects are important, the bulk of the fluid saturates by balancing shear and magnetic tension.
-As discussed at length in \citet{Vasil:2015}, when diffusive effects are not important, it is impossible to rearrange momentum without also rearranging magnetic fields.
+As discussed at length in \citet{Vasil:2015}, when diffusive effects are not important, it is impossible to rearrange momentum without also rearranging the magnetic field.
 His model demonstrates saturation without diffusive effects; our results show that outside of the boundary layers, a simultaneous rearrangement of momentum and field occurs.

### 2016-10-05 paper/thingap_mri_paper.tex
added james's edits

@@
 
-\title{The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor Couette flow}
+\title{The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor-Couette flow}
 \author{S. E. Clark\altaffilmark{1}}
@@
 The transport mechanism is likely turbulent, as molecular viscosity alone cannot account for the needed angular momentum transfer, and likely magnetic, as this turbulence is excited even in hydrodynamically stable disks (\citei{Shakura:1973wg}).
-Discovered by \citet{Chandrasekhar:1960wh} and \citet{Velikhov:1959} in global geometry, the MRI was subsequent rediscovered and applied to accretion disks by \citet{Balbus:1991vs}.
-Since then, the magnetorotational instability (MRI) remains the leading explanation for rapid angular momentum transport in astrophysical disks.
+Discovered by \citet{Chandrasekhar:1960wh} and \citet{Velikhov:1959} in a global geometry, the magnetorotational instability (MRI) was subsequently rediscovered and applied to accretion disks by \citet{Balbus:1991vs}.
+Since then, the MRI remains the leading explanation for rapid angular momentum transport in astrophysical disks.
 The instability in its simplest geometry arises when a differentially rotating disk is threaded by a vertical magnetic field.
 The presence of the magnetic field linearly destabilizes the disk gas, driving turbulence and angular momentum transport \citep[e.g.][]{2011ApJ...738...84H,2014MNRAS.438.2513P,2013MNRAS.435.2281P}.
-The MRI likely plays a role in a diverse host of astrophysical systems, including protoplanetary disks \citep[e.g.]{2015ApJ...798...84B} and black hole accretion disks \citep[e.g.][]{2013ApJ...769..156S}, as well as stellar interiors \citep{2015ApJ...799...85W}.
+The MRI likely plays a role in a diverse host of astrophysical systems, including protoplanetary disks \citep[e.g.][]{2015ApJ...798...84B} and black hole accretion disks \citep[e.g.][]{2013ApJ...769..156S}, as well as stellar interiors \citep[e.g.][]{2015ApJ...799...85W}.
 Despite its importance, many aspects of the MRI remain poorly understood.
@@
 MRI saturation has been tackled almost exclusively with simulation, with a few notable exceptions detailed below.
-In this work we investigate the weakly nonlinear behavior of the MRI.
+In this work we analytically investigate the weakly nonlinear saturation of the MRI.
 
@@
 The MRI is one such phenomenon: while there is a rich literature analytically examining the linear MRI, analytical treatments of the \textit{nonlinear} system are relatively few.
-The weakly nonlinear treatment of the MRI was pioneered by \citet{Knobloch:2005ba} and \citet[][hereafter URM07]{Umurhan:2007dz}.
-The latter authors undertook the first weakly nonlinear analysis of the MRI in a thin-gap Taylor Couette flow with strong dissipation (as is appropriate to liquid metal experiments), and found that the marginal MRI system approaches saturation in a manner analogous to that of Rayleigh-B\'enard convection.
+The weakly nonlinear treatment of the MRI was pioneered by \citet{Knobloch:2005ba} and \citet[][hereafter URM07; see also \citeyear{Umurhan:2007dz}]{Umurhan:2007hs}.
+The latter authors undertook the first weakly nonlinear analysis of the MRI in a thin-gap Taylor-Couette (TC) flow with strong dissipation (as is appropriate to liquid metal experiments), and found that the marginal MRI system approaches saturation in a manner analogous to that of Rayleigh-B\'enard convection.
 Weakly nonlinear analysis was instrumental in our understanding of Rayleigh-B\'enard convection saturation (\citei{Newell:1969}), and the similarities between convection and the thin-gap MRI are the result of important shared symmetries between the systems.
@@
 In this work we rederive the theory of URM07, and expand upon their findings.
-Our focus here is on fully characterizing the narrow gap MRI system, both for independent theoretical interest and to have a robust comparison point for extensions of this theory into more complicated geometries.
-In a companion paper we derive for the first time the weakly nonlinear theory of the standard and helical MRI in a wide gap Taylor Couette flow \citep[][hereafter Paper II]{Clark:2016}.
+Our focus here is on fully characterizing the thin-gap MRI system, both for independent theoretical interest and to have a robust comparison point for extensions of this theory into more complicated geometries.
+In a companion paper we derive for the first time the weakly nonlinear theory of the standard and helical MRI in a global, cylindrical TC flow \citep[][hereafter Paper II]{Clark:2016}.
 The thin- and wide-gap treatments complement one another theoretically, and both are important regimes for comparison with simulation.
 
-This work examines Taylor-Couette flow in the thin-gap regime, an idealization in which the radial extent of the channel is very small compared to its distance from the center of rotation, i.e.
+This work examines TC flow in the thin-gap regime, an idealization in which the radial extent of the channel is very small compared to its distance from the center of rotation, i.e.
 $(r_2 - r_1) \ll \frac{1}{2} (r_1 + r_2)$ where $r_1$ and $r_2$ are the radii of the inner and outer flow boundaries, respectively.
@@
 \citei{Vasil:2015} examines the weakly nonlinear MRI in a thin-gap regime in a minimal model, finding deep mathematical similarities between the MRI system and the elastodynamic instability of a buckling beam.
-We discuss his results and their relation to ours in section~\ref{sec:discussion}.
+We discuss these results and their relation to ours in section~\ref{sec:discussion}.
 Several authors have investigated the behavior of the MRI when the boundary conditions are shear periodic, and so the MRI has no mechanism by which to modify the background shear flow profile.
@@
 Analytical investigation of this case reveals that MRI saturation can be caused by parasitic Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime (\citei{Pessah:2010ic}).
-The theory of MRI channel mode parasites is robust (see also \citei{Pessah:2009gm}, \citei{Latter:2010iz}, \citei{Rembiasz:2016}), but their importance may be overestimated by the local approximation \citei{Latter:2015}, and not germane to global analyses like the one presented here.
-\citei{Latter:2015} gives a detailed analysis of the relation between local and global linear MRI modes.
+The theory of MRI channel mode parasites is robust (see also \citei{Pessah:2009gm}, \citei{Latter:2010iz}, \citei{Rembiasz:2016}), but their importance may be overestimated by the local approximation \citep{Latter:2015}, and not germane to global analyses like the one presented here.
+\citet{Latter:2015} gives a detailed analysis of the relation between local and global linear MRI modes.
 
@@
 %in units where the permittivity of free space is unity. 
-Equations \ref{eq:momentum} and \ref{eq:induction} are subject to the magnetic solenoid and incompressibility constraints,
+Equations \ref{eq:momentum} and \ref{eq:induction} are subject to the incompressibility and magnetic solenoid constraints,
+
+\beq\label{eq:incompressibility_constraint}
+\nabla \cdot \mathbf{u} = 0
+\eeq
 
 \beq\label{eq:solenoid_constraint}
-\nabla \cdot \mathbf{B} = 0
-\eeq
-
-\beq\label{eq:incompressibility_constraint}
-\nabla \cdot \mathbf{u} = 0.
+\nabla \cdot \mathbf{B} = 0.
 \eeq
@@
 We axisymmetrically perturb all three vector components of each of the fluid quantities.
-We nondimensionalize the equations, with lengths nondimensionalized by $L$, velocities by $\Omega_0 L$, magnetic fields by $B_0$, and pressure by $\Omega_0^2 L^2 \rho_0$, where $L$ is the channel width, $\Omega_0$ is the rotation rate at the center of the channel, and $\rho_0$ is the constant pressure in the base state (see Figure \ref{fig:setup}).
-We define the Reynolds number, $\reye \equiv {\Omega_0 L^2}/{\nu}$, magnetic Reynolds number, $\reym \equiv {\Omega_0 L^2}/{\eta}$, and a plasma beta parameter, $\beta \equiv {\Omega_0^2r_0^2}/{v_A^2}$, where the Alfv\'en sped $v_A$ is $v_A^2 = {B_0^2}/{\rho_0}$.
+We nondimensionalize the equations, with lengths nondimensionalized by $L$, time by $\Omega_0$, velocities by $\Omega_0 L$, magnetic fields by $B_0$, and pressure by $\Omega_0^2 L^2 \rho_0$, where $L$ is the channel width, $\Omega_0$ is the rotation rate at the center of the channel, and $\rho_0$ is the constant pressure in the base state (see Figure \ref{fig:setup}).
+We define the Reynolds number, $\reye \equiv {\Omega_0 L^2}/{\nu}$, magnetic Reynolds number, $\reym \equiv {\Omega_0 L^2}/{\eta}$, and a plasma beta parameter, $\beta \equiv {\Omega_0^2r_0^2}/{v_A^2}$, where the Alfv\'en speed $v_A$ is $v_A^2 = {B_0^2}/{\rho_0}$.
 The fluid symbols $\mathbf{u}$, $\mathbf{B}$, etc.
@@
 
-We define the flux function $A$ and streamfunction $\Psi$, where $A$ is the familiar two-dimensional vector potential.
-$A$ and $\Psi$ are scalar fields.
-The curl of $A$ and the curl of $\Psi$ are defined as the magnetic field and velocity, respectively, and so $A$ and $\Psi$ automatically satisfy our constraints (Equations \ref{eq:solenoid_constraint} and \ref{eq:incompressibility_constraint}).
+We define the streamfunction $\Psi$ and flux function $A$, where $A$ is the familiar two-dimensional vector potential.
+$\Psi$ and $A$ are scalar fields.
+The curl of $\Psi$ and the curl of $A$ are defined as the velocity and magnetic field perturbation, respectively, and so $\Psi$ and $A$ automatically satisfy our constraints (Equations \ref{eq:incompressibility_constraint} and \ref{eq:solenoid_constraint}).
 
-$A$ and $\Psi$ are thus related to the magnetic field and velocity perturbations, respectively, as
+$\Psi$ and $A$ are thus related to the velocity and magnetic field perturbations, respectively, as
+
+\beq
+\mathbf{u} \, = \, \left[\begin{matrix}
+\partial_z \Psi \\
+u_{y} \\
+-\partial_x \Psi \\
+\end{matrix}\right],\eeq
 
@@
 -\partial_xA \\
-\end{matrix}\right],\eeq
-
-\beq
-\mathbf{u} \, = \, \left[\begin{matrix}
-\partial_z \Psi \\
-u_{y} \\
--\partial_x \Psi \\
 \end{matrix}\right].\eeq
@@
 \eeq  
-and $q = - d \ln \Omega/ \ln R = 3/2$ is the dimensionless shear parameter such that the background velocity profile is $v_0 = q \Omega_0 x$.
+and $q \equiv - d \ln \Omega/ \ln R = 3/2$ is the dimensionless shear parameter defining a rotation profile $\Omega(r) = \Omega_0 (r/r_0)^{-q}$, such that the background velocity profile is $u_0 = -q \Omega_0 x$.
 
-Note that working in terms of the flux function raises the order of the first momentum equation.
+%Note that working in terms of the streamfunction raises the order of the first momentum equation.
 
 The weakly nonlinear regime is where the MRI system is nonlinearly unstable to only the most unstable mode of the linear solution.
-We find the marginal state, where the most unstable linear MRI mode neither grows nor decays, for a given set of dimensionless parameters.
+We find the marginal state, where the most unstable linear MRI mode neither grows nor decays, for a set of dimensionless parameters, and then destabilize the system.
 
@@
 
-$B_0$ appears in Equations \ref{eqset1} - \ref{eqset4} because it is the nondimensionalized background field strength.
-Because we nondimensionalize $B$ by the magnitude of the background field strength, $B_0 \equiv 1$.
-In order to study our system in the weakly nonlinear regime, we tune the background magnetic field down away from stability (recall that stronger vertical fields stabilize the MRI).
+%$B_0$ appears in Equations \ref{eqset1} - \ref{eqset4} because it is the nondimensionalized background field strength. 
+Because we nondimensionalize $B$ by the magnitude of the background field strength, $B_0 \equiv 1$ in Equations \ref{eqset1} - \ref{eqset4}.
+To excite the weakly nonlinear MRI, we tune the background magnetic field down away from stability (recall that stronger vertical fields stabilize the MRI).
 We do so by substituting $B = B_0\left(1 - \epsilon^2\right)$.
 The degree of departure from the marginal state is measured by the small parameter $\epsilon$.
-An $\mathcal{O}\left(\epsilon^2\right)$ weakening of the background magnetic field destabilizes a band of wave modes with width of $\mathcal{O}\left(\epsilon\right)$, which interact nonlinearly.
+An $\mathcal{O}\left(\epsilon^2\right)$ weakening of the background magnetic field destabilizes a finite band of wave modes with a width of $\mathcal{O}\left(\epsilon\right)$, which interact nonlinearly.
 
-The destabilizing substitution is made, and Equations \ref{eqset1} - \ref{eqset4} are rewritten such that the fluid variables are contained in a state vector $\mathbf{V} = \left[\Psi, u_y, A, B_y\right]^\mathrm{T}$.
+The destabilizing substitution is made, and Equations \ref{eqset1} - \ref{eqset4} are rewritten such that the fluid variables are contained in a state vector
+
+\beq
+\mathbf{V} = \left[\Psi, u_y, A, B_y\right]^\mathrm{T}.
+\eeq
+
 This yields the system of equations
@@
 We conduct a formal multiple scales analysis of this system.
-Our perturbations are characterized in terms of fast and slow-moving variables, in order to simultaneously track their evolution on two scales.
+Our perturbations are characterized in terms of fast- and slow-moving variables, that we treat as independent in order to simultaneously track the evolution of the system on two scales.
 The relative scalings of the fast and slow variables are chosen such that each of the temporal and spatial eigenvalues appear at the same lowest order in the linear dispersion relation (Appendix \ref{app:dispersion}).
@@
 %See Figure \ref{figscalings}. 
-Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic Taylor-Couette flow.
+Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic TC flow.
 Our $x$ dimension, the direction of angular momentum transport, is analogous to the direction of temperature transport in the convection problem.
 In analogy to these problems, we posit slow variation in both $Z$ and $T$.
-Each operator is expanded to reflect these scalings -- for instance, $\partial_z$ becomes $\partial_z + \epsilon\partial_Z$.
-The fast and slow variables are treated as independent in the perturbation series analysis that follows.
+Each operator in Equations \ref{eqset1} - \ref{eqset4} is expanded to reflect these scalings -- for instance, $\partial_z$ becomes $\partial_z + \epsilon\partial_Z$.
 
@@
 
-And our perturbed system is then expressed order by order as
+Our perturbed system is then expressed order by order as
 
@@
 \label{eq:gle}
-\partial_T \alpha = b \alpha + d \partial_Z^2 \alpha - c \alpha \left|\alpha^2\right|,
+\partial_T \alpha = b \alpha + h \partial_Z^2 \alpha - c \alpha \left|\alpha^2\right|,
 \eeq
@@
 Those authors investigated the behavior of this MRI system as a function of $\Pm$.
-We note that we reproduce the URM07 result that the analytic saturation amplitude scales as $\alpha^2 \propto \Pm^{4/3}$ in a thin-gap geometry.
+By analyzing the system over several orders of magnitude in $\Pm$, we reproduce the URM07 result that the analytic saturation amplitude scales as $\alpha_{saturation}^2 \propto \Pm^{4/3}$ in a thin-gap geometry when $\Pm \ll 1$.
+We also find, as in URM07, that the neutral mode amplitude $\beta$ is governed by a simple diffusion equation that decouples from the Ginzburg-Landau equation.
 
@@
 Figure~\ref{fig:sat_comp_vel} shows the saturated mean radial profiles of $u_0 - u_{y} = -q \Omega_0 x - u_{y}$ and each term in the steady state force balance (i.e.
-equation~\ref{eqset2} with $\partial_t u_{y} = 0$).
+Equation~\ref{eqset2} with $\partial_t u_{y} = 0$).
 In the bulk of the fluid away from the boundary layers, the saturated state shows reduced shear, with little diffusive contribution.
@@
 As discussed at length in \citet{Vasil:2015}, when diffusive effects are not important, it is impossible to rearrange momentum without also rearranging the magnetic field.
-His model demonstrates saturation without diffusive effects; our results show that outside of the boundary layers, a simultaneous rearrangement of momentum and field occurs.
+The \citet{Vasil:2015} model demonstrates saturation without diffusive effects; our results show that outside of the boundary layers, a simultaneous rearrangement of momentum and field occurs.
 In the boundary layers, the non-linear advection balances viscous dissipation.
@@
 In the high $\reye$ and  $\reym$  limit, \citet{Vasil:2015} derives an amplitude equation considerably different than the one found here.
-By averaging in the $z$ direction, he computes a mean-field equation with striking similarity to the buckling of an elastic beam under load.
-The most salient feature of his equation is its \emph{non-local} character.
-Unlike the present work, which focuses on Keplerian rotation profiles with $q = 3/2$ with a critical $B_0$, \citet{Vasil:2015} focuses on a fixed $B$ and a weakly destabilized shear profile.
-However, these differences are minor; the destabilizing parameter $\epsilon$ enters the analysis in the same quadratic proportion.
+By averaging in the $z$ direction, the author computes a mean-field equation with striking similarity to the buckling of an elastic beam under load.
+The most salient feature of this equation is its \emph{non-local} character.
+Unlike the present work, which focuses on Keplerian rotation profiles with $q = 3/2$ with a critical background magnetic field strength, \citet{Vasil:2015} focuses on a fixed field strength and a weakly destabilized shear profile.
+These differences are minor, however: the destabilizing parameter $\epsilon$ enters the analysis in the same quadratic proportion.
 Whether and how \citet{Vasil:2015}'s amplitude equation is equivalent to our own in the limit of dynamically important resistive and viscous effects is beyond the scope of this work.
-Nevertheless, he identifies the non-linear term responsible for saturation as consisting of flux and field transport and notes these are the only mechanisms that can produce saturation.
-Our results also demonstrate a combination of flux and field transport.
-This suggests that despite our formulation displaying different saturation dynamics (Ginzburg-Landau in our case; a network of coupled Duffing oscillators in his), there may indeed be an underlying unification.
+Nevertheless, the author identifies the non-linear term responsible for saturation as consisting of flux and field transport and notes these are the only mechanisms able to produce saturation.
+Our results likewise demonstrate a combination of flux and field transport in the comparable region of our domain.
+This suggests that despite our formulation displaying different saturation dynamics (Ginzburg-Landau in our case; a network of coupled Duffing oscillators in \citeauthor{Vasil:2015} \citeyear{Vasil:2015}), there may indeed be an underlying unification.
 
@@
 We evolve the system for $100 \Omega_0^{-1}$ in timesteps of $0.02 \Omega_0^{-1}$.
-Results are shown in Figures \ref{fig:IVP_2lambdacrit} and \ref{fig:IVP_10lambdacrit}.
+Results are shown in Figures \ref{fig:IVP_2lambdacrit} and \ref{fig:IVP_10lambdacrit}, where the amplitude and phase structure over the vertical domain is plotted for every $20$ timesteps.
 The system quickly organizes itself into rolls in $Z$ bounded by the analytic saturation amplitude $\alpha_s = \sqrt{b/c}$.
@@
 
-In this paper, we have constructed a weakly nonlinear analysis of the MRI using multiple-scales analysis, leading to a real Ginzburg-Landau equation for the nonlinear amplitude, confirming the previous results of \citet{Umurhan:2007hs}.
-We also confirm their $\alpha^2 \propto \Pm^{4/3} $ scaling results.
+In this paper we construct a weakly nonlinear analysis of the MRI using multiple-scales analysis, leading to a real Ginzburg-Landau equation for the nonlinear amplitude, confirming the previous results of \citet{Umurhan:2007hs}.
+We also confirm their results for the scaling of the analytic saturation amplitude with $\Pm$.
 We extend their results by constructing a detailed force and inductive balance for the saturated $u_y$ and $B_z$ components.
@@
 We demonstrate that complex patterns can organize the flow on long length scales $Z$, though the maximum magnitude of the amplitude $\alpha$ is well predicted by the steady state solution.
-In paper II, we make use of the techniques developed here to extend the weakly nonlinear analysis of the MRI to a full cylindrical geometry appropriate for a Taylor-Couette experiment.
+In Paper II, we make use of the techniques developed here to extend the weakly nonlinear analysis of the MRI to a full cylindrical geometry appropriate for a Taylor-Couette experiment.
 
@@
 %************* APPENDICES ************************%
-
 \clearpage
@@
 \eeq
+
+As noted by URM07, a second amplitude equation for a spatially constant azimuthal magnetic field mode arises from the terms in the $\mathcal{O}(\epsilon^3)$ equation which contain no $z$ dependence.
+This is a diffusion equation, so the neutral mode simply decays away.
 

### 2016-10-05 paper/widegap_mri_paper.tex
added james's edits

@@
 
-The magnetorotational instability (MRI) is widely belived to drive angular momentum transport in astrophysical disks.
-The MRI is a local instability excited by weak magnetic fields in differentially rotating fluids, and since first applied to an astrophysical context \citep{Balbus:1991vs} it has been widely invoked to explain accretion in protoplanetary disks \citep{Armitage:2010} and disks around black holes \citep{Blaes:2014}, as well as jet and wind launching (\citei{Lesur:2013}), anisotropic turbulence \citep{Murphy:2015tn}, and dynamo generation \citep{Brandenburg:1995, Vishniac:2009il}.
+The magnetorotational instability (MRI) is believed to drive angular momentum transport in astrophysical disks.
+The MRI is a local instability excited by weak magnetic fields in differentially rotating fluids, and since first applied to an astrophysical context \citep{Balbus:1991vs} it has been invoked to explain accretion in protoplanetary disks \citep{Armitage:2010} and disks around black holes \citep{Blaes:2014}, as well as jet and wind launching (\citei{Lesur:2013}), anisotropic turbulence \citep{Murphy:2015tn}, and dynamo generation \citep{Brandenburg:1995, Vishniac:2009il}.
 
@@
 The diversity of astrophysical systems which may be MRI unstable yields an enormous parameter space to be explored.
-In protoplanetary disks, for example, the behavior and evolution of the MRI--and even its very existence--may change drastically depending on the properties of the magnetic field, the disk composition, disk geometry, and so forth.
+In protoplanetary disks, for example, the behavior and evolution of the MRI -- and even its very existence -- may change drastically depending on the properties of the magnetic field, the disk composition, disk geometry, and so forth.
 Multiphysics numerical simulations of such systems are currently an area of intense focus, enabling the study of nonideal MHD effects, disk stratification, nonequilibrium chemistry, and other complex physics that does not lend itself easily to analytic study \citep[e.g.][among many others]{Fleming:2003fs,Bai:2011cm, Flock:2013,Suzuki:2014vh}.
@@
 The eigenvalue problem was solved for the widegap parameters listed in Table \ref{table:parameters}.
-Overlaid contours show growth rates at [ -8E-4, -1.3E-4, 1.3E-4, 8E-4, 1.5E-3], where dashed contours represent negative values.
+Overlaid contours show growth rates at [-8E-4, -1.3E-4, 1.3E-4, 8E-4, 1.5E-3], where dashed contours represent negative values.
 The gray dotted line shows the interpolated marginal stability curve.
@@
 Thus through judicious choice of cylinder rotation rates, one can set $q(r_0) = 3/2$, for quasi-Keplerian flow.
-Note that the narrow gap approximation imposes a linear shear (constant $q$), and thus the interaction of fluid perturbations with the base velocity profile differs significantly from the case considered here.
+Note that the narrow gap approximation imposes a linear shear (constant $q$), and so the interaction of fluid perturbations with the base velocity profile differs significantly from the case considered here.
 Our base velocity is
@@
 As detailed in \S\ref{sec:wnl_analysis}, we expand these equations in a perturbation series and solve them order by order using a pseudospectral code.
-The code solves partial differential equations of the form $\mathrm{M} \partial_t \mathbf{V} + \mathrm{L} \mathbf{V} = \mathbf{F}$, where $\mathrm{M}$ and $\mathrm{L}$ are matrices and $\mathbf{F}$ is a vector containing any inhomogenous terms.
-The nonlinear terms in our perturbation analysis become inhomogenous term inputs to the solver.
+The code solves partial differential equations of the form $\mathrm{M} \partial_t \mathbf{V} + \mathrm{L} \mathbf{V} = \mathbf{F}$, where $\mathrm{M}$ and $\mathrm{L}$ are matrices and $\mathbf{F}$ is a vector containing any nonhomogenous terms.
+The nonlinear terms in our perturbation analysis become nonhomogenous term inputs to the solver.
 
@@
 \includegraphics[width=\columnwidth]{../figures/widegap_totalstress_intstress.png}
-\caption{Total stress $\mathrm{T}(r, z) = u_r u_\phi - \frac{2}{\beta} B_r B_\phi$ , i.e.
+\caption{Total stress $\mathrm{T}(r, z) = u_r u_\phi - \frac{2}{\beta} B_r B_\phi$, i.e.
 the sum of the Reynolds and Maxwell stresses for the fiducial standard MRI parameters (top panel).
@@
 Figure \ref{fig:coefficients} shows the critical parameters $k_c$ and $\reym$ as a function of $\Pm$, as well as the GLE linear coefficient $b$ and the diffusion coefficient $d$.
-From Equation \ref{eq:gle} it is readily apparent that the asymptotic saturation amplitude is $\alpha_{s} = \pm \sqrt{b/c}$, and we plot the dependence of $\alpha_{s}$ on $\Pm$ in the bottom panel of \ref{fig:coefficients}.
+From Equation \ref{eq:gle} it is apparent that the asymptotic saturation amplitude is $\alpha_{s} = \pm \sqrt{b/c}$, and we plot the dependence of $\alpha_{s}$ on $\Pm$ in the bottom panel of Figure \ref{fig:coefficients}.
 Note that because $\reym$ is essentially constant as a function of $\Pm$, the saturation amplitude is equivalently sensitive to $\reye^{-1}$.
@@
 Figure \ref{fig:thin_wide_comparison} shows the structure of the third-order nonlinear term $N_{31}^{(A)}$ as a function of $\Pm$ for both the narrow and wide gap standard MRI.
-$N_{31}$ is the vector that determines the GLE coefficient $c$, and thus the scaling of the saturation amplitude because of the insensitivity of $b$ to $\Pm$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs} nad Paper I for the narrow gap equations).
+$N_{31}$ is the vector that determines the GLE coefficient $c$, and thus the scaling of the saturation amplitude because of the insensitivity of $b$ to $\Pm$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs} and Paper I for the narrow gap equations).
 Clearly, the width of the boundary layers scales with $\Pm$ in both the wide and narrow gap MRI.
@@
 The complex GLE may be unstable to traveling wave instabilities such as the Benjamin-Feir instability, a generalization of the Ekhaus instability.
-The complex GLE can also admit spatiotemporal chaos, and various classes of coherent structures (\citei{Aranson:2002}).Although a detailed description of the phase dynamics in the helical MRI is beyond the scope of this work, we note that such long-wavelength, long-timescale behavior may be observed in liquid metal helical MRI experiments.
+The complex GLE can also admit spatiotemporal chaos, and various classes of coherent structures (\citei{Aranson:2002}).
+Although a detailed description of the phase dynamics in the helical MRI is beyond the scope of this work, we note that such long-wavelength, long-timescale behavior may be observed in liquid metal helical MRI experiments.
 
@@
 Understanding the connection between local and global MRI modes is crucial for interpreting simulation results across domain geometries.
-(The relationship between local and global \textit{linear} MRI modes is investigated in \citet{Latter:2015}.) Phenomena such as saturation and the development of turbulence depend critically on the nature of the underlying MRI modes.
+(The relationship between local and global \textit{linear} MRI modes is investigated in \citeauthor{Latter:2015} \citeyear{Latter:2015}.) Phenomena such as saturation and the development of turbulence depend critically on the nature of the underlying MRI modes.
 The formalism presented here describes analytically the weakly nonlinear behavior of global MRI modes.
@@
 The Fredholm alternative states that if we consider a system of equations $\mathcal{L} \mathbf{V} = \mathbf{b}$ and its adjoint homogenous system $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$, only one of two conditions holds.
-Either there exists one and only one solution to the inhomogenous system, or the homogenous adjoint equation has a nontrivial solution.
+Either there exists one and only one solution to the nonhomogenous system, or the homogenous adjoint equation has a nontrivial solution.
 The relevant corollary arises as a consequence of the second condition: if $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$ has a nontrivial solution, then $\mathcal{L} \mathbf{V} = \mathbf{b}$ has a solution if and only if $\langle \mathbf{V}^\dagger | b \rangle = 0$.

### 2016-10-05 paper/thingap_submit/thingap_mri_paper_ApJ_submit.tex
thingap submitted materials incl all plots as eps or pdf

--- 2016-10-05 paper/thingap_mri_paper.tex
@@
 The magnetorotational instability (MRI) is a fundamental process of accretion disk physics, but its saturation mechanism remains poorly understood despite considerable theoretical and computational effort.
-We present a multiple-scales analysis of the non-ideal MRI in the weakly nonlinear regime -- that is, when the the most unstable MRI mode has a growth rate asymptotically approaching zero from above.
+We present a multiple scales analysis of the non-ideal MRI in the weakly nonlinear regime -- that is, when the most unstable MRI mode has a growth rate asymptotically approaching zero from above.
 Here, we develop our theory in a thin-gap, Cartesian channel.
@@
 We find the marginal state, where the most unstable linear MRI mode neither grows nor decays, for a set of dimensionless parameters, and then destabilize the system.
+We examine the system for fiducial parameters comparable to URM07, namely $\Pm = 1.0 \times 10^{-3}$, $\beta = 25$, $q = 1.5$.
+The system is marginal for a critical wavenumber $k_c = 0.75$ and a critical magnetic Reynolds number $\reym_c = 4.9$.
 
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_velocity_Pm_1E-3.png}
+\includegraphics[width=\textwidth]{thingap_streamfuncs_velocity_Pm_1E-3.png}
 \caption{First order (left), second order (center), and total (right) velocity perturbations.
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/thingap_streamfuncs_Bfield_Pm_1E-3.png}
+\includegraphics[width=\textwidth]{thingap_streamfuncs_Bfield_Pm_1E-3.png}
 \caption{As in Figure \ref{fig:allorders_velocity} but for the magnetic field.
@@
 By analyzing the system over several orders of magnitude in $\Pm$, we reproduce the URM07 result that the analytic saturation amplitude scales as $\alpha_{saturation}^2 \propto \Pm^{4/3}$ in a thin-gap geometry when $\Pm \ll 1$.
-We also find, as in URM07, that the neutral mode amplitude $\beta$ is governed by a simple diffusion equation that decouples from the Ginzburg-Landau equation.
+We also find, as in URM07, that the neutral mode amplitude $\beta(T, Z)$ is governed by a simple diffusion equation that decouples from the Ginzburg-Landau equation.
 
@@
 \centering
-\includegraphics[width=\columnwidth]{../figures/thingap_saturation_mechanism_plots_velocity_slice60.png}
-\caption{Top panel: Mean radial profile for $u_y^{saturated} - u_y^0$, the difference between the saturated azimuthal velocity profile and the initial azimuthal velocity profile (TC flow).
+\includegraphics[width=\columnwidth]{thingap_saturation_mechanism_plots_velocity_slice60.pdf}
+\caption{Top panel: radial profile for $u_y^{saturated} - u_y^0$, the difference between the saturated azimuthal velocity profile and the initial azimuthal velocity profile (TC flow).
 Bottom panel: each term in the steady state force balance (Equation \ref{eqset2} with $\partial_t u_{y} = 0$).
@@
 \centering
-\includegraphics[width=\columnwidth]{../figures/thingap_saturation_mechanism_plots_Bfield_slice60.png}
-\caption{Top panel: Mean radial profile for $B_z^{saturated}$, the saturated vertical magnetic field (black line).
+\includegraphics[width=\columnwidth]{thingap_saturation_mechanism_plots_Bfield_slice60.pdf}
+\caption{Top panel: radial profile for $B_z^{saturated}$, the saturated vertical magnetic field (black line).
 $B_z^0 = 1$ is the constant background magnetic field (gray line).
@@
 Here our focus is on a physical description of the saturation mechanism.
-Figure~\ref{fig:sat_comp_vel} shows the saturated mean radial profiles of $u_0 - u_{y} = -q \Omega_0 x - u_{y}$ and each term in the steady state force balance (i.e.
+Figure~\ref{fig:sat_comp_vel} shows saturated radial profiles of $u_0 - u_{y} = -q \Omega_0 x - u_{y}$ and each term in the steady state force balance (i.e.
 Equation~\ref{eqset2} with $\partial_t u_{y} = 0$).
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_2.png}
+\includegraphics[width=\textwidth]{IVP_plot_thingap_AA_phase_2lambdacrit_2.pdf}
 \caption{Evolution of the Ginzburg-Landau amplitude equation (Equation \ref{eq:gle}) on a Fourier $Z$ domain of length $2 \lambda_{crit}$, where $\lambda_{crit} = 2\pi/k_c$ is the critical wavelength of the system.
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/IVP_plot_thingap_AA_phase_2lambdacrit_10.png}%{allorders_velocity.eps}
+\includegraphics[width=\textwidth]{IVP_plot_thingap_AA_phase_2lambdacrit_10.pdf}%{allorders_velocity.eps}
 \caption{As in Figure \ref{fig:IVP_2lambdacrit} but for a $Z$ domain of length $10 \lambda_{crit}$.}\label{fig:IVP_10lambdacrit}
@@
 
-In this paper we construct a weakly nonlinear analysis of the MRI using multiple-scales analysis, leading to a real Ginzburg-Landau equation for the nonlinear amplitude, confirming the previous results of \citet{Umurhan:2007hs}.
+In this paper we construct a weakly nonlinear analysis of the MRI using multiple scales analysis, leading to a real Ginzburg-Landau equation for the nonlinear amplitude, confirming the previous results of \citet{Umurhan:2007hs}.
 We also confirm their results for the scaling of the analytic saturation amplitude with $\Pm$.

### 2016-10-05 python/plot_thingap_IVP.py

@@ -10,6 +10,8 @@ rc('text', usetex=True)
 rc('figure', dpi=120)
 
 
+lambdacrit = 2
+
 root = "/Users/susanclark/weakly_nonlinear_mri/data/"
 #fn = "Pm_0.00152830673266_thingap_GLE_IVP_gridnum128_init_0.hdf5"
 #fn = "Pm_0.00152830673266_thingap_GLE_IVP_gridnum128_init_1E-15.hdf5"
@@ -17,7 +19,10 @@ root = "/Users/susanclark/weakly_nonlinear_mri/data/"
 #fn = "Pm_0.00316227766017_thingap_GLE_IVP_gridnum128_init_0.hdf5"
 #fn = "Pm_0.000248162892284_thingap_GLE_IVP_gridnum128_init_0_nomean.hdf5"
 #fn = "Pm_0.000248162892284_thingap_GLE_IVP_gridnum128_init_0_noiselvl0.001.hdf5"
-fn = "IVP_thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_128_init_0_noiselvl0.001.hdf5"
+if lambdacrit == 10:
+    fn = "IVP_thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_128_init_0_noiselvl0.001.hdf5"
+elif lambdacrit == 2:
+    fn = "IVP_thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_128_init_0_noiselvl0.001_2.hdf5"
 
 
 #fn = "IVP_widegap_amplitude_parameters_Q_2.33_Rm_0.0015_Pm_1.00e-06_Omega1_313.55_Omega2_84.66_beta_0.02_xi_4.00.h5"

### 2016-10-05 python/plot_thingap_streamfuncs.py

@@ -14,7 +14,7 @@ import streamplot_uneven as su
 
 file_root = "/Users/susanclark/weakly_nonlinear_mri/data/"
 fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_128.h5"
-fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8738_Pm_1.00e-04_q_1.5_beta_25.00_gridnum_128.h5"
+#fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8738_Pm_1.00e-04_q_1.5_beta_25.00_gridnum_128.h5"
 obj = h5py.File(file_root + fn, "r")
 
 Q = obj.attrs['Q']
@@ -25,7 +25,7 @@ eps = 0.5
 
 # saturation amplitude -- for now just constant, coefficient-determined
 satamp = np.sqrt(obj.attrs['b']/obj.attrs['c']) #1
-satamp = 1
+#satamp = 1
 
 # create z grid
 nz = obj.attrs['gridnum']
@@ -82,7 +82,7 @@ def plot_o1(ax, obj, type="Bfield", labels=True, oplot=True, background=False, c
     if type is "velocity":
         cbarmax = np.max(V1_u.real)
         cbarmin = np.min(V1_u.real)
-        info = ax.pcolormesh(plotx, z, V1_u.real, cmap=cmap, vmin = cbarmin, vmax = cbarmax)
+        info = ax.pcolormesh(plotx, z, V1_u.real, cmap=cmap, vmin = cbarmin, vmax = cbarmax,linewidth=0,rasterized=True)
         divider = make_axes_locatable(ax)
         cax = divider.append_axes("right", size="5%", pad=0.05)  
         #cax = fig.add_axes([cbarleft, cbarvert, cbarwidth, cbarheight])  
@@ -114,7 +114,7 @@ def plot_o1(ax, obj, type="Bfield", labels=True, oplot=True, background=False, c
 
         cbarmax = np.max(V1_B.real)
         cbarmin = np.min(V1_B.real)
-        info = ax.pcolormesh(plotx, z, V1_B.real, cmap=cmap, vmin = cbarmin, vmax = cbarmax)
+        info = ax.pcolormesh(plotx, z, V1_B.real, cmap=cmap, vmin = cbarmin, vmax = cbarmax,linewidth=0,rasterized=True)
         divider = make_axes_locatable(ax)
[…]

### 2016-10-05 paper/thingap_submit/thingap_mri_paper_ApJ_submit.tex
widegap submitted version

@@
 where $P$ is the gas pressure, $\nu$ is the kinematic viscosity, $\eta$ is the microscopic diffusivity, $\nabla\Phi$ is the gravitational force per unit mass, and the current density is $\mathbf{J} = \nabla\times\mathbf{B}$.
-%in units where the permittivity of free space is unity. 
 Equations \ref{eq:momentum} and \ref{eq:induction} are subject to the incompressibility and magnetic solenoid constraints,
@@
 
-%and $\mathbf{\Psi}$ is defined analogously.
-
 Our final equation set is
@@
 
-%Note that working in terms of the streamfunction raises the order of the first momentum equation.
-
 The weakly nonlinear regime is where the MRI system is nonlinearly unstable to only the most unstable mode of the linear solution.
@@
 \centering
-\includegraphics[width=\textwidth]{thingap_streamfuncs_velocity_Pm_1E-3.png}
+\includegraphics[width=\textwidth]{thingap_streamfuncs_velocity_Pm_1E-3}
 \caption{First order (left), second order (center), and total (right) velocity perturbations.
@@
 We use a constant amplitude $\alpha = \alpha_{saturation}$ and a small parameter $\epsilon = 0.5$.}\label{fig:allorders_velocity}
-%\end{figure*}
 \vspace*{\floatsep}
-%\begin{figure*}[h!]
 \centering
-\includegraphics[width=\textwidth]{thingap_streamfuncs_Bfield_Pm_1E-3.png}
+\includegraphics[width=\textwidth]{thingap_streamfuncs_Bfield_Pm_1E-3}
 \caption{As in Figure \ref{fig:allorders_velocity} but for the magnetic field.
@@
 
-%$B_0$ appears in Equations \ref{eqset1} - \ref{eqset4} because it is the nondimensionalized background field strength. 
 Because we nondimensionalize $B$ by the magnitude of the background field strength, $B_0 \equiv 1$ in Equations \ref{eqset1} - \ref{eqset4}.
@@
 
-%See Figure \ref{figscalings}. 
 Note that these are the same scalings as apply to Rayleigh-B\'enard convection and hydrodynamic TC flow.
@@
 Note that the phase angle is wrapped on a $2\pi$ domain, such that $\pi$ = $-\pi$, as indicated by the circular colorbar.}\label{fig:IVP_2lambdacrit}
-%\end{figure*}
 
-%\begin{figure*}[h!]
 \centering
-\includegraphics[width=\textwidth]{IVP_plot_thingap_AA_phase_2lambdacrit_10.pdf}%{allorders_velocity.eps}
+\includegraphics[width=\textwidth]{IVP_plot_thingap_AA_phase_2lambdacrit_10.pdf}
 \caption{As in Figure \ref{fig:IVP_2lambdacrit} but for a $Z$ domain of length $10 \lambda_{crit}$.}\label{fig:IVP_10lambdacrit}
@@
 While still bounded by $\alpha_s$, the saturation amplitude exhibits a nonlinear phase geometry due to the nonlinear interaction of modes in $Z$.
-%The saturation amplitude, while not required to be trivially constant, has the same maximum magnitude as the steady-sate system.
 
@@
 
-%************* APPENDICES ************************%
 \clearpage
@@
 \end{document}
-
-%  LocalWords:  Magnetorotational Couette Oishi Lewiston Umurhan et
-%  LocalWords:  magnetorotational al Ginzburg accrete Shakura wg wh
-%  LocalWords:  hydrodynamically Velikhov Balbus protoplanetary ba ul
-%  LocalWords:  perturbative analytical multiscale Orszag Knobloch ic
-%  LocalWords:  URM enard Newell Vasil elastodynamic Pessah iz Alfv
-%  LocalWords:  Rembiasz axisymmetric Chebyshev diffusivity zA xA aa
-%  LocalWords:  incompressibility axisymmetrically nondimensionalize
-%  LocalWords:  nondimensionalized nondimensional streamfunction ey
-%  LocalWords:  Jacobian Colorbar azimuthal ansatz marginality Runge
-%  LocalWords:  wavenumber pseudospectral Dedalus solvability Kutta
-%  LocalWords:  axisymmetry advection Ebrahimi Keplerian timestep Im
-%  LocalWords:  timesteps crit arctan colorbar hs DGE LWS Mordecai Xu
-%  LocalWords:  Krommes Geoff Zweibel natexlab Bai Hawley Prager ApJ
-%  LocalWords:  Schnack Guan Krolik Rebecca Julien Fromang Faure JFM
-%  LocalWords:  MNRAS Gressel Parkin Bicknell Obergaulinger Cerd Dur
-%  LocalWords:  uller Aloy Schnittman Sunyaev Regev Menou RSPSA Sov
-%  LocalWords:  JETP Kagan Chatzopoulos

### 2016-10-05 paper/widegap_submit/widegap_mri_paper_ApJ_submit.tex
widegap submitted version

--- 2016-10-05 paper/widegap_mri_paper.tex
@@
 \centering
-\includegraphics[width=\columnwidth]{../figures/widegap_paramspace_crit_params_Pm_16E-7}
+\includegraphics[width=\columnwidth]{widegap_paramspace_crit_params_Pm_16E-7}
 \caption{Growth rates in the ($\reym, k_z$) plane.
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/widegap_orders1_2_Pm16E-7.pdf}
+\includegraphics[width=\textwidth]{widegap_orders1_2_Pm16E-7}
 \caption{Eigenfunctions of the first order equations, first order adjoint homogenous equations, and second order equations.
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/widegap_n2_n3_Pm16E-7.pdf}
+\includegraphics[width=\textwidth]{widegap_n2_n3_Pm16E-7.pdf}
 \caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/widegap_thingap_comparison_zoombox.png}
+\includegraphics[width=\textwidth]{widegap_thingap_comparison_zoombox}
 \caption{Nonlinear term $N_{31}^{(A)}$ for the wide gap (left) and narrow gap (right) standard MRI, where the wide gap is the TC flow considered in this work.
@@
 \centering
-\includegraphics[width=\columnwidth]{../figures/widegap_coeffs_satampscaling_fivepoints.png}
+\includegraphics[width=\columnwidth]{widegap_coeffs_satampscaling_fivepoints}
 \caption{Critical parameters $\reym_c$ and $k_c$, and coefficients of the Ginzburg-Landau equation (Equation \ref{eq:gle}) as a function of $\Pm$.
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/widegap_vel_Bfield_streamfuncs_512_final_wider.png}
+\includegraphics[width=\textwidth]{widegap_vel_Bfield_streamfuncs_512_final_wider}
 \caption{Perturbation structure for the velocity and magnetic field of the fiducial standard MRI case, including first and second order perturbations.
@@
 \centering
-\includegraphics[width=\columnwidth]{../figures/widegap_totalstress_intstress.png}
+\includegraphics[width=\columnwidth]{widegap_totalstress_intstress}
 \caption{Total stress $\mathrm{T}(r, z) = u_r u_\phi - \frac{2}{\beta} B_r B_\phi$, i.e.
@@
 \centering
-\includegraphics[width=\textwidth]{../figures/hmri_velocity_Bfield_streamfuncs.png}
+\includegraphics[width=\textwidth]{hmri_velocity_Bfield_streamfuncs}
 \caption{As in Figure \ref{fig:smri_streamfuncs} but for the fiducial helical MRI case, including first and second order perturbations.}

### 2017-02-08 paper/thingap_mri_paper.tex

@@
 \centering
-\includegraphics[trim=5cm 0cm 0cm 0cm, scale=.35]{setup_diagram.eps}
+\includegraphics[trim=5cm 0cm 0cm 0cm, scale=.35]{setup_diagram-eps-converted-to.pdf}
 \caption{Schematic diagram of our set-up, an axisymmetric thin-gap Taylor-Couette flow.
@@
 \includegraphics[width=\columnwidth]{../figures/thingap_saturation_mechanism_plots_velocity_slice60.png}
-\caption{Top panel: Mean radial profile for $u_y^{saturated} - u_y^0$, the difference between the saturated azimuthal velocity profile and the initial azimuthal velocity profile (TC flow).
+\caption{Top panel: radial profile for $u_y^{saturated} - u_y^0$, the difference between the saturated azimuthal velocity profile and the initial azimuthal velocity profile (TC flow).
 Bottom panel: each term in the steady state force balance (Equation \ref{eqset2} with $\partial_t u_{y} = 0$).
@@
 \includegraphics[width=\columnwidth]{../figures/thingap_saturation_mechanism_plots_Bfield_slice60.png}
-\caption{Top panel: Mean radial profile for $B_z^{saturated}$, the saturated vertical magnetic field (black line).
+\caption{Top panel: radial profile for $B_z^{saturated}$, the saturated vertical magnetic field (black line).
 $B_z^0 = 1$ is the constant background magnetic field (gray line).

### 2017-02-08 paper/thingap_submit/thingap_mri_paper_ApJ_submit.tex

@@
 \centering
-\includegraphics[width=\textwidth]{thingap_streamfuncs_velocity_Pm_1E-3}
+\includegraphics[width=\textwidth]{thingap_streamfuncs_velocity_Pm_1E-3-eps-converted-to.pdf}
 \caption{First order (left), second order (center), and total (right) velocity perturbations.
@@
 \centering
-\includegraphics[width=\textwidth]{thingap_streamfuncs_Bfield_Pm_1E-3}
+\includegraphics[width=\textwidth]{thingap_streamfuncs_Bfield_Pm_1E-3-eps-converted-to.pdf}
 \caption{As in Figure \ref{fig:allorders_velocity} but for the magnetic field.

### 2017-02-08 paper/widegap_submit/widegap_mri_paper_ApJ_submit.tex

@@
 \centering
-\includegraphics[width=\columnwidth]{widegap_paramspace_crit_params_Pm_16E-7}
+\includegraphics[width=\columnwidth]{widegap_paramspace_crit_params_Pm_16E-7-eps-converted-to.pdf}
 \caption{Growth rates in the ($\reym, k_z$) plane.
@@
 \centering
-\includegraphics[width=\textwidth]{widegap_thingap_comparison_zoombox}
+\includegraphics[width=\textwidth]{widegap_thingap_comparison_zoombox-eps-converted-to.pdf}
 \caption{Nonlinear term $N_{31}^{(A)}$ for the wide gap (left) and narrow gap (right) standard MRI, where the wide gap is the TC flow considered in this work.
@@
 \centering
-\includegraphics[width=\columnwidth]{widegap_coeffs_satampscaling_fivepoints}
+\includegraphics[width=\columnwidth]{widegap_coeffs_satampscaling_fivepoints-eps-converted-to.pdf}
 \caption{Critical parameters $\reym_c$ and $k_c$, and coefficients of the Ginzburg-Landau equation (Equation \ref{eq:gle}) as a function of $\Pm$.
@@
 \centering
-\includegraphics[width=\textwidth]{widegap_vel_Bfield_streamfuncs_512_final_wider}
+\includegraphics[width=\textwidth]{widegap_vel_Bfield_streamfuncs_512_final_wider-eps-converted-to.pdf}
 \caption{Perturbation structure for the velocity and magnetic field of the fiducial standard MRI case, including first and second order perturbations.
@@
 \centering
-\includegraphics[width=\columnwidth]{widegap_totalstress_intstress}
+\includegraphics[width=\columnwidth]{widegap_totalstress_intstress-eps-converted-to.pdf}
 \caption{Total stress $\mathrm{T}(r, z) = u_r u_\phi - \frac{2}{\beta} B_r B_\phi$, i.e.
@@
 \centering
-\includegraphics[width=\textwidth]{hmri_velocity_Bfield_streamfuncs}
+\includegraphics[width=\textwidth]{hmri_velocity_Bfield_streamfuncs-eps-converted-to.pdf}
 \caption{As in Figure \ref{fig:smri_streamfuncs} but for the fiducial helical MRI case, including first and second order perturbations.}

### 2017-02-17 paper/thingap_ref_response/thingap_mri_paper_ref_response_1.tex
new draft to respond to referee

--- 2017-02-08 paper/thingap_submit/thingap_mri_paper_ApJ_submit.tex
@@
 
-\beq
+\beq\label{eq:wallbcs}
 \Psi = \partial_x \Psi = u_y = A = \partial_x B_y = 0.
@@
 
+\subsection{\textbf{Shearing box and ambipolar diffusion}}
+
+Many studies of the MRI consider the instability in a shearing box, i.e.
+a wall-less local approximation that is meant to represent a small section of a disk.
+The shearing box obtains when Equations \ref{eqset1} - \ref{eqset4} are subjected to shear periodic radial boundary conditions rather than Equation \ref{eq:wallbcs}.
+The periodic nature of the shearing box allows us to decompose the fluid perturbations into Fourier modes proportional to $e^{i k_x x + i k_z z}$.
+This makes the shearing box MRI straightforward to treat analytically.
+However, as noted above, the fastest-growing linear MRI modes in the shearing box are also exact solutions of the nonlinear MRI equations -- that is, $J(\hat{\psi_0}, \hat{\psi_0}) = J(\hat{\psi_0}, \nabla^2 \hat{\psi_0}) = 0$ for $\hat{\psi_0} \propto e^{i k_x x + i k_z z}$.
+While this may be an appealing trait for analytic simplicity, it leads to the unphysical conclusion that the fastest growing modes will never nonlinearly interact.
+This `nonlinear property' will not be satisfied for two MRI modes with nonparallel wavenumbers, but with vertically periodic boundary conditions the most unstable mode has an axial vertical wavenumber.
+Thus a formal weakly nonlinear analysis cannot be conducted, as the most unstable mode will never nonlinearly interact with itself or its complex conjugate.
+Similarly, we cannot analytically examine interactions between MRI channel modes and damped eigenmodes belonging to other wave families.
+While the shearing box approximation allows
+
 \section{Discussion}
@@
 In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane, and the quadratic maximum is a consequence of the linear dispersion relation given in Appendix~\ref{app:dispersion}.
-In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well.
+\textbf{The geometry of the problem considered here mimics an idealized experimental setup, where the local MRI equations are subject to wall-like radial boundary conditions.} In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well.
+\textbf{How would the theory developed here change with the addition of vertical endcaps, as found in a real experiment?}%\textbf{Many studies of the MRI take place in the local shearing box, in which Equations \ref{eqset1} - \ref{eqset4} are solved subject to shear-periodic boundary conditions.
+This is meant }
 

### 2017-02-17 python/allorders_2.py

@@ -40,6 +40,12 @@ class MRI():
         self.gridnum = self.domain.bases[0].coeff_size
         self.x = self.domain.grid(0)
         
+        if self.domain.bases[0].__class__.__name__ == "Chebyshev":
+            self.setbcs = True
+        else:
+            self.setbcs = False
+            logger.info("Not running with Chebyshev basis; will not set boundary conditions.")
+        
         logger.info("MRI parameters: Q = {}; Rm = {}; Pm = {}; q = {}; beta = {}; norm = {}, Re = {}".format(self.Q, self.Rm, self.Pm, self.q, self.beta, norm, self.R))
         
         
@@ -56,21 +62,27 @@ class MRI():
             problem.add_bc("right(psi) = 0")
             problem.add_bc("left(psix) = 0")
             problem.add_bc("right(psix) = 0")
+            
+            # magnetic b.c.'s
+            problem.add_bc("left(A) = 0")
+            problem.add_bc("right(A) = 0")
+            problem.add_bc("left(Bx) = 0")
+            problem.add_bc("right(Bx) = 0")
         
         # free-slip b.c.'s
         else:
-            problem.add_bc("left(ux) = 0")
-            problem.add_bc("right(ux) = 0")
+            problem.add_bc("left(u) = 0")
+            problem.add_bc("right(u) = 0")
             problem.add_bc("left(psi) = 0")
             problem.add_bc("right(psi) = 0")
             problem.add_bc("left(psixx) = 0")
             problem.add_bc("right(psixx) = 0")
         
-        # magnetic b.c.'s
-        problem.add_bc("left(A) = 0")
-        problem.add_bc("right(A) = 0")
-        problem.add_bc("left(Bx) = 0")
-        problem.add_bc("right(Bx) = 0")
+            # magnetic b.c.'s
+            problem.add_bc("left(A) = 0")
+            problem.add_bc("right(A) = 0")
+            problem.add_bc("left(B) = 0")
+            problem.add_bc("right(B) = 0")
         
         return problem
 
@@ -261,9 +273,12 @@ class AdjointHomogenous(MRI):
         lv1.add_equation("dx(u) - ux = 0")
[…]

### 2017-02-17 paper/thingap_ref_response/thingap_mri_paper_ref_response_1.tex
added some text on the shearing box and AD

@@
 While this may be an appealing trait for analytic simplicity, it leads to the unphysical conclusion that the fastest growing modes will never nonlinearly interact.
-This `nonlinear property' will not be satisfied for two MRI modes with nonparallel wavenumbers, but with vertically periodic boundary conditions the most unstable mode has an axial vertical wavenumber.
+This `nonlinear property' will not be satisfied for two MRI modes with nonparallel wavenumbers, but with vertically periodic boundary conditions the most unstable mode has a strictly axial vertical wavenumber.
 Thus a formal weakly nonlinear analysis cannot be conducted, as the most unstable mode will never nonlinearly interact with itself or its complex conjugate.
 Similarly, we cannot analytically examine interactions between MRI channel modes and damped eigenmodes belonging to other wave families.
-While the shearing box approximation allows
+This is analytically tracked for other plasma instabilities by tracking the amplitudes of growing, marginal, and damped eigenmodes simultaneously (CITE).
+While the shearing box approximation allows the projection of the perturbed MRI equations into the basis set of linear eigenmodes, nonlinear coupling between modes will remain zero.
+
+The nonlinear property of primary MRI modes in the shearing box motivates the addition of radial boundaries, such that the nonlinear evolution of the weakly nonlinear MRI can be properly considered.
+It also begs the question of whether some additional nonlinear mechanism can be introduced such that the fastest-growing modes are no longer nonlinear solutions to the shearing box equations.
+It has already been shown that the Hall effect does not negate the nonlinear property of primary MRI modes (CITE kunz and balbus).
+However, it seems to have been overlooked in the literature that these linear modes are not solutions of the nonlinear ambipolar diffusion term,
+
+\beq
+\nabla \times ((\mathbf{J} \times \mathbf{B}) \times \mathbf{B}). 
+\eeq
 

### 2017-02-22 paper/thingap_ref_response/thingap_mri_paper_ref_response_1.tex
dispersion relation notebook for shearing box MRI with ambipolar diffusion

@@
 While this may be an appealing trait for analytic simplicity, it leads to the unphysical conclusion that the fastest growing modes will never nonlinearly interact.
-This `nonlinear property' will not be satisfied for two MRI modes with nonparallel wavenumbers, but with vertically periodic boundary conditions the most unstable mode has a strictly axial vertical wavenumber.
+This `nonlinear property' will not be satisfied for two MRI modes with nonparallel wavenumbers, but with vertically periodic boundary conditions and a vertical background magnetic field the most unstable mode has a strictly axial vertical wavenumber.
 Thus a formal weakly nonlinear analysis cannot be conducted, as the most unstable mode will never nonlinearly interact with itself or its complex conjugate.
@@
 It also begs the question of whether some additional nonlinear mechanism can be introduced such that the fastest-growing modes are no longer nonlinear solutions to the shearing box equations.
-It has already been shown that the Hall effect does not negate the nonlinear property of primary MRI modes (CITE kunz and balbus).
-However, it seems to have been overlooked in the literature that these linear modes are not solutions of the nonlinear ambipolar diffusion term,
+It has already been shown that the Hall effect does not negate the nonlinear property of primary MRI modes \citep{Kunz:2013}.
+However, it seems to have been overlooked in the literature that these linear modes are \textit{not} solutions of the nonlinear ambipolar diffusion term,
 
@@
 \eeq
+
+Furthermore, the radial wavenumber of the fastest-growing linear MRI mode in a shearing box with ambipolar diffusion is nonzero when a constant azimuthal background field is considered in addition to an axial one \citep{Kunz:2004ib}.
 
@@
 
+\bibitem[{Kunz \& Balbus(2004)}]{Kunz:2004ib}
+Kunz, M. W. \& Balbus, S. A. 2004, MNRAS 348, 355
+
+\bibitem[{Kunz \& Lesur(2013)}]{Kunz:2013}
+Kunz, M. W. \& Lesur, G. 2013, MNRAS 434, 2295
+
 \bibitem[{Latter et al.(2015)}]{Latter:2015}

### 2017-03-02 paper/thingap_ref_response/thingap_mri_paper_ref_response_1.tex
new paper draft

@@
 It has already been shown that the Hall effect does not negate the nonlinear property of primary MRI modes \citep{Kunz:2013}.
-However, it seems to have been overlooked in the literature that these linear modes are \textit{not} solutions of the nonlinear ambipolar diffusion term,
+However, it seems to have been overlooked in the literature that these linear modes are \textit{not} solutions of the nonlinear ambipolar diffusion term, which takes the form
 
@@
 Furthermore, the radial wavenumber of the fastest-growing linear MRI mode in a shearing box with ambipolar diffusion is nonzero when a constant azimuthal background field is considered in addition to an axial one \citep{Kunz:2004ib}.
+This means that, in the presence of ambipolar diffusion, we can derive the weakly nonlinear envelope equation for the MRI in the shearing box.
+Ambipolar diffusion adds both linear and nonlinear terms to Equations \ref{eq:ordere} to \ref{eq:ordere3}, but does not change their $Z$ or $T$ dependence.
+Thus, the slow-scale evolution of the MRI in a shearing box with ambipolar diffusion is also governed by a Ginzburg-Landau equation.
 

### 2017-03-02 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
new paper draft

--- 2017-03-02 paper/thingap_ref_response/thingap_mri_paper_ref_response_1.tex
@@
-\documentclass{emulateapj}
+%%
+%% Beginning of file 'sample61.tex'
+%%
+%% Modified 2016 September
+%%
+%% This is a sample manuscript marked up using the
+%% AASTeX v6.1 LaTeX 2e macros.
+%%
+%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
+%% (Copyright 2000-2015).  See the classfile for details.
+
+%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
+%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
+%% All of these external packages should already be present in the modern TeX 
+%% distributions.  If not they can also be obtained at www.ctan.org.
+
+%% The first piece of markup in an AASTeX v6.x document is the \documentclass
+%% command. LaTeX will ignore any data that comes before this command. The 
+%% documentclass can take an optional argument to modify the output style.
+%% The command below calls the preprint style  which will produce a tightly 
+%% typeset, one-column, single-spaced document.  It is the default and thus
+%% does not need to be explicitly stated.
+%%
+%%
+%% using aastex version 6.1
+\documentclass[twocolumn]{aastex61}
+
+\usepackage{amsmath}
+
 \newcommand{\beq}{\begin{equation}}
 \newcommand{\eeq}{\end{equation}}
-\newcommand{\hyd}    {{\rm H}}
-\newcommand{\vobs}{v_{\rm obs}}
-\newcommand{\hi}{H{\sc i}~}
-\newcommand{\hii}{H{\sc ii}~}
-\newcommand{\hia}{H{\sc i}}
-\newcommand{\hiia}{H{\sc ii}}
-
-\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
-\usepackage{gensymb}
-\usepackage{amsmath}
-\usepackage{color}
-
-\usepackage[titletoc,title]{appendix}
-
-\makeatletter
-\setlength{\@fptop}{0pt}
-\setlength{\@fpbot}{0pt}
-\setlength{\@fpsep}{0pt}
-\makeatother
-
-\newcommand{\citei}[1]{\citeauthor{#1} \citeyear{#1}}
-\newcommand{\unit}[1]{\textrm{ #1}}
-\newcommand{\kms}{km ${\rm s^{-1}}$}
-\newcommand{\kmsa}{km ${\rm s^{-1}}$}
-\newcommand{\citeia}[2]{\citeauthor{#1}~(\citeyear{#1};~#2)}
 
@@
 
-\renewcommand{\topfraction}{0.85}
-\renewcommand{\textfraction}{0.1}
-\renewcommand{\floatpagefraction}{0.75}
+%% The default is a single spaced, 10 point font, single spaced article.
+%% There are 5 other style options available via an optional argument. They
+%% can be envoked like this:
+%%
+%% \documentclass[argument]{aastex61}
+%% 
+%% where the arguement options are:
+%%
+%%  twocolumn   : two text columns, 10 point font, single spaced article.
+%%                This is the most compact and represent the final published
+%%                derived PDF copy of the accepted manuscript from the publisher
+%%  manuscript  : one text column, 12 point font, double spaced article.
+%%  preprint    : one text column, 12 point font, single spaced article.  
+%%  preprint2   : two text columns, 12 point font, single spaced article.
+%%  modern      : a stylish, single text column, 12 point font, article with
+%% 		  wider left and right margins. This uses the Daniel
+%% 		  Foreman-Mackey and David Hogg design.
+%%
+%% Note that you can submit to the AAS Journals in any of these 6 styles.
+%%
+%% There are other optional arguments one can envoke to allow other stylistic
+%% actions. The available options are:
+%%
+%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
+%%  tighten      : Makes baselineskip slightly smaller, only works with 
+%%                 the twocolumn substyle.
+%%  times        : uses times font instead of the default
+%%  linenumbers  : turn on lineno package.
+%%  trackchanges : required to see the revision mark up and print its output
+%%  longauthor   : Do not use the more compressed footnote style (default) for 
+%%                 the author/collaboration/affiliations. Instead print all
+%%                 affiliation information after each name. Creates a much
+%%                 long author list but may be desirable for short author papers
+%%
+%% these can be used in any combination, e.g.
+%%
+%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex61}
+
+%% AASTeX v6.* now includes \hyperref support. While we have built in specific
+%% defaults into the classfile you can manually override them with the
+%% \hypersetup command. For example,
+%%
+%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
+%%
+%% will change the color of the internal links to red, the links to the
+%% bibliography to green, the file links to cyan, and the external links to
+%% magenta. Additional information on \hyperref options can be found here:
+%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
+
+%% If you want to create your own macros, you can do so
+%% using \newcommand. Your macros should appear before
+%% the \begin{document} command.
+%%
+\newcommand{\vdag}{(v)^\dagger}
+\newcommand\aastex{AAS\TeX}
+\newcommand\latex{La\TeX}
+
+%% Reintroduced the \received and \accepted commands from AASTeX v5.2
+%\received{July 1, 2016}
+%\revised{September 27, 2016}
+%\accepted{\today}
+%% Command to document which AAS Journal the manuscript was submitted to.
+%% Adds "Submitted to " the arguement.
+\submitjournal{ApJ}
+
+%% Mark up commands to limit the number of authors on the front page.
+%% Note that in AASTeX v6.1 a \collaboration call (see below) counts as
+%% an author in this case.
+%
+%\AuthorCollaborationLimit=3
+%
+%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
+%% collaboration" on the front page of this example manuscript.
+%%
+%% Note that all of the author will be shown in the published article.
+%% This feature is meant to be used prior to acceptance to make the
+%% front end of a long author article more manageable. Please do not use
+%% this functionality for manuscripts with less than 20 authors. Conversely,
+%% please do use this when the number of authors exceeds 40.
+%%
+%% Use \allauthors at the manuscript end to show the full author list.
+%% This command should only be used with \AuthorCollaborationLimit is used.
+
+%% The following command can be used to set the latex table counters.  It
+%% is needed in this document because it uses a mix of latex tabular and
+%% AASTeX deluxetables.  In general it should not be needed.
+%\setcounter{table}{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% The following section outlines numerous optional output that
+%% can be displayed in the front matter or as running meta-data.
+%%
+%% If you wish, you may supply running head information, although
+%% this information may be modified by the editorial offices.
+\shorttitle{The weakly nonlinear magnetorotational instability}
+\shortauthors{Clark \& Oishi}
+%%
+%% You can add a light gray and diagonal water-mark to the first page 
+%% with this command:
+% \watermark{text}
+%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
+%% long you can control the water-mark size with:
+%  \setwatermarkfontsize{dimension}
+%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% This is the end of the preamble.  Indicate the beginning of the
+%% manuscript itself with \begin{document}.
 
@@
 \title{The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor-Couette flow}
-\author{S. E. Clark\altaffilmark{1}}
-\author{Jeffrey S. Oishi\altaffilmark{2,}\altaffilmark{3}}
 
-\altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
-\altaffiltext{2}{Department of Physics and Astronomy, Bates College, Lewiston, ME}
-\altaffiltext{3}{Department of Astrophysics, American Museum of Natural History, New York, NY}
+%% LaTeX will automatically break titles if they run longer than
+%% one line. However, you may use \\ to force a line break if
+%% you desire. In v6.1 you can include a footnote in the title.
 
+%% A significant change from earlier AASTEX versions is in the structure for 
+%% calling author and affilations. The change was necessary to implement 
+%% autoindexing of affilations which prior was a manual process that could 
+%% easily be tedious in large author manuscripts.
+%%
+%% The \author command is the same as before except it now takes an optional
+%% arguement which is the 16 digit ORCID. The syntax is:
+%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
+%%
+%% This will hyperlink the author name to the author's ORCID page. Note that
+%% during compilation, LaTeX will do some limited checking of the format of
+%% the ID to make sure it is valid.
+%%
+%% Use \affiliation for affiliation information. The old \affil is now aliased
+%% to \affiliation. AASTeX v6.1 will automatically index these in the header.
+%% When a duplicate is found its index will be the same as its previous entry.
+%%
+%% Note that \altaffilmark and \altaffiltext have been removed and thus 
+%% can not be used to document secondary affiliations. If they are used latex
+%% will issue a specific error message and quit. Please use multiple 
+%% \affiliation calls for to document more than one affiliation.
+%%
+%% The new \altaffiliation can be used to indicate some secondary information
+%% such as fellowships. This command produces a non-numeric footnote that is
+%% set away from the numeric \affiliation footnotes.  NOTE that if an
+%% \altaffiliation command is used it must come BEFORE the \affiliation call,
+%% right after the \author command, in order to place the footnotes in
+%% the proper location.
+%%
+%% Use \email to set provide email addresses. Each \email will appear on its
+%% own line so you can put multiple email address in one \email call. A new
+%% \correspondingauthor command is available in V6.1 to identify the
+%% corresponding author of the manuscript. It is the author's responsibility
+%% to make sure this name is also in the author list.
+%%
+%% While authors can be grouped inside the same \author and \affiliation
+%% commands it is better to have a single author for each. This allows for
+%% one to exploit all the new benefits and should make book-keeping easier.
+%%
+%% If done correctly the peer review system will be able to
+%% automatically put the author and affiliation information from the manuscript
+%% and save the corresponding author the trouble of entering it by hand.
+
+\correspondingauthor{S. E. Clark}
+\email{seclark@astro.columbia.edu}
+
+\author[0000-0002-7633-3376]{S. E. Clark}
+\affil{Department of Astronomy, Columbia University, New York, NY}
+
+\author{Jeffrey S. Oishi}
+\affiliation{Department of Physics and Astronomy, Bates College, Lewiston, ME}
+\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY}
+\nocollaboration
+
+%% Note that the \and command from previous versions of AASTeX is now
+%% depreciated in this version as it is no longer necessary. AASTeX 
+%% automatically takes care of all commas and "and"s between authors names.
+
+%% AASTeX 6.1 has the new \collaboration and \nocollaboration commands to
+%% provide the collaboration status of a group of authors. These commands 
+%% can be used either before or after the list of corresponding authors. The
+%% argument for \collaboration is the collaboration identifier. Authors are
+%% encouraged to surround collaboration identifiers with ()s. The 
+%% \nocollaboration command takes no argument and exists to indicate that
+%% the nearby authors are not part of surrounding collaborations.
+
+%% Mark off the abstract in the ``abstract'' environment. 
 \begin{abstract}
+
 The magnetorotational instability (MRI) is a fundamental process of accretion disk physics, but its saturation mechanism remains poorly understood despite considerable theoretical and computational effort.
@@
 We directly simulate the Ginzburg-Landau amplitude evolution for our system and demonstrate the pattern formation our model predicts on long length and time scales.
+
 \end{abstract}
 
-\section{Introduction}
+%% Keywords should appear after the \end{abstract} command. 
+%% See the online documentation for the full list of available subject
+%% keywords and the rules for their use.
+%\keywords{editorials, notices --- miscellaneous --- catalogs --- surveys}
 
+%% From the front matter, we move on to the body of the paper.
+%% Sections are demarcated by \section and \subsection, respectively.
+%% Observe the use of the LaTeX \label
+%% command after the \subsection to give a symbolic KEY to the
+%% subsection for cross-referencing in a \ref command.
+%% You can use LaTeX's \ref and \label commands to keep track of
+%% cross-references to sections, equations, tables, and figures.
+%% That way, if you change the order of any elements, LaTeX will
+%% automatically renumber them.
+
+%% We recommend that authors also use the natbib \citep
+%% and \citet commands to identify citations.  The citations are
+%% tied to the reference list via symbolic KEYs. The KEY corresponds
+%% to the KEY in the \bibitem in the reference list below.
+
+\section{Introduction} \label{sec:intro}
 For matter to accrete from a disk onto a central object, angular momentum must be transported radially outward in the disk.
-The transport mechanism is likely turbulent, as molecular viscosity alone cannot account for the needed angular momentum transfer, and likely magnetic, as this turbulence is excited even in hydrodynamically stable disks (\citei{Shakura:1973wg}).
+The transport mechanism is likely turbulent, as molecular viscosity alone cannot account for the needed angular momentum transfer, and likely magnetic, as this turbulence is excited even in hydrodynamically stable disks \citep{Shakura:1973wg}.
 Discovered by \citet{Chandrasekhar:1960wh} and \citet{Velikhov:1959} in a global geometry, the magnetorotational instability (MRI) was subsequently rediscovered and applied to accretion disks by \citet{Balbus:1991vs}.
@@
 The latter authors undertook the first weakly nonlinear analysis of the MRI in a thin-gap Taylor-Couette (TC) flow with strong dissipation (as is appropriate to liquid metal experiments), and found that the marginal MRI system approaches saturation in a manner analogous to that of Rayleigh-B\'enard convection.
-Weakly nonlinear analysis was instrumental in our understanding of Rayleigh-B\'enard convection saturation (\citei{Newell:1969}), and the similarities between convection and the thin-gap MRI are the result of important shared symmetries between the systems.
+Weakly nonlinear analysis was instrumental in our understanding of Rayleigh-B\'enard convection saturation \citep{Newell:1969}, and the similarities between convection and the thin-gap MRI are the result of important shared symmetries between the systems.
 The success of URM07 in modeling the MRI system near threshold merits further consideration, but we are unaware of any other attempts to expand upon their theoretical framework.
@@
 They consider a thin-gap regime as well, and so their theory may be considered the strongly nonlinear analogue to the one developed here.
-\citei{Vasil:2015} examines the weakly nonlinear MRI in a thin-gap regime in a minimal model, finding deep mathematical similarities between the MRI system and the elastodynamic instability of a buckling beam.
+\citet{Vasil:2015} examines the weakly nonlinear MRI in a thin-gap regime in a minimal model, finding deep mathematical similarities between the MRI system and the elastodynamic instability of a buckling beam.
 We discuss these results and their relation to ours in section~\ref{sec:discussion}.
 Several authors have investigated the behavior of the MRI when the boundary conditions are shear periodic, and so the MRI has no mechanism by which to modify the background shear flow profile.
-In this approximation linear MRI growth is dominated by channel modes, a type of MRI mode that, for periodic boundary conditions, are exact solutions of both the linear and nonlinear MRI equations (\citei{Goodman:1994ul}).
+In this approximation linear MRI growth is dominated by channel modes, a type of MRI mode that, for periodic boundary conditions, are exact solutions of both the linear and nonlinear MRI equations \citep{Goodman:1994ul}.
 In this regime the MRI saturates via parasitic instabilities, which feed off and destroy the primary MRI modes.
-Analytical investigation of this case reveals that MRI saturation can be caused by parasitic Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime (\citei{Pessah:2010ic}).
-The theory of MRI channel mode parasites is robust (see also \citei{Pessah:2009gm}, \citei{Latter:2010iz}, \citei{Rembiasz:2016}), but their importance may be overestimated by the local approximation \citep{Latter:2015}, and not germane to global analyses like the one presented here.
+Analytical investigation of this case reveals that MRI saturation can be caused by parasitic Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime \citep{Pessah:2010ic}.
+The theory of MRI channel mode parasites is robust \citep[see also]{Pessah:2009gm;Latter:2010iz;Rembiasz:2016}, but their importance may be overestimated by the local approximation \citep{Latter:2015}, and not germane to global analyses like the one presented here.
 \citet{Latter:2015} gives a detailed analysis of the relation between local and global linear MRI modes.
@@
 
-\beq\label{eq:momentum}
+\begin{eqnarray}\label{eq:momentum}
 \begin{split}
@@
 \end{split}
-\eeq
+\end{eqnarray}
 
@@
 
-\begin{multline}
+%\begin{widetext}
+\onecolumngrid
+\beq
 \label{eqset1}
-\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{y} \, = \\
-\, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi
-\end{multline}
+\partial_t \nabla^2 \Psi \, - \, 2 \partial_z u_{y} \, - \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, - \, \frac{1}{\reye}\nabla^4 \Psi \, = \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, - \, J\left(\Psi, \nabla^2 \Psi\right) \,
+\eeq
+%\end{eqnarray}
+%\end{widetext}
 
-\begin{multline}
+\beq
 \label{eqset2}
-\partial_t u_{y} \, + \, J\left(\Psi, u_{y}\right) \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \\
-\, \frac{2}{\beta}B_0\partial_z B_{y} \, + \, \frac{2}{\beta} J\left(A, B_{y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{y}
-\end{multline}
+\partial_t u_{y} \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \, \frac{2}{\beta}B_0\partial_z B_{y} \, + \, \frac{2}{\beta} J\left(A, B_{y}\right) \, - \, J\left(\Psi, u_{y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{y}
+\eeq
 
-\begin{multline}
+\beq
 \label{eqset3}
 \partial_t A \, = \, B_0 \partial_z \Psi \, + \, J\left(A, \Psi\right) \, + \, \frac{1}{\reym} \nabla^2 A
-\end{multline}
+\eeq
 
-\begin{multline}
+\beq
 \label{eqset4}
@@
 + \, \frac{1}{\reym} \nabla^2 B_{y}  \, - \, q \Omega_0 \partial_z A,
-\end{multline}
+\eeq
+\twocolumngrid
 
@@
 
-\begin{align}
+\begin{eqnarray}
 \mathcal{O}(\epsilon)&: \mathcal{L}\mathbf{V_1} + \mathcal{D}\partial_t \mathbf{V_1} = 0. \label{eq:ordere}\\
@@
 \mathcal{O}(\epsilon^3)&: \mathcal{L} \mathbf{V_3} + \mathcal{D}\partial_T \mathbf{V_1}  + \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} + \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} + \widetilde{\mathcal{G}}\mathbf{V_1} = \mathbf{N_3} \label{eq:ordere3}
-\end{align}
+\end{eqnarray}
 
@@
 We solve the radial component of each equation using the open source pseudospectral code Dedalus.
-We compute the radial components on a grid of Chebyshev polynomials, as is appropriate for bounded one-dimensional domains (e.g.
-\citei{Boyd:2001aa}).
+We compute the radial components on a grid of Chebyshev polynomials, as is appropriate for bounded one-dimensional domains \citep[e.g.]{Boyd:2001aa}.
 The nonuniform spacing of the Chebyshev grid allows us to resolve the boundary layers well on a 128-point grid.
@@
 
-\subsection{\textbf{Shearing box and ambipolar diffusion}}
+\subsection{\edit1{Shearing box and ambipolar diffusion}}
 
@@
 
+The Ginzburg-Landau form of the amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to the wavenumber \citep{Hoyle:2006}.
+In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane, and the quadratic maximum is a consequence of the linear dispersion relation given in Appendix~\ref{app:dispersion}.
+In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well.
+The Ginzburg-Landau equation arises due to symmetries in the local MRI equations, irrespective of the boundary conditions to which they are subjected.
+This means that the local MRI is able to saturate via nonlinear mode interaction so long as the primary MRI modes are not exact solutions of the nonlinear terms.
+This can be achieved by considering the effects of ambipolar diffusion when the boundary conditions are shear periodic, or by enforcing wall-like radial boundary conditions.
+Both constructions require the most unstable mode to have nonconstant radial structure.
+%So long as our system permits the most unstable mode to nonlinearly interact with
+
+%The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to the wavenumber \citep{Hoyle:2006}.
+In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane, and the quadratic maximum is a consequence of the linear dispersion relation given in Appendix~\ref{app:dispersion}.
+In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well.
+\textbf{How would the theory developed here change with the addition of vertical endcaps, as found in a real experiment?}%\textbf{Many studies of the MRI take place in the local shearing box, in which Equations \ref{eqset1} - \ref{eqset4} are solved subject to shear-periodic boundary conditions.
+This is meant }
+
 \section{Discussion}
@@
 
-The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to the wavenumber \citep{Hoyle:2006}.
+%The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to the wavenumber \citep{Hoyle:2006}.
 In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane, and the quadratic maximum is a consequence of the linear dispersion relation given in Appendix~\ref{app:dispersion}.
-\textbf{The geometry of the problem considered here mimics an idealized experimental setup, where the local MRI equations are subject to wall-like radial boundary conditions.} In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well.
+In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well.
 \textbf{How would the theory developed here change with the addition of vertical endcaps, as found in a real experiment?}%\textbf{Many studies of the MRI take place in the local shearing box, in which Equations \ref{eqset1} - \ref{eqset4} are solved subject to shear-periodic boundary conditions.
@@
 
-\clearpage
 \appendix
-
-\section{A. Detailed Equations}\label{app:matrices}
-
-Here we detail the perturbation analysis described in Section \ref{sec:wnl}. The perturbation series is described by Equations \ref{eq:ordere} - \ref{eq:ordere3}, where
-
-\beq
-\mathcal{L} = \mathcal{L}_0 + \mathcal{L}_1 \partial_z + \mathcal{L}_2 \partial_z^2 + \mathcal{L}_3 \partial_z^3 + \mathcal{L}_4 \partial_z^4,
-\eeq
-
-\beq
-\widetilde{\mathcal{L}}_1 =  \mathcal{L}_1 + 2\mathcal{L}_2\partial_z + 3\mathcal{L}_3\partial_z^2 + 4\mathcal{L}_4\partial_z^3
-\eeq
-
-\beq
-\widetilde{\mathcal{L}}_2 = \mathcal{L}_2 + 3\mathcal{L}_3\partial_z + 6\mathcal{L}_4\partial_z^2
-\eeq
-
-\beq
-\widetilde{\mathcal{G}} = - \mathcal{G} \partial_z - \mathcal{L}_3 \partial_z^3,
-\eeq
-
-and the constituent matrices are defined as
-
-\beq
-\mathcal{D} = \left[\begin{matrix}
-\nabla^2 & 0 & 0 & 0 \\
-0 & 1& 0 & 0 \\
-0 & 0 & 1 & 0\\
-0 & 0 & 0 & 1 \\
-\end{matrix}\right]
-\eeq
-
-\beq
-\mathcal{L}_0 = \left[\begin{matrix}
--\frac{1}{\reye}\partial_x^4 & 0 & 0 & 0 \\
-0 & -\frac{1}{\reye}\partial_x^2 & 0 &0 \\
-0 & 0 & -\frac{1}{\reym}\partial_x^2 & 0 \\
-0 & 0 & 0 & -\frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right]
-\eeq
-
-\beq
-\mathcal{L}_1 = \left[\begin{matrix}
-0 & -2 & -\frac{2}{\beta}\partial_x^2 & 0 \\
-(2-q)\Omega_0 & 0 & 0 & -\frac{2}{\beta} \\
--1 & 0 & 0 & 0 \\
-0 & -1 & q\Omega_0 & 0 \\ \end{matrix}\right] 
-\eeq
-
-\beq
-\mathcal{L}_2 = \left[\begin{matrix}
--2\frac{1}{\reye} \partial_x^2 & 0 & 0 & 0 \\
-0 & -\frac{1}{\reye} & 0 & 0 \\
-0 & 0 & -\frac{1}{\reym} & 0 \\
-0 & 0 & 0 & -\frac{1}{\reym} \\ \end{matrix}\right]
-\eeq
-
-\beq
-\mathcal{L}_3 = \left[\begin{matrix}
-0 & 0 & -\frac{2}{\beta} & 0 \\
-0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\ \end{matrix} \right]
-\eeq
-
-\beq
-\mathcal{L}_4 = \left[\begin{matrix}
--\frac{1}{\reye} & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 \\ \end{matrix}\right] 
-\eeq
-
-\beq
-\mathcal{G} = \left[\begin{matrix}
-0 & 0 & -\frac{2}{\beta}\partial_x^2 & 0 \\
-0 & 0 & 0 & -\frac{2}{\beta} \\
--1 & 0 & 0 & 0 \\
-0 & -1 & 0 & 0 \\ \end{matrix} \right]
-\eeq
-
-Once perturbed, the system is solved for successive orders of $\epsilon$ (Equations \ref{eq:ordere} - \ref{eq:ordere3}).
-$\mathcal{O}(\epsilon)$ is the linear system.
-At $\mathcal{O}(\epsilon^2)$, first-order MRI modes nonlinearly interact with themselves and with their complex conjugates, and so the term $\mathbf{N}_2$ in Equation \ref{eq:ordere2} has the form
-
-\beq
-\mathbf{N}_2 = |\alpha|^2 \mathbf{N}_{20} + \alpha^2 \mathbf{N}_{22} e^{2 i k_c z}
-\eeq
-
-(see Appendix \ref{app:nonlinear_terms} for the full form of $\mathbf{N}_{20}$ and $\mathbf{N}_{22}$).
-
-Note that, following the notation of \citet{Umurhan:2007hs}, the subscripts refer to $\epsilon$ order, $z$ order, successively, such that $\mathbf{N}_{22}$ is the second-order nonlinear term which corresponds to $e^{2 i k_c z}$ z-dependence.
-
-Equation \ref{eq:ordere2} is solved as three separate systems of equations, one for each possible $z$ resonance:
-
-\begin{align}
-\mathcal{L}\mathbf{V}_{20} & = \mathbf{N}_{20}\\
-\mathcal{L}\mathbf{V}_{21} & = - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_{11} \\
-\mathcal{L}\mathbf{V}_{22} & = \mathbf{N}_{22}
-\end{align}
-
-Finally, at $\mathcal{O}(\epsilon^3)$ we eliminate secular terms to close the system.
-Secular terms are terms which are resonant with the solution to the homogenous linear equation (Equation \ref{eq:ordere}), and which cause the higher-order solutions to grow without bound.
-The solvability criterion we enforce to eliminate these terms is the vanishing of the inner product of the solution to the adjoint linear homogenous equation $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$ with the nonhomogenous terms in Equation \ref{eq:ordere3}, namely
-
-\beq\label{eq:innerprod_gle}
-\langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle \partial_T \alpha + \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} \rangle \alpha +  \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} \rangle \partial_Z^2 \alpha = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle \alpha |\alpha|^2.
-\eeq
-
-This solvability criterion derives from a corollary to the Fredholm Alternative (see Paper II for a formal definition).
-
-Equation \ref{eq:innerprod_gle} can be rewritten as Equation \ref{eq:gle}, the Ginzburg-Landau equation, where the coefficients are
-
-\beq
-b = \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
-\eeq
-
-\beq
-h = \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
-\eeq
-
-and
-
-\beq
-c = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle. 
-\eeq
-
-We define the adjoint operator $\mathcal{L}^\dagger$ and solution $\mathbf{V}^\dagger$ as
-
-\beq
-\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \langle \mathcal{L}^\dagger \mathbf{V}^\dagger | \mathbf{V} \rangle,
-\eeq
-
-where the inner product is defined as
-
-\beq\label{eq:inner_product_def}
-\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \frac{k_c}{2\pi} \int_{-\pi/k_c}^{\pi/k_c} \int_{x_1}^{x_2} \mathbf{V}^{\dagger*} \cdot \mathcal{L} \mathbf{V} \, \mathrm{d} x \mathrm{d} z.
-\eeq
-
-The solution to the adjoint homogenous equation has the form
-
-\beq
-\mathbf{V^\dagger} = \mathbb{V}^\dagger(x) e^{i k_c z} + c.c.
-\eeq
-
-As noted by URM07, a second amplitude equation for a spatially constant azimuthal magnetic field mode arises from the terms in the $\mathcal{O}(\epsilon^3)$ equation which contain no $z$ dependence.
-This is a diffusion equation, so the neutral mode simply decays away.
-
-\section{B. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}
-
-At each order in our perturbation series, lower-order MRI modes nonlinearly interact.
-Thus there is a nonlinear term contribution at $\mathcal{O}(\epsilon^2)$ and $\mathcal{O}(\epsilon^3)$.
-Here we detail the form of these nonlinear terms.
-
-The overall nonlinear contribution to our system, written as a vector $\mathbf{N}$ in Equation \ref{eq:unperturbed_matrix_equations}, is
-
-\beq
-\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4)
-\eeq
-
-where
-
-\beq
-N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \frac{2}{\beta}J(A_1, \nabla^2 A_1) 
-\eeq
-
-\beq
-N_2^{(u)} = J(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_1) 
-\eeq
-
-\beq
-N_2^{(A)} = - J(A_1, \Psi_1) 
-\eeq
-
-\beq
-N_2^{(B)} =  J(\Psi_1, B_1) \, - \, J(A_1, u_1)
-\eeq
-
-and
-
-\beq
-\begin{split}
-N_3^{(\Psi)} & = J(\Psi_1, \nabla^2\Psi_2) \, - \, \frac{2}{\beta}J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) - \, \frac{2}{\beta}J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, \\
-& - \, 2 \frac{2}{\beta}J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \frac{2}{\beta}\widetilde{J}(A_1, \nabla^2 A_1)\\
-\end{split}
-\eeq
-
-\beq
-N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_2) \, - \, \frac{2}{\beta} J(A_2, B_1) \, - \, \frac{2}{\beta} \widetilde{J}(A_1, B_1)
-\eeq
-
-\beq
-N_3^{(A)} = - J(A_1, \Psi_2) \, - \, J(A_2, \Psi_1) \, - \, \widetilde{J}(A_1, \Psi_1)
-\eeq
-
-\beq
-N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1).\\
-\eeq
-
-$\mathbf{N_2}$ and $\mathbf{N_3}$ expand to become
-
-\beq
-\mathbf{N_2} = \alpha^2\mathbb{N}_{22} e^{i2 k_c z} + \left|\alpha\right|^2 \mathbb{N}_{20} + c.c.
-\eeq
-
-and
-
-\beq
-\mathbf{N_3} = \alpha^3 \mathbb{N}_{33} e^{i 3 k_c z} + \alpha\partial_Z\alpha \mathbb{N}_{32} e^{i 2 k_c z} + \alpha\left|\alpha\right|^2 \mathbb{N}_{31} e^{i k_c z} + \alpha \partial_Z  \beta \mathbb{\widetilde{N}}_{31} e^{i k_c z} + \alpha^*\partial_Z \alpha \mathbb{N}_{30} + c.c.
-\eeq
-
-The second order nonlinear terms are \\
-
-\beq
-\begin{split}
-N_{22}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11} - k_c^2 \partial_x \Psi_{11}\right) - \partial_x \Psi_{11} \cdot \left(i k_c \partial_x^2 \Psi_{11} - i k_c^3 \Psi_{11}\right) \\
-& + \frac{2}{\beta}\partial_x A_{11} \cdot \left(i k_c \partial_x^2 A_{11} - i k_c^3 A_{11}\right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11} - k_c^2 \partial_x A_{11}\right)
-\end{split}
-\eeq
-
-\beq
-\begin{split}
-N_{22}^{(u)} = & \, i k_c \Psi_{11} \cdot \partial_x u_{11} - \partial_x \Psi_{11} \cdot i k_c u_{11} - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11} + \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}
-\end{split}
-\eeq
-
-\beq
-N_{22}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11} + \partial_x A_{11} \cdot i k_c \Psi_{11}
-\eeq
-
-\beq
-N_{22}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11} - \partial_x \Psi_{11} \cdot i k_c B_{11} - i k_c A_{11} \cdot \partial_x u_{11} + \partial_x A_{11} \cdot i k_c u_{11}
-\eeq
-
-\beq
-\begin{split}
-N_{20}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - k_c^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i k_c^3 \Psi_{11}^* - i k_c \partial_x^2 \Psi_{11}^*\right) \\
-& +\frac{2}{\beta} \partial_x A_{11} \cdot \left(i k_c^3 A_{11}^* - i k_c \partial_x^2 A_{11}^* \right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
-\end{split}
-\eeq
-
-\beq
-N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* + \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11}^* - \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}^*
-\eeq
-
-\beq
-N_{20}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11}^* - \partial_x A_{11} \cdot i k_c \Psi_{11}^*
-\eeq
-
-\beq
-N_{20}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11}^* + \partial_x \Psi_{11} \cdot i k_c B_{11}^* - i k_c A_{11} \cdot \partial_x u_{11}^* - \partial_x A_{11} \cdot i k_c u_{11}^*
-\eeq \\
-
-and the third order nonlinear terms become \\
-
-\beq
-\begin{split}
-N_{31}^{(\Psi)} = &  \, i k_c \left(\Psi_{11} \cdot \partial_x^3 \Psi_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x^3\Psi_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x^3 \Psi_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot \partial_x^2 \Psi_{22}\right) \\
-& + i 8 k_c^3 \left(\partial_x\Psi_{11}^* \cdot \Psi_{22}\right) + i 4 k_c^3 \left(\Psi_{11}^* \cdot \partial_x \Psi_{22}\right) + \frac{2}{\beta}\left[ - i k_c \left(A_{11} \cdot \partial^3 A_{20}\right) - i k_c \left(A_{11} \cdot \partial_x^3 A_{20}^*\right) \right]\\
-& + \frac{2}{\beta}\left[i k_c \left(A_{11}^* \cdot \partial_x^3 A_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \partial_x^2 A_{22}\right) - i 8 k_c^3 \left(\partial_x A_{11}^* \cdot A_{22}\right) - i 4 k_c^3 \left(A_{11}^* \cdot \partial_x A_{22}\right)\right] \\
-& + i 2 k_c \left(\Psi_{22} \cdot \partial_x^3 \Psi_{11}^*\right) - i 2 k_c^3\left(\Psi_{22} \cdot \partial_x \Psi_{11}^* \right) - i k_c \left(\partial_x \Psi_{20} \cdot \partial_x^2 \Psi_{11}\right) + i k_c \left(\partial_x \Psi_{22} \cdot \partial_x^2 \Psi_{11}^*\right) \\
-& - i k_c \left(\partial_x \Psi_{20}^* \cdot \partial_x^2 \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20} \cdot \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20}^* \cdot \Psi_{11}\right) - i k_c^3 \left(\partial_x \Psi_{22} \cdot \Psi_{11}^*\right) \\
-& + \frac{2}{\beta} \left[ - i 2 k_c \left(A_{22} \cdot \partial_x^3 A_{11}^*\right) + i 2 k_c^3 \left(A_{22} \cdot \partial_x A_{11}^*\right) + i k_c \left(\partial_x A_{20} \cdot \partial_x^2 A_{11}\right) - i k_c \left(\partial_x A_{22} \cdot \partial_x^2 A_{11}^*\right) \right] \\
-& + \frac{2}{\beta} \left[ i k_c \left(\partial_x A_{20}^* \cdot \partial_x^2 A_{11}\right) - i k_c^3 \left(\partial_x A_{20} \cdot A_{11}\right) - i k_c^3 \left(\partial_x A_{20}^* \cdot A_{11}\right) + i k_c^3 \left(\partial_x A_{22} \cdot A_{11}^*\right)\right]
-\end{split}
-\eeq
-
-\beq
-\begin{split}
-N_{31}^{(u)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x u_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x u_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x u_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot u_{22}\right) \\
-& - i k_c \left(u_{11} \cdot \partial_x \Psi_{20} \right) - i k_c \left(u_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c \left(u_{11}^* \cdot \partial_x \Psi_{22} \right) + i 2 k_c \left(\partial_x u_{11}^* \cdot \Psi_{22}\right) \\
-& +\frac{2}{\beta}\left[- i k_c \left(A_{11} \cdot \partial_x B_{20}\right) - i k_c \left(A_{11} \cdot \partial_x B_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x B_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot B_{22}\right)\right] \\
-& +\frac{2}{\beta}\left[ i k_c \left(B_{11} \cdot \partial_x A_{20}\right) + i k_c \left(B_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(B_{11}^* \cdot \partial_x A_{20}\right) - i 2 k_c \left(\partial_x B_{11}^* \cdot A_{22}\right)\right]
-\end{split}
-\eeq
-
-\beq
-\begin{split}
-N_{31}^{(A)} =\, & -i k_c \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i k_c \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
- + i k_c \left(A_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \Psi_{22}\right) \\
- & + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
-\end{split}
-\eeq
-
-\beq
-\begin{split}
-N_{31}^{(B)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x B_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x B_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x B_{22}\right) - i 2 k_c\left(\partial_x \Psi_{11}^* \cdot B_{22}\right) \\
-& - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}\right) - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c\left(B_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x B_{11}^* \cdot \Psi_{22}\right) \\
-& - i k_c \left(A_{11} \cdot \partial_x u_{20}\right) - i k_c \left(A_{11} \cdot \partial_x u_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x u_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot u_{22}\right) \\
-& i k_c \left(u_{11} \cdot \partial_x A_{20} \right) + i k_c \left(u_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(u_{11}^* \cdot \partial_x A_{22}\right) - i 2 k_c \left(\partial_x u_{11}^* \cdot A_{22}\right)
-\end{split}
-\eeq
-
-\section{C. Linear dispersion relation}\label{app:dispersion}
-
-The linear dispersion relation, which determines the variable scalings in the multiple scales analysis.
-This relation is found by perturbing the linear system (Equation \ref{eq:ordere}) with a small perturbation of the form $e^{\sigma t + i k_x x + i k_z z}$.
-Note that the spatial eigenvalues appear as $k_z^2$ and $k_x^2$ at lowest order.
-
-\beq
-\begin{split}
-& \frac{B_{0}^{4} k_{x}^{2} k_{z}^{4}}{16 \pi^{2}} + \frac{B_{0}^{4} k_{z}^{6}}{16 \pi^{2}} - \frac{B_{0}^{2} \Omega_{0} k_{z}^{4} q}{2 \pi} - 2 \Omega_{0} k_{z}^{2} q \sigma^{2} - \frac{4 \sigma}{\reym} \Omega_{0} k_{x}^{2} k_{z}^{2} q - \frac{4 \sigma}{\reym} \Omega_{0} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{x}^{4} k_{z}^{2} q - \frac{4 \Omega_{0}}{\reym^{2}} k_{x}^{2} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{z}^{6} q \\
-& - k_{x}^{2} \sigma^{4} - k_{z}^{2} \sigma^{4} + 4 k_{z}^{2} \sigma^{2} - \frac{2 \sigma^{3}}{\reym} k_{x}^{4} - \frac{4 \sigma^{3}}{\reym} k_{x}^{2} k_{z}^{2} + \frac{8 \sigma}{\reym} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reym} k_{z}^{4} + \frac{8 \sigma}{\reym} k_{z}^{4} - \frac{k_{x}^{6} \sigma^{2}}{\reym^{2}} - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{4} k_{z}^{2} + \frac{4 k_{x}^{4}}{\reym^{2}} k_{z}^{2} \\
-& - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{2} k_{z}^{4} + \frac{8 k_{x}^{2}}{\reym^{2}} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reym^{2}} + \frac{4 k_{z}^{6}}{\reym^{2}} - \frac{2 \sigma^{3}}{\reye} k_{x}^{4} - \frac{4 \sigma^{3}}{\reye} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reye} k_{z}^{4} - \frac{4 k_{x}^{6} \sigma^{2}}{\reye \reym} -  \frac{12 k_{x}^{4} k_{z}^{2} \sigma^{2}}{\reye \reym} - \frac{12 k_{x}^{2} k_{z}^{4} \sigma^{2}}{\reye \reym} \\
-& - \frac{4 k_{z}^{6} \sigma^{2}}{\reye \reym} - \frac{2 k_{x}^{8} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye \reym^{2}} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye \reym^{2}} - \frac{2 k_{z}^{8} \sigma}{\reye \reym^{2}} - \frac{k_{x}^{6} \sigma^{2}}{\reye^{2}} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{4} k_{z}^{2} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{2} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reye^{2}} \\
-& - \frac{2 k_{x}^{8} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye^{2} \reym} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye^{2} \reym} - \frac{2 k_{z}^{8} \sigma}{\reye^{2} \reym} - \frac{k_{x}^{10}}{\reye^{2} \reym^{2}} - \frac{5 k_{x}^{8} k_{z}^{2}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{6} k_{z}^{4}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{4} k_{z}^{6}}{\reye^{2} \reym^{2}} \\
-& - \frac{5 k_{x}^{2} k_{z}^{8}}{\reye^{2} \reym^{2}} - \frac{k_{z}^{10}}{\reye^{2} \reym^{2}} = 0
-\end{split}
-\eeq
-
-\bibliographystyle{plain}
-
-\begin{thebibliography}{53}
-\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+\begin{thebibliography}{}
 
@@
 \bibitem[{Pessah \& Goodman(2009)}]{Pessah:2009gm}
-Pessah, M. E., \& Goodman, J., 2009, ApJ 698, L72
+Pessah, M. E., \& Goodman, J., 2009, \apj 698, L72
 
 \bibitem[{Rembiasz et al.(2016)}]{Rembiasz:2016}
-Rembiasz, T., Obergaulinger, M., Cerd\'a-Dur\'an, P., M\"uller, E., \& Aloy, M.A., 2016, MNRAS 456, 3782
+Rembiasz, T., Obergaulinger, M., Cerd\'a-Dur\'an, P., M\"uller, E., \& Aloy, M.A., 2016, \mnras 456, 3782
 
@@
 \bibitem[{Shakura \& Sunyaev(1973)}]{Shakura:1973wg}
-Shakura, N I and Sunyaev, R A, Astronomy \& Astrophysics, 1973, 24, 337
+Shakura, N. I. and Sunyaev, R. A., Astronomy \& Astrophysics, 1973, 24, 337
 
 \bibitem[{Umurhan et al.(2007a)}]{Umurhan:2007dz}
-Umurhan, O M, Regev, O, Menou, K, 2007, Phys. Rev. Letters, 98, 034501
+Umurhan, O. M., Regev, O., Menou, K., 2007, Phys. Rev. Letters, 98, 034501
 
 \bibitem[{Umurhan et al.(2007b)}]{Umurhan:2007hs}
-Umurhan, O M, Regev, O, Menou, K, 2007, Phys. Rev. E, 76, 036310
+Umurhan, O. M., Regev, O., Menou, K., 2007, Phys. Rev. E, 76, 036310
 
@@
 
+%% This command is needed to show the entire author+affilation list when
+%% the collaboration and author truncation commands are used.  It has to
+%% go at the end of the manuscript.
+%\allauthors
+
+%% Include this line if you are using the \added, \replaced, \deleted
+%% commands to see a summary list of all changes at the end of the article.
+%\listofchanges
+
 \end{document}
+
+% End of file `sample61.tex'.

### 2017-03-03 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
new title, a little more in new section

@@
 
-\title{The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor-Couette flow}
+%\title{The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor-Couette flow}
+\title{The Weakly Nonlinear Magnetorotational Instability in a local geometry}
 
@@
 It has already been shown that the Hall effect does not negate the nonlinear property of primary MRI modes \citep{Kunz:2013}.
-However, it seems to have been overlooked in the literature that these linear modes are \textit{not} solutions of the nonlinear ambipolar diffusion term, which takes the form
+However, it seems to have been overlooked in the literature that these linear modes are \textit{not} solutions of the nonlinear ambipolar diffusion term, which is proportional to
 
@@
 Both constructions require the most unstable mode to have nonconstant radial structure.
+Physically, this radial variation impedes the free exchange of angular momentum facilitated by the uniform stretching of channel modes.
 %So long as our system permits the most unstable mode to nonlinearly interact with

### 2017-03-04 paper/thingap_ref_response/response_to_referee.tex
draft response to referee

\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}

\begin{document}

\title{Response to Report on AAS02887 and AAS02889}% - received 11/13/16}
\author{S.E. Clark \& Jeffrey S. Oishi}

\maketitle

\textbf{We thank the reviewer for these thoughtful and thorough comments. Below please find the reviewer comments in plain text and our responses to these specific points in bold. First, an overarching point: our papers are separated into Paper I and Paper II, where Paper I considers the MRI equations in a local geometry and Paper II considers the global MRI equations. In the submitted version of Paper I, we considered only the local MRI equations with wall-like boundary conditions, the ``thin-gap Taylor Couette" setup considered in URM07. However, the WNL theory developed here applies regardless of the boundary conditions, as long as the most unstable mode meets certain conditions. In particular, the local equations considered here, when subjected to periodic boundary conditions, describe the ``shearing box" that is commonly considered in many simulations and analytic treatments of the MRI. In the resubmitted version of Paper I we identify conditions under which our theory applies to the shearing box: the additional consideration of the effects of ambipolar diffusion and a constant azimuthal component to the background field. We add a section (3.1) that details how these conditions allow the Ginzburg-Landau amplitude equation to obtain for the system with shear-periodic boundary conditions. We feel that this additional insight is a more direct connection between our work and most MRI simulations. We hope the referee will agree with the choice to use Paper I to focus on the local equations, while allowing Paper II to demonstrate for the first time that the weakly nonlinear theory developed for the local case is also applicable to a global, cylindrical geometry (as well as to consider the helical MRI). } \\

These papers investigate saturation of the magnetorotational (MRI) instability of a viscous, resistive, incompressible fluid in cylindrical geometry with rotation driven at the boundaries and parameters chosen such that the system is almost stable. Although Keplerian shear is imposed, the setup is more directly pertinent to laboratory experiments with liquid metals than to astrophysical accretion disks. Nevertheless, the authors make a good case that the work should be published in The Astrophysical Journal as a contribution to the MRI saturation problem, which is of great astrophysical importance.

Since these two papers are strongly linked, I am submitting a single, combined report. Although I recommend extensive revisions, I consider this an impressive piece of work by a young researcher who should be encouraged to continue working on this problem.

The following remarks apply to both papers:

\begin{itemize}				
\item One of the main ideas behind the “weakly nonlinear” approach is that near marginal stability there is only one or a very small band of unstable modes that self interact. There is a substantial body of work in the fusion plasma community that considers coupling to damped modes as an important ingredient in saturation. The names to look for are Hatch, Makwana, and Terry. I am not suggesting the authors redo their analysis, to include damped modes, but they should comment on this alternative, perhaps in the Introduction.

\item With regard to saturation through modification of the base flow, in the TC problem studied in these papers, the rotation law is enforced through viscous coupling of the interior fluid to the bounding cylinders. This introduces a timescale, the viscous diffusion time, with which the momentum transport driven by perturbations must compete. In an accretion disk the rotation law is enforced at every point directly by gravity. The authors should comment on how important they think this difference is to saturating the mode.

\textbf{In Paper II, the rotation law is enforced through viscous coupling to the boundaries. In Paper I, however, the rotation law is imposed directly at every point by the shear. }

\item The influence of boundary conditions and boundary layers is strong. To call out some examples: In Paper I, the amplitude ? converges to the constant value (b/c)1/2, independent of z. Is this really relevant to a lab experiment with vertical endcaps, or to a stratified disk? Have the authors experimented with a finite cylinder, even a very tall one? Likewise, the radial boundaries seem to play a role that a softer physical boundary might not, for example the boundary layers in Figure 2 of Paper II (in this case it would be useful to show blowups of the solution, as in Figure 4, and present evidence for numerical convergence). I would like to see more discussion of the role of these boundary conditions versus physical boundary conditions. Again, it is OK to model a lab experiment – but the connection to a natural system should be kept in mind.
\end{itemize}

4. Both papers are light on physical interpretation. This is especially an issue with Paper I – see below. I would particularly like to see:

A more extended discussion of how angular momentum transport works. After all, this is the primary raison d’etre for most studies of the MRI.

A more extensive discussion of the role of the neutral mode with amplitude ? that both URM and the present paper argue should be kept, but, at least to the order in ? considered, satisfies an amplitude equation that is completely decoupled from the MRI mode, with amplitude ?. See the comment on coupling of damped modes above, which may be related to this issue.
Neutral mode doesn’t exist in cylindrical geometry, yet we find same GLE -- can we just argue that it’s not important for saturation? 
The beta mode arises because a constant By solves the equations. However, this cannot non-linearly couple to any other modes, because all NL terms in the problem involve derivatives. Thus, there is no coupling to it, even though we discard couplings between the other damped eigenmode families in the thin-gap MHD system (that is, the non-MRI modes)

The saturated Bz profile shown in Figure 5 of Paper I is intriguing in light of the fact that the authors achieved instability by dialing down Bz from the marginally stable value. Would a radial average of the new Bz, which is stronger at the edges and weaker in the middle, reveal that it is marginally stable? Do the results bear on whether the MRI is a dynamo? There is no counterpart figure in Paper II; the authors should comment on dynamo action here, too.
No, axisymmetry / Cowling anti-dynamo theorem

5. A small point – the parameter called ? is not the standard definition; why not introduce the Cowling number?

6. Finally, would it be feasible to a compute a nonlinear solution with Dedalus (or some other platform) and see how accurate the perturbation theory really is, for different values of ?? Although the authors compare their results with Ebrahimi 2009, it would be desirable to have their own simulations for comparison.

Remarks on Paper I:

This paper follows URM07 very closely – how closely I did not appreciate until I looked at that paper. Unless the authors can substantially enhance their discussion of the physics, bring out details of how their approach differs from URM07 that escaped me, and/or more fully discuss the relationship between their work and other work on saturation such as that of Vasil 2015, Ebrahimi 2009, or the Hatch/Makwana/Terry approach mentioned above, it might be better to write one long paper that outlines the URM07 approach and applies it to the helical MRI. A secondary reason to suggest this is the strong overlap of the Introductions to Papers I and II.

\end{document}

### 2017-03-04 paper/thingap_ref_response/response_to_referee.tex
more ref response

@@
 In the resubmitted version of Paper I we identify conditions under which our theory applies to the shearing box: the additional consideration of the effects of ambipolar diffusion and a constant azimuthal component to the background field.
-We add a section (3.1) that details how these conditions allow the Ginzburg-Landau amplitude equation to obtain for the system with shear-periodic boundary conditions.
-We feel that this additional insight is a more direct connection between our work and most MRI simulations.
-We hope the referee will agree with the choice to use Paper I to focus on the local equations, while allowing Paper II to demonstrate for the first time that the weakly nonlinear theory developed for the local case is also applicable to a global, cylindrical geometry (as well as to consider the helical MRI).
-} \\
+We add a section (3.1) that details how these conditions allow the Ginzburg-Landau amplitude equation to obtain for the system with shear-periodic boundary conditions, as well as discussion throughout the text.
+This additional insight is a more direct connection between our work and most MRI simulations, in addition to experiments, and accordingly we are changing the title of Paper I to ``The weakly nonlinear magnetorotational instability in a local geometry" rather than ``in a thin-gap Taylor Couette flow".
+We hope the referee will agree with the choice to use Paper I to focus on the local equations, while allowing Paper II to demonstrate for the first time that the weakly nonlinear theory developed for the local case is also applicable to a global, cylindrical geometry, as well as to consider the helical MRI.
+Many of the referee's comments pushed us to consider applications of this theory beyond the connection to MRI experiments, and we detail more specific responses below.} \\
 

### 2017-03-06 paper/thingap_ref_response/response_to_referee.tex
lots more ref response

@@
 
-\begin{itemize}				
+\begin{enumerate}				
 \item One of the main ideas behind the “weakly nonlinear” approach is that near marginal stability there is only one or a very small band of unstable modes that self interact.
@@
 I am not suggesting the authors redo their analysis, to include damped modes, but they should comment on this alternative, perhaps in the Introduction.
+
+\textbf{Coupling to damped modes as explored in the plasma literature is not analytically possible in the walled TC flow, but can be analyzed in the shearing box when ambipolar diffusion is considered, as described in the new section of the text.}
 
@@
 Again, it is OK to model a lab experiment – but the connection to a natural system should be kept in mind.
-\end{itemize}
 
-4. Both papers are light on physical interpretation. This is especially an issue with Paper I – see below. I would particularly like to see:
+\item Both papers are light on physical interpretation. This is especially an issue with Paper I -- see below. I would particularly like to see:
 
-A more extended discussion of how angular momentum transport works. After all, this is the primary raison d’etre for most studies of the MRI.
+\begin{itemize}						 							
+\item A more extended discussion of how angular momentum transport works. After all, this is the primary raison d’etre for most studies of the MRI.
 
-A more extensive discussion of the role of the neutral mode with amplitude ?
-that both URM and the present paper argue should be kept, but, at least to the order in ?
-considered, satisfies an amplitude equation that is completely decoupled from the MRI mode, with amplitude ?.
+\item A more extensive discussion of the role of the neutral mode with amplitude $\beta$ that both URM and the present paper argue should be kept, but, at least to the order in $\epsilon$ considered, satisfies an amplitude equation that is completely decoupled from the MRI mode, with amplitude $\alpha$.
 See the comment on coupling of damped modes above, which may be related to this issue.
-Neutral mode doesn’t exist in cylindrical geometry, yet we find same GLE -- can we just argue that it’s not important for saturation? 
-The beta mode arises because a constant By solves the equations.
+
+\textbf{The neutral mode is a spatially constant azimuthal magnetic field.
+It cannot nonlinearly interact with any modes -- damped or otherwise -- because all of the nonlinearities involve derivatives.
+This remains true when ambipolar diffusion is considered.
+We have added two sentences to Section 3 to clarify this point.}
+
+%Neutral mode doesn’t exist in cylindrical geometry, yet we find same GLE -- can we just argue that it’s not important for saturation? 
+%The beta mode arises because a constant By solves the equations.
 However, this cannot non-linearly couple to any other modes, because all NL terms in the problem involve derivatives.
@@
 
-The saturated Bz profile shown in Figure 5 of Paper I is intriguing in light of the fact that the authors achieved instability by dialing down Bz from the marginally stable value.
+\item The saturated Bz profile shown in Figure 5 of Paper I is intriguing in light of the fact that the authors achieved instability by dialing down Bz from the marginally stable value.
 Would a radial average of the new Bz, which is stronger at the edges and weaker in the middle, reveal that it is marginally stable?
@@
 No, axisymmetry / Cowling anti-dynamo theorem
+\end{itemize}
 
-5. A small point – the parameter called ? is not the standard definition; why not introduce the Cowling number?
+\item A small point – the parameter called ? is not the standard definition; why not introduce the Cowling number?
 
-6.
-Finally, would it be feasible to a compute a nonlinear solution with Dedalus (or some other platform) and see how accurate the perturbation theory really is, for different values of ??
-Although the authors compare their results with Ebrahimi 2009, it would be desirable to have their own simulations for comparison.
+\item Finally, would it be feasible to a compute a nonlinear solution with Dedalus (or some other platform) and see how accurate the perturbation theory really is, for different values of ??
+Although the authors compare their results with Ebrahimi 2009, it would be desirable to have their own simulations for comparison.					\end{enumerate}
 

### 2017-03-06 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
lots more ref response

@@
 
-where $\alpha$ is a slowly-varying amplitude equation and $c.c.$ denotes the complex conjugate.
+where $\alpha(T,Z)$ is a slowly-varying amplitude and $c.c.$ denotes the complex conjugate.
 The $x$ dependence is contained in $\mathbb{V}_{11} = (\Psi_{11}, u_{11}, A_{11}, B_{11})^\mathbf{T}$, and must be solved subject to the radial boundary conditions.
@@
 The amplitude $\beta(T, Z)$ encodes the slow evolution of this mode.
+\edit1{This spatially constant mode cannot contribute to the nonlinear saturation of the MRI because all of the nonlinearities involve derivatives.
+The long-term evolution of $\beta(T, Z)$ is described by a simple diffusion equation that decouples from $\alpha(T, Z)$, and so we neglect it in what follows.} %We also find, as in URM07, that the neutral mode amplitude $\beta(T, Z)$ is governed by a simple diffusion equation that decouples from the Ginzburg-Landau equation.
 
@@
 By analyzing the system over several orders of magnitude in $\Pm$, we reproduce the URM07 result that the analytic saturation amplitude scales as $\alpha_{saturation}^2 \propto \Pm^{4/3}$ in a thin-gap geometry when $\Pm \ll 1$.
-We also find, as in URM07, that the neutral mode amplitude $\beta(T, Z)$ is governed by a simple diffusion equation that decouples from the Ginzburg-Landau equation.
+%We also find, as in URM07, that the neutral mode amplitude $\beta(T, Z)$ is governed by a simple diffusion equation that decouples from the Ginzburg-Landau equation.
 
@@
 However, as noted above, the fastest-growing linear MRI modes in the shearing box are also exact solutions of the nonlinear MRI equations -- that is, $J(\hat{\psi_0}, \hat{\psi_0}) = J(\hat{\psi_0}, \nabla^2 \hat{\psi_0}) = 0$ for $\hat{\psi_0} \propto e^{i k_x x + i k_z z}$.
-While this may be an appealing trait for analytic simplicity, it leads to the unphysical conclusion that the fastest growing modes will never nonlinearly interact.
+While this may be an appealing trait for analytic simplicity, it leads to the unphysical conclusion that the fastest growing modes will never nonlinearly interact \citep{Goodman:1994ul}.
 This `nonlinear property' will not be satisfied for two MRI modes with nonparallel wavenumbers, but with vertically periodic boundary conditions and a vertical background magnetic field the most unstable mode has a strictly axial vertical wavenumber.
@@
 Similarly, we cannot analytically examine interactions between MRI channel modes and damped eigenmodes belonging to other wave families.
-This is analytically tracked for other plasma instabilities by tracking the amplitudes of growing, marginal, and damped eigenmodes simultaneously (CITE).
+This is analytically tracked for other plasma instabilities by tracking the amplitudes of growing, marginal, and damped eigenmodes simultaneously \citep[e.g.][]{Makwana:2011}.
 While the shearing box approximation allows the projection of the perturbed MRI equations into the basis set of linear eigenmodes, nonlinear coupling between modes will remain zero.
@@
 Ambipolar diffusion adds both linear and nonlinear terms to Equations \ref{eq:ordere} to \ref{eq:ordere3}, but does not change their $Z$ or $T$ dependence.
+The constant azimuthal background field component does not contribute to any other terms in the local MRI equations.
 Thus, the slow-scale evolution of the MRI in a shearing box with ambipolar diffusion is also governed by a Ginzburg-Landau equation.
@@
 
+\bibitem[{Makwana et al.(2011)}]{Makwana:2011}
+Makwana, K.D., Terry, P.W., Kim, J.H., \& Hatch, D.R. 2011, \pop, 18, 1
+
 \bibitem[{Newell \& Whitehead(1969)}]{Newell:1969}

### 2017-03-08 paper/thingap_ref_response/response_to_referee.tex
more response, Bz and no dynamo comment

@@
 There is no counterpart figure in Paper II; the authors should comment on dynamo action here, too.
-No, axisymmetry / Cowling anti-dynamo theorem
+
+\textbf{Yes, the saturated Bz is marginally stable, and we have added a sentence to that effect in the text. We cannot comment on dynamo action because our setup is axisymmetric. }
+
 \end{itemize}
 
-\item A small point – the parameter called ? is not the standard definition; why not introduce the Cowling number?
+\item A small point – the parameter called $\beta$ is not the standard definition; why not introduce the Cowling number?
 
-\item Finally, would it be feasible to a compute a nonlinear solution with Dedalus (or some other platform) and see how accurate the perturbation theory really is, for different values of ??
+\item Finally, would it be feasible to a compute a nonlinear solution with Dedalus (or some other platform) and see how accurate the perturbation theory really is, for different values of $\epsilon$?
 Although the authors compare their results with Ebrahimi 2009, it would be desirable to have their own simulations for comparison.					\end{enumerate}
@@
 
+\textbf{}
+
 \end{document}

### 2017-03-08 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
more response, Bz and no dynamo comment

@@
 Here, the instability acts to push magnetic field toward the boundaries in both the bulk and the boundary layers.
-Note also that the total magnetic energy in the $B_z$ component is significantly larger than the initial condition.
+%Note also that the total magnetic energy in the $B_z$ component is significantly larger than the initial condition.
+\edit1{The radial average of the saturated $B_z$ is $B_0$, i.e. $B_z$ is marginally stable.}
  \citet{Ebrahimi:2009ey} considered the saturation of a single, strongly super-critical MRI mode allowed to interact non-linearly only with itself and the mean.
@@
 
+\clearpage
+\appendix
+
+\section{A. Detailed Equations}\label{app:matrices}
+
+Here we detail the perturbation analysis described in Section \ref{sec:wnl}. The perturbation series is described by Equations \ref{eq:ordere} - \ref{eq:ordere3}, where
+
+\beq
+\mathcal{L} = \mathcal{L}_0 + \mathcal{L}_1 \partial_z + \mathcal{L}_2 \partial_z^2 + \mathcal{L}_3 \partial_z^3 + \mathcal{L}_4 \partial_z^4,
+\eeq
+
+\beq
+\widetilde{\mathcal{L}}_1 =  \mathcal{L}_1 + 2\mathcal{L}_2\partial_z + 3\mathcal{L}_3\partial_z^2 + 4\mathcal{L}_4\partial_z^3
+\eeq
+
+\beq
+\widetilde{\mathcal{L}}_2 = \mathcal{L}_2 + 3\mathcal{L}_3\partial_z + 6\mathcal{L}_4\partial_z^2
+\eeq
+
+\beq
+\widetilde{\mathcal{G}} = - \mathcal{G} \partial_z - \mathcal{L}_3 \partial_z^3,
+\eeq
+
+and the constituent matrices are defined as
+
+\beq
+\mathcal{D} = \left[\begin{matrix}
+\nabla^2 & 0 & 0 & 0 \\
+0 & 1& 0 & 0 \\
+0 & 0 & 1 & 0\\
+0 & 0 & 0 & 1 \\
+\end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{L}_0 = \left[\begin{matrix}
+-\frac{1}{\reye}\partial_x^4 & 0 & 0 & 0 \\
+0 & -\frac{1}{\reye}\partial_x^2 & 0 &0 \\
+0 & 0 & -\frac{1}{\reym}\partial_x^2 & 0 \\
+0 & 0 & 0 & -\frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{L}_1 = \left[\begin{matrix}
+0 & -2 & -\frac{2}{\beta}\partial_x^2 & 0 \\
+(2-q)\Omega_0 & 0 & 0 & -\frac{2}{\beta} \\
+-1 & 0 & 0 & 0 \\
+0 & -1 & q\Omega_0 & 0 \\ \end{matrix}\right] 
+\eeq
+
+\beq
+\mathcal{L}_2 = \left[\begin{matrix}
+-2\frac{1}{\reye} \partial_x^2 & 0 & 0 & 0 \\
+0 & -\frac{1}{\reye} & 0 & 0 \\
+0 & 0 & -\frac{1}{\reym} & 0 \\
+0 & 0 & 0 & -\frac{1}{\reym} \\ \end{matrix}\right]
+\eeq
+
+\beq
+\mathcal{L}_3 = \left[\begin{matrix}
+0 & 0 & -\frac{2}{\beta} & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\ \end{matrix} \right]
+\eeq
+
+\beq
+\mathcal{L}_4 = \left[\begin{matrix}
+-\frac{1}{\reye} & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 \\ \end{matrix}\right] 
+\eeq
+
+\beq
+\mathcal{G} = \left[\begin{matrix}
+0 & 0 & -\frac{2}{\beta}\partial_x^2 & 0 \\
+0 & 0 & 0 & -\frac{2}{\beta} \\
+-1 & 0 & 0 & 0 \\
+0 & -1 & 0 & 0 \\ \end{matrix} \right]
+\eeq
+
+Once perturbed, the system is solved for successive orders of $\epsilon$ (Equations \ref{eq:ordere} - \ref{eq:ordere3}).
+$\mathcal{O}(\epsilon)$ is the linear system.
+At $\mathcal{O}(\epsilon^2)$, first-order MRI modes nonlinearly interact with themselves and with their complex conjugates, and so the term $\mathbf{N}_2$ in Equation \ref{eq:ordere2} has the form
+
+\beq
+\mathbf{N}_2 = |\alpha|^2 \mathbf{N}_{20} + \alpha^2 \mathbf{N}_{22} e^{2 i k_c z}
+\eeq
+
+(see Appendix \ref{app:nonlinear_terms} for the full form of $\mathbf{N}_{20}$ and $\mathbf{N}_{22}$).
+
+Note that, following the notation of \citet{Umurhan:2007hs}, the subscripts refer to $\epsilon$ order, $z$ order, successively, such that $\mathbf{N}_{22}$ is the second-order nonlinear term which corresponds to $e^{2 i k_c z}$ z-dependence.
+
+Equation \ref{eq:ordere2} is solved as three separate systems of equations, one for each possible $z$ resonance:
+
+\begin{align}
+\mathcal{L}\mathbf{V}_{20} & = \mathbf{N}_{20}\\
+\mathcal{L}\mathbf{V}_{21} & = - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V}_{11} \\
+\mathcal{L}\mathbf{V}_{22} & = \mathbf{N}_{22}
+\end{align}
+
+Finally, at $\mathcal{O}(\epsilon^3)$ we eliminate secular terms to close the system.
+Secular terms are terms which are resonant with the solution to the homogenous linear equation (Equation \ref{eq:ordere}), and which cause the higher-order solutions to grow without bound.
+The solvability criterion we enforce to eliminate these terms is the vanishing of the inner product of the solution to the adjoint linear homogenous equation $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$ with the nonhomogenous terms in Equation \ref{eq:ordere3}, namely
+
+\beq\label{eq:innerprod_gle}
+\langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle \partial_T \alpha + \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} \rangle \alpha +  \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} \rangle \partial_Z^2 \alpha = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle \alpha |\alpha|^2.
+\eeq
+
+This solvability criterion derives from a corollary to the Fredholm Alternative (see Paper II for a formal definition).
+
+Equation \ref{eq:innerprod_gle} can be rewritten as Equation \ref{eq:gle}, the Ginzburg-Landau equation, where the coefficients are
+
+\beq
+b = \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
+\eeq
+
+\beq
+h = \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
+\eeq
+
+and
+
+\beq
+c = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle. 
+\eeq
+
+We define the adjoint operator $\mathcal{L}^\dagger$ and solution $\mathbf{V}^\dagger$ as
+
+\beq
+\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \langle \mathcal{L}^\dagger \mathbf{V}^\dagger | \mathbf{V} \rangle,
+\eeq
+
+where the inner product is defined as
+
+\beq\label{eq:inner_product_def}
+\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \frac{k_c}{2\pi} \int_{-\pi/k_c}^{\pi/k_c} \int_{x_1}^{x_2} \mathbf{V}^{\dagger*} \cdot \mathcal{L} \mathbf{V} \, \mathrm{d} x \mathrm{d} z.
+\eeq
+
+The solution to the adjoint homogenous equation has the form
+
+\beq
+\mathbf{V^\dagger} = \mathbb{V}^\dagger(x) e^{i k_c z} + c.c.
+\eeq
+
+As noted by URM07, a second amplitude equation for a spatially constant azimuthal magnetic field mode arises from the terms in the $\mathcal{O}(\epsilon^3)$ equation which contain no $z$ dependence.
+This is a diffusion equation, so the neutral mode simply decays away.
+
+\section{B. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}
+
+At each order in our perturbation series, lower-order MRI modes nonlinearly interact.
+Thus there is a nonlinear term contribution at $\mathcal{O}(\epsilon^2)$ and $\mathcal{O}(\epsilon^3)$.
+Here we detail the form of these nonlinear terms.
+
+The overall nonlinear contribution to our system, written as a vector $\mathbf{N}$ in Equation \ref{eq:unperturbed_matrix_equations}, is
+
+\beq
+\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4)
+\eeq
+
+where
+
+\beq
+N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \frac{2}{\beta}J(A_1, \nabla^2 A_1) 
+\eeq
+
+\beq
+N_2^{(u)} = J(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_1) 
+\eeq
+
+\beq
+N_2^{(A)} = - J(A_1, \Psi_1) 
+\eeq
+
+\beq
+N_2^{(B)} =  J(\Psi_1, B_1) \, - \, J(A_1, u_1)
+\eeq
+
+and
+
+\beq
+\begin{split}
+N_3^{(\Psi)} & = J(\Psi_1, \nabla^2\Psi_2) \, - \, \frac{2}{\beta}J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) - \, \frac{2}{\beta}J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, \\
+& - \, 2 \frac{2}{\beta}J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \frac{2}{\beta}\widetilde{J}(A_1, \nabla^2 A_1)\\
+\end{split}
+\eeq
+
+\beq
+N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_2) \, - \, \frac{2}{\beta} J(A_2, B_1) \, - \, \frac{2}{\beta} \widetilde{J}(A_1, B_1)
+\eeq
+
+\beq
+N_3^{(A)} = - J(A_1, \Psi_2) \, - \, J(A_2, \Psi_1) \, - \, \widetilde{J}(A_1, \Psi_1)
+\eeq
+
+\beq
+N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1).\\
+\eeq
+
+$\mathbf{N_2}$ and $\mathbf{N_3}$ expand to become
+
+\beq
+\mathbf{N_2} = \alpha^2\mathbb{N}_{22} e^{i2 k_c z} + \left|\alpha\right|^2 \mathbb{N}_{20} + c.c.
+\eeq
+
+and
+
+\beq
+\mathbf{N_3} = \alpha^3 \mathbb{N}_{33} e^{i 3 k_c z} + \alpha\partial_Z\alpha \mathbb{N}_{32} e^{i 2 k_c z} + \alpha\left|\alpha\right|^2 \mathbb{N}_{31} e^{i k_c z} + \alpha \partial_Z  \beta \mathbb{\widetilde{N}}_{31} e^{i k_c z} + \alpha^*\partial_Z \alpha \mathbb{N}_{30} + c.c.
+\eeq
+
+The second order nonlinear terms are \\
+
+\beq
+\begin{split}
+N_{22}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11} - k_c^2 \partial_x \Psi_{11}\right) - \partial_x \Psi_{11} \cdot \left(i k_c \partial_x^2 \Psi_{11} - i k_c^3 \Psi_{11}\right) \\
+& + \frac{2}{\beta}\partial_x A_{11} \cdot \left(i k_c \partial_x^2 A_{11} - i k_c^3 A_{11}\right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11} - k_c^2 \partial_x A_{11}\right)
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{22}^{(u)} = & \, i k_c \Psi_{11} \cdot \partial_x u_{11} - \partial_x \Psi_{11} \cdot i k_c u_{11} - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11} + \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}
+\end{split}
+\eeq
+
+\beq
+N_{22}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11} + \partial_x A_{11} \cdot i k_c \Psi_{11}
+\eeq
+
+\beq
+N_{22}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11} - \partial_x \Psi_{11} \cdot i k_c B_{11} - i k_c A_{11} \cdot \partial_x u_{11} + \partial_x A_{11} \cdot i k_c u_{11}
+\eeq
+
+\beq
+\begin{split}
+N_{20}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - k_c^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i k_c^3 \Psi_{11}^* - i k_c \partial_x^2 \Psi_{11}^*\right) \\
+& +\frac{2}{\beta} \partial_x A_{11} \cdot \left(i k_c^3 A_{11}^* - i k_c \partial_x^2 A_{11}^* \right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
+\end{split}
+\eeq
+
+\beq
+N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* + \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11}^* - \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}^*
+\eeq
+
+\beq
+N_{20}^{(A)} = - i k_c A_{11} \cdot \partial_x \Psi_{11}^* - \partial_x A_{11} \cdot i k_c \Psi_{11}^*
+\eeq
+
+\beq
+N_{20}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11}^* + \partial_x \Psi_{11} \cdot i k_c B_{11}^* - i k_c A_{11} \cdot \partial_x u_{11}^* - \partial_x A_{11} \cdot i k_c u_{11}^*
+\eeq \\
+
+and the third order nonlinear terms become \\
+
+\beq
+\begin{split}
+N_{31}^{(\Psi)} = &  \, i k_c \left(\Psi_{11} \cdot \partial_x^3 \Psi_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x^3\Psi_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x^3 \Psi_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot \partial_x^2 \Psi_{22}\right) \\
+& + i 8 k_c^3 \left(\partial_x\Psi_{11}^* \cdot \Psi_{22}\right) + i 4 k_c^3 \left(\Psi_{11}^* \cdot \partial_x \Psi_{22}\right) + \frac{2}{\beta}\left[ - i k_c \left(A_{11} \cdot \partial^3 A_{20}\right) - i k_c \left(A_{11} \cdot \partial_x^3 A_{20}^*\right) \right]\\
+& + \frac{2}{\beta}\left[i k_c \left(A_{11}^* \cdot \partial_x^3 A_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \partial_x^2 A_{22}\right) - i 8 k_c^3 \left(\partial_x A_{11}^* \cdot A_{22}\right) - i 4 k_c^3 \left(A_{11}^* \cdot \partial_x A_{22}\right)\right] \\
+& + i 2 k_c \left(\Psi_{22} \cdot \partial_x^3 \Psi_{11}^*\right) - i 2 k_c^3\left(\Psi_{22} \cdot \partial_x \Psi_{11}^* \right) - i k_c \left(\partial_x \Psi_{20} \cdot \partial_x^2 \Psi_{11}\right) + i k_c \left(\partial_x \Psi_{22} \cdot \partial_x^2 \Psi_{11}^*\right) \\
+& - i k_c \left(\partial_x \Psi_{20}^* \cdot \partial_x^2 \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20} \cdot \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20}^* \cdot \Psi_{11}\right) - i k_c^3 \left(\partial_x \Psi_{22} \cdot \Psi_{11}^*\right) \\
+& + \frac{2}{\beta} \left[ - i 2 k_c \left(A_{22} \cdot \partial_x^3 A_{11}^*\right) + i 2 k_c^3 \left(A_{22} \cdot \partial_x A_{11}^*\right) + i k_c \left(\partial_x A_{20} \cdot \partial_x^2 A_{11}\right) - i k_c \left(\partial_x A_{22} \cdot \partial_x^2 A_{11}^*\right) \right] \\
+& + \frac{2}{\beta} \left[ i k_c \left(\partial_x A_{20}^* \cdot \partial_x^2 A_{11}\right) - i k_c^3 \left(\partial_x A_{20} \cdot A_{11}\right) - i k_c^3 \left(\partial_x A_{20}^* \cdot A_{11}\right) + i k_c^3 \left(\partial_x A_{22} \cdot A_{11}^*\right)\right]
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{31}^{(u)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x u_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x u_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x u_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot u_{22}\right) \\
+& - i k_c \left(u_{11} \cdot \partial_x \Psi_{20} \right) - i k_c \left(u_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c \left(u_{11}^* \cdot \partial_x \Psi_{22} \right) + i 2 k_c \left(\partial_x u_{11}^* \cdot \Psi_{22}\right) \\
+& +\frac{2}{\beta}\left[- i k_c \left(A_{11} \cdot \partial_x B_{20}\right) - i k_c \left(A_{11} \cdot \partial_x B_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x B_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot B_{22}\right)\right] \\
+& +\frac{2}{\beta}\left[ i k_c \left(B_{11} \cdot \partial_x A_{20}\right) + i k_c \left(B_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(B_{11}^* \cdot \partial_x A_{20}\right) - i 2 k_c \left(\partial_x B_{11}^* \cdot A_{22}\right)\right]
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{31}^{(A)} =\, & -i k_c \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i k_c \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
+ + i k_c \left(A_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \Psi_{22}\right) \\
+ & + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
+\end{split}
+\eeq
+
+\beq
+\begin{split}
+N_{31}^{(B)} = \, & i k_c \left(\Psi_{11} \cdot \partial_x B_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x B_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x B_{22}\right) - i 2 k_c\left(\partial_x \Psi_{11}^* \cdot B_{22}\right) \\
+& - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}\right) - i k_c \left(B_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c\left(B_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x B_{11}^* \cdot \Psi_{22}\right) \\
+& - i k_c \left(A_{11} \cdot \partial_x u_{20}\right) - i k_c \left(A_{11} \cdot \partial_x u_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x u_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot u_{22}\right) \\
+& i k_c \left(u_{11} \cdot \partial_x A_{20} \right) + i k_c \left(u_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(u_{11}^* \cdot \partial_x A_{22}\right) - i 2 k_c \left(\partial_x u_{11}^* \cdot A_{22}\right)
+\end{split}
+\eeq
+
+\section{C. Linear dispersion relation}\label{app:dispersion}
+
+The linear dispersion relation, which determines the variable scalings in the multiple scales analysis.
+This relation is found by perturbing the linear system (Equation \ref{eq:ordere}) with a small perturbation of the form $e^{\sigma t + i k_x x + i k_z z}$.
+Note that the spatial eigenvalues appear as $k_z^2$ and $k_x^2$ at lowest order.
+
+\beq
+\begin{split}
+& \frac{B_{0}^{4} k_{x}^{2} k_{z}^{4}}{16 \pi^{2}} + \frac{B_{0}^{4} k_{z}^{6}}{16 \pi^{2}} - \frac{B_{0}^{2} \Omega_{0} k_{z}^{4} q}{2 \pi} - 2 \Omega_{0} k_{z}^{2} q \sigma^{2} - \frac{4 \sigma}{\reym} \Omega_{0} k_{x}^{2} k_{z}^{2} q - \frac{4 \sigma}{\reym} \Omega_{0} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{x}^{4} k_{z}^{2} q - \frac{4 \Omega_{0}}{\reym^{2}} k_{x}^{2} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{z}^{6} q \\
+& - k_{x}^{2} \sigma^{4} - k_{z}^{2} \sigma^{4} + 4 k_{z}^{2} \sigma^{2} - \frac{2 \sigma^{3}}{\reym} k_{x}^{4} - \frac{4 \sigma^{3}}{\reym} k_{x}^{2} k_{z}^{2} + \frac{8 \sigma}{\reym} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reym} k_{z}^{4} + \frac{8 \sigma}{\reym} k_{z}^{4} - \frac{k_{x}^{6} \sigma^{2}}{\reym^{2}} - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{4} k_{z}^{2} + \frac{4 k_{x}^{4}}{\reym^{2}} k_{z}^{2} \\
+& - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{2} k_{z}^{4} + \frac{8 k_{x}^{2}}{\reym^{2}} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reym^{2}} + \frac{4 k_{z}^{6}}{\reym^{2}} - \frac{2 \sigma^{3}}{\reye} k_{x}^{4} - \frac{4 \sigma^{3}}{\reye} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reye} k_{z}^{4} - \frac{4 k_{x}^{6} \sigma^{2}}{\reye \reym} -  \frac{12 k_{x}^{4} k_{z}^{2} \sigma^{2}}{\reye \reym} - \frac{12 k_{x}^{2} k_{z}^{4} \sigma^{2}}{\reye \reym} \\
+& - \frac{4 k_{z}^{6} \sigma^{2}}{\reye \reym} - \frac{2 k_{x}^{8} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye \reym^{2}} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye \reym^{2}} - \frac{2 k_{z}^{8} \sigma}{\reye \reym^{2}} - \frac{k_{x}^{6} \sigma^{2}}{\reye^{2}} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{4} k_{z}^{2} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{2} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reye^{2}} \\
+& - \frac{2 k_{x}^{8} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye^{2} \reym} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye^{2} \reym} - \frac{2 k_{z}^{8} \sigma}{\reye^{2} \reym} - \frac{k_{x}^{10}}{\reye^{2} \reym^{2}} - \frac{5 k_{x}^{8} k_{z}^{2}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{6} k_{z}^{4}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{4} k_{z}^{6}}{\reye^{2} \reym^{2}} \\
+& - \frac{5 k_{x}^{2} k_{z}^{8}}{\reye^{2} \reym^{2}} - \frac{k_{z}^{10}}{\reye^{2} \reym^{2}} = 0
+\end{split}
+\eeq
+
 \appendix
@@
 \bibitem[{Kunz \& Balbus(2004)}]{Kunz:2004ib}
-Kunz, M. W. \& Balbus, S. A. 2004, MNRAS 348, 355
+Kunz, M. W. \& Balbus, S. A. 2004, \mnras 348, 355
 
 \bibitem[{Kunz \& Lesur(2013)}]{Kunz:2013}
-Kunz, M. W. \& Lesur, G. 2013, MNRAS 434, 2295
+Kunz, M. W. \& Lesur, G. 2013, \mnras 434, 2295
 
 \bibitem[{Latter et al.(2015)}]{Latter:2015}
-Latter, H.N., Fromang, S., \& Faure, J., 2015, MNRAS 453, 3257
+Latter, H.N., Fromang, S., \& Faure, J., 2015, \mnras 453, 3257
 
 \bibitem[{Latter et al.(2010)}]{Latter:2010iz}
-Latter, H.N., Fromang, S., \& Gressel, O., 2010, MNRAS 406, 848
+Latter, H.N., Fromang, S., \& Gressel, O., 2010, \mnras 406, 848
 
 \bibitem[{Makwana et al.(2011)}]{Makwana:2011}
-Makwana, K.D., Terry, P.W., Kim, J.H., \& Hatch, D.R. 2011, \pop, 18, 1
+Makwana, K.D., Terry, P.W., Kim, J.H., \& Hatch, D.R. 2011, Phys. of Plasmas, 18, 1
 

### 2017-03-09 paper/thingap_ref_response/response_to_referee.tex
minor correction to text

@@
 In the resubmitted version of Paper I we identify conditions under which our theory applies to the shearing box: the additional consideration of the effects of ambipolar diffusion and a constant azimuthal component to the background field.
-We add a section (3.1) that details how these conditions allow the Ginzburg-Landau amplitude equation to obtain for the system with shear-periodic boundary conditions, as well as discussion throughout the text.
+We add a section (3.1) that details how these conditions allow the Ginzburg-Landau amplitude equation to be obtained for the system with shear-periodic boundary conditions, as well as discussion throughout the text.
 This additional insight is a more direct connection between our work and most MRI simulations, in addition to experiments, and accordingly we are changing the title of Paper I to ``The weakly nonlinear magnetorotational instability in a local geometry" rather than ``in a thin-gap Taylor Couette flow".

### 2017-03-12 paper/widegap_ref_response/widegap_mri_paper_ref_response_apj.tex
widegap revised draft for referee

--- 2017-02-08 paper/widegap_submit/widegap_mri_paper_ApJ_submit.tex
@@
-\documentclass{emulateapj}
+%%
+%% Beginning of file 'sample61.tex'
+%%
+%% Modified 2016 September
+%%
+%% This is a sample manuscript marked up using the
+%% AASTeX v6.1 LaTeX 2e macros.
+%%
+%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
+%% (Copyright 2000-2015).  See the classfile for details.
+
+%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
+%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
+%% All of these external packages should already be present in the modern TeX 
+%% distributions.  If not they can also be obtained at www.ctan.org.
+
+%% The first piece of markup in an AASTeX v6.x document is the \documentclass
+%% command. LaTeX will ignore any data that comes before this command. The 
+%% documentclass can take an optional argument to modify the output style.
+%% The command below calls the preprint style  which will produce a tightly 
+%% typeset, one-column, single-spaced document.  It is the default and thus
+%% does not need to be explicitly stated.
+%%
+%%
+%% using aastex version 6.1
+\documentclass[twocolumn]{aastex61}
+
+\usepackage{amsmath}
+
 \newcommand{\beq}{\begin{equation}}
 \newcommand{\eeq}{\end{equation}}
-
-\newcommand{\hyd}    {{\rm H}}
-\newcommand{\vobs}{v_{\rm obs}}
-\newcommand{\hi}{H{\sc i}~}
-\newcommand{\hii}{H{\sc ii}~}
-\newcommand{\hia}{H{\sc i}}
-\newcommand{\hiia}{H{\sc ii}}
-
-\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
-\usepackage{gensymb}
-\usepackage{amsmath}
-\usepackage{color}
-\usepackage[titletoc,title]{appendix}
-
-\makeatletter
-\setlength{\@fptop}{0pt}
-\setlength{\@fpbot}{0pt}
-\setlength{\@fpsep}{0pt}
-\makeatother
-
-\newcommand{\citei}[1]{\citeauthor{#1} \citeyear{#1}}
-\newcommand{\unit}[1]{\textrm{ #1}}
-\newcommand{\kms}{km ${\rm s^{-1}}$}
-\newcommand{\kmsa}{km ${\rm s^{-1}}$}
-\newcommand{\citeia}[2]{\citeauthor{#1}~(\citeyear{#1};~#2)}
 
@@
 
-\renewcommand{\topfraction}{0.85}
-\renewcommand{\textfraction}{0.1}
-\renewcommand{\floatpagefraction}{0.75}
+%% The default is a single spaced, 10 point font, single spaced article.
+%% There are 5 other style options available via an optional argument. They
+%% can be envoked like this:
+%%
+%% \documentclass[argument]{aastex61}
+%% 
+%% where the arguement options are:
+%%
+%%  twocolumn   : two text columns, 10 point font, single spaced article.
+%%                This is the most compact and represent the final published
+%%                derived PDF copy of the accepted manuscript from the publisher
+%%  manuscript  : one text column, 12 point font, double spaced article.
+%%  preprint    : one text column, 12 point font, single spaced article.  
+%%  preprint2   : two text columns, 12 point font, single spaced article.
+%%  modern      : a stylish, single text column, 12 point font, article with
+%% 		  wider left and right margins. This uses the Daniel
+%% 		  Foreman-Mackey and David Hogg design.
+%%
+%% Note that you can submit to the AAS Journals in any of these 6 styles.
+%%
+%% There are other optional arguments one can envoke to allow other stylistic
+%% actions. The available options are:
+%%
+%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
+%%  tighten      : Makes baselineskip slightly smaller, only works with 
+%%                 the twocolumn substyle.
+%%  times        : uses times font instead of the default
+%%  linenumbers  : turn on lineno package.
+%%  trackchanges : required to see the revision mark up and print its output
+%%  longauthor   : Do not use the more compressed footnote style (default) for 
+%%                 the author/collaboration/affiliations. Instead print all
+%%                 affiliation information after each name. Creates a much
+%%                 long author list but may be desirable for short author papers
+%%
+%% these can be used in any combination, e.g.
+%%
+%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex61}
+
+%% AASTeX v6.* now includes \hyperref support. While we have built in specific
+%% defaults into the classfile you can manually override them with the
+%% \hypersetup command. For example,
+%%
+%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
+%%
+%% will change the color of the internal links to red, the links to the
+%% bibliography to green, the file links to cyan, and the external links to
+%% magenta. Additional information on \hyperref options can be found here:
+%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
+
+%% If you want to create your own macros, you can do so
+%% using \newcommand. Your macros should appear before
+%% the \begin{document} command.
+%%
+\newcommand{\vdag}{(v)^\dagger}
+\newcommand\aastex{AAS\TeX}
+\newcommand\latex{La\TeX}
+
+%% Reintroduced the \received and \accepted commands from AASTeX v5.2
+%\received{July 1, 2016}
+%\revised{September 27, 2016}
+%\accepted{\today}
+%% Command to document which AAS Journal the manuscript was submitted to.
+%% Adds "Submitted to " the arguement.
+\submitjournal{ApJ}
+
+%% Mark up commands to limit the number of authors on the front page.
+%% Note that in AASTeX v6.1 a \collaboration call (see below) counts as
+%% an author in this case.
+%
+%\AuthorCollaborationLimit=3
+%
+%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
+%% collaboration" on the front page of this example manuscript.
+%%
+%% Note that all of the author will be shown in the published article.
+%% This feature is meant to be used prior to acceptance to make the
+%% front end of a long author article more manageable. Please do not use
+%% this functionality for manuscripts with less than 20 authors. Conversely,
+%% please do use this when the number of authors exceeds 40.
+%%
+%% Use \allauthors at the manuscript end to show the full author list.
+%% This command should only be used with \AuthorCollaborationLimit is used.
+
+%% The following command can be used to set the latex table counters.  It
+%% is needed in this document because it uses a mix of latex tabular and
+%% AASTeX deluxetables.  In general it should not be needed.
+%\setcounter{table}{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% The following section outlines numerous optional output that
+%% can be displayed in the front matter or as running meta-data.
+%%
+%% If you wish, you may supply running head information, although
+%% this information may be modified by the editorial offices.
+\shorttitle{The weakly nonlinear magnetorotational instability}
+\shortauthors{Clark \& Oishi}
+%%
+%% You can add a light gray and diagonal water-mark to the first page 
+%% with this command:
+% \watermark{text}
+%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
+%% long you can control the water-mark size with:
+%  \setwatermarkfontsize{dimension}
+%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% This is the end of the preamble.  Indicate the beginning of the
+%% manuscript itself with \begin{document}.
 
 \begin{document}
+
+%\title{The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor-Couette flow}
 \title{The Weakly Nonlinear Magnetorotational Instability in a Global, Cylindrical Taylor-Couette Flow}
-\author{S. E. Clark\altaffilmark{1}}
-\author{Jeffrey S. Oishi\altaffilmark{2,}\altaffilmark{3}}
-\altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
-\altaffiltext{2}{Department of Physics and Astronomy, Bates College}
-\altaffiltext{3}{Department of Astrophysics, American Museum of Natural History, New York, NY}
-
+
+%% LaTeX will automatically break titles if they run longer than
+%% one line. However, you may use \\ to force a line break if
+%% you desire. In v6.1 you can include a footnote in the title.
+
+%% A significant change from earlier AASTEX versions is in the structure for 
+%% calling author and affilations. The change was necessary to implement 
+%% autoindexing of affilations which prior was a manual process that could 
+%% easily be tedious in large author manuscripts.
+%%
+%% The \author command is the same as before except it now takes an optional
+%% arguement which is the 16 digit ORCID. The syntax is:
+%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
+%%
+%% This will hyperlink the author name to the author's ORCID page. Note that
+%% during compilation, LaTeX will do some limited checking of the format of
+%% the ID to make sure it is valid.
+%%
+%% Use \affiliation for affiliation information. The old \affil is now aliased
+%% to \affiliation. AASTeX v6.1 will automatically index these in the header.
+%% When a duplicate is found its index will be the same as its previous entry.
+%%
+%% Note that \altaffilmark and \altaffiltext have been removed and thus 
+%% can not be used to document secondary affiliations. If they are used latex
+%% will issue a specific error message and quit. Please use multiple 
+%% \affiliation calls for to document more than one affiliation.
+%%
+%% The new \altaffiliation can be used to indicate some secondary information
+%% such as fellowships. This command produces a non-numeric footnote that is
+%% set away from the numeric \affiliation footnotes.  NOTE that if an
+%% \altaffiliation command is used it must come BEFORE the \affiliation call,
+%% right after the \author command, in order to place the footnotes in
+%% the proper location.
+%%
+%% Use \email to set provide email addresses. Each \email will appear on its
+%% own line so you can put multiple email address in one \email call. A new
+%% \correspondingauthor command is available in V6.1 to identify the
+%% corresponding author of the manuscript. It is the author's responsibility
+%% to make sure this name is also in the author list.
+%%
+%% While authors can be grouped inside the same \author and \affiliation
+%% commands it is better to have a single author for each. This allows for
+%% one to exploit all the new benefits and should make book-keeping easier.
+%%
+%% If done correctly the peer review system will be able to
+%% automatically put the author and affiliation information from the manuscript
+%% and save the corresponding author the trouble of entering it by hand.
+
+\correspondingauthor{S. E. Clark}
+\email{seclark@astro.columbia.edu}
+
+\author[0000-0002-7633-3376]{S. E. Clark}
+\affil{Department of Astronomy, Columbia University, New York, NY}
+
+\author{Jeffrey S. Oishi}
+\affiliation{Department of Physics and Astronomy, Bates College, Lewiston, ME}
+\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY}
+\nocollaboration
+
+%% Note that the \and command from previous versions of AASTeX is now
+%% depreciated in this version as it is no longer necessary. AASTeX 
+%% automatically takes care of all commas and "and"s between authors names.
+
+%% AASTeX 6.1 has the new \collaboration and \nocollaboration commands to
+%% provide the collaboration status of a group of authors. These commands 
+%% can be used either before or after the list of corresponding authors. The
+%% argument for \collaboration is the collaboration identifier. Authors are
+%% encouraged to surround collaboration identifiers with ()s. The 
+%% \nocollaboration command takes no argument and exists to indicate that
+%% the nearby authors are not part of surrounding collaborations.
+
+%% Mark off the abstract in the ``abstract'' environment. 
 \begin{abstract}
@@
 
+%% Keywords should appear after the \end{abstract} command. 
+%% See the online documentation for the full list of available subject
+%% keywords and the rules for their use.
+%\keywords{editorials, notices --- miscellaneous --- catalogs --- surveys}
+
+%% From the front matter, we move on to the body of the paper.
+%% Sections are demarcated by \section and \subsection, respectively.
+%% Observe the use of the LaTeX \label
+%% command after the \subsection to give a symbolic KEY to the
+%% subsection for cross-referencing in a \ref command.
+%% You can use LaTeX's \ref and \label commands to keep track of
+%% cross-references to sections, equations, tables, and figures.
+%% That way, if you change the order of any elements, LaTeX will
+%% automatically renumber them.
+
+%% We recommend that authors also use the natbib \citep
+%% and \citet commands to identify citations.  The citations are
+%% tied to the reference list via symbolic KEYs. The KEY corresponds
+%% to the KEY in the \bibitem in the reference list below. 
 \section{Introduction}
@@
 The magnetorotational instability (MRI) is believed to drive angular momentum transport in astrophysical disks.
-The MRI is a local instability excited by weak magnetic fields in differentially rotating fluids, and since first applied to an astrophysical context \citep{Balbus:1991vs} it has been invoked to explain accretion in protoplanetary disks \citep{Armitage:2010} and disks around black holes \citep{Blaes:2014}, as well as jet and wind launching (\citei{Lesur:2013}), anisotropic turbulence \citep{Murphy:2015tn}, and dynamo generation \citep{Brandenburg:1995, Vishniac:2009il}.
+The MRI is a local instability excited by weak magnetic fields in differentially rotating fluids, and since first applied to an astrophysical context \citep{Balbus:1991vs} it has been invoked to explain accretion in protoplanetary disks \citep{Armitage:2010} and disks around black holes \citep{Blaes:2014}, as well as jet and wind launching \citep{Lesur:2013}, anisotropic turbulence \citep{Murphy:2015tn}, and dynamo generation \citep{Brandenburg:1995, Vishniac:2009il}.
 
@@
 These linear modes are exact solutions to the \textit{nonlinear} local MRI equations.
-The shearing box MRI system avoids runaway growth by a secondary instability of the channel modes themselves (\citei{Goodman:1994ul}, \citei{Pessah:2010ic}).
+The shearing box MRI system avoids runaway growth by a secondary instability of the channel modes themselves \citep{Goodman:1994ul, Pessah:2010ic}.
 The growth of parasitic modes provides a saturation avenue for channel mode-dominated flows, yet this is unlikely to be the dominant saturation mechanism in laboratory experiments or astrophysical disks, as channel modes are artificially over-represented in the shearing box \citep{Latter:2015}.
 Thus while the shearing box may accurately approximate many features of the global MRI, the saturation mechanism may not be among them.
+\edit1{In \citet[][hereafter Paper I]{Clark:2016} we find that the fastest-growing MRI mode in the shearing box is not a channel mode when the effects of ambipolar diffusion are formally included.
+It is thus important to ask whether the symmetries that give rise to the weakly nonlinear saturation in the local geometry are also manifested in the global flow.}
 
@@
 Unfortunately, detection of the MRI has so far proven elusive.
-\citei{Sisan:2004ig} claimed to detect the MRI in a spherical Couette flow, but most likely detected unrelated MHD instabilities instead \citep{Hollerbach:2009ig,Gissinger:2011td}.
+\citet{Sisan:2004ig} claimed to detect the MRI in a spherical Couette flow, but most likely detected unrelated MHD instabilities instead \citep{Hollerbach:2009ig,Gissinger:2011td}.
 Most relevant to our work is the Princeton Plasma Physics Laboratory (PPPL) MRI Experiment, a liquid gallium TC flow with an axial magnetic field \citep{Ji:2001kd}.
-There has been a significant amount of theoretical work designed to complement the Princeton MRI experiment involving direct numerical simulation of the experimental conditions, much of it focused on the specific challenges in identifying MRI signatures despite spurious, apparatus-driven flows (e.g.
-\citei{Gissinger:2012gc}).
+There has been a significant amount of theoretical work designed to complement the Princeton MRI experiment involving direct numerical simulation of the experimental conditions, much of it focused on the specific challenges in identifying MRI signatures despite spurious, apparatus-driven flows \citep[e.g.][]{Gissinger:2012gc}.
 The vertical endcaps on a laboratory MRI apparatus drive meridional flows which both inhibit the excitement of MRI and obscure its detection.
-The Princeton MRI experiment employs split, independently rotating endcaps to mitigate these flows (\citei{Schartman:2009df}).
+The Princeton MRI experiment employs split, independently rotating endcaps to mitigate these flows \citep{Schartman:2009df}.
 Our work assumes an infinite vertical domain, an idealization that is theoretically expedient but experimentally impractical.
@@
 \centering
-\includegraphics[width=\textwidth]{widegap_orders1_2_Pm16E-7}
+%\includegraphics[width=\textwidth]{widegap_orders1_2_Pm16E-7}
 \caption{Eigenfunctions of the first order equations, first order adjoint homogenous equations, and second order equations.
@@
 Cartesian.
-Previous investigations into the weakly nonlinear behavior of the MRI have used this narrow gap approximation \citep{Umurhan:2007dz,Umurhan:2007hs,Clark:2016}.Building on our work in \citet[][hereafter Paper I]{Clark:2016}, here we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.
+Previous investigations into the weakly nonlinear behavior of the MRI have used this narrow gap approximation \citep{Umurhan:2007dz,Umurhan:2007hs,Clark:2016}.
+Building on our work in \edit1{Paper 1}, here we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.
 
@@
 However the rotation profiles needed to excite helical MRI may be steeper than Keplerian, depending on the boundary conditions, and so its role in astrophysical disks is currently a matter of debate \citep{Liu:2006,Rudiger:2007,Kirillov:2013}.
-Regardless of its astrophysical role, the helical MRI is significantly easier to excite in a laboratory setting than the standard MRI, and has already been detected by the Potsdam Rossendorf Magnetic Instability Experiment (PROMISE; \citei{Stefani:2006iv}, \citei{Stefani:2009hp}).
+Regardless of its astrophysical role, the helical MRI is significantly easier to excite in a laboratory setting than the standard MRI, and has already been detected by the Potsdam Rossendorf Magnetic Instability Experiment \citep[PROMISE;][]{Stefani:2006iv,Stefani:2009hp}.
 
@@
 \centering
-\includegraphics[width=\textwidth]{widegap_n2_n3_Pm16E-7.pdf}
+%\includegraphics[width=\textwidth]{widegap_n2_n3_Pm16E-7.pdf}
 \caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.
@@
 In this work we will focus our findings on two fiducial parameter sets, one for the standard MRI where $\xi = 0$ and one for the helical MRI.
-We choose the standard MRI parameters to be comparable to the case considered in \citei{Goodman:2002ix}.
-The helical MRI parameters were chosen to be comparable to \citei{Hollerbach:2005tr}.
+We choose the standard MRI parameters to be comparable to the case considered in \citet{Goodman:2002ix}.
+The helical MRI parameters were chosen to be comparable to \citet{Hollerbach:2005tr}.
 Our fiducial parameters are described in Table \ref{table:parameters}.
 
-\begin{widetext}
 Our perturbed system is 
-
-\begin{align}
-& \frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r^2}B_0 \xi \partial_z B_\phi \nonumber\\
-& - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi - \frac{2}{r^2}\partial_r\Psi) - \frac{1}{r^3} \nabla^2 \Psi + \frac{2}{r^4}\partial_r\Psi\right] = N^{(\Psi)} \label{eq:Psi_perturbed}
-\end{align}
-
-\begin{align}
-& \partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} ( \nabla^2 \uphi - \frac{1}{r^2} \uphi ) = N^{(u)} \label{eq:uphi_perturbed} \\
-& \partial_t A - B_0 \partial_z \Psi - \frac{1}{\reym} ( \nabla^2 A - \frac{2}{r} \partial_r A )= N^{(A)}  \label{eq:A_perturbed}\\
+\onecolumngrid
+\beq\label{eq:Psi_perturbed}
+\frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r^2}B_0 \xi \partial_z B_\phi - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi - \frac{2}{r^2}\partial_r\Psi) - \frac{1}{r^3} \nabla^2 \Psi + \frac{2}{r^4}\partial_r\Psi\right] = N^{(\Psi)}
+\eeq
+
+\beq \label{eq:uphi_perturbed}
+\partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} ( \nabla^2 \uphi - \frac{1}{r^2} \uphi ) = N^{(u)}
+\eeq
+
+\beq
+\partial_t A - B_0 \partial_z \Psi - \frac{1}{\reym} (\nabla^2 A - \frac{2}{r} \partial_r A) = N^{(A)}  \label{eq:A_perturbed}
+\eeq
+
+\beq  
 \label{eq:Bphi_perturbed}
-& \partial_t B_\phi + \frac{1}{r^2} u_0 \partial_z A - B_0 \partial_z u_\phi - \frac{1}{r} \partial_r u_0 \partial_z A - \frac{2}{r^3} B_0 \xi \partial_z \Psi - \frac{1}{\reym} (\nabla^2 B_\phi - \frac{1}{r^2} B_\phi ) = N^{(B)}
-\end{align}
+\partial_t B_\phi + \frac{1}{r^2} u_0 \partial_z A - B_0 \partial_z u_\phi - \frac{1}{r} \partial_r u_0 \partial_z A - \frac{2}{r^3} B_0 \xi \partial_z \Psi - \frac{1}{\reym} (\nabla^2 B_\phi - \frac{1}{r^2} B_\phi ) = N^{(B)}
+\eeq
 
@@
 \eeq
-\end{widetext}
+\twocolumngrid
 
@@
 
-and $B = 0$ at $r = r_1, r_2$ (see \citei{Willis:2002bh}). Here, $I_n$ and $K_n$ are the modified Bessel functions of the first and second kind, respectively.
+and $B = 0$ at $r = r_1, r_2$ \citep[see][]{Willis:2002bh}. Here, $I_n$ and $K_n$ are the modified Bessel functions of the first and second kind, respectively.
 
@@
 
-Just as in the weakly nonlinear analyses of \citei{Umurhan:2007hs} and Paper I, we tune the system away from marginality by taking $B_0 \rightarrow B_0\left(1 - \epsilon^2\right)$, where the small parameter $\epsilon \ll 1$.
+Just as in the weakly nonlinear analyses of \citet{Umurhan:2007hs} and Paper I, we tune the system away from marginality by taking $B_0 \rightarrow B_0\left(1 - \epsilon^2\right)$, where the small parameter $\epsilon \ll 1$.
 We parameterize scale separation as $Z = \epsilon z$ and $T = \epsilon^2 t$, where $Z$ and $T$ are slowly varying spatial and temporal scales, respectively.
@@
 a Ginzburg-Landau equation (GLE).
-The GLE governs the weakly nonlinear amplitude behavior in a wide range of physical systems, including the narrow gap MRI (\citei{Umurhan:2007hs}), Rayleigh-B\'enard convection (\citei{Newell:1969wr}), and hydrodynamic TC flow (e.g.
-\citei{Recktenwald:1993}).
-We emphasize that this is a model equation, valid only near marginality (\citei{Cross:1993el}).
+The GLE governs the weakly nonlinear amplitude behavior in a wide range of physical systems, including the narrow gap MRI \citep{Umurhan:2007hs}, Rayleigh-B\'enard convection \citep{Newell:1969wr}, and hydrodynamic TC flow \citep[e.g.][]{Recktenwald:1993}.
+We emphasize that this is a model equation, valid only near marginality \citep{Cross:1993el}.
 The dynamics of the GLE are determined by its coefficients, which are in turn determined by the linear eigenfunctions and nonlinear vectors plotted in Figures \ref{fig:linear_eigenfunctions} and \ref{fig:N2_N3}.
@@
 The real GLE is subject to several well-studied instabilities, including the Ekhaus and Zig-Zag instabilities.
-When the coefficients are complex, we have the complex GLE, a source of even richer phase dynamics than its real counterpart (see \citei{Aranson:2002} for a thorough review).
+When the coefficients are complex, we have the complex GLE, a source of even richer phase dynamics than its real counterpart \citep[see][]{Aranson:2002} for a thorough review.
 
@@
 They find that this amplitude dependence is driven by the $\Pm^{1/3}$ dependence of the linear boundary layer.
-Boundary layer analysis similarly reveals a $\nu^{1/3}$ dependence for the radial extent of the boundary layer in TC flow (\citei{Goodman:2002ix}).
+Boundary layer analysis similarly reveals a $\nu^{1/3}$ dependence for the radial extent of the boundary layer in TC flow \citep{Goodman:2002ix}.
 Figure \ref{fig:thin_wide_comparison} shows the structure of the third-order nonlinear term $N_{31}^{(A)}$ as a function of $\Pm$ for both the narrow and wide gap standard MRI.
-$N_{31}$ is the vector that determines the GLE coefficient $c$, and thus the scaling of the saturation amplitude because of the insensitivity of $b$ to $\Pm$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs} and Paper I for the narrow gap equations).
+$N_{31}$ is the vector that determines the GLE coefficient $c$, and thus the scaling of the saturation amplitude because of the insensitivity of $b$ to $\Pm$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citet{Umurhan:2007hs} and Paper I for the narrow gap equations).
 Clearly, the width of the boundary layers scales with $\Pm$ in both the wide and narrow gap MRI.
@@
 The complex GLE may be unstable to traveling wave instabilities such as the Benjamin-Feir instability, a generalization of the Ekhaus instability.
-The complex GLE can also admit spatiotemporal chaos, and various classes of coherent structures (\citei{Aranson:2002}).
+The complex GLE can also admit spatiotemporal chaos, and various classes of coherent structures \citep{Aranson:2002}.
 Although a detailed description of the phase dynamics in the helical MRI is beyond the scope of this work, we note that such long-wavelength, long-timescale behavior may be observed in liquid metal helical MRI experiments.
@@
 Here we make a phenomenological comparison to two other systems that give rise to a GLE.
-The first and perhaps most famous is Rayleigh-B\'enard convection, in which a fluid between two plates is heated from below (\citei{Newell:1969wr}).
+The first and perhaps most famous is Rayleigh-B\'enard convection, in which a fluid between two plates is heated from below \citep{Newell:1969wr}.
 If we take the plane of the fluid to be infinite in the horizontal plane, the system is initially translationally symmetric.
@@
 Zonal flows are axisymmetric structures, large-scale and long-lived, which form spontaneously out of turbulence.
-They have recently been observed in some numerical studies of the MRI, and have generated considerable interest for their possible role in planet formation in protoplanetary disks (\citei{Johansen:2009uj}, \citei{Kunz:2013}).
+They have recently been observed in some numerical studies of the MRI, and have generated considerable interest for their possible role in planet formation in protoplanetary disks \citep{Johansen:2009uj,Kunz:2013}.
 The present work is of course an idealized geometry, and we make no attempt to model a realistic protoplanetary disk environment.
@@
 This work provides a mathematical description of the MRI as a pattern-forming process, but much remains to be understood, particularly involving the application of this model system to realistic astrophysical disks.
-Our model is most directly relevant to TC flows, and we emphasize that laboratory MRI experiments stand poised to observe the MRI-driven pattern formation predicted here.
+\edit1{Paper I establishes that the GLE will arise in the shearing box approximation in the presence of ambipolar diffusion, and this work demonstrates that the pattern-forming behavior is not an artifact of the local geometry.
+The stage is thus set to apply this theory to more astrophysical conditions in either the global geometry or a local approximation.
+Of course our current} model is most directly relevant to TC flows, and we emphasize that laboratory MRI experiments stand poised to observe the MRI-driven pattern formation predicted here.
 
@@
 \end{document}
+
+% End of file `sample61.tex'.

### 2017-03-12 python/plot_thingap_streamfuncs.py

@@ -238,7 +238,7 @@ def plot_both(ax, obj, type="Bfield", labels=True, oplot=True, background=False,
             Bbothmag = np.sqrt(np.abs(Vboth_Br1**2) + np.abs(Vboth_Bz1**2))
             su.streamplot(ax, x, z, Vboth_Br1.real, Vboth_Bz1.real, linewidth = 2*Bbothmag/Bbothmag.max(), color = "black")
 
-type = "velocity"
+type = "Bfield"
 labels = True
 oplot = True
 background = False
@@ -262,4 +262,4 @@ fig.subplots_adjust(wspace=0.6)
 #plt.savefig("../figures/thingap_streamfuncs_"+type+"_Pm_1E-3.png", dpi=fig.dpi)
 #plt.savefig("../figures/thingap_streamfuncs_"+type+"_Pm_1E-4.png", dpi=fig.dpi)
 
-plt.savefig('../paper/thingap_submit/thingap_streamfuncs_velocity_Pm_1E-3.eps', dpi=100)
+#plt.savefig('../paper/thingap_submit/thingap_streamfuncs_velocity_Pm_1E-3.eps', dpi=100)

### 2017-03-12 paper/thingap_ref_response/response_to_referee.tex
widegap revised draft for referee
many more additions to text

@@
 
-\textbf{In Paper II, the rotation law is enforced through viscous coupling to the boundaries. In Paper I, however, the rotation law is imposed directly at every point by the shear. }
+\textbf{In Paper II, the rotation law is enforced through viscous coupling to the boundaries.
+In Paper I, however, the rotation law is imposed directly at every point by the shear.
+We have noted this point in the text.
+See also the newly extended discussion of the local flow with shear periodic boundary conditions.}
 
 \item The influence of boundary conditions and boundary layers is strong.
-To call out some examples: In Paper I, the amplitude ?
-converges to the constant value (b/c)1/2, independent of z.
+To call out some examples: In Paper I, the amplitude $\alpha$ converges to the constant value $(b/c)^{1/2}$, independent of z.
 Is this really relevant to a lab experiment with vertical endcaps, or to a stratified disk?
@@
 
-\textbf{Yes, the saturated Bz is marginally stable, and we have added a sentence to that effect in the text. We cannot comment on dynamo action because our setup is axisymmetric. }
+\textbf{Yes, the saturated Bz is marginally stable, and we have added a sentence to that effect in the text.
+We cannot comment on dynamo action in either paper because of the symmetries in our setup.
+No dynamo can arise from an axisymmetric system (e.g.
+Cowling's anti-dynamo theorem).}
 

### 2017-03-12 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
many more additions to text

@@
 
-\subsection{\edit1{Shearing box and ambipolar diffusion}}
-
-Many studies of the MRI consider the instability in a shearing box, i.e.
+\subsection{\edit1{Shearing box and ambipolar diffusion}}\label{sec:SBAD}
+
+\edit1{Many studies of the MRI consider the instability in a shearing box, i.e.
 a wall-less local approximation that is meant to represent a small section of a disk.
@@
 This is analytically tracked for other plasma instabilities by tracking the amplitudes of growing, marginal, and damped eigenmodes simultaneously \citep[e.g.][]{Makwana:2011}.
-While the shearing box approximation allows the projection of the perturbed MRI equations into the basis set of linear eigenmodes, nonlinear coupling between modes will remain zero.
-
-The nonlinear property of primary MRI modes in the shearing box motivates the addition of radial boundaries, such that the nonlinear evolution of the weakly nonlinear MRI can be properly considered.
+While the shearing box approximation allows the projection of the perturbed MRI equations into the basis set of linear eigenmodes, nonlinear coupling between modes will remain zero.}
+
+\edit1{The nonlinear property of primary MRI modes in the shearing box motivates the addition of radial boundaries, such that the nonlinear evolution of the weakly nonlinear MRI can be properly considered.
 It also begs the question of whether some additional nonlinear mechanism can be introduced such that the fastest-growing modes are no longer nonlinear solutions to the shearing box equations.
 It has already been shown that the Hall effect does not negate the nonlinear property of primary MRI modes \citep{Kunz:2013}.
-However, it seems to have been overlooked in the literature that these linear modes are \textit{not} solutions of the nonlinear ambipolar diffusion term, which is proportional to
+However, it seems to have been overlooked in the literature that these linear modes are \textit{not} solutions of the nonlinear ambipolar diffusion term, which is proportional to}
 
@@
 
-Furthermore, the radial wavenumber of the fastest-growing linear MRI mode in a shearing box with ambipolar diffusion is nonzero when a constant azimuthal background field is considered in addition to an axial one \citep{Kunz:2004ib}.
+\edit1{Furthermore, the radial wavenumber of the fastest-growing linear MRI mode in a shearing box with ambipolar diffusion is nonzero when a constant azimuthal background field is considered in addition to an axial one \citep{Kunz:2004ib}.
 This means that, in the presence of ambipolar diffusion, we can derive the weakly nonlinear envelope equation for the MRI in the shearing box.
@@
 The constant azimuthal background field component does not contribute to any other terms in the local MRI equations.
-Thus, the slow-scale evolution of the MRI in a shearing box with ambipolar diffusion is also governed by a Ginzburg-Landau equation.
-
-The Ginzburg-Landau form of the amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to the wavenumber \citep{Hoyle:2006}.
+Thus, the slow-scale evolution of the MRI in a shearing box with ambipolar diffusion is also governed by a Ginzburg-Landau equation.}
+
+\edit1{The Ginzburg-Landau form of the amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to the wavenumber \citep{Hoyle:2006}.
 In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane, and the quadratic maximum is a consequence of the linear dispersion relation given in Appendix~\ref{app:dispersion}.
@@
 Both constructions require the most unstable mode to have nonconstant radial structure.
-Physically, this radial variation impedes the free exchange of angular momentum facilitated by the uniform stretching of channel modes.
-%So long as our system permits the most unstable mode to nonlinearly interact with
+Physically, this radial variation impedes the free exchange of angular momentum facilitated by the uniform stretching of channel modes.}  %So long as our system permits the most unstable mode to nonlinearly interact with
 
@@
 
+\edit1{The weakly nonlinear theory predicts that the amplitude of the system is bounded by the saturation amplitude $\alpha_s = \sqrt{b/c}$, where $b$ and $c$ are coefficients corresponding to the linear growth term and nonlinear term of the Ginzburg-Landau equation, respectively.
+The coefficient $b$ comes from the interaction between the background magnetic field and the linear MRI solution.
+The coefficient $c$ describes the third-order nonlinear interaction between terms in the perturbation series.
+Physically, we see that the saturation amplitude is controlled by the strength of the mode interaction within our finite band of unstable modes.
+We stress that while the third-order nonlinear terms in the walled TC flow are strongly influenced by the boundary layers, this is not generically true of the MRI system.
+Indeed, in the shearing box MRI with ambipolar diffusion (the case sketched out in Section \ref{sec:SBAD}), boundary layers impossible in the shear periodic flow.
+In this case the third-order nonlinear behavior of the system includes three-mode interactions from the cubic nonlinearity in the ambipolar diffusion term.}
+
 \section{Conclusion}

### 2017-03-13 paper/thingap_ref_response/response_to_referee.tex
for numerical convergence plot

@@
 
+\textbf{The amplitude $\alpha$ is bounded by the value +/- $(b/c)^{1/2}$, although the amplitude does not necessarily converge to a constant value, as shown in Figures 6 and 7.
+We note, however, that although the value of the quantity $(b/c)^{1/2}$ is heavily influenced by the boundary layers in the walled TC flow, this is specific to the wall-like boundary conditions and not inherent to the theory.
+We have added a paragraph at the end of section 4 to elaborate on this point, discuss what the $(b/c)^{1/2}$ saturation amplitude means physically, and describe how the derived amplitude applies to the shearing box MRI.}
+
 \item Both papers are light on physical interpretation. This is especially an issue with Paper I -- see below. I would particularly like to see:

### 2017-03-13 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
for numerical convergence plot
replaced beta with Co

@@
 \newcommand{\Pm}{\mathrm{Pm}}
+\newcommand{\Co}{\mathrm{Co}}
 
@@
 Analytical investigation of this case reveals that MRI saturation can be caused by parasitic Kelvin-Helmholtz and tearing mode instabilities, depending on parameter regime \citep{Pessah:2010ic}.
-The theory of MRI channel mode parasites is robust \citep[see also]{Pessah:2009gm;Latter:2010iz;Rembiasz:2016}, but their importance may be overestimated by the local approximation \citep{Latter:2015}, and not germane to global analyses like the one presented here.
+The theory of MRI channel mode parasites is robust \citep[see also][]{Pessah:2009gm,Latter:2010iz,Rembiasz:2016}, but their importance may be overestimated by the local approximation \citep{Latter:2015}, and not germane to global analyses like the one presented here.
 \citet{Latter:2015} gives a detailed analysis of the relation between local and global linear MRI modes.
@@
 We then place our results in the ocntext of previous results from both analytic and computational studies in section~\ref{sec:discussion} and draw conclusions in section~\ref{sec:conclusion}.
-
-\section{Equations}\label{sec:equations}
 
@@
 
+\section{Equations}\label{sec:equations}
+
 The evolution of a conducting fluid is governed by the momentum and induction equations,
@@
 We nondimensionalize the equations, with lengths nondimensionalized by $L$, time by $\Omega_0$, velocities by $\Omega_0 L$, magnetic fields by $B_0$, and pressure by $\Omega_0^2 L^2 \rho_0$, where $L$ is the channel width, $\Omega_0$ is the rotation rate at the center of the channel, and $\rho_0$ is the constant pressure in the base state (see Figure \ref{fig:setup}).
-We define the Reynolds number, $\reye \equiv {\Omega_0 L^2}/{\nu}$, magnetic Reynolds number, $\reym \equiv {\Omega_0 L^2}/{\eta}$, and a plasma beta parameter, $\beta \equiv {\Omega_0^2r_0^2}/{v_A^2}$, where the Alfv\'en speed $v_A$ is $v_A^2 = {B_0^2}/{\rho_0}$.
+We define the Reynolds number, $\reye \equiv {\Omega_0 L^2}/{\nu}$, magnetic Reynolds number, $\reym \equiv {\Omega_0 L^2}/{\eta}$, and Cowling number, $\Co \equiv {2 v_A^2}/{\Omega_0^2r_0^2}$, where the Alfv\'en speed $v_A$ is $v_A^2 = {B_0^2}/{\rho_0}$.
+The fluid symbols $\mathbf{u}$, $\mathbf{B}$, etc.
+will henceforth be used to refer to the nondimensional, perturbed quantities.
+  %a plasma beta parameter, $\beta \equiv {\Omega_0^2r_0^2}/{v_A^2}$, where the Alfv\'en speed $v_A$ is $v_A^2 = {B_0^2}/{\rho_0}$.
 The fluid symbols $\mathbf{u}$, $\mathbf{B}$, etc.
@@
 \label{eqset1}
-\partial_t \nabla^2 \Psi \, - \, 2 \partial_z u_{y} \, - \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, - \, \frac{1}{\reye}\nabla^4 \Psi \, = \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, - \, J\left(\Psi, \nabla^2 \Psi\right) \,
+\partial_t \nabla^2 \Psi \, - \, 2 \partial_z u_{y} \, - \, \Co B_0 \partial_z \nabla^2 A \, - \, \frac{1}{\reye}\nabla^4 \Psi \, = \, \Co J\left(A, \nabla^2 A \right) \, - \, J\left(\Psi, \nabla^2 \Psi\right) \,
 \eeq
@@
 \label{eqset2}
-\partial_t u_{y} \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \, \frac{2}{\beta}B_0\partial_z B_{y} \, + \, \frac{2}{\beta} J\left(A, B_{y}\right) \, - \, J\left(\Psi, u_{y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{y}
+\partial_t u_{y} \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \, \Co B_0\partial_z B_{y} \, + \, \Co J\left(A, B_{y}\right) \, - \, J\left(\Psi, u_{y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{y}
 \eeq
@@
 We find the marginal state, where the most unstable linear MRI mode neither grows nor decays, for a set of dimensionless parameters, and then destabilize the system.
-We examine the system for fiducial parameters comparable to URM07, namely $\Pm = 1.0 \times 10^{-3}$, $\beta = 25$, $q = 1.5$.
+We examine the system for fiducial parameters comparable to URM07, namely $\Pm = 1.0 \times 10^{-3}$, $\Co = 0.08$, $q = 1.5$.
 The system is marginal for a critical wavenumber $k_c = 0.75$ and a critical magnetic Reynolds number $\reym_c = 4.9$.
@@
 
-\section{A. Detailed Equations}\label{app:matrices}
+\section{Detailed Equations}\label{app:matrices}
 
@@
 \mathcal{L}_1 = \left[\begin{matrix}
-0 & -2 & -\frac{2}{\beta}\partial_x^2 & 0 \\
-(2-q)\Omega_0 & 0 & 0 & -\frac{2}{\beta} \\
+0 & -2 & -\Co\partial_x^2 & 0 \\
+(2-q)\Omega_0 & 0 & 0 & -\Co \\
 -1 & 0 & 0 & 0 \\
@@
 \mathcal{L}_3 = \left[\begin{matrix}
-0 & 0 & -\frac{2}{\beta} & 0 \\
+0 & 0 & -\Co & 0 \\
 0 & 0 & 0 & 0 \\
@@
 \mathcal{G} = \left[\begin{matrix}
-0 & 0 & -\frac{2}{\beta}\partial_x^2 & 0 \\
-0 & 0 & 0 & -\frac{2}{\beta} \\
+0 & 0 & -\Co\partial_x^2 & 0 \\
+0 & 0 & 0 & -\Co \\
 -1 & 0 & 0 & 0 \\
@@
 
-\section{B. Expansion of Nonlinear Terms}\label{app:nonlinear_terms}
+\section{Expansion of Nonlinear Terms}\label{app:nonlinear_terms}
 
@@
 \beq
-N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \frac{2}{\beta}J(A_1, \nabla^2 A_1) 
-\eeq
-
-\beq
-N_2^{(u)} = J(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_1) 
+N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \Co J(A_1, \nabla^2 A_1) 
+\eeq
+
+\beq
+N_2^{(u)} = J(\Psi_1, u_1) \, - \, \Co J(A_1, B_1) 
 \eeq
@@
 \begin{split}
-N_3^{(\Psi)} & = J(\Psi_1, \nabla^2\Psi_2) \, - \, \frac{2}{\beta}J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) - \, \frac{2}{\beta}J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, \\
-& - \, 2 \frac{2}{\beta}J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \frac{2}{\beta}\widetilde{J}(A_1, \nabla^2 A_1)\\
+N_3^{(\Psi)} & = J(\Psi_1, \nabla^2\Psi_2) \, - \, \Co J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) - \, \Co J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, \\
+& - \, 2 \Co J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \Co \widetilde{J}(A_1, \nabla^2 A_1)\\
 \end{split}
@@
 \beq
-N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \frac{2}{\beta} J(A_1, B_2) \, - \, \frac{2}{\beta} J(A_2, B_1) \, - \, \frac{2}{\beta} \widetilde{J}(A_1, B_1)
+N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \Co J(A_1, B_2) \, - \, \Co J(A_2, B_1) \, - \, \Co \widetilde{J}(A_1, B_1)
 \eeq
@@
 N_{22}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11} - k_c^2 \partial_x \Psi_{11}\right) - \partial_x \Psi_{11} \cdot \left(i k_c \partial_x^2 \Psi_{11} - i k_c^3 \Psi_{11}\right) \\
-& + \frac{2}{\beta}\partial_x A_{11} \cdot \left(i k_c \partial_x^2 A_{11} - i k_c^3 A_{11}\right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11} - k_c^2 \partial_x A_{11}\right)
+& + \Co \partial_x A_{11} \cdot \left(i k_c \partial_x^2 A_{11} - i k_c^3 A_{11}\right) - \Co i k_c A_{11} \cdot \left(\partial_x^3 A_{11} - k_c^2 \partial_x A_{11}\right)
 \end{split}
@@
 \begin{split}
-N_{22}^{(u)} = & \, i k_c \Psi_{11} \cdot \partial_x u_{11} - \partial_x \Psi_{11} \cdot i k_c u_{11} - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11} + \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}
+N_{22}^{(u)} = & \, i k_c \Psi_{11} \cdot \partial_x u_{11} - \partial_x \Psi_{11} \cdot i k_c u_{11} - \Co i k_c A_{11} \cdot \partial_x B_{11} + \Co \partial_x A_{11} \cdot i k_c B_{11}
 \end{split}
@@
 N_{20}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - k_c^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i k_c^3 \Psi_{11}^* - i k_c \partial_x^2 \Psi_{11}^*\right) \\
-& +\frac{2}{\beta} \partial_x A_{11} \cdot \left(i k_c^3 A_{11}^* - i k_c \partial_x^2 A_{11}^* \right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
+& + \Co \partial_x A_{11} \cdot \left(i k_c^3 A_{11}^* - i k_c \partial_x^2 A_{11}^* \right) - \Co i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
 \end{split}
@@
 \beq
-N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* + \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11}^* - \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}^*
+N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* + \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \Co i k_c A_{11} \cdot \partial_x B_{11}^* - \Co \partial_x A_{11} \cdot i k_c B_{11}^*
 \eeq
@@
 N_{31}^{(\Psi)} = &  \, i k_c \left(\Psi_{11} \cdot \partial_x^3 \Psi_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x^3\Psi_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x^3 \Psi_{22}\right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot \partial_x^2 \Psi_{22}\right) \\
-& + i 8 k_c^3 \left(\partial_x\Psi_{11}^* \cdot \Psi_{22}\right) + i 4 k_c^3 \left(\Psi_{11}^* \cdot \partial_x \Psi_{22}\right) + \frac{2}{\beta}\left[ - i k_c \left(A_{11} \cdot \partial^3 A_{20}\right) - i k_c \left(A_{11} \cdot \partial_x^3 A_{20}^*\right) \right]\\
-& + \frac{2}{\beta}\left[i k_c \left(A_{11}^* \cdot \partial_x^3 A_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \partial_x^2 A_{22}\right) - i 8 k_c^3 \left(\partial_x A_{11}^* \cdot A_{22}\right) - i 4 k_c^3 \left(A_{11}^* \cdot \partial_x A_{22}\right)\right] \\
+& + i 8 k_c^3 \left(\partial_x\Psi_{11}^* \cdot \Psi_{22}\right) + i 4 k_c^3 \left(\Psi_{11}^* \cdot \partial_x \Psi_{22}\right) + \Co \left[ - i k_c \left(A_{11} \cdot \partial^3 A_{20}\right) - i k_c \left(A_{11} \cdot \partial_x^3 A_{20}^*\right) \right]\\
+& + \Co \left[i k_c \left(A_{11}^* \cdot \partial_x^3 A_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \partial_x^2 A_{22}\right) - i 8 k_c^3 \left(\partial_x A_{11}^* \cdot A_{22}\right) - i 4 k_c^3 \left(A_{11}^* \cdot \partial_x A_{22}\right)\right] \\
 & + i 2 k_c \left(\Psi_{22} \cdot \partial_x^3 \Psi_{11}^*\right) - i 2 k_c^3\left(\Psi_{22} \cdot \partial_x \Psi_{11}^* \right) - i k_c \left(\partial_x \Psi_{20} \cdot \partial_x^2 \Psi_{11}\right) + i k_c \left(\partial_x \Psi_{22} \cdot \partial_x^2 \Psi_{11}^*\right) \\
 & - i k_c \left(\partial_x \Psi_{20}^* \cdot \partial_x^2 \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20} \cdot \Psi_{11}\right) + i k_c^3 \left(\partial_x \Psi_{20}^* \cdot \Psi_{11}\right) - i k_c^3 \left(\partial_x \Psi_{22} \cdot \Psi_{11}^*\right) \\
-& + \frac{2}{\beta} \left[ - i 2 k_c \left(A_{22} \cdot \partial_x^3 A_{11}^*\right) + i 2 k_c^3 \left(A_{22} \cdot \partial_x A_{11}^*\right) + i k_c \left(\partial_x A_{20} \cdot \partial_x^2 A_{11}\right) - i k_c \left(\partial_x A_{22} \cdot \partial_x^2 A_{11}^*\right) \right] \\
-& + \frac{2}{\beta} \left[ i k_c \left(\partial_x A_{20}^* \cdot \partial_x^2 A_{11}\right) - i k_c^3 \left(\partial_x A_{20} \cdot A_{11}\right) - i k_c^3 \left(\partial_x A_{20}^* \cdot A_{11}\right) + i k_c^3 \left(\partial_x A_{22} \cdot A_{11}^*\right)\right]
+& + \Co \left[ - i 2 k_c \left(A_{22} \cdot \partial_x^3 A_{11}^*\right) + i 2 k_c^3 \left(A_{22} \cdot \partial_x A_{11}^*\right) + i k_c \left(\partial_x A_{20} \cdot \partial_x^2 A_{11}\right) - i k_c \left(\partial_x A_{22} \cdot \partial_x^2 A_{11}^*\right) \right] \\
+& + \Co \left[ i k_c \left(\partial_x A_{20}^* \cdot \partial_x^2 A_{11}\right) - i k_c^3 \left(\partial_x A_{20} \cdot A_{11}\right) - i k_c^3 \left(\partial_x A_{20}^* \cdot A_{11}\right) + i k_c^3 \left(\partial_x A_{22} \cdot A_{11}^*\right)\right]
 \end{split}
@@
 & - i k_c \left(u_{11} \cdot \partial_x \Psi_{20} \right) - i k_c \left(u_{11} \cdot \partial_x \Psi_{20}^*\right) + i k_c \left(u_{11}^* \cdot \partial_x \Psi_{22} \right) + i 2 k_c \left(\partial_x u_{11}^* \cdot \Psi_{22}\right) \\
-& +\frac{2}{\beta}\left[- i k_c \left(A_{11} \cdot \partial_x B_{20}\right) - i k_c \left(A_{11} \cdot \partial_x B_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x B_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot B_{22}\right)\right] \\
-& +\frac{2}{\beta}\left[ i k_c \left(B_{11} \cdot \partial_x A_{20}\right) + i k_c \left(B_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(B_{11}^* \cdot \partial_x A_{20}\right) - i 2 k_c \left(\partial_x B_{11}^* \cdot A_{22}\right)\right]
+& +\Co \left[- i k_c \left(A_{11} \cdot \partial_x B_{20}\right) - i k_c \left(A_{11} \cdot \partial_x B_{20}^*\right) + i k_c \left(A_{11}^* \cdot \partial_x B_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot B_{22}\right)\right] \\
+& + \Co \left[ i k_c \left(B_{11} \cdot \partial_x A_{20}\right) + i k_c \left(B_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(B_{11}^* \cdot \partial_x A_{20}\right) - i 2 k_c \left(\partial_x B_{11}^* \cdot A_{22}\right)\right]
 \end{split}
@@
 
-\section{C. Linear dispersion relation}\label{app:dispersion}
+\section{Linear dispersion relation}\label{app:dispersion}
 
@@
 
-\appendix
+%\appendix
 \begin{thebibliography}{}

### 2017-03-13 paper/widegap_ref_response/widegap_mri_paper_ref_response_apj.tex
many more additions to text
for numerical convergence plot
replaced beta with Co in widegap paper too

@@
   \hline
- & $\xi$ & $\Pm$ & $\beta$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ & radial magnetic b.c. \\ 
+ & $\xi$ & $\Pm$ & $\Co$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ & radial magnetic b.c. \\ 
   \hline\hline
-Standard MRI & 0 & 1.6E-6 & 41.2 & 0.121 & 0.33 & conducting \\ 
-Helical MRI & 4 & 1E-6 & 1.7E-2 & 0.27 & 0.5 & insulating\\ 
+%Standard MRI & 0 & 1.6E-6 & 41.2 & 0.121 & 0.33 & conducting \\ 
+Standard MRI & 0 & 1.6E-6 & 4.85E-2 & 0.121 & 0.33 & conducting \\ 
+Helical MRI & 4 & 1E-6 & 118 & 0.27 & 0.5 & insulating\\ 
    \hline
@@
 \centering
-%\includegraphics[width=\textwidth]{widegap_orders1_2_Pm16E-7}
+\includegraphics[width=\textwidth]{widegap_orders1_2_Pm16E-7}
 \caption{Eigenfunctions of the first order equations, first order adjoint homogenous equations, and second order equations.
@@
 \centering
-%\includegraphics[width=\textwidth]{widegap_n2_n3_Pm16E-7.pdf}
+\includegraphics[width=\textwidth]{widegap_n2_n3_Pm16E-7.pdf}
 \caption{Nonlinear terms $N_2$ and $N_3$ for our fiducial standard MRI parameters.
@@
 \beq\label{eq:Psi_perturbed}
-\frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r^2}B_0 \xi \partial_z B_\phi - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi - \frac{2}{r^2}\partial_r\Psi) - \frac{1}{r^3} \nabla^2 \Psi + \frac{2}{r^4}\partial_r\Psi\right] = N^{(\Psi)}
+\frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \Co \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \Co \frac{2}{r^2}B_0 \xi \partial_z B_\phi - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi - \frac{2}{r^2}\partial_r\Psi) - \frac{1}{r^3} \nabla^2 \Psi + \frac{2}{r^4}\partial_r\Psi\right] = N^{(\Psi)}
 \eeq
@@
 \beq \label{eq:uphi_perturbed}
-\partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \frac{2}{\beta} B_0 \partial_z B_\phi - \frac{1}{\reye} ( \nabla^2 \uphi - \frac{1}{r^2} \uphi ) = N^{(u)}
+\partial_t \uphi + \frac{1}{r^2} u_0 \partial_z \Psi + \frac{1}{r} \partial_r u_0 \partial_z \Psi - \Co B_0 \partial_z B_\phi - \frac{1}{\reye} ( \nabla^2 \uphi - \frac{1}{r^2} \uphi ) = N^{(u)}
 \eeq
@@
 \beq
-N^{(\Psi)} = - J(\Psi, \frac{1}{r^2} ( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi) ) + \frac{2}{\beta} J(A, \frac{1}{r^2} ( \nabla^2 A - \frac{2}{r} \partial_rA) ) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi  + \frac{2}{r} u_\phi \partial_z u_\phi
-\eeq
-\beq
-N^{(u)} = \frac{2}{\beta} \frac{1}{r} J(A, B_\phi) - \frac{1}{r} J(\Psi, \uphi) + \frac{2}{\beta}\frac{1}{r^2} B_\phi \partial_z A - \frac{1}{r^2} \uphi \partial_z \Psi 
+N^{(\Psi)} = - J(\Psi, \frac{1}{r^2} ( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi) ) + \Co J(A, \frac{1}{r^2} ( \nabla^2 A - \frac{2}{r} \partial_rA) ) - \Co \frac{2}{r}B_\phi \partial_z B_\phi  + \frac{2}{r} u_\phi \partial_z u_\phi
+\eeq
+\beq
+N^{(u)} = \Co \frac{1}{r} J(A, B_\phi) - \frac{1}{r} J(\Psi, \uphi) + \Co \frac{1}{r^2} B_\phi \partial_z A - \frac{1}{r^2} \uphi \partial_z \Psi 
 \eeq
@@
 $\Omega_1 = \Omega(r_1)$ is the rotation rate of the inner cylinder.
-We introduce the Reynolds number $\reye = \Omega_1 r_1^2/\nu$, the magnetic Reynolds number $\reym = \Omega_1 r_1^2 / \eta$, and a plasma beta parameter $\beta = \Omega_1^2 r_1^2 \rho_0/B_0^2$.
+We introduce the Reynolds number $\reye = \Omega_1 r_1^2/\nu$, the magnetic Reynolds number $\reym = \Omega_1 r_1^2 / \eta$, and a plasma beta parameter $\Co = 2 B_0^2 / \Omega_1^2 r_1^2 \rho_0$.
+%$\beta = \Omega_1^2 r_1^2 \rho_0/B_0^2$.
 Note that if we define the dimensional cylindrical coordinate $r = r_1(1 + \delta x)$, we recover the narrow gap approximation of the system in the limit $\delta \rightarrow 0$.
@@
 We find the marginal system as a function of the dimensionless parameters.
-The marginal stability curve for our standard MRI system is a hyperplane in $(\reym, \Pm, \beta, \Omega_2/\Omega_1, \mathrm{R}_1/\mathrm{R}_2$), but we hold all of these constant except for $\reym$.
+The marginal stability curve for our standard MRI system is a hyperplane in $(\reym, \Pm, \Co, \Omega_2/\Omega_1, \mathrm{R}_1/\mathrm{R}_2$), but we hold all of these constant except for $\reym$.
 To analyze the MRI system at marginality, we fix the parameters listed in Table \ref{table:parameters} and determine the critical $\reym$ and vertical wavenumber $k_z$ by repeatedly solving the linear MRI system to determine the smallest parameter values for which the fastest growing mode has zero growth rate.
@@
 \includegraphics[width=\columnwidth]{widegap_totalstress_intstress-eps-converted-to.pdf}
-\caption{Total stress $\mathrm{T}(r, z) = u_r u_\phi - \frac{2}{\beta} B_r B_\phi$, i.e.
+\caption{Total stress $\mathrm{T}(r, z) = u_r u_\phi - \Co B_r B_\phi$, i.e.
 the sum of the Reynolds and Maxwell stresses for the fiducial standard MRI parameters (top panel).
@@
 
-\section{A. Detailed Equations}\label{app:basic_equations}
+\section{Detailed Equations}\label{app:basic_equations}
 
@@
 \mathcal{L}_1 = \left[\begin{matrix}
-0 & -\frac{2}{r} u_0 & \frac{2}{\beta} (\frac{1}{r^2} \partial_r - \frac{1}{r}\partial_r^2) & 0 \\
-\frac{1}{r^2} u_0 + \frac{1}{r}\partial_r u_0 & 0 & 0 & -\frac{2}{\beta} \\
+0 & -\frac{2}{r} u_0 & \Co (\frac{1}{r^2} \partial_r - \frac{1}{r}\partial_r^2) & 0 \\
+\frac{1}{r^2} u_0 + \frac{1}{r}\partial_r u_0 & 0 & 0 & -\Co \\
 -1 & 0 & 0 & 0 \\
@@
 \mathcal{L}_3 = \left[\begin{matrix}
-0 & 0 & -\frac{2}{\beta}\frac{1}{r} & 0 \\
+0 & 0 & -\Co \frac{1}{r} & 0 \\
 0 & 0 & 0 & 0 \\
@@
 \mathcal{G} = \left[\begin{matrix}
-0 & 0 & \frac{2}{\beta}(\frac{1}{r^2}\partial_r - \frac{1}{r}\partial_r^2) & 0 \\
-0 & 0 & 0 & -\frac{2}{\beta} \\
+0 & 0 & \Co (\frac{1}{r^2}\partial_r - \frac{1}{r}\partial_r^2) & 0 \\
+0 & 0 & 0 & -\Co \\
 -1 & 0 & 0 & 0 \\
@@
 \mathcal{H} = \left[\begin{matrix}
-0 & 0 & 0 & \frac{2}{\beta} \frac{2}{r^2} \\
+0 & 0 & 0 & \Co \frac{2}{r^2} \\
 0 & 0 & 0 & 0 \\
@@
 -\frac{2}{r}u_0 & 0 & 0 & -1 \\
-\frac{2}{\beta}(\frac{1}{r^3} - \frac{1}{r^2}\partial_r - \frac{1}{r}\partial_r^2) & 0 & 0 & \frac{1}{r^2}u_0 - \frac{1}{r}\partial_r u_0 \\
-0 & -\frac{2}{\beta} & 0 & 0 \\
+\Co (\frac{1}{r^3} - \frac{1}{r^2}\partial_r - \frac{1}{r}\partial_r^2) & 0 & 0 & \frac{1}{r^2}u_0 - \frac{1}{r}\partial_r u_0 \\
+0 & -\Co & 0 & 0 \\
 \end{matrix}\right],
@@
 %%%%% NONLINEAR TERMS %%%%%%%%%%%
-\section{B. Nonlinear Terms}\label{app:nonlinear_terms}
+\section{Nonlinear Terms}\label{app:nonlinear_terms}
 
@@
 N_2^{\Psi}  = -J(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1) - J(\Psi_1, -\frac{2}{r^3}\partial_r\Psi_1)
-+ \frac{2}{\beta} J (A_1, \frac{1}{r^2} \nabla^2 A_1) + \frac{2}{\beta} J(A_1, -\frac{2}{r^3} \partial_r A_1) + \frac{2}{r} u_1 \partial_z u_1 - \frac{2}{\beta} \frac{2}{r} B_1 \partial_z B_1
++ \Co J (A_1, \frac{1}{r^2} \nabla^2 A_1) + \Co J(A_1, -\frac{2}{r^3} \partial_r A_1) + \frac{2}{r} u_1 \partial_z u_1 - \Co \frac{2}{r} B_1 \partial_z B_1
 \eeq
@@
 \begin{split}
-N_2^{u} = -\frac{1}{r} J\left(\Psi_1, u_1\right) + \frac{1}{r} \frac{2}{\beta} J\left(A_1, B_1\right) - \frac{1}{r^2} u_1 \partial_z \Psi_1 + \frac{2}{\beta}\frac{1}{r^2} B_1 \partial_z A_1
+N_2^{u} = -\frac{1}{r} J\left(\Psi_1, u_1\right) + \frac{1}{r} \Co J\left(A_1, B_1\right) - \frac{1}{r^2} u_1 \partial_z \Psi_1 + \Co \frac{1}{r^2} B_1 \partial_z A_1
 \end{split}
@@
 N_3^{\Psi} & = -J(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_2) - J(\Psi_2, \frac{1}{r^2} \nabla^2\Psi_1) - 2 J (\Psi_1, \frac{1}{r^2}\partial_Z\partial_z \Psi_1) - J(\Psi_1, -\frac{2}{r^3}\partial_r \Psi_2) - J(\Psi_2, -\frac{2}{r^3}\partial_r \Psi_1) \\
-& - \widetilde{J}(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1) - \widetilde{J} (\Psi_1, -\frac{2}{r^3}\partial_r \Psi_1) + \frac{2}{\beta} J(A_1, \frac{1}{r^2}\nabla^2 A_2) + \frac{2}{\beta} J(A_2, \frac{1}{r^2}\nabla^2 A_1) + \frac{4}{\beta} J(A_1, \frac{1}{r^2}\partial_Z\partial_z A_1) \\
-& + \frac{2}{\beta} J(A_1, -\frac{2}{r^3} \partial_r A_2 ) + \frac{2}{\beta} J(A_2, -\frac{2}{r^3} \partial_r A_1) + \frac{2}{\beta} \widetilde{J} (A_1, \frac{1}{r^2} \nabla^2 A_1) + \frac{2}{\beta} \widetilde{J} (A_1, -\frac{2}{r^3} \partial_r A_1) \\
-& + \frac{2}{r} u_1 \partial_z u_2 + \frac{2}{r} u_2 \partial_z u_1 + \frac{2}{r} u_1 \partial_Z u_1 - \frac{2}{\beta}\frac{2}{r} B_1\partial_z B_2 - \frac{2}{\beta}\frac{2}{r} B_2 \partial_z B_1 - \frac{2}{\beta} \frac{2}{r} B_1 \partial_Z B_1
+& - \widetilde{J}(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1) - \widetilde{J} (\Psi_1, -\frac{2}{r^3}\partial_r \Psi_1) + \Co J(A_1, \frac{1}{r^2}\nabla^2 A_2) + \Co J(A_2, \frac{1}{r^2}\nabla^2 A_1) + 2\Co J(A_1, \frac{1}{r^2}\partial_Z\partial_z A_1) \\
+& + \Co J(A_1, -\frac{2}{r^3} \partial_r A_2 ) + \Co J(A_2, -\frac{2}{r^3} \partial_r A_1) + \Co \widetilde{J} (A_1, \frac{1}{r^2} \nabla^2 A_1) + \Co \widetilde{J} (A_1, -\frac{2}{r^3} \partial_r A_1) \\
+& + \frac{2}{r} u_1 \partial_z u_2 + \frac{2}{r} u_2 \partial_z u_1 + \frac{2}{r} u_1 \partial_Z u_1 - \Co \frac{2}{r} B_1\partial_z B_2 - \Co \frac{2}{r} B_2 \partial_z B_1 - \Co \frac{2}{r} B_1 \partial_Z B_1
 \end{split}
@@
 \begin{split}
-N_3^u & = -\frac{1}{r}J\left(\Psi_1, u_2\right) - \frac{1}{r}J\left(\Psi_2, u_1\right) - \frac{1}{r}\widetilde{J} \left(\Psi_1, u_1\right) + \frac{1}{r}\frac{2}{\beta} J\left(A_1, B_2\right) + \frac{1}{r} \frac{2}{\beta} J\left(A_2, B_1\right) + \frac{1}{r}\frac{2}{\beta}\widetilde{J}\left(A_1, B_1\right) \\
-& - \frac{1}{r^2} u_1\partial_z \Psi_2 - \frac{1}{r^2} u_2 \partial_z \Psi_1 - \frac{1}{r^2} u_1 \partial_Z \Psi_1 + \frac{2}{\beta} \frac{1}{r^2} B_1 \partial_z A_2 + \frac{2}{\beta} \frac{1}{r^2} B_2 \partial_z A_1 + \frac{2}{\beta} \frac{1}{r^2} B_1 \partial_Z A_1
+N_3^u & = -\frac{1}{r}J\left(\Psi_1, u_2\right) - \frac{1}{r}J\left(\Psi_2, u_1\right) - \frac{1}{r}\widetilde{J} \left(\Psi_1, u_1\right) + \frac{1}{r}\Co J\left(A_1, B_2\right) + \frac{1}{r} \Co J\left(A_2, B_1\right) + \frac{1}{r}\Co \widetilde{J}\left(A_1, B_1\right) \\
+& - \frac{1}{r^2} u_1\partial_z \Psi_2 - \frac{1}{r^2} u_2 \partial_z \Psi_1 - \frac{1}{r^2} u_1 \partial_Z \Psi_1 + \Co \frac{1}{r^2} B_1 \partial_z A_2 + \Co \frac{1}{r^2} B_2 \partial_z A_1 + \Co \frac{1}{r^2} B_1 \partial_Z A_1
 \end{split}

### 2017-03-13 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
thingap -> local, clarifying intro text

@@
 We present a multiple scales analysis of the non-ideal MRI in the weakly nonlinear regime -- that is, when the most unstable MRI mode has a growth rate asymptotically approaching zero from above.
-Here, we develop our theory in a thin-gap, Cartesian channel.
+Here, we develop our theory in a \edit1{local}, Cartesian channel.
 Our results confirm the finding by Umurhan et al.
@@
 The latter authors undertook the first weakly nonlinear analysis of the MRI in a thin-gap Taylor-Couette (TC) flow with strong dissipation (as is appropriate to liquid metal experiments), and found that the marginal MRI system approaches saturation in a manner analogous to that of Rayleigh-B\'enard convection.
-Weakly nonlinear analysis was instrumental in our understanding of Rayleigh-B\'enard convection saturation \citep{Newell:1969}, and the similarities between convection and the thin-gap MRI are the result of important shared symmetries between the systems.
+Weakly nonlinear analysis was instrumental in our understanding of Rayleigh-B\'enard convection saturation \citep{Newell:1969}, and the similarities between convection and the \edit1{local} MRI are the result of important shared symmetries between the systems.
 The success of URM07 in modeling the MRI system near threshold merits further consideration, but we are unaware of any other attempts to expand upon their theoretical framework.
 In this work we rederive the theory of URM07, and expand upon their findings.
-Our focus here is on fully characterizing the thin-gap MRI system, both for independent theoretical interest and to have a robust comparison point for extensions of this theory into more complicated geometries.
+Our focus here is on fully characterizing the \edit1{local} MRI system, both for independent theoretical interest and to have a robust comparison point for extensions of this theory into more complicated geometries.
 In a companion paper we derive for the first time the weakly nonlinear theory of the standard and helical MRI in a global, cylindrical TC flow \citep[][hereafter Paper II]{Clark:2016}.
@@
 true Taylor-Couette flow) the base angular velocity is a function of $r$, where in the thin-gap approximation the shear flow reduces to a linear profile.
+\edit1{The equations of motion are thus identical to the MRI in a local shearing box, which differs from our fiducial setup only in the application of periodic boundary conditions.}
 
@@
 \citet{Latter:2015} gives a detailed analysis of the relation between local and global linear MRI modes.
-
-We begin with an overview of our basic model equations for the thin gap MRI in section~\ref{sec:equations} and then describe our weakly non-linear analysis and give basic results in section~\ref{sec:wnl}.
-We then place our results in the ocntext of previous results from both analytic and computational studies in section~\ref{sec:discussion} and draw conclusions in section~\ref{sec:conclusion}.
+\edit1{In this work we describe the applicability of our weakly nonlinear theory to shear-periodic boxes.
+We find that under certain conditions the weakly nonlinear mode interaction described here may provide an alternative MRI saturation mechanism in the shearing box that does not rely on parasitic instabilities.}
+
+We begin with an overview of our basic model equations for the \edit1{local} MRI in section~\ref{sec:equations} and then describe our weakly nonlinear analysis and give results \edit1{for the thin-gap TC flow} in section~\ref{sec:wnl}.
+\edit1{In Section \ref{sec:SBAD} we detail the conditions under which our theory applies to the case where the boundary conditions are shear periodic, namely the consideration of ambipolar diffusion.} We then place our results in the context of previous results from both analytic and computational studies in section~\ref{sec:discussion} and draw conclusions in section~\ref{sec:conclusion}.
 
@@
 
-\subsection{\edit1{Shearing box and ambipolar diffusion}}\label{sec:SBAD}
+\section{\edit1{Shearing box and ambipolar diffusion}}\label{sec:SBAD}
 
@@
 The \citet{Vasil:2015} model demonstrates saturation without diffusive effects; our results show that outside of the boundary layers, a simultaneous rearrangement of momentum and field occurs.
-In the boundary layers, the non-linear advection balances viscous dissipation.
+In the boundary layers, the nonlinear advection balances viscous dissipation.
 
@@
 \edit1{The radial average of the saturated $B_z$ is $B_0$, i.e. $B_z$ is marginally stable.}
- \citet{Ebrahimi:2009ey} considered the saturation of a single, strongly super-critical MRI mode allowed to interact non-linearly only with itself and the mean.
+ \citet{Ebrahimi:2009ey} considered the saturation of a single, strongly super-critical MRI mode allowed to interact nonlinearly only with itself and the mean.
 They considered two important cases, one in which the mean flow was forced to remain at its initial, quasi-Keplerian state for all time, and one in which the background flow was allowed to evolve.
@@
 Whether and how \citet{Vasil:2015}'s amplitude equation is equivalent to our own in the limit of dynamically important resistive and viscous effects is beyond the scope of this work.
-Nevertheless, the author identifies the non-linear term responsible for saturation as consisting of flux and field transport and notes these are the only mechanisms able to produce saturation.
+Nevertheless, the author identifies the nonlinear term responsible for saturation as consisting of flux and field transport and notes these are the only mechanisms able to produce saturation.
 Our results likewise demonstrate a combination of flux and field transport in the comparable region of our domain.

### 2017-03-13 python/allorders_2.py
changed sign of Gtwiddle in thingap text

@@ -47,7 +47,7 @@ class MRI():
             self.setbcs = False
             logger.info("Not running with Chebyshev basis; will not set boundary conditions.")
         
-        logger.info("MRI parameters: Q = {}; Rm = {}; Pm = {}; q = {}; beta = {}; norm = {}, Re = {}".format(self.Q, self.Rm, self.Pm, self.q, self.beta, norm, self.R))
+        logger.info("MRI parameters: Q = {}; Rm = {}; Pm = {}; q = {}; beta = {}; norm = {}, Re = {}, B0 = {}".format(self.Q, self.Rm, self.Pm, self.q, self.beta, norm, self.R, self.B0))
         
         
     def set_boundary_conditions(self, problem, noslip=True):
@@ -89,6 +89,7 @@ class MRI():
 
     def fastest_growing(self):
         gr, largest_eval_indx,freq  = self.EP.growth_rate({})
+        self.growth_rate = gr
         self.largest_eval_indx = largest_eval_indx
         self.EP.solver.set_state(largest_eval_indx)
     

### 2017-03-14 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
changed sign of Gtwiddle in thingap text
updated Oishi grant number
added discussion of growth rate with B0; added figure.

@@
-%%
 %% Beginning of file 'sample61.tex'
@@
 To excite the weakly nonlinear MRI, we tune the background magnetic field down away from stability (recall that stronger vertical fields stabilize the MRI).
-We do so by substituting $B = B_0\left(1 - \epsilon^2\right)$.
+We do so by substituting \edit1{$B = B_0\left(1 + \epsilon^2\right)$}.
 The degree of departure from the marginal state is measured by the small parameter $\epsilon$.
-An $\mathcal{O}\left(\epsilon^2\right)$ weakening of the background magnetic field destabilizes a finite band of wave modes with a width of $\mathcal{O}\left(\epsilon\right)$, which interact nonlinearly.
+An $\mathcal{O}\left(\epsilon^2\right)$ \edit1{strengthening} of the background magnetic field destabilizes a finite band of wave modes with a width of $\mathcal{O}\left(\epsilon\right)$, which interact nonlinearly.
+\edit1{We note that this definition of $\epsilon$ is opposite in sign to nearly all previous works \citep[e.g.][]{Umurhan:2007dz,Umurhan:2007hs}.
+Because in the ideal limit, the MRI can be tuned into instability by setting $B_0$ to its critical value and then \emph{decreasing} its value, it is natural to consider $\epsilon^2$ as a weakening of the background field.
+However, as we show in figure~\ref{fig:growth_rate_vs_B0}, for the dissipative case with $\eta, \nu \ne 0$, when all other parameters are critical, decreasing $B_0$ leads to stability, while increasing it pushes the system into instability.
+Figure~\ref{fig:growth_rate_vs_B0} is symmetric about $B_0 = 0$, as it must be, since the MRI is insensitive to the sign of the background field.
+There are several places at which the derivative of $\gamma$ appears discontinuous; this is not physical but rather reflects the fact that we define $\gamma$ as the growth rate of the most unstable mode.
+That is, it is the maximum real part of the eigenvalues of the linearized system (e.g.
+equation~\ref{eq:unperturbed_matrix_equations} with $\mathbf{N} = 0$).
+Because there are four wave families in rotating incompressible MHD, each modified differently by changing $B_0$, when the growth rates of the individual modes cross, there appear piecewise continuous solutions.
+We highlight one such point in the inset in figure~\ref{fig:growth_rate_vs_B0}, where the MRI mode becomes more stable than another mode which is always stable.
+Since all of these piecewise discontinuities are below $\gamma=0$, they do not affect the analysis here.}
+
+\begin{figure}[[h!]]
+  \centering
+  \includegraphics[width=\columnwidth]{growth_vs_B0.pdf}
+  \caption{Growth rate $\gamma$ as a function of background magnetic field strength $B_0$ at $\reym = \reym_c$, $k_z = k_c$, $\Pm = 10^{-3}$.
+Around the critical value $B_0 = 1.$, \emph{strengthening} $B_0$ tunes the system into instability, while \emph{decreasing} it leads to stability.
+The inset highlights the fact that $\gamma$ is determined by the maximum real eigenvalue of the system, which switches from one mode family to another as discussed in the text.}
+  \label{fig:growth_rate_vs_B0}
+\end{figure}
 
@@
 J.S.O.
-acknowledges support from NASA grant 15-LWS15\_2-0063.
+acknowledges support from NASA grant \edit1{NNX16AC92G}.
 The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
@@
 \beq
-\widetilde{\mathcal{G}} = - \mathcal{G} \partial_z - \mathcal{L}_3 \partial_z^3,
+\widetilde{\mathcal{G}} = \mathcal{G} \partial_z + \mathcal{L}_3 \partial_z^3,
 \eeq

### 2017-03-14 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
rearranged eqns

@@
 %% this information may be modified by the editorial offices.
-\shorttitle{The weakly nonlinear magnetorotational instability}
+\shorttitle{The weakly nonlinear magnetorotational instability in a local geometry}
 \shortauthors{Clark \& Oishi}
@@
 \label{eqset2}
-\partial_t u_{y} \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \, \Co B_0\partial_z B_{y} \, + \, \Co J\left(A, B_{y}\right) \, - \, J\left(\Psi, u_{y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{y}
+\partial_t u_{y} \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, - \, \Co B_0\partial_z B_{y} \, - \, \frac{1}{\reye} \nabla^2 u_{y} \, = \, \Co J\left(A, B_{y}\right) \, - \, J\left(\Psi, u_{y}\right)
 \eeq
@@
 \label{eqset3}
-\partial_t A \, = \, B_0 \partial_z \Psi \, + \, J\left(A, \Psi\right) \, + \, \frac{1}{\reym} \nabla^2 A
+\partial_t A \, - \, B_0 \partial_z \Psi \, - \, \frac{1}{\reym} \nabla^2 A \, = \, J\left(A, \Psi\right) 
 \eeq
@@
 \label{eqset4}
-\partial_t B_{y} \, = \, B_0 \partial_z u_{y} \, + \, J\left(A, u_{y}\right) \, - \, J\left(\Psi, B_{y}\right) \, \\
-+ \, \frac{1}{\reym} \nabla^2 B_{y}  \, - \, q \Omega_0 \partial_z A,
+\partial_t B_{y} \, + \, q \Omega_0 \partial_z A \, - \, B_0 \partial_z u_{y} \, - \, \frac{1}{\reym} \nabla^2 B_{y} \, = \, J\left(A, u_{y}\right) \, - \, J\left(\Psi, B_{y}\right)  ,
 \eeq
@@
 Because we nondimensionalize $B$ by the magnitude of the background field strength, $B_0 \equiv 1$ in Equations \ref{eqset1} - \ref{eqset4}.
-To excite the weakly nonlinear MRI, we tune the background magnetic field down away from stability (recall that stronger vertical fields stabilize the MRI).
+To excite the weakly nonlinear MRI, we tune the background magnetic field away from stability.
+%(recall that stronger vertical fields stabilize the MRI).
 We do so by substituting \edit1{$B = B_0\left(1 + \epsilon^2\right)$}.
@@
   \includegraphics[width=\columnwidth]{growth_vs_B0.pdf}
-  \caption{Growth rate $\gamma$ as a function of background magnetic field strength $B_0$ at $\reym = \reym_c$, $k_z = k_c$, $\Pm = 10^{-3}$.
+  \caption{\edit1{Growth rate $\gamma$ as a function of background magnetic field strength $B_0$ at $\reym = \reym_c$, $k_z = k_c$, $\Pm = 10^{-3}$.
 Around the critical value $B_0 = 1.$, \emph{strengthening} $B_0$ tunes the system into instability, while \emph{decreasing} it leads to stability.
-The inset highlights the fact that $\gamma$ is determined by the maximum real eigenvalue of the system, which switches from one mode family to another as discussed in the text.}
+The inset highlights the fact that $\gamma$ is determined by the maximum real eigenvalue of the system, which switches from one mode family to another as discussed in the text.}}
   \label{fig:growth_rate_vs_B0}

### 2017-03-14 python/plot_thingap_sim_wnl_energy.py

@@ -32,11 +32,14 @@ data0 = h5py.File(root+simname0+"_scalar_s1.h5", "r")
 wnl_file_root = "/Users/susanclark/weakly_nonlinear_MRI/data/"
 wnl_fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_128"
 wnl_fn_root = "../data/"
-nLz = 2
+nLz = 1
 
 wnl_TE = np.zeros(len(epss))
 wnl_BE = np.zeros(len(epss))
 wnl_KE = np.zeros(len(epss))
+wnl_Jdot_tot = np.zeros(len(epss))
+wnl_Jdot_R = np.zeros(len(epss))
+wnl_Jdot_M = np.zeros(len(epss))
 
 for i, _eps in enumerate(epss):
     wnl_data = h5py.File(wnl_fn_root + "zavg_quantities_"+str(int(nLz))+"Lz_eps"+str(_eps) + wnl_fn + ".h5", "r")
@@ -44,6 +47,10 @@ for i, _eps in enumerate(epss):
     wnl_TE[i] = wnl_data['TEint'].value.real
     wnl_BE[i] = wnl_data['BEint'].value.real
     wnl_KE[i] = wnl_data['KEint'].value.real
+    
+    wnl_Jdot_tot[i] = wnl_data['Jdot_tot'].value.real
+    wnl_Jdot_R[i] = wnl_data['Jdot_R'].value.real
+    wnl_Jdot_M[i] = wnl_data['Jdot_M'].value.real
 
 # values taken from Jeff's plot_energy.py
 t_orb = 2*np.pi
@@ -57,18 +64,23 @@ u_rms = np.append(data0['/tasks/vx_rms'][:,0,0], data['/tasks/vx_rms'][:,0,0])
 KE = np.append(data0['/tasks/KE'][:,0,0], data['/tasks/KE'][:,0,0])
 BE = np.append(data0['/tasks/BE'][:,0,0], data['/tasks/BE'][:,0,0])
 TE = KE + BE
-
+BxBy = np.append(data0['/tasks/BxBy'][:,0,0], data['/tasks/BxBy'][:,0,0])
+uxuy = np.append(data0['/tasks/uxuy'][:,0,0], data['/tasks/uxuy'][:,0,0])
+tot_stress = uxuy + BxBy
 
[…]

### 2017-03-14 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
fixed edit1 latex problem, it was choking on edit1 in a section header
dedalus acknowledged with software directive

@@
 
-\section{\edit1{Shearing box and ambipolar diffusion}}\label{sec:SBAD}
+\section{Shearing box and ambipolar diffusion}\label{sec:SBAD}
 
@@
 
+\section{Simulations}\label{sec:simulations}
+
+\edit1{We perform nonlinear, axisymmetric simulations of our fiducial setup using Dedalus.}
+
 \section{Discussion}
@@
 acknowledges support from NASA grant \edit1{NNX16AC92G}.
-The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
+\edit1{We thank the anonymous referee for thoughtful comments that greatly improved the manuscript.} The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
+\software{Dedalus: http://dedalus-project.org/}
 

### 2017-03-15 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
some sim text while we wait

@@
 
+\onecolumngrid
 Our final equation set is
@@
 %\begin{widetext}
-\onecolumngrid
+
 \beq
@@
 
-\edit1{We perform nonlinear, axisymmetric simulations of our fiducial setup using Dedalus.}
+\edit1{We perform two-dimensional, axisymmetric simulations of our fiducial setup using Dedalus.
+We solve the nonlinear Equations \ref{eqset1} - \ref{eqset4} with the background magnetic field set to $B_0 = 1 + \epsilon^2$ so that the system is MRI unstable.
+We initialize the runs with the linear eigenvalue solutions of the MRI equations multiplied by a small initial amplitude $A_0 = 1\times 10^{-3}$.
+We initialize the linear eigenvalue form for computational efficiency, as the simulation quickly enters its linear growth phase.
+We confirm that these simulations behave the same as simulations with noise initial conditions once they enter linear growth.
+We use a four-stage, third-order Runge-Kutta integration scheme to timestep the system until saturation.}
 

### 2017-03-15 python/plot_thingap_sim_wnl_energy.py

@@ -34,6 +34,12 @@ wnl_fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_2
 wnl_fn_root = "../data/"
 nLz = 1
 
+newfn =root+"MRI_run_Rm4.90e+00_eps5.00e-01_Pm1.00e-03_beta2.50e+01_Q7.50e-01_qsh1.50e+00_Omega1.00e+00_nx128_nz1024_Lz2_CFL_evalueIC/scalar_s1_p0.h5"
+newdata = h5py.File(newfn, "r")
+
+newfn2 = root+"MRI_run_Rm4.90e+00_eps1.00e-01_Pm1.00e-03_beta2.50e+01_Q7.50e-01_qsh1.50e+00_Omega1.00e+00_nx128_nz1024_Lz2_CFL/scalar_s1.h5"
+newdata2 = h5py.File(newfn2, "r")
+
 wnl_TE = np.zeros(len(epss))
 wnl_BE = np.zeros(len(epss))
 wnl_KE = np.zeros(len(epss))

### 2017-03-15 python/widegap/pleiades_run_hmri_amplitude_bro_ele.sh

@@ -1,6 +1,6 @@
 #PBS -S /bin/bash
 #PBS -N WidegapFinalLongLongBroEle
-#PBS -l select=2:ncpus=10:mpiprocs=10:model=bro_ele
+#PBS -l select=4:ncpus=5:mpiprocs=5:model=bro_ele
 #PBS -l walltime=120:00:00
 #PBS -j oe
 #PBS -q long

### 2017-03-15 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
spurious character

--- 2017-03-14 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
@@
 %% this information may be modified by the editorial offices.
-\shorttitle{The weakly nonlinear magnetorotational instability in a local geometry}
+\shorttitle{The weakly nonlinear magnetorotational instability}
 \shortauthors{Clark \& Oishi}
@@
 \label{eqset2}
-\partial_t u_{y} \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, - \, \Co B_0\partial_z B_{y} \, - \, \frac{1}{\reye} \nabla^2 u_{y} \, = \, \Co J\left(A, B_{y}\right) \, - \, J\left(\Psi, u_{y}\right)
+\partial_t u_{y} \, + \, \left(2 - q\right) \Omega_0 \partial_z \Psi \, = \, \Co B_0\partial_z B_{y} \, + \, \Co J\left(A, B_{y}\right) \, - \, J\left(\Psi, u_{y}\right) \, + \, \frac{1}{\reye} \nabla^2 u_{y}
 \eeq
@@
 \label{eqset3}
-\partial_t A \, - \, B_0 \partial_z \Psi \, - \, \frac{1}{\reym} \nabla^2 A \, = \, J\left(A, \Psi\right) 
+\partial_t A \, = \, B_0 \partial_z \Psi \, + \, J\left(A, \Psi\right) \, + \, \frac{1}{\reym} \nabla^2 A
 \eeq
@@
 \label{eqset4}
-\partial_t B_{y} \, + \, q \Omega_0 \partial_z A \, - \, B_0 \partial_z u_{y} \, - \, \frac{1}{\reym} \nabla^2 B_{y} \, = \, J\left(A, u_{y}\right) \, - \, J\left(\Psi, B_{y}\right)  ,
+\partial_t B_{y} \, = \, B_0 \partial_z u_{y} \, + \, J\left(A, u_{y}\right) \, - \, J\left(\Psi, B_{y}\right) \, \\
++ \, \frac{1}{\reym} \nabla^2 B_{y}  \, - \, q \Omega_0 \partial_z A,
 \eeq
@@
 Because we nondimensionalize $B$ by the magnitude of the background field strength, $B_0 \equiv 1$ in Equations \ref{eqset1} - \ref{eqset4}.
-To excite the weakly nonlinear MRI, we tune the background magnetic field away from stability.
-%(recall that stronger vertical fields stabilize the MRI).
+To excite the weakly nonlinear MRI, we tune the background magnetic field down away from stability (recall that stronger vertical fields stabilize the MRI).
 We do so by substituting \edit1{$B = B_0\left(1 + \epsilon^2\right)$}.
@@
 
-\begin{figure}[[h!]]
+\begin{figure}[[h!]
   \centering
   \includegraphics[width=\columnwidth]{growth_vs_B0.pdf}
-  \caption{\edit1{Growth rate $\gamma$ as a function of background magnetic field strength $B_0$ at $\reym = \reym_c$, $k_z = k_c$, $\Pm = 10^{-3}$.
+  \caption{Growth rate $\gamma$ as a function of background magnetic field strength $B_0$ at $\reym = \reym_c$, $k_z = k_c$, $\Pm = 10^{-3}$.
 Around the critical value $B_0 = 1.$, \emph{strengthening} $B_0$ tunes the system into instability, while \emph{decreasing} it leads to stability.
-The inset highlights the fact that $\gamma$ is determined by the maximum real eigenvalue of the system, which switches from one mode family to another as discussed in the text.}}
+The inset highlights the fact that $\gamma$ is determined by the maximum real eigenvalue of the system, which switches from one mode family to another as discussed in the text.}
   \label{fig:growth_rate_vs_B0}

### 2017-03-20 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
minor word changes; more edits on the bus.
added Oishi ORCID.
Edited simulations section; added some to conclusions.

--- 2017-03-14 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
@@
 
-\author{Jeffrey S. Oishi}
+\author[0000-0001-8531-6570]{Jeffrey S. Oishi}
 \affiliation{Department of Physics and Astronomy, Bates College, Lewiston, ME}
@@
 
-\begin{figure}[[h!]]
+\begin{figure}[h!]
   \centering
@@
 
-\edit1{We perform nonlinear, axisymmetric simulations of our fiducial setup using Dedalus.}
+\edit1{Here, we make a preliminary test of our weakly nonlinear theory by comparing it to direct numerical simulation.
+Using Dedalus, we  solved the full, nonlinear equations \ref{eqset1} - \ref{eqset4} with all parameters ($\reym$, $Q$) equal to their critical values except the background magnetic field, which we set to $B_0 = 1 + \epsilon^2$.
+We thus drive the system MRI unstable in the same way as in our theory.
+The computational requirements of low $\Pm$ simulations are quite intense in both time and space.
+Despite being virtually smooth, the solutions require a resolution of $192 \times 1536$ grid points at $\Pm = 10^{-2}$.
+Because the system has such a small growth rate, it takes hundreds of orbits for the system to reach saturation, as compared to the few orbits typical of high $\reym$ simulations \citep[e.g.][]{2007MNRAS.378.1471L}.
+As a result, we make our comparison at $\Pm = 10^{-2}$, which provides a good tradeoff between probing relatively low $\Pm$ while keeping the computational time for these exploratry simulations modest.
+}
+
+\edit1{We initialize the runs with the linear eigenvectors of the MRI unstable mode (also computed by Dedalus; see section~\ref{sec:wnl}) multiplied by an initial amplitude $A_0 = 10^{-3}$.
+Doing so requires considerably less run time, as the MRI unstable mode starts growing immediately from $A_0$.
+By contrast, initializing random noise in $\psi$ with amplitude $A_0$ would give the unstable mode a much smaller amplitude.
+Nevertheless, we have confirmed that simulations with eigenvector initial conditions have similar evolutions to those with noise initial conditions once each enter linear growth.
+}
+
+\textbf{SUSAN: SIMULATION RESULTS GO HERE}
 
@@
 In doing so, we find the saturated state is a complex balance in which reduction of shear and amplification and redistribution of $B_z$ combine to saturate the instability.
-We also perform direct numerical simulations of the amplitude equation.
-We demonstrate that complex patterns can organize the flow on long length scales $Z$, though the maximum magnitude of the amplitude $\alpha$ is well predicted by the steady state solution.
-In Paper II, we make use of the techniques developed here to extend the weakly nonlinear analysis of the MRI to a full cylindrical geometry appropriate for a Taylor-Couette experiment.
+\edit1{We also perform numerical simulations of the amplitude equation and a direct numerical simulation of the MRI system.} \edit1{Using the former, we demonstrate that complex patterns can organize the flow on long length scales $Z$, though the maximum magnitude of the amplitude $\alpha$ is well predicted by the steady state solution.
+The latter show that there is rough agreement for both total energy and average angular momentum transport between the weakly nonlinear theory and simulation for a representative case at $\Pm = 10^{-2}$.
+We defer a full comparison between theory and simulation to later work.} In Paper II, we make use of the techniques developed here to extend the weakly nonlinear analysis of the MRI to a full cylindrical geometry appropriate for a Taylor-Couette experiment.
 
@@
 
+\bibitem[Lesur \& Longaretti(2007)]{2007MNRAS.378.1471L} Lesur, G., \&
+  Longaretti, P.-Y.\ 2007, \mnras, 378, 1471
+
 \bibitem[{Makwana et al.(2011)}]{Makwana:2011}

### 2017-03-20 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
minor edit: begging the question is a subtly different thing.

@@
 \edit1{The nonlinear property of primary MRI modes in the shearing box motivates the addition of radial boundaries, such that the nonlinear evolution of the weakly nonlinear MRI can be properly considered.
-It also begs the question of whether some additional nonlinear mechanism can be introduced such that the fastest-growing modes are no longer nonlinear solutions to the shearing box equations.
+It also raises the question of whether some additional nonlinear mechanism can be introduced such that the fastest-growing modes are no longer nonlinear solutions to the shearing box equations.
 It has already been shown that the Hall effect does not negate the nonlinear property of primary MRI modes \citep{Kunz:2013}.

### 2017-03-22 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
unnorm 128 Pm1E-2 run

@@
 Because the system has such a small growth rate, it takes hundreds of orbits for the system to reach saturation, as compared to the few orbits typical of high $\reym$ simulations \citep[e.g.][]{2007MNRAS.378.1471L}.
-As a result, we make our comparison at $\Pm = 10^{-2}$, which provides a good tradeoff between probing relatively low $\Pm$ while keeping the computational time for these exploratry simulations modest.
+As a result, we make our comparison at $\Pm = 10^{-2}$, which provides a good tradeoff between probing relatively low $\Pm$ while keeping the computational time for these exploratory simulations modest.
 }

### 2017-03-22 python/plot_thingap_dissipation_investigation.py

@@ -14,29 +14,66 @@ file_root = "/Users/susanclark/weakly_nonlinear_MRI/data/"
 fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_128"
 #fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8738_Pm_1.00e-04_q_1.5_beta_25.00_gridnum_256_Anorm"
 #fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_256_Anorm"
-fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_256_Anorm"
+#fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_256_Anorm"
 
-fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.9000_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_256_Anorm"
+#fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.9000_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_256_Anorm"
+#fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.9000_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_128_Anorm"
+
+#fn = "thingap_amplitude_parameters_Q_0.76_Rm_4.9213_Pm_1.00e-02_q_1.5_beta_25.00_gridnum_128_Anorm"
+#fn = "thingap_amplitude_parameters_Q_0.84_Rm_5.3502_Pm_1.00e-01_q_1.5_beta_25.00_gridnum_128_Anorm"
+#fn = "thingap_amplitude_parameters_Q_0.76_Rm_4.9213_Pm_1.00e-02_q_1.5_beta_25.00_gridnum_128_Anorm"
+fn = "thingap_amplitude_parameters_Q_0.76_Rm_4.9213_Pm_1.00e-02_q_1.5_beta_25.00_gridnum_128_Anorm"
+
+
 
 obj = h5py.File(file_root + fn + ".h5", "r")
 
 Q = obj.attrs['Q']
+b = obj.attrs['b']
[…]

### 2017-03-22 python/plot_thingap_sim_wnl_energy.py

@@ -17,6 +17,7 @@ Rms = [5.7, 5.3, 5.1]
 simrootfns = [simroot1, simroot2, simroot3]
 
 simroot = simroot3
+restart = False
 
 # data after restart
 root = "/Users/susanclark/weakly_nonlinear_mri/data/simulations/"
@@ -30,7 +31,8 @@ data0 = h5py.File(root+simname0+"_scalar_s1.h5", "r")
 
 # data from WNL theory
 wnl_file_root = "/Users/susanclark/weakly_nonlinear_MRI/data/"
-wnl_fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_128"
+#wnl_fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00_gridnum_128"
+wnl_fn = "thingap_amplitude_parameters_Q_0.76_Rm_4.9213_Pm_1.00e-02_q_1.5_beta_25.00_gridnum_128_Anorm"
 wnl_fn_root = "../data/"
 nLz = 1
 

### 2017-03-22 python/run_amplitude.py

@@ -55,6 +55,7 @@ beta = 25.0
 Rm = 4.92129e+00
 Q = 7.55027e-01
 Pm=1.0E-2
+norm=False
 
 #Rm = 4.87384744169
 #Q = 0.746395815905
@@ -84,11 +85,12 @@ domain = de.Domain([x_basis], np.complex128, comm=MPI.COMM_SELF)
 print("running at gridnum", gridnum)
 
 fn_root = "../data/"
-fn = fn_root + "thingap_amplitude_parameters_Q_{:03.2f}_Rm_{:04.4f}_Pm_{:.2e}_q_{:02.1f}_beta_{:.2f}_gridnum_{}_Anorm.h5".format(Q, Rm, Pm, q, beta, gridnum)
+#fn = fn_root + "thingap_amplitude_parameters_Q_{:03.2f}_Rm_{:04.4f}_Pm_{:.2e}_q_{:02.1f}_beta_{:.2f}_gridnum_{}_Anorm.h5".format(Q, Rm, Pm, q, beta, gridnum)
+fn = fn_root + "thingap_amplitude_parameters_Q_{:03.2f}_Rm_{:04.4f}_Pm_{:.2e}_q_{:02.1f}_beta_{:.2f}_gridnum_{}_norm_{}.h5".format(Q, Rm, Pm, q, beta, gridnum, norm)
 
 print(fn)
 
-aa = AmplitudeAlpha(domain,Q = Q,Rm = Rm, Pm = Pm, q=q, beta=beta)
+aa = AmplitudeAlpha(domain,Q = Q,Rm = Rm, Pm = Pm, q=q, beta=beta, norm=norm)
 aa.print_coeffs()
 
 with h5py.File(fn,'w') as f:
@@ -216,6 +218,7 @@ with h5py.File(fn,'w') as f:
     f.attrs["beta"] = aa.beta
     f.attrs["gridnum"] = aa.gridnum
     f.attrs["q"] = aa.q
+    f.attrs["norm"] = aa.norm
     
     f.attrs["a"] = aa.a
     f.attrs["b"] = aa.b

### 2017-03-22 python/simulations/calc_transport_bulk.py

@@ -46,8 +46,8 @@ def calc_ang_mom_bulk(nt, data, fields, frac = None):
     bulk_right_edge = 0.5
     Lx_bulk = bulk_right_edge - bulk_left_edge
 
-    avg_Ma_stress_bulk = (Ma_stress_anti_zint.interpolate(x=bulk_right_edge) - Ma_stress_anti_zint.interpolate(x=bulk_right_edge)).evaluate()
-    avg_Re_stress_bulk = (Re_stress_anti_zint.interpolate(x=bulk_right_edge) - Re_stress_anti_zint.interpolate(x=bulk_right_edge)).evaluate()
+    avg_Ma_stress_bulk = (Ma_stress_anti_zint.interpolate(x=bulk_right_edge) - Ma_stress_anti_zint.interpolate(x=bulk_left_edge)).evaluate()
+    avg_Re_stress_bulk = (Re_stress_anti_zint.interpolate(x=bulk_right_edge) - Re_stress_anti_zint.interpolate(x=bulk_left_edge)).evaluate()
     avg_Ma_stress = avg_Ma_stress_bulk['g'][0,0]/(Lx_bulk*Lz)
     avg_Re_stress = avg_Re_stress_bulk['g'][0,0]/(Lx_bulk*Lz)
     

### 2017-03-22 python/widegap/plot_widegap_dissipation_investigation.py

@@ -12,6 +12,7 @@ rc('text', usetex=True)
 file_root = "/home/jsoishi/hg-projects/weakly_nonlinear_MRI/data/"
 #fn = "widegap_amplitude_parameters_Q_0.01_Rm_0.8403_Pm_1.00e-03_Omega1_313.55_Omega2_56.43_beta_25.00_xi_0.00_gridnum_128_Anorm.h5"
 fn = "widegap_amplitude_parameters_Q_0.90_Rm_3.3088_Pm_1.60e-06_Omega1_01.00_Omega2_00.12_beta_41.20_xi_0.00_gridnum_512_norm_True_Amidnorm_final"
+
 obj = h5py.File(file_root + fn + ".h5", "r")
 
 Q = obj.attrs['Q']

### 2017-03-22 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
added sim figure

@@
 
-\textbf{SUSAN: SIMULATION RESULTS GO HERE}
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{sim_theory_energy_jdot_Pm_1E-2_nonorm}
+\caption{\edit1{Average energy (left) and angular momentum transport (right) in the total, kinetic, and magnetic components of simulation data as a function of time.
+Gray lines show the weakly nonlinear theory values for each quantity.
+}} \label{fig:sim_theory_energy_jdot}
+\end{figure*}
+
+\edit1{We analyze the average energy and angular momentum transport in the simulation domain (Figure \ref{fig:sim_theory_energy_jdot}).  }
+
+%\textbf{SUSAN: SIMULATION RESULTS GO HERE}
 

### 2017-03-22 python/plot_thingap_dissipation_investigation.py

@@ -24,6 +24,7 @@ fn = "thingap_amplitude_parameters_Q_0.75_Rm_4.8790_Pm_1.00e-03_q_1.5_beta_25.00
 #fn = "thingap_amplitude_parameters_Q_0.76_Rm_4.9213_Pm_1.00e-02_q_1.5_beta_25.00_gridnum_128_Anorm"
 fn = "thingap_amplitude_parameters_Q_0.76_Rm_4.9213_Pm_1.00e-02_q_1.5_beta_25.00_gridnum_128_Anorm"
 
+fn = "thingap_amplitude_parameters_Q_0.76_Rm_4.9213_Pm_1.00e-02_q_1.5_beta_25.00_gridnum_128_norm_False" # non normalized
 
 
 obj = h5py.File(file_root + fn + ".h5", "r")

### 2017-03-22 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
more text edits around figures

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{thingap_stressfig_Pm_1E-2}
+\caption{\edit1{Total (Reynolds + Maxwell) stress in the domain as predicted from the weakly nonlinear theory.}} \label{fig:wnl_stress}
+\end{figure}
+
 \edit1{Many studies of the MRI consider the instability in a shearing box, i.e.
@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{sim_theory_energy_jdot_Pm_1E-2_nonorm}
+\caption{\edit1{Average energy (left) and angular momentum transport (right) in the total, kinetic, and magnetic components of simulation data as a function of time.
+Gray lines show the weakly nonlinear theory values for each quantity.
+}} \label{fig:sim_theory_energy_jdot}
+\end{figure*}
+
 \edit1{The nonlinear property of primary MRI modes in the shearing box motivates the addition of radial boundaries, such that the nonlinear evolution of the weakly nonlinear MRI can be properly considered.
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{sim_theory_energy_jdot_Pm_1E-2_nonorm}
-\caption{\edit1{Average energy (left) and angular momentum transport (right) in the total, kinetic, and magnetic components of simulation data as a function of time.
-Gray lines show the weakly nonlinear theory values for each quantity.
-}} \label{fig:sim_theory_energy_jdot}
-\end{figure*}
-
-\edit1{We analyze the average energy and angular momentum transport in the simulation domain (Figure \ref{fig:sim_theory_energy_jdot}).  }
+\edit1{We analyze the average energy and angular momentum transport in the simulation domain (Figure \ref{fig:sim_theory_energy_jdot}).
+We find good agreement between the theory and simulation values, especially considering that the weakly nonlinear theory applies rigorously to a channel of infinite height, while the simulation was carried out in a box with a vertical extent of only two critical wavelengths.
+Further comparisons between simulation and theory, including an analysis of the effect of the box height on the simulated flow, is deferred to future work.
+}
 

### 2017-03-22 python/plot_thingap_dissipation_investigation.py

@@ -7,6 +7,7 @@ import matplotlib.ticker as ticker
 from mpl_toolkits.axes_grid1 import make_axes_locatable, axes_size
 from matplotlib import rc
 import dedalus.public as de
+import copy
 rc('text', usetex=True)
 
 #file_root = "/home/joishi/hg-projects/weakly_nonlinear_mri/data/"
@@ -69,14 +70,23 @@ else:
     satamp_Z = final_alphaZ.reshape(ivpdata.attrs['gridnum'], 1)
     ivporconst = "IVP"
  
+ 
+# to avoid all ambiguity in z, take it from the sims.
+sim_fn_root = "../data/simulations/"
+simname = "MRI_run_Rm4.92e+00_eps5.00e-01_Pm1.00e-02_beta2.50e+01_Q7.60e-01_qsh1.50e+00_Omega1.00e+00_nx128_nz1024_Lz2_CFL_evalueIC"
+simfn = sim_fn_root + simname + "_plotfields.h5"
+simdata = h5py.File(simfn, "r")
+simz = simdata['z'].value
+ 
 beta = obj.attrs['beta']
 
 # create z grid
 #nz = obj.attrs['gridnum']
 nz = 1024#2048
-nLz = 2#2 # number of critical wavelengths
+nLz = 1#2 # number of critical wavelengths
 Lz = 2*np.pi/Q
-z = np.linspace(0, nLz*Lz, nz, endpoint=False)
+#z = np.linspace(0, nLz*Lz, nz, endpoint=False)
+z = copy.copy(simz)
 zz = z.reshape(nz, 1)
 
 dz = z[1] - z[0]
@@ -387,7 +397,7 @@ if save is True:
     print('saving output...')
 
     fn_root = "../data/"
-    out_fn = fn_root + "zavg_quantities_"+str(int(nLz))+"Lz_eps"+ str(eps) + fn + "_satamp_"+ivporconst+".h5"
+    out_fn = fn_root + "zavg_quantities_"+str(int(nLz))+"Lz_eps"+ str(eps) + fn + "_satamp_"+ivporconst+"_samez.h5"
     
     print(out_fn)
     

### 2017-03-22 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
minor edits and some additiona text.

--- 2017-03-22 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
@@
 \edit1{We note that this definition of $\epsilon$ is opposite in sign to nearly all previous works \citep[e.g.][]{Umurhan:2007dz,Umurhan:2007hs}.
-Because in the ideal limit, the MRI can be tuned into instability by setting $B_0$ to its critical value and then \emph{decreasing} its value, it is natural to consider $\epsilon^2$ as a weakening of the background field.
-However, as we show in figure~\ref{fig:growth_rate_vs_B0}, for the dissipative case with $\eta, \nu \ne 0$, when all other parameters are critical, decreasing $B_0$ leads to stability, while increasing it pushes the system into instability.
+Because in the ideal limit, the MRI can be tuned into instability by setting $B_0$ to its critical value and then \emph{decreasing} its value, it is natural to consider $\epsilon^2$ as a weakening of the background field \citep[as is done correctly in][for example]{Vasil:2015}.
+However, as we show in figure~\ref{fig:growth_rate_vs_B0}, for the \emph{dissipative} case with $\eta, \nu \ne 0$, when all other parameters are critical, decreasing $B_0$ leads to stability, while increasing it pushes the system into instability.
 Figure~\ref{fig:growth_rate_vs_B0} is symmetric about $B_0 = 0$, as it must be, since the MRI is insensitive to the sign of the background field.
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{thingap_stressfig_Pm_1E-2}
+\caption{\edit1{Total (Reynolds + Maxwell) stress in the domain as predicted from the weakly nonlinear theory.}} \label{fig:wnl_stress}
+\end{figure}
+
 \edit1{Many studies of the MRI consider the instability in a shearing box, i.e.
 a wall-less local approximation that is meant to represent a small section of a disk.
-The shearing box obtains when Equations \ref{eqset1} - \ref{eqset4} are subjected to shear periodic radial boundary conditions rather than Equation \ref{eq:wallbcs}.
+The shearing box is the limit in which Equations \ref{eqset1} - \ref{eqset4} are subjected to shear periodic radial boundary conditions rather than Equation \ref{eq:wallbcs} \citep[e.g.][]{2008A&A...481...21R}.
 The periodic nature of the shearing box allows us to decompose the fluid perturbations into Fourier modes proportional to $e^{i k_x x + i k_z z}$.
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{sim_theory_energy_jdot_Pm_1E-2_nonorm}
-\caption{\edit1{Average energy (left) and angular momentum transport (right) in the total, kinetic, and magnetic components of simulation data as a function of time.
-Gray lines show the weakly nonlinear theory values for each quantity.
-}} \label{fig:sim_theory_energy_jdot}
-\end{figure*}
-
-\edit1{We analyze the average energy and angular momentum transport in the simulation domain (Figure \ref{fig:sim_theory_energy_jdot}).  }
+\edit1{We analyze the average energy and angular momentum transport in the simulation domain (Figure \ref{fig:sim_theory_energy_jdot}).
+The theory slightly overpredicts both quantities, though with a reasonably good agreement considering that the weakly nonlinear theory applies rigorously to a channel of infinite height, while the simulation was carried out in a box with a vertical extent of only two critical wavelengths.
+We defer further comparisons between simulation and theory, including an analysis of the effect of the box height on the simulated flow, to future work.
+}
 
@@
 acknowledges support from NASA grant \edit1{NNX16AC92G}.
-\edit1{We thank the anonymous referee for thoughtful comments that greatly improved the manuscript.} The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
+\edit1{We thank the anonymous referee for thoughtful comments that greatly improved the manuscript.} \edit1{We also} thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
 \software{Dedalus: http://dedalus-project.org/}
@@
 
+\bibitem[Regev \& Umurhan(2008)]{2008A&A...481...21R} Regev, O., \& Umurhan, O.~M.\ 2008, \aap, 481, 21
+
 \bibitem[{Rembiasz et al.(2016)}]{Rembiasz:2016}

### 2017-03-22 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
save norm constant to test idea

@@
 \includegraphics[width=\columnwidth]{thingap_stressfig_Pm_1E-2}
-\caption{\edit1{Total (Reynolds + Maxwell) stress in the domain as predicted from the weakly nonlinear theory.}} \label{fig:wnl_stress}
+\caption{\edit1{Total (Reynolds + Maxwell) stress in the domain as predicted from the weakly nonlinear theory for \Pm=1E-2.}} \label{fig:wnl_stress}
 \end{figure}
@@
 We stress that while the third-order nonlinear terms in the walled TC flow are strongly influenced by the boundary layers, this is not generically true of the MRI system.
-Indeed, in the shearing box MRI with ambipolar diffusion (the case sketched out in Section \ref{sec:SBAD}), boundary layers impossible in the shear periodic flow.
+Indeed, in the shearing box MRI with ambipolar diffusion (the case sketched out in Section \ref{sec:SBAD}), boundary layers are impossible in the shear periodic flow.
 In this case the third-order nonlinear behavior of the system includes three-mode interactions from the cubic nonlinearity in the ambipolar diffusion term.}

### 2017-03-22 paper/widegap_ref_response/widegap_mri_paper_ref_response_apj.tex
changed sign of Htwiddle third order term in widegap paper
changed sign of Gtwiddle term in widegap paper
dedalus acknowledged with software directive
save norm constant to test idea

@@
 We solve the equations at each order using Dedalus, an open source pseudospectral code.
-We solve the radial portion of the eigenvectors on a basis of Chebyshev polynomials subject to our radial boundary conditions
+We solve the radial portion of the eigenvectors on a basis of Chebyshev polynomials subject to our radial boundary conditions.
+\edit1{We use a 512-component Chebyshev grid, and confirm numerical convergence at $1.5\times$ the resolution.
+Because }
 We solve Equation \ref{eq:ordere} as a linear eigenvalue problem, and Equation \ref{eq:ordere2} as a linear boundary value problem. 
@@
 The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
+\software{Dedalus: http://dedalus-project.org/}
 
@@
 \beq
-\langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle \partial_T \alpha + \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} - \xi \widetilde{\mathcal{H}} \mathbb{V}_{11} \rangle \alpha +  \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} + \xi \mathcal{H} \mathbb{V}_{21} \rangle \partial_Z^2 \alpha = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle \alpha |\alpha|^2,
+\langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle \partial_T \alpha + \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} + \xi \widetilde{\mathcal{H}} \mathbb{V}_{11} \rangle \alpha +  \langle \mathbb{V}^\dagger | \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} + \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} + \xi \mathcal{H} \mathbb{V}_{21} \rangle \partial_Z^2 \alpha = \langle \mathbb{V}^\dagger | \mathbf{N}_{31} \rangle \alpha |\alpha|^2,
 \eeq
@@
 \beq
-b = \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} - \xi \widetilde{\mathcal{H}} \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
+b = \langle  \mathbb{V}^\dagger | \widetilde{\mathcal{G}} \mathbb{V}_{11} + \xi \widetilde{\mathcal{H}} \mathbb{V}_{11} \rangle / \langle \mathbb{V}^\dagger | \mathcal{D} \mathbb{V}_{11} \rangle,
 \eeq

### 2017-03-22 python/allorders_2.py

@@ -180,6 +180,8 @@ class MRI():
         A['g'] = A['g']/norm
         B['g'] = B['g']/norm
         
+        self.normconst = norm
+        
         return psi, u, A, B
 
     def get_derivative(self, field):
@@ -386,6 +388,8 @@ class OrderE(MRI):
             self.psi, self.u, self.A, self.B = self.normalize_state_vector(self.psi, self.u, self.A, self.B)
         
             self.normpsi = self.psi
+        else:
+            self.normconst = None
             
         self.psi.name = "psi"
         self.u.name = "u"

### 2017-03-24 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
last edits

@@
 \includegraphics[width=\columnwidth]{thingap_stressfig_Pm_1E-2}
-\caption{\edit1{Total (Reynolds + Maxwell) stress in the domain as predicted from the weakly nonlinear theory for \Pm=1E-2.}} \label{fig:wnl_stress}
+\caption{\edit1{Total (Reynolds + Maxwell) stress in the domain as predicted from the weakly nonlinear theory for $\Pm=1E-2$.}} \label{fig:wnl_stress}
 \end{figure}
@@
 \centering
-\includegraphics[width=\textwidth]{sim_theory_energy_jdot_Pm_1E-2_nonorm}
+\includegraphics[width=\textwidth]{sim_theory_energy_jdot_Pm_1E-2_Bmaxnorm}
 \caption{\edit1{Average energy (left) and angular momentum transport (right) in the total, kinetic, and magnetic components of simulation data as a function of time.
@@
 \edit1{We analyze the average energy and angular momentum transport in the simulation domain (Figure \ref{fig:sim_theory_energy_jdot}).
-The theory slightly overpredicts both quantities, though with a reasonably good agreement considering that the weakly nonlinear theory applies rigorously to a channel of infinite height, while the simulation was carried out in a box with a vertical extent of only two critical wavelengths.
+The saturation amplitude predicted by the weakly nonlinear theory depends on the choice of normalization of the linear eigenvectors.
+The eigenvectors of the linear problem are only determined up to an arbitrary normalization, and the nonlinear coefficient of the Ginzburg-Landau equation is sensitive to this normalization.
+The undetermined factor is typically assigned by comparison with direct numerical simulation or laboratory experiment \citep[e.g.][]{Deyirmenjian:1997,Recktenwald:1993}.
+Here we determine the constant by requiring that the maximum amplitude of $B_y$ be equal in both theory and simulation.
+With this normalization choice all plotted quantities agree to within $\sim 25\%$.
+The theory and simulation are thus in reasonably good agreement considering that the weakly nonlinear theory applies rigorously to a channel of infinite height, while the simulation was carried out in a box with a vertical extent of only two critical wavelengths.
 We defer further comparisons between simulation and theory, including an analysis of the effect of the box height on the simulated flow, to future work.
 }
-
-%\textbf{SUSAN: SIMULATION RESULTS GO HERE}
 
@@
 
+\bibitem[{Deyirmenjian(1997)}]{Deyirmenjian:1997}
+Deyirmenjian, Z., Daya, A., \& Morris, S.W., 1997, Phys. Rev. E 56, 1706
+
 \bibitem[{Ebrahimi et al.(2009)}]{Ebrahimi:2009ey}
@@
 Pessah, M. E., \& Goodman, J., 2009, \apj 698, L72
+
+\bibitem[{Recktenwald et al.(1993)}]{Recktenwald:1993}
+Recktenwald, A., L\"ucke, M., M\"uller, H.W., 1993, PhRvE 48, 4444
 

### 2017-03-24 paper/widegap_ref_response/widegap_mri_paper_ref_response_apj.tex
last edits

@@
 \edit1{We use a 512-component Chebyshev grid, and confirm numerical convergence at $1.5\times$ the resolution.
-Because }
+This is sufficient to determine convergence because of the faster-than-exponential convergence of spectral methods.
+}
 We solve Equation \ref{eq:ordere} as a linear eigenvalue problem, and Equation \ref{eq:ordere2} as a linear boundary value problem. 

### 2017-03-24 python/plot_thingap_dissipation_investigation.py

@@ -27,6 +27,7 @@ fn = "thingap_amplitude_parameters_Q_0.76_Rm_4.9213_Pm_1.00e-02_q_1.5_beta_25.00
 
 fn = "thingap_amplitude_parameters_Q_0.76_Rm_4.9213_Pm_1.00e-02_q_1.5_beta_25.00_gridnum_128_norm_False" # non normalized
 
+fn = "thingap_amplitude_parameters_Q_0.76_Rm_4.9213_Pm_1.00e-02_q_1.5_beta_25.00_gridnum_128_norm_True_normconst2"
 
 obj = h5py.File(file_root + fn + ".h5", "r")
 
@@ -45,7 +46,9 @@ if constant_alpha:
     # saturation amplitude -- for now just constant, coefficient-determined
     satamp = np.sqrt(obj.attrs['b']/obj.attrs['c']) #1
     print('satamp is sqrt {}/{} = {}'.format(b, c, satamp))
-    #satamp = satamp - satamp + 0.07 #test
+    # satamp = np.sqrt(obj.attrs['b']/(obj.attrs['c']/obj.attrs['normconst']**2)) #test
+    # satamp = satamp*0 + (0.27929596842805482-4.6781424718396889e-14j)
+    satamp = satamp*0 +  0.23643163#0.23645746
     satamp_Z = 0.0 + 0j
     print('saturation amplitude = {}'.format(satamp))
     ivporconst = "const"
@@ -91,6 +94,19 @@ zz = z.reshape(nz, 1)
 
 dz = z[1] - z[0]
 
+# fit center of channel u
+#In [197]: np.where(simB == np.nanmax(simB))
+#Out[197]: (array([128, 640]), array([64, 64]))
+simB = simdata['B'].value
+#zcenterindx = np.int(nz/8.)
+#xcenterindx = np.int(len(x)/2.)
+zcenterindx = 640
+xcenterindx = 64
+zcenter = z[zcenterindx]
+xcenter = x[xcenterindx]
+simBcenter = simB[zcenterindx, xcenterindx]
+print('found simBcenter = {} at {}, {}'.format(simBcenter, zcenterindx, xcenterindx))
+
 eiqz = np.exp(1j*Q*zz)
 ei2qz = np.exp(2*1j*Q*zz)
 ei0qz = np.exp(0*1j*Q*zz)
@@ -113,7 +129,35 @@ B21_star = obj['B21'].value.conj()
[…]

### 2017-03-24 paper/thingap_ref_response/response_to_referee.tex
thingap -> local, clarifying intro text
referee response changes.

@@
 
-\textbf{Coupling to damped modes as explored in the plasma literature is not analytically possible in the walled TC flow, but can be analyzed in the shearing box when ambipolar diffusion is considered, as described in the new section of the text.}
+\textbf{Coupling to damped modes as explored in the plasma literature is not analytically possible in the walled TC flow, because the boundaries require numerical computation of the eigenvectors, precluding the simple identification of modes with branches of an algebraic dispersion relation.
+Such a program is beyond the scope of this work, though it is a very interesting possibility we had not previously considered.
+However, such inter-mode coupling \emph{can} be analyzed in the shearing box when ambipolar diffusion is considered, as described in the new section of the text.}
 
@@
 
+\textbf{We have added a figure to the thin-gap paper showing the morphology of the average angular momentum transport $\dot{J}$. }
+
 \item A more extensive discussion of the role of the neutral mode with amplitude $\beta$ that both URM and the present paper argue should be kept, but, at least to the order in $\epsilon$ considered, satisfies an amplitude equation that is completely decoupled from the MRI mode, with amplitude $\alpha$.
@@
 
+\textbf{We have made this change.}
+
 \item Finally, would it be feasible to a compute a nonlinear solution with Dedalus (or some other platform) and see how accurate the perturbation theory really is, for different values of $\epsilon$?
-Although the authors compare their results with Ebrahimi 2009, it would be desirable to have their own simulations for comparison.					\end{enumerate}
+Although the authors compare their results with Ebrahimi 2009, it would be desirable to have their own simulations for comparison.
+
+  \textbf{We have added a new section including a direct numerical simulation run with Dedalus, and made a preliminary comparison at modest Pm. }
+
+\end{enumerate}
 

### 2017-03-24 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
referee response changes.

--- 2017-03-22 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
@@
 \includegraphics[width=\columnwidth]{thingap_stressfig_Pm_1E-2}
-\caption{\edit1{Total (Reynolds + Maxwell) stress in the domain as predicted from the weakly nonlinear theory.}} \label{fig:wnl_stress}
+\caption{\edit1{Total (Reynolds + Maxwell) stress in the domain as predicted from the weakly nonlinear theory at $\Pm = 10^{-2}$.}} \label{fig:wnl_stress}
 \end{figure}
@@
 
+\edit1{Figure~\ref{fig:wnl_stress} shows the total stress $u_x u_y - \Co B_x B_y$ for the $\Pm = 10^{-2}$ model.
+The stress shows signficantly more structure throughout the domain than the  variables $u_x$, $u_y$, $B_x$ and $B_y$ that comprise it, demonstrating that a non-trivial correlation exists even in the weakly non-linear state.
+As in simulations at higher $\reym$, figure~\ref{fig:sim_theory_energy_jdot} shows that the Maxwell stress dominates over the Reynolds stresses even though the kinetic energy significantly exceeds the magnetic energy.
+}
+
 \section{Conclusion}

### 2017-03-24 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex

--- 2017-03-24 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
@@
 \includegraphics[width=\columnwidth]{thingap_stressfig_Pm_1E-2}
-\caption{\edit1{Total (Reynolds + Maxwell) stress in the domain as predicted from the weakly nonlinear theory for $\Pm=1E-2$.}} \label{fig:wnl_stress}
+\caption{\edit1{Total (Reynolds + Maxwell) stress in the domain as predicted from the weakly nonlinear theory at $\Pm = 10^{-2}$.}} \label{fig:wnl_stress}
 \end{figure}

### 2017-03-25 paper/thingap_ref_response/response_to_referee.tex
merged.

@@
 We note, however, that although the value of the quantity $(b/c)^{1/2}$ is heavily influenced by the boundary layers in the walled TC flow, this is specific to the wall-like boundary conditions and not inherent to the theory.
-We have added a paragraph at the end of section 4 to elaborate on this point, discuss what the $(b/c)^{1/2}$ saturation amplitude means physically, and describe how the derived amplitude applies to the shearing box MRI.}
+We have added a paragraph at the end of section 4 to elaborate on this point, discuss what the $(b/c)^{1/2}$ saturation amplitude means physically, and describe how the derived amplitude applies to the shearing box MRI.
+We also confirmed numerical convergence in the widegap geometry and added a sentence noting convergence in the text.}
 
@@
 
-\textbf{We have added a figure to the thin-gap paper showing the morphology of the average angular momentum transport $\dot{J}$. }
+\textbf{We have added a plot showing the morphology of the stress in the x-z domain (Fig 5), a comparison of angular momentum transport by Maxwell stresses, Reynolds stresses, and total stresses in both the weakly nonlinear theory and direct numerical simulations (Fig 6), and more discussion both in the "Discussion" section and throughout the text.}
+%\textbf{We have added a figure to the thin-gap paper showing the morphology of the average angular momentum transport $\dot{J}$. }
 
@@
 
+%\textbf{We have carried out a direct numerical situation of the thin-gap TC flow with Dedalus and have added a new section (``Simulations") discussing the results.}
+
   \textbf{We have added a new section including a direct numerical simulation run with Dedalus, and made a preliminary comparison at modest Pm. }
@@
 
-\textbf{}
+\textbf{A note on another change in Paper I: we realized that the parameterization of destabilization into the weakly nonlinear regime was incorrect in the previously submitted version of this paper, and in URM07: the background field must be raised, rather than lowered, by a factor of $\epsilon^2$ in order to destabilize the system.
+This is a facet of the nonideal MRI that is treated incorrectly in previous works, and we add text and a figure (Figure 4) to explain this point.
+We have also amended the appendices of both papers to reflect the correct tuning.
+}
+
+\textbf{We thank the referee again for useful comments and probing questions.
+We hope that the referee will agree that with the extensive additions, including a new section discussing the applicability of the weakly nonlinear theory to the shearing box, the addition of a direct numerical simulation, and three new plots, that these papers are suitable for publication in The Astrophysical Journal.
+}
 

### 2017-03-26 paper/thingap_ref_response/response_to_referee.tex
fixed acknowledgments

@@
 
-  \textbf{We have added a new section including a direct numerical simulation run with Dedalus, and made a preliminary comparison at modest Pm. }
+  \textbf{We have added a new section including a direct numerical simulation run with Dedalus, and made a preliminary comparison at modest Pm. See Section 5 and Figure 6.}
 
@@
 
-\textbf{A note on another change in Paper I: we realized that the parameterization of destabilization into the weakly nonlinear regime was incorrect in the previously submitted version of this paper, and in URM07: the background field must be raised, rather than lowered, by a factor of $\epsilon^2$ in order to destabilize the system.
-This is a facet of the nonideal MRI that is treated incorrectly in previous works, and we add text and a figure (Figure 4) to explain this point.
+\textbf{A note on another change in both papers: we realized that the parameterization of destabilization into the weakly nonlinear regime was incorrect in the previously submitted version of this paper, and in URM07: the background field must be raised, rather than lowered, by a factor of $\epsilon^2$ in order to destabilize the system.
+This is a facet of the nonideal MRI that is treated incorrectly in previous works, and we add text and a figure (Figure 4) to explain this point in Paper I.
 We have also amended the appendices of both papers to reflect the correct tuning.
-}
+Thanks to a fortuitous sign error the plots were correct for everything but the helical MRI saturated perturbation structure (Paper II, Figure 8), which has been corrected.
+(The correct parameterization changes the overall saturation amplitude derived from the weakly nonlinear theory, but not the character of the perturbation structure.) }
 

### 2017-03-26 paper/widegap_ref_response/widegap_mri_paper_ref_response_apj.tex
merged.
fixed acknowledgments

@@
 \centering
-\includegraphics[width=\textwidth]{widegap_vel_Bfield_streamfuncs_512_final_wider-eps-converted-to.pdf}
+\includegraphics[width=\textwidth]{widegap_vel_Bfield_streamfuncs_512_final_wider}
 \caption{Perturbation structure for the velocity and magnetic field of the fiducial standard MRI case, including first and second order perturbations.
@@
 \centering
-\includegraphics[width=\textwidth]{hmri_velocity_Bfield_streamfuncs-eps-converted-to.pdf}
+\includegraphics[width=\textwidth]{hmri_velocity_Bfield_streamfuncs_correct_satamp}
 \caption{As in Figure \ref{fig:smri_streamfuncs} but for the fiducial helical MRI case, including first and second order perturbations.}
@@
 J.S.O.
-acknowledges support from NASA grant 15-LWS15\_2-0063.
-The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
+acknowledges support from NASA grant \edit1{NNX16AC92G}.
+\edit1{We thank the anonymous referee for thoughtful comments.
+We also thank} Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
 \software{Dedalus: http://dedalus-project.org/}

### 2017-03-26 paper/thingap_ref_response/response_to_referee.tex
final edits for paper one

@@
 Below please find the reviewer comments in plain text and our responses to these specific points in bold.
+Corresponding changes are marked in the resubmitted texts using the aastex edit command, which makes them pink.
 First, an overarching point: our papers are separated into Paper I and Paper II, where Paper I considers the MRI equations in a local geometry and Paper II considers the global MRI equations.
@@
 In the resubmitted version of Paper I we identify conditions under which our theory applies to the shearing box: the additional consideration of the effects of ambipolar diffusion and a constant azimuthal component to the background field.
-We add a section (3.1) that details how these conditions allow the Ginzburg-Landau amplitude equation to be obtained for the system with shear-periodic boundary conditions, as well as discussion throughout the text.
+We add a section (Section 4) that details how these conditions allow the Ginzburg-Landau amplitude equation to be obtained for the system with shear-periodic boundary conditions, as well as discussion throughout the text.
 This additional insight is a more direct connection between our work and most MRI simulations, in addition to experiments, and accordingly we are changing the title of Paper I to ``The weakly nonlinear magnetorotational instability in a local geometry" rather than ``in a thin-gap Taylor Couette flow".
@@
 Such a program is beyond the scope of this work, though it is a very interesting possibility we had not previously considered.
-However, such inter-mode coupling \emph{can} be analyzed in the shearing box when ambipolar diffusion is considered, as described in the new section of the text.}
+However, such inter-mode coupling \emph{can} be analyzed in the shearing box when ambipolar diffusion is considered, as described in the new Section 4 in the text.}
 
@@
 
-\textbf{The amplitude $\alpha$ is bounded by the value +/- $(b/c)^{1/2}$, although the amplitude does not necessarily converge to a constant value, as shown in Figures 6 and 7.
+\textbf{The amplitude $\alpha$ is bounded by the value +/- $(b/c)^{1/2}$, although the amplitude does not necessarily converge to a constant value, as shown in Figures 9 and 10 (previously 6 and 7) in Paper I.
 We note, however, that although the value of the quantity $(b/c)^{1/2}$ is heavily influenced by the boundary layers in the walled TC flow, this is specific to the wall-like boundary conditions and not inherent to the theory.
-We have added a paragraph at the end of section 4 to elaborate on this point, discuss what the $(b/c)^{1/2}$ saturation amplitude means physically, and describe how the derived amplitude applies to the shearing box MRI.
-We also confirmed numerical convergence in the widegap geometry and added a sentence noting convergence in the text.}
+We have added a paragraph at the end of Paper I's section 6 to elaborate on this point, discuss what the $(b/c)^{1/2}$ saturation amplitude means physically, and describe how the derived amplitude applies to the shearing box MRI.
+We also confirmed numerical convergence in the Paper II geometry and added a sentence noting convergence in the text.}
 
@@
 
-\textbf{We have added a plot showing the morphology of the stress in the x-z domain (Fig 5), a comparison of angular momentum transport by Maxwell stresses, Reynolds stresses, and total stresses in both the weakly nonlinear theory and direct numerical simulations (Fig 6), and more discussion both in the "Discussion" section and throughout the text.}
+\textbf{We have added a plot showing the morphology of the stress in the x-z domain (Fig 5), a comparison of angular momentum transport by Maxwell stresses, Reynolds stresses, and total stresses in both the weakly nonlinear theory and direct numerical simulations (Fig 6), and more discussion both in the``Discussion" section and throughout the text.}
 %\textbf{We have added a figure to the thin-gap paper showing the morphology of the average angular momentum transport $\dot{J}$. }
@@
 
-\textbf{Yes, the saturated Bz is marginally stable, and we have added a sentence to that effect in the text.
+\textbf{Yes, the saturated Bz is marginally stable, and we have added a sentence to note this in the text.
 We cannot comment on dynamo action in either paper because of the symmetries in our setup.
@@
 \textbf{We thank the referee again for useful comments and probing questions.
-We hope that the referee will agree that with the extensive additions, including a new section discussing the applicability of the weakly nonlinear theory to the shearing box, the addition of a direct numerical simulation, and three new plots, that these papers are suitable for publication in The Astrophysical Journal.
+We hope that the referee will agree that with the extensive additions, including a new section discussing the applicability of the weakly nonlinear theory to the shearing box, the addition of a direct numerical simulation, and three new plots, that these papers are worthy of publication in The Astrophysical Journal.
 }

### 2017-03-26 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
fixed acknowledgments
final edits for paper one

@@
 (2007) that the perturbation amplitude follows a Ginzburg-Landau equation.
-We extend these results by performing a detailed force balance for the saturated azimuthal velocity and vertical magnetic field, demonstrating that even when diffusive effects are important, the bulk flow saturates via the combined processes of reducing the background shear and rearranging and strengthening the background vertical magnetic field.
+\edit1{We further find that the Ginzburg-Landau equation will arise for the local MRI system with shear-periodic boundary conditions when the effects of ambipolar diffusion are considered.} A detailed force balance for the saturated azimuthal velocity and vertical magnetic field demonstrates that even when diffusive effects are important, the bulk flow saturates via the combined processes of reducing the background shear and rearranging and strengthening the background vertical magnetic field.
 We directly simulate the Ginzburg-Landau amplitude evolution for our system and demonstrate the pattern formation our model predicts on long length and time scales.
+\edit1{We compare the weakly nonlinear theory results to a direct numerical simulation of the MRI in a thin-gap Taylor Couette flow.}
 
@@
 true Taylor-Couette flow) the base angular velocity is a function of $r$, where in the thin-gap approximation the shear flow reduces to a linear profile.
-\edit1{The equations of motion are thus identical to the MRI in a local shearing box, which differs from our fiducial setup only in the application of periodic boundary conditions.}
+\edit1{The equations of motion in thin-gap TC flow are thus identical to the MRI in a local shearing box, which differs from our fiducial setup only in the application of periodic boundary conditions.}
 
@@
 \citet{Vasil:2015} examines the weakly nonlinear MRI in a thin-gap regime in a minimal model, finding deep mathematical similarities between the MRI system and the elastodynamic instability of a buckling beam.
-We discuss these results and their relation to ours in section~\ref{sec:discussion}.
+We discuss these results and their relation to ours in Section~\ref{sec:discussion}.
 Several authors have investigated the behavior of the MRI when the boundary conditions are shear periodic, and so the MRI has no mechanism by which to modify the background shear flow profile.
@@
 
-We begin with an overview of our basic model equations for the \edit1{local} MRI in section~\ref{sec:equations} and then describe our weakly nonlinear analysis and give results \edit1{for the thin-gap TC flow} in section~\ref{sec:wnl}.
-\edit1{In Section \ref{sec:SBAD} we detail the conditions under which our theory applies to the case where the boundary conditions are shear periodic, namely the consideration of ambipolar diffusion.} We then place our results in the context of previous results from both analytic and computational studies in section~\ref{sec:discussion} and draw conclusions in section~\ref{sec:conclusion}.
+We begin with an overview of our basic model equations for the \edit1{local} MRI in Section~\ref{sec:equations} and then describe our weakly nonlinear analysis and give results \edit1{for the thin-gap TC flow} in Section~\ref{sec:wnl}.
+\edit1{In Section \ref{sec:SBAD} we detail the conditions under which our theory applies to the case where the boundary conditions are shear periodic, namely the consideration of ambipolar diffusion.
+We compare our results to a direct numerical simulation in Section \ref{sec:simulations}.} We then place our results in the context of previous results from both analytic and computational studies in Section~\ref{sec:discussion} and draw conclusions in Section~\ref{sec:conclusion}.
 
@@
 and $q \equiv - d \ln \Omega/ \ln R = 3/2$ is the dimensionless shear parameter defining a rotation profile $\Omega(r) = \Omega_0 (r/r_0)^{-q}$, such that the background velocity profile is $u_0 = -q \Omega_0 x$.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=\columnwidth]{growth_vs_B0.pdf}
+  \caption{\edit1{Growth rate $\gamma$ as a function of background magnetic field strength $B_0$ at $\reym = \reym_c$, $k_z = k_c$, $\Pm = 10^{-3}$.
+Around the critical value $B_0 = 1.$, \emph{strengthening} $B_0$ tunes the system into instability, while \emph{decreasing} it leads to stability.
+The inset highlights the fact that $\gamma$ is determined by the maximum real eigenvalue of the system, which switches from one mode family to another as discussed in the text.}}
+  \label{fig:growth_rate_vs_B0}
+\end{figure}
 
@@
 
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=\columnwidth]{growth_vs_B0.pdf}
-  \caption{\edit1{Growth rate $\gamma$ as a function of background magnetic field strength $B_0$ at $\reym = \reym_c$, $k_z = k_c$, $\Pm = 10^{-3}$.
-Around the critical value $B_0 = 1.$, \emph{strengthening} $B_0$ tunes the system into instability, while \emph{decreasing} it leads to stability.
-The inset highlights the fact that $\gamma$ is determined by the maximum real eigenvalue of the system, which switches from one mode family to another as discussed in the text.}}
-  \label{fig:growth_rate_vs_B0}
-\end{figure}
-
 The destabilizing substitution is made, and Equations \ref{eqset1} - \ref{eqset4} are rewritten such that the fluid variables are contained in a state vector
@@
 Similarly, we cannot analytically examine interactions between MRI channel modes and damped eigenmodes belonging to other wave families.
-This is analytically tracked for other plasma instabilities by tracking the amplitudes of growing, marginal, and damped eigenmodes simultaneously \citep[e.g.][]{Makwana:2011}.
+This is analytically examined for other plasma instabilities by tracking the amplitudes of growing, marginal, and damped eigenmodes simultaneously \citep[e.g.][]{Makwana:2011}.
 While the shearing box approximation allows the projection of the perturbed MRI equations into the basis set of linear eigenmodes, nonlinear coupling between modes will remain zero.}
@@
 
-\section{Simulations}\label{sec:simulations}
+\section{Direct Numerical Simulation}\label{sec:simulations}
 
@@
 }
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{thingap_saturation_mechanism_plots_velocity_slice60.pdf}
+\caption{Top panel: radial profile for $u_y^{saturated} - u_y^0$, the difference between the saturated azimuthal velocity profile and the initial azimuthal velocity profile (TC flow).
+Bottom panel: each term in the steady state force balance (Equation \ref{eqset2} with $\partial_t u_{y} = 0$).
+Saturated quantities are computed with $\alpha_{saturation} = \sqrt{b/c}$ and $\epsilon = 0.5$.
+The saturated state shows reduced shear in the bulk of the flow, outside of the boundary layers.} \label{fig:sat_comp_vel}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{thingap_saturation_mechanism_plots_Bfield_slice60.pdf}
+\caption{Top panel: radial profile for $B_z^{saturated}$, the saturated vertical magnetic field (black line).
+$B_z^0 = 1$ is the constant background magnetic field (gray line).
+Bottom panel: each term in the steady state inductive balance ($\partial_x$ of Equation \ref{eqset4} with $\partial_t B_y = 0$).
+Saturated quantities are computed with $\alpha_{saturation} = \sqrt{b/c}$ and $\epsilon = 0.5$.
+The saturated field is pushed to the radial domain boundaries.}\label{fig:sat_comp_Bfield}
+\end{figure}
 
@@
 \label{sec:discussion}
-
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{thingap_saturation_mechanism_plots_velocity_slice60.pdf}
-\caption{Top panel: radial profile for $u_y^{saturated} - u_y^0$, the difference between the saturated azimuthal velocity profile and the initial azimuthal velocity profile (TC flow).
-Bottom panel: each term in the steady state force balance (Equation \ref{eqset2} with $\partial_t u_{y} = 0$).
-Saturated quantities are computed with $\alpha_{saturation} = \sqrt{b/c}$ and $\epsilon = 0.5$.
-The saturated state shows reduced shear in the bulk of the flow, outside of the boundary layers.} \label{fig:sat_comp_vel}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{thingap_saturation_mechanism_plots_Bfield_slice60.pdf}
-\caption{Top panel: radial profile for $B_z^{saturated}$, the saturated vertical magnetic field (black line).
-$B_z^0 = 1$ is the constant background magnetic field (gray line).
-Bottom panel: each term in the steady state inductive balance ($\partial_x$ of Equation \ref{eqset4} with $\partial_t B_y = 0$).
-Saturated quantities are computed with $\alpha_{saturation} = \sqrt{b/c}$ and $\epsilon = 0.5$.
-The saturated field is pushed to the radial domain boundaries.}\label{fig:sat_comp_Bfield}
-\end{figure}
 
@@
 Figure~\ref{fig:sat_comp_Bfield} shows $B_z$ and the terms corresponding to steady state inductive balance ($\partial_x$ of Equation~\ref{eqset4} with $\partial_t B_{y} = 0$).
-Here, the instability acts to push magnetic field toward the boundaries in both the bulk and the boundary layers.
+Here, the instability acts to push the magnetic field toward the boundaries in both the bulk and the boundary layers.
 %Note also that the total magnetic energy in the $B_z$ component is significantly larger than the initial condition.
@@
 We simulate the evolution of the MRI amplitude equation by solving Equation \ref{eq:gle} on a Fourier basis in $Z$ using Dedalus.
-We initialize uniform random noise of amplitude $-10^{-3}$ to $+10^{-3}$ in $Z$, and timestep the system using a four-stage, fourth-order Runge-Kutta integrator.
+We initialize uniform random noise of amplitude $-10^{-3}$ to $+10^{-3}$ in $Z$, and timestep the system using a four-stage, third-order Runge-Kutta integrator.
 We evolve the system for $100 \Omega_0^{-1}$ in timesteps of $0.02 \Omega_0^{-1}$.
@@
 
-\edit1{Figure~\ref{fig:wnl_stress} shows the total stress $u_x u_y - \Co B_x B_y$ for the $\Pm = 10^{-2}$ model.
-The stress shows signficantly more structure throughout the domain than the  variables $u_x$, $u_y$, $B_x$ and $B_y$ that comprise it, demonstrating that a non-trivial correlation exists even in the weakly non-linear state.
+\edit1{Figure~\ref{fig:wnl_stress} shows the total stress $u_x u_y - \Co B_x B_y$ for the $\Pm = 10^{-2}$ model with $\epsilon = 0.5$.
+The stress shows significantly more structure throughout the domain than the variables $u_x$, $u_y$, $B_x$ and $B_y$ that comprise it, demonstrating that a non-trivial correlation exists even in the weakly non-linear state.
 As in simulations at higher $\reym$, figure~\ref{fig:sim_theory_energy_jdot} shows that the Maxwell stress dominates over the Reynolds stresses even though the kinetic energy significantly exceeds the magnetic energy.
@@
 In doing so, we find the saturated state is a complex balance in which reduction of shear and amplification and redistribution of $B_z$ combine to saturate the instability.
-\edit1{We also perform numerical simulations of the amplitude equation and a direct numerical simulation of the MRI system.} \edit1{Using the former, we demonstrate that complex patterns can organize the flow on long length scales $Z$, though the maximum magnitude of the amplitude $\alpha$ is well predicted by the steady state solution.
+\edit1{We perform numerical simulations of the amplitude equation and a direct numerical simulation of the MRI system.} \edit1{Using the former, we demonstrate that complex patterns can organize the flow on long length scales $Z$, though the maximum magnitude of the amplitude $\alpha$ is well predicted by the steady state solution.
 The latter show that there is rough agreement for both total energy and average angular momentum transport between the weakly nonlinear theory and simulation for a representative case at $\Pm = 10^{-2}$.
-We defer a full comparison between theory and simulation to later work.} In Paper II, we make use of the techniques developed here to extend the weakly nonlinear analysis of the MRI to a full cylindrical geometry appropriate for a Taylor-Couette experiment.
+We defer a full comparison between theory and simulation to later work.
+We describe the application of shear-periodic boundary conditions to the local MRI and find that with the inclusion of certain nonideal physical effects, namely ambipolar diffusion, our theory points to a new saturation avenue for the MRI in a shearing box.} In Paper II, we make use of the techniques developed here to extend the weakly nonlinear analysis of the MRI to a full cylindrical geometry appropriate for a Taylor-Couette experiment.
 
@@
 Chandrasekhar, S. 1960, Proceedings of the National Academy of Sciences of the United States of America, 46, 253
-\bibitem[{Clark \& Oishi(2016b)}]{Clark:2016}
-Clark, S.E. and Oishi, J.S., 2016, submitted
+\bibitem[{Clark \& Oishi(2017b)}]{Clark:2016}
+Clark, S.E. and Oishi, J.S., 2017, submitted
 

### 2017-03-26 paper/thingap_ref_response/response_to_referee.tex
okay final final version. proceeding to resubmit

@@
 \textbf{A note on another change in both papers: we realized that the parameterization of destabilization into the weakly nonlinear regime was incorrect in the previously submitted version of this paper, and in URM07: the background field must be raised, rather than lowered, by a factor of $\epsilon^2$ in order to destabilize the system.
-This is a facet of the nonideal MRI that is treated incorrectly in previous works, and we add text and a figure (Figure 4) to explain this point in Paper I.
+This is a facet of the nonideal MRI that is treated incorrectly in previous works, and we add text and a figure (Figure 2) to explain this point in Paper I.
 We have also amended the appendices of both papers to reflect the correct tuning.

### 2017-03-26 paper/widegap_ref_response/widegap_mri_paper_ref_response_apj.tex
okay final final version. proceeding to resubmit

@@
 Previous investigations into the weakly nonlinear behavior of the MRI have used this narrow gap approximation \citep{Umurhan:2007dz,Umurhan:2007hs,Clark:2016}.
-Building on our work in \edit1{Paper 1}, here we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.
+Building on our work in \edit1{Paper I}, here we undertake the first (to our knowledge) weakly nonlinear analysis of the MRI in the wide gap regime, where the channel width may be comparable to or larger than its distance from the center of rotation.
 
@@
 \edit1{We use a 512-component Chebyshev grid, and confirm numerical convergence at $1.5\times$ the resolution.
-This is sufficient to determine convergence because of the faster-than-exponential convergence of spectral methods.
-}
+This is sufficient to determine convergence because of the faster-than-exponential convergence of spectral methods \citep{Boyd:2001aa}.}
 We solve Equation \ref{eq:ordere} as a linear eigenvalue problem, and Equation \ref{eq:ordere2} as a linear boundary value problem. 
@@
 acknowledges support from NASA grant \edit1{NNX16AC92G}.
-\edit1{We thank the anonymous referee for thoughtful comments.
+\edit1{We thank the anonymous referee for many thoughtful comments on both this and Paper I.
 We also thank} Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
@@
 
+\bibitem[{Boyd(2001)}]{Boyd:2001aa}
+Boyd, J P, 2001, Chebyshev and Fourier Spectral Methods, New York, Dover
+
 \bibitem[{Brandenburg et al.(1995)}]{Brandenburg:1995}
@@
 
-\bibitem[{Clark \& Oishi(2016a)}]{Clark:2016}
-Clark, S.E. and Oishi, J.S., 2016, in prep
+\bibitem[{Clark \& Oishi(2017a)}]{Clark:2016}
+Clark, S.E. and Oishi, J.S., 2017, submitted
 
